### 2024-02-08 Masterarbeit_Robotik_Pamuk.tex
timeline

%
% FH Technikum Wien
% !TEX encoding = UTF-8 Unicode
%
% Erstellung von Master- und Bachelorarbeiten an der FH Technikum Wien mit Hilfe von LaTeX und der Klasse TWBOOK
%
% Um ein eigenes Dokument zu erstellen, müssen Sie folgendes ergänzen:
% 1) Mit \documentclass[..] einstellen: Master- oder Bachelorarbeit, Studiengang und Sprache
% 2) Mit \newcommand{\FHTWCitationType}.. Zitierstandard festlegen (wird in der Regel vom Studiengang vorgegeben - bitte erfragen)
% 3) Deckblatt, Kurzfassung, etc. ausfüllen
% 4) und die Arbeit schreiben (die verwendeten Literaturquellen in Literatur.bib eintragen)
%
% Getestet mit TeXstudio mit Zeichenkodierung ISO-8859-1 (=ansinew/latin1) und MikTex unter Windows
% Zu beachten ist, dass die Kodierung der Datei mit der Kodierung des paketes inputenc zusammen passt!
% Die Kodierung der Datei twbook.cls MUSS ANSI betragen!
% Bei der Verwendung von UTF8 muss dnicht nur die Kodierung des Dokuments auf UTF8 gestellt sein, sondern auch die des BibTex-Files!
%
% Bugreports und Feedback bitte per E-Mail an latex@technikum-wien.at
%
% Versionen
% *) V0.7: 9.1.2015, RO: Modeline angepasst und verschoben
% *) V0.6: 10.10.2014, RO: Weitere Anpassung an die UK
% *) V0.5: 8.8.2014, WK: Literaturquellen überarbeitet und angepasst
% *) V0.4: 4.8.2014, WK: Initalversion in SVN eingespielt
%
\documentclass[MMR,Master,english]{twbook}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%
% Hier biblatex & Biber konfigurieren; Vergessen Sie nicht, dass Sie biber verwenden müssen um eine Bibliothek zu erzeugen
%
\usepackage[backend=biber, style=numeric]{biblatex}
\addbibresource{Literatur.bib}

%
% Bei Bedarf bitte hier die Syntax-Highlightings anpassen
%
\usepackage[final]{listings}
\lstset{captionpos=b, numberbychapter=false,caption=\lstname,frame=single, numbers=left, stepnumber=1, numbersep=2pt, xleftmargin=15pt, framexleftmargin=15pt, numberstyle=\tiny, tabsize=3, columns=fixed, basicstyle={\fontfamily{pcr}\selectfont\footnotesize}, keywordstyle=\bfseries, commentstyle={\color[gray]{0.33}\itshape}, stringstyle=\color[gray]{0.25}, breaklines, breakatwhitespace, breakautoindent}
\lstloadlanguages{[ANSI]C, C++, [gnu]make, gnuplot, Matlab}

%Formatieren des Quellcodeverzeichnisses
\makeatletter
% Setzen der Bezeichnungen für das Quellcodeverzeichnis/Abkürzungsverzeichnis in Abhängigkeit von der eingestellten Sprache
\providecommand\listacroname{}
\@ifclasswith{twbook}{english}
{%
    \renewcommand\lstlistingname{Code}
    \renewcommand\lstlistlistingname{List of Code}
    \renewcommand\listacroname{List of Abbreviations}
}{%
    \renewcommand\lstlistingname{Quellcode}
    \renewcommand\lstlistlistingname{Quellcodeverzeichnis}
    \renewcommand\listacroname{Abkürzungsverzeichnis}
}
% Wenn die Option listof=entryprefix gewählt wurde, Definition des Entyprefixes für das Quellcodeverzeichnis. Definition des Macros listoflolentryname analog zu listoflofentryname und listoflotentryname der KOMA-Klasse
\@ifclasswith{scrbook}{listof=entryprefix}
{%
    \newcommand\listoflolentryname\lstlistingname
}{%
}
\makeatother
\newcommand{\listofcode}{\phantomsection\lstlistoflistings}

% Die nachfolgenden Pakete stellen sonst nicht benötigte Features zur Verfügung
\usepackage{blindtext}

%
% Einträge für Deckblatt, Kurzfassung, etc.
%
\title{Arbeitstitel\\Arbeitstitel}
\author{Halil Pamuk, BSc}
\studentnumber{51842568}
%\author{Titel Vorname Name, Titel\and{}Titel Vorname Name, Titel}
%\studentnumber{XXXXXXXXXXXXXXX\and{}XXXXXXXXXXXXXXX}
\supervisor{Sebastian Rauh, MSc. BEng}
%\supervisor[Begutachter]{Titel Vorname Name, Titel}
%\supervisor[Begutachterin]{Titel Vorname Name, Titel}
%\secondsupervisor{Titel Vorname Name, Titel}
%\secondsupervisor[Begutachter]{Titel Vorname Name, Titel}
%\secondsupervisor[Begutachterinnen]{Titel Vorname Name, Titel}
\place{Wien}
\kurzfassung{\blindtext}
\schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
\outline{\blindtext}
\keywords{Keyword1, Keyword2, Keyword3, Keyword4}
%\acknowledgements{\blindtext}

\begin{document}

\maketitle

%
% .. und hier beginnt die eigentliche Arbeit. Viel Erfolg beim Verfassen!
%
\chapter{Erste Überschrift der Ebene 1 (chapter)}
\blinddocument

\blindmathpaper

\section{Erste Überschrift Tiefe 2 (section)}
\blindtext

\subsection{Erste Überschrift Tiefe 3 (subsection)}
\blindtext

\subsubsection{Erste Überschrift Tiefe 4 (subsubsection)}
\blindtext

\chapter{Zweite Überschrift der Tiefe 1 (chapter)}
\blindtext

\section{Zweite Überschrift Tiefe 2 (section)}
\blindtext

\section{Zweite Überschrift Tiefe 2 (section)}
\blindtext

\subsection{Zweite Überschrift Tiefe 3 (subsection)}
\blindtext

\subsection{Dritte Überschrift Tiefe 3 (subsection)}
\blindtext

\subsubsection{Zweite Überschrift Tiefe 4 (subsubsection)}
\blindtext

\noindent Querverweise werden in \LaTeX{} automatisch erzeugt und verwaltet, damit sie leicht aktualisiert werden können. Hier wird zum Beispiel auf Abbildung \ref{Abb1} verwiesen.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{img/buchruecken}
\caption{Beispiel für die Beschriftung eines Buchrückens.}\label{Abb1}
\end{figure}
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{img/buchruecken}
\caption{2. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb2}
\end{figure}

Und hier ist ein Verweis auf Tabelle \ref{tab1}. Das gezeigte Tabellenformat ist nur ein Beispiel. Tabellen können individuell gestaltet werden.

\begin{table}[!htbp]
\centering
\caption{Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab1}
\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
Datum & Thema & Raum\\\hline
20.08.2008 & Graphentheorie & HS 3.13\\
01.10.2008 & Biomathematik & HS 1.05\\\hline
\end{tabular}
\end{table}
\begin{table}[!htbp]
\centering
\caption{2. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab2}
\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
Datum & Thema & Raum\\\hline
20.08.2008 & Graphentheorie & HS 3.13\\
01.10.2008 & Biomathematik & HS 1.05\\\hline
\end{tabular}
\end{table}

Hier wird auf die Formel \ref{Gl1} verwiesen.

\begin{align}
x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl1}
\end{align}
\begin{align}
x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl2}
\end{align}

\begin{lstlisting}[language=C++,name={1. Beispiel},label={sc:bsp:1}]
#include <iostream>

void SayHello(void)
{
    // Kommentar
    cout << "Hello World!" << endl;
}

int main(int argc, char **argv)
{
    SayHello();
    return 0;
}
\end{lstlisting}

Literaturverweise sollten automatisch verwaltet werden, vor allem, wenn es viele Quellenverweise gibt. Beispiele sind  \cite{Ko05a}, \cite{Ko05b}, \cite{MiGo05}, \cite{TeGo14}, \cite{HuHa07}, \cite{HuZi10}, \cite{ZiKu07}, \cite{He07}, \cite{SIE11}, \cite{SIE14}, \cite{ISO98}, \cite{ATM11}, \cite{Hu11}, \cite{Po10}. Das verwendete Zitierformat (bzw.~das Format des Literaturverzeichnisses) ist entspechend der Vorgaben der Studiengänge zu wählen.
Es wird dringend empfohlen, BibTeX~zu verwenden (wie in diesem Beispiel).

\chapter{Dritte Überschrift der Tiefe 1 (chapter)}
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{img/buchruecken}
\caption{3. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb3}
\end{figure}
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{img/buchruecken}
\caption{4. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb4}
\end{figure}

\begin{table}[!htbp]
\centering
\caption{3. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab3}
\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
Datum & Thema & Raum\\\hline
20.08.2008 & Graphentheorie & HS 3.13\\
01.10.2008 & Biomathematik & HS 1.05\\\hline
\end{tabular}
\end{table}
\begin{table}[!htbp]
\centering
\caption{4. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab4}
\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
Datum & Thema & Raum\\\hline
20.08.2008 & Graphentheorie & HS 3.13\\
01.10.2008 & Biomathematik & HS 1.05\\\hline
\end{tabular}
\end{table}

\begin{align}
x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl3}
\end{align}
\begin{align}
x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl4}
\end{align}
\begin{lstlisting}[language=C++,name={2. Beispiel},label={sc:bsp:2}]
#include <iostream>

void SayHello(void)
{
    // Kommentar
    cout << "Hello World!" << endl;
}

int main(int argc, char **argv)
{
    SayHello();
    return 0;
}
\end{lstlisting}

%
% Hier beginnen die Verzeichnisse.
%
\clearpage
\printbibliography
\clearpage

\cite{SIE14}

% Das Abbildungsverzeichnis
\listoffigures
\clearpage

% Das Tabellenverzeichnis
\listoftables
\clearpage

% Das Quellcodeverzeichnis
\listofcode
\clearpage

\phantomsection
\addcontentsline{toc}{chapter}{\listacroname}
\chapter*{\listacroname}
\begin{acronym}[XXXXX]
    \acro{ABC}[ABC]{Alphabet}
    \acro{WWW}[WWW]{world wide web}
    \acro{ROFL}[ROFL]{Rolling on floor laughing}
\end{acronym}

%
% Hier beginnt der Anhang.
%
\clearpage
\appendix
\chapter{Anhang A}
\clearpage
\chapter{Anhang B}
\end{document}

### 2024-03-10 Masterarbeit_Robotik_Pamuk.tex
Leitfaden Masterarbeit und Gliederung

@@
 %
-\title{Arbeitstitel\\Arbeitstitel}
+\title{Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die
+Einhaltung der Echtzeit}
 \author{Halil Pamuk, BSc}
@@
 \place{Wien}
-\kurzfassung{\blindtext}
+\kurzfassung{Erstellung einer Echtzeit-Robotersteuerungsplattform unter Verwendung von Salamander OS, Xenomai, QEMU
+und PCV-521 in der Yocto-Umgebung. Die Plattform basiert auf Salamander OS und nutzt Xenomai für Echtzeit-
+Funktionen. Dazu muss im ersten Schritt die Virtualisierungsplattform evaluiert werden. (QEMU, Hyper-V, Virtual
+Box, etc.) Als weiterer Schritt folgt die Anbindung eines Roboters über eine VARAN-Bus Schnittstelle. Das
+gesamte System wird in der Yocto-Umgebung erstellt und konfiguriert.
+Das Hauptziel der Arbeit ist es, herauszufinden, wie die Integration von Echtzeit-Funktionen und effizienten
+Kommunikationssystemen in eine Robotersteuerungsplattform die Reaktionszeit und Zuverlässigkeit von
+Roboteranwendungen verbessern kann}
 \schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
 \outline{\blindtext}
-\keywords{Keyword1, Keyword2, Keyword3, Keyword4}
+\keywords{Echtzeit, Virtualisierung, Xenomai, VARAN}
 %\acknowledgements{\blindtext}
@@
 %
-\chapter{Erste Überschrift der Ebene 1 (chapter)}
-\blinddocument
+%Die folgende Gliederung von Bachelor- und Masterarbeiten ist für die Robotik-Studiengänge
+%verpflichtend vorgeschrieben.
+%
+%- Deckblatt
+%- Eidesstattliche Erklärung (digital signiert)
+%- Kurzfassung und Abstract mit Schlüsselwörtern/Keywords
+%- Danksagung (optional)
+%- Inhaltsverzeichnis
 
+%- Einleitung (die genannten Punkte müssen keine eigenen Unterkapitel sein, müssen aber aus der Einleitung klar hervorkommen):
+%   - Stand der Technik
+%   - Problem- und Aufgabenstellung
+%   - Zielsetzung
+%- Hauptteil -> Gliederung je nach Thema
+%   - Methodik
+%   - Resultate
+%   - Wirtschaftliche Betrachtung (optional)
+%   - Diskussion
+%- Zusammenfassung und Ausblick
+
+%- Literaturverzeichnis
+%- Abbildungsverzeichnis
+%- Tabellenverzeichnis
+%- Abkürzungsverzeichnis (optional)
+%- Glossar (optional)
+%- Anhang (optional)
+\chapter{Einleitung}
 \blindmathpaper
-
-\section{Erste Überschrift Tiefe 2 (section)}
+\clearpage
+\section{Stand der Technik}
+\blindtext
+\clearpage
+\section{Problem- und Aufgabenstellung}
+\blindtext
+\clearpage
+\section{Zielsetzung}
 \blindtext
 
-\subsection{Erste Überschrift Tiefe 3 (subsection)}
-\blindtext
-
-\subsubsection{Erste Überschrift Tiefe 4 (subsubsection)}
-\blindtext
-
-\chapter{Zweite Überschrift der Tiefe 1 (chapter)}
-\blindtext
-
-\section{Zweite Überschrift Tiefe 2 (section)}
-\blindtext
-
-\section{Zweite Überschrift Tiefe 2 (section)}
-\blindtext
-
-\subsection{Zweite Überschrift Tiefe 3 (subsection)}
-\blindtext
-
-\subsection{Dritte Überschrift Tiefe 3 (subsection)}
-\blindtext
-
-\subsubsection{Zweite Überschrift Tiefe 4 (subsubsection)}
-\blindtext
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+\chapter{Methodik}
 
@@
 
-\chapter{Dritte Überschrift der Tiefe 1 (chapter)}
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.5\linewidth]{img/buchruecken}
-\caption{3. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb3}
-\end{figure}
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.5\linewidth]{img/buchruecken}
-\caption{4. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb4}
-\end{figure}
-
-\begin{table}[!htbp]
-\centering
-\caption{3. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab3}
-\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
-Datum & Thema & Raum\\\hline
-20.08.2008 & Graphentheorie & HS 3.13\\
-01.10.2008 & Biomathematik & HS 1.05\\\hline
-\end{tabular}
-\end{table}
-\begin{table}[!htbp]
-\centering
-\caption{4. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab4}
-\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
-Datum & Thema & Raum\\\hline
-20.08.2008 & Graphentheorie & HS 3.13\\
-01.10.2008 & Biomathematik & HS 1.05\\\hline
-\end{tabular}
-\end{table}
-
-\begin{align}
-x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl3}
-\end{align}
-\begin{align}
-x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl4}
-\end{align}
-\begin{lstlisting}[language=C++,name={2. Beispiel},label={sc:bsp:2}]
-#include <iostream>
-
-void SayHello(void)
-{
-    // Kommentar
-    cout << "Hello World!" << endl;
-}
-
-int main(int argc, char **argv)
-{
-    SayHello();
-    return 0;
-}
-\end{lstlisting}
+\clearpage
+\chapter{Hauptteil}
+\clearpage
+\chapter{Resultate}
+\clearpage
+\chapter{Diskussion}
+\clearpage
+\chapter{Zusammenfassung und Ausblick}
 
@@
 \clearpage
-
-\cite{SIE14}
 

### 2024-03-19 Masterarbeit_Robotik_Pamuk.tex
kernel_processes.md, kernelshark, start thesis

@@
 \schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
-\outline{\blindtext}
+\outline{Abstract}
 \keywords{Echtzeit, Virtualisierung, Xenomai, VARAN}
@@
 %- Anhang (optional)
-\chapter{Einleitung}
-\blindmathpaper
+\chapter{Introduction}
+
 \clearpage
-\section{Stand der Technik}
-\blindtext
+\section{State of the art}
+
 \clearpage
-\section{Problem- und Aufgabenstellung}
-\blindtext
+\section{Problem and task definition}
+In robotics, accurate and timely control is crucial to ensure precise movements and reliable interaction with the environment.
+However, existing robot control systems often have limitations in terms of real-time capabilities and communication efficiency, which can have a detrimental effect on response time and reliability.
+The challenge is to develop a powerful real-time robot control platform that overcomes these limitations and improves the performance of robotic applications.
+
 \clearpage
-\section{Zielsetzung}
-\blindtext
+\section{Objective}
+
+The main objective of this work is to create a real-time robot control platform that integrates Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
+By combining these technologies, a platform is to be created that provides real-time functions and efficient communication systems.
+This integration is expected to lead to a significant improvement in the response time and reliability of robot applications.
 
@@
 \clearpage
-\chapter{Methodik}
+\chapter{Methodology}
 
-\noindent Querverweise werden in \LaTeX{} automatisch erzeugt und verwaltet, damit sie leicht aktualisiert werden können. Hier wird zum Beispiel auf Abbildung \ref{Abb1} verwiesen.
+The methodology of this thesis consists of several steps.
+First, a comprehensive literature review is conducted to understand the current trends and challenges in real-time robot control.
+Based on the literature study, a suitable virtualisation platform is selected.
+After the selection of the virtualisation platform, the robot control platform is implemented.
+This step includes the installation and configuration of Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
+Once the platform has been implemented, the robot is connected via a VARAN bus interface.
+Finally, the platform is evaluated to determine how the integration of real-time functions and efficient communication systems improves the response time and reliability of robot applications.
 
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.5\linewidth]{img/buchruecken}
-\caption{Beispiel für die Beschriftung eines Buchrückens.}\label{Abb1}
-\end{figure}
-\begin{figure}[!htbp]
-\centering
-\includegraphics[width=0.5\linewidth]{img/buchruecken}
-\caption{2. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb2}
-\end{figure}
+- Evaluation der Virtualisierungsplattform: 
+Ich werde verschiedene Virtualisierungsplattformen wie QEMU, Hyper-V, Virtual Box usw. evaluieren. Dies ist ein wichtiger Schritt, um die beste Plattform für meine Anforderungen zu finden.
 
-Und hier ist ein Verweis auf Tabelle \ref{tab1}. Das gezeigte Tabellenformat ist nur ein Beispiel. Tabellen können individuell gestaltet werden.
+- Erstellung und Konfiguration des Systems in der Yocto-Umgebung: 
+Ich werde das Yocto-Framework verwenden, um mein Embedded Linux System zu erstellen und zu konfigurieren.
+Yocto bietet viele Tools und Funktionen, die mir bei der Erstellung und Konfiguration meines Systems helfen können.
 
-\begin{table}[!htbp]
-\centering
-\caption{Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab1}
-\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
-Datum & Thema & Raum\\\hline
-20.08.2008 & Graphentheorie & HS 3.13\\
-01.10.2008 & Biomathematik & HS 1.05\\\hline
-\end{tabular}
-\end{table}
-\begin{table}[!htbp]
-\centering
-\caption{2. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab2}
-\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
-Datum & Thema & Raum\\\hline
-20.08.2008 & Graphentheorie & HS 3.13\\
-01.10.2008 & Biomathematik & HS 1.05\\\hline
-\end{tabular}
-\end{table}
+- Verbesserung der Reaktionszeit und Zuverlässigkeit von Roboteranwendungen: 
+Mein Hauptziel ist es, herauszufinden, wie die Integration von Echtzeitfunktionen und effizienten Kommunikationssystemen die Reaktionszeit und Zuverlässigkeit von Roboteranwendungen verbessern kann.
+Ich strebe an, die Leistung und Zuverlässigkeit meiner Roboteranwendungen zu verbessern, indem ich ihre Fähigkeit verbessere, in Echtzeit auf Ereignisse zu reagieren.
 
-Hier wird auf die Formel \ref{Gl1} verwiesen.
-
-\begin{align}
-x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl1}
-\end{align}
-\begin{align}
-x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl2}
-\end{align}
-
-\begin{lstlisting}[language=C++,name={1. Beispiel},label={sc:bsp:1}]
-#include <iostream>
-
-void SayHello(void)
-{
-    // Kommentar
-    cout << "Hello World!" << endl;
-}
-
-int main(int argc, char **argv)
-{
-    SayHello();
-    return 0;
-}
-\end{lstlisting}
-
-Literaturverweise sollten automatisch verwaltet werden, vor allem, wenn es viele Quellenverweise gibt.
-Beispiele sind  \cite{Ko05a}, \cite{Ko05b}, \cite{MiGo05}, \cite{TeGo14}, \cite{HuHa07}, \cite{HuZi10}, \cite{ZiKu07}, \cite{He07}, \cite{SIE11}, \cite{SIE14}, \cite{ISO98}, \cite{ATM11}, \cite{Hu11}, \cite{Po10}.
-Das verwendete Zitierformat (bzw.~das Format des Literaturverzeichnisses) ist entspechend der Vorgaben der Studiengänge zu wählen.
-Es wird dringend empfohlen, BibTeX~zu verwenden (wie in diesem Beispiel).
+- Anbindung eines Roboters über eine VARAN-Bus Schnittstelle:
+Ich plane, einen Roboter in mein System zu integrieren.
+Ich werde eine VARAN-Bus Schnittstelle verwenden, um eine schnelle und zuverlässige Kommunikation zwischen dem Roboter und dem Steuerungssystem zu gewährleisten.
 

### 2024-03-19 bkp.tex
kernel_processes.md, kernelshark, start thesis

--- 2024-03-19 Masterarbeit_Robotik_Pamuk.tex
@@
 \schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
-\outline{Abstract}
+\outline{\blindtext}
 \keywords{Echtzeit, Virtualisierung, Xenomai, VARAN}
@@
 %- Anhang (optional)
-\chapter{Introduction}
-
+\chapter{Einleitung}
+\blindmathpaper
 \clearpage
-\section{State of the art}
-
+\section{Stand der Technik}
+\blindtext
 \clearpage
-\section{Problem and task definition}
+\section{Problem- und Aufgabenstellung}
 In robotics, accurate and timely control is crucial to ensure precise movements and reliable interaction with the environment.
@@
 \clearpage
-\section{Objective}
+\section{Zielsetzung}
 
@@
 \clearpage
-\chapter{Methodology}
+\chapter{Methodik}
 
@@
 Ich werde eine VARAN-Bus Schnittstelle verwenden, um eine schnelle und zuverlässige Kommunikation zwischen dem Roboter und dem Steuerungssystem zu gewährleisten.
+
+\noindent Querverweise werden in \LaTeX{} automatisch erzeugt und verwaltet, damit sie leicht aktualisiert werden können. Hier wird zum Beispiel auf Abbildung \ref{Abb1} verwiesen.
+
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.5\linewidth]{img/buchruecken}
+\caption{Beispiel für die Beschriftung eines Buchrückens.}\label{Abb1}
+\end{figure}
+\begin{figure}[!htbp]
+\centering
+\includegraphics[width=0.5\linewidth]{img/buchruecken}
+\caption{2. Beispiel für die Beschriftung eines Buchrückens.}\label{Abb2}
+\end{figure}
+
+Und hier ist ein Verweis auf Tabelle \ref{tab1}. Das gezeigte Tabellenformat ist nur ein Beispiel. Tabellen können individuell gestaltet werden.
+
+\begin{table}[!htbp]
+\centering
+\caption{Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab1}
+\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
+Datum & Thema & Raum\\\hline
+20.08.2008 & Graphentheorie & HS 3.13\\
+01.10.2008 & Biomathematik & HS 1.05\\\hline
+\end{tabular}
+\end{table}
+\begin{table}[!htbp]
+\centering
+\caption{2. Semesterplan der Lehrveranstaltung \glqq Angewandte Mathematik\grqq.}\label{tab2}
+\begin{tabular}{| p{0.3\linewidth} | p{0.3\linewidth} | p{0.3\linewidth} |}\hline
+Datum & Thema & Raum\\\hline
+20.08.2008 & Graphentheorie & HS 3.13\\
+01.10.2008 & Biomathematik & HS 1.05\\\hline
+\end{tabular}
+\end{table}
+
+Hier wird auf die Formel \ref{Gl1} verwiesen.
+
+\begin{align}
+x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl1}
+\end{align}
+\begin{align}
+x = -\frac{p}{2}\pm\sqrt{\frac{p^2}{4}-q}\label{Gl2}
+\end{align}
+
+\begin{lstlisting}[language=C++,name={1. Beispiel},label={sc:bsp:1}]
+#include <iostream>
+
+void SayHello(void)
+{
+    // Kommentar
+    cout << "Hello World!" << endl;
+}
+
+int main(int argc, char **argv)
+{
+    SayHello();
+    return 0;
+}
+\end{lstlisting}
+
+Literaturverweise sollten automatisch verwaltet werden, vor allem, wenn es viele Quellenverweise gibt.
+Beispiele sind  \cite{Ko05a}, \cite{Ko05b}, \cite{MiGo05}, \cite{TeGo14}, \cite{HuHa07}, \cite{HuZi10}, \cite{ZiKu07}, \cite{He07}, \cite{SIE11}, \cite{SIE14}, \cite{ISO98}, \cite{ATM11}, \cite{Hu11}, \cite{Po10}.
+Das verwendete Zitierformat (bzw.~das Format des Literaturverzeichnisses) ist entspechend der Vorgaben der Studiengänge zu wählen.
+Es wird dringend empfohlen, BibTeX~zu verwenden (wie in diesem Beispiel).
 

### 2024-03-19 kernel-patch/check_smp_activity.sh
Deployed 2b71b7c1 with MkDocs version: 1.5.3

@@ -1,0 +1,12 @@
+#!/bin/bash
+for IRQ in /proc/irq/*; do
+    if [ -f "$IRQ/smp_affinity" ]; then
+        # Read the current smp_affinity
+        AFFINITY=$(cat "$IRQ/smp_affinity")
+        # Check if the bit for CPU17 is set
+        if (( (0x$AFFINITY & (1 << 17)) != 0 )); then
+            # Print the IRQ number
+            echo "${IRQ#/proc/irq/}"
+        fi
+    fi
+done

### 2024-03-19 resources/kernel/compare_2_files.py

@@ -35,4 +35,4 @@ def compare_files(file1, file2, output_file):
     print(f"Comparison results have been written to {output_file}")
 
 # Call the function with the file names
-compare_files("defconfig", "output.txt", "output_comparison.txt")
+compare_files("6.5.0", "6.8.0", "output_comparison.txt")

### 2024-03-19 clean.sh
Added zotero and clean.sh

@@ -1,0 +1,1 @@
+find . -type f \( -name "*.aux" -o -name "*.bbl" -o -name "*.blg" -o -name "*.idx" -o -name "*.ind" -o -name "*.lof" -o -name "*.lot" -o -name "*.out" -o -name "*.toc" -o -name "*.acn" -o -name "*.acr" -o -name "*.alg" -o -name "*.glg" -o -name "*.glo" -o -name "*.gls" -o -name "*.ist" -o -name "*.fls" -o -name "*.log" -o -name "*.fdb_latexmk" -o -name "*.lol" -o -name "*.synctex.gz" -o -name "*.bcf" -o -name "*.run.xml" \) -exec rm -f {} \;

### 2024-03-19 Masterarbeit_Robotik_Pamuk.tex
Added zotero and clean.sh
Added progress bar

@@
 
-The methodology of this thesis consists of several steps.
+In diesem Abschnitt werden sämtliche theoretischen Konzepte und Randbedingungen sowie praktische Methoden, die zur Erreichung der Zielsetzung dieser Masterarbeit beigetragen haben, detailliert beschrieben.
+
+Diese Masterarbeit wurde in Zusammenarbeit mit der Firma SIGMATEK erstellt.
+
 First, a comprehensive literature review is conducted to understand the current trends and challenges in real-time robot control.
@@
 \clearpage
-\chapter{Resultate}
+\chapter{Results}
 \clearpage
-\chapter{Diskussion}
+\chapter{Discussion}
 \clearpage
-\chapter{Zusammenfassung und Ausblick}
+\chapter{Summary and Outlook}
+I am using \cite{blackadarHistoricalReviewCauses2016} for citing.
 

### 2024-03-19 Masterarbeit_Robotik_Pamuk.tex
Ubuntu Pro

@@
 \chapter{Introduction}
+Diese Masterarbeit wurde in der Firma SIGMATEK GmbH \& Co KG erstellt.
+
+Firma SIGMATEK
+
+Yocto
+
+Salamander 4 - harte Echtzeit
+
+Xenomai
+
+QEMU
+
+Trace-cmd
+
+Kernelshark
+
+VARAN-Bus PCV-521
+
+Windows
+Ubuntu
+WSL
 
@@
 
-Diese Masterarbeit wurde in Zusammenarbeit mit der Firma SIGMATEK erstellt.
+Zu Beginn der Arbeit wurde eine ausführliche Analyse der einzelnen Virtualisierungsmöglichkeiten von Salamander 4 durchgeführt.
+Im Besonderen wurde hier die Virtualisierungsperformance von Ubuntu 22.04, Windows 10 und WSL unter QEMU verglichen.
 
@@
 Based on the literature study, a suitable virtualisation platform is selected.
+
 After the selection of the virtualisation platform, the robot control platform is implemented.

### 2024-03-21 masterthesis-documentation/docs/salamander4/trace-cmd/analysis/test/count_lines.py
kvm_exit plot

@@ -15,27 +15,30 @@ def count_lines_with_word(file_path, word, total_lines):
 # Usage
 file_path = 'kvm_exit_log.txt'  # replace with your file path
 total_lines = sum(1 for line in open(file_path))
-print(f"Total lines: {total_lines}.")
+print(f"Total exits: {total_lines}.")
 
 words = ['APIC_WRITE', 
-         'EXTERNAL_INTERRUPT', 
          'HLT', 
          'EPT_MISCONFIG',
          'PREEMPTION_TIMER',
-         'PAUSE_INSTRUCTION',
-         'EPT_VIOLATION',
+         'EXTERNAL_INTERRUPT', 
          'IO_INSTRUCTION',
          'EOI_INDUCED',
-         'MSR_READ',
-         'CPUID'    
+         'EPT_VIOLATION',
+         'PAUSE_INSTRUCTION',
+         'CPUID',   
+         'MSR_READ'
          ]  # replace with the words you want to search for
 
-word_counts = []
+counts = []
 for word in words:
     total_lines, word_lines = count_lines_with_word(file_path, word, total_lines)
-    print(f"The word '{word}' appears in {word_lines} lines.")
-    print(f"Remaining lines: {total_lines}.")
-    word_counts.append(word_lines)
+    print(f"'{word}': {word_lines} times")
+    counts.append((word, word_lines))
+
+# Sort words and counts by count in descending order
+counts.sort(key=lambda x: x[1], reverse=True)
+words, word_counts = zip(*counts)  # unzip the sorted pairs
 
 # Set the figure size
 plt.figure(figsize=(10, 6))
@@ -50,7 +53,7 @@ for bar in bars:
 
 plt.xlabel('Reasons')
 plt.ylabel('Count')
-plt.title('Word Counts in File')
+plt.title('Exit Reason Count')
 plt.xticks(rotation=90, fontsize='small')  # Adjust font size for readability
 
 # Adjust bottom margin
@@ -60,5 +63,7 @@ plt.subplots_adjust(bottom=0.4)  # Increase the bottom margin
 #plt.xlim([0, len(words)])  # Set x-axis length
 plt.ylim([0, max(word_counts)*1.2])  # Set y-axis length
 
-plt.show()
+# Save the figure
+plt.savefig('plot.png', bbox_inches='tight')
 
+plt.show()

### 2024-03-21 Masterarbeit_Robotik_Pamuk.tex
kvm_exit plot
add xenomai to .tex

@@
 The main objective of this work is to create a real-time robot control platform that integrates Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
-By combining these technologies, a platform is to be created that provides real-time functions and efficient communication systems.
+
 This integration is expected to lead to a significant improvement in the response time and reliability of robot applications.
@@
 \clearpage 
-\chapter{Hauptteil}
+\chapter{Salamander 4}
+Salamander 4 ist das proprietäre Betriebssystem. Es basiert auf der Linux-Architektur und integriert zusätzlich Xenomai, eine Echtzeit-Entwicklungsumgebung \cite{XenomaiXenomai}.
+
+Durch die Kombination dieser Technologien bietet Salamander OS sowohl die Flexibilität und Offenheit von Linux als auch die Echtzeitfähigkeiten von Xenomai, was es zu einer idealen Lösung für unsere spezifischen Anforderungen macht.”
+
 \clearpage

### 2024-03-21 Masterarbeit_Robotik_Pamuk.tex
.gitignore
windows pdf to docs

--- 2024-03-19 Masterarbeit_Robotik_Pamuk.tex
@@
 \usepackage[T1]{fontenc}
-
 %
@@
 % Bei Bedarf bitte hier die Syntax-Highlightings anpassen
+%
+\usepackage{float}
+%
+% Platzierungsoptionen 
 %
@@
 \clearpage 
-\chapter{Hauptteil}
+\chapter{Salamander 4 Operating System}
+Salamander 4 ist das proprietäre Betriebssystem.
+Es basiert auf der der Linux-Version 5.15.94 und integriert zusätzlich Xenomai 3.2, eine Echtzeit-Entwicklungsumgebung \cite{Xenomai}.
+Es handelt sich um ein 64-Bit-System, was auf die x86\_64-Architektur hinweist.
+Das Echtzeitverhalten wird durch die Verwendung von Symmetric Multi-Processing (SMP) und Preemptive Scheduling (PREEMPT) erreicht.
+Darüber hinaus nutzt es IRQPIPE, um Interrupts in einer Weise zu verarbeiten, die die Echtzeitanforderungen des Systems erfüllt.
+Xenomai besteht aus 3 Teilen.
+Diese sind der Tabelle \ref{tab:what_is_xenomai} zu entnehmen
+
+\begin{table}[!h]
+    \centering
+    \setlength{\tabcolsep}{0.5em} % for the horizontal padding
+    {\renewcommand{\arraystretch}{1.2}% for the vertical padding
+    \begin{tabular}{|c|p{0.6\textwidth}|}
+        \hline
+        \textbf{Teil} & \textbf{Beschreibung} \\ \hline
+        i-pipe  & Kernelerweiterung für das Domain-Konzept \\ \hline
+        Xenomai Kernel & Benutzt die i-pipe, und hängt sich als root-Domain ein \\ \hline
+        Xenomai User &  Programme (LRT) verwenden diese Bibliothek, um Xenomai Funktionen verwenden zu können. \\ \hline
+    \end{tabular}}
+  \end{table}
+
+\noindent Xenomai beruht auf einem Domain-Konzept, das bedeutet, dass alle IRQ an die erste Domain gesendet werden. (root – Domain / Xenomai)
+Nur wenn diese nichts mehr zu tun hat, dann darf die 2. Domain arbeiten.
+Das bedeutet, erst wenn alle Xenomai Task in einem Wartezustand sind, arbeiten die Linux-Tasks.
+
+\noindent In der Regel unterbricht der Prozessor beim IRQ-Handling seine aktuellen Aktivitäten, um einen Interrupt zu bearbeiten, während die IRQ-Behandlung von Xenomai einen Interrupt-Pipeline-Mechanismus verwendet, der das gleichzeitige Abrufen und Vorbereiten eines anderen Interrupts ermöglicht, während ein Interrupt bearbeitet wird, was die Leistung verbessert und die Latenzzeit verringert.
+
+\noindent Was Xenomai4 von seinem Vorgänger Xenomai3 unterscheidet, ist die vollständige Neugestaltung der Ausführungsphase mit hoher Priorität.
+Dies geschah aus Gründen der Portabilität und Wartungsfreundlichkeit: I-pipe - die zweite Iteration der ursprünglichen Adeos-Interrupt-Pipeline - wurde vollständig durch Dovetail ersetzt.
+
+\begin{table}[H]
+    \centering
+    \caption[Kurz]{Lang}
+    \label{tab:domain_specific_functions}
+    \begin{tabular}{|c|c|}
+        \hline
+        \textbf{Xenomai spezifische Funktionen} & \textbf{Linux spezifische Funktionen} \\ \hline
+        Tasks                                   & Dateizugriffe                           \\ \hline
+        Mutexes, Semaphoren, Events             & Netzwerk                                \\ \hline
+    \end{tabular}
+\end{table}
+
+\noindent Ein Aufruf dieser Funktionen erfordert die entsprechende Domain. Wenn der Task in der falschen Domain läuft, dann wird ein Domain-Wechsel forciert.
+Ein Domainwechsel von Xenomai nach Linux geht relativ einfach.
+Aber der Wechsel von Linux nach Xenomai braucht Unterstützung, und dafür ist die Hilfe des Gatekeepers notwendig.
+Das bedeutet, der Gatekeeper hilft einem Task von Linux nach Xenomai zu wechseln.
+
+\section{Task priorities}
+Es gibt grundsätzlich 4 Gruppen
+
+\begin{table}[ht]
+    \centering
+    \caption{Übersicht der Prioritätsgruppen und ihrer Beziehungen}
+    \label{tab:priorities}
+    \begin{tabular}{|c|c|}
+        \hline
+        \textbf{Prioritätsgruppe} & \textbf{Bereich}  \\ \hline
+        Xenomai Priorität & 0 bis 99  \\ \hline
+        Linux RT Priorität & 1 bis 99 \\ \hline
+        Linux (Nice Level) Priorität & -20 bis 19  \\ \hline
+        RTK Priorität & 0 bis 14 \\ \hline
+    \end{tabular}
+
+\end{table}
+
+\section{Memory Management}
+Es gibt verschiedene Speicherbereiche
+
+Linux/System/Programm Speicher
+	Der Speicher, den Linux und Programme belegt haben.
+	Dieser Speicher ist intern in viele Teile aufgeteilt. (DMA, …)
+
+LRT-Heap Speicher
+	Speicher den der LRT verwendet, oder welcher über ein CIL Funktionen
+	angefordert wird.
+
+App Heap, App Code, …
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.8\columnwidth]{img/RAM_Memory_management.png}
+    \caption[Kurz]{Lang}
+    \label{fig:memory_management}
+  \end{figure}
+
 \clearpage

### 2024-03-25 Masterarbeit_Robotik_Pamuk.tex
Literature search
Draft: Introduction, additional papers

@@
 \usepackage{float}
+\usepackage{comment}
 %
@@
 \chapter{Introduction}
+In today's industrial production and automation, robot systems are well established and of crucial importance.
+Robots must react to their environment and perform time-critical tasks within strict time constraints.
+Delays or errors can have catastrophic consequences in some cases.
+Traditional operating systems, such as Windows or Linux, are often not suitable for these types of real-time requirements as they cannot guarantee deterministic execution times.
+Therefore, real-time operating systems are required that are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
+
+\bigskip \noindent The core component of an RTOS that enables real-time capabilities is the kernel.
+The kernel is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior.
+It employs preemptive scheduling mechanisms to allow high-priority tasks to preempt lower-priority tasks, ensuring that time-critical tasks are not delayed.
+The kernel also implements priority-based scheduling algorithms, such as Rate Monotonic Scheduling (RMS) or Earliest Deadline First (EDF), to schedule tasks based on their priorities and timing constraints.
+Additionally, RTOS kernels are designed to minimize interrupt latency, which is crucial for real-time applications that require immediate response to external events.
+
+\bigskip \noindent In these RTOS systems, task scheduling is based on so-called priority-based preemptive scheduling.
+Each task in a software application is assigned a priority.
+A higher priority means that a faster response is required.
+Pre-emptive task scheduling ensures a very fast response.
+Preemptive means that the scheduler can stop a currently running task at any point if it recognises that another task needs to be executed immediately.
+The basic rule on which priority-based preemptive scheduling is based is that the task with the highest priority that is ready to run is always the task that must be executed.
+So if both a task with a lower priority and a task with a higher priority are ready to run, the scheduler ensures that the task with the higher priority runs first.
+The lower priority task is only executed once the higher priority task has been processed.
+Real-time systems are usually categorised as either soft or hard real-time systems.
+The difference lies exclusively in the consequences of a violation of the time limits.
+
+\bigskip \noindent Hard real-time is when the system stops operating if a deadline is missed, which can have catastrophic consequences.
+These consequences include endangering people, damaging machinery, impacting the health and integrity of the environment.
+Typical examples of this are some control systems in aeroplanes or cars, e.g.
+internal combustion engines.
+
+\bigskip \noindent Soft real-time exists when a system continues to function even if it cannot perform the tasks within a specified time.
+If the system has missed the deadline, this has no critical consequences.
+The system continues to run, although it does so with undesirably lower output quality.
+Examples of this are multimedia systems where the occasional deviation from a playback rate of 25 frames/sec.
+leads to stuttering or similar.
+
+\begin{comment}
+In der heutigen industriellen Fertigung und Automatisierung sind Robotersysteme sehr etabliert und von entscheidender Bedeutung.
+Roboter müssen auf ihre Umgebung reagieren und zeitkritische Aufgaben innerhalb strenger Zeitvorgaben ausführen.
+Verzögerungen oder Fehler können in einigen Fällen katastrophale Konseqenzen haben.
+Traditionelle Betriebssysteme, wie Windows oder Linux sind für diese Art von Echtzeitanforderungen oft nicht geeignet, da sie keine deterministischen Ausführungszeiten garantieren können.
+Daher sind Echtzeit-Betriebssysteme nötig, die speziell dafür konzipiert sind, auf Ereignisse innerhalb fester Zeitschranken zu reagieren und Prozesse mit hoher Priorität bevorzugt auszuführen.
+Bei diesen RTOS-Systemen basiert das Task-Scheduling auf dem sogenannten prioritätsbasierten präemptiven Scheduling.
+Jeder Task in einer Software-Applikation ist eine Priorität zuzuweisen.
+Eine höhere Priorität bedeutet, dass eine schnellere Reaktion erforderlich ist.
+Durch das präemptive Task-Scheduling wird eine sehr schnelle Reaktion sichergestellt.
+Präemptiv bedeutet, dass der Scheduler eine gerade laufende Task an jeder Stelle anhalten kann, wenn er erkennt, dass eine andere Task sofort ausgeführt werden muss.
+Die Grundregel, auf der das prioritätsbasierte präemptive Scheduling basiert, besagt, dass die ausführbereite Task mit der höchsten Priorität immer die Task ist, die ausgeführt werden muss.
+Wenn also sowohl eine Task mit niedriger als auch eine Task mit höherer Priorität ausführbereit sind, sorgt der Scheduler dafür, dass zuerst die Task mit der höheren Priorität läuft.
+Die Task mit niedrigerer Priorität wird erst dann ausgeführt, wenn die höherpriore Task verarbeitet wurde.
+
+\noindent Echtzeitsysteme werden in der Regel entweder als weiches oder als hartes Echtzeitsystem kategorisiert. Der Unterschied liegt ausschließlich in den Folgen einer Verletzung der Zeitschranken.
+
+\noindent Harte Echtzeit liegt vor, wenn das System seinen Betrieb einstellt, wenn ein Termin verpasst wurde, was katastrophale Folgen haben kann.
+Zu diesen Folgen zählen die Gefährdung von Menschen, die Beschädigung von Maschinen, also Auswirkungen auf Gesundheit und Unversehrtheit der Umgebung.
+Typische Beispiele hierfür sind einige Steuerungssysteme im Flugzeug oder im Auto, z.B.
+bei der Verbrennungsmaschine.
+
+\noindent Weiche Echtzeit besteht, wenn ein System weiter funktionsbereit ist, selbst wenn es die Aufgaben innerhalb einer festgelegten Zeit nicht ausführen kann.
+Wenn das System den Termin verpasst hat, hat dies keine kritischen Konsequenzen.
+Das System läuft weiter, obwohl es dies mit unerwünscht geringerer Ausgabenqualität tut.
+Beispiele hierfür sind Multimediasysteme, bei denen das gelegentlich Abweichen von einer Abspielrate von 25 Bildern/sek.
+zu einem Ruckeln o.ä.
+führt.
+\end{comment}
+
+\clearpage
+\section{Relevant SIGMATEK information}
 Diese Masterarbeit wurde in der Firma SIGMATEK GmbH \& Co KG erstellt.
-
-Firma SIGMATEK
-
-Yocto
-
-Salamander 4 - harte Echtzeit
-
-Xenomai
-
-QEMU
-
-Trace-cmd
-
-Kernelshark
-
-VARAN-Bus PCV-521
-
-Windows
-Ubuntu
-WSL
 
@@
 \section{State of the art}
+\clearpage
 
-\clearpage
 \section{Problem and task definition}
@@
 \clearpage 
-\chapter{Salamander 4 Operating System}
+\chapter{Salamander Operating System}
 Salamander 4 ist das proprietäre Betriebssystem.
@@
     \centering
+    \caption[Kurz]{Lang}
+    \label{tab:what_is_xenomai}
     \setlength{\tabcolsep}{0.5em} % for the horizontal padding
@@
 \chapter{Results}
+Firma SIGMATEK
+
+Yocto
+
+Salamander 4 - harte Echtzeit
+
+Xenomai
+
+QEMU
+
+Trace-cmd
+
+Kernelshark
+
+VARAN-Bus PCV-521
+
+Windows
+Ubuntu
+WSL
 \clearpage

### 2024-03-25 Masterarbeit_Robotik_Pamuk.tex
PDF structure and links

@@
 \clearpage
-\section{Relevant SIGMATEK information}
+\section{Thesis Context}
 Diese Masterarbeit wurde in der Firma SIGMATEK GmbH \& Co KG erstellt.
@@
 Firma SIGMATEK
-
 Yocto
-
 Salamander 4 - harte Echtzeit
-
 Xenomai
-
 QEMU
-
 Trace-cmd
-
 Kernelshark
-
 VARAN-Bus PCV-521
-
 Windows

### 2024-03-25 Masterarbeit_Robotik_Pamuk.tex
pics

@@
 
-\begin{comment}
-In der heutigen industriellen Fertigung und Automatisierung sind Robotersysteme sehr etabliert und von entscheidender Bedeutung.
+%In der heutigen industriellen Fertigung und Automatisierung sind Robotersysteme sehr etabliert und von entscheidender Bedeutung.
 Roboter müssen auf ihre Umgebung reagieren und zeitkritische Aufgaben innerhalb strenger Zeitvorgaben ausführen.
@@
 
-\noindent Echtzeitsysteme werden in der Regel entweder als weiches oder als hartes Echtzeitsystem kategorisiert. Der Unterschied liegt ausschließlich in den Folgen einer Verletzung der Zeitschranken.
+%\noindent Echtzeitsysteme werden in der Regel entweder als weiches oder als hartes Echtzeitsystem kategorisiert. Der Unterschied liegt ausschließlich in den Folgen einer Verletzung der Zeitschranken.
 
-\noindent Harte Echtzeit liegt vor, wenn das System seinen Betrieb einstellt, wenn ein Termin verpasst wurde, was katastrophale Folgen haben kann.
+%\noindent Harte Echtzeit liegt vor, wenn das System seinen Betrieb einstellt, wenn ein Termin verpasst wurde, was katastrophale Folgen haben kann.
 Zu diesen Folgen zählen die Gefährdung von Menschen, die Beschädigung von Maschinen, also Auswirkungen auf Gesundheit und Unversehrtheit der Umgebung.
@@
 
-\noindent Weiche Echtzeit besteht, wenn ein System weiter funktionsbereit ist, selbst wenn es die Aufgaben innerhalb einer festgelegten Zeit nicht ausführen kann.
+%\noindent Weiche Echtzeit besteht, wenn ein System weiter funktionsbereit ist, selbst wenn es die Aufgaben innerhalb einer festgelegten Zeit nicht ausführen kann.
 Wenn das System den Termin verpasst hat, hat dies keine kritischen Konsequenzen.
@@
 führt.
-\end{comment}
 
@@
 \section{State of the art}
+
 \clearpage
@@
 
+  Eine LASAL CPU besteht aus den folgenden Software-Modulen:
+  \begin{itemize}
+    \item Operating system
+    \item Loader
+    \item Hardware-Klassen
+  \end{itemize}
+
+  Die Schnittstelle zwischen den einzelnen Modulen wird in Abbildung \ref{fig:lasal_cpu} durch einen Pfeil gekennzeichnet.
+
+  \begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.6\columnwidth]{img/Software-Struktur einer LASAL CPU.png}
+    \caption[Kurz]{Lang}
+    \label{fig:lasal_cpu}
+  \end{figure}
+
 \clearpage
+
+\chapter{Real-Time Virtualisation}
+
+Salamander 4
+
+Trace-cmd was used for tracing the Linux kernel.
+It enables kernel events and function calls to be recorded and analysed in real time.
+With trace-cmd, developers and system administrators can gain detailed insights into system behaviour and identify bottlenecks or performance problems.
+
+\section{Kernelshark}
+
 \chapter{Results}
-Firma SIGMATEK
-Yocto
-Salamander 4 - harte Echtzeit
-Xenomai
-QEMU
-Trace-cmd
-Kernelshark
-VARAN-Bus PCV-521
-Windows
-Ubuntu
-WSL
+Firma SIGMATEK\newline
+Yocto\newline
+Salamander 4\newline
+Xenomai\newline
+QEMU\newline
+Trace-cmd\newline
+Kernelshark\newline
+VARAN-Bus PCV-521\newline
+Windows\newline
+Ubuntu\newline
+WSL\newline
 \clearpage

### 2024-03-25 Masterarbeit_Robotik_Pamuk.tex
trace-cmd and kernelshark

@@
 \usepackage{comment}
+\usepackage{csquotes}
 %
@@
 
+\bigskip \noindent Trace-cmd was used for tracing the Linux kernel.
+It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
+Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency \cite{Tracecmd}.
+
+\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool.
+It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
+By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed \cite{KernelShark}.
+
+\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel.
+It enables low-latency and deterministic execution of time-critical tasks.
+Xenomai introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
+A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
+The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times\cite{Xenomai}.
+
 \clearpage 
@@
 
-\begin{figure}[!h]
+\begin{figure}[H]
     \centering
@@
     \centering
-    \includegraphics[width=0.6\columnwidth]{img/Software-Struktur einer LASAL CPU.png}
+    \includegraphics[width=0.5\columnwidth]{img/Software-Struktur einer LASAL CPU.png}
     \caption[Kurz]{Lang}
@@
 
-\chapter{Real-Time Virtualisation}
+\chapter{Comparison of Real-Time Latency}
+
+\section{Salamander 4 Bare Metal}
+
+\section{Salamander 4 Virtualisation}
 
 Salamander 4
-
-Trace-cmd was used for tracing the Linux kernel.
-It enables kernel events and function calls to be recorded and analysed in real time.
-With trace-cmd, developers and system administrators can gain detailed insights into system behaviour and identify bottlenecks or performance problems.
 
@@
 \chapter{Summary and Outlook}
-I am using \cite{blackadarHistoricalReviewCauses2016} for citing.
 

### 2024-03-26 Masterarbeit_Robotik_Pamuk.tex
plot python

@@
 %
-\usepackage[backend=biber, style=numeric]{biblatex}
+\usepackage[backend=biber, style=numeric, sorting=none]{biblatex}
 \addbibresource{Literatur.bib}
-
 %
@@
 %
+% own 
 \usepackage{float}
@@
 \usepackage{csquotes}
+\usepackage{pgfplots}
 %
@@
 \usepackage{blindtext}
-
 %
@@
 
-- Verbesserung der Reaktionszeit und Zuverlässigkeit von Roboteranwendungen: 
-Mein Hauptziel ist es, herauszufinden, wie die Integration von Echtzeitfunktionen und effizienten Kommunikationssystemen die Reaktionszeit und Zuverlässigkeit von Roboteranwendungen verbessern kann.
-Ich strebe an, die Leistung und Zuverlässigkeit meiner Roboteranwendungen zu verbessern, indem ich ihre Fähigkeit verbessere, in Echtzeit auf Ereignisse zu reagieren.
-
 - Anbindung eines Roboters über eine VARAN-Bus Schnittstelle:
@@
 A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
-The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times\cite{Xenomai}.
+The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times \cite{Xenomai}.
 
@@
 \chapter{Salamander Operating System}
-Salamander 4 ist das proprietäre Betriebssystem.
-Es basiert auf der der Linux-Version 5.15.94 und integriert zusätzlich Xenomai 3.2, eine Echtzeit-Entwicklungsumgebung \cite{Xenomai}.
-Es handelt sich um ein 64-Bit-System, was auf die x86\_64-Architektur hinweist.
-Das Echtzeitverhalten wird durch die Verwendung von Symmetric Multi-Processing (SMP) und Preemptive Scheduling (PREEMPT) erreicht.
-Darüber hinaus nutzt es IRQPIPE, um Interrupts in einer Weise zu verarbeiten, die die Echtzeitanforderungen des Systems erfüllt.
-Xenomai besteht aus 3 Teilen.
-Diese sind der Tabelle \ref{tab:what_is_xenomai} zu entnehmen
+
+Salamander 4 is the proprietary operating system of SIGMATEK.
+It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment \cite{Xenomai}.
+Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture.
+The real-time behaviour is achieved through the use of Symmetric Multi-Processing (SMP) and Preemptive Scheduling (PREEMPT).
+In addition, it utilises IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
+The output of the command \texttt{uname -a} can be observed in code \ref{lst:uname_a}.
+
+\vspace{1em}
+
+\begin{lstlisting}[caption={Ein Beispiel für einen Linux-Befehl}, label={lst:uname_a}]
+root@sigmatek-core2:~# uname -a
+Linux sigmatek-core2 5.15.94 #1 SMP PREEMPT IRQPIPE Tue Feb 14 18:18:05 UTC 2023 x86_64 GNU/Linux
+\end{lstlisting}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\columnwidth]{img/kvm_exits.png}
+    \caption[Kurz]{Lang}
+    \label{fig:kvm_exits}
+  \end{figure}
+
+Xenomai consists of 3 parts. These can be found in the table \ref{tab:what_is_xenomai}.
 
@@
 \chapter{Comparison of Real-Time Latency}
+We tried to measure the latencies under 2 conditions:
+IDLE and CPU-STRESSED.
+In the IDLE condition, in order to create such test environment, some non-essential user processes were eliminated.
+While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
+The commands to kill these processes are written in the test script.
+
+\begin{table}[H]
+    \centering
+    \caption[Kurz]{Lang}
+    \label{tab:kernel_configuration}
+    \begin{tabular}{|c|c|}
+        \hline
+        \textbf{Kernel} & \textbf{Patches} \\ \hline
+        . & .  \\ \hline
+        . & .  \\ \hline
+    \end{tabular}
+\end{table}
 
@@
 \section{Salamander 4 Virtualisation}
+In Xenomai 3, you need:
 
-Salamander 4
+The Linux kernel with the Dovetail interface, which allows the kernel to handle real-time tasks with low latency.
+The Cobalt real-time core, which is responsible for scheduling and executing real-time tasks.
+In Xenomai 4, you need:
+
+The Linux kernel with the EVL interface, which, like Dovetail, enables the kernel to handle real-time tasks with low latency.
+The libevl library, which provides a user interface to the EVL core, allowing applications to use the real-time capabilities provided by EVL.
+So, while both versions require a Linux kernel with a real-time interface (Dovetail for Xenomai 3, EVL for Xenomai 4), Xenomai 3 uses a real-time core (Cobalt), while Xenomai 4 uses a library (libevl) to provide applications with access to its real-time capabilities.
 
 \section{Kernelshark}
+
+\begin{lstlisting}[language=C++,name={1. Beispiel},label={sc:bsp:1}]
+    #include <iostream>
+
+    void SayHello(void)
+    {
+        // Kommentar
+        cout << "Hello World!" << endl;
+    }
+
+    int main(int argc, char **argv)
+    {
+        SayHello();
+        return 0;
+    }
+    \end{lstlisting}
 

### 2024-03-26 masterthesis-documentation/docs/sigmatek/xenomai/print_numbers.py

@@ -1,0 +1,9 @@
+# Open the file in write mode
+with open("numbers.txt", "w") as file:
+    # Write numbers from 1 to 600
+    for number in range(1, 601):
+        file.write(str(number) + "\n")
+
+# Print a success message
+print("The numbers from 1 to 600 are successfully saved to numbers.txt file.")
+

### 2024-03-26 sigmatek/trace-cmd/analysis/test/kvm_exit_number.py
Deployed 5643f3f with MkDocs version: 1.5.3

@@ -44,7 +44,7 @@ words, word_counts = zip(*counts)  # unzip the sorted pairs
 plt.figure(figsize=(10, 6))
 
 # Plotting
-bars = plt.bar(words, word_counts)
+bars = plt.bar(words, word_counts, color='blue')  # Added color='blue'
 
 # Labeling the bars with their counts
 for bar in bars:
@@ -64,6 +64,8 @@ plt.subplots_adjust(bottom=0.4)  # Increase the bottom margin
 plt.ylim([0, max(word_counts)*1.2])  # Set y-axis length
 
 # Save the figure
-plt.savefig('plot.png', bbox_inches='tight')
+plt.savefig('../../../../../../img/kvm_exits.png', bbox_inches='tight')
+plt.savefig('kvm_exits.png', bbox_inches='tight')
 
-plt.show()
+# Print a success message
+print("The plot was successfully saved to kvm_exits.png")

### 2024-03-26 sigmatek/xenomai/latency_no_taskset.py

@@ -1,0 +1,27 @@
+import matplotlib.pyplot as plt
+
+# Data points
+x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,445,446,447,448,449,450,451,452,453,454,455,456,457,458,459,460,461,462,463,464,465,466,467,468,469,470,47
[…]

### 2024-03-26 sigmatek/xenomai/latency_taskset.py

@@ -1,0 +1,27 @@
+import matplotlib.pyplot as plt
+
+# Data points
+x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,445,446,447,448,449,450,451,452,453,454,455,456,457,458,459,460,461,462,463,464,465,466,467,468,469,470,471
[…]

### 2024-03-26 masterthesis-documentation/docs/sigmatek/xenomai/latency_no_taskset.py
Add Python to Readme

@@ -18,8 +18,8 @@ plt.xlim([0, 600])
 plt.ylim([0, 5000])
 
 # Save the plot to a file
-plt.savefig("latency.png")
-plt.savefig("../../../../img/latency.png")
+plt.savefig("latency_no_taskset.png")
+plt.savefig("../../../../img/latency_no_taskset.png")
 
 # Print a success message
 print("The plot was successfully saved to latency.png")

### 2024-03-27 sigmatek/xenomai/latency_hardware.py
Deployed 7e4b31c with MkDocs version: 1.5.3

@@ -1,0 +1,28 @@
+import matplotlib.pyplot as plt
+
+# Data points
+x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,41
[…]

### 2024-03-27 sigmatek/xenomai/latency_no_taskset.py

@@ -2,6 +2,7 @@ import matplotlib.pyplot as plt
 
 # Data points
 x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,445,446,447,448,449,450,451,452,453,454,455,456,457,458,459,460,461,462,463,464,465,466,467,468,469,4
[…]

### 2024-03-27 Masterarbeit_Robotik_Pamuk.tex
Add Python to Readme
Insert Images and write thesis

@@
 \clearpage
-\section{Thesis Context}
-Diese Masterarbeit wurde in der Firma SIGMATEK GmbH \& Co KG erstellt.
+\section{Application Context}
+This master's thesis was written at SIGMATEK GmbH \& Co KG \cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.
 
@@
 
-\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool.
+\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool \cite{KernelShark}.
 It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
-By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed \cite{KernelShark}.
+By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
 
@@
 A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
-The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times \cite{Xenomai}.
+The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times \cite{XenomaiXenomai}.
 
@@
 Salamander 4 is the proprietary operating system of SIGMATEK.
-It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment \cite{Xenomai}.
+It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment \cite{XenomaiXenomai}.
 Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture.
@@
 \vspace{1em}
-
-\begin{lstlisting}[caption={Ein Beispiel für einen Linux-Befehl}, label={lst:uname_a}]
+\begin{minipage}{0.95\columnwidth}
+\begin{lstlisting}[language=bash,name={System information},label={lst:uname_a}]
 root@sigmatek-core2:~# uname -a
@@
 \end{lstlisting}
+\end{minipage}
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\columnwidth]{img/kvm_exits.png}
-    \caption[Kurz]{Lang}
-    \label{fig:kvm_exits}
-  \end{figure}
-
-Xenomai consists of 3 parts. These can be found in the table \ref{tab:what_is_xenomai}.
+Xenomai consists of 3 parts. These can be found in the Table \ref{tab:what_is_xenomai}.
 
@@
     \centering
-    \caption[Kurz]{Lang}
+    \caption[Xenomai architecture]{Xenomai architecture}
     \label{tab:what_is_xenomai}
@@
     \centering
-    \caption[Kurz]{Lang}
+    \caption[Domain specific functions]{Domain specific functions}
     \label{tab:domain_specific_functions}
@@
     \centering
-    \caption{Übersicht der Prioritätsgruppen und ihrer Beziehungen}
+    \caption{Overview of the priority groups and their relationships}
     \label{tab:priorities}
@@
     \includegraphics[width=0.8\columnwidth]{img/RAM_Memory_management.png}
-    \caption[Kurz]{Lang}
+    \caption[Memory Management]{Memory Management}
     \label{fig:memory_management}
@@
     \centering
-    \includegraphics[width=0.5\columnwidth]{img/Software-Struktur einer LASAL CPU.png}
-    \caption[Kurz]{Lang}
+    \includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
+    \caption[LASAL CPU]{LASAL CPU}
     \label{fig:lasal_cpu}
@@
 \chapter{Comparison of Real-Time Latency}
-We tried to measure the latencies under 2 conditions:
-IDLE and CPU-STRESSED.
-In the IDLE condition, in order to create such test environment, some non-essential user processes were eliminated.
+In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4.
+For this purpose, the latency tool of the Xenomai test suite was used.
+The latency was measured under two conditions, idle and CPU-stressed.
+The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.
+
+Vorgehensweise von \cite{linPerformanceEvaluationXenomai}
+
+%In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
 While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
@@
     \centering
-    \caption[Kurz]{Lang}
+    \caption[Kernel and Patches]{Kernel and Patches}
     \label{tab:kernel_configuration}
@@
 
+After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
+
 \section{Salamander 4 Bare Metal}
+Salamander 4 Bare Metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system, including
 
 \section{Salamander 4 Virtualisation}
+In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
+It was developped using Yocto, an open source project that allows customised Linux distributions to be created for embedded systems \cite{WelcomeYoctoProject}.
+The virtualisation runs in a QEMU environment, which is an open source tool for hardware virtualisation \cite{QEMU}.
+With the help of the script depicted in code \ref{lst:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
+This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host.
+Upon generating the necessary files, Yocto generates a QEMU folder with the following components shown in code \ref{lst:ls}.
+
+\vspace{1em}
+\begin{minipage}{\linewidth}
+    \begin{lstlisting}[language=bash,name={Contents of QEMU folder for Salamander 4},label={lst:ls}]
+    sigma_ibo@localhost:~/Desktop/salamander-image$ ls -1
+    bzImage
+    drive-c
+    ovmf.code.qcow2
+    qemu_def_bkp.sh
+    qemu_def.sh
+    salamander-image-sigmatek-core2.ext4
+    stek-drive-c-image-sigmatek-core2.tar.gz
+    vmlinux
+    \end{lstlisting}
+    \end{minipage}
+
+\vspace{1em}
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[language=bash,name={QEMU script for starting Salamander 4 virtualisation},label={lst:qemu_def}]
+    #!/bin/sh
+
+    if  [ ! -d drive-c/ ]; then
+            echo "Filling drive-c/"
+            mkdir drive-c/
+            tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
+    fi
+
+    exec  qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
+    -m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
+    -append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable" \
+    -net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
+    -fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
+    -device vhost-vsock-pci,guest-cid=3,id=vsock0 \
+    -drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
+    -no-reboot -nographic
+\end{lstlisting}
+\end{minipage}
+
+Here is a description of the used components:
+\begin{itemize}
+    \item \textbf{bzImage}: Compressed Linux kernel image, loaded by QEMU at system start.
+    \item \textbf{ovmf.code.qcow2}: Firmware file for QEMU, enables UEFI boot process.
+    \item \textbf{qemu\_def.sh}: Shell script, starts QEMU with correct parameters to boot Salamamder 4 OS.
+    \item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Archive containing files for C drive, unpacked and copied to drive-c/ directory by qemu\_def.sh script.
+    \item \textbf{drive-c}: Directory, serves as C drive for QEMU system, created and filled by qemu\_def.sh script..
+    \item \textbf{salamander-image-sigmatek-core2.ext4}: Root file system for Salamander 4 OS, used as hard drive for QEMU system.
+    \item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging, contains debugging symbols not present in bzImage.
+\end{itemize}
+
+ %   \begin{table}[ht]
+ %       \centering
+ %       \caption{QEMU script components}
+ %       \label{tab:qemu_components}
+ %       \begin{tabular}{|l|p{8cm}|}
+ %           \hline
+ %           \textbf{File} & \textbf{Description} \\
+ %           \hline
+ %           bzImage & Compressed Linux kernel image, loaded by QEMU at system start. \\
+ %           \hline
+ %           ovmf.code.qcow2 & Firmware file for QEMU, enables UEFI boot process. \\
+ %           \hline
+ %           qemu\_def.sh & Shell script, starts QEMU with correct parameters to boot Salamamder 4 OS. \\
+ %           \hline
+ %           stek-drive-c-image-sigmatek-core2.tar.gz & Archive containing files for C drive, unpacked and copied to drive-c/ directory by qemu\_def.sh script. \\
+ %           \hline
+ %           drive-c & Directory, serves as C drive for QEMU system, created and filled by qemu\_def.sh script. \\
+ %           \hline
+ %           salamander-image-sigmatek-core2.ext4 & Root file system for Salamander 4 OS, used as hard drive for QEMU system. \\
+ %           \hline
+ %           vmlinux & Uncompressed Linux kernel image, typically used for debugging, contains debugging symbols not present in bzImage. \\
+ %           \hline
+ %       \end{tabular}
+ %   
+ %   \end{table}
+
+ Figure \ref{fig:latency_hardware}
+ \begin{figure}[H]
+   \centering
+   \includegraphics[width=0.8\columnwidth]{img/latency_hardware.png}
+   \caption[Latency hardware]{Latency hardware}
+   \label{fig:latency_hardware}
+ \end{figure}
+
+ Figure \ref{fig:latency_no_taskset}
+
+ \begin{figure}[H]
+     \centering
+     \includegraphics[width=0.8\columnwidth]{img/latency_no_taskset.png}
+     \caption[Latency no taskset]{Latency no taskset}
+     \label{fig:latency_no_taskset}
+   \end{figure}
+
+   Figure \ref{fig:latency_taskset}
+   \begin{figure}[H]
+     \centering
+     \includegraphics[width=0.8\columnwidth]{img/latency_taskset.png}
+     \caption[Latency taskset]{Latency taskset}
+     \label{fig:latency_taskset}
+   \end{figure}
+
+ Figure \ref{fig:kvm_exits}
+ \begin{figure}[H]
+     \centering
+     \includegraphics[width=0.8\columnwidth]{img/kvm_exits.png}
+     \caption[kvm exits]{kvm exits}
+     \label{fig:kvm_exits}
+   \end{figure}
+
+   Figure \ref{fig:gnuplot_latency_hardware}
+   \begin{figure}[H]
+       \centering
+       \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/hardware/gnuplot_latency_hardware.png}
+       \caption[gnuplot latency hardware]{gnuplot latency hardware}
+       \label{fig:gnuplot_latency_hardware}
+     \end{figure}
+
+     Figure \ref{fig:gnuplot_latency_no_taskset}
+     \begin{figure}[H]
+         \centering
+         \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/no_taskset/gnuplot_latency_no_taskset.png}
+         \caption[gnuplot latency no taskset]{gnuplot latency no taskset}
+         \label{fig:gnuplot_latency_no_taskset}
+       \end{figure}
+
+   Figure \ref{fig:gnuplot_latency_with_taskset}
+   \begin{figure}[H]
+       \centering
+       \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/with_taskset/gnuplot_latency_with_taskset.png}
+       \caption[gnuplot latency with taskset]{gnuplot latency with taskset}
+       \label{fig:gnuplot_latency_with_taskset}
+     \end{figure}
+
+\clearpage
+\chapter{Results}
+
+VARAN-Bus PCV-521\newline
+Windows\newline
+Ubuntu\newline
+WSL\newline\newline\newline
+
 In Xenomai 3, you need:
@@
 In Xenomai 4, you need:
-
 The Linux kernel with the EVL interface, which, like Dovetail, enables the kernel to handle real-time tasks with low latency.
@@
 
-\section{Kernelshark}
-
-\begin{lstlisting}[language=C++,name={1. Beispiel},label={sc:bsp:1}]
-    #include <iostream>
-
-    void SayHello(void)
-    {
-        // Kommentar
-        cout << "Hello World!" << endl;
-    }
-
-    int main(int argc, char **argv)
-    {
-        SayHello();
-        return 0;
-    }
-    \end{lstlisting}
-
-\chapter{Results}
-Firma SIGMATEK\newline
-Yocto\newline
-Salamander 4\newline
-Xenomai\newline
-QEMU\newline
-Trace-cmd\newline
-Kernelshark\newline
-VARAN-Bus PCV-521\newline
-Windows\newline
-Ubuntu\newline
-WSL\newline
 \clearpage
@@
 \begin{acronym}[XXXXX]
-    \acro{ABC}[ABC]{Alphabet}
-    \acro{WWW}[WWW]{world wide web}
-    \acro{ROFL}[ROFL]{Rolling on floor laughing}
+    \acro{CPU}[CPU]{Central Processing Unit}
+    \acro{QEMU}[QEMU]{Quick Emulator}
 \end{acronym}

### 2024-03-27 masterthesis-documentation/docs/sigmatek/xenomai/print_numbers.py

@@ -1,5 +1,5 @@
 # Open the file in write mode
-with open("numbers.txt", "w") as file:
+with open("print_numbers.txt", "w") as file:
     # Write numbers from 1 to 600
     for number in range(1, 601):
         file.write(str(number) + "\n")

### 2024-03-27 masterthesis-documentation/docs/sigmatek/xenomai/latency_taskset.py
Insert Images and write thesis

@@ -2,6 +2,7 @@ import matplotlib.pyplot as plt
 
 # Data points
 x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,445,446,447,448,449,450,451,452,453,454,455,456,457,458,459,460,461,462,463,464,465,466,467,468,469,470,471,472,4
[…]

### 2024-03-27 sigmatek/xenomai/latency_default.py
Deployed 10ec6d4 with MkDocs version: 1.5.3

@@ -19,10 +19,10 @@ plt.xlim([0, 600])
 plt.ylim([0, 5000])
 
 # Save the plot to a file
-plt.savefig("latency_no_taskset.png")
-plt.savefig("../../../../img/latency_no_taskset.png")
+plt.savefig("latency_default.png")
+plt.savefig("../../../../img/latency_default.png")
 
 # Print a success message
-print("The plot was successfully saved to latency.png")
+print("The plot was successfully saved to latency_default.png")
 
 

### 2024-03-27 sigmatek/xenomai/latency_taskset.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 # Data points
 x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,445,446,447,448,449,450,451,452,453,454,455,456,457,458,45
[…]

### 2024-03-27 Masterarbeit_Robotik_Pamuk.tex
latency_default

@@
      \centering
-     \includegraphics[width=0.8\columnwidth]{img/latency_no_taskset.png}
+     \includegraphics[width=0.8\columnwidth]{img/latency_default.png}
      \caption[Latency no taskset]{Latency no taskset}
@@
 
- Figure \ref{fig:kvm_exits}
+ Figure \ref{fig:kvm_exits} shows 5 seconds of sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all
+
  \begin{figure}[H]
@@
          \centering
-         \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/no_taskset/gnuplot_latency_no_taskset.png}
+         \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/default/gnuplot_latency_default.png}
          \caption[gnuplot latency no taskset]{gnuplot latency no taskset}

### 2024-03-27 Masterarbeit_Robotik_Pamuk.tex
max_

@@
 
-   Figure \ref{fig:gnuplot_latency_hardware}
+   Figure \ref{fig:gnuplot_max_latency_hardware}
    \begin{figure}[H]
        \centering
-       \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/hardware/gnuplot_latency_hardware.png}
+       \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/hardware/gnuplot_max_latency_hardware.png}
        \caption[gnuplot latency hardware]{gnuplot latency hardware}
-       \label{fig:gnuplot_latency_hardware}
+       \label{fig:gnuplot_max_latency_hardware}
      \end{figure}
 
-     Figure \ref{fig:gnuplot_latency_no_taskset}
+     Figure \ref{fig:gnuplot_max_latency_no_taskset}
      \begin{figure}[H]
          \centering
-         \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/default/gnuplot_latency_default.png}
+         \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/default/gnuplot_max_latency_default.png}
          \caption[gnuplot latency no taskset]{gnuplot latency no taskset}
-         \label{fig:gnuplot_latency_no_taskset}
+         \label{fig:gnuplot_max_latency_no_taskset}
        \end{figure}
 
-   Figure \ref{fig:gnuplot_latency_with_taskset}
+   Figure \ref{fig:gnuplot_max_latency_with_taskset}
    \begin{figure}[H]
        \centering
-       \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/with_taskset/gnuplot_latency_with_taskset.png}
+       \includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/with_taskset/gnuplot_max_latency_with_taskset.png}
        \caption[gnuplot latency with taskset]{gnuplot latency with taskset}
-       \label{fig:gnuplot_latency_with_taskset}
+       \label{fig:gnuplot_max_latency_with_taskset}
      \end{figure}

### 2024-03-27 sigmatek/xenomai/max_latency_default.py
Deployed fb2c4e9 with MkDocs version: 1.5.3

@@ -19,10 +19,10 @@ plt.xlim([0, 600])
 plt.ylim([0, 5000])
 
 # Save the plot to a file
-plt.savefig("latency_default.png")
-plt.savefig("../../../../img/latency_default.png")
+plt.savefig("max_latency_default.png")
+plt.savefig("../../../../img/max_latency_default.png")
 
 # Print a success message
-print("The plot was successfully saved to latency_default.png")
+print("The plot was successfully saved to max_latency_default.png")
 
 

### 2024-03-27 sigmatek/xenomai/max_latency_hardware.py

@@ -19,10 +19,10 @@ plt.xlim([0, 600])
 plt.ylim([0, 15])
 
 # Save the plot to a file
-plt.savefig("latency_hardware.png")
-plt.savefig("../../../../img/latency_hardware.png")
+plt.savefig("max_latency_hardware.png")
+plt.savefig("../../../../img/max_latency_hardware.png")
 
 # Print a success message
-print("The plot was successfully saved to latency_hardware.png")
+print("The plot was successfully saved to max_latency_hardware.png")
 
 

### 2024-03-27 sigmatek/xenomai/max_latency_taskset.py

@@ -19,10 +19,10 @@ plt.xlim([0, 600])
 plt.ylim([0, 500])
 
 # Save the plot to a file
-plt.savefig("latency_taskset.png")
-plt.savefig("../../../../img/latency_taskset.png")
+plt.savefig("max_latency_taskset.png")
+plt.savefig("../../../../img/max_latency_taskset.png")
 
 # Print a success message
-print("The plot was successfully saved to latency_taskset.png")
+print("The plot was successfully saved to max_latency_taskset.png")
 
 

### 2024-03-27 sigmatek/trace-cmd/analysis/test/default/kvm_exit_default_number.py
Deployed b7b08dc with MkDocs version: 1.5.3

@@ -1,0 +1,71 @@
+import matplotlib.pyplot as plt
+
+def count_lines_with_word(file_path, word, total_lines):
+    with open(file_path, 'r') as file:
+        lines = file.readlines()
+
+    count = 0
+    for line in lines:
+        if word in line:
+            count += 1
+
+    total_lines -= count
+    return total_lines, count
+
+# Usage
+file_path = 'kvm_exit_default_log.txt'  # replace with your file path
+total_lines = sum(1 for line in open(file_path))
+print(f"Total exits: {total_lines}.")
+
+words = ['APIC_WRITE', 
+         'HLT', 
+         'EPT_MISCONFIG',
+         'PREEMPTION_TIMER',
+         'EXTERNAL_INTERRUPT', 
+         'IO_INSTRUCTION',
+         'EOI_INDUCED',
+         'EPT_VIOLATION',
+         'PAUSE_INSTRUCTION',
+         'CPUID',   
+         'MSR_READ'
+         ]  # replace with the words you want to search for
+
+counts = []
+for word in words:
+    total_lines, word_lines = count_lines_with_word(file_path, word, total_lines)
+    print(f"'{word}': {word_lines} times")
+    counts.append((word, word_lines))
+
+# Sort words and counts by count in descending order
+counts.sort(key=lambda x: x[1], reverse=True)
+words, word_counts = zip(*counts)  # unzip the sorted pairs
+
+# Set the figure size
+plt.figure(figsize=(10, 6))
+
+# Plotting
+bars = plt.bar(words, word_counts, color='blue')  # Added color='blue'
+
+# Labeling the bars with their counts
+for bar in bars:
+    yval = bar.get_height()
+    plt.text(bar.get_x() + bar.get_width()/2, yval + 0.05, yval, ha='center', va='bottom')
+
+plt.xlabel('Reasons')
+plt.ylabel('Count')
+plt.title('Exit Reason Count')
+plt.xticks(rotation=90, fontsize='small')  # Adjust font size for readability
+
+# Adjust bottom margin
+plt.subplots_adjust(bottom=0.4)  # Increase the bottom margin
+
+# Set x and y axis lengths
+#plt.xlim([0, len(words)])  # Set x-axis length
+plt.ylim([0, max(word_counts)*1.2])  # Set y-axis length
+
+# Save the figure
+plt.savefig('../../../../../../../img/kvm_exits_default.png', bbox_inches='tight')
[…]

### 2024-03-27 Masterarbeit_Robotik_Pamuk.tex
Write masterthesis 27/03

@@
 \begin{minipage}{0.95\columnwidth}
-\begin{lstlisting}[language=bash,name={System information},label={lst:uname_a}]
+\begin{lstlisting}[name={System information},label={lst:uname_a}]
 root@sigmatek-core2:~# uname -a
@@
 
+Figure \ref{fig:max_latency_hardware} shows latency of hardware Salamander4.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.8\columnwidth]{img/max_latency_hardware.png}
+  \caption[Latency hardware]{Latency hardware}
+  \label{fig:max_latency_hardware}
+\end{figure}
+
 \section{Salamander 4 Virtualisation}
@@
 \begin{minipage}{\linewidth}
-    \begin{lstlisting}[language=bash,name={Contents of QEMU folder for Salamander 4},label={lst:ls}]
+    \begin{lstlisting}[name={Contents of QEMU folder for Salamander 4},label={lst:ls}]
     sigma_ibo@localhost:~/Desktop/salamander-image$ ls -1
@@
 \begin{minipage}{\linewidth}
-\begin{lstlisting}[language=bash,name={QEMU script for starting Salamander 4 virtualisation},label={lst:qemu_def}]
+\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualisation},label={lst:qemu_def}]
     #!/bin/sh
@@
 
- Figure \ref{fig:latency_hardware}
- \begin{figure}[H]
-   \centering
-   \includegraphics[width=0.8\columnwidth]{img/latency_hardware.png}
-   \caption[Latency hardware]{Latency hardware}
-   \label{fig:latency_hardware}
- \end{figure}
+\clearpage
+\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is noted bound to a specific CPU core.
+This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
+As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on user level.
+However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
+Consequently, these kernel tasks and interrupts can still utilize the CPU.
 
- Figure \ref{fig:latency_no_taskset}
+Figure \ref{fig:max_latency_default} shows latency of QEMU default Salamander4.
 
@@
      \centering
-     \includegraphics[width=0.8\columnwidth]{img/latency_default.png}
+     \includegraphics[width=0.75\columnwidth]{img/max_latency_default.png}
      \caption[Latency no taskset]{Latency no taskset}
-     \label{fig:latency_no_taskset}
+     \label{fig:max_latency_default}
    \end{figure}
 
-   Figure \ref{fig:latency_taskset}
+   Figure \ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
    \begin{figure}[H]
      \centering
-     \includegraphics[width=0.8\columnwidth]{img/latency_taskset.png}
+     \includegraphics[width=0.75\columnwidth]{img/max_latency_taskset.png}
      \caption[Latency taskset]{Latency taskset}
-     \label{fig:latency_taskset}
+     \label{fig:max_latency_taskset}
    \end{figure}
 
- Figure \ref{fig:kvm_exits} shows 5 seconds of sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all
+\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
+However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
+Consequently, these kernel tasks and interrupts can still utilize the CPU.
+This led to the investigation of the causes for the observed high and inconsistent latency.
+The guest operates within the kvm\_entry and \texttt{kvm\_exit} events of the host.
+Kernelshark revealed a high frequency of \texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
+This frequent switching hinders the guest\'s ability to run continuously, thereby increasing the virtualization latency.
+To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
+Specifically, the causes for \texttt{kvm\_exit} events were analyzed.
+The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
+The results in Figure \ref{fig:kvm_exits} were obtained.
+Additionally, table \ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
 
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|l|p{0.62\textwidth}|}
+  \hline
+  \textbf{Exit Reason} & \textbf{Description}\\
+  \hline
+  APIC\_WRITE & Triggered when the guest writes to its APIC. \\ \hline
+  EXTERNAL\_INTERRUPT & Triggered by external hardware interrupts. \\ \hline
+  HLT & Triggered when the guest executes the HLT instruction. \\ \hline
+  EPT\_MISCONFIG & Triggered by a misconfiguration in the EPT. \\ \hline
+  PREEMPTION\_TIMER & Triggered when the host's preemption timer expires. \\ \hline
+  PAUSE\_INSTRUCTION & Triggered when the PAUSE instruction is executed. \\ \hline
+  EPT\_VIOLATION & Triggered by a violation of the EPT permission settings. \\ \hline
+  IO\_INSTRUCTION & Triggered when the guest executes an I/O instruction. \\ \hline
+  EOI\_INDUCED & Triggered when an EOI signal is sent to the APIC. \\ \hline
+  MSR\_READ & Triggered when the guest reads from a MSR. \\ \hline
+  CPUID & Triggered when the guest executes the CPUID instruction.  \\ \hline
+  \end{tabular}
+  \caption{Description of kvm\_exit reasons}
+  \label{tab:kvm_exit}
+  \end{table}
+
+\clearpage
+  Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency with CPU islation.
   \begin{figure}[H]
     \centering
-     \includegraphics[width=0.8\columnwidth]{img/kvm_exits.png}
+    \includegraphics[width=1.0\columnwidth]{img/kvm_exits.png}
     \caption[kvm exits]{kvm exits}
@@
 
-   Figure \ref{fig:gnuplot_max_latency_hardware}
+  Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=1.0\columnwidth]{img/kvm_exits_default.png}
+    \caption[kvm exits default]{kvm exits default}
+    \label{fig:kvm_exits_default}
+  \end{figure}
+  \clearpage
+
+\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
+This explains why there are fewer \texttt{kvm\_exit} events.
+However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a \texttt{kvm\_exit} and kvm\_entry.
+
+\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in \texttt{kvm\_exit} events.
+This is because every context switch takes place at hypervisor level.
+Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.
+
+\noindent In the process of analyzing the \texttt{kvm\_exit} events, several reasons for these exits were identified.
+The most frequent among these were the \texttt{APIC\_WRITE} and \texttt{HLT} events.
+The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
+The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
+Other significant but less frequent events included \texttt{EXTERNAL\_INTERRUPT} and \texttt{IO\_INSTRUCTION}.
+These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
+Events such as \texttt{EPT\_MISCONFIG} and \texttt{PREEMPTION\_TIMER} were also noted.
+These could potentially signal issues with memory management and the host's scheduling of the guest.
+While events like \texttt{PAUSE\_INSTRUCTION}, \texttt{EPT\_VIOLATION}, \texttt{EOI\_INDUCED}, \texttt{MSR\_READ}, and \texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
+
+The gnuplot latency is visible in Figure \ref{fig:gnuplot_max_max_latency_hardware}
    \begin{figure}[H]
@@
        \caption[gnuplot latency hardware]{gnuplot latency hardware}
-       \label{fig:gnuplot_max_latency_hardware}
+       \label{fig:gnuplot_max_max_latency_hardware}
      \end{figure}
 
-     Figure \ref{fig:gnuplot_max_latency_no_taskset}
+     Figure \ref{fig:gnuplot_max_max_latency_default}
      \begin{figure}[H]
@@
          \caption[gnuplot latency no taskset]{gnuplot latency no taskset}
-         \label{fig:gnuplot_max_latency_no_taskset}
+         \label{fig:gnuplot_max_max_latency_default}
        \end{figure}

### 2024-03-31 Masterarbeit_Robotik_Pamuk.tex
Application Context and  Produktkatalog

@@
 A higher priority means that a faster response is required.
-Pre-emptive task scheduling ensures a very fast response.
-Preemptive means that the scheduler can stop a currently running task at any point if it recognises that another task needs to be executed immediately.
+Preemptive task scheduling ensures a very fast response.
+Preemptive means that the scheduler can stop a currently running task at any point if it recognizes that another task needs to be executed immediately.
 The basic rule on which priority-based preemptive scheduling is based is that the task with the highest priority that is ready to run is always the task that must be executed.
@@
 The lower priority task is only executed once the higher priority task has been processed.
-Real-time systems are usually categorised as either soft or hard real-time systems.
+Real-time systems are usually categorized as either soft or hard real-time systems.
 The difference lies exclusively in the consequences of a violation of the time limits.
@@
 \bigskip \noindent Hard real-time is when the system stops operating if a deadline is missed, which can have catastrophic consequences.
-These consequences include endangering people, damaging machinery, impacting the health and integrity of the environment.
-Typical examples of this are some control systems in aeroplanes or cars, e.g.
-internal combustion engines.
-
-\bigskip \noindent Soft real-time exists when a system continues to function even if it cannot perform the tasks within a specified time.
+Soft real-time exists when a system continues to function even if it cannot perform the tasks within a specified time.
 If the system has missed the deadline, this has no critical consequences.
 The system continues to run, although it does so with undesirably lower output quality.
-Examples of this are multimedia systems where the occasional deviation from a playback rate of 25 frames/sec.
-leads to stuttering or similar.
-
-%In der heutigen industriellen Fertigung und Automatisierung sind Robotersysteme sehr etabliert und von entscheidender Bedeutung.
-Roboter müssen auf ihre Umgebung reagieren und zeitkritische Aufgaben innerhalb strenger Zeitvorgaben ausführen.
-Verzögerungen oder Fehler können in einigen Fällen katastrophale Konseqenzen haben.
-Traditionelle Betriebssysteme, wie Windows oder Linux sind für diese Art von Echtzeitanforderungen oft nicht geeignet, da sie keine deterministischen Ausführungszeiten garantieren können.
-Daher sind Echtzeit-Betriebssysteme nötig, die speziell dafür konzipiert sind, auf Ereignisse innerhalb fester Zeitschranken zu reagieren und Prozesse mit hoher Priorität bevorzugt auszuführen.
-Bei diesen RTOS-Systemen basiert das Task-Scheduling auf dem sogenannten prioritätsbasierten präemptiven Scheduling.
-Jeder Task in einer Software-Applikation ist eine Priorität zuzuweisen.
-Eine höhere Priorität bedeutet, dass eine schnellere Reaktion erforderlich ist.
-Durch das präemptive Task-Scheduling wird eine sehr schnelle Reaktion sichergestellt.
-Präemptiv bedeutet, dass der Scheduler eine gerade laufende Task an jeder Stelle anhalten kann, wenn er erkennt, dass eine andere Task sofort ausgeführt werden muss.
-Die Grundregel, auf der das prioritätsbasierte präemptive Scheduling basiert, besagt, dass die ausführbereite Task mit der höchsten Priorität immer die Task ist, die ausgeführt werden muss.
-Wenn also sowohl eine Task mit niedriger als auch eine Task mit höherer Priorität ausführbereit sind, sorgt der Scheduler dafür, dass zuerst die Task mit der höheren Priorität läuft.
-Die Task mit niedrigerer Priorität wird erst dann ausgeführt, wenn die höherpriore Task verarbeitet wurde.
-
-%\noindent Echtzeitsysteme werden in der Regel entweder als weiches oder als hartes Echtzeitsystem kategorisiert. Der Unterschied liegt ausschließlich in den Folgen einer Verletzung der Zeitschranken.
-
-%\noindent Harte Echtzeit liegt vor, wenn das System seinen Betrieb einstellt, wenn ein Termin verpasst wurde, was katastrophale Folgen haben kann.
-Zu diesen Folgen zählen die Gefährdung von Menschen, die Beschädigung von Maschinen, also Auswirkungen auf Gesundheit und Unversehrtheit der Umgebung.
-Typische Beispiele hierfür sind einige Steuerungssysteme im Flugzeug oder im Auto, z.B.
-bei der Verbrennungsmaschine.
-
-%\noindent Weiche Echtzeit besteht, wenn ein System weiter funktionsbereit ist, selbst wenn es die Aufgaben innerhalb einer festgelegten Zeit nicht ausführen kann.
-Wenn das System den Termin verpasst hat, hat dies keine kritischen Konsequenzen.
-Das System läuft weiter, obwohl es dies mit unerwünscht geringerer Ausgabenqualität tut.
-Beispiele hierfür sind Multimediasysteme, bei denen das gelegentlich Abweichen von einer Abspielrate von 25 Bildern/sek.
-zu einem Ruckeln o.ä.
-führt.
 
@@
 This master's thesis was written at SIGMATEK GmbH \& Co KG \cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.
+This section provides necessary information about the starting point and boundary conditions of the thesis, as the work should be reproducible with the means used.
+
+SIGMATEK uses its own customized Linux distribution to be run on their self-manufactured CPUs, namely Salamander 4. This operating system
+
+, to employ hard real-time in their Salamander4 is virtualized through third party services.
+
+The details of this operating system are explained in section \ref{sec:salamander4}
 
@@
 \clearpage
-\chapter{Salamander Operating System}
+\chapter{Salamander Operating System}\label{sec:salamander4}
 
@@
 	\end{tabular}
-
 \end{table}
@@
 		\hline
-  \textbf{Exit Reason} & \textbf{Description}\\
+		\textbf{Exit Reason} & \textbf{Description}                                     \\
 		\hline

### 2024-04-02 Masterarbeit_Robotik_Pamuk.tex
bcc, problem_solution, application context

@@
 This master's thesis was written at SIGMATEK GmbH \& Co KG \cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}. 
-This section provides necessary information about the starting point and boundary conditions of the thesis, as the work should be reproducible with the means used.
-
-SIGMATEK uses its own customized Linux distribution to be run on their self-manufactured CPUs, namely Salamander 4. This operating system
-
-, to employ hard real-time in their Salamander4 is virtualized through third party services.
-
-The details of this operating system are explained in section \ref{sec:salamander4}
+%This section provides necessary information about the starting point and boundary conditions of the thesis, as the work should be reproducible with the means used. 
+SIGMATEK uses its own customized Linux distribution to be run on their self-manufactured CPUs, namely Salamander 4.
+This operating system employs hard real-time with latency requirements between 20 and 50 \textmu s.
+The goal is to virtualize Salamander 4 and approach the performance of bare metal CPUs.
+Salamander 4 is virtualized through a third party service, QEMU.
+The details of this operating system are explained in section \ref{sec:salamander4}.
 
@@
 \section{Problem and task definition}
-In robotics, accurate and timely control is crucial to ensure precise movements and reliable interaction with the environment.
-However, existing robot control systems often have limitations in terms of real-time capabilities and communication efficiency, which can have a detrimental effect on response time and reliability.
-The challenge is to develop a powerful real-time robot control platform that overcomes these limitations and improves the performance of robotic applications.
 
@@
 
-This integration is expected to lead to a significant improvement in the response time and reliability of robot applications.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In diesem Abschnitt werden sämtliche theoretischen Konzepte und Randbedingungen sowie praktische Methoden, die zur Erreichung der Zielsetzung dieser Masterarbeit beigetragen haben, detailliert beschrieben.
+This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
 
@@
 The real-time behaviour is achieved through the use of Symmetric Multi-Processing (SMP) and Preemptive Scheduling (PREEMPT).
-In addition, it utilises IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
+In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
 The output of the command \texttt{uname -a} can be observed in code \ref{lst:uname_a}.

### 2024-04-04 Masterarbeit_Robotik_Pamuk.tex
APIC_WRITE

@@
 \section{State of the art}
-
 \clearpage
@@
 This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
+
+\begin{comment}
 
@@
 Ich werde eine VARAN-Bus Schnittstelle verwenden, um eine schnelle und zuverlässige Kommunikation zwischen dem Roboter und dem Steuerungssystem zu gewährleisten.
+
+\end{comment}
 
@@
 The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
-The results in Figure \ref{fig:kvm_exits} were obtained.
+The results in Figure \ref{fig:kvm_exit_taskset} were obtained.
 Additionally, table \ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exits.png}
+	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_taskset.png}
 	\caption[kvm exits]{kvm exits}
-	\label{fig:kvm_exits}
+	\label{fig:kvm_exit_taskset}
 \end{figure}
@@
 
-The gnuplot latency is visible in Figure \ref{fig:gnuplot_max_max_latency_hardware}
+The gnuplot latency is visible in Figure \ref{fig:gnuplot_max_latency_hardware}
 \begin{figure}[H]
@@
 	\caption[gnuplot latency hardware]{gnuplot latency hardware}
-	\label{fig:gnuplot_max_max_latency_hardware}
+	\label{fig:gnuplot_max_latency_hardware}
 \end{figure}
@@
 
+\chapter{Latency Reduction}
+\section{CPU isolation}
 \clearpage
+\section{User and Kerneltasks}
+\clearpage
+\section{IRQ}
+\clearpage
+
+\chapter{KVM exit reasons}
+\section{APIC\_WRITE}
+
+The Advanced Programmable Interrupt Controller (APIC) is responsible for the distribution of interrupts in x86 and Itanium-based computer systems.
+It consists of two main components: the I/O APIC and the local APICs.
+The I/O APIC receives interrupt requests from devices and distributes them as messages to the Local APICs.
+The Local APICs then forward the highest-priority interrupt to the CPU core.
+The APIC offers many advantages, including more inputs for interrupts, a flexible configuration, definable priorities and support for message-signalled interrupts.
+
+\cite{lutsykPipelinedMulticoreMachine2020}
+%https://www.wikiwand.com/de/Advanced_Programmable_Interrupt_Controller
+
+\subsection{APIC virtualization}
+
+Newer Intel processors offer hardware virtualization of the Advanced Programmable Interrupt Controller (APICv).
+APICv improves virtualized AMD64 and Intel 64 guest performance by allowing the guest to directly access the APIC, dramatically cutting down interrupt latencies and the number of virtual machine exits caused by the APIC.
+This feature is used by default in newer Intel processors and improves I/O performance.
+
+\clearpage
+\section{HLT}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{EPT\_MISCONFIG}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{PREEMPTION\_TIMER}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{EXTERNAL\_INTERRUPT}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{IO\_INSTRUCTION}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{EOI\_INDUCED}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{EPT\_VIOLATION}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{PAUSE\_INSTRUCTION}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{CPUID}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+\section{MSR\_READ}
+\subsection{Description}
+\subsection{Occurence Reduction}
+\clearpage
+
 \chapter{Results}
+
+\begin{comment}
 
@@
 So, while both versions require a Linux kernel with a real-time interface (Dovetail for Xenomai 3, EVL for Xenomai 4), Xenomai 3 uses a real-time core (Cobalt), while Xenomai 4 uses a library (libevl) to provide applications with access to its real-time capabilities.
+\end{comment}
 

### 2024-04-05 Masterarbeit_Robotik_Pamuk.tex
defconfig, vapic

@@
 \usepackage{pgfplots}
+\usepackage{soul}
+\usepackage{color}
+\usepackage{xcolor}  % Required for custom colors
+\sethlcolor{blue}
 %
@@
 	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Archive containing files for C drive, unpacked and copied to drive-c/ directory by qemu\_def.sh script.
-	\item \textbf{drive-c}: Directory, serves as C drive for QEMU system, created and filled by qemu\_def.sh script..
+	\item \textbf{drive-c}: Directory serving as C drive for QEMU system, created and filled by qemu\_def.sh script.
 	\item \textbf{salamander-image-sigmatek-core2.ext4}: Root file system for Salamander 4 OS, used as hard drive for QEMU system.
@@
 \section{CPU isolation}
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={User and Kernel Tasks},label={lst:pse},escapeinside={(*@}{@*)}]
+		sigma_ibo@sigma-ibo:~$ cat /sys/devices/system/cpu/isolated
+		19
+		sigma_ibo@sigma-ibo:~$ ps -e -o pid,psr,comm | awk '$2 == 19'
+				92  19 cpuhp/19
+				93  19 idle_inject/19
+ 	    		94  19 migration/19
+ 	    		95  19 ksoftirqd/19
+ 	    		97  19 kworker/19:0H-events_highpri
+ 	 		 1025  19 irq/205-iwlwifi:queue_7
+ 	 		17448  19 kworker/19:1H-kblockd
+ 	 		17499  19 kworker/19:2-events
+ 	 		18761  19 kworker/19:3-events
+		   (*@\colorbox{yellow}{21401 19 qemu-system-x86}@*)
+\end{lstlisting}
+\end{minipage}
+
 \clearpage

### 2024-04-05 masterthesis-documentation/docs/sigmatek/QEMU/qemu_def_7vapic.sh
defconfig, vapic

@@ -1,0 +1,16 @@
+#!/bin/sh
+
+if  [ ! -d drive-c/ ]; then
+        echo "Filling drive-c/"
+        mkdir drive-c/
+        tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
+fi
+
+exec taskset -c 19 qemu-system-x86_64 -M pc,accel=kvm -cpu host,kvm=off,hv_relaxed,hv_vapic -kernel ./bzImage \
+-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
+-append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable vapic" \
+-net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
+-fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
+-device vhost-vsock-pci,guest-cid=3,id=vsock0 \
+-drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
+-no-reboot -nographic

### 2024-04-05 masterthesis-documentation/docs/latency_reduction/check_CPU_IRQ_usage.sh
check_CPU_IRQ_usage

@@ -1,0 +1,32 @@
+#!/bin/bash
+# Get the IRQ number from the command-line argument
+IRQ=$1
+# Get the number of CPUs
+num_cpus=$(nproc)
+# Initialize an empty array to store the CPU numbers
+CPUs=()
+# Check if the smp_affinity file exists for this IRQ
+if [ -f "/proc/irq/$IRQ/smp_affinity" ]; then
+    # Read the current smp_affinity
+    AFFINITY=$(cat "/proc/irq/$IRQ/smp_affinity")
+    # Iterate over each CPU
+    for ((cpu=0; cpu<num_cpus; cpu++)); do
+        # Check if the bit for the current CPU is set
+        if (( (0x$AFFINITY & (1 << cpu)) != 0 )); then
+            # Add the CPU number to the array
+            CPUs+=("$cpu")
+        fi
+    done
+    # Add the isolated CPU to the list
+    CPUs+=("19")
+    # Sort the array
+    IFS=$'\n' sorted=($(sort -n <<<"${CPUs[*]}"))
+    # Print the IRQ number
+    echo "IRQ $IRQ is being used by:"
+    # Print the sorted CPU numbers on separate lines
+    for cpu in "${sorted[@]}"; do
+        echo "$cpu"
+    done
+else
+    echo "IRQ $IRQ does not exist or does not have an smp_affinity file."
+fi

### 2024-04-05 Masterarbeit_Robotik_Pamuk.tex
Write masterthesis

--- 2024-04-04 Masterarbeit_Robotik_Pamuk.tex
@@
 %- Anhang (optional)
-\chapter{Introduction}
+\chapter{Introduction}\label{cha:introduction}
 In today's industrial production and automation, robot systems are well established and of crucial importance.
@@
 Salamander 4 is virtualized through a third party service, QEMU.
-The details of this operating system are explained in section \ref{sec:salamander4}.
+The details of this operating system are explained in chapter \ref{cha:salamander4}.
 
@@
 \clearpage
-
 \section{Problem and task definition}
-
 \clearpage
@@
 \clearpage
-\chapter{Methodology}
+
+\chapter{Methodology}\label{cha:methodology}
 
@@
 \clearpage
-\chapter{Salamander Operating System}\label{sec:salamander4}
+
+\chapter{Salamander 4}\label{cha:salamander4}
 
@@
 In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
-The output of the command \texttt{uname -a} can be observed in code \ref{lst:uname_a}.
+The output of the command \texttt{uname -a} can be observed in code \ref{output:uname_a}.
 
@@
 \begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={System information},label={lst:uname_a}]
+	\begin{lstlisting}[name={System information},label={output:uname_a}]
 root@sigmatek-core2:~# uname -a
@@
 
-\section{Task priorities}
+\clearpage
+\section{Salamander 4 Description}
+
+\subsection{Task priorities}
 Es gibt grundsätzlich 4 Gruppen
@@
 \end{table}
+\clearpage
 
-\section{Memory Management}
+\subsection{Memory Management}
 Es gibt verschiedene Speicherbereiche
@@
 
-\chapter{Comparison of Real-Time Latency}
-In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4.
-For this purpose, the latency tool of the Xenomai test suite was used.
-The latency was measured under two conditions, idle and CPU-stressed.
-The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.
-
-Vorgehensweise von \cite{linPerformanceEvaluationXenomai}
-
-%In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
-While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
-The commands to kill these processes are written in the test script.
-
-\begin{table}[H]
-	\centering
-	\caption[Kernel and Patches]{Kernel and Patches}
-	\label{tab:kernel_configuration}
-	\begin{tabular}{|c|c|}
-		\hline
-		\textbf{Kernel} & \textbf{Patches} \\ \hline
-		.               & .                \\ \hline
-		.               & .                \\ \hline
-	\end{tabular}
-\end{table}
-
-After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
-
 \section{Salamander 4 Bare Metal}
@@
 \end{figure}
+\clearpage
 
@@
 The virtualisation runs in a QEMU environment, which is an open source tool for hardware virtualisation \cite{QEMU}.
-With the help of the script depicted in code \ref{lst:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
+With the help of the script depicted in code \ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
 This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host.
@@
     ovmf.code.qcow2
-    qemu_def_bkp.sh
     qemu_def.sh
@@
 \begin{minipage}{\linewidth}
-	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualisation},label={lst:qemu_def}]
+	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualisation},label={script:qemu_def}]
     #!/bin/sh
@@
 \end{minipage}
-
+\clearpage
 Here is a description of the used components:
@@
 
-\clearpage
 \noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is noted bound to a specific CPU core.
@@
 \end{figure}
-
-\chapter{Latency Reduction}
-\section{CPU isolation}
-\clearpage
-\section{User and Kerneltasks}
-\clearpage
-\section{IRQ}
 \clearpage
 
-\chapter{KVM exit reasons}
-\section{APIC\_WRITE}
+\section{Latency Comparison}
+In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4.
+For this purpose, the latency tool of the Xenomai test suite was used.
+The latency was measured under two conditions, idle and CPU-stressed.
+The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.
+
+Vorgehensweise von \cite{linPerformanceEvaluationXenomai}
+
+%In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
+While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
+The commands to kill these processes are written in the test script.
+
+\begin{table}[H]
+	\centering
+	\caption[Kernel and Patches]{Kernel and Patches}
+	\label{tab:kernel_configuration}
+	\begin{tabular}{|c|c|}
+		\hline
+		\textbf{Kernel} & \textbf{Patches} \\ \hline
+		.               & .                \\ \hline
+		.               & .                \\ \hline
+	\end{tabular}
+\end{table}
+
+After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
+\clearpage
+
+\chapter{Real-Time Latency Reduction}\label{cha:latency_reduction}
+\section{CPU isolation}
+Isolating CPUs inolves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
+Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process.
+
+Output \ref{output:pse} shows
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={User and Kernel Tasks},label={output:pse},escapeinside={(*@}{@*)}]
+		sigma_ibo@sigma-ibo:~$ cat /sys/devices/system/cpu/isolated
+		19
+		sigma_ibo@sigma-ibo:~$ ps -e -o pid,psr,comm | awk '$2 == 19'
+				92  19 cpuhp/19
+				93  19 idle_inject/19
+ 	    		94  19 migration/19
+ 	    		95  19 ksoftirqd/19
+ 	    		97  19 kworker/19:0H-events_highpri
+ 	 		 1025  19 irq/205-iwlwifi:queue_7
+ 	 		17448  19 kworker/19:1H-kblockd
+ 	 		17499  19 kworker/19:2-events
+ 	 		18761  19 kworker/19:3-events
+			21401  19 qemu-system-x86
+\end{lstlisting}
+\end{minipage}
+
+\clearpage
+\section{Kernel tuning}
+\clearpage
+\section{IRQ affinity}
+
+The IRQ affinity can be configured so that interrupts are handled on CPUs other than CPU 19. This helps reduce the number of kernel threads that are executed on CPU 19.
+
+The script \ref{script:irq_affinity} checks the smp\_affinity for each IRQ on a specific CPU and prints the IRQ numbers that are allowed to run on that CPU, sorted in ascending order.
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={Read IRQ affinity per CPU on system},label={script:irq_affinity}]
+		#!/bin/bash
+		# Get the CPU number from the command-line argument
+		CPU=$1
+		echo -n "CPU $CPU: "
+		for IRQ in /proc/irq/*; do
+			if [ -f "$IRQ/smp_affinity" ]; then
+				# Read the current smp_affinity
+				AFFINITY=$(cat "$IRQ/smp_affinity")
+				# Check if the bit for the current CPU is set
+				if (( (0x$AFFINITY & (1 << CPU)) != 0 )); then
+					# Print the IRQ number without a newline, followed by a space
+					echo -n "${IRQ#/proc/irq/} "
+				fi
+			fi
+		done | sort -n
+		# Print a final newline
+		echo
+\end{lstlisting}
+\end{minipage}
+
+Output \ref{output:irq_affinity} shows the output of the script above for CPU 19.
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+\begin{lstlisting}[name={Output},label={output:irq_affinity}, breaklines=true]
+	sigma_ibo@sigma-ibo:~$ ./check_smp_affinity.sh 19
+	CPU 19 IRQ affinity:
+	0
+	2
+	3
+	4
+	5
+	6
+	7
+	10
+	11
+	13
+	15
+	130
+	172
+	173
+	189
+	192
+\end{lstlisting}
+\end{minipage}
+
+\clearpage
+\section{KVM exit reasons}\label{sec:kvm_exit_reasons}
+\subsection{APIC\_WRITE}
 
@@
 
-\subsection{APIC virtualization}
+\subsubsection{APIC virtualization}
 
@@
 \clearpage
-\section{HLT}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{HLT}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{EPT\_MISCONFIG}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{EPT\_MISCONFIG}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{PREEMPTION\_TIMER}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{PREEMPTION\_TIMER}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{EXTERNAL\_INTERRUPT}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{EXTERNAL\_INTERRUPT}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{IO\_INSTRUCTION}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{IO\_INSTRUCTION}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{EOI\_INDUCED}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{EOI\_INDUCED}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{EPT\_VIOLATION}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{EPT\_VIOLATION}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{PAUSE\_INSTRUCTION}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{PAUSE\_INSTRUCTION}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{CPUID}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{CPUID}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
-\section{MSR\_READ}
-\subsection{Description}
-\subsection{Occurence Reduction}
+\subsection{MSR\_READ}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
 \clearpage
 
-\chapter{Results}
+\chapter{Results}\label{cha:results}
 
@@
 \clearpage
-\chapter{Discussion}
+\chapter{Discussion}\label{cha:discussion}
 \clearpage
-\chapter{Summary and Outlook}
+\chapter{Summary and Outlook}\label{cha:summary_and_outlook}
 
@@
 	\acro{QEMU}[QEMU]{Quick Emulator}
+	\acro{IRQ}[Interrupt Request]{Interrupt Request}
 \end{acronym}
@@
 \appendix
+
 \chapter{Anhang A}
 \clearpage
+
 \chapter{Anhang B}
+
 \end{document}

### 2024-04-05 latency_reduction/check_smp_affinity.sh
Deployed 9939e2d with MkDocs version: 1.5.3

@@ -16,5 +16,9 @@ for IRQ in /proc/irq/*; do
 done
 # Sort the array
 IFS=$'\n' sorted=($(sort -n <<<"${IRQs[*]}"))
-# Print the sorted IRQ numbers on the same line, separated by ","
-IFS=','; echo "CPU $CPU: ${sorted[*]}"
+# Print the CPU number
+echo "CPU $CPU IRQ affinity:"
+# Print the sorted IRQ numbers on separate lines
+for irq in "${sorted[@]}"; do
+    echo "$irq"
+done

### 2024-04-05 latency_reduction/check_smp_affinity_all.py

@@ -51,12 +51,12 @@ def summarize_irqs(irqs):
     return ','.join(map(str, ranges))
 
 # Open the log file
-with open("irq_affinity.log", "w") as f:
+with open("check_smp_affinity_all.log", "w") as f:
     # Print the IRQs for each CPU
     for cpu in range(num_cpus):
         # Write the output to the log file
         f.write(f'CPU {cpu}: [{summarize_irqs(irqs_per_cpu[cpu])}] -> total of {len(irqs_per_cpu[cpu])} IRQs\n')
 
 # Print a success message
-print("The IRQs for each CPU have been successfully written to irq_affinity.log")
+print("The IRQs for each CPU have been successfully written to check_smp_affinity_all.log")
 

### 2024-04-15 Masterarbeit_Robotik_Pamuk.tex
Interrupt Requests Handling

@@
 
+Figure \ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.75\columnwidth]{img/max_latency_vapic.png}
+	\caption[Latency taskset]{Latency taskset}
+	\label{fig:max_latency_vapic}
+\end{figure}
+
 \noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
@@
 \clearpage
-Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency with CPU islation.
+Figure \ref{fig:kvm_exit_taskset} shows \texttt{kvm\_exit} frequency with CPU islation.
 \begin{figure}[H]
@@
 	\label{fig:kvm_exits_default}
+\end{figure}
+
+\clearpage
+
+Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
+	\caption[kvm exits default]{kvm exits default}
+	\label{fig:kvm_exit_vapic}
 \end{figure}
@@
 \end{figure}
+
+Figure \ref{fig:gnuplot_max_latency_vapic}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/vapic/gnuplot_max_latency_vapic.png}
+	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
+	\label{fig:gnuplot_max_latency_vapic}
+\end{figure}
+
 \clearpage
@@
 \section{CPU isolation}
-Isolating CPUs inolves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process.
-
-Output \ref{output:pse} shows
+Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
+Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section \ref{sec:irq_handling}.
+Output \ref{output:pse} shows the user and kernel tasks that run on CPU 19.
+After the isolation, user tasks other than the QEMU process have been removed from running on this CPU.
+Only few critical kernel threads that are tied to this CPU still take CPU time.
 
@@
 \clearpage
-\section{Kernel tuning}
-\clearpage
-\section{IRQ affinity}
-
-The IRQ affinity can be configured so that interrupts are handled on CPUs other than CPU 19. This helps reduce the number of kernel threads that are executed on CPU 19.
-
-The script \ref{script:irq_affinity} checks the smp\_affinity for each IRQ on a specific CPU and prints the IRQ numbers that are allowed to run on that CPU, sorted in ascending order.
+\section{Interrupt Requests Handling}\label{sec:irq_handling}
+Once the CPUs were isolated, Interrupt Requests Handling was the next step.
+Interrupt Requests are used to send a signal to the CPU, prompting it to 'interrupt' its current task and divert its attention to another task.
+This allows hardware devices to communicate with the CPU through frequent context switches, which can lead to performance degradation, especially in high-performance computing or real-time scenarios.
+To mitigate this, the IRQs needed to be removed from the isolated CPUs.
+This was done by manipulating a file in the proc filesystem, namely \texttt{/proc/irq/default\_smp\_affinity}.
+The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format.
+Each bit in this mask corresponds to a CPU in the system.
+The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the significance increases towards the left until CPU19.
+In a system with 20 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be FFFFF.
+The script in code \ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.
 
@@
 \begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={Read IRQ affinity per CPU on system},label={script:irq_affinity}]
+	\begin{lstlisting}[name={Check distribution of Interrupt Requests across each CPU},label={script:smp_affinity}]
 		#!/bin/bash
+		# Check if a command-line argument is provided
+		if [ -z "$1" ]; then
+			echo "Please provide a CPU number as a command-line argument."
+			exit 1
+		fi
 		# Get the CPU number from the command-line argument
 		CPU=$1
-		echo -n "CPU $CPU: "
+		# Initialize an empty array to store the IRQ numbers
+		IRQs=()
 		for IRQ in /proc/irq/*; do
@@
 				if (( (0x$AFFINITY & (1 << CPU)) != 0 )); then
-					# Print the IRQ number without a newline, followed by a space
-					echo -n "${IRQ#/proc/irq/} "
+					# Add the IRQ number to the array
+					IRQs+=("${IRQ#/proc/irq/}")
 				fi
 			fi
-		done | sort -n
-		# Print a final newline
-		echo
+		done
+		# Sort the array
+		IFS=$'\n' sorted=($(sort -n <<<"${IRQs[*]}"))
+		# Print the CPU number
+		echo "CPU $CPU IRQ affinity:"
+		# Print the sorted IRQ numbers on separate lines
+		for irq in "${sorted[@]}"; do
+			echo "$irq"
+		done
+
 \end{lstlisting}
@@
 \begin{minipage}{0.95\columnwidth}
-\begin{lstlisting}[name={Output},label={output:irq_affinity}, breaklines=true]
+\begin{lstlisting}[name={Output of smp_affinity for CPU 19},label={output:irq_affinity}, breaklines=true]
 	sigma_ibo@sigma-ibo:~$ ./check_smp_affinity.sh 19
@@
 	15
-	130
+	131
 	172
-	173
+	188
 	189
-	192
+	195
 \end{lstlisting}
 \end{minipage}
+
+By changing the values in
+
+, I could effectively control the assignment of IRQs, ensuring they would not be handled by the isolated CPUs.
+This manipulation was done carefully, considering the implications and ensuring the stability and integrity of the system.
+
+Through this approach, I was able to create a more streamlined and efficient environment for the execution of specific tasks on the isolated CPUs.
+By reducing the interruptions caused by IRQs, the isolated CPUs were able to focus more on their assigned tasks, potentially leading to improved performance and more predictable execution times.
+
+\section{Kernel tuning}
 
@@
 	\acro{QEMU}[QEMU]{Quick Emulator}
-	\acro{IRQ}[Interrupt Request]{Interrupt Request}
+	\acro{IRQ}[IRQ]{Interrupt Request}
 \end{acronym}

### 2024-04-16 Masterarbeit_Robotik_Pamuk.tex
Updated analyze_trace.py to include PID

@@
 
+\begin{table}[h!]
+	\centering
+	\begin{minipage}{.5\textwidth}
+	\centering
+	\caption{Host report (Total of 445.908)}
+	\label{tab:results_host_report}
+	\begin{tabular}{|c|c|c|}
+	\hline
+	PID & Task & Count \\ \hline
+	182579 & qemu-system-x86 & 302.748 \\ \hline
+	0 & \textless{}idle\textgreater{} & 112.911 \\ \hline
+	182618 & vhost & 21.204 \\ \hline
+	182572 & qemu-system-x86 & 7.597 \\ \hline
+	182755 & qemu-system-x86 & 644 \\ \hline
+	182754 & qemu-system-x86 & 643 \\ \hline
+	181870 & kworker/19:1 & 139 \\ \hline
+	3820 & kworker/19:1H & 16 \\ \hline
+	94 & migration/19 & 6 \\ \hline
+	\end{tabular}
+	\end{minipage}%
+	\begin{minipage}{.5\textwidth}
+	\centering
+	\caption{Guest report (Total of 362.370)}
+	\label{tab:results_guest_report}
+	\begin{tabular}{|c|c|c|}
+	\hline
+	PID & Task & Count \\ \hline
+	0 & \textless{}idle\textgreater{} & 150.744 \\ \hline
+	331 & LRT-Main & 56.697 \\ \hline
+	377 & trace-cmd & 48.507 \\ \hline
+	346 & CLI & 26.426 \\ \hline
+	378 & kthreadd & 25.837 \\ \hline
+	340 & MainTaskLow & 19.291 \\ \hline
+	339 & \textless{}...\textgreater{} & 9.980 \\ \hline
+	34 & MainTaskHigh & 9.185 \\ \hline
+	327 & LE-Logger & 4.965 \\ \hline
+	369 & kworker/0:0 & 2.793 \\ \hline
+	321 & kWorker-LRT & 2.542 \\ \hline
+	328 & LRTMgr-Main & 1.651 \\ \hline
+	34 & LrtMgrCyclic & 1.220 \\ \hline
+	332 & cobalt\_printf & 1.112 \\ \hline
+	325 & LE-System & 534 \\ \hline
+	343 & TCP-Listen & 187 \\ \hline
+	15 & rcu\_preempt & 162 \\ \hline
+	25 & kcompactd0 & 122 \\ \hline
+	58 & kworker/0:1H & 96 \\ \hline
+	63 & kworker/u2:2 & 89 \\ \hline
+	8 & jbd2/sda-8 & 86 \\ \hline
+	22 & kworker/0:1 & 56 \\ \hline
+	1 & init & 31 \\ \hline
+	2 & kthreadd & 25 \\ \hline
+	375 & trace-cmd & 24 \\ \hline
+	14 & ksoftirqd/0 & 8 \\ \hline
+	\end{tabular}
+	\end{minipage}
+	\end{table}
+
 \noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
@@
 
+\subsection{Generic Ubuntu}
+
+\subsection{Real-Time Ubuntu}
+
 %In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
@@
 The commands to kill these processes are written in the test script.
-
-\begin{table}[H]
-	\centering
-	\caption[Kernel and Patches]{Kernel and Patches}
-	\label{tab:kernel_configuration}
-	\begin{tabular}{|c|c|}
-		\hline
-		\textbf{Kernel} & \textbf{Patches} \\ \hline
-		.               & .                \\ \hline
-		.               & .                \\ \hline
-	\end{tabular}
-\end{table}
 
@@
 To mitigate this, the IRQs needed to be removed from the isolated CPUs.
-This was done by manipulating a file in the proc filesystem, namely \texttt{/proc/irq/default\_smp\_affinity}.
+This was done by manipulating a file in the proc filesystem, namely \texttt{/proc/irq/<IRQ>/smp\_affinity}.
 The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format.
@@
 
-By changing the values in
-
-, I could effectively control the assignment of IRQs, ensuring they would not be handled by the isolated CPUs.
-This manipulation was done carefully, considering the implications and ensuring the stability and integrity of the system.
-
-Through this approach, I was able to create a more streamlined and efficient environment for the execution of specific tasks on the isolated CPUs.
-By reducing the interruptions caused by IRQs, the isolated CPUs were able to focus more on their assigned tasks, potentially leading to improved performance and more predictable execution times.
+\noindent By changing the values of the \texttt{smp\_affinity} files of the respective IRQs, the assignment of IRQs was controlled so that they would not be handled by the isolated CPU.
+This reduced the interruptions caused by IRQs and the isolated CPUs were able to focus more on their assigned tasks.
 

### 2024-04-16 masterthesis-documentation/docs/sigmatek/trace-cmd/analysis/test/analyze_trace.py

@@ -8,6 +8,9 @@ def count_tasks(input_filename):
     # Initialize a dictionary to store the counts
     task_counts = defaultdict(int)
 
+    # Initialize a dictionary to store the PIDs
+    task_pids = {}
+
     # Initialize a variable to store the total count of all events
     total_count = 0
 
@@ -29,6 +32,10 @@ def count_tasks(input_filename):
 
             # Update the count for this task
             task_counts[task] += 1
+            
+            # Extract the PID from the task and store it
+            pid = task.split("-")[-1] if "-" in task else "N/A"
+            task_pids[task] = pid
 
             # Update the total count of all events
             total_count += 1
@@ -45,41 +52,52 @@ def count_tasks(input_filename):
         f.write(f"Total count of all events: {total_count:,}\n\n".replace(",", ".")) 
 
         # Write the table headers
-        f.write("| Task | Count |\n")
-        f.write("| --- | --- |\n")
+        f.write("| Task | PID | Count |\n")
+        f.write("| --- | --- | --- |\n")
 
-        # Write the counts to the output file
+        # Write the counts and PIDs to the output file
         for task, count in sorted_tasks:
-            # Escape < and > with a backslash
-            task = task.replace("<", "\\<").replace(">", "\\>")
-            f.write(f"| {task} | {count:,} |\n".replace(",", "."))
+            # Extract the task name without the PID
+            task_name = task.rsplit("-", 1)[0] if "-" in task else task
 
+            # Escape < and > with a backslash
+            escaped_task_name = task_name.replace("<", "\\<").replace(">", "\\>")
+
+            pid = task_pids[task]
+            f.write(f"| {escaped_task_name} | {pid} | {count:,} |\n".replace(",", "."))
+        
     # Plotting
     tasks, counts = zip(*sorted_tasks)  # Unpack the sorted tasks and counts
-    plt.figure(figsize=(20, 15))
-    bars = plt.bar(tasks, counts, color='blue')
-
+    plt.figure(figsize=(10, 7))  # Adjust the figure size to better fit the horizontal plot
+    bars = plt.barh(tasks, counts, color='blue')  # Use barh instead of bar for a horizontal plot
+    
     # Labeling the bars with their counts
+    max_count = max(counts)
     for bar in bars:
[…]

### 2024-04-16 sigmatek/trace-cmd/analysis/test/analyze_trace.py
Deployed 23f02ce with MkDocs version: 1.5.3

@@ -1,15 +1,25 @@
+import os
 import sys
+import matplotlib.pyplot as plt
 from collections import defaultdict
+import matplotlib.ticker as ticker
 
 def count_tasks(input_filename):
     # Initialize a dictionary to store the counts
     task_counts = defaultdict(int)
+
+    # Initialize a dictionary to store the PIDs
+    task_pids = {}
 
     # Initialize a variable to store the total count of all events
     total_count = 0
 
     with open(input_filename, 'r') as f:
         for line in f:
+            # Skip lines that start with "cpus=1" or "cpus=20" or "000:" or "010:"
+            if line.startswith("cpus=1") or line.startswith("cpus=20") or line.startswith("000:") or line.startswith("010:"):
+                continue
+
             # Split the line into words
             words = line.split()
 
@@ -22,6 +32,10 @@ def count_tasks(input_filename):
 
             # Update the count for this task
             task_counts[task] += 1
+            
+            # Extract the PID from the task and store it
+            pid = task.split("-")[-1] if "-" in task else "N/A"
+            task_pids[task] = pid
 
             # Update the total count of all events
             total_count += 1
@@ -29,14 +43,61 @@ def count_tasks(input_filename):
     # Sort the tasks by count in descending order
     sorted_tasks = sorted(task_counts.items(), key=lambda x: x[1], reverse=True)
 
-    # Use f-string to include the input_filename in the output filename
-    with open(f'results_{input_filename}', 'w') as f:
-        # Write the total count of all events to the output file
-        f.write(f"Total count of all events: {total_count}\n") 
+    # Remove the .txt extension from the input filename
+    base_filename = os.path.splitext(input_filename)[0]
 
-        # Write the counts to the output file
+    # Use f-string to include the base_filename in the output filename
+    with open(f'results_{base_filename}.md', 'w') as f:
+        # Write the total count of all events to the output file
+        f.write(f"Total count of all events: {total_count:,}\n\n".replace(",", ".")) 
+
+        # Write the table headers
+        f.write("| Task | PID | Count |\n")
+        f.write("| --- | --- | --- |\n")
+
+        # Write the counts and PIDs to the output file
         for task, count in sorted_tasks:
-            f.write(f"{task}: {count}\n")
[…]

### 2024-04-16 sigmatek/trace-cmd/analysis/test/kvm_exit_count.py

@@ -1,0 +1,73 @@
+import matplotlib.pyplot as plt
+
+def count_lines_with_word(file_path, word, total_lines):
+    with open(file_path, 'r') as file:
+        lines = file.readlines()
+
+    count = 0
+    for line in lines:
+        if word in line:
+            count += 1
+
+    total_lines -= count
+    return total_lines, count
+
+# Usage
+file_path = 'kvm_exit_count.txt'  # replace with your file path
+total_lines = sum(1 for line in open(file_path))
+print(f"Total exits: {total_lines}.")
+total = total_lines
+
+words = ['APIC_WRITE', 
+         'HLT', 
+         'EPT_MISCONFIG',
+         'PREEMPTION_TIMER',
+         'EXTERNAL_INTERRUPT', 
+         'IO_INSTRUCTION',
+         'EOI_INDUCED',
+         'EPT_VIOLATION',
+         'PAUSE_INSTRUCTION',
+         'CPUID',   
+         'MSR_READ',
+         'MSR_WRITE'
+         ]  # replace with the words you want to search for
+
+counts = []
+for word in words:
+    total_lines, word_lines = count_lines_with_word(file_path, word, total_lines)
+    print(f"'{word}': {word_lines} times")
+    counts.append((word, word_lines))
+
+# Sort words and counts by count in descending order
+counts.sort(key=lambda x: x[1], reverse=True)
+words, word_counts = zip(*counts)  # unzip the sorted pairs
+
+# Set the figure size
+plt.figure(figsize=(10, 6))
+
+# Plotting
+bars = plt.bar(words, word_counts, color='blue')  # Added color='blue'
+
+# Labeling the bars with their counts
+for bar in bars:
+    yval = bar.get_height()
+    plt.text(bar.get_x() + bar.get_width()/2, yval + 0.05, yval, ha='center', va='bottom')
+
+plt.xlabel('Reasons')
+plt.ylabel('Count')
+plt.title(f'Exit Reason Count (Total exits: {total})')  # Add total exits to the title
+plt.xticks(rotation=90, fontsize='small')  # Adjust font size for readability
+
+# Adjust bottom margin
+plt.subplots_adjust(bottom=0.4)  # Increase the bottom margin
+
+# Set x and y axis lengths
+#plt.xlim([0, len(words)])  # Set x-axis length
+plt.ylim([0, max(word_counts)*1.2])  # Set y-axis length
+
+# Save the figure
[…]

### 2024-04-18 Masterarbeit_Robotik_Pamuk.tex
add papers

@@
 
+\noindent In the following, the host and guest tasks along with their impact on system latency are briefly described.
+
+\begin{itemize} 
+	\item \textbf{qemu-system-x86}: Part of the QEMU process and specifically, this task emulates x86 systems.
+In table \ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
+	\item \textbf{<idle>}: This represents the idle time of the CPU, hence it is not being used by any process. 
+	\item \textbf{vhost}: A kernel module which improves virtio (virtual input/output) performance by handling virtqueues in the kernel, thereby reducing context switches and system calls. 
+	\item \textbf{kworker/19:1}: A kernel worker thread created by the Linux kernel, kworker/19:1 performs work in response to system events.
+The number after the slash and colon indicate the CPU core and internal ID of the worker thread, respectively.
+	\item \textbf{kworker/19:1H}: Similar to kworker/19:1, kworker/19:1H is a kernel worker thread, with the ‘H’ suggesting that this thread handles hardware interrupts. 
+	\item \textbf{migration/19}: The migration process is a kernel process that balances load across CPU cores by moving threads from one CPU to another.
+The number after the slash indicates the CPU core to which the migration process is bound.
+\end{itemize}
+
+ qemu-system-x86 
+ <idle> 
+ vhost 
+ qemu-system-x86 
+ qemu-system-x86 
+ qemu-system-x86 
+ kworker/19:1 
+ kworker/19:1H 
+ migration/19
+
+\begin{comment}
 \noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
@@
 Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.
+\end{comment}
 

### 2024-04-26 Masterarbeit_Robotik_Pamuk.tex
latency_reduction_steps
url

@@
 % own 
+\usepackage{hyperref}
 \usepackage{float}
@@
 	\centering
-	\caption{Host report (Total of 445.908)}
+	\caption{Host report CPU19 (Total of 445.908)}
 	\label{tab:results_host_report}
@@
 In table \ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
-	\item \textbf{<idle>}: This represents the idle time of the CPU, hence it is not being used by any process. 
-	\item \textbf{vhost}: A kernel module which improves virtio (virtual input/output) performance by handling virtqueues in the kernel, thereby reducing context switches and system calls. 
+	\item \textbf{<idle>}: This represents the idle time of the CPU, hence it is not being used by any process, allowing to save power.
+The system halts until the next interrupt, which could be a timer interrupt, I/O interrupt, etc.
+	\item \textbf{vhost}: A kernel module which improves virtual input/output (virtio) performance by handling virtqueues in the kernel, thereby reducing context switches and system calls. 
 	\item \textbf{kworker/19:1}: A kernel worker thread created by the Linux kernel, kworker/19:1 performs work in response to system events.
@@
 The number after the slash indicates the CPU core to which the migration process is bound.
+(URL: \url{https://elixir.bootlin.com/linux/latest/source/kernel/sched/core.c#L2325})
+
 \end{itemize}
-
- qemu-system-x86 
- <idle> 
- vhost 
- qemu-system-x86 
- qemu-system-x86 
- qemu-system-x86 
- kworker/19:1 
- kworker/19:1H 
- migration/19
 

### 2024-04-30 masterthesis-documentation/docs/sigmatek/trace-cmd/analysis/test/report.sh
merge and failed_reason

@@ -20,6 +20,6 @@ python analyze_trace.py guest_report.txt && echo "analyze_trace for guest_report
 wait
 # Move and Copy elements in directory 
 mv *.txt *.md *.png $1
-cp start_kernelshark.sh *.dat $1
+cp merge.py failed_reason.py start_kernelshark.sh *.dat $1
 exit 0
 

### 2024-04-30 Masterarbeit_Robotik_Pamuk.tex
write thesis, richard

@@
 
-Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
-	\caption[kvm exits default]{kvm exits default}
-	\label{fig:kvm_exit_vapic}
-\end{figure}
-\clearpage
-
 \begin{table}[h!]
@@
 
+\section{Latency Comparison}
+In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4.
+For this purpose, the latency tool of the Xenomai test suite was used.
+The latency was measured under two conditions, idle and CPU-stressed.
+The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.
+
+Vorgehensweise von \cite{linPerformanceEvaluationXenomai}
+
+\subsection{Generic Ubuntu}
+
+\subsection{Real-Time Ubuntu}
+
+%In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
+While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
+The commands to kill these processes are written in the test script.
+
+After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
+\clearpage
+
+\chapter{Real-Time Performance Optimizations and Latency Reduction}\label{cha:latency_reduction}
+\section{Host OS Optimization}
+
+\subsection{CPU governor}
+
+Figure \ref{fig:power_saver_kvm_exit} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/power_saver/kvm_exit_count.png}
+	\caption[kvm exits default]{power\_saver kvm\_exit\_count}
+	\label{fig:power_saver_kvm_exit}
+\end{figure}
+\clearpage
+
+Figure \ref{fig:power_saver_kvm_host} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/power_saver/results_host_report.png}
+	\caption[kvm exits default]{power\_saver host report}
+	\label{fig:power_saver_kvm_host}
+\end{figure}
+\clearpage
+
+Figure \ref{fig:power_saver_kvm_guest} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/power_saver/results_guest_report.png}
+	\caption[kvm exits default]{power\_saver guest report}
+	\label{fig:power_saver_kvm_guest}
+\end{figure}
+\clearpage
+
+Figure \ref{fig:balanced_kvm_exit} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/balanced/kvm_exit_count.png}
+	\caption[kvm exits default]{balanced kvm\_exit\_count}
+	\label{fig:balanced_kvm_exit}
+\end{figure}
+\clearpage
+
+Figure \ref{fig:balanced_kvm_host} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/balanced/results_host_report.png}
+	\caption[kvm exits default]{balanced host report}
+	\label{fig:balanced_kvm_host}
+\end{figure}
+\clearpage
+
+Figure \ref{fig:balanced_kvm_guest} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/balanced/results_guest_report.png}
+	\caption[kvm exits default]{balanced guest report}
+	\label{fig:balanced_kvm_guest}
+\end{figure}
+\clearpage
+
+Figure \ref{fig:performance_kvm_exit} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/performance/kvm_exit_count.png}
+	\caption[kvm exits default]{performance \_exit\_count}
+	\label{fig:performance_kvm_exit}
+\end{figure}
+\clearpage
+
+Figure \ref{fig:performance_kvm_host} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/performance/results_host_report.png}
+	\caption[kvm exits default]{performance host report}
+	\label{fig:performance_kvm_host}
+\end{figure}
+\clearpage
+
+Figure \ref{fig:performance_kvm_guest} shows ...
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/performance/results_guest_report.png}
+	\caption[kvm exits default]{balanced guest report}
+	\label{fig:performance_kvm_guest}
+\end{figure}
+\clearpage
+
 \begin{comment}
@@
 
-\section{Latency Comparison}
-In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4.
-For this purpose, the latency tool of the Xenomai test suite was used.
-The latency was measured under two conditions, idle and CPU-stressed.
-The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.
-
-Vorgehensweise von \cite{linPerformanceEvaluationXenomai}
-
-\subsection{Generic Ubuntu}
-
-\subsection{Real-Time Ubuntu}
-
-%In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
-While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
-The commands to kill these processes are written in the test script.
-
-After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
-\clearpage
-
-\chapter{Real-Time Latency Reduction}\label{cha:latency_reduction}
-\section{CPU isolation}
+\subsection{CPU isolation}
 Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
@@
 \clearpage
-\section{Interrupt Requests Handling}\label{sec:irq_handling}
+
+\subsection{Interrupt Requests Handling}\label{sec:irq_handling}
 Once the CPUs were isolated, Interrupt Requests Handling was the next step.
@@
 
-\section{Kernel tuning}
+\subsection{Kernel tuning}
+
+\section{Guest OS Optimization}
+
+\section{Virtual Machine Configuration Optimizations}
 
 \clearpage
+
 \section{KVM exit reasons}\label{sec:kvm_exit_reasons}
+
 \subsection{APIC\_WRITE}
@@
 The Advanced Programmable Interrupt Controller (APIC) is responsible for the distribution of interrupts in x86 and Itanium-based computer systems.
-It consists of two main components: the I/O APIC and the local APICs.
-The I/O APIC receives interrupt requests from devices and distributes them as messages to the Local APICs.
-The Local APICs then forward the highest-priority interrupt to the CPU core.
-The APIC offers many advantages, including more inputs for interrupts, a flexible configuration, definable priorities and support for message-signalled interrupts.
-
-\cite{lutsykPipelinedMulticoreMachine2020}
-%https://www.wikiwand.com/de/Advanced_Programmable_Interrupt_Controller
-
-\subsubsection{APIC virtualization}
-
-Newer Intel processors offer hardware virtualization of the Advanced Programmable Interrupt Controller (APICv).
+An APIC\_WRITE occurs when a guest operating system attempts to write to the APIC registers.
+Since the APIC is a physical hardware component, KVM must intercept this operation and cause a VM exit.
+To avoid this, newer Intel processors offer hardware virtualization of the Advanced Programmable Interrupt Controller (APICv).
 APICv improves virtualized AMD64 and Intel 64 guest performance by allowing the guest to directly access the APIC, dramatically cutting down interrupt latencies and the number of virtual machine exits caused by the APIC.
 This feature is used by default in newer Intel processors and improves I/O performance.
+
+This can be done by setting the apic flag to 'v' in the VM configuration file.
+
+Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.
+Comparing this to the previous Figures \ref{fig:kvm_exit_taskset} and \ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
+
+(URL: \url{https://www.qemu.org/docs/master/system/i386/hyperv.html})
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
+	\caption[kvm exits default]{kvm exits default}
+	\label{fig:kvm_exit_vapic}
+\end{figure}
+\clearpage
 
@@
 \subsubsection{Occurence Reduction}
+\clearpage
+
+\chapter{To Include}\label{cha:to_include}
+
+The Figure \ref{fig:latency_comparison} below compares non-optimized guest latency with optimized guest latency and includes optimized bare-metal latency as a reference.
+The data shows that a 40\% reduction in QD1 latency is achievable through system tuning.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/latency_comparison.png}
+	\caption[latency comparison]{latency comparison}
+	\label{fig:latency_comparison}
+\end{figure}
 \clearpage

### 2024-05-02 Masterarbeit_Robotik_Pamuk.tex
Figure captions

@@
 	\includegraphics[width=1.0\columnwidth]{img/power_saver/kvm_exit_count.png}
-	\caption[kvm exits default]{power\_saver kvm\_exit\_count}
+	\caption[power\_saver kvm\_exit\_count]{power\_saver kvm\_exit\_count}
 	\label{fig:power_saver_kvm_exit}
@@
 	\includegraphics[width=1.0\columnwidth]{img/power_saver/results_host_report.png}
-	\caption[kvm exits default]{power\_saver host report}
+	\caption[power\_saver host report]{power\_saver host report}
 	\label{fig:power_saver_kvm_host}
@@
 	\includegraphics[width=1.0\columnwidth]{img/power_saver/results_guest_report.png}
-	\caption[kvm exits default]{power\_saver guest report}
+	\caption[power\_saver guest report]{power\_saver guest report}
 	\label{fig:power_saver_kvm_guest}
@@
 	\includegraphics[width=1.0\columnwidth]{img/balanced/kvm_exit_count.png}
-	\caption[kvm exits default]{balanced kvm\_exit\_count}
+	\caption[balanced kvm\_exit\_count]{balanced kvm\_exit\_count}
 	\label{fig:balanced_kvm_exit}
@@
 	\includegraphics[width=1.0\columnwidth]{img/balanced/results_host_report.png}
-	\caption[kvm exits default]{balanced host report}
+	\caption[balanced host report]{balanced host report}
 	\label{fig:balanced_kvm_host}
@@
 	\includegraphics[width=1.0\columnwidth]{img/balanced/results_guest_report.png}
-	\caption[kvm exits default]{balanced guest report}
+	\caption[balanced guest report]{balanced guest report}
 	\label{fig:balanced_kvm_guest}
@@
 	\includegraphics[width=1.0\columnwidth]{img/performance/kvm_exit_count.png}
-	\caption[kvm exits default]{performance \_exit\_count}
+	\caption[performance \_exit\_count]{performance \_exit\_count}
 	\label{fig:performance_kvm_exit}
@@
 	\includegraphics[width=1.0\columnwidth]{img/performance/results_host_report.png}
-	\caption[kvm exits default]{performance host report}
+	\caption[performance host report]{performance host report}
 	\label{fig:performance_kvm_host}
@@
 	\includegraphics[width=1.0\columnwidth]{img/performance/results_guest_report.png}
-	\caption[kvm exits default]{balanced guest report}
+	\caption[balanced guest report]{balanced guest report}
 	\label{fig:performance_kvm_guest}

### 2024-05-06 masterthesis-documentation/docs/sigmatek/trace-cmd/analysis/test/analyze_tasks.py
cp841, report_hardware, literature

@@ -49,7 +49,7 @@ def count_tasks(input_filename):
     # Use f-string to include the base_filename in the output filename
     with open(f'tasks_{base_filename}.md', 'w') as f:
         # Write the total count of all events to the output file
-        f.write(f"Total count of all events: {total_count:,}\n\n".replace(",", ".")) 
+        f.write(f"Total count of all tasks: {total_count:,}\n\n".replace(",", ".")) 
 
         # Write the table headers
         f.write("| Task | PID | Count |\n")

### 2024-05-06 masterthesis-documentation/docs/sigmatek/trace-cmd/analysis/test/analyze_tasks.py
analysis/hardware

@@ -17,7 +17,7 @@ def count_tasks(input_filename):
     with open(input_filename, 'r') as f:
         for line in f:
             # Skip lines that start with "cpus=1" or "cpus=20" or "000:" or "010:"
-            if line.startswith("cpus=1") or line.startswith("cpus=20") or line.startswith("000:") or line.startswith("010:"):
+            if line.startswith("cpus=1") or line.startswith("cpus=2") or line.startswith("cpus=20") or line.startswith("000:") or line.startswith("010:"):
                 continue
 
             # Split the line into words

### 2024-05-06 Masterarbeit_Robotik_Pamuk.tex
Organize content

@@
 \chapter{Salamander 4}\label{cha:salamander4}
+This chapter breifly describes the Salamander 4 operating system by SIGMATEK.
 
@@
 
-Xenomai consists of 3 parts. These can be found in the Table \ref{tab:what_is_xenomai}.
+\noindent Salamander 4 is powered by SIGMATEK's CPU and is comprised of the following software modules:
+
+ \begin{itemize}
+	\item \textbf{Operating system}
+	\item \textbf{Loader}
+	\item \textbf{Hardware classes}
+	\item \textbf{Application}
+\end{itemize}
+
+\noindent The interfaces between the individual modules are indicated by an arrow in Figure \ref{fig:lasal_cpu}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
+	\caption[LASAL CPU]{LASAL CPU}
+	\label{fig:lasal_cpu}
+\end{figure}
+\clearpage
+\section{Xenomai}
+
+\noindent Xenomai consists of 3 parts. These can be found in the Table \ref{tab:what_is_xenomai}.
 
@@
 \clearpage
-\section{Salamander 4 Description}
-
-\subsection{Task priorities}
+
+\section{Task priorities}
 Es gibt grundsätzlich 4 Gruppen
@@
 
-\subsection{Memory Management}
+\section{Memory Management}
 Es gibt verschiedene Speicherbereiche
@@
 
-Eine LASAL CPU besteht aus den folgenden Software-Modulen:
-\begin{itemize}
-	\item Operating system
-	\item Loader
-	\item Hardware-Klassen
-\end{itemize}
-
-Die Schnittstelle zwischen den einzelnen Modulen wird in Abbildung \ref{fig:lasal_cpu} durch einen Pfeil gekennzeichnet.
-
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
-	\caption[LASAL CPU]{LASAL CPU}
-	\label{fig:lasal_cpu}
-\end{figure}
-
-\clearpage
-
+\clearpage
+
+\chapter{Initial Real-Time Latency}\label{cha:initial-real-time-latency}
+
+\clearpage
 \section{Salamander 4 Bare Metal}
@@
 	\includegraphics[width=0.75\columnwidth]{img/max_latency_vapic.png}
-	\caption[Latency taskset]{Latency taskset}
+	\caption[Latency vapic]{Latency vapic}
 	\label{fig:max_latency_vapic}
@@
 
+\subsection{Generic Ubuntu}
+
+\subsection{Real-Time Ubuntu}
+
+%In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
+While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
+The commands to kill these processes are written in the test script.
+
+After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
+
+\clearpage
 \section{Latency Comparison}
@@
 
-\subsection{Generic Ubuntu}
-
-\subsection{Real-Time Ubuntu}
-
-%In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
-While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
-The commands to kill these processes are written in the test script.
-
-After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
+\clearpage
+\section{KVM exit reasons}\label{sec:kvm_exit_reasons}
+
+\subsection{APIC\_WRITE}
+
+The Advanced Programmable Interrupt Controller (APIC) is responsible for the distribution of interrupts in x86 and Itanium-based computer systems.
+An APIC\_WRITE occurs when a guest operating system attempts to write to the APIC registers.
+Since the APIC is a physical hardware component, KVM must intercept this operation and cause a VM exit.
+To avoid this, newer Intel processors offer hardware virtualization of the Advanced Programmable Interrupt Controller (APICv).
+APICv improves virtualized AMD64 and Intel 64 guest performance by allowing the guest to directly access the APIC, dramatically cutting down interrupt latencies and the number of virtual machine exits caused by the APIC.
+This feature is used by default in newer Intel processors and improves I/O performance.
+
+This can be done by setting the apic flag to 'v' in the VM configuration file.
+
+Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.
+Comparing this to the previous Figures \ref{fig:kvm_exit_taskset} and \ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
+
+(URL: \url{https://www.qemu.org/docs/master/system/i386/hyperv.html})
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
+	\caption[kvm exits default]{kvm exits default}
+	\label{fig:kvm_exit_vapic}
+\end{figure}
+\clearpage
+
+\clearpage
+\subsection{HLT}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
+\clearpage
+\subsection{EPT\_MISCONFIG}
+\subsubsection{Description}
+\subsubsection{Occurence Reduction}
+\clearpage
+\subsection{PREEMPTION\_TIMER}
+\clearpage
+\subsection{EXTERNAL\_INTERRUPT}
+\clearpage
+\subsection{IO\_INSTRUCTION}
+\clearpage
+\subsection{EOI\_INDUCED}
+\clearpage
+\subsection{EPT\_VIOLATION}
+\clearpage
+\subsection{PAUSE\_INSTRUCTION}
+\clearpage
+\subsection{CPUID}
+\clearpage
+\subsection{MSR\_READ}
 \clearpage
@@
 \section{Host OS Optimization}
+
+\subsection{Tasks and Events}
 
@@
 
+\subsection{Tasks and Events}
+
 \section{Virtual Machine Configuration Optimizations}
@@
 
-\section{KVM exit reasons}\label{sec:kvm_exit_reasons}
-
-\subsection{APIC\_WRITE}
-
-The Advanced Programmable Interrupt Controller (APIC) is responsible for the distribution of interrupts in x86 and Itanium-based computer systems.
-An APIC\_WRITE occurs when a guest operating system attempts to write to the APIC registers.
-Since the APIC is a physical hardware component, KVM must intercept this operation and cause a VM exit.
-To avoid this, newer Intel processors offer hardware virtualization of the Advanced Programmable Interrupt Controller (APICv).
-APICv improves virtualized AMD64 and Intel 64 guest performance by allowing the guest to directly access the APIC, dramatically cutting down interrupt latencies and the number of virtual machine exits caused by the APIC.
-This feature is used by default in newer Intel processors and improves I/O performance.
-
-This can be done by setting the apic flag to 'v' in the VM configuration file.
-
-Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.
-Comparing this to the previous Figures \ref{fig:kvm_exit_taskset} and \ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
-
-(URL: \url{https://www.qemu.org/docs/master/system/i386/hyperv.html})
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
-	\caption[kvm exits default]{kvm exits default}
-	\label{fig:kvm_exit_vapic}
-\end{figure}
-\clearpage
-
-\clearpage
-\subsection{HLT}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{EPT\_MISCONFIG}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{PREEMPTION\_TIMER}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{EXTERNAL\_INTERRUPT}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{IO\_INSTRUCTION}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{EOI\_INDUCED}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{EPT\_VIOLATION}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{PAUSE\_INSTRUCTION}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{CPUID}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
-\subsection{MSR\_READ}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
-\clearpage
+\vspace{1em}
+\begin{minipage}{\linewidth}
+	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualisation},label={script:qemu_configuration_optimizations}]
+    #!/bin/sh
+
+    if  [ ! -d drive-c/ ]; then
+            echo "Filling drive-c/"
+            mkdir drive-c/
+            tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
+    fi
+
+    exec  qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
+    -m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
+    -append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable" \
+    -net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
+    -fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
+    -device vhost-vsock-pci,guest-cid=3,id=vsock0 \
+    -drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
+    -no-reboot -nographic
+\end{lstlisting}
+\end{minipage}
 
@@
 \end{figure}
+\clearpage
+
+\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
+\cite{Tracecmd}
+\cite{KernelShark}
+\cite{XenomaiXenomai}
+\cite{WelcomeYoctoProject}
+\cite{QEMU}
+\cite{linPerformanceEvaluationXenomai}
+\cite{abediRTNFPredictableLatency2019}
+\cite{broskyShieldedProcessorsGuaranteeing2003}
+\cite{casiniLatencyAnalysisVirtualization2021}
+\cite{cinqueVirtualizingMixedCriticalitySystems2022}
+\cite{garcia-vallsChallengesRealtimeVirtualization2014}
+\cite{guStateoftheArtSurveyRealTime2012}
+\cite{HardRealTime}
+\cite{javierperezHowRealTime2022}
+\cite{kirovaImpactModernVirtualization2019}
+\cite{kiszkaLinuxRealTimeHypervisor}
+\cite{lutsykPipelinedMulticoreMachine2020}
+\cite{malallahComprehensiveStudyKernel2021}
+\cite{masrurVMBasedRealTimeServices2010}
+\cite{mckenneyRealTimeVs}
+\cite{pielAsymmetricSchedulingLoad2006}
+\cite{queirozTestingLimitsGeneralpurpose2023}
+\cite{reghenzaniRealTimeLinuxKernel2020}
+\cite{yoonRealTimePerformanceAnalysis}
 \clearpage

### 2024-05-13 masterthesis-documentation/docs/sigmatek/trace-cmd/analysis/test/report.sh
qemu 2cpus

@@ -21,7 +21,7 @@ python analyze_tasks.py host_report.txt && echo "analyze_tasks for host_report c
 python analyze_tasks.py guest_report.txt && echo "analyze_tasks for guest_report completed"&
 python analyze_events.py host_report.txt && echo "analyze_events for host_report completed"&
 python analyze_events.py guest_report.txt && echo "analyze_events for guest_report completed"&
-python kvm_exit_count.py && echo "kvm_exit_count plot completed"&
+python kvm_exit_count.py kvm_exit_count.txt && echo "kvm_exit_count plot completed"&
 wait
 # Move and Copy elements in directory 
 mv *.txt $1/
@@ -34,4 +34,5 @@ cp merge.py failed_reason.py start_kernelshark.sh *.dat $1
 cd $1
 python merge.py && echo "host_report and guest_report merged"
 mv kvm_exit_count.* kvm_exits
+rm merge.py
 exit 0

### 2024-05-14 sigmatek/latency_reduction/IRQaffinity/remove_irqs_from_CPU.sh
Deployed 70e4292 with MkDocs version: 1.6.0

@@ -1,0 +1,31 @@
+#!/bin/bash
+
+# Check if a command-line argument is provided
+if [ -z "$1" ]; then
+    echo "Please provide a CPU number as a command-line argument."
+    exit 1
+fi
+
+# Get the CPU number from the command-line argument
+CPU=$1
+
+# Run the check_smp_affinity.sh script and get the IRQs
+IRQs=$(./check_smp_affinity.sh $CPU | grep -o '[0-9]\+')
+
+# Initialize an empty array to store the IRQs that could not be removed
+failed_IRQs=()
+
+# Loop over the IRQs
+for IRQ in $IRQs; do
+    # Try to change the smp_affinity
+    echo 1fff | sudo tee /proc/irq/$IRQ/smp_affinity > /dev/null 2>&1
+    # If the command failed, add the IRQ to the failed_IRQs array
+    if [ $? -ne 0 ]; then
+        failed_IRQs+=($IRQ)
+    fi
+done
+
+# Check if there were any failed IRQs
+if [ ${#failed_IRQs[@]} -ne 0 ]; then
+    echo "IRQs ${failed_IRQs[@]} could not be removed from CPU $CPU."
+fi

### 2024-05-14 masterthesis-documentation/docs/sigmatek/latency_reduction/IRQaffinity/check_CPU_IRQ_usage.sh
Hardware and OS configuration checklist

@@ -27,7 +27,7 @@ if [ -f "/proc/irq/$IRQ/smp_affinity" ]; then
     # Sort the array
     IFS=$'\n' sorted=($(sort -n <<<"${CPUs[*]}"))
     # Print the IRQ number
-    echo "IRQ $IRQ is being used by:"
+    echo "IRQ $IRQ is being used by CPU:"
     # Print the sorted CPU numbers on separate lines
     for cpu in "${sorted[@]}"; do
         echo "$cpu"

### 2024-05-16 Masterarbeit_Robotik_Pamuk.tex
compare kernels
qemu 2cpus
Hardware and OS configuration checklist
links

@@
 \clearpage
-
 \section{Task priorities}
@@
 \clearpage
-
 \section{Memory Management}
@@
 \clearpage
-
 \section{Salamander 4 Virtualisation}
@@
 \subsection{Generic Ubuntu}
-
+\clearpage
 \subsection{Real-Time Ubuntu}
@@
 \section{KVM exit reasons}\label{sec:kvm_exit_reasons}
-
 \subsection{APIC\_WRITE}
@@
 \subsection{HLT}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
+
 \clearpage
 \subsection{EPT\_MISCONFIG}
-\subsubsection{Description}
-\subsubsection{Occurence Reduction}
+
 \clearpage
 \subsection{PREEMPTION\_TIMER}
+
 \clearpage
 \subsection{EXTERNAL\_INTERRUPT}
+
 \clearpage
 \subsection{IO\_INSTRUCTION}
+
 \clearpage
 \subsection{EOI\_INDUCED}
+
 \clearpage
 \subsection{EPT\_VIOLATION}
+
 \clearpage
 \subsection{PAUSE\_INSTRUCTION}
+
 \clearpage
 \subsection{CPUID}
+
 \clearpage
 \subsection{MSR\_READ}
+
 \clearpage
@@
 \section{Host OS Optimization}
-
 \subsection{Tasks and Events}
-
+\clearpage
 \subsection{CPU governor}
@@
 
-Figure \ref{fig:gnuplot_max_max_latency_default}
+Figure \ref{fig:gnuplot_max_latency_default}
 \begin{figure}[H]
@@
 	\caption[gnuplot latency no taskset]{gnuplot latency no taskset}
-	\label{fig:gnuplot_max_max_latency_default}
-\end{figure}
-
-Figure \ref{fig:gnuplot_max_latency_with_taskset}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/with_taskset/gnuplot_max_latency_with_taskset.png}
+	\label{fig:gnuplot_max_latency_default}
+\end{figure}
+
+Figure \ref{fig:gnuplot_max_latency_taskset}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/taskset/gnuplot_max_latency_taskset.png}
 	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
-	\label{fig:gnuplot_max_latency_with_taskset}
+	\label{fig:gnuplot_max_latency_taskset}
 \end{figure}
@@
 \clearpage
-
 \subsection{CPU isolation}
@@
 \clearpage
-
 \subsection{Interrupt Requests Handling}\label{sec:irq_handling}
@@
 This reduced the interruptions caused by IRQs and the isolated CPUs were able to focus more on their assigned tasks.
-
+\clearpage
 \subsection{Kernel tuning}
+\clearpage
 
 \section{Guest OS Optimization}
-
+\clearpage
 \subsection{Tasks and Events}
+\clearpage
 
 \section{Virtual Machine Configuration Optimizations}
-
-\clearpage
 
@@
 \end{minipage}
+\clearpage
+
+\chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
+
+\section{VARAN}
+\clearpage
+\section{Robotic Application}
+\clearpage
 
@@
 \cite{yoonRealTimePerformanceAnalysis}
+\cite{adamPerformanceAssessmentLinux2021}
 \clearpage
@@
 \end{comment}
-
-\clearpage
+\clearpage
+
 \chapter{Discussion}\label{cha:discussion}
 \clearpage
+
 \chapter{Summary and Outlook}\label{cha:summary_and_outlook}
@@
 \addcontentsline{toc}{chapter}{\listacroname}
+
 \chapter*{\listacroname}

### 2024-05-16 masterthesis-documentation/docs/sigmatek/salamander4/latency_reduction/kernel-patch/compare_2_files.py
papers and configs

@@ -40,4 +40,4 @@ def compare_files(file1, file2, output_file):
     print(f"Comparison results have been written to {output_file}")
 
 # Call the function with the file names
-compare_files("config-5.15.0-1062-realtime", "config-6.5.0-28-generic", "config-comparison.txt")
+compare_files("config/config-6.8.0", "config/config-6.5.0-35-generic", "config-comparison.txt")

### 2024-05-21 Masterarbeit_Robotik_Pamuk.tex
papers and configs
minor stuff

@@
 % 3) Deckblatt, Kurzfassung, etc. ausfüllen
-% 4) und die Arbeit schreiben (die verwendeten Literaturquellen in Literatur.bib eintragen)
+% 4) und die Arbeit schreiben (die verwendeten Literaturquellen in Literature.bib eintragen)
 %
@@
 \usepackage[backend=biber, style=numeric, sorting=none]{biblatex}
-\addbibresource{Literatur.bib}
+\addbibresource{Literature.bib}
 %
@@
 
-(URL: \url{https://www.qemu.org/docs/master/system/i386/hyperv.html})
-
 \begin{figure}[H]
@@
 
-\chapter{Real-Time Performance Optimizations and Latency Reduction}\label{cha:latency_reduction}
-\section{Host OS Optimization}
+\chapter{Real-Time Performance Tuning}\label{cha:real-time_tuning}
+\section{BIOS Configurations}
+\section{Kernel Configurations} 
+\section{Host OS Configurations}
 \subsection{Tasks and Events}
@@
 \clearpage
-\subsection{Kernel tuning}
-\clearpage
-
-\section{Guest OS Optimization}
+\subsection{Real-time patch}
+\clearpage
+\section{QEMU-KVM Configurations}
+\section{Guest OS Configurations}
 \clearpage
@@
 \clearpage
-
-\section{Virtual Machine Configuration Optimizations}
 
@@
 \clearpage
+
+Configurations for guest OS in the real-time VM .......................................................... 13
 

### 2024-05-21 masterthesis-documentation/docs/sigmatek/salamander4/latency_reduction/kernel-patch/compare_2_files.py
links and compare_2_files.py

@@ -1,4 +1,15 @@
-def compare_files(file1, file2, output_file):
+import os
+import sys
+import json
+
+def compare_files(file1, file2):
+    # Get the base names of the files
+    file1_basename = os.path.basename(file1)
+    file2_basename = os.path.basename(file2)
+
+    # Create the output file name
+    output_file = f"{file1_basename}_VS_{file2_basename}.json"
+
     # Read the lines from the first file
     with open(file1, 'r') as f1:
         lines1 = set(line.strip() for line in f1)
@@ -8,36 +19,35 @@ def compare_files(file1, file2, output_file):
         lines2 = set(line.strip() for line in f2)
 
     # Find the lines common to both files
-    common_lines = lines1 & lines2
+    common_lines = list(lines1 & lines2)
 
     # Find the lines only in the first file
-    lines_only_in_file1 = lines1 - lines2
+    lines_only_in_file1 = list(lines1 - lines2)
 
     # Find the lines only in the second file
-    lines_only_in_file2 = lines2 - lines1
+    lines_only_in_file2 = list(lines2 - lines1)
 
-    # Write the results to the output file
+    # Prepare the results in a dictionary
+    results = {
+        "total common lines": len(common_lines),
+        f"both in {file1_basename} and {file2_basename}": common_lines,
+        f"count only in {file1_basename}": len(lines_only_in_file1),
+        f"only in {file1_basename}": lines_only_in_file1,
+        f"count only in {file2_basename}": len(lines_only_in_file2),
+        f"only in {file2_basename}": lines_only_in_file2
+    }
+
+    # Write the results to the output file in JSON format
     with open(output_file, 'w') as out:
-        out.write(f"Total common lines: {len(common_lines)}\n")
-        out.write("Lines both files have in common:\n")
-        for line in common_lines:
-            out.write(line + "\n")
-       
+        json.dump(results, out, indent=4)
 
-        out.write("\n" + "#"*90 + "\n" + "#"*90 + "\n" + "#"*90 + "\n")
-        out.write(f"Total lines only in {file1}: {len(lines_only_in_file1)}\n")
-        out.write("Lines only " + file1 + " has:\n")
-        for line in lines_only_in_file1:
-            out.write(line + "\n")
-        
-
-        out.write("\n" + "#"*90 + "\n" + "#"*90 + "\n" + "#"*90 + "\n")
-        out.write(f"Total lines only in {file2}: {len(lines_only_in_file2)}\n")
[…]

### 2024-05-27 Masterarbeit_Robotik_Pamuk.tex
links and compare_2_files.py
Write thesis

@@
 %This section provides necessary information about the starting point and boundary conditions of the thesis, as the work should be reproducible with the means used. 
-SIGMATEK uses its own customized Linux distribution to be run on their self-manufactured CPUs, namely Salamander 4.
-This operating system employs hard real-time with latency requirements between 20 and 50 \textmu s.
-The goal is to virtualize Salamander 4 and approach the performance of bare metal CPUs.
-Salamander 4 is virtualized through a third party service, QEMU.
+SIGMATEK uses its own customized Linux distribution, namely Salamander 4, to be run on their self-manufactured CPUs.
+Salamander 4 system employs hard real-time with Xenomai 3 and requires a worst latency value between 20 and 50 \textmu s.
+The goal is to virtualize Salamander 4 and approach the performance of bare metal.
+Salamander 4 is built with Yocto and virtualized through QEMU/KVM.
 The details of this operating system are explained in chapter \ref{cha:salamander4}.
-
-\clearpage
+Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-bare-metal}  demonstrate the inital real-time latency values gathered for bare metal and virtualisation.
+
 \section{State of the art}
-\clearpage
+
 \section{Problem and task definition}
-\clearpage
+
 \section{Objective}
@@
 It enables low-latency and deterministic execution of time-critical tasks.
-Xenomai introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
+Xenomai 3 introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
 A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
@@
 
+\begin{table}[h]
+	\centering
+	\caption[Testbed Configuration]{Testbed Configuration}
+	\label{tab:testbed_configuration}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|l|l|}
+	\hline
+	\textbf{CPU} & i7-8665U quad core, 1.9 GHz \\ \hline
+	\textbf{Memory} & 2x SO-DIMM DDR4-2400/2133 MHZ, 64GB \\ \hline
+	\textbf{GPU} & Integrated GPU \\ \hline
+	\textbf{BIOS} & Version WL37R107 \\ \hline
+	\textbf{OS} & Ubuntu 18.04 \\
+	\hline
+	\end{tabular}}
+	\end{table}
+
 \clearpage
@@
 \chapter{Salamander 4}\label{cha:salamander4}
-This chapter breifly describes the Salamander 4 operating system by SIGMATEK.
-
+This chapter briefly describes the Salamander 4 operating system by SIGMATEK.
 Salamander 4 is the proprietary operating system of SIGMATEK.
@@
 In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
-The output of the command \texttt{uname -a} can be observed in code \ref{output:uname_a}.
+The output of the command \texttt{uname -a} can be observed in code \ref{output:uname_a_virt}.
 
@@
 \begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={System information},label={output:uname_a}]
+	\begin{lstlisting}[name={System information},label={output:uname_a_virt}]
 root@sigmatek-core2:~# uname -a
@@
 
-\noindent Salamander 4 is powered by SIGMATEK's CPU and is comprised of the following software modules:
+\noindent Salamander 4 is powered by SIGMATEK's CP 841 \cite{CPUEinheitenSIGMATEK} and is comprised of the following software modules:
 
  \begin{itemize}
-	\item \textbf{Operating system}
-	\item \textbf{Loader}
-	\item \textbf{Hardware classes}
-	\item \textbf{Application}
+	\item \textbf{Operating system}: The operating system in a LASAL CPU manages the hardware and software resources of the system.
+It is provided in a completely PC-compatible manner, working with a standard PC BIOS.
+	\item \textbf{Loader}: The loader is a part of the operating system that is responsible for loading programs from executables into memory, preparing them for execution and then executing them.
+	\item \textbf{Hardware classes}: Hardware classes in LASAL represent the different types of hardware components that can be controlled by the LASAL CPU.
+They provide a way to organize and manage the hardware components in a modular and reusable manner.
+The graphical hardware editor in LASAL allows for a true-to-detail simulation of the actual hardware.
+	\item \textbf{Application}: Applications are developed using LASAL CLASS 2 \cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
 \end{itemize}
 
+\clearpage
 \noindent The interfaces between the individual modules are indicated by an arrow in Figure \ref{fig:lasal_cpu}.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
-	\caption[LASAL CPU]{LASAL CPU}
+	\includegraphics[width=0.6\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
+	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU}
 	\label{fig:lasal_cpu}
 \end{figure}
-\clearpage
-\section{Xenomai}
-
-\noindent Xenomai consists of 3 parts. These can be found in the Table \ref{tab:what_is_xenomai}.
-
-\begin{table}[!h]
-	\centering
-	\caption[Xenomai architecture]{Xenomai architecture}
-	\label{tab:what_is_xenomai}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|c|p{0.6\textwidth}|}
-			\hline
-			\textbf{Teil}  & \textbf{Beschreibung}                                                                  \\ \hline
-			i-pipe         & Kernelerweiterung für das Domain-Konzept                                               \\ \hline
-			Xenomai Kernel & Benutzt die i-pipe, und hängt sich als root-Domain ein                                 \\ \hline
-			Xenomai User   & Programme (LRT) verwenden diese Bibliothek, um Xenomai Funktionen verwenden zu können. \\ \hline
-		\end{tabular}}
-\end{table}
-
-\noindent Xenomai beruht auf einem Domain-Konzept, das bedeutet, dass alle IRQ an die erste Domain gesendet werden. (root – Domain / Xenomai)
-Nur wenn diese nichts mehr zu tun hat, dann darf die 2. Domain arbeiten.
-Das bedeutet, erst wenn alle Xenomai Task in einem Wartezustand sind, arbeiten die Linux-Tasks.
-
-\noindent In der Regel unterbricht der Prozessor beim IRQ-Handling seine aktuellen Aktivitäten, um einen Interrupt zu bearbeiten, während die IRQ-Behandlung von Xenomai einen Interrupt-Pipeline-Mechanismus verwendet, der das gleichzeitige Abrufen und Vorbereiten eines anderen Interrupts ermöglicht, während ein Interrupt bearbeitet wird, was die Leistung verbessert und die Latenzzeit verringert.
-
-\noindent Was Xenomai4 von seinem Vorgänger Xenomai3 unterscheidet, ist die vollständige Neugestaltung der Ausführungsphase mit hoher Priorität.
-Dies geschah aus Gründen der Portabilität und Wartungsfreundlichkeit: I-pipe - die zweite Iteration der ursprünglichen Adeos-Interrupt-Pipeline - wurde vollständig durch Dovetail ersetzt.
-
-\begin{table}[H]
-	\centering
-	\caption[Domain specific functions]{Domain specific functions}
-	\label{tab:domain_specific_functions}
-	\begin{tabular}{|c|c|}
-		\hline
-		\textbf{Xenomai spezifische Funktionen} & \textbf{Linux spezifische Funktionen} \\ \hline
-		Tasks                                   & Dateizugriffe                         \\ \hline
-		Mutexes, Semaphoren, Events             & Netzwerk                              \\ \hline
-	\end{tabular}
-\end{table}
-
-\noindent Ein Aufruf dieser Funktionen erfordert die entsprechende Domain. Wenn der Task in der falschen Domain läuft, dann wird ein Domain-Wechsel forciert.
-Ein Domainwechsel von Xenomai nach Linux geht relativ einfach.
-Aber der Wechsel von Linux nach Xenomai braucht Unterstützung, und dafür ist die Hilfe des Gatekeepers notwendig.
-Das bedeutet, der Gatekeeper hilft einem Task von Linux nach Xenomai zu wechseln.
-
-\clearpage
-\section{Task priorities}
-Es gibt grundsätzlich 4 Gruppen
-
-\begin{table}[ht]
-	\centering
-	\caption{Overview of the priority groups and their relationships}
-	\label{tab:priorities}
-	\begin{tabular}{|c|c|}
-		\hline
-		\textbf{Prioritätsgruppe}    & \textbf{Bereich} \\ \hline
-		Xenomai Priorität            & 0 bis 99         \\ \hline
-		Linux RT Priorität           & 1 bis 99         \\ \hline
-		Linux (Nice Level) Priorität & -20 bis 19       \\ \hline
-		RTK Priorität                & 0 bis 14         \\ \hline
-	\end{tabular}
-\end{table}
-\clearpage
-\section{Memory Management}
-Es gibt verschiedene Speicherbereiche
-
-Linux/System/Programm Speicher
-Der Speicher, den Linux und Programme belegt haben.
-Dieser Speicher ist intern in viele Teile aufgeteilt. (DMA, …)
-
-LRT-Heap Speicher
-Speicher den der LRT verwendet, oder welcher über ein CIL Funktionen
-angefordert wird.
-
-App Heap, App Code, …
+
+\noindent For the sake of completeness, figure \ref{fig:memory_management} displays the memory management of Salamander 4.
+LRT stands for Lasal Runtime and creates an execution environment where applications developed using the LASAL Class 2 can run, providing defined real-time functions, data types, and other constructs tailored for real-time programming.
 
@@
 \clearpage
+\section{Xenomai}
+\noindent Xenomai 3 is a real-time framework that offers two paths to real-time performance.
+The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in figure \ref{fig:cobalt}.
+Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\columnwidth]{img/x3-cobalt-interfaces.png}
+	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces}
+	\label{fig:cobalt}
+\end{figure}
+
+\bigskip \noindent  The second approach, called Mercury and shown in figure \ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
+Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\columnwidth]{img/x3-mercury-interfaces.png}
+	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces}
+	\label{fig:mercury}
+\end{figure}
+
+\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
+
+\clearpage
 
@@
 \clearpage
-\section{Salamander 4 Bare Metal}
-Salamander 4 Bare Metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system, including
+\section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
+Salamander 4 Bare Metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system.
+The output of the command \texttt{uname -a} can be observed in code \ref{output:uname_a_hw}.
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={System information},label={output:uname_a_hw}]
+root@sigmatek-core2:~# uname -a 
+Linux sigmatek-core2 5.15.94 #1 SMP PREEMPT IRQPIPE Tue Feb 14 18:18:05 UTC 2023 x86_64 GNU/Linux
+\end{lstlisting}
+\end{minipage}
 
@@
 \clearpage
-\section{Salamander 4 Virtualisation}
+\section{Salamander 4 Virtualisation}\label{sec:salamander4-virtualisation}
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/hardware/gnuplot_max_latency_hardware.png}
+	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
 	\caption[gnuplot latency hardware]{gnuplot latency hardware}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/default/gnuplot_max_latency_default.png}
+	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
 	\caption[gnuplot latency no taskset]{gnuplot latency no taskset}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/taskset/gnuplot_max_latency_taskset.png}
+	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
 	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}

### 2024-05-27 Masterarbeit_Robotik_Pamuk.tex
cite

@@
 This chapter briefly describes the Salamander 4 operating system by SIGMATEK.
-Salamander 4 is the proprietary operating system of SIGMATEK.
+
+\section{General}
+\noindent Salamander 4 is the proprietary operating system of SIGMATEK.
 It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment \cite{XenomaiXenomai}.
@@
 \clearpage
-\section{KVM exit reasons}\label{sec:kvm_exit_reasons}
-\subsection{APIC\_WRITE}
+\chapter{KVM exit reasons}\label{cha:kvm_exit_reasons}
+\section{APIC\_WRITE}
 
@@
 \end{figure}
-\clearpage
-
-\clearpage
-\subsection{HLT}
-
-\clearpage
-\subsection{EPT\_MISCONFIG}
-
-\clearpage
-\subsection{PREEMPTION\_TIMER}
-
-\clearpage
-\subsection{EXTERNAL\_INTERRUPT}
-
-\clearpage
-\subsection{IO\_INSTRUCTION}
-
-\clearpage
-\subsection{EOI\_INDUCED}
-
-\clearpage
-\subsection{EPT\_VIOLATION}
-
-\clearpage
-\subsection{PAUSE\_INSTRUCTION}
-
-\clearpage
-\subsection{CPUID}
-
-\clearpage
-\subsection{MSR\_READ}
+
+\section{HLT}
+\section{EPT\_MISCONFIG}
+\section{PREEMPTION\_TIMER}
+\section{EXTERNAL\_INTERRUPT}
+\section{IO\_INSTRUCTION}
+\section{EOI\_INDUCED}
+\section{EPT\_VIOLATION}
+\section{PAUSE\_INSTRUCTION}
+\section{CPUID}
+\section{MSR\_READ}
 
@@
 \subsection{Tasks and Events}
-\clearpage
 \subsection{CPU governor}
@@
 This reduced the interruptions caused by IRQs and the isolated CPUs were able to focus more on their assigned tasks.
-\clearpage
+
 \subsection{Real-time patch}
@@
 \section{Guest OS Configurations}
-\clearpage
-\subsection{Tasks and Events}
 \clearpage
@@
 
-Configurations for guest OS in the real-time VM .......................................................... 13
-
 \chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
@@
 \section{VARAN}
-\clearpage
 \section{Robotic Application}
@@
 \cite{XenomaiXenomai}
+\cite{CPUEinheitenSIGMATEK}
+\cite{EngineeringToolLASAL}
 \cite{WelcomeYoctoProject}
@@
 \cite{linPerformanceEvaluationXenomai}
-\cite{abediRTNFPredictableLatency2019}
 \cite{broskyShieldedProcessorsGuaranteeing2003}
-\cite{casiniLatencyAnalysisVirtualization2021}
 \cite{cinqueVirtualizingMixedCriticalitySystems2022}
@@
 \cite{adamPerformanceAssessmentLinux2021}
+\cite{adamRealTimePerformanceResponse2021}
+\cite{deoliveiraDemystifyingRealTimeLinux}
+\cite{RealTimePerformanceTuning2022}
 \clearpage

### 2024-05-28 Masterarbeit_Robotik_Pamuk.tex
write thesis

@@
 \schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
-\outline{Abstract}
+\outline{Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-virtualisation}  demonstrate the inital real-time latency values gathered for bare metal and virtualisation. 
+}
 \keywords{Echtzeit, Virtualisierung, Xenomai, VARAN}
@@
 The details of this operating system are explained in chapter \ref{cha:salamander4}.
-Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-bare-metal}  demonstrate the inital real-time latency values gathered for bare metal and virtualisation.
 
@@
 
-\section{General}
+\section{Structure}
 \noindent Salamander 4 is the proprietary operating system of SIGMATEK.
@@
 	\centering
-	\includegraphics[width=0.6\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
+	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
 	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU}
@@
 
-\noindent For the sake of completeness, figure \ref{fig:memory_management} displays the memory management of Salamander 4.
+\section{Memory Management}
+\noindent For the sake of completeness, Figure \ref{fig:memory_management} displays the memory management of Salamander 4.
 LRT stands for Lasal Runtime and creates an execution environment where applications developed using the LASAL Class 2 can run, providing defined real-time functions, data types, and other constructs tailored for real-time programming.
@@
 \noindent Xenomai 3 is a real-time framework that offers two paths to real-time performance.
-The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in figure \ref{fig:cobalt}.
+The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure \ref{fig:cobalt}.
 Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
@@
 
-\bigskip \noindent  The second approach, called Mercury and shown in figure \ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
+\bigskip \noindent  The second approach, called Mercury and shown in Figure \ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
 Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
@@
 
-\clearpage
+As a starting point, initial latency values of both the bare metal and virtualization versions were measured with the latency tool of the xenomai tool suite.
+Salamander 4 Bare Metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system.
+Salamander 4 virtualisation refers to a virtual version of the Salamander 4 hardware platform, achieved through QEMU/KVM.
+Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-virtualisation} specify the details of the measurements for both versions.
+In the further course, the aim was to bring the latency values of the virtualization closer to those of the hardware.
+
 \section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
-Salamander 4 Bare Metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system.
-The output of the command \texttt{uname -a} can be observed in code \ref{output:uname_a_hw}.
+The output of the command \texttt{uname -a} for Salamander 4 bare metal is shown in code \ref{output:uname_a_hw}.
 
@@
 \begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={System information},label={output:uname_a_hw}]
+	\begin{lstlisting}[name={Salamander 4 bare metal system information},label={output:uname_a_hw}]
 root@sigmatek-core2:~# uname -a 
@@
 
-Figure \ref{fig:max_latency_hardware} shows latency of hardware Salamander4.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/max_latency_hardware.png}
+\noindent As a reference point, the latency program was executed on Salamander 4 bare metal for a duration of 10 minutes.
+The complete command used was \texttt{latency -h -s -T 600}.
+Figure \ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/max_latency_hardware.png}
 	\caption[Latency hardware]{Latency hardware}
@@
 \end{figure}
-\clearpage
+
+Figure \ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
+	\caption[gnuplot latency hardware]{gnuplot latency hardware}
+	\label{fig:gnuplot_max_latency_hardware}
+\end{figure}
+
+\clearpage
+
+\noindent The statistics obtained from this measurement are provided in Table \ref{tab:latency_statistics}.
+It gives an overview of the average, maximum, minimum latency, and the standard deviation of the latency values in microseconds.
+
+\begin{table}[h]
+	\centering
+	\begin{tabular}{|c|c|}\hline
+	\textbf{Statistic} & \textbf{Value (in \textmu s)} \\\hline
+	Average Latency & 4.06 \\\hline
+	Maximum Latency & 10.71 \\\hline
+	Minimum Latency & 2.82 \\\hline
+	Standard Deviation & 0.85 \\\hline
+	\end{tabular}
+	\caption{Latency Statistics in microseconds}
+	\label{tab:latency_statistics}
+	\end{table}
+
 \section{Salamander 4 Virtualisation}\label{sec:salamander4-virtualisation}
@@
 
+Here is a description of the used components:
+\begin{itemize}
+	\item \textbf{bzImage}: Compressed Linux kernel image, loaded by QEMU at system start.
+	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU, enables UEFI boot process.
+	\item \textbf{qemu\_def.sh}: Shell script, starts QEMU with correct parameters to boot Salamamder 4 OS.
+	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Archive containing files for C drive, unpacked and copied to drive-c/ directory by qemu\_def.sh script.
+	\item \textbf{drive-c}: Directory serving as C drive for QEMU system, created and filled by qemu\_def.sh script.
+	\item \textbf{salamander-image-sigmatek-core2.ext4}: Root file system for Salamander 4 OS, used as hard drive for QEMU system.
+	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging, contains debugging symbols not present in bzImage.
+\end{itemize}
+
+\noindent The inital QEMU script after the custom Yocto build and the starting point for this work is shown in Code \ref{script:qemu_def}.
+This script will be adjusted in the course of this thesis in order to accompany real-time performance tunings.
+
 \vspace{1em}
@@
 \end{minipage}
-\clearpage
-Here is a description of the used components:
-\begin{itemize}
-	\item \textbf{bzImage}: Compressed Linux kernel image, loaded by QEMU at system start.
-	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU, enables UEFI boot process.
-	\item \textbf{qemu\_def.sh}: Shell script, starts QEMU with correct parameters to boot Salamamder 4 OS.
-	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Archive containing files for C drive, unpacked and copied to drive-c/ directory by qemu\_def.sh script.
-	\item \textbf{drive-c}: Directory serving as C drive for QEMU system, created and filled by qemu\_def.sh script.
-	\item \textbf{salamander-image-sigmatek-core2.ext4}: Root file system for Salamander 4 OS, used as hard drive for QEMU system.
-	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging, contains debugging symbols not present in bzImage.
-\end{itemize}
 
@@
 
-\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is noted bound to a specific CPU core.
+\clearpage
+\noindent Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code \ref{script:qemu_def} and no further adjustments for 10 minutes, the following latency values in Figure \ref{fig:max_latency_default} were collected.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.75\columnwidth]{img/max_latency_default.png}
+	\caption[Latency no taskset]{Latency no taskset}
+	\label{fig:max_latency_default}
+\end{figure}
+
+\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is not bound to a specific CPU core.
 This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
@@
 However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
-Consequently, these kernel tasks and interrupts can still utilize the CPU.
-
-Figure \ref{fig:max_latency_default} shows latency of QEMU default Salamander4.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.75\columnwidth]{img/max_latency_default.png}
-	\caption[Latency no taskset]{Latency no taskset}
-	\label{fig:max_latency_default}
-\end{figure}
+Consequently, these kernel tasks and interrupts can still utilize the CPU \cite{maPerformanceTuningKVMbased}.
 
@@
 The results in Figure \ref{fig:kvm_exit_taskset} were obtained.
-Additionally, table \ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
+Additionally, Table \ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
 
@@
 	\item \textbf{qemu-system-x86}: Part of the QEMU process and specifically, this task emulates x86 systems.
-In table \ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
+In Table \ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
 	\item \textbf{<idle>}: This represents the idle time of the CPU, hence it is not being used by any process, allowing to save power.
@@
 \subsection{Generic Ubuntu}
+
 \clearpage
@@
 While events like \texttt{PAUSE\_INSTRUCTION}, \texttt{EPT\_VIOLATION}, \texttt{EOI\_INDUCED}, \texttt{MSR\_READ}, and \texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
-
-The gnuplot latency is visible in Figure \ref{fig:gnuplot_max_latency_hardware}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
-	\caption[gnuplot latency hardware]{gnuplot latency hardware}
-	\label{fig:gnuplot_max_latency_hardware}
-\end{figure}
 
@@
 
+\begin{table}[h]
+	\centering
+	\begin{tabular}{|p{0.1\textwidth}|p{0.45\textwidth}|p{0.45\textwidth}|}
+	\hline
+	\textbf{Number} & \textbf{Book} & \textbf{Citation} \\
+	\hline
+	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core.
+\\
+	\hline
+	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core.
+\\
+	\hline
+	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core.
+\\
+	\hline
+	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core.
+\\
+	\hline
+	\end{tabular}
+	\caption{Your Caption}
+	\label{tab:my_label}
+	\end{table}
+
 \cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
@@
 \cite{linPerformanceEvaluationXenomai}
+\cite{maPerformanceTuningKVMbased}
 \cite{broskyShieldedProcessorsGuaranteeing2003}
@@
 \cite{RealTimePerformanceTuning2022}
+
 \clearpage

### 2024-05-28 masterthesis-documentation/docs/sigmatek/xenomai/0hardware/max_latency_hardware/max_latency_hardware.py
write thesis

@@ -23,7 +23,7 @@ plt.figure(figsize=(10, 6))
 plt.plot(x, y,'-', color='blue')
 
 # Set the title and labels
-plt.title('Latency of Operating System')
+#plt.title('Latency of Operating System')
 plt.xlabel('Time [s]')
 plt.ylabel('Latency [µs]')
 
@@ -41,8 +41,8 @@ print("The plot was successfully saved to max_latency_hardware.png")
 # Write the statistics to a file
 with open('max_latency_hardware_statistics.txt', 'w') as f:
     f.write(f"Average latency: {round(average_y, 2)}us\n") 
-    f.write(f"Max latency: {max_y}us\n")
-    f.write(f"Min latency: {min_y}us\n")
+    f.write(f"Max latency: {round(max_y, 2)}us\n")
+    f.write(f"Min latency: {round(min_y, 2)}us\n")
     f.write(f"Standard Deviation: {round(std_y, 2)}us\n")
 
 # Print a success message

### 2024-05-29 masterthesis-documentation/docs/sigmatek/xenomai/1default/max_latency_default/max_latency_default.py
minor

@@ -23,7 +23,7 @@ plt.figure(figsize=(10, 6))
 plt.plot(x, y,'-', color='blue')
 
 # Set the title and labels
-plt.title('Latency of Operating System')
+#plt.title('Latency of Operating System')
 plt.xlabel('Time [s]')
 plt.ylabel('Latency [µs]')
 
@@ -41,8 +41,8 @@ print("The plot was successfully saved to max_latency_default.png")
 # Write the statistics to a file
 with open('max_latency_default_statistics.txt', 'w') as f:
     f.write(f"Average latency: {round(average_y, 2)}us\n") 
-    f.write(f"Max latency: {max_y}us\n")
-    f.write(f"Min latency: {min_y}us\n")
+    f.write(f"Max latency: {round(max_y, 2)}us\n")
+    f.write(f"Min latency: {round(min_y, 2)}us\n")
     f.write(f"Standard Deviation: {round(std_y, 2)}us\n")
 
 # Print a success message

### 2024-05-29 Masterarbeit_Robotik_Pamuk.tex
Write thesis

@@
 
+\clearpage
+
+The system configuration is shown in Table \ref{tab:testbed_configuration}
+
 \begin{table}[h]
 	\centering
-	\caption[Testbed Configuration]{Testbed Configuration}
+	\caption[System configuration]{System configuration}
 	\label{tab:testbed_configuration}
@@
 	\hline
-	\textbf{CPU} & i7-8665U quad core, 1.9 GHz \\ \hline
-	\textbf{Memory} & 2x SO-DIMM DDR4-2400/2133 MHZ, 64GB \\ \hline
-	\textbf{GPU} & Integrated GPU \\ \hline
-	\textbf{BIOS} & Version WL37R107 \\ \hline
-	\textbf{OS} & Ubuntu 18.04 \\
+	\textbf{CPU} & 13th Gen Intel(R) Core(TM) i7-13800H \\ \hline
+	\textbf{Memory} & 2x 16GB SO-DIMM DDR5-5600 MT/s, 32GB \\ \hline
+	\textbf{GPU} & NVIDIA RTX A500 Laptop GPU \\ \hline
+	\textbf{BIOS} & Dell Version 1.12.0 \\ \hline
+	\textbf{OS} & Ubuntu 22.04.4 LTS \\
 	\hline
@@
 
-\clearpage
-
+\noindent Figure \ref{fig:lstopo} is the output of the \texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
+	\caption[Hardware topology]{Hardware topology}
+	\label{fig:lstopo}
+\end{figure}
+
+\clearpage
 \chapter{Salamander 4}\label{cha:salamander4}
@@
 	\centering
+	\caption{Latency Statistics in microseconds}
+	\label{tab:latency_statistics}
 	\begin{tabular}{|c|c|}\hline
@@
 	\end{tabular}
-	\caption{Latency Statistics in microseconds}
-	\label{tab:latency_statistics}
 	\end{table}
@@
 \noindent The inital QEMU script after the custom Yocto build and the starting point for this work is shown in Code \ref{script:qemu_def}.
-This script will be adjusted in the course of this thesis in order to accompany real-time performance tunings.
+This script is used to start QEMU with correct parameters to boot Salamamder 4 OS.
+It will be adjusted in chapter \ref{cha:real-time_tuning} in order to accompany real-time performance tunings.
 
@@
 
-%   \begin{table}[ht]
-%       \centering
-%       \caption{QEMU script components}
-%       \label{tab:qemu_components}
-%       \begin{tabular}{|l|p{8cm}|}
-%           \hline
-%           \textbf{File} & \textbf{Description} \\
-%           \hline
-%           bzImage & Compressed Linux kernel image, loaded by QEMU at system start. \\
-%           \hline
-%           ovmf.code.qcow2 & Firmware file for QEMU, enables UEFI boot process. \\
-%           \hline
-%           qemu\_def.sh & Shell script, starts QEMU with correct parameters to boot Salamamder 4 OS. \\
-%           \hline
-%           stek-drive-c-image-sigmatek-core2.tar.gz & Archive containing files for C drive, unpacked and copied to drive-c/ directory by qemu\_def.sh script. \\
-%           \hline
-%           drive-c & Directory, serves as C drive for QEMU system, created and filled by qemu\_def.sh script. \\
-%           \hline
-%           salamander-image-sigmatek-core2.ext4 & Root file system for Salamander 4 OS, used as hard drive for QEMU system. \\
-%           \hline
-%           vmlinux & Uncompressed Linux kernel image, typically used for debugging, contains debugging symbols not present in bzImage. \\
-%           \hline
-%       \end{tabular}
-%   
-%   \end{table}
-
-\clearpage
+\noindent The script is run on a generic Ubuntu 22.04.4 system. The kernel version and other details are presented in Code \ref{output:uname_a_ub}, using the \texttt{uname -a} command.
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={Ubuntu 22.04.4 system information},label={output:uname_a_ub}]
+root@sigmatek-core2:~# uname -a 
+Linux sigma-ibo 6.5.0-35-generic #35~22.04.1-Ubuntu SMP PREEMPT_DYNAMIC Tue May  7 09:00:52 UTC 2 x86_64 x86_64 x86_64 GNU/Linux
+\end{lstlisting}
+\end{minipage}
+
 \noindent Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code \ref{script:qemu_def} and no further adjustments for 10 minutes, the following latency values in Figure \ref{fig:max_latency_default} were collected.
@@
 
-\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is not bound to a specific CPU core.
-This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
-As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on user level.
+Figure \ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
+	\caption[gnuplot latency hardware]{gnuplot latency hardware}
+	\label{fig:gnuplot_max_latency_default}
+\end{figure}
+
+\noindent The statistics obtained from this measurement are provided in Table \ref{tab:x}.
+It gives an overview of the average, maximum, minimum latency, and the standard deviation of the latency values in microseconds.
+
+\clearpage
+
+\begin{table}[h]
+	\centering
+	\caption{Latency Statistics default in microseconds}
+	\label{tab:x}
+	\begin{tabular}{|c|c|}\hline
+	\textbf{Statistic} & \textbf{Value (in \textmu s)} \\\hline
+	Average Latency & 174.5us \\\hline
+	Maximum Latency & 4070.02us \\\hline
+	Minimum Latency & 6.96us \\\hline
+	Standard Deviation & 359.27us \\\hline
+	\end{tabular}
+	\end{table}
+
+\clearpage
+\chapter{Real-Time Performance Tuning}\label{cha:real-time_tuning}
+\section{BIOS Configurations}
+\begin{table}[h]
+	\centering
+	\caption{BIOS Configurations}
+	\begin{tabular}{|l|l|}
+	\hline
+	\textbf{Option} & \textbf{Status} \\
+	\hline
+	Hyper Threading & Disabled \\
+	\hline
+	Intel SpeedStep® & Disabled \\
+	\hline
+	Intel® Speed Shift Technology & Disabled \\
+	\hline
+	C States & Disabled \\
+	\hline
+	VT-d & Enabled \\
+	\hline
+	\end{tabular}
+	\end{table}
+
+\section{Kernel Configurations}
+
+The kernel command-line parameters [6] are shown in Code \ref{script:kernel_cli} below. Some of them are essential for real-time performance.
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={Kernel Configuration},label={script:kernel_cli}]
+		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 nohz_full=4 default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable noht audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup nohz=on no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
+\end{lstlisting}
+\end{minipage}
+
+\noindent In the following, these settings along with their impact on system latency are briefly described.
+
+\begin{itemize}
+	\item \textbf{isolcpus=4}: This isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned.
+	\item \textbf{rcu\_nocbs=4}: Moves the RCU (Read-Copy-Update) callback handling off of CPU 4.
+	\item \textbf{nohz\_full=4}: Makes CPU 4 tickless, reducing the timer interrupts and thus can improve performance for specific applications.
+	\item \textbf{default\_hugepagesz=1G, hugepagesz=1G, hugepages=8}: Sets the default huge page size to 1GB, and reserves 8 huge pages at boot.
+	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit) for hardware that supports it.
+	\item \textbf{rdt=l3cat}: Enables the L3 Cache Allocation Technology.
+	\item \textbf{nmi\_watchdog=0}: Disables the NMI watchdog, which can free up a bit of system resources.
+	\item \textbf{idle=poll}: Changes the idle loop of the CPU to actively poll.
+	\item \textbf{clocksource=tsc, tsc=reliable}: Sets the clocksource to TSC (Time Stamp Counter) and marks it as reliable.
+	\item \textbf{noht}: Disables Hyper-Threading.
+	\item \textbf{audit=0}: Disables the Linux audit system.
+	\item \textbf{skew\_tick=1}: Enables a mode to reduce timer interrupt overhead.
+	\item \textbf{intel\_pstate=disable}: Disables the intel\_pstate driver.
+	\item \textbf{intel.max\_cstate=0, intel\_idle.max\_cstate=0, processor.max\_cstate=0, processor\_idle.max\_cstate=0}: Disables deeper C-states, which can help reduce latency.
+	\item \textbf{nosoftlockup}: Disables the soft lockup detector.
+	\item \textbf{nohz=on}: Enables tickless operation system-wide.
+	\item \textbf{no\_timer\_check}: Disables the check for broken timer interrupt sources.
+	\item \textbf{nospectre\_v2 spectre\_v2\_user=off}: Disables mitigations for the Spectre v2 vulnerability.
+	\item \textbf{kvm.kvmclock\_periodic\_sync=N, kvm\_intel.ple\_gap=0}: These are KVM (Kernel-based Virtual Machine) related parameters.
+	\item \textbf{irqaffinity=0}: Sets the default IRQ affinity.
+\end{itemize}
+
+\section{Host OS Configurations}
+\subsection{CPU isolation}
+
 However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
@@
 
-Figure \ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.75\columnwidth]{img/max_latency_taskset.png}
-	\caption[Latency taskset]{Latency taskset}
-	\label{fig:max_latency_taskset}
-\end{figure}
-
-Figure \ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.75\columnwidth]{img/max_latency_vapic.png}
-	\caption[Latency vapic]{Latency vapic}
-	\label{fig:max_latency_vapic}
-\end{figure}
-
-\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
-However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
-Consequently, these kernel tasks and interrupts can still utilize the CPU.
-This led to the investigation of the causes for the observed high and inconsistent latency.
-The guest operates within the kvm\_entry and \texttt{kvm\_exit} events of the host.
-Kernelshark revealed a high frequency of \texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
-This frequent switching hinders the guest\'s ability to run continuously, thereby increasing the virtualization latency.
-To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
-Specifically, the causes for \texttt{kvm\_exit} events were analyzed.
-The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
-The results in Figure \ref{fig:kvm_exit_taskset} were obtained.
-Additionally, Table \ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
-
-\begin{table}[H]
-	\centering
-	\begin{tabular}{|l|p{0.62\textwidth}|}
-		\hline
-		\textbf{Exit Reason} & \textbf{Description}                                     \\
-		\hline
-		APIC\_WRITE          & Triggered when the guest writes to its APIC.             \\ \hline
-		EXTERNAL\_INTERRUPT  & Triggered by external hardware interrupts.               \\ \hline
-		HLT                  & Triggered when the guest executes the HLT instruction.   \\ \hline
-		EPT\_MISCONFIG       & Triggered by a misconfiguration in the EPT.              \\ \hline
-		PREEMPTION\_TIMER    & Triggered when the host's preemption timer expires.      \\ \hline
-		PAUSE\_INSTRUCTION   & Triggered when the PAUSE instruction is executed.        \\ \hline
-		EPT\_VIOLATION       & Triggered by a violation of the EPT permission settings. \\ \hline
-		IO\_INSTRUCTION      & Triggered when the guest executes an I/O instruction.    \\ \hline
-		EOI\_INDUCED         & Triggered when an EOI signal is sent to the APIC.        \\ \hline
-		MSR\_READ            & Triggered when the guest reads from a MSR.               \\ \hline
-		CPUID                & Triggered when the guest executes the CPUID instruction. \\ \hline
-	\end{tabular}
-	\caption{Description of kvm\_exit reasons}
-	\label{tab:kvm_exit}
-\end{table}
-
-\clearpage
-Figure \ref{fig:kvm_exit_taskset} shows \texttt{kvm\_exit} frequency with CPU islation.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_taskset.png}
-	\caption[kvm exits]{kvm exits}
-	\label{fig:kvm_exit_taskset}
-\end{figure}
-
-Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_default.png}
-	\caption[kvm exits default]{kvm exits default}
-	\label{fig:kvm_exits_default}
-\end{figure}
-
-\clearpage
-
-\begin{table}[h!]
-	\centering
-	\begin{minipage}{.5\textwidth}
-	\centering
-	\caption{Host report CPU19 (Total of 445.908)}
-	\label{tab:results_host_report}
-	\begin{tabular}{|c|c|c|}
-	\hline
-	PID & Task & Count \\ \hline
-	182579 & qemu-system-x86 & 302.748 \\ \hline
-	0 & \textless{}idle\textgreater{} & 112.911 \\ \hline
-	182618 & vhost & 21.204 \\ \hline
-	182572 & qemu-system-x86 & 7.597 \\ \hline
-	182755 & qemu-system-x86 & 644 \\ \hline
-	182754 & qemu-system-x86 & 643 \\ \hline
-	181870 & kworker/19:1 & 139 \\ \hline
-	3820 & kworker/19:1H & 16 \\ \hline
-	94 & migration/19 & 6 \\ \hline
-	\end{tabular}
-	\end{minipage}%
-	\begin{minipage}{.5\textwidth}
-	\centering
-	\caption{Guest report (Total of 362.370)}
-	\label{tab:results_guest_report}
-	\begin{tabular}{|c|c|c|}
-	\hline
-	PID & Task & Count \\ \hline
-	0 & \textless{}idle\textgreater{} & 150.744 \\ \hline
-	331 & LRT-Main & 56.697 \\ \hline
-	377 & trace-cmd & 48.507 \\ \hline
-	346 & CLI & 26.426 \\ \hline
-	378 & kthreadd & 25.837 \\ \hline
-	340 & MainTaskLow & 19.291 \\ \hline
-	339 & \textless{}...\textgreater{} & 9.980 \\ \hline
-	34 & MainTaskHigh & 9.185 \\ \hline
-	327 & LE-Logger & 4.965 \\ \hline
-	369 & kworker/0:0 & 2.793 \\ \hline
-	321 & kWorker-LRT & 2.542 \\ \hline
-	328 & LRTMgr-Main & 1.651 \\ \hline
-	34 & LrtMgrCyclic & 1.220 \\ \hline
-	332 & cobalt\_printf & 1.112 \\ \hline
-	325 & LE-System & 534 \\ \hline
-	343 & TCP-Listen & 187 \\ \hline
-	15 & rcu\_preempt & 162 \\ \hline
-	25 & kcompactd0 & 122 \\ \hline
-	58 & kworker/0:1H & 96 \\ \hline
-	63 & kworker/u2:2 & 89 \\ \hline
-	8 & jbd2/sda-8 & 86 \\ \hline
-	22 & kworker/0:1 & 56 \\ \hline
-	1 & init & 31 \\ \hline
-	2 & kthreadd & 25 \\ \hline
-	375 & trace-cmd & 24 \\ \hline
-	14 & ksoftirqd/0 & 8 \\ \hline
-	\end{tabular}
-	\end{minipage}
-	\end{table}
-
-\noindent In the following, the host and guest tasks along with their impact on system latency are briefly described.
-
-\begin{itemize} 
-	\item \textbf{qemu-system-x86}: Part of the QEMU process and specifically, this task emulates x86 systems.
-In Table \ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
-	\item \textbf{<idle>}: This represents the idle time of the CPU, hence it is not being used by any process, allowing to save power.
-The system halts until the next interrupt, which could be a timer interrupt, I/O interrupt, etc.
-	\item \textbf{vhost}: A kernel module which improves virtual input/output (virtio) performance by handling virtqueues in the kernel, thereby reducing context switches and system calls. 
-	\item \textbf{kworker/19:1}: A kernel worker thread created by the Linux kernel, kworker/19:1 performs work in response to system events.
-The number after the slash and colon indicate the CPU core and internal ID of the worker thread, respectively.
-	\item \textbf{kworker/19:1H}: Similar to kworker/19:1, kworker/19:1H is a kernel worker thread, with the ‘H’ suggesting that this thread handles hardware interrupts. 
-	\item \textbf{migration/19}: The migration process is a kernel process that balances load across CPU cores by moving threads from one CPU to another.
-The number after the slash indicates the CPU core to which the migration process is bound.
-(URL: \url{https://elixir.bootlin.com/linux/latest/source/kernel/sched/core.c#L2325})
-
-\end{itemize}
-
-\subsection{Generic Ubuntu}
-
-\clearpage
-\subsection{Real-Time Ubuntu}
-
-%In the idle condition, in order to create such test environment, some non-essential user processes were eliminated.
-While in the CPU-STRESSED condition, these process are also forcely killed, and we used the stress utility to keep the CPU busy.
-The commands to kill these processes are written in the test script.
-
-After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
-
-\clearpage
-\section{Latency Comparison}
-In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4.
-For this purpose, the latency tool of the Xenomai test suite was used.
-The latency was measured under two conditions, idle and CPU-stressed.
-The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.
-
-Vorgehensweise von \cite{linPerformanceEvaluationXenomai}
-
-\clearpage
-\chapter{KVM exit reasons}\label{cha:kvm_exit_reasons}
-\section{APIC\_WRITE}
-
-The Advanced Programmable Interrupt Controller (APIC) is responsible for the distribution of interrupts in x86 and Itanium-based computer systems.
-An APIC\_WRITE occurs when a guest operating system attempts to write to the APIC registers.
-Since the APIC is a physical hardware component, KVM must intercept this operation and cause a VM exit.
-To avoid this, newer Intel processors offer hardware virtualization of the Advanced Programmable Interrupt Controller (APICv).
-APICv improves virtualized AMD64 and Intel 64 guest performance by allowing the guest to directly access the APIC, dramatically cutting down interrupt latencies and the number of virtual machine exits caused by the APIC.
-This feature is used by default in newer Intel processors and improves I/O performance.
-
-This can be done by setting the apic flag to 'v' in the VM configuration file.
-
-Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.
-Comparing this to the previous Figures \ref{fig:kvm_exit_taskset} and \ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
-	\caption[kvm exits default]{kvm exits default}
-	\label{fig:kvm_exit_vapic}
-\end{figure}
-
-\section{HLT}
-\section{EPT\_MISCONFIG}
-\section{PREEMPTION\_TIMER}
-\section{EXTERNAL\_INTERRUPT}
-\section{IO\_INSTRUCTION}
-\section{EOI\_INDUCED}
-\section{EPT\_VIOLATION}
-\section{PAUSE\_INSTRUCTION}
-\section{CPUID}
-\section{MSR\_READ}
-
-\clearpage
-
-\chapter{Real-Time Performance Tuning}\label{cha:real-time_tuning}
-\section{BIOS Configurations}
-\section{Kernel Configurations} 
-\section{Host OS Configurations}
-\subsection{Tasks and Events}
-\subsection{CPU governor}
-
-Figure \ref{fig:power_saver_kvm_exit} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/power_saver/kvm_exit_count.png}
-	\caption[power\_saver kvm\_exit\_count]{power\_saver kvm\_exit\_count}
-	\label{fig:power_saver_kvm_exit}
-\end{figure}
-\clearpage
-
-Figure \ref{fig:power_saver_kvm_host} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/power_saver/results_host_report.png}
-	\caption[power\_saver host report]{power\_saver host report}
-	\label{fig:power_saver_kvm_host}
-\end{figure}
-\clearpage
-
-Figure \ref{fig:power_saver_kvm_guest} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/power_saver/results_guest_report.png}
-	\caption[power\_saver guest report]{power\_saver guest report}
-	\label{fig:power_saver_kvm_guest}
-\end{figure}
-\clearpage
-
-Figure \ref{fig:balanced_kvm_exit} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/balanced/kvm_exit_count.png}
-	\caption[balanced kvm\_exit\_count]{balanced kvm\_exit\_count}
-	\label{fig:balanced_kvm_exit}
-\end{figure}
-\clearpage
-
-Figure \ref{fig:balanced_kvm_host} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/balanced/results_host_report.png}
-	\caption[balanced host report]{balanced host report}
-	\label{fig:balanced_kvm_host}
-\end{figure}
-\clearpage
-
-Figure \ref{fig:balanced_kvm_guest} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/balanced/results_guest_report.png}
-	\caption[balanced guest report]{balanced guest report}
-	\label{fig:balanced_kvm_guest}
-\end{figure}
-\clearpage
-
-Figure \ref{fig:performance_kvm_exit} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/performance/kvm_exit_count.png}
-	\caption[performance \_exit\_count]{performance \_exit\_count}
-	\label{fig:performance_kvm_exit}
-\end{figure}
-\clearpage
-
-Figure \ref{fig:performance_kvm_host} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/performance/results_host_report.png}
-	\caption[performance host report]{performance host report}
-	\label{fig:performance_kvm_host}
-\end{figure}
-\clearpage
-
-Figure \ref{fig:performance_kvm_guest} shows ...
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/performance/results_guest_report.png}
-	\caption[balanced guest report]{balanced guest report}
-	\label{fig:performance_kvm_guest}
-\end{figure}
-\clearpage
-
-\begin{comment}
-\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
-This explains why there are fewer \texttt{kvm\_exit} events.
-However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a \texttt{kvm\_exit} and kvm\_entry.
-
-\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in \texttt{kvm\_exit} events.
-This is because every context switch takes place at hypervisor level.
-Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.
-\end{comment}
-
-\noindent In the process of analyzing the \texttt{kvm\_exit} events, several reasons for these exits were identified.
-The most frequent among these were the \texttt{APIC\_WRITE} and \texttt{HLT} events.
-The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
-The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
-Other significant but less frequent events included \texttt{EXTERNAL\_INTERRUPT} and \texttt{IO\_INSTRUCTION}.
-These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
-Events such as \texttt{EPT\_MISCONFIG} and \texttt{PREEMPTION\_TIMER} were also noted.
-These could potentially signal issues with memory management and the host's scheduling of the guest.
-While events like \texttt{PAUSE\_INSTRUCTION}, \texttt{EPT\_VIOLATION}, \texttt{EOI\_INDUCED}, \texttt{MSR\_READ}, and \texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
-
-Figure \ref{fig:gnuplot_max_latency_default}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
-	\caption[gnuplot latency no taskset]{gnuplot latency no taskset}
-	\label{fig:gnuplot_max_latency_default}
-\end{figure}
-
-Figure \ref{fig:gnuplot_max_latency_taskset}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
-	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
-	\label{fig:gnuplot_max_latency_taskset}
-\end{figure}
-
-Figure \ref{fig:gnuplot_max_latency_vapic}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/vapic/gnuplot_max_latency_vapic.png}
-	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
-	\label{fig:gnuplot_max_latency_vapic}
-\end{figure}
-
-\clearpage
-\subsection{CPU isolation}
 Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
@@
 		sigma_ibo@sigma-ibo:~$ cat /sys/devices/system/cpu/isolated
-		19
-		sigma_ibo@sigma-ibo:~$ ps -e -o pid,psr,comm | awk '$2 == 19'
-				92  19 cpuhp/19
-				93  19 idle_inject/19
- 	    		94  19 migration/19
- 	    		95  19 ksoftirqd/19
- 	    		97  19 kworker/19:0H-events_highpri
- 	 		 1025  19 irq/205-iwlwifi:queue_7
- 	 		17448  19 kworker/19:1H-kblockd
- 	 		17499  19 kworker/19:2-events
- 	 		18761  19 kworker/19:3-events
-			21401  19 qemu-system-x86
+		4
+		sigma_ibo@sigma-ibo:~$ ps axHo psr,pid,lwp,args,policy,nice,rtprio | awk '$1 == 4'
+		4      38      38 [cpuhp/4]                   TS    0      -
+		4      39      39 [idle_inject/4]             FF    -     50
+		4      40      40 [migration/4]               FF    -     99
+		4      41      41 [ksoftirqd/4]               TS    0      -
+		4      42      42 [kworker/4:0-events]        TS    0      -
+		4      43      43 [kworker/4:0H-events_highpr TS  -20      -
+		4     505     505 [irq/189-iwlwifi:queue_1]   FF    -     50
+		4     519     519 [irq/203-iwlwifi:exception] FF    -     50
+		4    3008    3008 [kworker/4:1H]              TS  -20      -
+		4    3022    3022 [kworker/4:2-rcu_gp]        TS    0      -	  
 \end{lstlisting}
@@
 
-\clearpage
 \subsection{Interrupt Requests Handling}\label{sec:irq_handling}
@@
 
-Output \ref{output:irq_affinity} shows the output of the script above for CPU 19.
-
 \vspace{1em}
@@
 
-\subsection{Real-time patch}
-\clearpage
+\subsection{Disable simultaneous multithreading}
+Disable
+\subsection{Disable dynamic frequency scaling}
+\subsection{Disable RT throttling}
+\subsection{No unexpected RT processes are running on your system}
+\subsection{IRQ affinity}
+\subsection{RCU CPU offloading}
+\subsection{Suppress rcu cpu stall}
+\subsection{Maybe?}
+\subsection{Start QEMU normally and give all QEMU threads rt-priority}
+\subsection{Kill all running user processes}
+\subsection{Set CPU Affinity for systemd services}
+\subsection{Cache Isolation for CPU and GPU}
+\subsection{Set CPU Affinity of IRQ thread to CPU 0}
+\subsection{Set Device Driver Work Queue to CPU 0}
+\subsection{Disable Machine Check}
+\subsection{Stop Certain Services}
+
 \section{QEMU-KVM Configurations}
+\subsection{Tune lapic timer advance}
+\subsection{Set QEMU options for real-time VM}
+\subsection{Set CPU affinity and scheduling policy of QEMU CPU threads}
+\subsection{Passthrough PCI devices into the VM}
+
 \section{Guest OS Configurations}
+
 \clearpage
@@
 
+\clearpage
+
+\chapter{KVM exit reasons}\label{cha:kvm_exit_reasons}
+
+\section{APIC\_WRITE}
+\section{HLT}
+\section{EPT\_MISCONFIG}
+\section{PREEMPTION\_TIMER}
+\section{EXTERNAL\_INTERRUPT}
+\section{IO\_INSTRUCTION}
+\section{EOI\_INDUCED}
+\section{EPT\_VIOLATION}
+\section{PAUSE\_INSTRUCTION}
+\section{CPUID}
+\section{MSR\_READ}
+
+\clearpage
+
 \chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
@@
 
+\begin{comment}
 \chapter{To Include}\label{cha:to_include}
+\section{Latency Comparison}
+In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4.
+For this purpose, the latency tool of the Xenomai test suite was used.
+The latency was measured under two conditions, idle and CPU-stressed.
+The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.
+
+Vorgehensweise von \cite{linPerformanceEvaluationXenomai}
+\section{else}
+Figure \ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.75\columnwidth]{img/max_latency_vapic.png}
+	\caption[Latency vapic]{Latency vapic}
+	\label{fig:max_latency_vapic}
+\end{figure}
+
+After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
+
+\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is not bound to a specific CPU core.
+This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
+As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on user level.
+
+Figure \ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.75\columnwidth]{img/max_latency_taskset.png}
+	\caption[Latency taskset]{Latency taskset}
+	\label{fig:max_latency_taskset}
+\end{figure}
+
+\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
+However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
+Consequently, these kernel tasks and interrupts can still utilize the CPU.
+This led to the investigation of the causes for the observed high and inconsistent latency.
+The guest operates within the kvm\_entry and \texttt{kvm\_exit} events of the host.
+Kernelshark revealed a high frequency of \texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
+This frequent switching hinders the guest\'s ability to run continuously, thereby increasing the virtualization latency.
+To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
+Specifically, the causes for \texttt{kvm\_exit} events were analyzed.
+The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
+The results in Figure \ref{fig:kvm_exit_taskset} were obtained.
+Additionally, Table \ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|l|p{0.62\textwidth}|}
+		\hline
+		\textbf{Exit Reason} & \textbf{Description}                                     \\
+		\hline
+		APIC\_WRITE          & Triggered when the guest writes to its APIC.             \\ \hline
+		EXTERNAL\_INTERRUPT  & Triggered by external hardware interrupts.               \\ \hline
+		HLT                  & Triggered when the guest executes the HLT instruction.   \\ \hline
+		EPT\_MISCONFIG       & Triggered by a misconfiguration in the EPT.              \\ \hline
+		PREEMPTION\_TIMER    & Triggered when the host's preemption timer expires.      \\ \hline
+		PAUSE\_INSTRUCTION   & Triggered when the PAUSE instruction is executed.        \\ \hline
+		EPT\_VIOLATION       & Triggered by a violation of the EPT permission settings. \\ \hline
+		IO\_INSTRUCTION      & Triggered when the guest executes an I/O instruction.    \\ \hline
+		EOI\_INDUCED         & Triggered when an EOI signal is sent to the APIC.        \\ \hline
+		MSR\_READ            & Triggered when the guest reads from a MSR.               \\ \hline
+		CPUID                & Triggered when the guest executes the CPUID instruction. \\ \hline
+	\end{tabular}
+	\caption{Description of kvm\_exit reasons}
+	\label{tab:kvm_exit}
+\end{table}
+
+\clearpage
+Figure \ref{fig:kvm_exit_taskset} shows \texttt{kvm\_exit} frequency with CPU islation.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_taskset.png}
+	\caption[kvm exits]{kvm exits}
+	\label{fig:kvm_exit_taskset}
+\end{figure}
+
+Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_default.png}
+	\caption[kvm exits default]{kvm exits default}
+	\label{fig:kvm_exits_default}
+\end{figure}
+
+\clearpage
+
+\begin{table}[h!]
+	\centering
+	\begin{minipage}{.5\textwidth}
+	\centering
+	\caption{Host report CPU19 (Total of 445.908)}
+	\label{tab:results_host_report}
+	\begin{tabular}{|c|c|c|}
+	\hline
+	PID & Task & Count \\ \hline
+	182579 & qemu-system-x86 & 302.748 \\ \hline
+	0 & \textless{}idle\textgreater{} & 112.911 \\ \hline
+	182618 & vhost & 21.204 \\ \hline
+	182572 & qemu-system-x86 & 7.597 \\ \hline
+	182755 & qemu-system-x86 & 644 \\ \hline
+	182754 & qemu-system-x86 & 643 \\ \hline
+	181870 & kworker/19:1 & 139 \\ \hline
+	3820 & kworker/19:1H & 16 \\ \hline
+	94 & migration/19 & 6 \\ \hline
+	\end{tabular}
+	\end{minipage}%
+	\begin{minipage}{.5\textwidth}
+	\centering
+	\caption{Guest report (Total of 362.370)}
+	\label{tab:results_guest_report}
+	\begin{tabular}{|c|c|c|}
+	\hline
+	PID & Task & Count \\ \hline
+	0 & \textless{}idle\textgreater{} & 150.744 \\ \hline
+	331 & LRT-Main & 56.697 \\ \hline
+	377 & trace-cmd & 48.507 \\ \hline
+	346 & CLI & 26.426 \\ \hline
+	378 & kthreadd & 25.837 \\ \hline
+	340 & MainTaskLow & 19.291 \\ \hline
+	339 & \textless{}...\textgreater{} & 9.980 \\ \hline
+	34 & MainTaskHigh & 9.185 \\ \hline
+	327 & LE-Logger & 4.965 \\ \hline
+	369 & kworker/0:0 & 2.793 \\ \hline
+	321 & kWorker-LRT & 2.542 \\ \hline
+	328 & LRTMgr-Main & 1.651 \\ \hline
+	34 & LrtMgrCyclic & 1.220 \\ \hline
+	332 & cobalt\_printf & 1.112 \\ \hline
+	325 & LE-System & 534 \\ \hline
+	343 & TCP-Listen & 187 \\ \hline
+	15 & rcu\_preempt & 162 \\ \hline
+	25 & kcompactd0 & 122 \\ \hline
+	58 & kworker/0:1H & 96 \\ \hline
+	63 & kworker/u2:2 & 89 \\ \hline
+	8 & jbd2/sda-8 & 86 \\ \hline
+	22 & kworker/0:1 & 56 \\ \hline
+	1 & init & 31 \\ \hline
+	2 & kthreadd & 25 \\ \hline
+	375 & trace-cmd & 24 \\ \hline
+	14 & ksoftirqd/0 & 8 \\ \hline
+	\end{tabular}
+	\end{minipage}
+	\end{table}
+
+\noindent In the following, the host and guest tasks along with their impact on system latency are briefly described.
+
+\begin{itemize} 
+	\item \textbf{qemu-system-x86}: Part of the QEMU process and specifically, this task emulates x86 systems.
+In Table \ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
+	\item \textbf{<idle>}: This represents the idle time of the CPU, hence it is not being used by any process, allowing to save power.
+The system halts until the next interrupt, which could be a timer interrupt, I/O interrupt, etc.
+	\item \textbf{vhost}: A kernel module which improves virtual input/output (virtio) performance by handling virtqueues in the kernel, thereby reducing context switches and system calls. 
+	\item \textbf{kworker/19:1}: A kernel worker thread created by the Linux kernel, kworker/19:1 performs work in response to system events.
+The number after the slash and colon indicate the CPU core and internal ID of the worker thread, respectively.
+	\item \textbf{kworker/19:1H}: Similar to kworker/19:1, kworker/19:1H is a kernel worker thread, with the ‘H’ suggesting that this thread handles hardware interrupts. 
+	\item \textbf{migration/19}: The migration process is a kernel process that balances load across CPU cores by moving threads from one CPU to another.
+The number after the slash indicates the CPU core to which the migration process is bound.
+(URL: \url{https://elixir.bootlin.com/linux/latest/source/kernel/sched/core.c#L2325})
+
+\end{itemize}
 
@@
 
+	Figure \ref{fig:power_saver_kvm_exit} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/power_saver/kvm_exit_count.png}
+		\caption[power\_saver kvm\_exit\_count]{power\_saver kvm\_exit\_count}
+		\label{fig:power_saver_kvm_exit}
+	\end{figure}
+	\clearpage
+
+	Figure \ref{fig:power_saver_kvm_host} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/power_saver/results_host_report.png}
+		\caption[power\_saver host report]{power\_saver host report}
+		\label{fig:power_saver_kvm_host}
+	\end{figure}
+	\clearpage
+
+	Figure \ref{fig:power_saver_kvm_guest} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/power_saver/results_guest_report.png}
+		\caption[power\_saver guest report]{power\_saver guest report}
+		\label{fig:power_saver_kvm_guest}
+	\end{figure}
+	\clearpage
+
+	Figure \ref{fig:balanced_kvm_exit} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/balanced/kvm_exit_count.png}
+		\caption[balanced kvm\_exit\_count]{balanced kvm\_exit\_count}
+		\label{fig:balanced_kvm_exit}
+	\end{figure}
+	\clearpage
+
+	Figure \ref{fig:balanced_kvm_host} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/balanced/results_host_report.png}
+		\caption[balanced host report]{balanced host report}
+		\label{fig:balanced_kvm_host}
+	\end{figure}
+	\clearpage
+
+	Figure \ref{fig:balanced_kvm_guest} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/balanced/results_guest_report.png}
+		\caption[balanced guest report]{balanced guest report}
+		\label{fig:balanced_kvm_guest}
+	\end{figure}
+	\clearpage
+
+	Figure \ref{fig:performance_kvm_exit} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/performance/kvm_exit_count.png}
+		\caption[performance \_exit\_count]{performance \_exit\_count}
+		\label{fig:performance_kvm_exit}
+	\end{figure}
+	\clearpage
+
+	Figure \ref{fig:performance_kvm_host} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/performance/results_host_report.png}
+		\caption[performance host report]{performance host report}
+		\label{fig:performance_kvm_host}
+	\end{figure}
+	\clearpage
+
+	Figure \ref{fig:performance_kvm_guest} shows ...
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{img/performance/results_guest_report.png}
+		\caption[balanced guest report]{balanced guest report}
+		\label{fig:performance_kvm_guest}
+	\end{figure}
+	\clearpage
+
+	\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
+This explains why there are fewer \texttt{kvm\_exit} events.
+However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a \texttt{kvm\_exit} and kvm\_entry.
+
+	\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in \texttt{kvm\_exit} events.
+This is because every context switch takes place at hypervisor level.
+Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.
+
+	Figure \ref{fig:gnuplot_max_latency_default}
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
+		\caption[gnuplot latency no taskset]{gnuplot latency no taskset}
+		\label{fig:gnuplot_max_latency_default}
+	\end{figure}
+
+	Figure \ref{fig:gnuplot_max_latency_taskset}
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
+		\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
+		\label{fig:gnuplot_max_latency_taskset}
+	\end{figure}
+
+	Figure \ref{fig:gnuplot_max_latency_vapic}
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/vapic/gnuplot_max_latency_vapic.png}
+		\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
+		\label{fig:gnuplot_max_latency_vapic}
+	\end{figure}
+
+	\noindent In the process of analyzing the \texttt{kvm\_exit} events, several reasons for these exits were identified.
+The most frequent among these were the \texttt{APIC\_WRITE} and \texttt{HLT} events.
+The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
+The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
+Other significant but less frequent events included \texttt{EXTERNAL\_INTERRUPT} and \texttt{IO\_INSTRUCTION}.
+These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
+Events such as \texttt{EPT\_MISCONFIG} and \texttt{PREEMPTION\_TIMER} were also noted.
+These could potentially signal issues with memory management and the host's scheduling of the guest.
+While events like \texttt{PAUSE\_INSTRUCTION}, \texttt{EPT\_VIOLATION}, \texttt{EOI\_INDUCED}, \texttt{MSR\_READ}, and \texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
+
+	The Advanced Programmable Interrupt Controller (APIC) is responsible for the distribution of interrupts in x86 and Itanium-based computer systems.
+An APIC\_WRITE occurs when a guest operating system attempts to write to the APIC registers.
+Since the APIC is a physical hardware component, KVM must intercept this operation and cause a VM exit.
+To avoid this, newer Intel processors offer hardware virtualization of the Advanced Programmable Interrupt Controller (APICv).
+APICv improves virtualized AMD64 and Intel 64 guest performance by allowing the guest to directly access the APIC, dramatically cutting down interrupt latencies and the number of virtual machine exits caused by the APIC.
+This feature is used by default in newer Intel processors and improves I/O performance.
+
+This can be done by setting the apic flag to 'v' in the VM configuration file.
+
+Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.
+Comparing this to the previous Figures \ref{fig:kvm_exit_taskset} and \ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
+	\caption[kvm exits default]{kvm exits default}
+	\label{fig:kvm_exit_vapic}
+\end{figure}
+
+\end{comment}
+
+\chapter{Results}\label{cha:results}
 \cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
@@
 \cite{QEMU}
+\cite{maPerformanceTuningKVMbased}
 \cite{linPerformanceEvaluationXenomai}
-\cite{maPerformanceTuningKVMbased}
 \cite{broskyShieldedProcessorsGuaranteeing2003}
@@
 \cite{RealTimePerformanceTuning2022}
-
-\clearpage
-
-\chapter{Results}\label{cha:results}
+\clearpage
 

### 2024-06-03 Masterarbeit_Robotik_Pamuk.tex
secnumdepth, tocdepth
minor
latency from source

@@
 \sethlcolor{blue}
+\setcounter{secnumdepth}{3} % in text 
+\setcounter{tocdepth}{3} % contents
+
 %
@@
 \clearpage
+
 \chapter{Salamander 4}\label{cha:salamander4}
@@
 	\includegraphics[width=1.0\columnwidth]{img/max_latency_hardware.png}
-	\caption[Latency hardware]{Latency hardware}
+	\caption[Latency in hardware]{Latency in hardware}
 	\label{fig:max_latency_hardware}
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
-	\caption[gnuplot latency hardware]{gnuplot latency hardware}
+	\caption[Variation in latency of hardware]{Variation in latency of hardware}
 	\label{fig:gnuplot_max_latency_hardware}
@@
 	\includegraphics[width=0.75\columnwidth]{img/max_latency_default.png}
-	\caption[Latency no taskset]{Latency no taskset}
+	\caption[Latency with default settings]{Latency with default settings}
 	\label{fig:max_latency_default}
@@
 
-Figure \ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time.
+Figure \ref{fig:gnuplot_max_latency_default} depicts the variation in latency with default settings over the course of said time.
 
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
-	\caption[gnuplot latency hardware]{gnuplot latency hardware}
+	\caption[Variation in latency with default settings]{Variation in latency with default settings}
 	\label{fig:gnuplot_max_latency_default}
@@
 \clearpage
+
 \chapter{Real-Time Performance Tuning}\label{cha:real-time_tuning}
 \section{BIOS Configurations}
+BIOS is firmware used to perform hardware initialization during the booting process.
+Some BIOS settings can have a significant impact on the real-time performance of the system [2].
+The configurations in Table 4 are tuned for real-time performance.
+
 \begin{table}[h]
@@
 	\end{table}	
-
 \section{Kernel Configurations}
@@
 	\item \textbf{no\_timer\_check}: Disables the check for broken timer interrupt sources.
-	\item \textbf{nospectre\_v2 spectre\_v2\_user=off}: Disables mitigations for the Spectre v2 vulnerability.
+	\item \textbf{nospectre\_v2, spectre\_v2\_user=off}: Disables mitigations for the Spectre v2 vulnerability.
 	\item \textbf{kvm.kvmclock\_periodic\_sync=N, kvm\_intel.ple\_gap=0}: These are KVM (Kernel-based Virtual Machine) related parameters.
@@
 \end{itemize}
-
 \section{Host OS Configurations}
@@
 \end{minipage}
-
 \subsection{Interrupt Requests Handling}\label{sec:irq_handling}
@@
 This reduced the interruptions caused by IRQs and the isolated CPUs were able to focus more on their assigned tasks.
-
-\subsection{Disable simultaneous multithreading}
-Disable
 \subsection{Disable dynamic frequency scaling}
@@
 \subsection{Stop Certain Services}
-
 \section{QEMU-KVM Configurations}
@@
 
+\clearpage
 \section{Guest OS Configurations}
-
-\clearpage
-
 \vspace{1em}
@@
 \end{minipage}
-\clearpage
-
 \clearpage
@@
 \begin{comment}
+
 \chapter{To Include}\label{cha:to_include}

### 2024-06-03 Masterarbeit_Robotik_Pamuk.tex
max_latency

@@
 \section{Xenomai}
-\noindent Xenomai 3 is a real-time framework that offers two paths to real-time performance.
+\noindent Xenomai 3 \cite{XenomaiXenomai} is a real-time framework that offers two paths to real-time performance.
 The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure \ref{fig:cobalt}.
@@
 	\textbf{Statistic} & \textbf{Value (in \textmu s)} \\\hline
-	Average Latency & 174.5us \\\hline
-	Maximum Latency & 4070.02us \\\hline
-	Minimum Latency & 6.96us \\\hline
-	Standard Deviation & 359.27us \\\hline
+	Average Latency & 46.22us \\\hline
+	Maximum Latency & 707.62us \\\hline
+	Minimum Latency & 15.59us \\\hline
+	Standard Deviation & 52.13us \\\hline
 	\end{tabular}
 	\end{table}
-
 \clearpage

### 2024-06-03 masterthesis-documentation/docs/sigmatek/xenomai/1default/max_latency_default/max_latency_default.py

@@ -4,7 +4,8 @@ import numpy as np
 # Data points
 x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,445,446,447,448,449,450,451,452,453,454,455,456,457,458,459,4
[…]

### 2024-06-03 Masterarbeit_Robotik_Pamuk.tex
change latency program

@@
 \setcounter{tocdepth}{3} % contents
+\usepackage{longtable}
+\usepackage{pdflscape}
+\usepackage{rotating}
 
@@
 
+Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
+Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section \ref{sec:irq_handling}.
+
+For CPU isolation, the \texttt{isolcpus} function was used to isolate a peroformance CPU from the general scheduling algorithms of the operating system.
+This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks or processes.
 However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
 Consequently, these kernel tasks and interrupts can still utilize the CPU \cite{maPerformanceTuningKVMbased}.
-
-Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section \ref{sec:irq_handling}.
-Output \ref{output:pse} shows the user and kernel tasks that run on CPU 19.
+Output \ref{output:pse} shows the user and kernel tasks that run on CPU 4.
 After the isolation, user tasks other than the QEMU process have been removed from running on this CPU.
@@
 		4      41      41 [ksoftirqd/4]               TS    0      -
-		4      42      42 [kworker/4:0-events]        TS    0      -
 		4      43      43 [kworker/4:0H-events_highpr TS  -20      -
@@
 		4     519     519 [irq/203-iwlwifi:exception] FF    -     50
-		4    3008    3008 [kworker/4:1H]              TS  -20      -
-		4    3022    3022 [kworker/4:2-rcu_gp]        TS    0      -	  
+		4    3008    3008 [kworker/4:1H-kblockd]      TS  -20      -
+		4  177779  177779 [kworker/4:2-events]        TS    0      -
+		4  448807  448807 [kworker/4:1-events]        TS    0      -		
 \end{lstlisting}
@@
 \subsection{Interrupt Requests Handling}\label{sec:irq_handling}
-Once the CPUs were isolated, Interrupt Requests Handling was the next step.
+Once the CPUs were isolated, Interrupt Requests handling was the next step.
 Interrupt Requests are used to send a signal to the CPU, prompting it to 'interrupt' its current task and divert its attention to another task.
 This allows hardware devices to communicate with the CPU through frequent context switches, which can lead to performance degradation, especially in high-performance computing or real-time scenarios.
-To mitigate this, the IRQs needed to be removed from the isolated CPUs.
-This was done by manipulating a file in the proc filesystem, namely \texttt{/proc/irq/<IRQ>/smp\_affinity}.
+To mitigate this, the IRQs needed to be removed from the isolated CPU.
+This was done by manipulating a file in the \texttt{proc} filesystem, namely \texttt{/proc/irq/<IRQ>/smp\_affinity}.
 The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format.
@@
 The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the significance increases towards the left until CPU19.
-In a system with 20 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be FFFFF.
+In a system with 14 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be 3FFF.
 The script in code \ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.

### 2024-06-04 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] chapter 4 finished

@@
 \schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
-\outline{Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-virtualisation}  demonstrate the inital real-time latency values gathered for bare metal and virtualisation. 
+\outline{Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-virtualization}  demonstrate the inital real-time latency values gathered for bare metal and virtualization. 
 }
@@
 First, a comprehensive literature review is conducted to understand the current trends and challenges in real-time robot control.
-Based on the literature study, a suitable virtualisation platform is selected.
-
-After the selection of the virtualisation platform, the robot control platform is implemented.
+Based on the literature study, a suitable virtualization platform is selected.
+
+After the selection of the virtualization platform, the robot control platform is implemented.
 This step includes the installation and configuration of Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
@@
 \clearpage
+
 \section{Xenomai}
@@
 As a starting point, initial latency values of both the bare metal and virtualization versions were measured with the latency tool of the xenomai tool suite.
-Salamander 4 Bare Metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system.
-Salamander 4 virtualisation refers to a virtual version of the Salamander 4 hardware platform, achieved through QEMU/KVM.
-Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-virtualisation} specify the details of the measurements for both versions.
-In the further course, the aim was to bring the latency values of the virtualization closer to those of the hardware.
+Salamander 4 bare metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system.
+Salamander 4 virtualization refers to a virtual version of the Salamander 4 hardware platform, achieved through QEMU/KVM.
+Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-virtualization} specify the details of the measurements for both versions.
+In the further course, the aim was to bring the latency values of the virtualization closer to those of the hardware and guarantee deterministic and reliable behavior.
 
@@
 \noindent As a reference point, the latency program was executed on Salamander 4 bare metal for a duration of 10 minutes.
-The complete command used was \texttt{latency -h -s -T 600}.
+The complete command used was \texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file \texttt{gnuplot.txt}.
 Figure \ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.
@@
 	\includegraphics[width=1.0\columnwidth]{img/max_latency_hardware.png}
-	\caption[Latency in hardware]{Latency in hardware}
+	\caption[Latency of Salamander 4 bare metal]{Latency of Salamander 4 bare metal}
 	\label{fig:max_latency_hardware}
@@
 
-Figure \ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time.
+Figure \ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time. Since the data varies strongly, a logarithmic scale was used for both axes.
 
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
-	\caption[Variation in latency of hardware]{Variation in latency of hardware}
+	\caption[Variation in latency of Salamander 4 bare metal]{Variation in latency of Salamander 4 bare metal}
 	\label{fig:gnuplot_max_latency_hardware}
@@
 	\begin{tabular}{|c|c|}\hline
-	\textbf{Statistic} & \textbf{Value (in \textmu s)} \\\hline
+	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
 	Average Latency & 4.06 \\\hline
@@
 
-\section{Salamander 4 Virtualisation}\label{sec:salamander4-virtualisation}
+\section{Salamander 4 Virtualization}\label{sec:salamander4-virtualization}
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
-It was developped using Yocto, an open source project that allows customised Linux distributions to be created for embedded systems \cite{WelcomeYoctoProject}.
-The virtualisation runs in a QEMU environment, which is an open source tool for hardware virtualisation \cite{QEMU}.
-With the help of the script depicted in code \ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
-This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host.
-Upon generating the necessary files, Yocto generates a QEMU folder with the following components shown in code \ref{lst:ls}.
+It was developed using Yocto, an open-source project that allows customised Linux distributions to be created for embedded systems \cite{WelcomeYoctoProject}.
+Upon generating the necessary files, Yocto provides a QEMU folder with the following components shown in code \ref{lst:ls}.
+QEMU is the environment in which the virtualization runs, as it is an open-source tool for hardware virtualization \cite{QEMU}.
 
@@
 
-Here is a description of the used components:
+\noindent With the help of the script depicted in code \ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
+This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host.
+The following is a description of the components used for the virtualization of Salamander 4:
 \begin{itemize}
-	\item \textbf{bzImage}: Compressed Linux kernel image, loaded by QEMU at system start.
-	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU, enables UEFI boot process.
-	\item \textbf{qemu\_def.sh}: Shell script, starts QEMU with correct parameters to boot Salamamder 4 OS.
-	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Archive containing files for C drive, unpacked and copied to drive-c/ directory by qemu\_def.sh script.
+	\item \textbf{bzImage}: Compressed Linux kernel image that is loaded by QEMU at system start. "bz" stands for big-zipped.
 	\item \textbf{drive-c}: Directory serving as C drive for QEMU system, created and filled by qemu\_def.sh script.
-	\item \textbf{salamander-image-sigmatek-core2.ext4}: Root file system for Salamander 4 OS, used as hard drive for QEMU system.
-	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging, contains debugging symbols not present in bzImage.
+	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process.
+\texttt{OVMF} stands for Open Virtual Machine Firmware, \texttt{qcow2} is a format for disk image files used by QEMU, it stands for "QEMU Copy On Write version 2".
+	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code \ref{script:qemu_def}.
+	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform.
+It uses the \texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
+	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Compressed tarball containing a pre-configured environment for the Salamander 4 OS.
+It is unpacked and sets up the \texttt{drive-c/} directory with system and log files in the \texttt{qemu\_def.sh} script.
+	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging.
 \end{itemize}
 
+\begin{comment}
+	https://www.baeldung.com/linux/kernel-images
+\end{comment}
+
 \noindent The inital QEMU script after the custom Yocto build and the starting point for this work is shown in Code \ref{script:qemu_def}.
-This script is used to start QEMU with correct parameters to boot Salamamder 4 OS.
+This script is used to start QEMU with required parameters to boot Salamamder 4 OS.
 It will be adjusted in chapter \ref{cha:real-time_tuning} in order to accompany real-time performance tunings.
@@
 \begin{minipage}{\linewidth}
-	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualisation},label={script:qemu_def}]
+	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualization},label={script:qemu_def}]
 	#!/bin/sh
@@
 	-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
-    -append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable" \
+	-append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4" \
 	-net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
@@
 
-\noindent The script is run on a generic Ubuntu 22.04.4 system. The kernel version and other details are presented in Code \ref{output:uname_a_ub}, using the \texttt{uname -a} command.
+\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in chapter \ref{cha:methodology}.
+The kernel version and other details are presented in Code \ref{output:uname_a_ub}, using the \texttt{uname -a} command.
 
@@
 
-Figure \ref{fig:gnuplot_max_latency_default} depicts the variation in latency with default settings over the course of said time.
+Figure \ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time with default settings.
 
@@
 
-\clearpage
-
 \begin{table}[h]
@@
 	\begin{tabular}{|c|c|}\hline
-	\textbf{Statistic} & \textbf{Value (in \textmu s)} \\\hline
-	Average Latency & 46.22us \\\hline
-	Maximum Latency & 707.62us \\\hline
-	Minimum Latency & 15.59us \\\hline
-	Standard Deviation & 52.13us \\\hline
+	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
+	Average Latency & 46.22 \\\hline
+	Maximum Latency & 707.62 \\\hline
+	Minimum Latency & 15.59 \\\hline
+	Standard Deviation & 52.13 \\\hline
 	\end{tabular}
 	\end{table}
-\clearpage
-
+
+\noindent Comparing these values to those of bare metal, it is evident that there is a significant initial gap in the statistics.
+A maximum latency of 707.62~\textmu s is not tolerable for the system and needs to be tuned.
+
+\clearpage
 \chapter{Real-Time Performance Tuning}\label{cha:real-time_tuning}
@@
 \begin{minipage}{\linewidth}
-	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualisation},label={script:qemu_configuration_optimizations}]
+	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualization},label={script:qemu_configuration_optimizations}]
     #!/bin/sh
@@
 The latency was measured under two conditions, idle and CPU-stressed.
-The goal was to optimize the latency of the virtualisation of Salamander 4 OS to closely match that of the bare metal version.
+The goal was to optimize the latency of the virtualization of Salamander 4 OS to closely match that of the bare metal version.
 
@@
 
-Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.
+Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualization.
 Comparing this to the previous Figures \ref{fig:kvm_exit_taskset} and \ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.

### 2024-06-04 Masterarbeit_Robotik_Pamuk.tex
minor

@@
 
-\noindent The statistics obtained from this measurement are provided in Table \ref{tab:latency_statistics}.
+\noindent The statistics obtained from this measurement are provided in Table \ref{tab:latency_statistics_bare_metal}.
 It gives an overview of the average, maximum, minimum latency, and the standard deviation of the latency values in microseconds.
@@
 	\centering
-	\caption{Latency Statistics in microseconds}
-	\label{tab:latency_statistics}
+	\caption{Latency statistics of Salamander 4 bare metal in microseconds}
+	\label{tab:latency_statistics_bare_metal}
 	\begin{tabular}{|c|c|}\hline
@@
 
-\noindent The statistics obtained from this measurement are provided in Table \ref{tab:x}.
+\noindent The statistics obtained from this measurement are provided in Table \ref{tab:latency_statistics_virt}.
 It gives an overview of the average, maximum, minimum latency, and the standard deviation of the latency values in microseconds.
@@
 	\centering
-	\caption{Latency Statistics default in microseconds}
-	\label{tab:x}
+	\caption{Latency statistics of default Salamander 4 virtualization in microseconds}
+	\label{tab:latency_statistics_virt}
 	\begin{tabular}{|c|c|}\hline
@@
 BIOS is firmware used to perform hardware initialization during the booting process.
-Some BIOS settings can have a significant impact on the real-time performance of the system [2].
+Some BIOS settings can have a significant impact on the real-time performance of the system.
 The configurations in Table 4 are tuned for real-time performance.
@@
 
-The kernel command-line parameters [6] are shown in Code \ref{script:kernel_cli} below. Some of them are essential for real-time performance.
+The kernel command-line parameters are shown in Code \ref{script:kernel_cli} below. Some of them are essential for real-time performance.
 

### 2024-06-04 Masterarbeit_Robotik_Pamuk.tex
~

@@
 \schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
-\outline{Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-virtualization}  demonstrate the inital real-time latency values gathered for bare metal and virtualization. 
+\outline{Sections~\ref{sec:salamander4-bare-metal} and~\ref{sec:salamander4-virtualization}  demonstrate the inital real-time latency values gathered for bare metal and virtualization. 
 }
@@
 \section{Application Context}
-This master's thesis was written at SIGMATEK GmbH \& Co KG \cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.
+This master's thesis was written at SIGMATEK GmbH \& Co KG~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.
 %This section provides necessary information about the starting point and boundary conditions of the thesis, as the work should be reproducible with the means used. 
@@
 Salamander 4 is built with Yocto and virtualized through QEMU/KVM.
-The details of this operating system are explained in chapter \ref{cha:salamander4}.
+The details of this operating system are explained in chapter~\ref{cha:salamander4}.
 
@@
 It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
-Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency \cite{Tracecmd}.
-
-\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool \cite{KernelShark}.
+Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency~\cite{Tracecmd}.
+
+\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
 It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
@@
 A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
-The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times \cite{XenomaiXenomai}.
-
-\clearpage
-
-The system configuration is shown in Table \ref{tab:testbed_configuration}
+The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times~\cite{XenomaiXenomai}.
+
+\clearpage
+
+The system configuration is shown in Table~\ref{tab:testbed_configuration}
 
@@
 	\hline
-	\textbf{CPU} & 13th Gen Intel(R) Core(TM) i7-13800H \\ \hline
-	\textbf{Memory} & 2x 16GB SO-DIMM DDR5-5600 MT/s, 32GB \\ \hline
+	\textbf{CPU} & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
+	\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB \\ \hline
 	\textbf{GPU} & NVIDIA RTX A500 Laptop GPU \\ \hline
@@
 
-\noindent Figure \ref{fig:lstopo} is the output of the \texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
+\noindent Figure~\ref{fig:lstopo} is the output of the \texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
 \begin{figure}[H]
@@
 \noindent Salamander 4 is the proprietary operating system of SIGMATEK.
-It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment \cite{XenomaiXenomai}.
+It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment~\cite{XenomaiXenomai}.
 Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture.
@@
 In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
-The output of the command \texttt{uname -a} can be observed in code \ref{output:uname_a_virt}.
+The output of the command \texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
 
@@
 
-\noindent Salamander 4 is powered by SIGMATEK's CP 841 \cite{CPUEinheitenSIGMATEK} and is comprised of the following software modules:
+\noindent Salamander 4 is powered by SIGMATEK's CP 841~\cite{CPUEinheitenSIGMATEK} and is comprised of the following software modules:
 
@@
 The graphical hardware editor in LASAL allows for a true-to-detail simulation of the actual hardware.
-	\item \textbf{Application}: Applications are developed using LASAL CLASS 2 \cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
+	\item \textbf{Application}: Applications are developed using LASAL CLASS 2~\cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
 \end{itemize}
@@
 \clearpage
-\noindent The interfaces between the individual modules are indicated by an arrow in Figure \ref{fig:lasal_cpu}.
+\noindent The interfaces between the individual modules are indicated by an arrow in Figure~\ref{fig:lasal_cpu}.
 
@@
 \section{Memory Management}
-\noindent For the sake of completeness, Figure \ref{fig:memory_management} displays the memory management of Salamander 4.
+\noindent For the sake of completeness, Figure~\ref{fig:memory_management} displays the memory management of Salamander 4.
 LRT stands for Lasal Runtime and creates an execution environment where applications developed using the LASAL Class 2 can run, providing defined real-time functions, data types, and other constructs tailored for real-time programming.
@@
 \section{Xenomai}
-\noindent Xenomai 3 \cite{XenomaiXenomai} is a real-time framework that offers two paths to real-time performance.
-The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure \ref{fig:cobalt}.
+\noindent Xenomai 3~\cite{XenomaiXenomai} is a real-time framework that offers two paths to real-time performance.
+The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
 Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
@@
 
-\bigskip \noindent  The second approach, called Mercury and shown in Figure \ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
+\bigskip \noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
 Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
@@
 Salamander 4 virtualization refers to a virtual version of the Salamander 4 hardware platform, achieved through QEMU/KVM.
-Sections \ref{sec:salamander4-bare-metal} and \ref{sec:salamander4-virtualization} specify the details of the measurements for both versions.
+Sections~\ref{sec:salamander4-bare-metal} and~\ref{sec:salamander4-virtualization} specify the details of the measurements for both versions.
 In the further course, the aim was to bring the latency values of the virtualization closer to those of the hardware and guarantee deterministic and reliable behavior.
@@
 \section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
-The output of the command \texttt{uname -a} for Salamander 4 bare metal is shown in code \ref{output:uname_a_hw}.
+The output of the command \texttt{uname -a} for Salamander 4 bare metal is shown in code~\ref{output:uname_a_hw}.
 
@@
 The complete command used was \texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file \texttt{gnuplot.txt}.
-Figure \ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.
+Figure~\ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.
 
@@
 
-Figure \ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time. Since the data varies strongly, a logarithmic scale was used for both axes.
+Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time. Since the data varies strongly, a logarithmic scale was used for both axes.
 
@@
 
-\noindent The statistics obtained from this measurement are provided in Table \ref{tab:latency_statistics_bare_metal}.
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_statistics_bare_metal}.
 It gives an overview of the average, maximum, minimum latency, and the standard deviation of the latency values in microseconds.
@@
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
-It was developed using Yocto, an open-source project that allows customised Linux distributions to be created for embedded systems \cite{WelcomeYoctoProject}.
-Upon generating the necessary files, Yocto provides a QEMU folder with the following components shown in code \ref{lst:ls}.
-QEMU is the environment in which the virtualization runs, as it is an open-source tool for hardware virtualization \cite{QEMU}.
+It was developed using Yocto, an open-source project that allows customised Linux distributions to be created for embedded systems~\cite{WelcomeYoctoProject}.
+Upon generating the necessary files, Yocto provides a QEMU folder with the following components shown in code~\ref{lst:ls}.
+QEMU is the environment in which the virtualization runs, as it is an open-source tool for hardware virtualization~\cite{QEMU}.
 
@@
 
-\noindent With the help of the script depicted in code \ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
+\noindent With the help of the script depicted in code~\ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
 This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host.
@@
 \texttt{OVMF} stands for Open Virtual Machine Firmware, \texttt{qcow2} is a format for disk image files used by QEMU, it stands for "QEMU Copy On Write version 2".
-	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code \ref{script:qemu_def}.
+	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code~\ref{script:qemu_def}.
 	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform.
@@
 
-\noindent The inital QEMU script after the custom Yocto build and the starting point for this work is shown in Code \ref{script:qemu_def}.
+\noindent The inital QEMU script after the custom Yocto build and the starting point for this work is shown in Code~\ref{script:qemu_def}.
 This script is used to start QEMU with required parameters to boot Salamamder 4 OS.
-It will be adjusted in chapter \ref{cha:real-time_tuning} in order to accompany real-time performance tunings.
+It will be adjusted in chapter~\ref{cha:real-time_tuning} in order to accompany real-time performance tunings.
 
@@
 
-\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in chapter \ref{cha:methodology}.
-The kernel version and other details are presented in Code \ref{output:uname_a_ub}, using the \texttt{uname -a} command.
+\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in chapter~\ref{cha:methodology}.
+The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the \texttt{uname -a} command.
 
@@
 
-\noindent Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code \ref{script:qemu_def} and no further adjustments for 10 minutes, the following latency values in Figure \ref{fig:max_latency_default} were collected.
+\noindent Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, the following latency values in Figure~\ref{fig:max_latency_default} were collected.
 
@@
 
-Figure \ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time with default settings.
+Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time with default settings.
 
@@
 
-\noindent The statistics obtained from this measurement are provided in Table \ref{tab:latency_statistics_virt}.
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_statistics_virt}.
 It gives an overview of the average, maximum, minimum latency, and the standard deviation of the latency values in microseconds.
@@
 \clearpage
+
 \chapter{Real-Time Performance Tuning}\label{cha:real-time_tuning}
+
+In this chapter, the significant initial gap in latency statistics between the virtualized system and the bare metal system is tackled.
+For this reason, an extensive tuning process is carried out.
+This involves configurations spanning the BIOS, kernel, host OS, QEMU-KVM virtualization layer, and the Salamander 4 OS itself.
+The individual configurations will be discussed in detail and the modifications will be justified with a clear explanation.
+The goal is to bring the latency of the virtualized system closer to that of the bare metal system, thereby ensuring deterministic behavior under real-time constraints.
+
 \section{BIOS Configurations}
-BIOS is firmware used to perform hardware initialization during the booting process.
-Some BIOS settings can have a significant impact on the real-time performance of the system.
-The configurations in Table 4 are tuned for real-time performance.
+
+BIOS stands for Basic Input/Output System.
+It abstracts the hardware and enables basic functions of a computer during the booting process, such as starting the operating system and loading other software.
+Since the BIOS is embedded very deep, its configuration can significantly influence the real-time performance of the system.
+Table~\ref{tab:bios_configuration} illustrates the specific BIOS settings that have been adjusted for the purpose of real-time performance.
 
@@
 	\caption{BIOS Configurations}
+	\label{tab:bios_configuration}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
 	\begin{tabular}{|l|l|}
@@
 	\hline
-	\end{tabular}
+	\end{tabular}}
 	\end{table}
+
+	\noindent In the following, these settings along with their impact on system latency are briefly described.
+
+	\begin{itemize}
+		\item \textbf{Hyper Threading}: Explain
+		\item \textbf{Intel SpeedStep}: Explain
+		\item \textbf{Intel® Speed Shift Technology}: Explain
+		\item \textbf{C States}: Explain
+		\item \textbf{VT-d}: Explain
+	\end{itemize}
+
+\clearpage
+
 \section{Kernel Configurations}
 
-The kernel command-line parameters are shown in Code \ref{script:kernel_cli} below. Some of them are essential for real-time performance.
+The kernel command-line parameters are shown in Code~\ref{script:kernel_cli} below. Some of them are essential for real-time performance.
 
@@
 Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section \ref{sec:irq_handling}.
+Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section~\ref{sec:irq_handling}.
 
@@
 However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
-Consequently, these kernel tasks and interrupts can still utilize the CPU \cite{maPerformanceTuningKVMbased}.
-Output \ref{output:pse} shows the user and kernel tasks that run on CPU 4.
+Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}.
+Output~\ref{output:pse} shows the user and kernel tasks that run on CPU 4.
 After the isolation, user tasks other than the QEMU process have been removed from running on this CPU.
@@
 In a system with 14 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be 3FFF.
-The script in code \ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.
+The script in code~\ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.
 
@@
 
-Vorgehensweise von \cite{linPerformanceEvaluationXenomai}
+Vorgehensweise von~\cite{linPerformanceEvaluationXenomai}
 \section{else}
-Figure \ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
+Figure~\ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
 \begin{figure}[H]
@@
 
-Figure \ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
+Figure~\ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
 \begin{figure}[H]
@@
 The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
-The results in Figure \ref{fig:kvm_exit_taskset} were obtained.
-Additionally, Table \ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
+The results in Figure~\ref{fig:kvm_exit_taskset} were obtained.
+Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
 
@@
 \clearpage
-Figure \ref{fig:kvm_exit_taskset} shows \texttt{kvm\_exit} frequency with CPU islation.
+Figure~\ref{fig:kvm_exit_taskset} shows \texttt{kvm\_exit} frequency with CPU islation.
 \begin{figure}[H]
@@
 
-Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
+Figure~\ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
 \begin{figure}[H]
@@
 	\item \textbf{qemu-system-x86}: Part of the QEMU process and specifically, this task emulates x86 systems.
-In Table \ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
+In Table~\ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
 	\item \textbf{<idle>}: This represents the idle time of the CPU, hence it is not being used by any process, allowing to save power.
@@
 
-The Figure \ref{fig:latency_comparison} below compares non-optimized guest latency with optimized guest latency and includes optimized bare-metal latency as a reference.
+The Figure~\ref{fig:latency_comparison} below compares non-optimized guest latency with optimized guest latency and includes optimized bare-metal latency as a reference.
 The data shows that a 40\% reduction in QD1 latency is achievable through system tuning.
@@
 
-	Figure \ref{fig:power_saver_kvm_exit} shows ...
+	Figure~\ref{fig:power_saver_kvm_exit} shows ...
 
@@
 
-	Figure \ref{fig:power_saver_kvm_host} shows ...
+	Figure~\ref{fig:power_saver_kvm_host} shows ...
 
@@
 
-	Figure \ref{fig:power_saver_kvm_guest} shows ...
+	Figure~\ref{fig:power_saver_kvm_guest} shows ...
 
@@
 
-	Figure \ref{fig:balanced_kvm_exit} shows ...
+	Figure~\ref{fig:balanced_kvm_exit} shows ...
 
@@
 
-	Figure \ref{fig:balanced_kvm_host} shows ...
+	Figure~\ref{fig:balanced_kvm_host} shows ...
 
@@
 
-	Figure \ref{fig:balanced_kvm_guest} shows ...
+	Figure~\ref{fig:balanced_kvm_guest} shows ...
 
@@
 
-	Figure \ref{fig:performance_kvm_exit} shows ...
+	Figure~\ref{fig:performance_kvm_exit} shows ...
 
@@
 
-	Figure \ref{fig:performance_kvm_host} shows ...
+	Figure~\ref{fig:performance_kvm_host} shows ...
 
@@
 
-	Figure \ref{fig:performance_kvm_guest} shows ...
+	Figure~\ref{fig:performance_kvm_guest} shows ...
 
@@
 
-	Figure \ref{fig:gnuplot_max_latency_default}
+	Figure~\ref{fig:gnuplot_max_latency_default}
 	\begin{figure}[H]
@@
 
-	Figure \ref{fig:gnuplot_max_latency_taskset}
+	Figure~\ref{fig:gnuplot_max_latency_taskset}
 	\begin{figure}[H]
@@
 
-	Figure \ref{fig:gnuplot_max_latency_vapic}
+	Figure~\ref{fig:gnuplot_max_latency_vapic}
 	\begin{figure}[H]
@@
 
-Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualization.
-Comparing this to the previous Figures \ref{fig:kvm_exit_taskset} and \ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
+Figure~\ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualization.
+Comparing this to the previous Figures~\ref{fig:kvm_exit_taskset} and~\ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
 
@@
 \chapter{Results}\label{cha:results}
-\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
-\cite{Tracecmd}
-\cite{KernelShark}
-\cite{XenomaiXenomai}
-\cite{CPUEinheitenSIGMATEK}
-\cite{EngineeringToolLASAL}
-\cite{WelcomeYoctoProject}
-\cite{QEMU}
-\cite{maPerformanceTuningKVMbased}
-\cite{linPerformanceEvaluationXenomai}
-\cite{broskyShieldedProcessorsGuaranteeing2003}
-\cite{cinqueVirtualizingMixedCriticalitySystems2022}
-\cite{garcia-vallsChallengesRealtimeVirtualization2014}
-\cite{guStateoftheArtSurveyRealTime2012}
-\cite{HardRealTime}
-\cite{javierperezHowRealTime2022}
-\cite{kirovaImpactModernVirtualization2019}
-\cite{kiszkaLinuxRealTimeHypervisor}
-\cite{lutsykPipelinedMulticoreMachine2020}
-\cite{malallahComprehensiveStudyKernel2021}
-\cite{masrurVMBasedRealTimeServices2010}
-\cite{mckenneyRealTimeVs}
-\cite{pielAsymmetricSchedulingLoad2006}
-\cite{queirozTestingLimitsGeneralpurpose2023}
-\cite{reghenzaniRealTimeLinuxKernel2020}
-\cite{yoonRealTimePerformanceAnalysis}
-\cite{adamPerformanceAssessmentLinux2021}
-\cite{adamRealTimePerformanceResponse2021}
-\cite{deoliveiraDemystifyingRealTimeLinux}
-\cite{RealTimePerformanceTuning2022}
-\clearpage
+~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
+~\cite{Tracecmd}
+~\cite{KernelShark}
+~\cite{XenomaiXenomai}
+~\cite{CPUEinheitenSIGMATEK}
+~\cite{EngineeringToolLASAL}
+~\cite{WelcomeYoctoProject}
+~\cite{QEMU}
+~\cite{maPerformanceTuningKVMbased}
+~\cite{linPerformanceEvaluationXenomai}
+~\cite{broskyShieldedProcessorsGuaranteeing2003}
+~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
+~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
+~\cite{guStateoftheArtSurveyRealTime2012}
+~\cite{HardRealTime}
+~\cite{javierperezHowRealTime2022}
+~\cite{kirovaImpactModernVirtualization2019}
+~\cite{kiszkaLinuxRealTimeHypervisor}
+~\cite{lutsykPipelinedMulticoreMachine2020}
+~\cite{malallahComprehensiveStudyKernel2021}
+~\cite{masrurVMBasedRealTimeServices2010}
+~\cite{mckenneyRealTimeVs}
+~\cite{pielAsymmetricSchedulingLoad2006}
+~\cite{queirozTestingLimitsGeneralpurpose2023}
+~\cite{reghenzaniRealTimeLinuxKernel2020}
+~\cite{yoonRealTimePerformanceAnalysis}
+~\cite{adamPerformanceAssessmentLinux2021}
+~\cite{adamRealTimePerformanceResponse2021}
+~\cite{deoliveiraDemystifyingRealTimeLinux}
+~\cite{RealTimePerformanceTuning2022}
+~\clearpage
 
@@
 \chapter{Summary and Outlook}\label{cha:summary_and_outlook}
-
 %

### 2024-06-04 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] BIOS Configurations finished

@@
 	\begin{itemize}
-		\item \textbf{Hyper Threading}: Explain
-		\item \textbf{Intel SpeedStep}: Explain
-		\item \textbf{Intel® Speed Shift Technology}: Explain
-		\item \textbf{C States}: Explain
-		\item \textbf{VT-d}: Explain
+		\item \textbf{Hyper Threading}: Hyper-Threading allows CPUs to process two threads simultaneously instead of just one.
+When it is enabled on the host, this allows the parallelisation of tasks and increases performance.
+However, in a real-time system like the guest Salamander 4, this can lead to increased latencies due to contention between threads.
+In order to ensure more deterministic behavior in the guest, it is disabled on the host.
+		\item \textbf{Intel SpeedStep}: This technology dynamically adjusts the clock speed of the CPU based on workload.
+These dynamic changes can lead to unpredictable latencies in a real-time system.
+It is also disabled on the host to maintain a constant CPU speed.
+		\item \textbf{Intel® Speed Shift Technology}: Similar to SpeedStep, Speed Shift allows the processor to directly control its frequency and voltage.
+This can lead to unpredictable latencies, too.
+Hence, it is also disabled on the host.
+		\item \textbf{C States}: These are low-power idle states where the clock frequency and voltage of the CPU are reduced.
+Transitioning between C-states can cause variable latencies.
+To prevent this from happening, C-states are disabled on the host.
+		\item \textbf{VT-d}: Direct access to physical devices from within virtual machines is possible when VT-d is enabled on the host.
+This can help reduce latencies associated with I/O operations in the virtual machine.
+It is therefore enabled on the host.
 	\end{itemize}
-
-\clearpage
 

### 2024-06-05 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Kernel Configurations 10/19 finished

@@
 \begin{itemize}
-	\item \textbf{bzImage}: Compressed Linux kernel image that is loaded by QEMU at system start. "bz" stands for big-zipped.
+	\item \textbf{bzImage}: Compressed Linux kernel image that is loaded by QEMU at system start. ``bz`` stands for big-zipped.
 	\item \textbf{drive-c}: Directory serving as C drive for QEMU system, created and filled by qemu\_def.sh script.
@@
 	\begin{lstlisting}[name={Kernel Configuration},label={script:kernel_cli}]
-		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 nohz_full=4 default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable noht audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup nohz=on no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
+		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 nohz_full=4 default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup nohz=on no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
 \end{lstlisting}
@@
 \begin{itemize}
-	\item \textbf{isolcpus=4}: This isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned.
-	\item \textbf{rcu\_nocbs=4}: Moves the RCU (Read-Copy-Update) callback handling off of CPU 4.
-	\item \textbf{nohz\_full=4}: Makes CPU 4 tickless, reducing the timer interrupts and thus can improve performance for specific applications.
-	\item \textbf{default\_hugepagesz=1G, hugepagesz=1G, hugepages=8}: Sets the default huge page size to 1GB, and reserves 8 huge pages at boot.
-	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit) for hardware that supports it.
-	\item \textbf{rdt=l3cat}: Enables the L3 Cache Allocation Technology.
+	\item \textbf{isolcpus=4}: Isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned.
+CPU isolation is explained in detail in section \ref{sec:cpu_isolation}.
+	\item \textbf{rcu\_nocbs=4}: The Linux kernel uses a synchronization mechanism called RCU, or Read-Copy-Update.
+It lets writers update the data in a way that guarantees readers will always see the same version while enabling multiple readers to access shared data without locks.
+The RCU subsystem uses callback functions that need to be invoked once readers are done with the data they accessed.
+By default, these callbacks are handled by the CPUs that executed the read-side critical sections.
+This parameter offloads RCU callback handling from CPU 4 to other CPUs.
+CPU 4 remains dedicated to high-priority tasks which helps in reducing latency.
+	\item \textbf{nohz\_full=4}: Makes CPU core 4 "tickless", meaning the kernel tries to avoid sending periodic scheduling-clock interrupts to the CPU when there are no runnable tasks.
+This lowers latency by reducing unnecessary wake-ups but may increase power consumption because the CPU is not able to enter a low-power state when idle.
+Additionally, timer interrupts cannot be fully eliminated because certain events, such as incoming interrupts or task activations, can still cause the kernel to send timer interrupts to the tickless CPU.
+	\item \textbf{default\_hugepagesz=1G, hugepagesz=1G, hugepages=8}: Huge pages are large contiguous areas of memory that can be used by applications and the kernel, instead of the traditional 4KB small pages.
+The default huge page size is set to 1GB, and 8 huge pages of 1GB size are reserved at boot.
+This pre-allocation makes sure that these large memory regions are available to be used by the kernel or applications, without having to dynamically allocate and potentially fail.
+	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit), which connects a Direct Memory Access (DMA)-capable I/O bus, such as graphics cards and network adapters, to the main memory.
+It can enhance device performance by allowing these devices to directly access and use memory, which is especially helpful when these devices are virtualized.
+	\item \textbf{rdt=l3cat}: Activates the L3 Cache Allocation Technology (L3 CAT) feature of Intel’s Resource Director Technology (RDT).
+Unlike L1 and L2 caches, where each core has its fixed capacity, L3 cache is a shared pool among multiple cores.
+L3 CAT is a mechanism that controls the amount of L3 cache that a process can use.
+By controlling cache allocation, it can prevent a single process from monopolizing the L3 cache, which is particularly beneficial in virtualized environments, where multiple virtual machines share the same physical host.
+\begin{comment}
 \item \textbf{nmi\_watchdog=0}: Disables the NMI watchdog, which can free up a bit of system resources.
-	\item \textbf{idle=poll}: Changes the idle loop of the CPU to actively poll.
+\end{comment}
+\item \textbf{idle=poll}: Changes the CPU’s idle loop behavior to active polling. Instead of entering a low-power state when idle, the CPU continuously polls for new 
+tasks. This can reduce task start latency in real-time systems, but it increases power consumption.
+\begin{comment}
 \item \textbf{clocksource=tsc, tsc=reliable}: Sets the clocksource to TSC (Time Stamp Counter) and marks it as reliable.
-	\item \textbf{noht}: Disables Hyper-Threading.
+\end{comment}
 	\item \textbf{audit=0}: Disables the Linux audit system.
+When it is enabled, it generates log entries for security-relevant events, which is a slow operation since they are written to disk.
+If there are a large number of such events, the audit system can consume significant CPU time and I/O bandwidth which could lead to higher latency.
+\begin{comment}
 \item \textbf{skew\_tick=1}: Enables a mode to reduce timer interrupt overhead.
@@
 \item \textbf{intel.max\_cstate=0, intel\_idle.max\_cstate=0, processor.max\_cstate=0, processor\_idle.max\_cstate=0}: Disables deeper C-states, which can help reduce latency.
-	\item \textbf{nosoftlockup}: Disables the soft lockup detector.
+\end{comment}
+	\item \textbf{nosoftlockup}: Disables the soft lockup detector in the Linux kernel.
+A soft lockup is when a CPU is busy executing kernel code for a long period of time without giving other tasks a chance to run.
+Especially threads with SCHED\_FIFO policy occupy the CPU for an extensive duration.
+This is detected and reported by the soft lockup detector, hence it is disabled to prevent these unnecessary warnings.
+\begin{comment}
 \item \textbf{nohz=on}: Enables tickless operation system-wide.
@@
 \item \textbf{kvm.kvmclock\_periodic\_sync=N, kvm\_intel.ple\_gap=0}: These are KVM (Kernel-based Virtual Machine) related parameters.
-	\item \textbf{irqaffinity=0}: Sets the default IRQ affinity.
+\end{comment}
+	\item \textbf{irqaffinity=0}: Sets the default Interrupt Request affinity to none.
+This means that no CPU core is preferred over another for handling IRQs.
+Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs.
+
 \end{itemize}
+
 \section{Host OS Configurations}
-\subsection{CPU isolation}
+\subsection{CPU isolation}\label{sec:cpu_isolation}
 
@@
 
-For CPU isolation, the \texttt{isolcpus} function was used to isolate a peroformance CPU from the general scheduling algorithms of the operating system.
+For CPU isolation, the \texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
 This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks or processes.
@@
 Once the CPUs were isolated, Interrupt Requests handling was the next step.
-Interrupt Requests are used to send a signal to the CPU, prompting it to 'interrupt' its current task and divert its attention to another task.
+Interrupt Requests are used to send a signal to the CPU, prompting it to `interrupt' its current task and divert its attention to another task.
 This allows hardware devices to communicate with the CPU through frequent context switches, which can lead to performance degradation, especially in high-performance computing or real-time scenarios.
@@
 
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
-\begin{lstlisting}[name={Output of smp_affinity for CPU 19},label={output:irq_affinity}, breaklines=true]
-	sigma_ibo@sigma-ibo:~$ ./check_smp_affinity.sh 19
-	CPU 19 IRQ affinity:
-	0
-	2
-	3
-	4
-	5
-	6
-	7
-	10
-	11
-	13
-	15
-	131
-	172
-	188
-	189
-	195
-\end{lstlisting}
-\end{minipage}
-
 \noindent By changing the values of the \texttt{smp\_affinity} files of the respective IRQs, the assignment of IRQs was controlled so that they would not be handled by the isolated CPU.
@@
 \section{Guest OS Configurations}
-\vspace{1em}
-\begin{minipage}{\linewidth}
-	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualization},label={script:qemu_configuration_optimizations}]
-    #!/bin/sh
-
-    if  [ ! -d drive-c/ ]; then
-            echo "Filling drive-c/"
-            mkdir drive-c/
-            tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
-    fi
-
-    exec  qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
-    -m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
-    -append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable" \
-    -net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
-    -fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
-    -device vhost-vsock-pci,guest-cid=3,id=vsock0 \
-    -drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
-    -no-reboot -nographic
-\end{lstlisting}
-\end{minipage}
+
 \clearpage

### 2024-06-06 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Kernel Configurations 16/19 finished

@@
 %- Anhang (optional)
+
 \chapter{Introduction}\label{cha:introduction}
@@
 
-Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time. Since the data varies strongly, a logarithmic scale was used for both axes.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
-	\caption[Variation in latency of Salamander 4 bare metal]{Variation in latency of Salamander 4 bare metal}
-	\label{fig:gnuplot_max_latency_hardware}
-\end{figure}
-
-\clearpage
-
 \noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_statistics_bare_metal}.
@@
 	\end{table}
+
+\noindent Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time. Since the data varies strongly, a logarithmic scale was used for both axes.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
+	\caption[Variation in latency of Salamander 4 bare metal]{Variation in latency of Salamander 4 bare metal}
+	\label{fig:gnuplot_max_latency_hardware}
+\end{figure}
 
@@
 
-Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time with default settings.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
-	\caption[Variation in latency with default settings]{Variation in latency with default settings}
-	\label{fig:gnuplot_max_latency_default}
-\end{figure}
-
 \noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_statistics_virt}.
@@
 
-\noindent Comparing these values to those of bare metal, it is evident that there is a significant initial gap in the statistics.
+\noindent Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time with default settings.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
+	\caption[Variation in latency with default settings]{Variation in latency with default settings}
+	\label{fig:gnuplot_max_latency_default}
+\end{figure}
+
+\noindent Comparing these values to those of bare metal in Figure \ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in the statistics.
 A maximum latency of 707.62~\textmu s is not tolerable for the system and needs to be tuned.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/01combined/gnuplot_combined_max_latency.png}
+	\caption[Comparison of variation in latency between hardware and virtualization]{Comparison of variation in latency between hardware and virtualization}
+	\label{fig:gnuplot_max_latency_combined}
+\end{figure}
 
@@
 	\begin{lstlisting}[name={Kernel Configuration},label={script:kernel_cli}]
-		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 nohz_full=4 default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup nohz=on no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
+		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 nohz_full=4 nohz=on default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
 \end{lstlisting}
@@
 Additionally, timer interrupts cannot be fully eliminated because certain events, such as incoming interrupts or task activations, can still cause the kernel to send timer interrupts to the tickless CPU.
+	\item \textbf{nohz=on}: Sets all CPUs to tickless mode system-wide.
 	\item \textbf{default\_hugepagesz=1G, hugepagesz=1G, hugepages=8}: Huge pages are large contiguous areas of memory that can be used by applications and the kernel, instead of the traditional 4KB small pages.
@@
 This pre-allocation makes sure that these large memory regions are available to be used by the kernel or applications, without having to dynamically allocate and potentially fail.
-	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit), which connects a Direct Memory Access (DMA)-capable I/O bus, such as graphics cards and network adapters, to the main memory.
+	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit), which connects a DMA (Direct Memory Access)-capable I/O bus, such as graphics cards and network adapters, to the main memory.
 It can enhance device performance by allowing these devices to directly access and use memory, which is especially helpful when these devices are virtualized.
-	\item \textbf{rdt=l3cat}: Activates the L3 Cache Allocation Technology (L3 CAT) feature of Intel’s Resource Director Technology (RDT).
+	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel’s RDT (Resource Director Technology).
 Unlike L1 and L2 caches, where each core has its fixed capacity, L3 cache is a shared pool among multiple cores.
@@
 By controlling cache allocation, it can prevent a single process from monopolizing the L3 cache, which is particularly beneficial in virtualized environments, where multiple virtual machines share the same physical host.
-\begin{comment}
-\item \textbf{nmi\_watchdog=0}: Disables the NMI watchdog, which can free up a bit of system resources.
-\end{comment}
-\item \textbf{idle=poll}: Changes the CPU’s idle loop behavior to active polling. Instead of entering a low-power state when idle, the CPU continuously polls for new 
-tasks. This can reduce task start latency in real-time systems, but it increases power consumption.
-\begin{comment}
-\item \textbf{clocksource=tsc, tsc=reliable}: Sets the clocksource to TSC (Time Stamp Counter) and marks it as reliable.
-\end{comment}
+	\item \textbf{nmi\_watchdog=0}: Disables the NMI (Non-Maskable Interrupt) watchdog, which is a debugging feature of the Linux kernel.
+It works by periodically generating non-maskable interrupts.
+If the system does not respond to these interrupts within a certain timeframe, the NMI watchdog concludes that the system has hung and generates a system dump for debugging.
+This constant monitoring consumes CPU cycles and can introduce undesirable latency in real-time systems.
+	\item \textbf{idle=poll}: Changes the CPU’s idle loop behavior to active polling.
+Instead of entering a low-power state when idle, the CPU continuously polls for new tasks.
+This can reduce task start latency in real-time systems, but it increases power consumption.
+	\item \textbf{clocksource=tsc, tsc=reliable}: TSC (Time Stamp Counter) is a high-resolution timer provided by most x86 processors that counts the number of CPU cycles since it was last reset.
+Accurate timekeeping is crucial, particularly for real-time systems.
+These parameters set the clocksource to TSC and mark it as a reliable source of timekeeping, meaning it increments at a consistent rate and does not stop when the processor is idle.
 	\item \textbf{audit=0}: Disables the Linux audit system.
@@
 If there are a large number of such events, the audit system can consume significant CPU time and I/O bandwidth which could lead to higher latency.
-\begin{comment}
-\item \textbf{skew\_tick=1}: Enables a mode to reduce timer interrupt overhead.
-\item \textbf{intel\_pstate=disable}: Disables the intel\_pstate driver.
-\item \textbf{intel.max\_cstate=0, intel\_idle.max\_cstate=0, processor.max\_cstate=0, processor\_idle.max\_cstate=0}: Disables deeper C-states, which can help reduce latency.
-\end{comment}
+	\item \textbf{skew\_tick=1}: Enables a mode in the Linux kernel that reduces timer interrupt overhead.
+Normally, timer interrupts happen simultaneously on all CPUs, resulting in all CPUs to exit their low-power states at once.
+This can lead to increased contention for system resources.
+When enabled, the kernel offsets the timer interrupts on different CPUs, spreading them out over time.
+	\item \textbf{intel\_pstate=disable}: Disables the Intel P-state driver, which is a part of the Linux kernel that handles power management for Intel CPUs.
+It controls the frequency of the CPU by scaling it up when demand is high and scaling it down to save power when demand is low.
+This dynamic frequency scaling is disabled because it leads to increased latencies for real-time systems.
+	\item \textbf{intel.max\_cstate=0, intel\_idle.max\_cstate=0, processor.max\_cstate=0, processor\_idle.max\_cstate=0}: These parameters disable deeper C-states (CPU power saving states).
+Normally, when a CPU is idle, it can enter various C-states, with higher-numbered states representing deeper sleep states that save more power but take longer to wake up from.
+In real-time systems, these wake-up delays can be problematic.
+Disabling them leads keeps the CPUs ready to respond quickly to new tasks and helps in reducing latency.
 	\item \textbf{nosoftlockup}: Disables the soft lockup detector in the Linux kernel.
@@
 \begin{comment}
-\item \textbf{nohz=on}: Enables tickless operation system-wide.
 \item \textbf{no\_timer\_check}: Disables the check for broken timer interrupt sources.
@@
 Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs.
-
+Section \ref{sec:irq_handling} dives deeper into Interrupt Request affinity.
 \end{itemize}
+
+\clearpage
 

### 2024-06-07 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Kernel Configurations finished

@@
 This is detected and reported by the soft lockup detector, hence it is disabled to prevent these unnecessary warnings.
-\begin{comment}
 	\item \textbf{no\_timer\_check}: Disables the check for broken timer interrupt sources.
-\item \textbf{nospectre\_v2, spectre\_v2\_user=off}: Disables mitigations for the Spectre v2 vulnerability.
+Broken timer interrupt sources are problems with hardware or software that prevent timer interrupts from working as intended.
+Such a timer may not generate interrupts at the expected rate or at all.
+The kernel skips the checks for these broken timer interrupt sources, which can cause unnecessary overhead in a real-time system where every CPU cycle counts.
+	\item \textbf{nospectre\_v2, spectre\_v2\_user=off}: These parameters disable mitigations for the Spectre v2 vulnerability.
+Spectre v2 is a hardware vulnerability that affects many modern microprocessors and can allow malicious programs to access sensitive data they are not supposed to.
+While this is necessary for security, it has an impact on performance and should be turned off in controlled environments where the risk of exploitation is low.
 	\item \textbf{kvm.kvmclock\_periodic\_sync=N, kvm\_intel.ple\_gap=0}: These are KVM (Kernel-based Virtual Machine) related parameters.
-\end{comment}
+They disable the periodic synchronization of the kvmclock and set the gap between PLE (Pause Loop Exiting) events to 0.
+The kvmclock is a paravirtualized clock source provided by KVM to its guest OS and disabling this reduces latency introduced by clock synchronization.
+By setting the gap to 0, the virtual machine exits the pause loop immediately, which can reduce latency in spinlock-intensive workloads.
 	\item \textbf{irqaffinity=0}: Sets the default Interrupt Request affinity to none.
@@
 
-\clearpage
-
 \section{Host OS Configurations}
@@
 
-For CPU isolation, the \texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
+\noindent For CPU isolation, the \texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
 This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks or processes.

### 2024-06-07 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Host Configurations 5/13 finished

@@
 
-\noindent Figure~\ref{fig:lstopo} is the output of the \texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
+\noindent Figure~\ref{fig:lstopo} is the output of the ~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
 \begin{figure}[H]
@@
 In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
-The output of the command \texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
+The output of the command ~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
 
@@
 \section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
-The output of the command \texttt{uname -a} for Salamander 4 bare metal is shown in code~\ref{output:uname_a_hw}.
+The output of the command ~\texttt{uname -a} for Salamander 4 bare metal is shown in code~\ref{output:uname_a_hw}.
 
@@
 \noindent As a reference point, the latency program was executed on Salamander 4 bare metal for a duration of 10 minutes.
-The complete command used was \texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file \texttt{gnuplot.txt}.
+The complete command used was ~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file ~\texttt{gnuplot.txt}.
 Figure~\ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.
@@
 	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process.
-\texttt{OVMF} stands for Open Virtual Machine Firmware, \texttt{qcow2} is a format for disk image files used by QEMU, it stands for "QEMU Copy On Write version 2".
+~\texttt{OVMF} stands for Open Virtual Machine Firmware, ~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for "QEMU Copy On Write version 2".
 	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code~\ref{script:qemu_def}.
 	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform.
-It uses the \texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
+It uses the ~\texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
 	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Compressed tarball containing a pre-configured environment for the Salamander 4 OS.
-It is unpacked and sets up the \texttt{drive-c/} directory with system and log files in the \texttt{qemu\_def.sh} script.
+It is unpacked and sets up the ~\texttt{drive-c/} directory with system and log files in the ~\texttt{qemu\_def.sh} script.
 	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging.
@@
 \noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in chapter~\ref{cha:methodology}.
-The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the \texttt{uname -a} command.
+The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the ~\texttt{uname -a} command.
 
@@
 
-\section{BIOS Configurations}
+\section{BIOS Configurations}\label{sec:bios_configurations}
 
@@
 
-\section{Kernel Configurations}
+\section{Kernel Configurations}\label{sec:kernel_configurations}
 
@@
 	\item \textbf{isolcpus=4}: Isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned.
-CPU isolation is explained in detail in section \ref{sec:cpu_isolation}.
+CPU isolation is explained in detail in section \ref{subsec:cpu_isolation}.
 	\item \textbf{rcu\_nocbs=4}: The Linux kernel uses a synchronization mechanism called RCU, or Read-Copy-Update.
@@
 Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs.
-Section \ref{sec:irq_handling} dives deeper into Interrupt Request affinity.
+Section \ref{subsec:irq_handling} dives deeper into Interrupt Request affinity.
 \end{itemize}
 
-\section{Host OS Configurations}
-\subsection{CPU isolation}\label{sec:cpu_isolation}
+\section{Host OS Configurations}\label{sec:host_configurations}
+\subsection{CPU isolation}\label{subsec:cpu_isolation}
 
 Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section~\ref{sec:irq_handling}.
-
-\noindent For CPU isolation, the \texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
+Also, modifying the ~\texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section~\ref{subsec:irq_handling}.
+
+\noindent For CPU isolation, the ~\texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
 This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks or processes.
@@
 After the isolation, user tasks other than the QEMU process have been removed from running on this CPU.
-Only few critical kernel threads that are tied to this CPU still take CPU time.
+Only few per-CPU kernel threads that are tied to this CPU still take CPU time.
 
@@
 		4      41      41 [ksoftirqd/4]               TS    0      -
-		4      43      43 [kworker/4:0H-events_highpr TS  -20      -
-		4     505     505 [irq/189-iwlwifi:queue_1]   FF    -     50
-		4     519     519 [irq/203-iwlwifi:exception] FF    -     50
-		4    3008    3008 [kworker/4:1H-kblockd]      TS  -20      -
-		4  177779  177779 [kworker/4:2-events]        TS    0      -
-		4  448807  448807 [kworker/4:1-events]        TS    0      -		
+		4      42      42 [kworker/4:0-events]        TS    0      -
+		4      43      43 [kworker/4:0H-kblockd]      TS  -20      -
+		4     153     153 [kworker/4:1-events]        TS    0      -
+		4   81649   81649 qemu-system-x86_64 -M pc,ac TS    0      -
+		4   81649   81654 qemu-system-x86_64 -M pc,ac TS    0      -
+		4   81649   81676 qemu-system-x86_64 -M pc,ac TS    0      -
+		4   81649   81702 qemu-system-x86_64 -M pc,ac TS    0      -
+		4   81649   82185 qemu-system-x86_64 -M pc,ac TS    0      -
+		4   81649   82187 qemu-system-x86_64 -M pc,ac TS    0      -
+		4   82134   82134 [kworker/4:1H-kblockd]      TS  -20      -
 \end{lstlisting}
 \end{minipage}
-\subsection{Interrupt Requests Handling}\label{sec:irq_handling}
-Once the CPUs were isolated, Interrupt Requests handling was the next step.
-Interrupt Requests are used to send a signal to the CPU, prompting it to `interrupt' its current task and divert its attention to another task.
-This allows hardware devices to communicate with the CPU through frequent context switches, which can lead to performance degradation, especially in high-performance computing or real-time scenarios.
-To mitigate this, the IRQs needed to be removed from the isolated CPU.
-This was done by manipulating a file in the \texttt{proc} filesystem, namely \texttt{/proc/irq/<IRQ>/smp\_affinity}.
-The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format.
-Each bit in this mask corresponds to a CPU in the system.
-The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the significance increases towards the left until CPU19.
-In a system with 14 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be 3FFF.
-The script in code~\ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.
+\subsection{Interrupt Affinity}\label{subsec:irq_handling}
+Once the CPUs were isolated, interrupt requests handling was the next step.
+The purpose of interrupt requests is to inform the CPU to stop working on a certain job and start working on another.
+This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems.
+The ~\texttt{/proc/interrupts} file can be monitored using the command ~\texttt{watch -d -c cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
+The IRQs needed to be removed from the isolated CPU by manipulating the ~\texttt{/proc/irq/<IRQ>/smp\_affinity} files.
+The value in the ~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU.
+The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU13).
+When there are 14 CPUs available, the default value for ~\texttt{smp\_affinity} would be 3FFF.
+Removing CPU4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111).
+The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU.
+By changing the values of the ~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled.
+However, as the proc filesystem resets to its default state after each reboot, manually changing numerous IRQ files is tedious and time-consuming.
+This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot.
 
@@
 \begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={Check distribution of Interrupt Requests across each CPU},label={script:smp_affinity}]
+	\begin{lstlisting}[name={Check distribution of interrupt requests across each CPU},label={script:smp_affinity}]
+		import os
+		import pandas as pd
+		from tabulate import tabulate
+
+		# Get the number of CPUs
+		num_cpus = os.cpu_count()
+
+		# Initialize a dictionary to store the CPUs for each IRQ
+		irqs = {}
+
+		# Iterate over each IRQ
+		for irq in os.listdir('/proc/irq'):
+			# Check if the smp_affinity file exists for this IRQ
+			if os.path.isfile(f'/proc/irq/{irq}/smp_affinity'):
+				# Read the current smp_affinity
+				with open(f'/proc/irq/{irq}/smp_affinity', 'r') as f:
+					affinity = int(f.read().strip(), 16)
+				# Initialize an empty list to store the CPUs for this IRQ
+				cpus = []
+				# Iterate over each CPU
+				for cpu in range(num_cpus):
+					# Check if the bit for the current CPU is set
+					if ((affinity & (1 << cpu)) != 0):
+						# Add the CPU to the list for this IRQ
+						cpus.append(cpu)
+				# Sort the list of CPUs
+				cpus.sort()
+				# Add the list of CPUs to the dictionary for this IRQ
+				irqs[irq] = cpus
+
+		# Create a DataFrame to store the table
+		df = pd.DataFrame(index=sorted(irqs.keys(), key=int), columns=range(num_cpus))
+
+		# Fill the DataFrame with 'x' where a CPU is assigned to an IRQ
+		for irq, cpus in irqs.items():
+			for cpu in cpus:
+				df.loc[irq, cpu] = 'x'
+
+		# Replace NaN values with empty strings
+		df.fillna('', inplace=True)
+
+		# Print the table in pipe format
+		print(tabulate(df, headers='keys', tablefmt='pipe', showindex=True))
+
+		# Convert the DataFrame to a markdown table
+		markdown_table = df.to_markdown()
+
+		# Write the markdown table to a file
+		with open('table_CPU_IRQ.md', 'w') as f:
+			f.write(markdown_table)
+\end{lstlisting}
+\end{minipage}
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={Change IRQ assignment of a CPU},label={script:change_smp_affinity}]
 		#!/bin/bash
+
 		# Check if a command-line argument is provided
@@
 		fi
+
 		# Get the CPU number from the command-line argument
 		CPU=$1
-		# Initialize an empty array to store the IRQ numbers
-		IRQs=()
-		for IRQ in /proc/irq/*; do
-			if [ -f "$IRQ/smp_affinity" ]; then
-				# Read the current smp_affinity
-				AFFINITY=$(cat "$IRQ/smp_affinity")
-				# Check if the bit for the current CPU is set
-				if (( (0x$AFFINITY & (1 << CPU)) != 0 )); then
-					# Add the IRQ number to the array
-					IRQs+=("${IRQ#/proc/irq/}")
-				fi
+
+		# Define the mask values
+		declare -A mask_values
+		mask_values=( [0]="3ffe" [1]="3ffd" [2]="3ffb" [3]="3ff7" [4]="3fef" [5]="3fdf" [6]="3fbf" [7]="3f7f" [8]="3eff" [9]="3dff" [10]="3bff" [11]="37ff" [12]="2fff" [13]="17ff")
+
+		# Run the check_smp_affinity.sh script and get the IRQs
+		IRQs=$(./check_smp_affinity.sh $CPU | grep -o '[0-9]\+')
+
+		# Initialize an empty array to store the IRQs that could not be removed
+		failed_IRQs=()
+		# Initialize an empty array to store the IRQs that were successfully removed
+		succeeded_IRQs=()
+
+		# Loop over the IRQs
+		for IRQ in $IRQs; do
+			# Try to change the smp_affinity
+			echo ${mask_values[$CPU]} | sudo tee /proc/irq/$IRQ/smp_affinity > /dev/null 2>&1
+			# If the command failed, add the IRQ to the failed_IRQs array
+			if [ $? -ne 0 ]; then
+				failed_IRQs+=($IRQ)
+			else
+				succeeded_IRQs+=($IRQ)
 			fi
 		done
-		# Sort the array
-		IFS=$'\n' sorted=($(sort -n <<<"${IRQs[*]}"))
-		# Print the CPU number
-		echo "CPU $CPU IRQ affinity:"
-		# Print the sorted IRQ numbers on separate lines
-		for irq in "${sorted[@]}"; do
-			echo "$irq"
-		done
-
+
+		# Check if there were any failed IRQs
+		if [ ${#failed_IRQs[@]} -ne 0 ]; then
+			echo "IRQs ${failed_IRQs[@]} could not be removed from CPU $CPU."
+		fi
+
+		# Check if there were any successful IRQs
+		if [ ${#succeeded_IRQs[@]} -ne 0 ]; then
+			# Remove the first entry from the succeeded_IRQs array
+			succeeded_IRQs=("${succeeded_IRQs[@]:1}")
+			echo "IRQs ${succeeded_IRQs[@]} were removed from CPU $CPU."
+		fi		
 \end{lstlisting}
 \end{minipage}
-
-\noindent By changing the values of the \texttt{smp\_affinity} files of the respective IRQs, the assignment of IRQs was controlled so that they would not be handled by the isolated CPU.
-This reduced the interruptions caused by IRQs and the isolated CPUs were able to focus more on their assigned tasks.
-\subsection{Disable dynamic frequency scaling}
 \subsection{Disable RT throttling}
-\subsection{No unexpected RT processes are running on your system}
-\subsection{IRQ affinity}
-\subsection{RCU CPU offloading}
+If a real-time task consumes 100\% of the CPU time, the system may become unresponsive as a whole.
+This happens because an RT process is constantly using the CPU and the Linux scheduler will not schedule other non-RT processes in the meantime.
+To prevent complete system lockups, the kernel has a function to throttle RT processes if they consume 0.95 seconds out of every 1 second of CPU time.
+It does this by pausing the process for the remaining 0.05 seconds, which is not desired because this could result in missed deadlines.
+RT throttling can be disabled by  writing the value -1 to the ~\texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file.
+This change also needs to be made permanent because the proc filesystem resets to its default state after each reboot.
+For this purpose, the line ~\texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the ~\texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters.
+This reduces the potential for missed deadlines.
+\subsection{No unexpected RT processes}
+It is important that there are no other unexpected real time processes running on the system.
+A high priority RT process could occasionally be created by the host operating system at boot, hence disabling the offending RT process is advised if it is not needed.
+The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
 \subsection{Suppress rcu cpu stall}
+As already mentioned in section \ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
+A stall occurs when an RCU read-side critical section runs for an extended period of time, preventing the RCU grace period from completing.
+By writing 1 to the \texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file, these warnings are suppressed.
+This helps eliminate potential interruptions and CPU load that are caused by these warnings, thereby reduce the potential for increased latency.
+This setting is also not persistent across reboots, so the command needs to be added to a startup script.
 \subsection{Maybe?}
+
 \subsection{Start QEMU normally and give all QEMU threads rt-priority}
-\subsection{Kill all running user processes}
+
+\subsection{Cache Isolation for CPU and GPU}
+
+\subsection{Set CPU Affinity of IRQ thread to CPU 0}
+
+\subsection{Set Device Driver Work Queue to CPU 0}
+
 \subsection{Set CPU Affinity for systemd services}
-\subsection{Cache Isolation for CPU and GPU}
-\subsection{Set CPU Affinity of IRQ thread to CPU 0}
-\subsection{Set Device Driver Work Queue to CPU 0}
+
+\subsection{Stop Certain Services}
+
 \subsection{Disable Machine Check}
-\subsection{Stop Certain Services}
-\section{QEMU-KVM Configurations}
+
+There are other configurations that may be relevant depending on your use case, some of which are documented in [this talk](https://www.youtube.com/watch?v=NrjXEaTSyrw) and [this other talk](https://www.youtube.com/watch?v=w3yT8zJe0Uw).
+Additionally, quality-of-life configurations, such the variables in /etc/security/limits.conf, may need to be tuned as well.
+
+\clearpage
+
+\section{QEMU-KVM Configurations}\label{sec:guest_configurations}
 \subsection{Tune lapic timer advance}
@@
 This led to the investigation of the causes for the observed high and inconsistent latency.
-The guest operates within the kvm\_entry and \texttt{kvm\_exit} events of the host.
-Kernelshark revealed a high frequency of \texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
+The guest operates within the kvm\_entry and ~\texttt{kvm\_exit} events of the host.
+Kernelshark revealed a high frequency of ~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
 This frequent switching hinders the guest\'s ability to run continuously, thereby increasing the virtualization latency.
 To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
-Specifically, the causes for \texttt{kvm\_exit} events were analyzed.
+Specifically, the causes for ~\texttt{kvm\_exit} events were analyzed.
 The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
 The results in Figure~\ref{fig:kvm_exit_taskset} were obtained.
-Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed \texttt{kvm\_exit} events.
+Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed ~\texttt{kvm\_exit} events.
 
@@
 \clearpage
-Figure~\ref{fig:kvm_exit_taskset} shows \texttt{kvm\_exit} frequency with CPU islation.
+Figure~\ref{fig:kvm_exit_taskset} shows ~\texttt{kvm\_exit} frequency with CPU islation.
 \begin{figure}[H]
@@
 
-Figure~\ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
+Figure~\ref{fig:kvm_exits_default} shows ~\texttt{kvm\_exit} frequency without CPU islation.
 \begin{figure}[H]
@@
 	\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
-This explains why there are fewer \texttt{kvm\_exit} events.
-However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a \texttt{kvm\_exit} and kvm\_entry.
-
-	\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in \texttt{kvm\_exit} events.
+This explains why there are fewer ~\texttt{kvm\_exit} events.
+However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a ~\texttt{kvm\_exit} and kvm\_entry.
+
+	\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in ~\texttt{kvm\_exit} events.
 This is because every context switch takes place at hypervisor level.
@@
 
-	\noindent In the process of analyzing the \texttt{kvm\_exit} events, several reasons for these exits were identified.
-The most frequent among these were the \texttt{APIC\_WRITE} and \texttt{HLT} events.
+	\noindent In the process of analyzing the ~\texttt{kvm\_exit} events, several reasons for these exits were identified.
+The most frequent among these were the ~\texttt{APIC\_WRITE} and ~\texttt{HLT} events.
 The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
 The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
-Other significant but less frequent events included \texttt{EXTERNAL\_INTERRUPT} and \texttt{IO\_INSTRUCTION}.
+Other significant but less frequent events included ~\texttt{EXTERNAL\_INTERRUPT} and ~\texttt{IO\_INSTRUCTION}.
 These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
-Events such as \texttt{EPT\_MISCONFIG} and \texttt{PREEMPTION\_TIMER} were also noted.
+Events such as ~\texttt{EPT\_MISCONFIG} and ~\texttt{PREEMPTION\_TIMER} were also noted.
 These could potentially signal issues with memory management and the host's scheduling of the guest.
-While events like \texttt{PAUSE\_INSTRUCTION}, \texttt{EPT\_VIOLATION}, \texttt{EOI\_INDUCED}, \texttt{MSR\_READ}, and \texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
+While events like ~\texttt{PAUSE\_INSTRUCTION}, ~\texttt{EPT\_VIOLATION}, ~\texttt{EOI\_INDUCED}, ~\texttt{MSR\_READ}, and ~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
 
@@
 
-Figure~\ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualization.
+Figure~\ref{fig:kvm_exit_vapic} shows ~\texttt{kvm\_exit} frequency with APIC virtualization.
 Comparing this to the previous Figures~\ref{fig:kvm_exit_taskset} and~\ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.

### 2024-06-10 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Host Configurations 5/12 finished

@@
 \maketitle
-
 %
@@
 	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process.
-~\texttt{OVMF} stands for Open Virtual Machine Firmware, ~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for "QEMU Copy On Write version 2".
+~\texttt{OVMF} stands for Open Virtual Machine Firmware, ~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for ``QEMU Copy On Write version 2''.
 	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code~\ref{script:qemu_def}.
@@
 CPU 4 remains dedicated to high-priority tasks which helps in reducing latency.
-	\item \textbf{nohz\_full=4}: Makes CPU core 4 "tickless", meaning the kernel tries to avoid sending periodic scheduling-clock interrupts to the CPU when there are no runnable tasks.
+	\item \textbf{nohz\_full=4}: Makes CPU core 4 ``tickless'', meaning the kernel tries to avoid sending periodic scheduling-clock interrupts to the CPU when there are no runnable tasks.
 This lowers latency by reducing unnecessary wake-ups but may increase power consumption because the CPU is not able to enter a low-power state when idle.
@@
 The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
-\subsection{Suppress rcu cpu stall}
+\subsection{Disable RCU CPU stall warnings}
 As already mentioned in section \ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
-A stall occurs when an RCU read-side critical section runs for an extended period of time, preventing the RCU grace period from completing.
-By writing 1 to the \texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file, these warnings are suppressed.
-This helps eliminate potential interruptions and CPU load that are caused by these warnings, thereby reduce the potential for increased latency.
+An RCU CPU stall in the Linux kernel can occur due to several reasons.
+These include a CPU looping in an RCU read-side critical section, a CPU looping with interrupts disabled, a CPU looping with preemption disabled, or a CPU not getting around to less urgent tasks, known as “bottom halves”.
+The number of seconds the kernel should wait before checking for stalled CPUs and reporting a stall warning can be set via the  ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_timeout} file.
+These warnings can be suppressed alltogether to reduce the potential for increased latency by writing ``1'' to the ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file.
 This setting is also not persistent across reboots, so the command needs to be added to a startup script.
-\subsection{Maybe?}
-
-\subsection{Start QEMU normally and give all QEMU threads rt-priority}
+\subsection{RT-priority}
 
@@
 Additionally, quality-of-life configurations, such the variables in /etc/security/limits.conf, may need to be tuned as well.
+
+``korrekt''
+
+\enquote{This LaTeX.}
 
@@
 \clearpage
-
 % Das Abbildungsverzeichnis

### 2024-06-10 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Host Configurations 5/10 finished

@@
 Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-Also, modifying the ~\texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section~\ref{subsec:irq_handling}.
-
-\noindent For CPU isolation, the ~\texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
-This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks or processes.
+For CPU isolation, the ~\texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
+This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks.
 However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
-Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}.
+Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}, including systemd services.
+To prevent systemd services from running on an isolated CPU, the CPU affinity can be set the with line \texttt{CPUAffinity=0 1 2 3 5 6 7 8 9 10 11 12 13} in the \texttt{/etc/systemd/system.conf} file to indicate the CPUs that systemd-services are allowed to run on.
+Every CPU other than the isolated CPU 4 is allowed.
 Output~\ref{output:pse} shows the user and kernel tasks that run on CPU 4.
@@
 When there are 14 CPUs available, the default value for ~\texttt{smp\_affinity} would be 3FFF.
-Removing CPU4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111).
+Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111).
 The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU.
@@
 This reduces the potential for missed deadlines.
-\subsection{No unexpected RT processes}
-It is important that there are no other unexpected real time processes running on the system.
-A high priority RT process could occasionally be created by the host operating system at boot, hence disabling the offending RT process is advised if it is not needed.
-The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
 \subsection{Disable RCU CPU stall warnings}
@@
 \subsection{RT-priority} 
-
-\subsection{Cache Isolation for CPU and GPU}
-
+\subsection{No unexpected RT processes}
+It is important that there are no other unexpected real time processes running on the system.
+A high priority RT process could occasionally be created by the host operating system at boot, hence disabling the offending RT process is advised if it is not needed.
+The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
 \subsection{Set CPU Affinity of IRQ thread to CPU 0}
@@
 \subsection{Set Device Driver Work Queue to CPU 0}
-
-\subsection{Set CPU Affinity for systemd services}
 

### 2024-06-11 Masterarbeit_Robotik_Pamuk.tex
cpuaffinity.sh

@@
 \section{Host OS Configurations}\label{sec:host_configurations}
-\subsection{CPU isolation}\label{subsec:cpu_isolation}
+\subsection{CPU affinity and isolation}\label{subsec:cpu_isolation}
 
@@
 
+Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads running on CPU 4.
+This can be done through finding out the respective process IDs through \texttt{ps -e | grep 'irq/.*nvme'} and then executing \texttt{sudo taskset -a -p -c 0 <PID>} with the PID obtained from the previous command.
+
 \vspace{1em}

### 2024-06-11 masterthesis-documentation/docs/sigmatek/xenomai/print_numbers.py

@@ -1,7 +1,7 @@
 # Open the file in write mode
 with open("print_numbers.txt", "w") as file:
     # Write numbers from 1 to 600
-    for number in range(1, 95661):
+    for number in range(1, 50943):
         file.write(str(number) + "\n")
 
 # Print a success message

### 2024-06-11 masterthesis-documentation/docs/sigmatek/salamander4/latency_reduction/IRQaffinity/cpuaffinity.sh
cpuaffinity.sh

@@ -1,0 +1,37 @@
+#!/bin/bash
+
+# Replace 'process_name1' and 'process_name2' with the names of your processes
+process_name1="irq/157-nvme0q5"
+process_name2="irq/172-nvme1q5"
+
+# Initialize a variable to track success status
+success=true
+
+# Find the PID of the first process
+pid1=$(pgrep -f $process_name1)
+
+# Check if the PID was found
+if [ -n "$pid1" ]; then
+    # Set the CPU affinity of the process to CPU 0
+    sudo taskset -p -c 0 $pid1 > /dev/null
+else
+    echo "Process $process_name1 not found."
+    success=false
+fi
+
+# Find the PID of the second process
+pid2=$(pgrep -f $process_name2)
+
+# Check if the PID was found
+if [ -n "$pid2" ]; then
+    # Set the CPU affinity of the process to CPU 0
+    sudo taskset -p -c 0 $pid2 > /dev/null
+else
+    echo "Process $process_name2 not found."
+    success=false
+fi
+
+# Print a success message if no errors occurred
+if $success; then
+    echo "CPU affinity successfully changed."
+fi

### 2024-06-11 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] Host Configurations finished

--- 2024-06-10 Masterarbeit_Robotik_Pamuk.tex
@@
 
-\begin{table}[h]
+\begin{table}[H]
 	\centering
@@
 	\label{tab:latency_statistics_bare_metal}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
 	\begin{tabular}{|c|c|}\hline
@@
 	Standard Deviation & 0.85 \\\hline
-	\end{tabular}
+	\end{tabular}}
 	\end{table}
@@
 	\label{tab:latency_statistics_virt}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
 	\begin{tabular}{|c|c|}\hline
@@
 	Standard Deviation & 52.13 \\\hline
-	\end{tabular}
+	\end{tabular}}
 	\end{table}
@@
 This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot.
+Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads.
+This can be done through finding out the respective process IDs through \texttt{ps -e | grep 'irq/.*nvme'} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.
 
@@
 \subsection{RT-priority} 
+Having a real-time kernel itself is a crucial part of achieving deterministic behavior, but not enough to take full advantage of the real-time capabilities.
+One key aspect of this are real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}.
+In essence, Table \ref{tab:scheduling_priorities} lists minimum and maximum priorities for different scheduling policies in Linux.
+
+\begin{table}[H]
+	\centering
+	\caption{Minimum and maximum priorities for different scheduling policies}
+	\label{tab:scheduling_priorities}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|l|l|l|}
+	\hline
+	\textbf{Scheduling Policy} & \textbf{Min Priority} & \textbf{Max Priority} \\ \hline
+	SCHED\_OTHER               & 0                     & 0                     \\ \hline
+	SCHED\_FIFO                & 1                     & 99                    \\ \hline
+	SCHED\_RR                  & 1                     & 99                    \\ \hline
+	SCHED\_BATCH               & 0                     & 0                     \\ \hline
+	SCHED\_IDLE                & 0                     & 0                     \\ \hline
+	SCHED\_DEADLINE            & 0                     & 0                     \\ \hline
+	\end{tabular}}
+	\end{table}
+
+\noindent \texttt{SCHED\_FIFO} allows deterministic, high-priority execution of critical tasks without being preempted by lower-priority processes.
+The virtual machine can either be started with \texttt{chrt -f <PRIO>} or adjusted at a later point with \texttt{chrt -f <PRIO> <PID>}.
+
 \subsection{No unexpected RT processes}
@@
 The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
-\subsection{Set CPU Affinity of IRQ thread to CPU 0}
-
-\subsection{Set Device Driver Work Queue to CPU 0}
-
+\subsection{Set Device Driver Work Queue}
+The device driver work queue allows time-consuming tasks to be offloaded to be processed later in a separate kernel thread.
+By setting the work queue away from CPU 4 to another CPU, it is free to handle real-time tasks without being interrupted by these work queue tasks.
+This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
 \subsection{Stop Certain Services}
-
+Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{table:stop_servies} can be further sources for random latency and unnecessary overhead.
+Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means they will not be able to interrupt the CPU with their tasks.
+
+\begin{table}[H]
+	\centering
+	\caption{Description of Services}
+	\label{table:stop_servies}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|l|l|}
+	\hline
+	\textbf{Service} & \textbf{Description} \\
+	\hline
+	irqbalance.service & Distributes hardware interrupts across CPUs \\\hline
+	thermald.service & A daemon that prevents overheating \\\hline
+	wpa\_supplicant.service & A service for wireless network devices \\
+	\hline
+	\end{tabular}}
+	\end{table}
 \subsection{Disable Machine Check}
-
-There are other configurations that may be relevant depending on your use case, some of which are documented in [this talk](https://www.youtube.com/watch?v=NrjXEaTSyrw) and [this other talk](https://www.youtube.com/watch?v=w3yT8zJe0Uw).
-Additionally, quality-of-life configurations, such the variables in /etc/security/limits.conf, may need to be tuned as well.
-
-``korrekt''
-
-\enquote{This LaTeX.}
-
-\clearpage
+Machine checks report hardware errors and these checks can cause interruptions and increase latency.
+Hence it is best to disable them in real-time scenarios by writing a ``0'' to the \texttt{/sys/devices/system/machinecheck/machinecheck0/check\_interval} file.
 
@@
 \clearpage
+
 \section{Guest OS Configurations}
 
-\clearpage
+\section{Other configurations}
+There are other configurations that may be relevant depending on the case.
+Linux Kernel Developer Steven Rostedt explains all relevant aspects of a real-time system that must be considered in~\cite{kernelrecipesKernelRecipes20162016} and gives insight for finding sources of latency on the linux system in~\cite{thelinuxfoundationFindingSourcesLatency2020}.
+A Checklist for Writing Linux Real-Time Applications is provided by John Ogness in \cite{thelinuxfoundationChecklistWritingLinux2020}.
+Every layer of the system stack must be deterministic to ensure predictable and reliable latency, including hardware, operating system, middleware and drivers, and the application software.
+\cite{HOWTOBuildRTapplication} and \cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real time preemption patch in great detail.
+Various hardware and software tunings are mentioned in~\cite{KVMQemuVirtualization} and~\cite{RealTimePerformanceTuning2022}.
 
@@
 \chapter{To Include}\label{cha:to_include}
+``korrekt''
+
+\enquote{This LaTeX.}
+
 \section{Latency Comparison}
@@
 
-\begin{table}[h!]
+\begin{table}[H]
 	\centering

### 2024-06-11 Masterarbeit_Robotik_Pamuk.tex
[DRAFT] QEMU/KVM Configurations finished

@@
 Delays or errors can have catastrophic consequences in some cases.
-Traditional operating systems, such as Windows or Linux, are often not suitable for these types of real-time requirements as they cannot guarantee deterministic execution times.
+Traditional operating systems, such as Windows or Ubuntu, are often not suitable for these types of real-time requirements as they cannot guarantee deterministic execution times.
 Therefore, real-time operating systems are required that are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
@@
 The goal is to virtualize Salamander 4 and approach the performance of bare metal.
-Salamander 4 is built with Yocto and virtualized through QEMU/KVM.
+Salamander 4 is built with Yocto and virtualized through Quick Emulator/Kernel-based Virtual Machine (QEMU/KVM).
 The details of this operating system are explained in chapter~\ref{cha:salamander4}.
@@
 
-\bigskip \noindent Trace-cmd was used for tracing the Linux kernel.
+\bigskip \noindent Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
 It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
-Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency~\cite{Tracecmd}.
+Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
 
@@
 
-\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel.
+\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel~\cite{XenomaiXenomai}.
 It enables low-latency and deterministic execution of time-critical tasks.
@@
 A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
-The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times~\cite{XenomaiXenomai}.
+The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
 
@@
 \clearpage
-\noindent The interfaces between the individual modules are indicated by an arrow in Figure~\ref{fig:lasal_cpu}.
+\noindent These modules and the interfaces (indicated by an arrow) between them are shown in Figure~\ref{fig:lasal_cpu}.
 
@@
 	\includegraphics[width=0.6\columnwidth]{img/x3-cobalt-interfaces.png}
-	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces}
+	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
 	\label{fig:cobalt}
@@
 	\includegraphics[width=0.6\columnwidth]{img/x3-mercury-interfaces.png}
-	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces}
+	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
 	\label{fig:mercury}
@@
 \section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
-The output of the command ~\texttt{uname -a} for Salamander 4 bare metal is shown in code~\ref{output:uname_a_hw}.
+The output of the command ~\texttt{uname -a} for Salamander 4 bare metal is shown in Code~\ref{output:uname_a_hw}.
 
@@
 It was developed using Yocto, an open-source project that allows customised Linux distributions to be created for embedded systems~\cite{WelcomeYoctoProject}.
-Upon generating the necessary files, Yocto provides a QEMU folder with the following components shown in code~\ref{lst:ls}.
+Upon generating the necessary files, Yocto provides a QEMU folder with the following components shown in Code~\ref{lst:ls}.
 QEMU is the environment in which the virtualization runs, as it is an open-source tool for hardware virtualization~\cite{QEMU}.
@@
 
-\noindent With the help of the script depicted in code~\ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
+\noindent With the help of the script depicted in Code~\ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
 This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host.
-The following is a description of the components used for the virtualization of Salamander 4:
+The following is a description of the components used for the virtualization of Salamander 4.
+\clearpage
 \begin{itemize}
@@
 
+\clearpage
+
 \noindent Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time with default settings.
@@
 For this reason, an extensive tuning process is carried out.
-This involves configurations spanning the BIOS, kernel, host OS, QEMU-KVM virtualization layer, and the Salamander 4 OS itself.
+This involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
 The individual configurations will be discussed in detail and the modifications will be justified with a clear explanation.
@@
 
-The kernel command-line parameters are shown in Code~\ref{script:kernel_cli} below. Some of them are essential for real-time performance.
+The kernel command-line parameters are shown in Code~\ref{script:kernel_cli} below.
 
@@
 \section{Host OS Configurations}\label{sec:host_configurations}
+The host OS needs to provide an environment where the guest OS can operate in real-time.
+This entails a number of host-side adjustments to reduce interruptions and reduce latency.
+In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest OS is provided.
 \subsection{CPU affinity and isolation}\label{subsec:cpu_isolation}
@@
 \end{minipage}
+\subsection{RT-priority} 
+Having a real-time kernel itself is a crucial part of achieving deterministic behavior, but not enough to take full advantage of the real-time capabilities.
+One key aspect of this are real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}.
+In essence, Table \ref{tab:scheduling_priorities} lists minimum and maximum priorities for different scheduling policies in Linux.
+
+\begin{table}[H]
+	\centering
+	\caption{Minimum and maximum priorities for different scheduling policies}
+	\label{tab:scheduling_priorities}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|l|l|l|}
+	\hline
+	\textbf{Scheduling Policy} & \textbf{Min Priority} & \textbf{Max Priority} \\ \hline
+	SCHED\_OTHER               & 0                     & 0                     \\ \hline
+	SCHED\_FIFO                & 1                     & 99                    \\ \hline
+	SCHED\_RR                  & 1                     & 99                    \\ \hline
+	SCHED\_BATCH               & 0                     & 0                     \\ \hline
+	SCHED\_IDLE                & 0                     & 0                     \\ \hline
+	SCHED\_DEADLINE            & 0                     & 0                     \\ \hline
+	\end{tabular}}
+	\end{table}
+
+\noindent \texttt{SCHED\_FIFO} allows deterministic, high-priority execution of critical tasks without being preempted by lower-priority processes.
+The virtual machine can either be started with \texttt{chrt -f <PRIO>} or adjusted at a later point with \texttt{chrt -f <PRIO> <PID>}.
+It is also important that there are no other unexpected real time processes running on the system concurrently.
 \subsection{Disable RT throttling}
@@
 This reduces the potential for missed deadlines.
+\subsection{Set Device Driver Work Queue}
+The device driver work queue allows time-consuming tasks to be offloaded to be processed later in a separate kernel thread.
+By setting the work queue away from CPU 4 to another CPU, it is free to handle real-time tasks without being interrupted by these work queue tasks.
+This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
 \subsection{Disable RCU CPU stall warnings}
@@
 This setting is also not persistent across reboots, so the command needs to be added to a startup script.
-\subsection{RT-priority} 
-Having a real-time kernel itself is a crucial part of achieving deterministic behavior, but not enough to take full advantage of the real-time capabilities.
-One key aspect of this are real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}.
-In essence, Table \ref{tab:scheduling_priorities} lists minimum and maximum priorities for different scheduling policies in Linux.
+\subsection{Stop Certain Services}
+Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{tab:stop_servies} can be further sources for random latency and unnecessary overhead.
+Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means they will not be able to interrupt the CPU with their tasks.
 
@@
 	\centering
-	\caption{Minimum and maximum priorities for different scheduling policies}
-	\label{tab:scheduling_priorities}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-	\begin{tabular}{|l|l|l|}
-	\hline
-	\textbf{Scheduling Policy} & \textbf{Min Priority} & \textbf{Max Priority} \\ \hline
-	SCHED\_OTHER               & 0                     & 0                     \\ \hline
-	SCHED\_FIFO                & 1                     & 99                    \\ \hline
-	SCHED\_RR                  & 1                     & 99                    \\ \hline
-	SCHED\_BATCH               & 0                     & 0                     \\ \hline
-	SCHED\_IDLE                & 0                     & 0                     \\ \hline
-	SCHED\_DEADLINE            & 0                     & 0                     \\ \hline
-	\end{tabular}}
-	\end{table}
-
-\noindent \texttt{SCHED\_FIFO} allows deterministic, high-priority execution of critical tasks without being preempted by lower-priority processes.
-The virtual machine can either be started with \texttt{chrt -f <PRIO>} or adjusted at a later point with \texttt{chrt -f <PRIO> <PID>}.
-
-\subsection{No unexpected RT processes}
-It is important that there are no other unexpected real time processes running on the system.
-A high priority RT process could occasionally be created by the host operating system at boot, hence disabling the offending RT process is advised if it is not needed.
-The process of disabling these RT processes is usually described in the documentation for the chosen Linux distribution.
-\subsection{Set Device Driver Work Queue}
-The device driver work queue allows time-consuming tasks to be offloaded to be processed later in a separate kernel thread.
-By setting the work queue away from CPU 4 to another CPU, it is free to handle real-time tasks without being interrupted by these work queue tasks.
-This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
-\subsection{Stop Certain Services}
-Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{table:stop_servies} can be further sources for random latency and unnecessary overhead.
-Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means they will not be able to interrupt the CPU with their tasks.
-
-\begin{table}[H]
-	\centering
 	\caption{Description of Services}
-	\label{table:stop_servies}
+	\label{tab:stop_servies}
 	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
@@
 
-\section{QEMU-KVM Configurations}\label{sec:guest_configurations}
+\section{QEMU/KVM Configurations}\label{sec:guest_configurations}
+KVM allows the guest OS to run directly on the hardware, bypassing the need for traditional emulation which can introduce delays.
+QEMU, when used with KVM, provides hardware-assisted virtualization, which also lowers latency in the guest OS.
+
 \subsection{Tune lapic timer advance}
+The Local Advanced Programmable Interrupt Controller (LAPIC) is a built-in timer that handles the delivery of interrupts to the CPU.
+It generates interrupts at a rate.
+This rate can be tuned in the \texttt{/sys/module/kvm/parameters/lapic\_timer\_advance\_ns} file to reduce the frequency of interrupts and therefore decrease the latency of the guest VM.
+The default value is ``-1'', which means that the kernel will automatically calculate an appropriate advance for the timer.
+Here, it is set to the value ``7500''.
+Hence,the timer interrupt will be delivered 7500 nanoseconds earlier than it is actually due.
+This gives the VM more time to handle it.
+
 \subsection{Set QEMU options for real-time VM}
-\subsection{Set CPU affinity and scheduling policy of QEMU CPU threads}
-\subsection{Passthrough PCI devices into the VM}
-
-\clearpage
+QEMU provides several options that can be used to improve the real-time performance of the guest VM. Table \ref{tab:qemu_options} briefly explains these options.
+
+\begin{table}[H]
+    \centering
+    \caption{QEMU options for real-time performance}
+    \label{tab:qemu_options}
+    \setlength{\tabcolsep}{0.5em} % for the horizontal padding
+    {\renewcommand{\arraystretch}{1.2}% for the vertical padding
+    \begin{tabular}{|p{6.5cm}|p{8cm}|}
+    \hline
+    \textbf{QEMU Option} & \textbf{Description} \\\hline
+    \texttt{-object memory-backend-ram,}\newline\texttt{id=ram0,size=4G,prealloc=on} & Locks the memory of the VM to 4GB and prevents \newline it from being swapped out to disk \\\hline
+    \texttt{-mem-prealloc} \newline \texttt{-mem-path /dev/hugepages/} & Enables the use of hugepages and improves \newline memory access \\\hline
+    \end{tabular}}
+    \end{table}
+
+\noindent Code \ref{script:qemu_def_tuned} shows the final QEMU script used to start the Salamander 4 virtualization, including these options.
+
+	\vspace{1em}
+	\begin{minipage}{\linewidth}
+		\begin{lstlisting}[name={Tuned QEMU script for starting Salamander 4 virtualization},label={script:qemu_def_tuned}]
+		#!/bin/sh
+
+		if  [ ! -d drive-c/ ]; then
+				echo "Filling drive-c/"
+				mkdir drive-c/
+				tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
+		fi
+
+		exec taskset -c 4 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
+		-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
+		-append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable nohlt idle=poll quiet xeno_hal.smi=1 xenomai.smi=1 threadirqs" \
+		-net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
+		-fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
+		-device vhost-vsock-pci,guest-cid=3,id=vsock0 \
+		-drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
+		-cpu host,hv-passthrough \
+		-object memory-backend-ram,id=ram0,size=4G,prealloc=on \
+		-mem-prealloc -mem-path /dev/hugepages \
+		-no-reboot -nographic
+	\end{lstlisting}
+	\end{minipage}
 
@@
 
+\clearpage
+
+\chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
+
+\section{VARAN}
+\section{Robotic Application}
+\clearpage
+
+\begin{comment}
+
 \chapter{KVM exit reasons}\label{cha:kvm_exit_reasons}
-
 \section{APIC\_WRITE}
@@
 \section{MSR\_READ}
-
-\clearpage
-
-\chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
-
-\section{VARAN}
-\section{Robotic Application}
-\clearpage
-
-\begin{comment}
 
@@
 ~\cite{maPerformanceTuningKVMbased}
-~\cite{linPerformanceEvaluationXenomai}
+~\cite{LinuxProcessPriorities}
+~\cite{kernelrecipesKernelRecipes20162016}
+~\cite{thelinuxfoundationFindingSourcesLatency2020}
+~\cite{thelinuxfoundationChecklistWritingLinux2020}
+~\cite{HOWTOBuildRTapplication}
+~\cite{RealtimeProgrammingLinux}
+~\cite{KVMQemuVirtualization}
+~\cite{RealTimePerformanceTuning2022}
+~\cite{adamPerformanceAssessmentLinux2021}
+~\cite{adamRealTimePerformanceResponse2021}
 ~\cite{broskyShieldedProcessorsGuaranteeing2003}
 ~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
+~\cite{deoliveiraDemystifyingRealTimeLinux}
 ~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
@@
 ~\cite{kiszkaLinuxRealTimeHypervisor}
+~\cite{linPerformanceEvaluationXenomai}
 ~\cite{lutsykPipelinedMulticoreMachine2020}
@@
 ~\cite{yoonRealTimePerformanceAnalysis}
-~\cite{adamPerformanceAssessmentLinux2021}
-~\cite{adamRealTimePerformanceResponse2021}
-~\cite{deoliveiraDemystifyingRealTimeLinux}
-~\cite{RealTimePerformanceTuning2022}
 ~\clearpage

### 2024-06-12 Masterarbeit_Robotik_Pamuk.tex
Added to kernel and host config

@@
 	\begin{lstlisting}[name={Kernel Configuration},label={script:kernel_cli}]
-		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 nohz_full=4 nohz=on default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
+		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 rcu_nocb_poll nohz_full=4 nohz=on default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
 \end{lstlisting}
@@
 CPU 4 remains dedicated to high-priority tasks which helps in reducing latency.
+	\item \textbf{rcu\_nocb\_poll}: This is used together with \texttt{rcu\_nocbs} and causes the system to actively poll for RCU callbacks to invoke, instead of waiting for the next RCU grace period.
+This reduces latency.
 	\item \textbf{nohz\_full=4}: Makes CPU core 4 ``tickless'', meaning the kernel tries to avoid sending periodic scheduling-clock interrupts to the CPU when there are no runnable tasks.
@@
 This reduces the potential for missed deadlines.
+\subsection{Disable timer migration}
+Timer migration allows timers to be moved from one CPU to another, which means the kernel can balance load across multiple CPUs.
+In a real-time system, this can introduce latency and jitter.
+To disable timer migration, the value ``0'' needs to be written to the ~\texttt{/proc/sys/kernel/timer\_migration} file.
+This change also needs to be made permanent by writing the line ~\texttt{kernel.timer\_migration = 0} to the ~\texttt{/etc/sysctl.conf} file.
+This reduces the amount of context switches and interrupts.
 \subsection{Set Device Driver Work Queue}

### 2024-06-12 masterthesis-documentation/docs/sigmatek/xenomai/print_numbers.py

@@ -1,7 +1,7 @@
 # Open the file in write mode
 with open("print_numbers.txt", "w") as file:
     # Write numbers from 1 to 600
-    for number in range(1, 50943):
+    for number in range(1, 53608):
         file.write(str(number) + "\n")
 
 # Print a success message

### 2024-06-13 Masterarbeit_Robotik_Pamuk.tex
cite

@@
 ~\cite{kiszkaLinuxRealTimeHypervisor}
-~\cite{linPerformanceEvaluationXenomai}
+~\cite{huang2015performance}
 ~\cite{lutsykPipelinedMulticoreMachine2020}

### 2024-07-29 Masterarbeit_Robotik_Pamuk.tex
pdfs and imgs, robotic application

@@
 \chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
-
-\section{VARAN}
+This chapter compares the latency of the Salamander 4 operating system before and after the real-time performance tunings.
+Additionally, these results are contrasted with the latency of Salamander 4 running on bare metal hardware to understand how closely the performance of the virtualized system can match that of the bare metal system.
+The experimental setup includes a six-axis mini-robot, illustrated in figure \ref{fig:mini_robot}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\columnwidth]{img/experiment/mini_robot.jpg}
+	\caption[Mini-robot of the experiment]{Mini-robot of the experiment}
+	\label{fig:mini_robot}
+\end{figure}
+
+\noindent In order to move this robot, it needs to get signals  
+The program is written in Lasal Class 2 and will be applied to all three versions to measure the reaction time of the robot to the specific commands.
+
+\ref{fig:hardware_tree}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\columnwidth]{img/experiment/hardware_tree.png}
+	\caption[Hardware tree]{Hardware tree}
+	\label{fig:hardware_tree}
+\end{figure}
+
+\ref{fig:virt_tree}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\columnwidth]{img/experiment/virt_tree.png}
+	\caption[Virt tree]{Virt tree}
+	\label{fig:virt_tree}
+\end{figure}
+
+\begin{comment}
+	https://www.sigmatek-automation.com/en/products/accessories/controls-hmis/pcv-522/
+\end{comment}
+
 \section{Robotic Application}
+
 \clearpage

### 2024-07-30 Masterarbeit_Robotik_Pamuk.tex
Real-Time Robotic Application

--- 2024-06-13 Masterarbeit_Robotik_Pamuk.tex
@@
 	\centering
-	\includegraphics[width=0.6\columnwidth]{img/x3-cobalt-interfaces.png}
+	\includegraphics[width=0.5\columnwidth]{img/x3-cobalt-interfaces.png}
 	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
@@
 	\centering
-	\includegraphics[width=0.6\columnwidth]{img/x3-mercury-interfaces.png}
+	\includegraphics[width=0.5\columnwidth]{img/x3-mercury-interfaces.png}
 	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
@@
 Linux Kernel Developer Steven Rostedt explains all relevant aspects of a real-time system that must be considered in~\cite{kernelrecipesKernelRecipes20162016} and gives insight for finding sources of latency on the linux system in~\cite{thelinuxfoundationFindingSourcesLatency2020}.
-A Checklist for Writing Linux Real-Time Applications is provided by John Ogness in \cite{thelinuxfoundationChecklistWritingLinux2020}.
-Every layer of the system stack must be deterministic to ensure predictable and reliable latency, including hardware, operating system, middleware and drivers, and the application software.
-\cite{HOWTOBuildRTapplication} and \cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real time preemption patch in great detail.
+A Checklist for Writing Linux Real-Time Applications is provided by John Ogness in~\cite{thelinuxfoundationChecklistWritingLinux2020}.
+Every layer of the system stack must be deterministic to ensure predictable and reliable latency, including hardware, operating system, middleware and drivers, and the application software.~\cite{HOWTOBuildRTapplication} and~\cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real time preemption patch in great detail.
 Various hardware and software tunings are mentioned in~\cite{KVMQemuVirtualization} and~\cite{RealTimePerformanceTuning2022}.
@@
 \chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
-
-\section{VARAN}
+This chapter compares the latency of the Salamander 4 operating system before and after the real-time performance tunings.
+Additionally, these results are contrasted with the latency of Salamander 4 running on bare metal hardware to understand how closely the performance of the virtualization can match that of the bare metal.
+The experimental setup includes a six-axis mini-robot, illustrated in Figure \ref{fig:mini_robot}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\columnwidth]{img/experiment/mini_robot.jpg}
+	\caption[Mini-robot of the experiment]{Mini-robot of the experiment}
+	\label{fig:mini_robot}
+\end{figure}
+
+\noindent The drive system of the robot arm consists of six MG996R digital servo motors~\cite{MG996RServoMotor}, equipped with a metal gearbox.
+The motor is able to rotate in a range of approximately 180 degrees and its position can be controlled with a high degree of accuracy.
+Each servo motor has three wires that need to be connected as shown in Figure \ref{fig:motor_wires}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\columnwidth]{img/experiment/motor_wires.png}
+	\caption[MG996R Servo Motor]{MG996R Servo Motor~\cite{MG996RServoMotor}}
+	\label{fig:motor_wires}
+\end{figure}
+
+\noindent To drive the motor, it has to be powered using the red and brown wires and can be controlled by sending PWM signals to the orange wire.
+In this experiment, this PWM signal is generated by the proprietary PW 022 pulse width module of Sigmatek~\cite{DigitalOutputSIGMATEK}.
+Its connector layout is illustrated in Figure \ref{fig:pw022_connectors}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.5\columnwidth]{img/experiment/pw022_connectors.png}
+	\caption[PW 022 pulse width module connector layout]{PW 022 pulse width module connector layout~\cite{DigitalOutputSIGMATEK}}
+	\label{fig:pw022_connectors}
+\end{figure}
+
+\noindent The module has two +24 V switching PWM outputs with an adjustable frequency for controlling inductive loads.
+Since the mentioned servo motors operate between 4.8 volts and 7.2 volts~\cite{MG996RServoMotor}, this voltage needed to be reduced through resistors before supplying it to a servo motor.
+For this purpose, two resistors with resistances of 2500 kiloohms and 1000 kiloohms were connected in series.
+The voltage across each resistor is calculated using Ohm's law, which is given by equation \ref{eq:ohms_law} below.
+\begin{equation}
+	\label{eq:ohms_law}
+	V = I \cdot R
+\end{equation}	
+In this equation, \(V\) represents the voltage across the resistor, \(I\) is the current flowing through the resistor, and \(R\) is the resistance of the resistor.
+The resulting voltages across the resistors are as follows:
+\begin{itemize}
+\item The voltage across the 2.5 kiloohm resistor is approximately 17.14 volts.
+\item The voltage across the 1 kiloohm resistor is approximately 6.86 volts. This voltage was then supplied to the control wire of servo motor 1 of the mini-robot.
+\end{itemize}
+
+\noindent Connecting a second servo motor of the mini-robot means repeating the process of reducing the voltage through resistors for the second PWM output of the PW 022 module.
+
+\noindent The program was written in Lasal Class 2 and will be applied to all three mentioned versions of Salamander 4 to measure the reaction time of the robot to the specific commands.
+Prior to examining the software program, the next two subsections briefly explain the setup of each version of the experiment.
+
+\subsection{Setup of Hardware Salamander 4}
+
+In this version, Salamander 4 runs on the CP 841~\cite{CPUEinheitenSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
+The PW 022 module is directly mounted on the CPU via the S-DIAS bus~\cite{SDIASSIGMATEK} and communicates over the hard real-time capable Ethernet VARAN with 100 Mbit/s.
+The setup is visible on Figure \ref{fig:hardware_tree}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.5\columnwidth]{img/experiment/hardware_tree.png}
+	\caption[Hardware tree]{Hardware tree}
+	\label{fig:hardware_tree}
+\end{figure}
+
+\subsection{Setup of QEMU Salamander 4}
+To connect QEMU with the PWM module, the PCV 522 VARAN Manager PCI Insert Card~\cite{ControlsHMIsSIGMATEK} needed to be plugged into the working station PC.
+An additional Varan Connection module VI 021~\cite{InterfacesSplittersSIGMATEK} was required to connect the PCV 522 module to the PW 022 module that generates the signal to move the mini-robot.
+This setup is depicted in Figure \ref{fig:virt_tree}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.5\columnwidth]{img/experiment/virt_tree.png}
+	\caption[Virtualization tree]{Virtualization tree}
+	\label{fig:virt_tree}
+\end{figure}
+
 \section{Robotic Application}
+
 \clearpage
@@
 ~\cite{RealTimePerformanceTuning2022}
+~\cite{MG996RServoMotor}
+~\cite{DigitalOutputSIGMATEK}
+~\cite{SDIASSIGMATEK}
+~\cite{ControlsHMIsSIGMATEK}
+~\cite{InterfacesSplittersSIGMATEK}
 ~\cite{adamPerformanceAssessmentLinux2021}
@@
 ~\cite{yoonRealTimePerformanceAnalysis}
-~\clearpage
+~\cite{CPUUnitsSIGMATEK}
+\clearpage
 

### 2024-08-01 Masterarbeit_Robotik_Pamuk.tex
Setup QEMU

@@
     \textbf{QEMU Option} & \textbf{Description} \\\hline
-    \texttt{-object memory-backend-ram,}\newline\texttt{id=ram0,size=4G,prealloc=on} & Locks the memory of the VM to 4GB and prevents \newline it from being swapped out to disk \\\hline
+    \texttt{-object memory-backend-ram,}\newline\texttt{id=ram0,size=4G,prealloc=on} & Locks the memory of the VM to 4GB and prevents it from being swapped out to disk \\\hline
     \texttt{-mem-prealloc} \newline \texttt{-mem-path /dev/hugepages/} & Enables the use of hugepages and improves \newline memory access \\\hline
@@
 		-drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
-		-cpu host,hv-passthrough \
 		-object memory-backend-ram,id=ram0,size=4G,prealloc=on \
@@
 \noindent Connecting a second servo motor of the mini-robot means repeating the process of reducing the voltage through resistors for the second PWM output of the PW 022 module.
-
-\noindent The program was written in Lasal Class 2 and will be applied to all three mentioned versions of Salamander 4 to measure the reaction time of the robot to the specific commands.
+The connection between the PW 022 module and the motor of the mini-robot is demonstrated in Figure \ref{fig:pw022_minirobot_connection}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{img/experiment/pw022_minirobot.png}
+	\caption[Connection between PW 022 and mini-robot]{Connection between PW 022 and mini-robot~\cite{MG996RDigitalServo}}
+	\label{fig:pw022_minirobot_connection}
+\end{figure}
+
+\begin{comment}
+	battery -> https://www.flaticon.com/de/kostenloses-icon/autobatterie_4238252?related_id=4238254
+\end{comment}
+
+\noindent The program was written in Lasal Class 2 and was applied to all three mentioned versions of Salamander 4 to measure the reaction time of the robot to the specific commands.
 Prior to examining the software program, the next two subsections briefly explain the setup of each version of the experiment.
@@
 \subsection{Setup of Hardware Salamander 4}
-
 In this version, Salamander 4 runs on the CP 841~\cite{CPUEinheitenSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
-The PW 022 module is directly mounted on the CPU via the S-DIAS bus~\cite{SDIASSIGMATEK} and communicates over the hard real-time capable Ethernet VARAN with 100 Mbit/s.
-The setup is visible on Figure \ref{fig:hardware_tree}
+The PW 022 module is directly mounted on the CPU via the S-DIAS bus and communicates over the hard real-time capable Ethernet VARAN with 100 Mbit/s~\cite{SDIASSIGMATEK}.
+The setup is visible on Figure \ref{fig:hardware_tree}.
 
@@
 \subsection{Setup of QEMU Salamander 4}
-To connect QEMU with the PWM module, the PCV 522 VARAN Manager PCI Insert Card~\cite{ControlsHMIsSIGMATEK} needed to be plugged into the working station PC.
-An additional Varan Connection module VI 021~\cite{InterfacesSplittersSIGMATEK} was required to connect the PCV 522 module to the PW 022 module that generates the signal to move the mini-robot.
+In the virtualized setup of the experiment, the VCPU functionality of QEMU is used to connect QEMU with the PWM module.
+In order to achieve that, the PCV 522 VARAN Manager PCI Insert Card~\cite{ControlsHMIsSIGMATEK}, which serves as a bridge between the PC and the rest of the setup, needed to be plugged into the PC.
+The command \texttt{lspci -nn} lists all PCI devices along with their vendor and device IDs.
+In this case, the command for binding the PCV 522 module to the VFIO-PCI driver was \texttt{sudo sh -c 'echo "5112 2200" > /sys/bus/pci/drivers/vfio-pci/new\_id'} with vendor ID 5112 and device ID 2200.
+To verify that the device has been successfully bound to the VFIO-PCI driver, \texttt{lspci -v} can be used.
+On top of the binding process, the QEMU script from the previous sections also needed to be modified to include the PCV 522 VARAN Manager PCI Insert Card.
+This is done by adding \texttt{-device\ vfio-pci,host=03:00.0 } to the QEMU script, where \texttt{03:00.0} refers to the PCI address of the device, with \texttt{03} being the bus number, \texttt{00} the device number, and \texttt{0} the function number.
+The final script can be seen in Code~\ref{script:qemu_def_pci}.
+
+\vspace{1em}
+\begin{minipage}{\linewidth}
+	\begin{lstlisting}[name={Include PCI in QEMU script for Salamander 4 virtualization},label={script:qemu_def_pci}]
+	#!/bin/sh
+
+	if  [ ! -d drive-c/ ]; then
+			echo "Filling drive-c/"
+			mkdir drive-c/
+			tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
+	fi
+
+	exec taskset -c 4 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
+	-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
+	-append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable nohlt idle=poll quiet xeno_hal.smi=1 xenomai.smi=1 threadirqs" \
+	-net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
+	-fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
+	-device vhost-vsock-pci,guest-cid=3,id=vsock0 \
+	-drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
+	-object memory-backend-ram,id=ram0,size=4G,prealloc=on \
+	-mem-prealloc -mem-path /dev/hugepages \
+	-device vfio-pci,host=03:00.0 \
+	-no-reboot -nographic
+\end{lstlisting}
+\end{minipage}
+
+\noindent An additional Varan Connection module VI 021~\cite{InterfacesSplittersSIGMATEK} was required to enable the connection between the PCV 522 module and the PW 022 module that generates the signal to move the mini-robot.
 This setup is depicted in Figure \ref{fig:virt_tree}.
@@
 
+\noindent Essentially, the CP 841 CPU is being emulated by the VCPU functionality of QEMU, allowing the PCV 522 VARAN Manager PCI Insert Card and the Varan Connection module VI 021 to interact as if they were communicating with a physical CPU unit.
+
+\clearpage
 \section{Robotic Application}

### 2024-08-05 Masterarbeit_Robotik_Pamuk.tex
text-based environment

@@
 Every CPU other than the isolated CPU 4 is allowed.
-Output~\ref{output:pse} shows the user and kernel tasks that run on CPU 4.
+Code~\ref{output:pse} shows the user and kernel tasks that run on CPU 4.
 After the isolation, user tasks other than the QEMU process have been removed from running on this CPU.
@@
 This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems.
-The ~\texttt{/proc/interrupts} file can be monitored using the command ~\texttt{watch -d -c cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
+The ~\texttt{/proc/interrupts} file can be monitored using the command ~\texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
 The IRQs needed to be removed from the isolated CPU by manipulating the ~\texttt{/proc/irq/<IRQ>/smp\_affinity} files.
@@
 By changing the values of the ~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled.
-However, as the proc filesystem resets to its default state after each reboot, manually changing numerous IRQ files is tedious and time-consuming.
-This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot.
-Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads.
-This can be done through finding out the respective process IDs through \texttt{ps -e | grep 'irq/.*nvme'} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.
 
@@
 \end{minipage}
+
+\noindent However, as the proc filesystem resets to its default state after each reboot, manually changing numerous IRQ files is tedious and time-consuming.
+This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot.
+Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads.
+This can be done through finding out the respective process IDs through \texttt{ps -e | grep irq/.*nvme} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.
 
@@
 Hence it is best to disable them in real-time scenarios by writing a ``0'' to the \texttt{/sys/devices/system/machinecheck/machinecheck0/check\_interval} file.
+
+\subsection{Boot into text-based environment}
+The Graphical User Interface (GUI) consumes great amounts of system resources.
+It often runs different background processes that are not essential for real-time systems and hence increase latency.
+These resources can be freed up by switching to a text-based environment.
+This way, processing power and memory can be allocated to critical real-time tasks which in turn can lead to lower latency and deterministic behavior.
+The command \texttt{systemctl set-default multi-user.target} and a following reboot allow for booting into a text-based environment.
+For the sake of completeness, the graphical interface can be brought back through the command \texttt{systemctl set-default graphical.target} and a following reboot.
 

### 2024-08-05 Masterarbeit_Robotik_Pamuk.tex
Remove from to include and add to rt performance tuning

@@
 
+\bigskip \noindent It is important to mention that a real-time system implies determinism at every component of the system ranging from hardware over the kernel to the software and the application on it.
+As already established earlier, the guest OS Salamamder 4 runs Xenomai and is hence real-time capable.
+The host system also needs to be aware of the real-time determinism in order to achieve highest possible reliability.
+The very step of achieving the goal of reducing latency is to apply the preempt-rt patch \cite{RealtimePreempt_rt_versionsWiki} to the host system.
+This patch is a set of modifications to the Linux kernel with the goal of making it fully preemptible.
+This means, almost all parts of the kernel can be preempted and higher-priority tasks can interrupt lower-priority ones.
+This significantly reduces the time it takes for high-priority tasks to start executing after an event \cite{RealtimeKernelPatchset}.
+The patch also includes support for high-resolution timers and make the system more predictable, which are crucial for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling \cite{rostedtInternalsRTPatch}.
+
+\bigskip \noindent Nevertheless, a real-time kernel alone does not make a system truly “real-time” \cite{WhatRealtimeLinux}.
+Additional modifications are required to achieve this.
+After each tuning, the latency test of Xenomai will be executed to evaluate the improved latency of the system.
+
 \section{BIOS Configurations}\label{sec:bios_configurations}
@@
 	\end{itemize}
+
+After the BIOS configurations,
 
@@
 \begin{comment}
-
-\chapter{KVM exit reasons}\label{cha:kvm_exit_reasons}
-\section{APIC\_WRITE}
-\section{HLT}
-\section{EPT\_MISCONFIG}
-\section{PREEMPTION\_TIMER}
-\section{EXTERNAL\_INTERRUPT}
-\section{IO\_INSTRUCTION}
-\section{EOI\_INDUCED}
-\section{EPT\_VIOLATION}
-\section{PAUSE\_INSTRUCTION}
-\section{CPUID}
-\section{MSR\_READ}
-
 \chapter{To Include}\label{cha:to_include}
@@
 Vorgehensweise von~\cite{linPerformanceEvaluationXenomai}
-\section{else}
+\section{Latency vapic}
 Figure~\ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
@@
 
+\section{Latency taskset}
 Figure~\ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
@@
 
+\section{kvm exits}
 \begin{table}[H]
@@
 
+\section{Host and Guest tasks}
 \begin{table}[H]
@@
 
+\section{Citation table}
 \begin{table}[H]
@@
 	\hline
-	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core.
-\\
-	\hline
-	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core.
-\\
-	\hline
-	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core.
-\\
-	\hline
-	\cite{maPerformanceTuningKVMbased} & Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System & It means that all threads, except for some per-core kernel threads, were prevented from running on the second processor core.
-\\
+	\cite{maPerformanceTuningKVMbased} & book & citation \\
+	\hline
+	\cite{maPerformanceTuningKVMbased} & book & citation \\
+	\hline
+	\cite{maPerformanceTuningKVMbased} & book & citation \\
+	\hline
+	\cite{maPerformanceTuningKVMbased} & book & citation \\
 	\hline
@@
 
-	Figure~\ref{fig:power_saver_kvm_exit} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/power_saver/kvm_exit_count.png}
-		\caption[power\_saver kvm\_exit\_count]{power\_saver kvm\_exit\_count}
-		\label{fig:power_saver_kvm_exit}
-	\end{figure}
-	\clearpage
-
-	Figure~\ref{fig:power_saver_kvm_host} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/power_saver/results_host_report.png}
-		\caption[power\_saver host report]{power\_saver host report}
-		\label{fig:power_saver_kvm_host}
-	\end{figure}
-	\clearpage
-
-	Figure~\ref{fig:power_saver_kvm_guest} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/power_saver/results_guest_report.png}
-		\caption[power\_saver guest report]{power\_saver guest report}
-		\label{fig:power_saver_kvm_guest}
-	\end{figure}
-	\clearpage
-
-	Figure~\ref{fig:balanced_kvm_exit} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/balanced/kvm_exit_count.png}
-		\caption[balanced kvm\_exit\_count]{balanced kvm\_exit\_count}
-		\label{fig:balanced_kvm_exit}
-	\end{figure}
-	\clearpage
-
-	Figure~\ref{fig:balanced_kvm_host} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/balanced/results_host_report.png}
-		\caption[balanced host report]{balanced host report}
-		\label{fig:balanced_kvm_host}
-	\end{figure}
-	\clearpage
-
-	Figure~\ref{fig:balanced_kvm_guest} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/balanced/results_guest_report.png}
-		\caption[balanced guest report]{balanced guest report}
-		\label{fig:balanced_kvm_guest}
-	\end{figure}
-	\clearpage
-
-	Figure~\ref{fig:performance_kvm_exit} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/performance/kvm_exit_count.png}
-		\caption[performance \_exit\_count]{performance \_exit\_count}
-		\label{fig:performance_kvm_exit}
-	\end{figure}
-	\clearpage
-
-	Figure~\ref{fig:performance_kvm_host} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/performance/results_host_report.png}
-		\caption[performance host report]{performance host report}
-		\label{fig:performance_kvm_host}
-	\end{figure}
-	\clearpage
-
-	Figure~\ref{fig:performance_kvm_guest} shows ...
-
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=1.0\columnwidth]{img/performance/results_guest_report.png}
-		\caption[balanced guest report]{balanced guest report}
-		\label{fig:performance_kvm_guest}
-	\end{figure}
-	\clearpage
-
 	\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
@@
 
-	Figure~\ref{fig:gnuplot_max_latency_default}
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
-		\caption[gnuplot latency no taskset]{gnuplot latency no taskset}
-		\label{fig:gnuplot_max_latency_default}
-	\end{figure}
-
+\section{gnuplot latency with taskset}
 	Figure~\ref{fig:gnuplot_max_latency_taskset}
@@
 		\label{fig:gnuplot_max_latency_taskset}
-	\end{figure}
-
-	Figure~\ref{fig:gnuplot_max_latency_vapic}
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/vapic/gnuplot_max_latency_vapic.png}
-		\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
-		\label{fig:gnuplot_max_latency_vapic}
 	\end{figure}
@@
 
-	The Advanced Programmable Interrupt Controller (APIC) is responsible for the distribution of interrupts in x86 and Itanium-based computer systems.
-An APIC\_WRITE occurs when a guest operating system attempts to write to the APIC registers.
-Since the APIC is a physical hardware component, KVM must intercept this operation and cause a VM exit.
-To avoid this, newer Intel processors offer hardware virtualization of the Advanced Programmable Interrupt Controller (APICv).
-APICv improves virtualized AMD64 and Intel 64 guest performance by allowing the guest to directly access the APIC, dramatically cutting down interrupt latencies and the number of virtual machine exits caused by the APIC.
-This feature is used by default in newer Intel processors and improves I/O performance.
-
-This can be done by setting the apic flag to 'v' in the VM configuration file.
-
-Figure~\ref{fig:kvm_exit_vapic} shows ~\texttt{kvm\_exit} frequency with APIC virtualization.
-Comparing this to the previous Figures~\ref{fig:kvm_exit_taskset} and~\ref{fig:kvm_exits_default}, it can be observed that an APIC\_Write no longer occurs.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
-	\caption[kvm exits default]{kvm exits default}
-	\label{fig:kvm_exit_vapic}
-\end{figure}
-
+\clearpage
 \end{comment}
-
 \chapter{Results}\label{cha:results}
@@
 ~\cite{QEMU}
+~\cite{RealtimePreempt_rt_versionsWiki}
+~\cite{RealtimeKernelPatchset}
+~\cite{rostedtInternalsRTPatch}
+~\cite{WhatRealtimeLinux}
 ~\cite{maPerformanceTuningKVMbased}
@@
 ~\cite{DigitalOutputSIGMATEK}
+~\cite{MG996RDigitalServo}
 ~\cite{SDIASSIGMATEK}
@@
 
-\begin{comment}
-
-VARAN-Bus PCV-521\newline
-Windows\newline
-Ubuntu\newline
-WSL\newline\newline\newline
-
-In Xenomai 3, you need:
-
-The Linux kernel with the Dovetail interface, which allows the kernel to handle real-time tasks with low latency.
-The Cobalt real-time core, which is responsible for scheduling and executing real-time tasks.
-In Xenomai 4, you need:
-The Linux kernel with the EVL interface, which, like Dovetail, enables the kernel to handle real-time tasks with low latency.
-The libevl library, which provides a user interface to the EVL core, allowing applications to use the real-time capabilities provided by EVL.
-So, while both versions require a Linux kernel with a real-time interface (Dovetail for Xenomai 3, EVL for Xenomai 4), Xenomai 3 uses a real-time core (Cobalt), while Xenomai 4 uses a library (libevl) to provide applications with access to its real-time capabilities.
-\end{comment}
-\clearpage
-
 \chapter{Discussion}\label{cha:discussion}

### 2024-08-05 Masterarbeit_Robotik_Pamuk.tex
Figure and Table after each Tuning

@@
 	\centering
-	\includegraphics[width=0.75\columnwidth]{img/max_latency_default.png}
+	\includegraphics[width=1.0\columnwidth]{img/max_latency_default.png}
 	\caption[Latency with default settings]{Latency with default settings}
@@
 
-After the BIOS configurations,
+After the BIOS configurations, Figure \ref{fig:max_latency_rt}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/3rt/max_latency_rt/max_latency_rt.png}
+	\caption[After BIOS configurations]{After BIOS configurations}
+	\label{fig:max_latency_rt}
+\end{figure}
+
+\begin{table}[H]
+	\centering
+	\caption[Latency statistics of Salamander 4 after BIOS configurations]{Latency statistics of Salamander 4 virtualization after BIOS configurations in microseconds}
+	\label{tab:latency_statistics_virt}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|c|c|}\hline
+	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
+	Average Latency & 17.68 \\\hline
+	Maximum Latency & 32.21 \\\hline
+	Minimum Latency & 8.005 \\\hline
+	Standard Deviation & 6.1 \\\hline
+	\end{tabular}}
+	\end{table}
 
@@
 
+After the Kernel configurations, Figure \ref{fig:max_latency_rt_kernelparam}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/max_latency_rt_kernelparam/max_latency_rt_kernelparam.png}
+	\caption[After Kernel configurations]{After Kernel configurations}
+	\label{fig:max_latency_rt_kernelparam}
+\end{figure}
+
+\begin{table}[H]
+	\centering
+	\caption[Latency statistics of Salamander 4 after Kernel configurations]{Latency statistics of Salamander 4 virtualization after Kernel configurations in microseconds}
+	\label{tab:latency_statistics_virt}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|c|c|}\hline
+	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
+	Average Latency & 14.00 \\\hline
+	Maximum Latency & 21.694 \\\hline
+	Minimum Latency & 6.351 \\\hline
+	Standard Deviation & 1.44 \\\hline
+	\end{tabular}}
+	\end{table}
+
 \section{Host OS Configurations}\label{sec:host_configurations}
@@
 
+After the Host configurations, Figure \ref{fig:max_latency_rt_kernelparam_host}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/5rt_kernelparam_host/max_latency_rt_kernelparam_host/max_latency_rt_kernelparam_host.png}
+	\caption[After Host configurations]{After Host configurations}
+	\label{fig:max_latency_rt_kernelparam_host}
+\end{figure}
+
+\begin{table}[H]
+	\centering
+	\caption[Latency statistics of Salamander 4 after Host configurations]{Latency statistics of Salamander 4 virtualization after Host configurations in microseconds}
+	\label{tab:latency_statistics_virt}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|c|c|}\hline
+	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
+	Average Latency & 12.61 \\\hline
+	Maximum Latency & 17.041 \\\hline
+	Minimum Latency & 7.872 \\\hline
+	Standard Deviation & 1.8 \\\hline
+	\end{tabular}}
+	\end{table}
+
 \section{QEMU/KVM Configurations}\label{sec:guest_configurations}
@@
 	\end{minipage}
+
+	After the QEMU configurations, Figure \ref{fig:max_latency_rt_kernelparam_host_qemu}
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/6rt_kernelparam_host_qemu/max_latency_rt_kernelparam_host_qemu/max_latency_rt_kernelparam_host_qemu.png}
+		\caption[After QEMU configurations]{After QEMU configurations}
+		\label{fig:max_latency_rt_kernelparam_host_qemu}
+	\end{figure}
+
+	\begin{table}[H]
+		\centering
+		\caption[Latency statistics of Salamander 4 after QEMU configurations]{Latency statistics of Salamander 4 virtualization after QEMU configurations in microseconds}
+		\label{tab:latency_statistics_virt}
+		\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+		{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+		\begin{tabular}{|c|c|}\hline
+		\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
+		Average Latency & 16.26 \\\hline
+		Maximum Latency & 17.134 \\\hline
+		Minimum Latency & 14.532 \\\hline
+		Standard Deviation & 0.3 \\\hline
+		\end{tabular}}
+		\end{table}
 
@@
 \appendix
-
+\begin{comment}
 \chapter{Anhang A}
 \clearpage
-
 \chapter{Anhang B}
+\end{comment}
 

### 2024-08-05 Masterarbeit_Robotik_Pamuk.tex
~\cite

@@
 The host system also needs to be aware of the real-time determinism in order to achieve highest possible reliability.
-The very step of achieving the goal of reducing latency is to apply the preempt-rt patch \cite{RealtimePreempt_rt_versionsWiki} to the host system.
+The very step of achieving the goal of reducing latency is to apply the preempt-rt patch~\cite{RealtimePreempt_rt_versionsWiki} to the host system.
 This patch is a set of modifications to the Linux kernel with the goal of making it fully preemptible.
 This means, almost all parts of the kernel can be preempted and higher-priority tasks can interrupt lower-priority ones.
-This significantly reduces the time it takes for high-priority tasks to start executing after an event \cite{RealtimeKernelPatchset}.
-The patch also includes support for high-resolution timers and make the system more predictable, which are crucial for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling \cite{rostedtInternalsRTPatch}.
-
-\bigskip \noindent Nevertheless, a real-time kernel alone does not make a system truly “real-time” \cite{WhatRealtimeLinux}.
+This significantly reduces the time it takes for high-priority tasks to start executing after an event~\cite{RealtimeKernelPatchset}.
+The patch also includes support for high-resolution timers and make the system more predictable, which are crucial for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling~\cite{rostedtInternalsRTPatch}.
+
+\bigskip \noindent Nevertheless, a real-time kernel alone does not make a system truly “real-time”~\cite{WhatRealtimeLinux}.
 Additional modifications are required to achieve this.

### 2024-08-06 Masterarbeit_Robotik_Pamuk.tex
organize

@@
 
-\noindent Figure~\ref{fig:lstopo} is the output of the ~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
+\noindent Figure~\ref{fig:lstopo} is the output of the~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
 \begin{figure}[H]
@@
 In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
-The output of the command ~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
+The output of the command~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
 
@@
 \section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
-The output of the command ~\texttt{uname -a} for Salamander 4 bare metal is shown in Code~\ref{output:uname_a_hw}.
+The output of the command~\texttt{uname -a} for Salamander 4 bare metal is shown in Code~\ref{output:uname_a_hw}.
 
@@
 \noindent As a reference point, the latency program was executed on Salamander 4 bare metal for a duration of 10 minutes.
-The complete command used was ~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file ~\texttt{gnuplot.txt}.
+The complete command used was~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file~\texttt{gnuplot.txt}.
 Figure~\ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.
@@
 	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process.
-~\texttt{OVMF} stands for Open Virtual Machine Firmware, ~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for ``QEMU Copy On Write version 2''.
+~\texttt{OVMF} stands for Open Virtual Machine Firmware,~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for ``QEMU Copy On Write version 2''.
 	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code~\ref{script:qemu_def}.
 	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform.
-It uses the ~\texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
+It uses the~\texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
 	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Compressed tarball containing a pre-configured environment for the Salamander 4 OS.
-It is unpacked and sets up the ~\texttt{drive-c/} directory with system and log files in the ~\texttt{qemu\_def.sh} script.
+It is unpacked and sets up the~\texttt{drive-c/} directory with system and log files in the~\texttt{qemu\_def.sh} script.
 	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging.
@@
 \noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in chapter~\ref{cha:methodology}.
-The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the ~\texttt{uname -a} command.
+The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the~\texttt{uname -a} command.
 
@@
 
-After the BIOS configurations, Figure \ref{fig:max_latency_rt}
+\textcolor{red}{After the BIOS configurations, Figure \ref{fig:max_latency_rt}}
 
@@
 
-After the Kernel configurations, Figure \ref{fig:max_latency_rt_kernelparam}
+\textcolor{red}{After the Kernel configurations, Figure \ref{fig:max_latency_rt_kernelparam}}
 
@@
 Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-For CPU isolation, the ~\texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
+For CPU isolation, the~\texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
 This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks.
@@
 This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems.
-The ~\texttt{/proc/interrupts} file can be monitored using the command ~\texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
-The IRQs needed to be removed from the isolated CPU by manipulating the ~\texttt{/proc/irq/<IRQ>/smp\_affinity} files.
-The value in the ~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU.
+The~\texttt{/proc/interrupts} file can be monitored using the command~\texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
+The IRQs needed to be removed from the isolated CPU by manipulating the~\texttt{/proc/irq/<IRQ>/smp\_affinity} files.
+The value in the~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU.
 The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU13).
-When there are 14 CPUs available, the default value for ~\texttt{smp\_affinity} would be 3FFF.
+When there are 14 CPUs available, the default value for~\texttt{smp\_affinity} would be 3FFF.
 Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111).
 The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU.
-By changing the values of the ~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled.
+By changing the values of the~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled.
 
@@
 It does this by pausing the process for the remaining 0.05 seconds, which is not desired because this could result in missed deadlines.
-RT throttling can be disabled by  writing the value -1 to the ~\texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file.
+RT throttling can be disabled by  writing the value -1 to the~\texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file.
 This change also needs to be made permanent because the proc filesystem resets to its default state after each reboot.
-For this purpose, the line ~\texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the ~\texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters.
+For this purpose, the line~\texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the~\texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters.
 This reduces the potential for missed deadlines.
@@
 In a real-time system, this can introduce latency and jitter.
-To disable timer migration, the value ``0'' needs to be written to the ~\texttt{/proc/sys/kernel/timer\_migration} file.
-This change also needs to be made permanent by writing the line ~\texttt{kernel.timer\_migration = 0} to the ~\texttt{/etc/sysctl.conf} file.
+To disable timer migration, the value ``0'' needs to be written to the~\texttt{/proc/sys/kernel/timer\_migration} file.
+This change also needs to be made permanent by writing the line~\texttt{kernel.timer\_migration = 0} to the~\texttt{/etc/sysctl.conf} file.
 This reduces the amount of context switches and interrupts.
@@
 The number of seconds the kernel should wait before checking for stalled CPUs and reporting a stall warning can be set via the ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_timeout} file.
-These warnings can be suppressed alltogether to reduce the potential for increased latency by writing ``1'' to the ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file.
+These warnings can be suppressed alltogether to reduce the potential for increased latency by writing ``1'' to the~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file.
 This setting is also not persistent across reboots, so the command needs to be added to a startup script.
@@
 
-After the Host configurations, Figure \ref{fig:max_latency_rt_kernelparam_host}
+\textcolor{red}{After the Host configurations, Figure \ref{fig:max_latency_rt_kernelparam_host}}
 
@@
 
-	After the QEMU configurations, Figure \ref{fig:max_latency_rt_kernelparam_host_qemu}
+	\textcolor{red}{After the QEMU configurations, Figure \ref{fig:max_latency_rt_kernelparam_host_qemu}}
 
@@
 A Checklist for Writing Linux Real-Time Applications is provided by John Ogness in~\cite{thelinuxfoundationChecklistWritingLinux2020}.
-Every layer of the system stack must be deterministic to ensure predictable and reliable latency, including hardware, operating system, middleware and drivers, and the application software.~\cite{HOWTOBuildRTapplication} and~\cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real time preemption patch in great detail.
+Every layer of the system stack must be deterministic to ensure predictable and reliable latency, including hardware, operating system, middleware and drivers, and the application software.
+~\cite{HOWTOBuildRTapplication} and~\cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real time preemption patch in great detail.
 Various hardware and software tunings are mentioned in~\cite{KVMQemuVirtualization} and~\cite{RealTimePerformanceTuning2022}.
@@
 
-\begin{comment}
 \chapter{To Include}\label{cha:to_include}
@@
 
-\enquote{This LaTeX.}
+\noindent \enquote{This LaTeX.}
 
@@
 Vorgehensweise von~\cite{linPerformanceEvaluationXenomai}
-\section{Latency vapic}
-Figure~\ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.75\columnwidth]{img/max_latency_vapic.png}
-	\caption[Latency vapic]{Latency vapic}
-	\label{fig:max_latency_vapic}
-\end{figure}
 
@@
 As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on user level.
+
+\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
+This explains why there are fewer~\texttt{kvm\_exit} events.
+However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a~\texttt{kvm\_exit} and kvm\_entry.
+
+\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in~\texttt{kvm\_exit} events.
+This is because every context switch takes place at hypervisor level.
+Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.
 
@@
 This led to the investigation of the causes for the observed high and inconsistent latency.
-The guest operates within the kvm\_entry and ~\texttt{kvm\_exit} events of the host.
-Kernelshark revealed a high frequency of ~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
+The guest operates within the~\texttt{kvm\_entry} and~\texttt{kvm\_exit} events of the host.
+Kernelshark revealed a high frequency of~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
 This frequent switching hinders the guest\'s ability to run continuously, thereby increasing the virtualization latency.
 To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
-Specifically, the causes for ~\texttt{kvm\_exit} events were analyzed.
+Specifically, the causes for~\texttt{kvm\_exit} events were analyzed.
 The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
 The results in Figure~\ref{fig:kvm_exit_taskset} were obtained.
-Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed ~\texttt{kvm\_exit} events.
+Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed~\texttt{kvm\_exit} events.
 
@@
 
-\clearpage
-Figure~\ref{fig:kvm_exit_taskset} shows ~\texttt{kvm\_exit} frequency with CPU islation.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_taskset.png}
-	\caption[kvm exits]{kvm exits}
-	\label{fig:kvm_exit_taskset}
-\end{figure}
-
-Figure~\ref{fig:kvm_exits_default} shows ~\texttt{kvm\_exit} frequency without CPU islation.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_default.png}
-	\caption[kvm exits default]{kvm exits default}
-	\label{fig:kvm_exits_default}
-\end{figure}
+\noindent In the process of analyzing the~\texttt{kvm\_exit} events, several reasons for these exits were identified.
+The most frequent among these were the~\texttt{APIC\_WRITE} and~\texttt{HLT} events.
+The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
+The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
+Other significant but less frequent events included~\texttt{EXTERNAL\_INTERRUPT} and~\texttt{IO\_INSTRUCTION}.
+These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
+Events such as~\texttt{EPT\_MISCONFIG} and~\texttt{PREEMPTION\_TIMER} were also noted.
+These could potentially signal issues with memory management and the host's scheduling of the guest.
+While events like~\texttt{PAUSE\_INSTRUCTION},~\texttt{EPT\_VIOLATION},~\texttt{EOI\_INDUCED},~\texttt{MSR\_READ}, and~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
 
@@
 
-The Figure~\ref{fig:latency_comparison} below compares non-optimized guest latency with optimized guest latency and includes optimized bare-metal latency as a reference.
+The Figure below (deleted) compares non-optimized guest latency with optimized guest latency and includes optimized bare-metal latency as a reference.
 The data shows that a 40\% reduction in QD1 latency is achievable through system tuning.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/latency_comparison.png}
-	\caption[latency comparison]{latency comparison}
-	\label{fig:latency_comparison}
-\end{figure}
-\clearpage
 
@@
 	\end{table}
-
-	\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
-This explains why there are fewer ~\texttt{kvm\_exit} events.
-However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a ~\texttt{kvm\_exit} and kvm\_entry.
-
-	\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in ~\texttt{kvm\_exit} events.
-This is because every context switch takes place at hypervisor level.
-Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.
-
-\section{gnuplot latency with taskset}
-	Figure~\ref{fig:gnuplot_max_latency_taskset}
-	\begin{figure}[H]
-		\centering
-		\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
-		\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
-		\label{fig:gnuplot_max_latency_taskset}
-	\end{figure}
-
-	\noindent In the process of analyzing the ~\texttt{kvm\_exit} events, several reasons for these exits were identified.
-The most frequent among these were the ~\texttt{APIC\_WRITE} and ~\texttt{HLT} events.
-The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
-The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
-Other significant but less frequent events included ~\texttt{EXTERNAL\_INTERRUPT} and ~\texttt{IO\_INSTRUCTION}.
-These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
-Events such as ~\texttt{EPT\_MISCONFIG} and ~\texttt{PREEMPTION\_TIMER} were also noted.
-These could potentially signal issues with memory management and the host's scheduling of the guest.
-While events like ~\texttt{PAUSE\_INSTRUCTION}, ~\texttt{EPT\_VIOLATION}, ~\texttt{EOI\_INDUCED}, ~\texttt{MSR\_READ}, and ~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
-
-\clearpage
-\end{comment}
+\clearpage
+
 \chapter{Results}\label{cha:results}

### 2024-08-06 Masterarbeit_Robotik_Pamuk.tex
kvm_exit

@@
 
-As a starting point, initial latency values of both the bare metal and virtualization versions were measured with the latency tool of the xenomai tool suite.
+As a starting point, initial latency values of both the bare metal and virtualization versions were measured with the \texttt{latency} tool of the xenomai tool suite.
 Salamander 4 bare metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system.
@@
 
-\clearpage
-
 \noindent Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time with default settings.
@@
 Additional modifications are required to achieve this.
+The next sections will deal with these real-time performance tunings.
+The very first step in this regard is the isolation of a CPU to dedicate it solely to the desired real-time task.
+CPU isolation and affinity will be covered in detail later in section~\ref{subsec:cpu_isolation}, but its impact will be briefly demonstrated here.
+
+Figure~\ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/max_latency_taskset.png}
+	\caption[Latency taskset]{Latency taskset}
+	\label{fig:max_latency_taskset}
+\end{figure}
+
 After each tuning, the latency test of Xenomai will be executed to evaluate the improved latency of the system.
+
+\textcolor{red}{After Article, Figure \ref{fig:max_latency_rt}}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/3rt/max_latency_rt/max_latency_rt.png}
+	\caption[After Article configurations]{After Article configurations}
+	\label{fig:max_latency_rt}
+\end{figure}
+
+\begin{table}[H]
+	\centering
+	\caption[Latency statistics of Salamander 4 after BIOS configurations]{Latency statistics of Salamander 4 virtualization after BIOS configurations in microseconds}
+	\label{tab:latency_statistics_virt}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|c|c|}\hline
+	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
+	Average Latency & 17.68 \\\hline
+	Maximum Latency & 32.21 \\\hline
+	Minimum Latency & 8.005 \\\hline
+	Standard Deviation & 6.1 \\\hline
+	\end{tabular}}
+	\end{table}
 
@@
 	\end{itemize}
-
-\textcolor{red}{After the BIOS configurations, Figure \ref{fig:max_latency_rt}}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/3rt/max_latency_rt/max_latency_rt.png}
-	\caption[After BIOS configurations]{After BIOS configurations}
-	\label{fig:max_latency_rt}
-\end{figure}
-
-\begin{table}[H]
-	\centering
-	\caption[Latency statistics of Salamander 4 after BIOS configurations]{Latency statistics of Salamander 4 virtualization after BIOS configurations in microseconds}
-	\label{tab:latency_statistics_virt}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-	\begin{tabular}{|c|c|}\hline
-	\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
-	Average Latency & 17.68 \\\hline
-	Maximum Latency & 32.21 \\\hline
-	Minimum Latency & 8.005 \\\hline
-	Standard Deviation & 6.1 \\\hline
-	\end{tabular}}
-	\end{table}
 
@@
 
-\textcolor{red}{After the Kernel configurations, Figure \ref{fig:max_latency_rt_kernelparam}}
+\textcolor{red}{After the BIOS and Kernel configurations, Figure \ref{fig:max_latency_rt_kernelparam}}
 
@@
 	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/max_latency_rt_kernelparam/max_latency_rt_kernelparam.png}
-	\caption[After Kernel configurations]{After Kernel configurations}
+	\caption[After BIOS and Kernel configurations]{After BIOS and Kernel configurations}
 	\label{fig:max_latency_rt_kernelparam}
@@
 \end{minipage}
+\subsection{KVM entry and KVM exit}\label{subsec:kvm_entry_exit}
+\textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
+However, the isolcpus function only isolates CPUs from the scheduler, meaning that user-space processes won't be scheduled on these CPUs, but kernel tasks and interrupts can still utilize the CPU.This led to the investigation of the causes for the observed high and inconsistent latency.
+The guest operates within the~\texttt{kvm\_entry} and~\texttt{kvm\_exit} events of the host.
+Kernelshark revealed a high frequency of~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
+This frequent switching hinders the guest's ability to run continuously, thereby increasing the virtualization latency.}
+
+\textcolor{red}{To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
+Specifically, the causes for~\texttt{kvm\_exit} events were analyzed.
+The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
+The results in Figure~\ref{fig:kvm_exit_taskset} were obtained.
+Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed~\texttt{kvm\_exit} events.}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|l|p{0.62\textwidth}|}
+		\hline
+		\textbf{Exit Reason} & \textbf{Description}                                     \\ \hline
+		APIC\_WRITE          & Triggered when the guest writes to its APIC.             \\ \hline
+		EXTERNAL\_INTERRUPT  & Triggered by external hardware interrupts.               \\ \hline
+		HLT                  & Triggered when the guest executes the HLT instruction.   \\ \hline
+		EPT\_MISCONFIG       & Triggered by a misconfiguration in the EPT.              \\ \hline
+		PREEMPTION\_TIMER    & Triggered when the host's preemption timer expires.      \\ \hline
+		PAUSE\_INSTRUCTION   & Triggered when the PAUSE instruction is executed.        \\ \hline
+		EPT\_VIOLATION       & Triggered by a violation of the EPT permission settings. \\ \hline
+		IO\_INSTRUCTION      & Triggered when the guest executes an I/O instruction.    \\ \hline
+		EOI\_INDUCED         & Triggered when an EOI signal is sent to the APIC.        \\ \hline
+		MSR\_READ            & Triggered when the guest reads from a MSR.               \\ \hline
+		CPUID                & Triggered when the guest executes the CPUID instruction. \\ \hline
+	\end{tabular}
+	\caption{Description of kvm\_exit reasons}
+	\label{tab:kvm_exit}
+\end{table}
+
+\textcolor{red}{\noindent In the process of analyzing the~\texttt{kvm\_exit} events, several reasons for these exits were identified.
+The most frequent among these were the~\texttt{APIC\_WRITE} and~\texttt{HLT} events.
+The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
+The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
+Other significant but less frequent events included~\texttt{EXTERNAL\_INTERRUPT} and~\texttt{IO\_INSTRUCTION}.
+These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
+Events such as~\texttt{EPT\_MISCONFIG} and~\texttt{PREEMPTION\_TIMER} were also noted.
+These could potentially signal issues with memory management and the host's scheduling of the guest.
+While events like~\texttt{PAUSE\_INSTRUCTION},~\texttt{EPT\_VIOLATION},~\texttt{EOI\_INDUCED},~\texttt{MSR\_READ}, and~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.}
+
+\textcolor{red}{\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is not bound to a specific CPU core.
+This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
+As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on the user level.
+Without CPU isolation, context switches take place at the operating system level and not at the hypervisor level.
+This explains why there are fewer~\texttt{kvm\_exit} events.
+However, this will also, as previously shown, lead to higher latency, as context switches at the operating system level generally take longer than a~\texttt{kvm\_exit} and ~\texttt{kvm\_entry}.
+When the CPU was dedicated to the QEMU process, on the other hand, there was a significant increase in~\texttt{kvm\_exit} events.
+This is because every context switch takes place at the hypervisor level.
+Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.}
+
 \subsection{Interrupt Affinity}\label{subsec:irq_handling}
@@
 
+\begin{comment}
 \chapter{To Include}\label{cha:to_include}
+
 ``korrekt''
@@
 
-\section{Latency Comparison}
-In the initial phase, a comparative latency analysis was conducted between the hardware version and the virtualized version of Salamander 4.
-For this purpose, the latency tool of the Xenomai test suite was used.
-The latency was measured under two conditions, idle and CPU-stressed.
-The goal was to optimize the latency of the virtualization of Salamander 4 OS to closely match that of the bare metal version.
-
-Vorgehensweise von~\cite{linPerformanceEvaluationXenomai}
-
+The latency was measured under two conditions, idle and CPU-stressed. Vorgehensweise von~\cite{linPerformanceEvaluationXenomai}
+
+\section{Trace-cmd and Kernelshark}
 After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
-
-\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is not bound to a specific CPU core.
-This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
-As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on user level.
-
-\noindent Without CPU isolation, context switches take place at operating system level and not at hypervisor level.
-This explains why there are fewer~\texttt{kvm\_exit} events.
-However, this will also, as previously shown, lead to higher latency, as context switches at operating system level generally take longer than a~\texttt{kvm\_exit} and kvm\_entry.
-
-\noindent When the CPU was dedicated to the QEMU process, on the other hand,there was a significant increase in~\texttt{kvm\_exit} events.
-This is because every context switch takes place at hypervisor level.
-Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.
-
-\section{Latency taskset}
-Figure~\ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.75\columnwidth]{img/max_latency_taskset.png}
-	\caption[Latency taskset]{Latency taskset}
-	\label{fig:max_latency_taskset}
-\end{figure}
-
-\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
-However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
-Consequently, these kernel tasks and interrupts can still utilize the CPU.
-This led to the investigation of the causes for the observed high and inconsistent latency.
-The guest operates within the~\texttt{kvm\_entry} and~\texttt{kvm\_exit} events of the host.
-Kernelshark revealed a high frequency of~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
-This frequent switching hinders the guest\'s ability to run continuously, thereby increasing the virtualization latency.
-To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
-Specifically, the causes for~\texttt{kvm\_exit} events were analyzed.
-The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
-The results in Figure~\ref{fig:kvm_exit_taskset} were obtained.
-Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed~\texttt{kvm\_exit} events.
-
-\section{kvm exits}
-\begin{table}[H]
-	\centering
-	\begin{tabular}{|l|p{0.62\textwidth}|}
-		\hline
-		\textbf{Exit Reason} & \textbf{Description}                                     \\
-		\hline
-		APIC\_WRITE          & Triggered when the guest writes to its APIC.             \\ \hline
-		EXTERNAL\_INTERRUPT  & Triggered by external hardware interrupts.               \\ \hline
-		HLT                  & Triggered when the guest executes the HLT instruction.   \\ \hline
-		EPT\_MISCONFIG       & Triggered by a misconfiguration in the EPT.              \\ \hline
-		PREEMPTION\_TIMER    & Triggered when the host's preemption timer expires.      \\ \hline
-		PAUSE\_INSTRUCTION   & Triggered when the PAUSE instruction is executed.        \\ \hline
-		EPT\_VIOLATION       & Triggered by a violation of the EPT permission settings. \\ \hline
-		IO\_INSTRUCTION      & Triggered when the guest executes an I/O instruction.    \\ \hline
-		EOI\_INDUCED         & Triggered when an EOI signal is sent to the APIC.        \\ \hline
-		MSR\_READ            & Triggered when the guest reads from a MSR.               \\ \hline
-		CPUID                & Triggered when the guest executes the CPUID instruction. \\ \hline
-	\end{tabular}
-	\caption{Description of kvm\_exit reasons}
-	\label{tab:kvm_exit}
-\end{table}
-
-\noindent In the process of analyzing the~\texttt{kvm\_exit} events, several reasons for these exits were identified.
-The most frequent among these were the~\texttt{APIC\_WRITE} and~\texttt{HLT} events.
-The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
-The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
-Other significant but less frequent events included~\texttt{EXTERNAL\_INTERRUPT} and~\texttt{IO\_INSTRUCTION}.
-These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
-Events such as~\texttt{EPT\_MISCONFIG} and~\texttt{PREEMPTION\_TIMER} were also noted.
-These could potentially signal issues with memory management and the host's scheduling of the guest.
-While events like~\texttt{PAUSE\_INSTRUCTION},~\texttt{EPT\_VIOLATION},~\texttt{EOI\_INDUCED},~\texttt{MSR\_READ}, and~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.
-
-\clearpage
 
@@
 \clearpage
+\end{comment}
 

### 2024-08-09 Masterarbeit_Robotik_Pamuk.tex
PW 022 three times

@@
 \chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
-This chapter compares the latency of the Salamander 4 operating system before and after the real-time performance tunings.
-Additionally, these results are contrasted with the latency of Salamander 4 running on bare metal hardware to understand how closely the performance of the virtualization can match that of the bare metal.
+This chapter compares the latency of the Salamander 4 operating system before and after the real-time performance tunings with the latency of Salamander 4 running on bare metal hardware.
+This comparison is done with the aid of a program that will be explained shortly.
+The goal is to understand how closely the performance of the virtualization can match that of the bare metal.
 The experimental setup includes a six-axis mini-robot, illustrated in Figure \ref{fig:mini_robot}.
@@
 
-	exec taskset -c 4 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
+	exec taskset -c 4-5 chrt -f 99 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
 	-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \

### 2024-08-11 Masterarbeit_Robotik_Pamuk.tex
introduction 1.1

@@
 Delays or errors can have catastrophic consequences in some cases.
-Traditional operating systems, such as Windows or Ubuntu, are often not suitable for these types of real-time requirements as they cannot guarantee deterministic execution times.
-Therefore, real-time operating systems are required that are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
-
-\bigskip \noindent The core component of an RTOS that enables real-time capabilities is the kernel.
-The kernel is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior.
+Modern operating systems nowadays are categorized into two types:  general-purpose operating systems (GPOS) and real-time operating systems (RTOS).
+The key difference is whether the system is time-critical or not~\cite{canbazPerformanceAnalysisRealtime2022}.
+In a traditional GPOS, such as Windows or Linux-based distributions, a high-priority task cannot interrupt a kernel operation, which makes them not suitable for real-time requirements, as they cannot guarantee deterministic execution times.
+However, in an RTOS, a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation.
+RTOS are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
+Many Linux distributions can function as both GPOS and RTOS with kernel modifications.
+
+\section{Real-Time Operating Systems}
+\bigskip \noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.The core component of an RTOS that enables real-time capabilities is the kernel.
+It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior.
 It employs preemptive scheduling mechanisms to allow high-priority tasks to preempt lower-priority tasks, ensuring that time-critical tasks are not delayed.
-The kernel also implements priority-based scheduling algorithms, such as Rate Monotonic Scheduling (RMS) or Earliest Deadline First (EDF), to schedule tasks based on their priorities and timing constraints.
-Additionally, RTOS kernels are designed to minimize interrupt latency, which is crucial for real-time applications that require immediate response to external events.
-
-\bigskip \noindent In these RTOS systems, task scheduling is based on so-called priority-based preemptive scheduling.
+Additionally, RTOS kernels are designed to allocate and manage memory resources efficiently and minimize interrupt latency, which is crucial for real-time applications that require immediate response to external events~\cite{wangRealtimeEmbeddedSystems2017}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\columnwidth]{img/rtos_structure.png}
+	\caption[RTOS Structure]{RTOS Structure~\cite{wangRealtimeEmbeddedSystems2017}}
+	\label{fig:rtos_structure}
+\end{figure}
+
+\bigskip \noindent In these real-time operating system, task scheduling is based on so-called priority-based preemptive scheduling~\cite{buttazzoHardRealtimeComputing2024}.
 Each task in a software application is assigned a priority.
@@
 The lower priority task is only executed once the higher priority task has been processed.
-Real-time systems are usually categorized as either soft or hard real-time systems.
+
+\bigskip \noindent Real-time systems are usually classified as either soft or hard real-time systems ~\cite{lipariRealTimeSchedulingHard}.
 The difference lies exclusively in the consequences of a violation of the time limits.
-
-\bigskip \noindent Hard real-time is when the system stops operating if a deadline is missed, which can have catastrophic consequences.
+Hard real-time is when the system stops operating if a deadline is missed, which can have catastrophic consequences.
 Soft real-time exists when a system continues to function even if it cannot perform the tasks within a specified time.
@@
 \clearpage
-\section{Application Context}
-This master's thesis was written at SIGMATEK GmbH \& Co KG~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.
-%This section provides necessary information about the starting point and boundary conditions of the thesis, as the work should be reproducible with the means used. 
-SIGMATEK uses its own customized Linux distribution, namely Salamander 4, to be run on their self-manufactured CPUs.
-Salamander 4 system employs hard real-time with Xenomai 3 and requires a worst latency value between 20 and 50 \textmu s.
-The goal is to virtualize Salamander 4 and approach the performance of bare metal.
-Salamander 4 is built with Yocto and virtualized through Quick Emulator/Kernel-based Virtual Machine (QEMU/KVM).
-The details of this operating system are explained in chapter~\ref{cha:salamander4}.
+
+\section{Application context and boundaries}
+
+This section defines the boundary conditions within which this master's thesis was composed.
+This work was written at SIGMATEK GmbH \& Co KG~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}, a company that researches, develops, produces, and distributes automation technology for industrial machinery and plant engineering.
+SIGMATEK uses its own custom Linux-based operating system, to be run on their self-manufactured CPUs.
+This operating system will be referred to as Salamander 4 in this work.
+The details of this OS are explained in chapter~\ref{cha:salamander4}.
+Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open source project that helps develop custom Linux-based systems, as discussed in Section~\ref{sec:yocto}.
+Salamander 4 is virtualized through Quick emulator (QEMU), an open-source emulator and virtualizer that allows to run different operating systems on a computer, as described in Section~\ref{sec:qemu}.
+Salamander 4 employs hard real-time with Xenomai 3~\cite{XenomaiXenomai}, a real-time framework for Linux that enables deterministic and low-latency performance, detailed in Section~\ref{sec:xenomai}.
+The goal is to virtualize Salamander 4 and approach the performance of bare metal through real-time performance tunings, as shown in chapter~\ref{cha:real-time_tuning}, focussing on latency requirements set by SIGMATEK.
 
@@
 
+\chapter{Theoretical Foundations}
+\section{Yocto}\label{sec:yocto}
+
+\section{QEMU}\label{sec:qemu}
+
+\section{Xenomai}\label{sec:xenomai}
+\noindent Xenomai 3~\cite{XenomaiXenomai} is a real-time framework that offers two paths to real-time performance.
+The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
+Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.5\columnwidth]{img/x3-cobalt-interfaces.png}
+	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
+	\label{fig:cobalt}
+\end{figure}
+
+\bigskip \noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
+Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.5\columnwidth]{img/x3-mercury-interfaces.png}
+	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
+	\label{fig:mercury}
+\end{figure}
+
+\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
+
+\clearpage
+
 The main objective of this work is to create a real-time robot control platform that integrates Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
@@
 
-\chapter{Methodology}\label{cha:methodology}
-
-This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
-
-\begin{comment}
-
-Zu Beginn der Arbeit wurde eine ausführliche Analyse der einzelnen Virtualisierungsmöglichkeiten von Salamander 4 durchgeführt.
-Im Besonderen wurde hier die Virtualisierungsperformance von Ubuntu 22.04, Windows 10 und WSL unter QEMU verglichen.
-
-First, a comprehensive literature review is conducted to understand the current trends and challenges in real-time robot control.
-Based on the literature study, a suitable virtualization platform is selected.
-
-After the selection of the virtualization platform, the robot control platform is implemented.
-This step includes the installation and configuration of Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
-Once the platform has been implemented, the robot is connected via a VARAN bus interface.
-Finally, the platform is evaluated to determine how the integration of real-time functions and efficient communication systems improves the response time and reliability of robot applications.
-
-- Evaluation der Virtualisierungsplattform:
-Ich werde verschiedene Virtualisierungsplattformen wie QEMU, Hyper-V, Virtual Box usw. evaluieren. Dies ist ein wichtiger Schritt, um die beste Plattform für meine Anforderungen zu finden.
-
-- Erstellung und Konfiguration des Systems in der Yocto-Umgebung:
-Ich werde das Yocto-Framework verwenden, um mein Embedded Linux System zu erstellen und zu konfigurieren.
-Yocto bietet viele Tools und Funktionen, die mir bei der Erstellung und Konfiguration meines Systems helfen können.
-
-- Anbindung eines Roboters über eine VARAN-Bus Schnittstelle:
-Ich plane, einen Roboter in mein System zu integrieren.
-Ich werde eine VARAN-Bus Schnittstelle verwenden, um eine schnelle und zuverlässige Kommunikation zwischen dem Roboter und dem Steuerungssystem zu gewährleisten.
-
-\end{comment}
-
-\bigskip \noindent Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
-It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
-Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
-
-\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
-It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
-By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
-
-\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel~\cite{XenomaiXenomai}.
-It enables low-latency and deterministic execution of time-critical tasks.
-Xenomai 3 introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
-A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
-The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
-
-\clearpage
-
-The system configuration is shown in Table~\ref{tab:testbed_configuration}
-
-\begin{table}[H]
-	\centering
-	\caption[System configuration]{System configuration}
-	\label{tab:testbed_configuration}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-	\begin{tabular}{|l|l|}
-	\hline
-	\textbf{CPU} & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
-	\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB \\ \hline
-	\textbf{GPU} & NVIDIA RTX A500 Laptop GPU \\ \hline
-	\textbf{BIOS} & Dell Version 1.12.0 \\ \hline
-	\textbf{OS} & Ubuntu 22.04.4 LTS \\
-	\hline
-	\end{tabular}}
-	\end{table}
-
-\noindent Figure~\ref{fig:lstopo} is the output of the~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
-	\caption[Hardware topology]{Hardware topology}
-	\label{fig:lstopo}
-\end{figure}
-
-\clearpage
-
-\chapter{Salamander 4}\label{cha:salamander4}
+\section{Salamander 4}\label{cha:salamander4}
 This chapter briefly describes the Salamander 4 operating system by SIGMATEK.
 
-\section{Structure}
+\subsection{Structure}
 \noindent Salamander 4 is the proprietary operating system of SIGMATEK.
@@
 
-\section{Memory Management}
+\subsection{Memory Management}
 \noindent For the sake of completeness, Figure~\ref{fig:memory_management} displays the memory management of Salamander 4.
@@
 
-\section{Xenomai}
-\noindent Xenomai 3~\cite{XenomaiXenomai} is a real-time framework that offers two paths to real-time performance.
-The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
-Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.5\columnwidth]{img/x3-cobalt-interfaces.png}
-	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
-	\label{fig:cobalt}
-\end{figure}
-
-\bigskip \noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
-Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.5\columnwidth]{img/x3-mercury-interfaces.png}
-	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
-	\label{fig:mercury}
-\end{figure}
-
-\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
+\section{VARAN-Bus}
+
+\chapter{Related Work}\label{cha:related_work}
+
+\chapter{Methodology}\label{cha:methodology}
+
+This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
+
+\begin{comment}
+
+Zu Beginn der Arbeit wurde eine ausführliche Analyse der einzelnen Virtualisierungsmöglichkeiten von Salamander 4 durchgeführt.
+Im Besonderen wurde hier die Virtualisierungsperformance von Ubuntu 22.04, Windows 10 und WSL unter QEMU verglichen.
+
+First, a comprehensive literature review is conducted to understand the current trends and challenges in real-time robot control.
+Based on the literature study, a suitable virtualization platform is selected.
+
+After the selection of the virtualization platform, the robot control platform is implemented.
+This step includes the installation and configuration of Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
+Once the platform has been implemented, the robot is connected via a VARAN bus interface.
+Finally, the platform is evaluated to determine how the integration of real-time functions and efficient communication systems improves the response time and reliability of robot applications.
+
+- Evaluation der Virtualisierungsplattform:
+Ich werde verschiedene Virtualisierungsplattformen wie QEMU, Hyper-V, Virtual Box usw. evaluieren. Dies ist ein wichtiger Schritt, um die beste Plattform für meine Anforderungen zu finden.
+
+- Erstellung und Konfiguration des Systems in der Yocto-Umgebung:
+Ich werde das Yocto-Framework verwenden, um mein Embedded Linux System zu erstellen und zu konfigurieren.
+Yocto bietet viele Tools und Funktionen, die mir bei der Erstellung und Konfiguration meines Systems helfen können.
+
+- Anbindung eines Roboters über eine VARAN-Bus Schnittstelle:
+Ich plane, einen Roboter in mein System zu integrieren.
+Ich werde eine VARAN-Bus Schnittstelle verwenden, um eine schnelle und zuverlässige Kommunikation zwischen dem Roboter und dem Steuerungssystem zu gewährleisten.
+
+\end{comment}
+
+\bigskip \noindent Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
+It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
+Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
+
+\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
+It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
+By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
+
+\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel~\cite{XenomaiXenomai}.
+It enables low-latency and deterministic execution of time-critical tasks.
+Xenomai 3 introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
+A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
+The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
+
+\clearpage
+
+The system configuration is shown in Table~\ref{tab:testbed_configuration}
+
+\begin{table}[H]
+	\centering
+	\caption[System configuration]{System configuration}
+	\label{tab:testbed_configuration}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+	\begin{tabular}{|l|l|}
+	\hline
+	\textbf{CPU} & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
+	\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB \\ \hline
+	\textbf{GPU} & NVIDIA RTX A500 Laptop GPU \\ \hline
+	\textbf{BIOS} & Dell Version 1.12.0 \\ \hline
+	\textbf{OS} & Ubuntu 22.04.4 LTS \\
+	\hline
+	\end{tabular}}
+	\end{table}
+
+\noindent Figure~\ref{fig:lstopo} is the output of the~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices. 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
+	\caption[Hardware topology]{Hardware topology}
+	\label{fig:lstopo}
+\end{figure}
 
@@
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
-It was developed using Yocto, an open-source project that allows customised Linux distributions to be created for embedded systems~\cite{WelcomeYoctoProject}.
+It was developed using Yocto, an open-source project that allows customised Linux distributions to be created for embedded systems.
 Upon generating the necessary files, Yocto provides a QEMU folder with the following components shown in Code~\ref{lst:ls}.
@@
 
-\noindent Comparing these values to those of bare metal in Figure \ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in the statistics.
+\noindent Comparing these values to those of bare metal in Figure~\ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in the statistics.
 A maximum latency of 707.62~\textmu s is not tolerable for the system and needs to be tuned.
@@
 
-\textcolor{red}{After Article, Figure \ref{fig:max_latency_rt}}
+\textcolor{red}{After Article, Figure~\ref{fig:max_latency_rt}}
 
@@
 	\item \textbf{isolcpus=4}: Isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned.
-CPU isolation is explained in detail in section \ref{subsec:cpu_isolation}.
+CPU isolation is explained in detail in section~\ref{subsec:cpu_isolation}.
 	\item \textbf{rcu\_nocbs=4}: The Linux kernel uses a synchronization mechanism called RCU, or Read-Copy-Update.
@@
 Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs.
-Section \ref{subsec:irq_handling} dives deeper into Interrupt Request affinity.
+Section~\ref{subsec:irq_handling} dives deeper into Interrupt Request affinity.
 \end{itemize}
 
-\textcolor{red}{After the BIOS and Kernel configurations, Figure \ref{fig:max_latency_rt_kernelparam}}
+\textcolor{red}{After the BIOS and Kernel configurations, Figure~\ref{fig:max_latency_rt_kernelparam}}
 
@@
 One key aspect of this are real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}.
-In essence, Table \ref{tab:scheduling_priorities} lists minimum and maximum priorities for different scheduling policies in Linux.
+In essence, Table~\ref{tab:scheduling_priorities} lists minimum and maximum priorities for different scheduling policies in Linux.
 
@@
 \subsection{Disable RCU CPU stall warnings}
-As already mentioned in section \ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
+As already mentioned in section~\ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
 An RCU CPU stall in the Linux kernel can occur due to several reasons.
@@
 
-\textcolor{red}{After the Host configurations, Figure \ref{fig:max_latency_rt_kernelparam_host}}
+\textcolor{red}{After the Host configurations, Figure~\ref{fig:max_latency_rt_kernelparam_host}}
 
@@
 \subsection{Set QEMU options for real-time VM}
-QEMU provides several options that can be used to improve the real-time performance of the guest VM. Table \ref{tab:qemu_options} briefly explains these options.
+QEMU provides several options that can be used to improve the real-time performance of the guest VM. Table~\ref{tab:qemu_options} briefly explains these options.
 
@@
 
-\noindent Code \ref{script:qemu_def_tuned} shows the final QEMU script used to start the Salamander 4 virtualization, including these options.
+\noindent Code~\ref{script:qemu_def_tuned} shows the final QEMU script used to start the Salamander 4 virtualization, including these options.
 
@@
 
-	\textcolor{red}{After the QEMU configurations, Figure \ref{fig:max_latency_rt_kernelparam_host_qemu}}
+	\textcolor{red}{After the QEMU configurations, Figure~\ref{fig:max_latency_rt_kernelparam_host_qemu}}
 
@@
 The goal is to understand how closely the performance of the virtualization can match that of the bare metal.
-The experimental setup includes a six-axis mini-robot, illustrated in Figure \ref{fig:mini_robot}.
+The experimental setup includes a six-axis mini-robot, illustrated in Figure~\ref{fig:mini_robot}.
 
@@
 The motor is able to rotate in a range of approximately 180 degrees and its position can be controlled with a high degree of accuracy.
-Each servo motor has three wires that need to be connected as shown in Figure \ref{fig:motor_wires}.
+Each servo motor has three wires that need to be connected as shown in Figure~\ref{fig:motor_wires}.
 
@@
 In this experiment, this PWM signal is generated by the proprietary PW 022 pulse width module of Sigmatek~\cite{DigitalOutputSIGMATEK}.
-Its connector layout is illustrated in Figure \ref{fig:pw022_connectors}.
+Its connector layout is illustrated in Figure~\ref{fig:pw022_connectors}.
 
@@
 For this purpose, two resistors with resistances of 2500 kiloohms and 1000 kiloohms were connected in series.
-The voltage across each resistor is calculated using Ohm's law, which is given by equation \ref{eq:ohms_law} below.
+The voltage across each resistor is calculated using Ohm's law, which is given by equation~\ref{eq:ohms_law} below.
 \begin{equation}
@@
 \noindent Connecting a second servo motor of the mini-robot means repeating the process of reducing the voltage through resistors for the second PWM output of the PW 022 module.
-The connection between the PW 022 module and the motor of the mini-robot is demonstrated in Figure \ref{fig:pw022_minirobot_connection}.
+The connection between the PW 022 module and the motor of the mini-robot is demonstrated in Figure~\ref{fig:pw022_minirobot_connection}.
 
@@
 The PW 022 module is directly mounted on the CPU via the S-DIAS bus and communicates over the hard real-time capable Ethernet VARAN with 100 Mbit/s~\cite{SDIASSIGMATEK}.
-The setup is visible on Figure \ref{fig:hardware_tree}.
+The setup is visible on Figure~\ref{fig:hardware_tree}.
 
@@
 \noindent An additional Varan Connection module VI 021~\cite{InterfacesSplittersSIGMATEK} was required to enable the connection between the PCV 522 module and the PW 022 module that generates the signal to move the mini-robot.
-This setup is depicted in Figure \ref{fig:virt_tree}.
+This setup is depicted in Figure~\ref{fig:virt_tree}.
 

### 2024-08-12 Masterarbeit_Robotik_Pamuk.tex
Application Context 1.2

@@
 
+\hypersetup{
+    colorlinks=true, % Aktiviert farbige Links
+    linkcolor={rgb,255:red,0;green,0;blue,255}, % Interne Links (z.B. Sections)
+    citecolor=orange, % Zitate
+    urlcolor={rgb,255:red,128;green,0;blue,128} % URLs
+}
+
 %
@@
 \lstset{captionpos=b, numberbychapter=false,caption=\lstname,frame=single, numbers=left, stepnumber=1, numbersep=2pt, xleftmargin=15pt, framexleftmargin=15pt, numberstyle=\tiny, tabsize=3, columns=fixed, basicstyle={\fontfamily{pcr}\selectfont\footnotesize}, keywordstyle=\bfseries, commentstyle={\color[gray]{0.33}\itshape}, stringstyle=\color[gray]{0.25}, breaklines, breakatwhitespace, breakautoindent}
-\lstloadlanguages{[ANSI]C, C++, [gnu]make, gnuplot, Matlab}
+\lstloadlanguages{[ANSI]C, C++, python, [gnu]make, gnuplot, Matlab}
 
@@
 \kurzfassung{Erstellung einer Echtzeit-Robotersteuerungsplattform unter Verwendung von Salamander OS, Xenomai, QEMU
-und PCV-521 in der Yocto-Umgebung. Die Plattform basiert auf Salamander OS und nutzt Xenomai für Echtzeit-
+und PCV-522 in der Yocto-Umgebung. Die Plattform basiert auf Salamander OS und nutzt Xenomai für Echtzeit-
 Funktionen. Dazu muss im ersten Schritt die Virtualisierungsplattform evaluiert werden. (QEMU, Hyper-V, Virtual
@@
 The key difference is whether the system is time-critical or not~\cite{canbazPerformanceAnalysisRealtime2022}.
-In a traditional GPOS, such as Windows or Linux-based distributions, a high-priority task cannot interrupt a kernel operation, which makes them not suitable for real-time requirements, as they cannot guarantee deterministic execution times.
-However, in an RTOS, a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation.
-RTOS are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
-Many Linux distributions can function as both GPOS and RTOS with kernel modifications.
+In a traditional GPOS, such as Windows or Linux-based distributions, a high-priority task cannot interrupt a kernel operation, as illustrated in Figure \ref{fig:kernel_generic}.
+This makes them not suitable for real-time requirements, as they cannot guarantee deterministic execution times.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.75\columnwidth]{img/introduction/kernel_generic.png}
+	\caption[Non-preemptible kernel]{Non-preemptible kernel \cite{WhatRealtimeLinuxa}}
+	\label{fig:kernel_generic}
+\end{figure}
+
+\noindent However, in an RTOS, a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation, as shown in Figure \ref{fig:kernel_rt} RTOS are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.75\columnwidth]{img/introduction/kernel_rt.png}
+	\caption[Non-preemptible kernel]{Non-preemptible kernel \cite{WhatRealtimeLinuxa}}
+	\label{fig:kernel_rt}
+\end{figure}
+
+\noindent Many Linux distributions can function as both GPOS and RTOS with kernel modifications.
 
 \section{Real-Time Operating Systems}
-\bigskip \noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.The core component of an RTOS that enables real-time capabilities is the kernel.
+\noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.The core component of an RTOS that enables real-time capabilities is the kernel.
 It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior.
@@
 	\centering
-	\includegraphics[width=0.6\columnwidth]{img/rtos_structure.png}
+	\includegraphics[width=0.50\columnwidth]{img/introduction/rtos_structure.png}
 	\caption[RTOS Structure]{RTOS Structure~\cite{wangRealtimeEmbeddedSystems2017}}
@@
 
-\bigskip \noindent In these real-time operating system, task scheduling is based on so-called priority-based preemptive scheduling~\cite{buttazzoHardRealtimeComputing2024}.
+\noindent In these real-time operating system, task scheduling is based on so-called priority-based preemptive scheduling~\cite{buttazzoHardRealtimeComputing2024}.
 Each task in a software application is assigned a priority.
@@
 
-\clearpage
-
-\section{Application context and boundaries}
-
-This section defines the boundary conditions within which this master's thesis was composed.
-This work was written at SIGMATEK GmbH \& Co KG~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}, a company that researches, develops, produces, and distributes automation technology for industrial machinery and plant engineering.
+\section{Application Context}
+
+This section briefly defines the context and scope within which this master's thesis was composed, highlighting the specific contributions and requirements set by SIGMATEK GmbH \& Co KG~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.
+
+\begin{itemize}
+	\item This work was written at SIGMATEK GmbH \& Co KG, a company that researches, develops, produces, and distributes automation technology for industrial machinery and plant engineering.
 SIGMATEK uses its own custom Linux-based operating system, to be run on their self-manufactured CPUs.
-This operating system will be referred to as Salamander 4 in this work.
-The details of this OS are explained in chapter~\ref{cha:salamander4}.
-Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open source project that helps develop custom Linux-based systems, as discussed in Section~\ref{sec:yocto}.
-Salamander 4 is virtualized through Quick emulator (QEMU), an open-source emulator and virtualizer that allows to run different operating systems on a computer, as described in Section~\ref{sec:qemu}.
-Salamander 4 employs hard real-time with Xenomai 3~\cite{XenomaiXenomai}, a real-time framework for Linux that enables deterministic and low-latency performance, detailed in Section~\ref{sec:xenomai}.
-The goal is to virtualize Salamander 4 and approach the performance of bare metal through real-time performance tunings, as shown in chapter~\ref{cha:real-time_tuning}, focussing on latency requirements set by SIGMATEK.
-
-\section{State of the art}
-
-\section{Problem and task definition}
+This operating system will be referred to as ``Salamander 4'' in this work.
+The details of Salamander 4 are explained in chapter~\ref{cha:salamander4}.
+	\item Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open-source project that helps develop custom Linux-based systems. Yocto is discussed in Section~\ref{sec:yocto}. 
+	\item Salamander 4 is virtualized through Quick emulator (QEMU), an open-source emulator and virtualizer that allows to run different operating systems on a computer QEMU is described in Section~\ref{sec:qemu}.
+	\item Salamander 4 employs hard real-time with Xenomai 3~\cite{XenomaiXenomai}, a real-time framework for Linux that enables deterministic and low-latency performance.
+Xenomai 3 is detailed in Section~\ref{sec:xenomai}.
+	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal through real-time performance tunings, focussing on latency requirements set by SIGMATEK. 
+	These modifications are subject of chapter~\ref{cha:real-time_tuning}.
+	\item SIGMATEK uses the VARAN bus for communication between machines and systems, a real-time Ethernet network that was specially developed for industrial automation.
+Its operating principle is outlined in Section \ref{sec:varan}.
+	\item For the purpose of replacing the physical CPU, SIGMATEK provided a PCI Insert Card module PCV 522, that was plugged into the PC, as explained in Section \ref{sec:setup_experiment_virtualized}.
+This component could then be used to communicate with the Input/Output peripherals, just like the physical CPU could when the I/O modules were attached to it.
+Subsequently, the PCV 522 module of SIGMATEK was connected with the Pulse Width Module PW 022 of SIGMATEK over the VARAN bus and the VI 021 module of SIGMATEK serving as the power supply.
+	\item The robotic application from chapter \ref{cha:robotic_application} was written in Lasal Class 2, an object oriented programming tool of SIGMATEK with client-server technology and graphic representation.
+	\end{itemize}
+
+\section{State of the Art}
+
+\subsection{Virtualization}
+Yocto primarily uses
+
+At the beginning of the work, a detailed analysis of the individual virtualization options of Salamander 4 was carried out.
+In particular, the virtualization performance of Ubuntu 22.04, Windows 10 and WSL under QEMU was compared.
+
+\subsection{Real-time Kernel}
+\subsection{}
+
+\section{Problem and Task Definition}
 
 \section{Objective}
+
+The main objective of this work is to create a real-time robot control platform that integrates Salamander OS, Xenomai, QEMU and PCV-522 in the Yocto environment.
+
+\clearpage
 
@@
 \section{Xenomai}\label{sec:xenomai}
-\noindent Xenomai 3~\cite{XenomaiXenomai} is a real-time framework that offers two paths to real-time performance.
+\noindent Xenomai 3~\cite{XenomaiXenomai} is an open-source real-time framework that offers two paths to real-time performance.
 The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
@@
 	\centering
-	\includegraphics[width=0.5\columnwidth]{img/x3-cobalt-interfaces.png}
+	\includegraphics[width=0.5\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
 	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
@@
 
-\bigskip \noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
+\noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
 Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
@@
 	\centering
-	\includegraphics[width=0.5\columnwidth]{img/x3-mercury-interfaces.png}
+	\includegraphics[width=0.5\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
 	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
@@
 
-\clearpage
-
-The main objective of this work is to create a real-time robot control platform that integrates Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
+\section{VARAN-Bus}\label{sec:varan}
 
@@
 
-\clearpage
 \noindent These modules and the interfaces (indicated by an arrow) between them are shown in Figure~\ref{fig:lasal_cpu}.
@@
 
-\section{VARAN-Bus}
-
 \chapter{Related Work}\label{cha:related_work}
 
+\clearpage
+
 \chapter{Methodology}\label{cha:methodology}
@@
 This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
-
-\begin{comment}
-
-Zu Beginn der Arbeit wurde eine ausführliche Analyse der einzelnen Virtualisierungsmöglichkeiten von Salamander 4 durchgeführt.
-Im Besonderen wurde hier die Virtualisierungsperformance von Ubuntu 22.04, Windows 10 und WSL unter QEMU verglichen.
-
-First, a comprehensive literature review is conducted to understand the current trends and challenges in real-time robot control.
-Based on the literature study, a suitable virtualization platform is selected.
-
-After the selection of the virtualization platform, the robot control platform is implemented.
-This step includes the installation and configuration of Salamander OS, Xenomai, QEMU and PCV-521 in the Yocto environment.
-Once the platform has been implemented, the robot is connected via a VARAN bus interface.
-Finally, the platform is evaluated to determine how the integration of real-time functions and efficient communication systems improves the response time and reliability of robot applications.
-
-- Evaluation der Virtualisierungsplattform:
-Ich werde verschiedene Virtualisierungsplattformen wie QEMU, Hyper-V, Virtual Box usw. evaluieren. Dies ist ein wichtiger Schritt, um die beste Plattform für meine Anforderungen zu finden.
-
-- Erstellung und Konfiguration des Systems in der Yocto-Umgebung:
-Ich werde das Yocto-Framework verwenden, um mein Embedded Linux System zu erstellen und zu konfigurieren.
-Yocto bietet viele Tools und Funktionen, die mir bei der Erstellung und Konfiguration meines Systems helfen können.
-
-- Anbindung eines Roboters über eine VARAN-Bus Schnittstelle:
-Ich plane, einen Roboter in mein System zu integrieren.
-Ich werde eine VARAN-Bus Schnittstelle verwenden, um eine schnelle und zuverlässige Kommunikation zwischen dem Roboter und dem Steuerungssystem zu gewährleisten.
-
-\end{comment}
 
@@
 The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
-
-\clearpage
 
@@
 In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest OS is provided.
-\subsection{CPU affinity and isolation}\label{subsec:cpu_isolation}
+\subsection{CPU affinity and Isolation}\label{subsec:cpu_isolation}
 
@@
 \end{minipage}
-\subsection{KVM entry and KVM exit}\label{subsec:kvm_entry_exit}
+\subsection{KVM Entry and KVM Exit}\label{subsec:kvm_entry_exit}
 \textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
@@
 \end{minipage}
-\subsection{RT-priority} 
+\subsection{RT-Priority} 
 Having a real-time kernel itself is a crucial part of achieving deterministic behavior, but not enough to take full advantage of the real-time capabilities.
@@
 It is also important that there are no other unexpected real time processes running on the system concurrently.
-\subsection{Disable RT throttling}
+\subsection{Disable RT Throttling}
 If a real-time task consumes 100\% of the CPU time, the system may become unresponsive as a whole.
@@
 This reduces the potential for missed deadlines.
-\subsection{Disable timer migration}
+\subsection{Disable Timer Migration}
 Timer migration allows timers to be moved from one CPU to another, which means the kernel can balance load across multiple CPUs.
@@
 This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
-\subsection{Disable RCU CPU stall warnings}
+\subsection{Disable RCU CPU Stall Warnings}
 As already mentioned in section~\ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
@@
 
-\subsection{Tune lapic timer advance}
+\subsection{Tune LAPIC Timer Advance}
 The Local Advanced Programmable Interrupt Controller (LAPIC) is a built-in timer that handles the delivery of interrupts to the CPU.
@@
 
-\subsection{Set QEMU options for real-time VM}
+\subsection{Set QEMU Options for real-time VM}
 QEMU provides several options that can be used to improve the real-time performance of the guest VM. Table~\ref{tab:qemu_options} briefly explains these options.
@@
 
-\section{Other configurations}
+\section{Other Configurations}
 There are other configurations that may be relevant depending on the case.
@@
 
-\chapter{Real-Time Robotic Application}\label{cha:real-time-testing}
+\chapter{Real-Time Robotic Application}\label{cha:robotic_application}
 This chapter compares the latency of the Salamander 4 operating system before and after the real-time performance tunings with the latency of Salamander 4 running on bare metal hardware.
@@
 
-\subsection{Setup of Hardware Salamander 4}
+\section{Setup of Hardware Salamander 4}
 In this version, Salamander 4 runs on the CP 841~\cite{CPUEinheitenSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
@@
 
-\subsection{Setup of QEMU Salamander 4}
+\section{Setup of QEMU Salamander 4}\label{sec:setup_experiment_virtualized}
 In the virtualized setup of the experiment, the VCPU functionality of QEMU is used to connect QEMU with the PWM module.

### 2024-08-12 Masterarbeit_Robotik_Pamuk.tex
related work in red

@@
 At the beginning of the work, a detailed analysis of the individual virtualization options of Salamander 4 was carried out.
+
 In particular, the virtualization performance of Ubuntu 22.04, Windows 10 and WSL under QEMU was compared.
@@
 \subsection{Real-time Kernel}
-\subsection{}
+
+\subsection{Related Work}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Demystifying the Real-Time Linux Scheduling Latency” by Daniel Bristot de Oliveira and colleagues explores the intricacies of Linux as a real-time operating system, specifically focusing on the PREEMPT\_RT variant.
+The authors critique the traditional cyclictest tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
+To address these issues, the paper leverages the Thread Synchronization Model of Linux to derive a set of properties and rules that define the kernel’s behavior from a scheduling perspective12.
+This model helps in establishing a sound bound to the scheduling latency by considering all possible sources of delays in synchronization events34.
+The authors also introduce an efficient tracing method to observe kernel events with minimal overhead, resulting in a practical tool for deriving reliable scheduling latency bounds5.
+The paper concludes with an experimental analysis comparing cyclictest and the proposed method, demonstrating that the latter can find sound bounds faster and with acceptable overheads6.
+This work contributes significantly to the understanding and improvement of real-time performance in Linux systems.}
+
+\bigskip \noindent \textcolor{red}{The white paper, “Real-Time Performance Tuning Best Practice Guidelines for KVM-Based Virtual Machines,” provides comprehensive guidelines for optimizing real-time performance in KVM-based virtual machines on Intel Architecture (IA) platforms1.
+It begins by explaining the importance of real-time systems, which must respond to events within strict time constraints, and the relevance of tuning these systems for applications running in virtual machines (VMs), particularly in the manufacturing sector2.
+The document details the testbed configuration used for performance evaluation, including an industrial PC based on the Intel Whiskey Lake-U platform with specific BIOS, kernel, and host configurations3.
+Key BIOS settings such as disabling Hyper-Threading and Intel SpeedStep, and enabling Intel VMX and VT-d, are recommended to enhance real-time performance.
+Kernel configurations include using a real-time kernel, setting specific kernel parameters, and isolating CPUs for real-time workloads.
+Host configurations involve setting CPU affinity for system services, cache isolation, and disabling certain services like irqbalance and thermald.
+The paper also outlines QEMU-KVM configurations, such as tuning lapic timer advance and setting CPU affinity for QEMU threads.
+For the guest OS in the real-time VM, similar configurations to the host are recommended, with additional tuning for sched\_rt\_runtime\_us.
+The performance evaluation section describes a rigorous testing process, including running heavy workloads alongside the real-time VM, which demonstrated significantly enhanced real-time performance with latency in the range of tens of microseconds4.
+The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads5.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System” explores the integration of real-time operating systems (RTOS) with general-purpose operating systems (GPOS) using Kernel-Based Virtual Machine (KVM) technology.
+The authors, Ruhui Ma, Fanfu Zhou, Erzhou Zhu, and Haibing Guan, propose an embedded real-time virtualization architecture that combines VxWorks (RTOS) and Linux (GPOS) on a multi-core platform1.
+This architecture aims to address issues such as responsiveness, system performance, and CPU resource utilization.
+The paper details several performance tuning methods, including CPU shielding, prioritization, and a novel co-scheduling algorithm, to achieve sub-millisecond interrupt response latency on the guest VxWorks2.
+Experimental results demonstrate that these tuning methods significantly improve real-time responsiveness and system performance.
+The study concludes that with careful system setups, such as applying RT patches and real-time tunings on the host Linux, the proposed architecture is suitable for certain embedded real-time scenarios34.
+The work is supported by various Chinese national foundations and highlights the potential of virtualization technology in enhancing the capabilities of embedded systems.}
+
+\bigskip \noindent \textcolor{red}{The paper "‘Real Time’ vs.
+‘Real Fast’: How to Choose?"1 by Paul E.
+McKenney from IBM Linux Technology Center explores the distinction between real-time and real-fast computing in the context of Linux systems.
+It provides a comprehensive analysis to help developers make informed design choices between achieving quick start times (real-time) and fast completion times (real-fast)2.
+The paper uses two example applications: an embedded fuel-injection system and a Linux kernel build, to illustrate the differences3.
+The fuel-injection scenario, which requires precise timing, benefits from real-time Linux, while the kernel build, which prioritizes throughput, performs better on non-real-time Linux.
+The paper discusses factors influencing the choice, such as workload duration, CPU utilization, and memory usage.
+It also addresses the overheads associated with real-time Linux, including increased locking and interrupt handling costs45.
+The conclusion emphasizes using the right tool for the job and suggests ongoing work to reduce the performance penalty of real-time Linux, potentially making it suitable for a broader range of workloads6.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Real-Time Performance Analysis in Linux-Based Robotic Systems” by Hobin Yoon, Jungmoo Song, and Jamee Lee from Samsung Advanced Institute of Technology explores the feasibility of using real-time flavored Linux for controlling complex humanoid robots1.
+The authors emphasize the importance of precise control periods for mobile and humanoid robots, which require timely responses to external stimuli to maintain balance and smooth operation.
+The study introduces the adoption of EtherCAT technology for real-time communication among distributed devices, highlighting its flexible topology, simple configuration, and cost-effectiveness2.
+The paper details the design and implementation of the Robot Control Kernel Subsystem (RCKS), which integrates EtherCAT master and slave devices, and discusses various real-time performance tunings applied to the Linux kernel to achieve deterministic timing.
+Experimental results demonstrate that with appropriate real-time tunings, including CPU shielding, memory locking, and spinning nanosleep, Linux can meet the stringent timing requirements of robotic systems, achieving control periods within acceptable limits even under heavy system load.
+The authors conclude that Linux, with its real-time enhancements, is a viable solution for deploying in advanced robotic systems, offering both real-time capabilities and the benefits of a general-purpose OS3.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Performance Evaluation of Xenomai 3” by Ching-Chun (Jim) Huang, Chan-Hsiang Lin, and Che-Kang Wu, presents an in-depth analysis of Xenomai 3, a real-time framework for Linux.
+Xenomai 3 operates either as a dual-kernel system alongside Linux or natively over mainline Linux kernels, enhanced by PREEMPT RT for stricter response times1.
+The study evaluates Xenomai 3’s performance, particularly in dual-kernel configuration, using ARM Cortex-A series benchmarks.
+The authors highlight Xenomai 3’s optimizations in RTOS API emulation, thread-to-thread communications, and reduced overhead.
+The paper compares Xenomai 3 with its predecessor, Xenomai 2, and the PREEMPT RT-enhanced Linux kernel, demonstrating superior maximum response times to external interrupts in dual-kernel setups.
+The study employs BeagleBone Black hardware and various kernel configurations to measure latencies under idle and CPU-stressed conditions2.
+Results indicate that Xenomai 3 generally outperforms Xenomai 2 and PREEMPT RT, except for a slight degradation in timer-IRQ latency under stress.
+The paper concludes that Xenomai 3 offers significant improvements in real-time performance and flexibility, making it a robust choice for real-time applications.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Performance Analysis Towards a KVM-Based Embedded Real-Time Virtualization Architecture” explores the challenges and solutions associated with integrating real-time operating systems (RTOS) with general-purpose operating systems (GPOS) in embedded systems using virtualization technology.
+The authors propose an architecture based on the Kernel-Based Virtual Machine (KVM) that combines VxWorks (an RTOS) and Linux (a GPOS) on a multi-core platform12.
+The primary focus is on evaluating how KVM affects the interrupt-response times of VxWorks when running as a guest OS3.
+The study highlights the shift from single-core to multi-core processors in embedded systems and the need for legacy RTOS to adapt to multi-core environments4.
+The paper discusses hardware-assisted virtualization, particularly Intel’s VT technology, and the implementation of KVM as a hypervisor within the Linux kernel.
+Through experimental setups and performance tuning methods like prioritization and CPU shielding, the authors demonstrate that sub-millisecond interrupt response latency can be achieved, making the proposed architecture suitable for certain embedded real-time applications5.
+The results indicate that with careful system configurations, KVM can effectively support real-time performance requirements in embedded systems.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Performance Analysis Towards a KVM-Based Embedded Real-Time Virtualization Architecture” explores the challenges and solutions associated with integrating real-time operating systems (RTOS) with general-purpose operating systems (GPOS) on multi-core processors in embedded systems.
+The authors propose an architecture based on the Kernel-Based Virtual Machine (KVM) that combines VxWorks (an RTOS) and Linux (a GPOS) to leverage the benefits of both systems.
+The study focuses on evaluating the impact of KVM on the interrupt-response times of VxWorks when running as a guest OS.
+By applying real-time performance tuning methods on the host Linux, the authors demonstrate that sub-millisecond interrupt response latency can be achieved, making the architecture suitable for certain embedded real-time scenarios.
+The paper also discusses hardware-assisted virtualization, the role of KVM as a hypervisor, and the experimental setup used to measure the performance impacts.
+The results indicate that with appropriate system configurations, such as prioritization and CPU shielding, the proposed architecture can effectively meet real-time performance requirements.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Real-time capabilities in the standard Linux Kernel: How to enable and use them?” explores the transformation of the Linux operating system into a real-time system using the PREEMPT-RT patches1.
+Originally designed as a general-purpose OS, Linux has gained popularity in real-time applications due to its low cost and open-source nature2.
+The authors discuss the importance of real-time systems, which must meet strict temporal requirements to ensure system correctness.
+They differentiate between hard and soft real-time systems, emphasizing that Linux, with the PREEMPT-RT patch, can achieve soft real-time performance3.
+The PREEMPT-RT patch, maintained by Ingo Molnar and Thomas Gleixner, allows almost the entire kernel to be preempted, enhancing its real-time capabilities4.
+Key kernel configurations and runtime precautions are highlighted to maintain real-time performance.
+The paper concludes that while Linux can achieve real-time performance, it is still not as deterministic as commercial real-time operating systems like QNX and Windows CE5.
+However, with proper configuration, Linux can be a viable option for real-time applications.}
+
+\bigskip \noindent \textcolor{red}{This paper presents a modular platform for industrial automation that leverages open-source components to run both non-critical activities and real-time control tasks on the same hardware, thereby reducing overall hardware costs.
+The platform uses a hypervisor to provide isolation, allowing concurrent execution of real-time control code and design tools.
+The authors detail the software architecture, including the hypervisor and Real-Time Operating System (RTOS), and discuss the necessary extensions for practical application.
+They address the challenge of interference on shared hardware resources, which can degrade real-time performance.
+The paper also includes experimental results demonstrating the platform’s effectiveness and performance.
+The authors highlight the benefits of using open-source software to avoid expensive commercial solutions and emphasize the importance of multi-core processors and hardware-assisted virtualization in industrial automation.
+The work is supported by the European Commission through the I-MECH ECSEL project.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\bigskip \noindent \textcolor{red}{Shielded Processors: Guaranteeing Sub-millisecond Response in Standard Linux by Steve Brosky and Steve Rotolo of Concurrent Computer Corporation explores the implementation and benefits of shielded processors in RedHawk Linux for real-time applications.
+The paper highlights the limitations of standard Linux in providing deterministic performance due to networking and graphics activities.
+Shielded CPUs, a model developed over a decade of experience, offload operating system overhead to a subset of CPUs, ensuring more deterministic execution and interrupt response times.
+The authors describe the modifications made to the Linux kernel, including CPU affinity settings and the disabling of local timer interrupts, to create a shielded environment.
+Benchmark tests demonstrate significant improvements in execution determinism and interrupt response times, with jitter reduced to 1.87\% on shielded CPUs compared to 26.17\% on standard Linux.
+The paper concludes that shielded processors can guarantee interrupt response times of less than 30 microseconds, even under heavy system load, making them suitable for applications with hard real-time deadlines.
+The work underscores the potential of using commodity operating systems like Linux for high-performance real-time applications by leveraging shielded processors.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Real-Time Performance and Response Latency Measurements of Linux Kernels on Single-Board Computers” by George K.
+Adam from the CSLab Computer Systems Laboratory, University of Thessaly, investigates the real-time performance of Linux kernels with the PREEMPT\_RT patch on embedded devices like the BeagleBoard and Raspberry Pi1.
+The research focuses on measuring latencies in real-time applications at both user and kernel space using newly developed software modules tailored for these experiments2.
+The study highlights that the PREEMPT\_RT patch significantly improves the real-time performance of Linux kernels, reducing worst-case latencies to around 160 µs, making these systems more suitable for real-time applications3.
+The methodology involves response and periodic task models, and the results suggest that the proposed measurement approach can be applied to other Linux-based platforms4.
+The paper also discusses the challenges and solutions encountered during the experiments, emphasizing the importance of real-time capabilities in embedded systems for applications ranging from consumer electronics to industrial automation.}
+
+\bigskip \noindent \textcolor{red}{Asymmetric Scheduling and Load Balancing for Real-Time on Linux SMP
+Authors: Eric Piel, Philippe Marquet, Julien Soula, and Jean-Luc Dekeyser12
+The paper presents ARTiS, a real-time extension of the GNU/Linux scheduler designed for SMP (Symmetric Multi-Processors) systems3.
+ARTiS aims to guarantee the preemption of processors for real-time tasks by assigning specific processors to these tasks4.
+This system ensures low latency by migrating non-preemptible tasks away from real-time processors5.
+ARTiS also incorporates load-balancing strategies to fully utilize SMP systems without wasting resources.
+The implementation of ARTiS involves modifications to the Linux scheduler, resulting in significant improvements in latency compared to the standard Linux scheduler6.
+The paper details the performance evaluation of ARTiS, highlighting its ability to maintain low interrupt latency even under high system load, making it suitable for hard real-time applications.
+The authors propose further enhancements, including the integration of real-time scheduling policies like EDF (Earliest Deadline First) and RM (Rate Monotonic)7.}
+
+\bigskip \noindent \textcolor{red}{Title: A Comprehensive Study of Kernel (Issues and Concepts) in Different Operating Systems12
+Authors: Hayfaa Subhi Malallah, Subhi R. M. Zeebaree, Rizgar R. Zebari, Mohammed A. M. Sadeeq, Zainab Salih Ageed, Ibrahim Mahmood Ibrahim, Hajar Maseeh Yasin, Karwan Jameel Merceedi
+Summary: This paper provides an in-depth comparative analysis of various operating systems (OS) focusing on their kernels, which are the core components managing system resources and providing essential services.
+The study examines the strengths and limitations of popular OS such as iOS, Android, Mac, Windows, and Linux3.
+It highlights that Linux, Android, and Windows 10 are particularly noted for their stability, compatibility, and reliability4.
+The paper discusses the evolution of OS kernels, including microkernels, monolithic kernels, and hybrid kernels, and their respective advantages and challenges.
+It also addresses modern developments in OS technology, such as multicore processors, cloud computing, and the Internet of Things (IoT), and their impact on kernel design and performance.
+The authors emphasize the importance of security in kernel development, noting that vulnerabilities can lead to significant system compromises.
+The paper concludes with a discussion on the future directions of OS kernel research and development, advocating for continuous improvement in performance, security, and adaptability to emerging technologies.}
+
+\bigskip \noindent \textcolor{red}{Title: Testing the limits of general-purpose hypervisors for real-time control systems
+Authors: Rui Queiroz, Tiago J. Cruz, Paulo Simões (University of Coimbra)
+Summary: This paper explores the feasibility of using general-purpose off-the-shelf hypervisors to virtualize real-time control systems within the context of Industry 4.0.
+The authors address the need for flexibility, security, and resilience in automation infrastructures while maintaining real-time requirements.
+They discuss the potential benefits of virtualizing servers and control devices in Industrial and Automation Control Systems (IACS), but also highlight the challenges due to the focus of general-purpose hypervisors on maximizing throughput, often at the expense of determinism and increased latency.
+The paper presents an empirical evaluation of various hypervisors, concluding that while some are capable of handling typical real-time workloads, this capability cannot be generalized to all types of real-time systems.
+The authors provide a comprehensive analysis of cyber-physical systems (CPS), control devices, and real-time systems, including their latency and jitter requirements.
+They also delve into the specifics of real-time operating systems (RTOS) and mixed-criticality systems, and review different virtualization techniques, comparing commercial off-the-shelf hypervisors with real-time hypervisors.
+The study emphasizes the importance of ensuring both spatial and temporal isolation in virtualized environments to maintain predictable real-time performance.
+It also discusses the role of scheduling policies in real-time applications and the challenges of I/O sharing between multiple virtual machines.
+The authors conclude that while general-purpose hypervisors show promise, further research and development are needed to fully realize their potential in real-time control systems.}
+
+\bigskip \noindent \textcolor{red}{“Testing the limits of general-purpose hypervisors for real-time control systems” by Rui Queiroz, Tiago J. Cruz, and Paulo Simões from the University of Coimbra:
+With the advent of Industry 4.0, there is a growing need for flexibility, security, and resilience in industrial automation infrastructures while maintaining real-time requirements.
+This paper explores the feasibility of using general-purpose off-the-shelf hypervisors to virtualize cyber-physical systems’ servers and control devices.
+The authors, Rui Queiroz, Tiago J.
+Cruz, and Paulo Simões, present an empirical evaluation of these hypervisors, focusing on their ability to handle real-time workloads.
+The study highlights the challenges posed by the need for determinism and low latency, which are often compromised by hypervisors designed to maximize throughput.
+The paper concludes that while some hypervisors can manage typical real-time workloads, this capability is not universal across all types of real-time systems.
+The research underscores the importance of understanding the limits and capabilities of conventional hypervisors to ensure they meet the specific requirements of industrial automation control systems.}
+
+\bigskip \noindent \textcolor{red}{Title: A State-of-the-Art Survey on Real-Time Issues in Embedded Systems Virtualization
+Authors: Zonghua Gu and Qingling Zhao
+Summary: This paper provides a comprehensive survey on real-time issues in virtualization for embedded systems, authored by Zonghua Gu and Qingling Zhao.
+Virtualization, widely accepted in server and cloud computing, has been increasingly applied to embedded systems with stringent timing constraints.
+The paper discusses various virtualization systems, including KVM, Xen, and L4, and their applications in embedded systems such as avionics, industrial automation, and mobile phones.
+It highlights the importance of real-time performance, security, and dependability in these systems.
+The survey covers different types of virtualization, including Type-1 (bare-metal) and Type-2 (hosted) virtualization, as well as full and para-virtualization.
+It also addresses specific real-time issues like scheduling, partitioning, and the Lock-Holder Preemption problem.
+The paper concludes with a discussion on the future directions and challenges in the field of real-time virtualization for embedded systems.}
+
+\bigskip \noindent \textcolor{red}{Title: How Real (Time) Are Virtual PLCs?1
+Authors: Diogenes Javier Perez, Josef Waltl, Laurin Prenzel, Sebastian Steinhorst
+This paper explores the concept of virtual programmable logic controllers (vPLCs) as a solution to the limitations of traditional hardware-based PLCs in industrial automation.
+The authors argue that hardware PLCs are costly to maintain and lack the flexibility and scalability needed for modern applications, especially those involving artificial intelligence and machine learning2.
+The proposed vPLC operates on a commercial off-the-shelf (COTS) server using a real-time hypervisor, which allows it to manage multiple virtual machines and handle resource-intensive tasks efficiently34.
+The paper details the architecture of the vPLC, which includes a real-time operating system and a hypervisor to ensure low latency and deterministic behavior.
+The authors conducted various tests to compare the performance of vPLCs with traditional SoftPLCs, focusing on metrics such as cycle time, jitter, and response time.
+The results indicate that vPLCs can achieve performance on par with traditional PLCs for applications requiring response times below 10 milliseconds5.
+The study concludes that while further improvements are needed, vPLCs offer significant advantages in terms of flexibility, scalability, and cost-efficiency for future industrial systems6.}
+
+\bigskip \noindent \textcolor{red}{Impact of Modern Virtualization Methods on Timing Precision and Performance of High-Speed Applications
+Authors: Veronika Kirova, Kirill Karpov, Eduard Siemens, Irina Zander, Oksana Vasylenko, Dmitry Kachan, Sergii Maksymov
+This research investigates the precision of timing methods and their efficiency in various virtual environments, focusing on their impact on high-speed network applications.
+The study examines how timer hardware is shared among CPU- and I/O-bound tasks on both virtualized and bare operating systems.
+By replacing timing methods within an application for estimating available path bandwidth, the authors analyze the impact on estimation accuracy.
+The findings highlight that timer overhead and precision are crucial for high-performance network applications, with low-precision timing methods leading to degradation in virtual environments.
+The paper confirms that using precise timing operations and AvB estimation methods can overcome inefficiencies in standard time-related operations.
+The research evaluates five different virtual environments to determine the most optimal for high-speed network applications, providing a comprehensive analysis of the impact of virtualization on timing precision and performance.}
+
+\bigskip \noindent \textcolor{red}{“Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane:
+The paper “Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane provides an in-depth examination of Real Time Operating Systems (RTOS) and their critical role in embedded systems, particularly in telecommunication applications such as telephony, navigation, and military signaling systems.
+RTOS is essential for applications that require strict deadline constraints and the ability to handle multiple tasks simultaneously1.
+The paper highlights the key features of RTOS, including timing considerations, task prioritization, and communication and synchronization between tasks2.
+The document delves into various design aspects necessary for developing an RTOS, such as task management, multitasking, context switching, task priority, and kernel functions3.
+It explains how tasks are created, prioritized, and managed within an RTOS, emphasizing the importance of efficient CPU utilization and the prevention of data corruption through mutual exclusion.
+The paper also discusses the implementation of RTOS, covering topics like programming language selection, task stack management, and the initialization process.
+Furthermore, the paper presents a practical example of designing and implementing an RTOS for a Philips 8052 microcontroller, demonstrating the application of theoretical concepts in real-world scenarios.
+The successful implementation of applications such as a column controller, digital calculator, and dancing LEDs showcases the versatility and effectiveness of RTOS in managing complex embedded systems4.
+Overall, the paper provides a comprehensive overview of RTOS, offering valuable insights into its architecture, design, and implementation, making it a significant contribution to the field of embedded systems5.}
+
+\bigskip \noindent \textcolor{red}{The paper “Challenges in Real-Time Virtualization and Predictable Cloud Computing” explores the integration of real-time systems with cloud computing and virtualization technologies.
+The authors, Marisol García-Valls, Tommaso Cucinotta, and Chenyang Lu, identify the significant technical challenges in supporting real-time applications within cloud environments1.
+They highlight that while cloud computing offers benefits such as reduced operational costs, server consolidation, and elastic resource provisioning, it struggles to meet the demands of soft real-time applications like online video streaming, cloud-based gaming, and telecommunication management2.
+These applications require predictable performance in open, shared, and virtualized environments3.
+The paper discusses the evolution of cloud computing, emphasizing the shift from local computing resources to distributed cloud-based models.
+It outlines the benefits of virtualization, including functional execution isolation, easier management, and the coexistence of legacy and new applications4.
+However, the authors note that the multi-tenant nature of cloud computing introduces challenges in providing stable and predictable performance due to resource sharing among multiple virtual machines (VMs).
+The authors delve into various virtualization techniques, such as full virtualization, hardware-assisted virtualization, para-virtualization, OS-level virtualization, and application-level virtualization, each with its performance characteristics and suitability for real-time applications.
+They also address the importance of real-time scheduling and resource management in virtualized environments, highlighting the need for hierarchical scheduling and the role of real-time hypervisors.
+The paper concludes by presenting potential future research directions to enhance the predictability and performance of cloud-based real-time applications, emphasizing the need for improved resource management frameworks and real-time virtualization techniques.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “VM-Based Real-Time Services for Automotive Control Applications” is authored by Alejandro Masrur, Sebastian Drossler, Thomas Pfeuffer, and Samarjit Chakraborty.
+It explores the use of Virtual Machines (VMs) to provide real-time services in automotive control applications12.
+The authors utilize the Xen hypervisor to design a real-time control loop on top of a virtualization layer3.
+They first analyze a typical Xen configuration and identify issues when used for real-time applications4.
+The paper highlights that the SEDF (Simple Earliest Deadline First) scheduler in Xen can be improved with minimal modifications to enhance worst-case performance5.
+To further reduce latency and jitter, the authors propose a new scheduler for Xen that prioritizes real-time VMs using a fixed-priority policy6.
+This approach ensures that timing constraints in automotive control applications are met7.
+The paper includes extensive experiments to illustrate the effectiveness of the proposed solutions.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\bigskip \noindent \textcolor{red}{Title: Hard Real Time Linux* Using Xenomai* on Intel® Multi-Core Processors12
+Author: Amarpreet Singh Ugal, BIOS/Firmware Engineer, Intel Corporation
+Summary: This document explores the implementation of hard real-time capabilities in Linux using Xenomai on Intel® multi-core processors.
+Traditional Linux is not a hard real-time operating system because it cannot guarantee that tasks will meet strict deadlines34.
+To address this, the document discusses the use of a high-priority real-time kernel that runs between the hardware and standard Linux.
+This real-time kernel executes real-time tasks and suspends normal Linux processes during this period56.
+The real-time scheduler treats the standard Linux kernel as an idle task, allowing real-time tasks to preempt normal processes78.
+Xenomai, which uses an interrupt pipeline from the ADEOS project, is highlighted as a solution for managing interrupts and ensuring minimal latency and jitter.
+The document also covers the installation and configuration of Xenomai on CentOS 5.2, including kernel patching and testing with various tools like switchbench, switchtest, cyclictest, clocktest, and latency.
+The detailed steps and configurations provided aim to help developers achieve hard real-time performance on Intel® Core™ i7 multi-core processors.}
+
+\bigskip \noindent \textcolor{red}{Title: Performance Assessment of Linux Kernels with PREEMPT\_RT on ARM-Based Embedded Devices
+Authors: George K. Adam, Nikos Petrellis, Lambros T. Doulos
+Summary: This research investigates the real-time performance of Linux kernels and distributions with the PREEMPT\_RT real-time patch on ARM-based embedded devices, specifically BeagleBoard and Raspberry Pi.
+The study aims to provide novel insights into Linux real-time performance by evaluating various metrics such as response latency, periodic task execution, and maximum sustained frequency using heuristic methods and the cyclictest benchmark tool.
+The authors introduce a new response task model and develop specific software measurement modules to assess the performance.
+The results demonstrate that the PREEMPT\_RT patch significantly improves real-time performance compared to default Linux kernels, particularly in reducing worst-case latencies.
+This makes ARM-based devices running Linux with PREEMPT\_RT more suitable for time-sensitive embedded control systems.
+The research also highlights the growing use of Linux in embedded systems for real-time applications, driven by its open-source nature, performance, and adaptability to various hardware systems.
+The methodology and findings of this study could be applied to other Linux-based real-time platforms, contributing to the broader field of real-time systems research.}
+
+\bigskip \noindent \textcolor{red}{Title: Real-Time I/O System for Many-core Embedded Systems 1 Author: Zhe Jiang
+In this thesis, Zhe Jiang addresses the critical need for time predictability in modern real-time embedded systems, particularly focusing on I/O operations.
+The proposed solution, BlueIO, is a hardware-implemented real-time I/O virtualization system designed to meet the requirements of predictability, timing accuracy, enhanced performance, scalability, parallel access, and isolation simultaneously2.
+BlueIO integrates key functionalities of I/O virtualization and low-layer I/O drivers through the Virtualized Complicated Device Controller (VCDC) and a clock cycle-level timing-accurate I/O controller, the GPIO Command Processor (GPIOCP)3.
+This system supports abstract virtualized access to I/O devices from the software domain, providing isolation and concurrent access to I/O operations, thereby improving performance45.
+The thesis also introduces BlueVisor, a scalable real-time hardware hypervisor built upon BlueIO, enabling predictable virtualization on CPU, memory, and I/O, along with fast interrupt handling and inter-virtual machine communication67.
+The work demonstrates significant improvements in I/O performance and low running costs across different operating systems, with a hardware consumption analysis showing linear scalability with the number of CPUs and I/O devices.}
+
+\bigskip \noindent \textcolor{red}{Title: Real-Time Operating Systems1
+Author: Jiacun Wang, Monmouth University
+Summary: This chapter provides an in-depth overview of Real-Time Operating Systems (RTOS), which are crucial for embedded systems requiring timely and predictable responses.
+The chapter begins by outlining the main functions of general-purpose operating systems, such as process management, memory management, interrupts management, multitasking, file system management, and I/O management.
+It then delves into the specific characteristics of RTOS kernels, emphasizing the need for predictable timing behavior, priority scheduling, intertask communication, and resource sharing.
+The chapter also discusses clocks and timers, asynchronous I/O, and memory locking as essential features of RTOS.
+Finally, it provides examples of widely used RTOS products, including LynxOS, OSE, QNX, VxWorks, and Windows Embedded Compact, highlighting their unique features and applications in various industries such as aerospace, automotive, and telecommunications.}
+
+\bigskip \noindent \textcolor{red}{Title: Real-Time Scheduling: From Hard to Soft Real-Time Systems1
+Authors: Giuseppe Lipari (Université de Lille) and Luigi Palopoli (Università di Trento)
+Summary: This paper explores the evolution of real-time systems, traditionally divided into hard real-time and soft real-time categoriesHard real-time systems are critical, where missing a deadline can lead to catastrophic consequences, such as in automotive brake systems2.
+Soft real-time systems, on the other hand, aim to optimize the quality of service, like multimedia players that can tolerate occasional delays without severe impact.
+The authors argue that the boundary between these two categories is becoming increasingly blurred.
+They introduce the concept of time-predictability and criticality to understand the origins of real-time deadlines.
+The paper presents the resource reservation framework as a method for scheduling both hard and soft real-time tasks.
+This framework enhances system robustness and resource utilization by isolating the temporal behavior of tasks.
+The paper also discusses the Constant Bandwidth Server (CBS) algorithm, which integrates with the EDF scheduler to manage execution budgets and periods, ensuring temporal isolation3.
+The authors highlight the importance of designing systems that remain consistent even when deadlines are missed, emphasizing the need for robust software development practices.
+In conclusion, the paper demonstrates how resource reservation techniques can be applied to classical control systems, improving performance and robustness4.
+The authors provide a comprehensive overview of existing techniques and tools for soft real-time scheduling and analysis, making a strong case for the integration of these methods in modern real-time systems5.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Performance analysis of real-time and general-purpose operating systems for path planning of the multi-robot systems” by Seçkin Canbaz and Gökhan Erdemir explores the performance differences between real-time operating systems (RTOS) and general-purpose operating systems (GPOS) in the context of multi-robot systems.
+The study focuses on two Linux distributions, Ubuntu and Pardus, and evaluates their effectiveness as both GPOS and RTOS for path planning tasks12.
+The authors highlight the critical distinction between GPOS and RTOS, emphasizing that RTOS can preempt low-priority tasks with high-priority ones, even during kernel calls, which is not possible in GPOS.
+The research involves using the turtlesim simulation environment to perform path tracking tasks with robot groups of varying sizes.
+The experiments compare CPU usage and processing times across the different operating systems.
+Results indicate that Ubuntu GPOS generally outperforms other configurations in terms of lower CPU usage and shorter processing times3.
+However, RTOS configurations, while slower, offer higher system security and reduced error margins due to their strict time management.
+The study concludes that while GPOS may be preferable for performance-centric applications, RTOS is essential for tasks with high time dependency.
+The authors suggest that the choice between GPOS and RTOS should consider factors such as potential failure conditions, process security, and processing time requirements.
+This research provides valuable insights into the suitability of different operating systems for multi-robot applications, contributing to the broader field of robotics and autonomous systems.}
+
+\bigskip \noindent \textcolor{red}{Title: The Real-Time Linux Kernel: A Survey on PREEMPT\_RT12
+Authors: Federico Reghenzani, Giuseppe Massari, and William Fornaciari from Politecnico di Milano3
+Summary: This survey explores the evolution and current state of the PREEMPT\_RT patch for the Linux kernel, which aims to enhance the real-time capabilities of Linux.
+The authors discuss the increasing demand for real-time applications and the challenges posed by using Commercial-Off-The-Shelf (COTS) hardware in such environments.
+The PREEMPT\_RT patch modifies the existing Linux kernel to improve predictability and reduce latencies, making it suitable for real-time applications.
+The survey covers various approaches to real-time Linux, including cokernel approaches like RTLinux, Xenomai, and RTAI, and compares them with the PREEMPT\_RT patch4.
+The authors highlight the benefits of using Linux in real-time systems, such as the availability of rich hardware support and a well-established programming environment.
+They also discuss the challenges of integrating real-time capabilities into a general-purpose operating system and the ongoing efforts to address these issues.
+The survey concludes with an overview of industrial and scientific use cases that have benefited from the PREEMPT\_RT patch, emphasizing its significance in fields like aerospace, automotive, and high-performance computing.}
 
@@
 \end{figure}
-
-\clearpage
-
-\chapter{Related Work}\label{cha:related_work}
 

### 2024-08-13 Masterarbeit_Robotik_Pamuk.tex
Related Work 1.3 First 3

@@
 
-\section{State of the Art}
-
-\subsection{Virtualization}
+\section{Related Work and State of the Art}
+
+\begin{comment}
 Yocto primarily uses
@@
 In particular, the virtualization performance of Ubuntu 22.04, Windows 10 and WSL under QEMU was compared.
-
-\subsection{Real-time Kernel}
-
-\subsection{Related Work}
-
-\bigskip \noindent \textcolor{red}{The paper titled “Demystifying the Real-Time Linux Scheduling Latency” by Daniel Bristot de Oliveira and colleagues explores the intricacies of Linux as a real-time operating system, specifically focusing on the PREEMPT\_RT variant.
+\end{comment}
+
+This master’s thesis builds upon a number of previous studies in the field of virtualization of real-time systems and the inherent latency.
+In this section, the most influential studies are presented and their relevance to the work is addressed.
+
+\bigskip \noindent \citeauthor{perneelRealtimeCapabilitiesStandard2015} \cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general purpose operating system without real-time applications in mind.
+They talk about the evolution of real-time behavior in the standard Linux kernel and state that it has recently gained popularity among the real-time community, largely due to its open-source nature and stability.
+The authors explain how to transform Linux into a real-time operating system by implementing the PREEMPT-RT patches.
+The paper also emphasizes some kernel configurations that ensure that the system's real-time behavior is preserved during runtime.
+Even though these modifications to the kernel enable soft real-time performance, the authors underline the fundamental rule in real-time software, that latency improvements have a negative impact on throughput and kernel performance.
+This is due to the added overhead of the \texttt{CONFIG\_PREEMPT\_RT} option.
+They concluded with their Linux build and a testing application that Linux can be a viable option for real-time applications, if it is correctly configured.
+
+\bigskip \noindent In the course of writing this thesis, the white paper "Real-Time Performance Tuning Best Practice Guidelines for KVM-Based Virtual Machines" \cite{RealTimePerformanceTuning2022} by Intel has been instrumental.
+Given that the working computer used for this work is equipped with an Intel processor, the guidelines presented in this document were particularly relevant and valuable for optimizing real-time performance in KVM-based virtual machines.
+Chapter \ref{cha:real-time_tuning} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
+The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads.
+
+\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased} \cite{maPerformanceTuningKVMbased} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
+When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS.
+Once the interrupt is delivered to the guest RTOS, it incurs additional latency since there is also the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started.
+These stages are the primary sources of latencies caused by KVM and depicted in Figure \ref{fig:six_stages_kvm}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/six_stages_kvm.png}
+	\caption[Six stages for the delivery of the interrupt to the guest]{Six stages for the delivery of the interrupt to the guest, followed by the guest IRT \cite{maPerformanceTuningKVMbased}}
+	\label{fig:six_stages_kvm}
+\end{figure}
+
+Each of these stages is breifly described below:
+
+\begin{enumerate}
+    \item \textbf{Guest Code}: The guest OS is executing its normal operations.
+    \item \textbf{Interrupt Happens}
+    \item \textbf{VM Exit}: The host processes the VM exit, which involves handling the interrupt.
+    \item \textbf{After VM Exit}: The host performs necessary operations after the VM exit.
+    \item \textbf{Host ISR}: The host executes its ISR to handle the interrupt.
+    \item \textbf{Host Scheduling}: The host schedules the guest VM to resume execution.
+    \item \textbf{Before VM Entry}: The host prepares to re-enter the guest VM.
+    \item \textbf{VM Entry}: The host completes the process of re-entering the guest VM.
+    \item \textbf{Guest IRT}: The guest OS starts executing the ISR to handle the interrupt.
+\end{enumerate}
+
+\begin{comment}
+and then apply real-time tunings on the host Linux and analyze 
+how they can ease certain latencies incurred by harmful workloads.
+
+here are two approaches to raising the priority of the guest RTOS process, specifically I/O thread and VCPU threads. On one hand, there is a Linux system call sched_ 
+setscheduler() that can be invoked by KVM to set priority when creating those threads. 
+On the other, Linux shell provides a command chrt that can be used by users to change 
+priority after start-up.
+\end{comment}
+
+\bigskip \noindent
+\textcolor{red}{"Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System”}
+\textcolor{red}{explores the integration of real-time operating systems (RTOS) with general-purpose operating systems (GPOS) using Kernel-Based Virtual Machine (KVM) technology.
+The authors, Ruhui Ma, Fanfu Zhou, Erzhou Zhu, and Haibing Guan, propose an embedded real-time virtualization architecture that combines VxWorks (RTOS) and Linux (GPOS) on a multi-core platform1.
+This architecture aims to address issues such as responsiveness, system performance, and CPU resource utilization.
+The paper details several performance tuning methods, including CPU shielding, prioritization, and a novel co-scheduling algorithm, to achieve sub-millisecond interrupt response latency on the guest VxWorks2.
+Experimental results demonstrate that these tuning methods significantly improve real-time responsiveness and system performance.
+The study concludes that with careful system setups, such as applying RT patches and real-time tunings on the host Linux, the proposed architecture is suitable for certain embedded real-time scenarios34.
+The work is supported by various Chinese national foundations and highlights the potential of virtualization technology in enhancing the capabilities of embedded systems.}
+
+\clearpage
+
+\textcolor{red}{"Demystifying the Real-Time Linux Scheduling Latency”}
+\bigskip \noindent \textcolor{red}{Daniel Bristot de Oliveira and colleagues \cite{deoliveiraDemystifyingRealTimeLinux} explores the intricacies of Linux as a real-time operating system, specifically focusing on the PREEMPT\_RT variant.
 The authors critique the traditional cyclictest tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
@@
 This work contributes significantly to the understanding and improvement of real-time performance in Linux systems.}
-
-\bigskip \noindent \textcolor{red}{The white paper, “Real-Time Performance Tuning Best Practice Guidelines for KVM-Based Virtual Machines,” provides comprehensive guidelines for optimizing real-time performance in KVM-based virtual machines on Intel Architecture (IA) platforms1.
-It begins by explaining the importance of real-time systems, which must respond to events within strict time constraints, and the relevance of tuning these systems for applications running in virtual machines (VMs), particularly in the manufacturing sector2.
-The document details the testbed configuration used for performance evaluation, including an industrial PC based on the Intel Whiskey Lake-U platform with specific BIOS, kernel, and host configurations3.
-Key BIOS settings such as disabling Hyper-Threading and Intel SpeedStep, and enabling Intel VMX and VT-d, are recommended to enhance real-time performance.
-Kernel configurations include using a real-time kernel, setting specific kernel parameters, and isolating CPUs for real-time workloads.
-Host configurations involve setting CPU affinity for system services, cache isolation, and disabling certain services like irqbalance and thermald.
-The paper also outlines QEMU-KVM configurations, such as tuning lapic timer advance and setting CPU affinity for QEMU threads.
-For the guest OS in the real-time VM, similar configurations to the host are recommended, with additional tuning for sched\_rt\_runtime\_us.
-The performance evaluation section describes a rigorous testing process, including running heavy workloads alongside the real-time VM, which demonstrated significantly enhanced real-time performance with latency in the range of tens of microseconds4.
-The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads5.}
-
-\bigskip \noindent \textcolor{red}{The paper titled “Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System” explores the integration of real-time operating systems (RTOS) with general-purpose operating systems (GPOS) using Kernel-Based Virtual Machine (KVM) technology.
-The authors, Ruhui Ma, Fanfu Zhou, Erzhou Zhu, and Haibing Guan, propose an embedded real-time virtualization architecture that combines VxWorks (RTOS) and Linux (GPOS) on a multi-core platform1.
-This architecture aims to address issues such as responsiveness, system performance, and CPU resource utilization.
-The paper details several performance tuning methods, including CPU shielding, prioritization, and a novel co-scheduling algorithm, to achieve sub-millisecond interrupt response latency on the guest VxWorks2.
-Experimental results demonstrate that these tuning methods significantly improve real-time responsiveness and system performance.
-The study concludes that with careful system setups, such as applying RT patches and real-time tunings on the host Linux, the proposed architecture is suitable for certain embedded real-time scenarios34.
-The work is supported by various Chinese national foundations and highlights the potential of virtualization technology in enhancing the capabilities of embedded systems.}
 
@@
 The results indicate that with careful system configurations, KVM can effectively support real-time performance requirements in embedded systems.}
-
-\bigskip \noindent \textcolor{red}{The paper titled “Performance Analysis Towards a KVM-Based Embedded Real-Time Virtualization Architecture” explores the challenges and solutions associated with integrating real-time operating systems (RTOS) with general-purpose operating systems (GPOS) on multi-core processors in embedded systems.
-The authors propose an architecture based on the Kernel-Based Virtual Machine (KVM) that combines VxWorks (an RTOS) and Linux (a GPOS) to leverage the benefits of both systems.
-The study focuses on evaluating the impact of KVM on the interrupt-response times of VxWorks when running as a guest OS.
-By applying real-time performance tuning methods on the host Linux, the authors demonstrate that sub-millisecond interrupt response latency can be achieved, making the architecture suitable for certain embedded real-time scenarios.
-The paper also discusses hardware-assisted virtualization, the role of KVM as a hypervisor, and the experimental setup used to measure the performance impacts.
-The results indicate that with appropriate system configurations, such as prioritization and CPU shielding, the proposed architecture can effectively meet real-time performance requirements.}
-
-\bigskip \noindent \textcolor{red}{The paper titled “Real-time capabilities in the standard Linux Kernel: How to enable and use them?” explores the transformation of the Linux operating system into a real-time system using the PREEMPT-RT patches1.
-Originally designed as a general-purpose OS, Linux has gained popularity in real-time applications due to its low cost and open-source nature2.
-The authors discuss the importance of real-time systems, which must meet strict temporal requirements to ensure system correctness.
-They differentiate between hard and soft real-time systems, emphasizing that Linux, with the PREEMPT-RT patch, can achieve soft real-time performance3.
-The PREEMPT-RT patch, maintained by Ingo Molnar and Thomas Gleixner, allows almost the entire kernel to be preempted, enhancing its real-time capabilities4.
-Key kernel configurations and runtime precautions are highlighted to maintain real-time performance.
-The paper concludes that while Linux can achieve real-time performance, it is still not as deterministic as commercial real-time operating systems like QNX and Windows CE5.
-However, with proper configuration, Linux can be a viable option for real-time applications.}
 
@@
 \chapter{Results}\label{cha:results}
+
+Online
+
 ~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
@@
 ~\cite{WhatRealtimeLinux}
-~\cite{maPerformanceTuningKVMbased}
 ~\cite{LinuxProcessPriorities}
@@
 ~\cite{KVMQemuVirtualization}
-~\cite{RealTimePerformanceTuning2022}
 ~\cite{MG996RServoMotor}
@@
 ~\cite{InterfacesSplittersSIGMATEK}
-~\cite{adamPerformanceAssessmentLinux2021}
+~\cite{lutsykPipelinedMulticoreMachine2020}
+~\cite{CPUUnitsSIGMATEK}
+
+\noindent Papers
+
+~\cite{RealTimePerformanceTuning2022}
+~\cite{maPerformanceTuningKVMbased}
+~\cite{kiszkaLinuxRealTimeHypervisor}
+~\cite{deoliveiraDemystifyingRealTimeLinux}
+~\cite{mckenneyRealTimeVs}
+~\cite{yoonRealTimePerformanceAnalysis}
+~\cite{huang2015performance}
+~\cite{junzhangPerformanceAnalysisKVMBased2010}
+~\cite{perneelRealtimeCapabilitiesStandard2015}
+~\cite{scordinoRealTimeVirtualizationIndustrial2020}
+
+~\cite{broskyShieldedProcessorsGuaranteeing2003}
 ~\cite{adamRealTimePerformanceResponse2021}
-~\cite{broskyShieldedProcessorsGuaranteeing2003}
+~\cite{pielAsymmetricSchedulingLoad2006}
+~\cite{malallahComprehensiveStudyKernel2021}
+~\cite{queirozTestingLimitsGeneralpurpose2023}
 ~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
-~\cite{deoliveiraDemystifyingRealTimeLinux}
-~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
 ~\cite{guStateoftheArtSurveyRealTime2012}
-~\cite{HardRealTime}
 ~\cite{javierperezHowRealTime2022}
 ~\cite{kirovaImpactModernVirtualization2019}
-~\cite{kiszkaLinuxRealTimeHypervisor}
-~\cite{huang2015performance}
-~\cite{lutsykPipelinedMulticoreMachine2020}
-~\cite{malallahComprehensiveStudyKernel2021}
+~\cite{REALTIMEOPERATING}
+~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
 ~\cite{masrurVMBasedRealTimeServices2010}
-~\cite{mckenneyRealTimeVs}
-~\cite{pielAsymmetricSchedulingLoad2006}
-~\cite{queirozTestingLimitsGeneralpurpose2023}
+
+~\cite{HardRealTime}
+~\cite{adamPerformanceAssessmentLinux2021}
+~\cite{jiangRealTimeSystemManycore}
+~\cite{wangRealtimeEmbeddedSystems2017}
+~\cite{lipariRealTimeSchedulingHard}
+~\cite{canbazPerformanceAnalysisRealtime2022}
 ~\cite{reghenzaniRealTimeLinuxKernel2020}
-~\cite{yoonRealTimePerformanceAnalysis}
-~\cite{CPUUnitsSIGMATEK}
 \clearpage

### 2024-08-14 Masterarbeit_Robotik_Pamuk.tex
Related Work 1.3 Papers read

@@
 \usepackage{rotating}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+\usetikzlibrary{positioning}
+\usepackage[none]{hyphenat}
 
@@
 
-\section{Application Context}
+\section{Application Context}\label{sec:application_context}
 
@@
 This operating system will be referred to as ``Salamander 4'' in this work.
-The details of Salamander 4 are explained in chapter~\ref{cha:salamander4}.
+The details of Salamander 4 are explained in Chapter~\ref{cha:salamander4}.
 	\item Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open-source project that helps develop custom Linux-based systems. Yocto is discussed in Section~\ref{sec:yocto}. 
@@
 	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal through real-time performance tunings, focussing on latency requirements set by SIGMATEK. 
-	These modifications are subject of chapter~\ref{cha:real-time_tuning}.
+	These modifications are subject of Chapter~\ref{cha:real-time_tuning}.
+	\item The latency tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to use by SIGMATEK. 
 	\item SIGMATEK uses the VARAN bus for communication between machines and systems, a real-time Ethernet network that was specially developed for industrial automation.
@@
 Subsequently, the PCV 522 module of SIGMATEK was connected with the Pulse Width Module PW 022 of SIGMATEK over the VARAN bus and the VI 021 module of SIGMATEK serving as the power supply.
-	\item The robotic application from chapter \ref{cha:robotic_application} was written in Lasal Class 2, an object oriented programming tool of SIGMATEK with client-server technology and graphic representation.
+	\item The robotic application from Chapter \ref{cha:robotic_application} was written in Lasal Class 2, an object oriented programming tool of SIGMATEK with client-server technology and graphic representation.
 	\end{itemize}
@@
 
-\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased} \cite{maPerformanceTuningKVMbased} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
+\bigskip \noindent \citeauthor{yoonRealTimePerformanceAnalysis} \cite{yoonRealTimePerformanceAnalysis} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
+They implement EtherCAT for real-time communication of distributed devices whereas this thesis uses Varan.
+The authors highlight the importance of timely data processing in robots that collect data from their environment and respond accordingly through their sensors and actuators.
+The authors conclude with an experiment that their robotic system could achieve deterministic control of actuators and sensors even under heavy system load through the use of Linux with the real-time preemption patch.
+Their research has similarities to this master's thesis in terms of real-time tunings for robot control.
+However, this thesis includes also the virtualization aspect that is central.
+
+\bigskip \noindent Virtualization offers the great benefit of reducing overall hardware costs, since all non-critical activities and real-time control tasks can run on the same hardware.
+\citeauthor{scordinoRealTimeVirtualizationIndustrial2020} \cite{scordinoRealTimeVirtualizationIndustrial2020} take advantage of this, but also address the challenge of interference on shared hardware resources, which can degrade real-time performance.
+
+\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased} \cite{maPerformanceTuningKVMbased} \cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
 When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS.
-Once the interrupt is delivered to the guest RTOS, it incurs additional latency since there is also the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started.
-These stages are the primary sources of latencies caused by KVM and depicted in Figure \ref{fig:six_stages_kvm}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/six_stages_kvm.png}
-	\caption[Six stages for the delivery of the interrupt to the guest]{Six stages for the delivery of the interrupt to the guest, followed by the guest IRT \cite{maPerformanceTuningKVMbased}}
-	\label{fig:six_stages_kvm}
-\end{figure}
-
-Each of these stages is breifly described below:
-
-\begin{enumerate}
-    \item \textbf{Guest Code}: The guest OS is executing its normal operations.
-    \item \textbf{Interrupt Happens}
-    \item \textbf{VM Exit}: The host processes the VM exit, which involves handling the interrupt.
-    \item \textbf{After VM Exit}: The host performs necessary operations after the VM exit.
-    \item \textbf{Host ISR}: The host executes its ISR to handle the interrupt.
-    \item \textbf{Host Scheduling}: The host schedules the guest VM to resume execution.
-    \item \textbf{Before VM Entry}: The host prepares to re-enter the guest VM.
-    \item \textbf{VM Entry}: The host completes the process of re-entering the guest VM.
-    \item \textbf{Guest IRT}: The guest OS starts executing the ISR to handle the interrupt.
-\end{enumerate}
-
-\begin{comment}
-and then apply real-time tunings on the host Linux and analyze 
-how they can ease certain latencies incurred by harmful workloads.
-
-here are two approaches to raising the priority of the guest RTOS process, specifically I/O thread and VCPU threads. On one hand, there is a Linux system call sched_ 
-setscheduler() that can be invoked by KVM to set priority when creating those threads. 
-On the other, Linux shell provides a command chrt that can be used by users to change 
-priority after start-up.
-\end{comment}
-
-\bigskip \noindent
-\textcolor{red}{"Performance Tuning Towards a KVM-based Embedded Real-Time Virtualization System”}
-\textcolor{red}{explores the integration of real-time operating systems (RTOS) with general-purpose operating systems (GPOS) using Kernel-Based Virtual Machine (KVM) technology.
-The authors, Ruhui Ma, Fanfu Zhou, Erzhou Zhu, and Haibing Guan, propose an embedded real-time virtualization architecture that combines VxWorks (RTOS) and Linux (GPOS) on a multi-core platform1.
-This architecture aims to address issues such as responsiveness, system performance, and CPU resource utilization.
-The paper details several performance tuning methods, including CPU shielding, prioritization, and a novel co-scheduling algorithm, to achieve sub-millisecond interrupt response latency on the guest VxWorks2.
-Experimental results demonstrate that these tuning methods significantly improve real-time responsiveness and system performance.
-The study concludes that with careful system setups, such as applying RT patches and real-time tunings on the host Linux, the proposed architecture is suitable for certain embedded real-time scenarios34.
-The work is supported by various Chinese national foundations and highlights the potential of virtualization technology in enhancing the capabilities of embedded systems.}
-
-\clearpage
-
-\textcolor{red}{"Demystifying the Real-Time Linux Scheduling Latency”}
-\bigskip \noindent \textcolor{red}{Daniel Bristot de Oliveira and colleagues \cite{deoliveiraDemystifyingRealTimeLinux} explores the intricacies of Linux as a real-time operating system, specifically focusing on the PREEMPT\_RT variant.
-The authors critique the traditional cyclictest tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
-To address these issues, the paper leverages the Thread Synchronization Model of Linux to derive a set of properties and rules that define the kernel’s behavior from a scheduling perspective12.
-This model helps in establishing a sound bound to the scheduling latency by considering all possible sources of delays in synchronization events34.
-The authors also introduce an efficient tracing method to observe kernel events with minimal overhead, resulting in a practical tool for deriving reliable scheduling latency bounds5.
-The paper concludes with an experimental analysis comparing cyclictest and the proposed method, demonstrating that the latter can find sound bounds faster and with acceptable overheads6.
-This work contributes significantly to the understanding and improvement of real-time performance in Linux systems.}
-
-\bigskip \noindent \textcolor{red}{The paper "‘Real Time’ vs.
-‘Real Fast’: How to Choose?"1 by Paul E.
-McKenney from IBM Linux Technology Center explores the distinction between real-time and real-fast computing in the context of Linux systems.
-It provides a comprehensive analysis to help developers make informed design choices between achieving quick start times (real-time) and fast completion times (real-fast)2.
-The paper uses two example applications: an embedded fuel-injection system and a Linux kernel build, to illustrate the differences3.
-The fuel-injection scenario, which requires precise timing, benefits from real-time Linux, while the kernel build, which prioritizes throughput, performs better on non-real-time Linux.
-The paper discusses factors influencing the choice, such as workload duration, CPU utilization, and memory usage.
-It also addresses the overheads associated with real-time Linux, including increased locking and interrupt handling costs45.
-The conclusion emphasizes using the right tool for the job and suggests ongoing work to reduce the performance penalty of real-time Linux, potentially making it suitable for a broader range of workloads6.}
-
-\bigskip \noindent \textcolor{red}{The paper titled “Real-Time Performance Analysis in Linux-Based Robotic Systems” by Hobin Yoon, Jungmoo Song, and Jamee Lee from Samsung Advanced Institute of Technology explores the feasibility of using real-time flavored Linux for controlling complex humanoid robots1.
-The authors emphasize the importance of precise control periods for mobile and humanoid robots, which require timely responses to external stimuli to maintain balance and smooth operation.
-The study introduces the adoption of EtherCAT technology for real-time communication among distributed devices, highlighting its flexible topology, simple configuration, and cost-effectiveness2.
-The paper details the design and implementation of the Robot Control Kernel Subsystem (RCKS), which integrates EtherCAT master and slave devices, and discusses various real-time performance tunings applied to the Linux kernel to achieve deterministic timing.
-Experimental results demonstrate that with appropriate real-time tunings, including CPU shielding, memory locking, and spinning nanosleep, Linux can meet the stringent timing requirements of robotic systems, achieving control periods within acceptable limits even under heavy system load.
-The authors conclude that Linux, with its real-time enhancements, is a viable solution for deploying in advanced robotic systems, offering both real-time capabilities and the benefits of a general-purpose OS3.}
-
-\bigskip \noindent \textcolor{red}{The paper titled “Performance Evaluation of Xenomai 3” by Ching-Chun (Jim) Huang, Chan-Hsiang Lin, and Che-Kang Wu, presents an in-depth analysis of Xenomai 3, a real-time framework for Linux.
-Xenomai 3 operates either as a dual-kernel system alongside Linux or natively over mainline Linux kernels, enhanced by PREEMPT RT for stricter response times1.
-The study evaluates Xenomai 3’s performance, particularly in dual-kernel configuration, using ARM Cortex-A series benchmarks.
-The authors highlight Xenomai 3’s optimizations in RTOS API emulation, thread-to-thread communications, and reduced overhead.
-The paper compares Xenomai 3 with its predecessor, Xenomai 2, and the PREEMPT RT-enhanced Linux kernel, demonstrating superior maximum response times to external interrupts in dual-kernel setups.
-The study employs BeagleBone Black hardware and various kernel configurations to measure latencies under idle and CPU-stressed conditions2.
-Results indicate that Xenomai 3 generally outperforms Xenomai 2 and PREEMPT RT, except for a slight degradation in timer-IRQ latency under stress.
-The paper concludes that Xenomai 3 offers significant improvements in real-time performance and flexibility, making it a robust choice for real-time applications.}
-
-\bigskip \noindent \textcolor{red}{The paper titled “Performance Analysis Towards a KVM-Based Embedded Real-Time Virtualization Architecture” explores the challenges and solutions associated with integrating real-time operating systems (RTOS) with general-purpose operating systems (GPOS) in embedded systems using virtualization technology.
-The authors propose an architecture based on the Kernel-Based Virtual Machine (KVM) that combines VxWorks (an RTOS) and Linux (a GPOS) on a multi-core platform12.
-The primary focus is on evaluating how KVM affects the interrupt-response times of VxWorks when running as a guest OS3.
-The study highlights the shift from single-core to multi-core processors in embedded systems and the need for legacy RTOS to adapt to multi-core environments4.
-The paper discusses hardware-assisted virtualization, particularly Intel’s VT technology, and the implementation of KVM as a hypervisor within the Linux kernel.
-Through experimental setups and performance tuning methods like prioritization and CPU shielding, the authors demonstrate that sub-millisecond interrupt response latency can be achieved, making the proposed architecture suitable for certain embedded real-time applications5.
-The results indicate that with careful system configurations, KVM can effectively support real-time performance requirements in embedded systems.}
-
-\bigskip \noindent \textcolor{red}{This paper presents a modular platform for industrial automation that leverages open-source components to run both non-critical activities and real-time control tasks on the same hardware, thereby reducing overall hardware costs.
-The platform uses a hypervisor to provide isolation, allowing concurrent execution of real-time control code and design tools.
-The authors detail the software architecture, including the hypervisor and Real-Time Operating System (RTOS), and discuss the necessary extensions for practical application.
-They address the challenge of interference on shared hardware resources, which can degrade real-time performance.
-The paper also includes experimental results demonstrating the platform’s effectiveness and performance.
-The authors highlight the benefits of using open-source software to avoid expensive commercial solutions and emphasize the importance of multi-core processors and hardware-assisted virtualization in industrial automation.
-The work is supported by the European Commission through the I-MECH ECSEL project.}
+Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), since there is also the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started.
+These stages are the primary sources of latencies caused by KVM, they are breifly described in Figure \ref{fig:kvm_stages}.
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}[node distance = 0.45cm, auto,
+		block/.style={
+			rectangle,
+			draw,
+			fill=blue!20, 
+			text width=35em, 
+			rounded corners, 
+			minimum height=3.5em,
+			align=center,
+			inner xsep=2em,
+			inner ysep=0.5em 
+		},
+		line/.style={draw, -latex'}
+	]
+		% Place nodes
+		\node [block] (init) {\textbf{Guest Code}\\ The guest OS is executing its normal operations.};
+		\node [block, below=of init] (vmexit) {\textbf{VM Exit}\\ External interrupt triggers VM exit. Hardware stores interrupt info in VMCS.};
+		\node [block, below=of vmexit] (aftervmexit) {\textbf{After VM Exit}\\ Control returns to kernel space. KVM retrieves interrupt info from VMCS,\\ marks it as pending for guest.};
+		\node [block, below=of aftervmexit] (hostisr) {\textbf{Host ISR}\\ Host Linux handles the interrupt.};
+		\node [block, below=of hostisr] (hostscheduling) {\textbf{Host Scheduling}\\ Host kernel invokes scheduler. Ideally, guest RTOS process isn't preempted.};
+		\node [block, below=of hostscheduling] (beforevmentry) {\textbf{Before VM Entry}\\ KVM's kernel module prepares for VM entry.
+Detects pending interrupt,\\ injects virtual interrupt into guest by writing to VMCS.};
+		\node [block, below=of beforevmentry] (vmentry) {\textbf{VM Entry}\\ Hardware emulates conventional interrupt handling for guest based on VMCS info.};
+		\node [block, below=of vmentry] (guestirt) {\textbf{Guest IRT}\\ Guest's interrupt handler is invoked to service the interrupt.};
+		% Draw edges
+		\path [line] (init) -- (vmexit);
+		\path [line] (vmexit) -- (aftervmexit);
+		\path [line] (aftervmexit) -- (hostisr);
+		\path [line] (hostisr) -- (hostscheduling);
+		\path [line] (hostscheduling) -- (beforevmentry);
+		\path [line] (beforevmentry) -- (vmentry);
+		\path [line] (vmentry) -- (guestirt);
+	\end{tikzpicture}
+	\caption{Flowchart of operations in a virtualized environment}
+	\label{fig:kvm_stages}
+\end{figure}
+
+\noindent The paper also details several performance tuning methods, including CPU shielding and prioritization, which have been applied in this work.
+
+\bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor} \cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
+Most importantly for this master's thesis, the author warns of starvation when the priorities of QEMU’s threads are raised and lifted into a real-time scheduling class.
+Starvation is when a process does not get the resources it needs for a long time because the resources are being allocated to other processes.
+A way to reduce this risk is by not using more virtual CPUs (VCPUs) than there are actual processor cores.
+Moreover, Linux has a feature that can limit how much CPU time all real-time tasks can use in the host system, which was an important information for the real-time tunings of this thesis later.
+
+\bigskip \noindent \citeauthor{mckenneyRealTimeVs} \cite{mckenneyRealTimeVs} addresses the overheads associated with real-time Linux when using KVM and QEMU.
+The author lists several sources of overhead, including memory locking to avoid page-fault latencies, increased overhead of locking and interrupts due to more-aggressive preemption, threaded interrupts that permit long-running interrupt handlers to be preempted by high-priority real-time processes, real-time task scheduling that requires global scheduling, high-resolution timers for tens-of-microseconds accuracy and precision, and preemptible RCU which has slightly higher read-side overhead than Classic RCU.
+These information helped understand the trade-offs involved in real-time Linux systems.
+
+\bigskip \noindent \citeauthor{deoliveiraDemystifyingRealTimeLinux} \cite{deoliveiraDemystifyingRealTimeLinux} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
+This is the main reason why this tool was not used to measure the scheduling latency in this work.
+Instead, the \texttt{latency} tool was used,as was determined in Chapter \ref{sec:application_context} .
 
@@
 The survey concludes with an overview of industrial and scientific use cases that have benefited from the PREEMPT\_RT patch, emphasizing its significance in fields like aerospace, automotive, and high-performance computing.}
+
+\bigskip \noindent \textcolor{red}{The paper titled “Performance Evaluation of Xenomai 3” by Ching-Chun (Jim) Huang, Chan-Hsiang Lin, and Che-Kang Wu, presents an in-depth analysis of Xenomai 3, a real-time framework for Linux.
+Xenomai 3 operates either as a dual-kernel system alongside Linux or natively over mainline Linux kernels, enhanced by PREEMPT RT for stricter response times1.
+The study evaluates Xenomai 3’s performance, particularly in dual-kernel configuration, using ARM Cortex-A series benchmarks.
+The authors highlight Xenomai 3’s optimizations in RTOS API emulation, thread-to-thread communications, and reduced overhead.
+The paper compares Xenomai 3 with its predecessor, Xenomai 2, and the PREEMPT RT-enhanced Linux kernel, demonstrating superior maximum response times to external interrupts in dual-kernel setups.
+The study employs BeagleBone Black hardware and various kernel configurations to measure latencies under idle and CPU-stressed conditions2.
+Results indicate that Xenomai 3 generally outperforms Xenomai 2 and PREEMPT RT, except for a slight degradation in timer-IRQ latency under stress.
+The paper concludes that Xenomai 3 offers significant improvements in real-time performance and flexibility, making it a robust choice for real-time applications.}
 
@@
 This script is used to start QEMU with required parameters to boot Salamamder 4 OS.
-It will be adjusted in chapter~\ref{cha:real-time_tuning} in order to accompany real-time performance tunings.
+It will be adjusted in Chapter~\ref{cha:real-time_tuning} in order to accompany real-time performance tunings.
 
@@
 
-\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in chapter~\ref{cha:methodology}.
+\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in Chapter~\ref{cha:methodology}.
 The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the~\texttt{uname -a} command.
@@
 The very first step in this regard is the isolation of a CPU to dedicate it solely to the desired real-time task.
-CPU isolation and affinity will be covered in detail later in section~\ref{subsec:cpu_isolation}, but its impact will be briefly demonstrated here.
+CPU isolation and affinity will be covered in detail later in Section~\ref{subsec:cpu_isolation}, but its impact will be briefly demonstrated here.
 
@@
 	\caption[Latency statistics of Salamander 4 after BIOS configurations]{Latency statistics of Salamander 4 virtualization after BIOS configurations in microseconds}
-	\label{tab:latency_statistics_virt}
+	\label{tab:latency_statistics_bios}
 	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
@@
 	\item \textbf{isolcpus=4}: Isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned.
-CPU isolation is explained in detail in section~\ref{subsec:cpu_isolation}.
+CPU isolation is explained in detail in Section~\ref{subsec:cpu_isolation}.
 	\item \textbf{rcu\_nocbs=4}: The Linux kernel uses a synchronization mechanism called RCU, or Read-Copy-Update.
@@
 	\caption[Latency statistics of Salamander 4 after Kernel configurations]{Latency statistics of Salamander 4 virtualization after Kernel configurations in microseconds}
-	\label{tab:latency_statistics_virt}
+	\label{tab:latency_statistics_kernel}
 	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
@@
 \subsection{Disable RCU CPU Stall Warnings}
-As already mentioned in section~\ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
+As already mentioned in Section~\ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
 An RCU CPU stall in the Linux kernel can occur due to several reasons.
@@
 	\caption[Latency statistics of Salamander 4 after Host configurations]{Latency statistics of Salamander 4 virtualization after Host configurations in microseconds}
-	\label{tab:latency_statistics_virt}
+	\label{tab:latency_statistics_host}
 	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
@@
 		\caption[Latency statistics of Salamander 4 after QEMU configurations]{Latency statistics of Salamander 4 virtualization after QEMU configurations in microseconds}
-		\label{tab:latency_statistics_virt}
+		\label{tab:latency_statistics_qemu}
 		\setlength{\tabcolsep}{0.5em} % for the horizontal padding
@@
 \chapter{Summary and Outlook}\label{cha:summary_and_outlook}
+\section{Trace-cmd \& Kernelshark}
+
 %

### 2024-08-15 Masterarbeit_Robotik_Pamuk.tex
Related Work 1.3 Papers read+

@@
 \usepackage{tikz}
+\usepackage{setspace}
 \usetikzlibrary{shapes,arrows}
@@
 \noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.The core component of an RTOS that enables real-time capabilities is the kernel.
-It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior.
+It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior \cite{malallahComprehensiveStudyKernel2021}.
 It employs preemptive scheduling mechanisms to allow high-priority tasks to preempt lower-priority tasks, ensuring that time-critical tasks are not delayed.
@@
 \section{Application Context}\label{sec:application_context}
-
 This section briefly defines the context and scope within which this master's thesis was composed, highlighting the specific contributions and requirements set by SIGMATEK GmbH \& Co KG~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.
@@
 \section{Related Work and State of the Art}
-
-\begin{comment}
-Yocto primarily uses
-
-At the beginning of the work, a detailed analysis of the individual virtualization options of Salamander 4 was carried out.
-
-In particular, the virtualization performance of Ubuntu 22.04, Windows 10 and WSL under QEMU was compared.
-\end{comment}
-
 This master’s thesis builds upon a number of previous studies in the field of virtualization of real-time systems and the inherent latency.
@@
 They concluded with their Linux build and a testing application that Linux can be a viable option for real-time applications, if it is correctly configured.
+Similarly, \citeauthor{adamRealTimePerformanceResponse2021} \cite{adamRealTimePerformanceResponse2021} applied the PREEMPT\_RT patch on both Raspberry Pi 3 and BeagleBone Black and achieved significantly lower latencies compared to standard Linux kernels.
 
@@
 
+\bigskip \noindent \citeauthor{taccariEmbeddedRealTimeVirtualization2014} \cite{taccariEmbeddedRealTimeVirtualization2014}, \citeauthor{sandstromVirtualizationTechnologiesEmbedded2013} \cite{sandstromVirtualizationTechnologiesEmbedded2013}, and \citeauthor{guStateoftheArtSurveyRealTime2012} \cite{guStateoftheArtSurveyRealTime2012} all discuss and compare the current state-of-the-art in virtualization technologies on embedded real-time platforms.
+
+highlighting similar advancements and challenges in their respective works.
+
+\bigskip \noindent 
+\citeauthor{garcia-vallsChallengesRealtimeVirtualization2014} \cite{garcia-vallsChallengesRealtimeVirtualization2014}
+
+\citeauthor{cinqueVirtualizingMixedCriticalitySystems2022} \cite{cinqueVirtualizingMixedCriticalitySystems2022}
+
+\citeauthor{reghenzaniRealTimeLinuxKernel2020} \cite{reghenzaniRealTimeLinuxKernel2020}
+
 \bigskip \noindent Virtualization offers the great benefit of reducing overall hardware costs, since all non-critical activities and real-time control tasks can run on the same hardware.
 \citeauthor{scordinoRealTimeVirtualizationIndustrial2020} \cite{scordinoRealTimeVirtualizationIndustrial2020} take advantage of this, but also address the challenge of interference on shared hardware resources, which can degrade real-time performance.
+In addition, \citeauthor{javierperezHowRealTime2022} \cite{javierperezHowRealTime2022} state that hardware-based PLCs are costly to maintain and lack the flexibility needed for modern, resource-intensive applications like Machine Learning (ML) or Artificial Intelligence (AI).
 
@@
 Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), since there is also the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started.
-These stages are the primary sources of latencies caused by KVM, they are breifly described in Figure \ref{fig:kvm_stages}.
-
-\begin{figure}[H]
-	\centering
-	\begin{tikzpicture}[node distance = 0.45cm, auto,
+These stages are the primary sources of latencies caused by KVM, they are briefly described in Figure \ref{fig:kvm_stages}.
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}[node distance = 0.35cm, auto, % Adjusted node distance
 		block/.style={
@@
 			align=center,
-			inner xsep=2em,
-			inner ysep=0.5em 
+			inner sep=0.2em % Adjusted inner separation
 		},
@@
 		% Place nodes
-		\node [block] (init) {\textbf{Guest Code}\\ The guest OS is executing its normal operations.};
-		\node [block, below=of init] (vmexit) {\textbf{VM Exit}\\ External interrupt triggers VM exit. Hardware stores interrupt info in VMCS.};
-		\node [block, below=of vmexit] (aftervmexit) {\textbf{After VM Exit}\\ Control returns to kernel space. KVM retrieves interrupt info from VMCS,\\ marks it as pending for guest.};
-		\node [block, below=of aftervmexit] (hostisr) {\textbf{Host ISR}\\ Host Linux handles the interrupt.};
-		\node [block, below=of hostisr] (hostscheduling) {\textbf{Host Scheduling}\\ Host kernel invokes scheduler. Ideally, guest RTOS process isn't preempted.};
-		\node [block, below=of hostscheduling] (beforevmentry) {\textbf{Before VM Entry}\\ KVM's kernel module prepares for VM entry.
-Detects pending interrupt,\\ injects virtual interrupt into guest by writing to VMCS.};
-		\node [block, below=of beforevmentry] (vmentry) {\textbf{VM Entry}\\ Hardware emulates conventional interrupt handling for guest based on VMCS info.};
-		\node [block, below=of vmentry] (guestirt) {\textbf{Guest IRT}\\ Guest's interrupt handler is invoked to service the interrupt.};
+		\node [block] (init) {\scriptsize \textbf{Guest Code}\\  The guest OS is executing its normal operations.};
+		\node [block, below=of init] (vmexit) {\scriptsize \textbf{VM Exit}\\  External interrupt triggers VM exit. Hardware stores interrupt info in VMCS.};
+		\node [block, below=of vmexit] (aftervmexit) {\scriptsize \textbf{After VM Exit}\\  Control returns to kernel space.
+KVM retrieves interrupt info from VMCS,\\\vspace{-0.5em} marks it as pending for guest.};
+		\node [block, below=of aftervmexit] (hostisr) {\scriptsize \textbf{Host ISR}\\  Host Linux handles the interrupt.};
+		\node [block, below=of hostisr] (hostscheduling) {\scriptsize \textbf{Host Scheduling}\\  Host kernel invokes scheduler. Ideally, guest RTOS process isn't preempted.};
+		\node [block, below=of hostscheduling] (beforevmentry) {\scriptsize \textbf{Before VM Entry}\\  KVM's kernel module prepares for VM entry.
+Detects pending interrupt,\\\vspace{-0.5em} injects virtual interrupt into guest by writing to VMCS.};
+		\node [block, below=of beforevmentry] (vmentry) {\scriptsize \textbf{VM Entry}\\  Hardware emulates conventional interrupt handling for guest based on VMCS info.};
+		\node [block, below=of vmentry] (guestirt) {\scriptsize \textbf{Guest IRT}\\  Guest's interrupt handler is invoked to service the interrupt.};
 		% Draw edges
@@
 
+\bigskip \noindent \citeauthor{broskyShieldedProcessorsGuaranteeing2003} \cite{broskyShieldedProcessorsGuaranteeing2003} report in their work about the shielded CPU model for enhancing real-time performance in symmetric multiprocessor systems.
+By dedicating one or more CPUs to high-priority processes and interrupts, the goal is to provide deterministic execution of real-time applications and interrupt responses.
+The authors mention that there is no user interface for setting process CPU affinity in the standard Linux, while there is one for setting interrupt CPU affinity via the \texttt{/proc/irq/*/smp\_affinity} files.
+These files were used in this masther's thesis.
+According to \citeauthor{broskyShieldedProcessorsGuaranteeing2003}, the shielded CPU model is especially useful for tasks that require guaranteed interrupt response times, very fast interrupt responses and high-frequency, deterministic execution.
+
 \bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor} \cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
@@
 This is the main reason why this tool was not used to measure the scheduling latency in this work.
-Instead, the \texttt{latency} tool was used,as was determined in Chapter \ref{sec:application_context} .
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bigskip \noindent \textcolor{red}{Shielded Processors: Guaranteeing Sub-millisecond Response in Standard Linux by Steve Brosky and Steve Rotolo of Concurrent Computer Corporation explores the implementation and benefits of shielded processors in RedHawk Linux for real-time applications.
-The paper highlights the limitations of standard Linux in providing deterministic performance due to networking and graphics activities.
-Shielded CPUs, a model developed over a decade of experience, offload operating system overhead to a subset of CPUs, ensuring more deterministic execution and interrupt response times.
-The authors describe the modifications made to the Linux kernel, including CPU affinity settings and the disabling of local timer interrupts, to create a shielded environment.
-Benchmark tests demonstrate significant improvements in execution determinism and interrupt response times, with jitter reduced to 1.87\% on shielded CPUs compared to 26.17\% on standard Linux.
-The paper concludes that shielded processors can guarantee interrupt response times of less than 30 microseconds, even under heavy system load, making them suitable for applications with hard real-time deadlines.
-The work underscores the potential of using commodity operating systems like Linux for high-performance real-time applications by leveraging shielded processors.}
+Instead, the \texttt{latency} tool was used,as was determined in Chapter \ref{sec:application_context}.
 
@@
 The paper includes extensive experiments to illustrate the effectiveness of the proposed solutions.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2024-08-15 Masterarbeit_Robotik_Pamuk.tex
xenomai and no hyphenat

@@
 \usetikzlibrary{positioning}
-\usepackage[none]{hyphenat}
 
@@
 	\centering
-	\includegraphics[width=0.5\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
+	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
 	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
@@
 	\centering
-	\includegraphics[width=0.5\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
+	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
 	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}

### 2024-08-16 Masterarbeit_Robotik_Pamuk.tex
Related Work 1.3 Papers read++

@@
 
-\bigskip \noindent \citeauthor{taccariEmbeddedRealTimeVirtualization2014} \cite{taccariEmbeddedRealTimeVirtualization2014}, \citeauthor{sandstromVirtualizationTechnologiesEmbedded2013} \cite{sandstromVirtualizationTechnologiesEmbedded2013}, and \citeauthor{guStateoftheArtSurveyRealTime2012} \cite{guStateoftheArtSurveyRealTime2012} all discuss and compare the current state-of-the-art in virtualization technologies on embedded real-time platforms.
-
-highlighting similar advancements and challenges in their respective works.
+\bigskip \noindent \citeauthor{sandstromVirtualizationTechnologiesEmbedded2013} \cite{sandstromVirtualizationTechnologiesEmbedded2013}, and \citeauthor{taccariEmbeddedRealTimeVirtualization2014} \cite{taccariEmbeddedRealTimeVirtualization2014} and  discuss and compare the current state-of-the-art in virtualization technologies with a focus on embedded real-time platforms.
+The latter emphasizes, that research on real-time systems is focusing on virtualization and multicore scheduling and underlines that flexibility and cost are key metrics.
+Both works acknowledge the great benefits of virtualization, that is, reducing overall hardware costs, since all non-critical activities and real-time control tasks can run on the same hardware.
+In addition, \citeauthor{javierperezHowRealTime2022} \cite{javierperezHowRealTime2022} state that hardware-based Programmable Logic Controllers (PLCs) are costly to maintain and lack the flexibility needed for modern, resource-intensive applications like Machine Learning (ML) or Artificial Intelligence (AI).
+
+\bigskip \noindent Despite the promising advantages of virtualization, one must be aware of the potential drawbacks of it, too.
+\citeauthor{guStateoftheArtSurveyRealTime2012} \cite{guStateoftheArtSurveyRealTime2012} highlight issues like lock-holder preemption and task-grain scheduling.
+\citeauthor{garcia-vallsChallengesRealtimeVirtualization2014} \cite{garcia-vallsChallengesRealtimeVirtualization2014} identify challenges in supporting real-time applications in the cloud, including resource management, quality of service guarantees, temporal and spatial isolation, and network performance.
+\citeauthor{scordinoRealTimeVirtualizationIndustrial2020} \cite{scordinoRealTimeVirtualizationIndustrial2020} address the challenge of interference on shared hardware resources, which can degrade real-time performance.
 
 \bigskip \noindent
-\citeauthor{garcia-vallsChallengesRealtimeVirtualization2014} \cite{garcia-vallsChallengesRealtimeVirtualization2014}
-
 \citeauthor{cinqueVirtualizingMixedCriticalitySystems2022} \cite{cinqueVirtualizingMixedCriticalitySystems2022}
@@
 \citeauthor{reghenzaniRealTimeLinuxKernel2020} \cite{reghenzaniRealTimeLinuxKernel2020}
-
-\bigskip \noindent Virtualization offers the great benefit of reducing overall hardware costs, since all non-critical activities and real-time control tasks can run on the same hardware.
-\citeauthor{scordinoRealTimeVirtualizationIndustrial2020} \cite{scordinoRealTimeVirtualizationIndustrial2020} take advantage of this, but also address the challenge of interference on shared hardware resources, which can degrade real-time performance.
-In addition, \citeauthor{javierperezHowRealTime2022} \cite{javierperezHowRealTime2022} state that hardware-based PLCs are costly to maintain and lack the flexibility needed for modern, resource-intensive applications like Machine Learning (ML) or Artificial Intelligence (AI).
 
@@
 These files were used in this masther's thesis.
-According to \citeauthor{broskyShieldedProcessorsGuaranteeing2003}, the shielded CPU model is especially useful for tasks that require guaranteed interrupt response times, very fast interrupt responses and high-frequency, deterministic execution.
+According to the authors, the shielded CPU model is especially useful for tasks that require guaranteed interrupt response times, very fast interrupt responses and high-frequency, deterministic execution.
 
@@
 This is the main reason why this tool was not used to measure the scheduling latency in this work.
-Instead, the \texttt{latency} tool was used,as was determined in Chapter \ref{sec:application_context}.
-
-\bigskip \noindent \textcolor{red}{The paper titled “Real-Time Performance and Response Latency Measurements of Linux Kernels on Single-Board Computers” by George K.
-Adam from the CSLab Computer Systems Laboratory, University of Thessaly, investigates the real-time performance of Linux kernels with the PREEMPT\_RT patch on embedded devices like the BeagleBoard and Raspberry Pi1.
-The research focuses on measuring latencies in real-time applications at both user and kernel space using newly developed software modules tailored for these experiments2.
-The study highlights that the PREEMPT\_RT patch significantly improves the real-time performance of Linux kernels, reducing worst-case latencies to around 160 µs, making these systems more suitable for real-time applications3.
-The methodology involves response and periodic task models, and the results suggest that the proposed measurement approach can be applied to other Linux-based platforms4.
-The paper also discusses the challenges and solutions encountered during the experiments, emphasizing the importance of real-time capabilities in embedded systems for applications ranging from consumer electronics to industrial automation.}
+Instead, the \texttt{latency} tool was used, as was determined in Chapter \ref{sec:application_context}.
 
@@
 The paper concludes with a discussion on the future directions and challenges in the field of real-time virtualization for embedded systems.}
-
-\bigskip \noindent \textcolor{red}{Title: How Real (Time) Are Virtual PLCs?1
-Authors: Diogenes Javier Perez, Josef Waltl, Laurin Prenzel, Sebastian Steinhorst
-This paper explores the concept of virtual programmable logic controllers (vPLCs) as a solution to the limitations of traditional hardware-based PLCs in industrial automation.
-The authors argue that hardware PLCs are costly to maintain and lack the flexibility and scalability needed for modern applications, especially those involving artificial intelligence and machine learning2.
-The proposed vPLC operates on a commercial off-the-shelf (COTS) server using a real-time hypervisor, which allows it to manage multiple virtual machines and handle resource-intensive tasks efficiently34.
-The paper details the architecture of the vPLC, which includes a real-time operating system and a hypervisor to ensure low latency and deterministic behavior.
-The authors conducted various tests to compare the performance of vPLCs with traditional SoftPLCs, focusing on metrics such as cycle time, jitter, and response time.
-The results indicate that vPLCs can achieve performance on par with traditional PLCs for applications requiring response times below 10 milliseconds5.
-The study concludes that while further improvements are needed, vPLCs offer significant advantages in terms of flexibility, scalability, and cost-efficiency for future industrial systems6.}
 

### 2024-08-16 Masterarbeit_Robotik_Pamuk.tex
change color

@@
 
-\bigskip \noindent \textcolor{red}{Asymmetric Scheduling and Load Balancing for Real-Time on Linux SMP
+\bigskip \noindent \textcolor{orange}{Asymmetric Scheduling and Load Balancing for Real-Time on Linux SMP
 	Authors: Eric Piel, Philippe Marquet, Julien Soula, and Jean-Luc Dekeyser12
@@
 
-\bigskip \noindent \textcolor{red}{“Testing the limits of general-purpose hypervisors for real-time control systems” by Rui Queiroz, Tiago J. Cruz, and Paulo Simões from the University of Coimbra:
-	With the advent of Industry 4.0, there is a growing need for flexibility, security, and resilience in industrial automation infrastructures while maintaining real-time requirements.
-This paper explores the feasibility of using general-purpose off-the-shelf hypervisors to virtualize cyber-physical systems’ servers and control devices.
-The authors, Rui Queiroz, Tiago J.
-Cruz, and Paulo Simões, present an empirical evaluation of these hypervisors, focusing on their ability to handle real-time workloads.
-The study highlights the challenges posed by the need for determinism and low latency, which are often compromised by hypervisors designed to maximize throughput.
-The paper concludes that while some hypervisors can manage typical real-time workloads, this capability is not universal across all types of real-time systems.
-The research underscores the importance of understanding the limits and capabilities of conventional hypervisors to ensure they meet the specific requirements of industrial automation control systems.}
-
-\bigskip \noindent \textcolor{red}{Title: A State-of-the-Art Survey on Real-Time Issues in Embedded Systems Virtualization
-	Authors: Zonghua Gu and Qingling Zhao
-	Summary: This paper provides a comprehensive survey on real-time issues in virtualization for embedded systems, authored by Zonghua Gu and Qingling Zhao.
-Virtualization, widely accepted in server and cloud computing, has been increasingly applied to embedded systems with stringent timing constraints.
-The paper discusses various virtualization systems, including KVM, Xen, and L4, and their applications in embedded systems such as avionics, industrial automation, and mobile phones.
-It highlights the importance of real-time performance, security, and dependability in these systems.
-The survey covers different types of virtualization, including Type-1 (bare-metal) and Type-2 (hosted) virtualization, as well as full and para-virtualization.
-It also addresses specific real-time issues like scheduling, partitioning, and the Lock-Holder Preemption problem.
-The paper concludes with a discussion on the future directions and challenges in the field of real-time virtualization for embedded systems.}
-
 \bigskip \noindent \textcolor{red}{Impact of Modern Virtualization Methods on Timing Precision and Performance of High-Speed Applications
@@
 
-\bigskip \noindent \textcolor{red}{“Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane:
+\bigskip \noindent \textcolor{orange}{“Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane:
 	The paper “Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane provides an in-depth examination of Real Time Operating Systems (RTOS) and their critical role in embedded systems, particularly in telecommunication applications such as telephony, navigation, and military signaling systems.
@@
 
-\bigskip \noindent \textcolor{red}{The paper “Challenges in Real-Time Virtualization and Predictable Cloud Computing” explores the integration of real-time systems with cloud computing and virtualization technologies.
-The authors, Marisol García-Valls, Tommaso Cucinotta, and Chenyang Lu, identify the significant technical challenges in supporting real-time applications within cloud environments1.
-They highlight that while cloud computing offers benefits such as reduced operational costs, server consolidation, and elastic resource provisioning, it struggles to meet the demands of soft real-time applications like online video streaming, cloud-based gaming, and telecommunication management2.
-These applications require predictable performance in open, shared, and virtualized environments3.
-	The paper discusses the evolution of cloud computing, emphasizing the shift from local computing resources to distributed cloud-based models.
-It outlines the benefits of virtualization, including functional execution isolation, easier management, and the coexistence of legacy and new applications4.
-However, the authors note that the multi-tenant nature of cloud computing introduces challenges in providing stable and predictable performance due to resource sharing among multiple virtual machines (VMs).
-	The authors delve into various virtualization techniques, such as full virtualization, hardware-assisted virtualization, para-virtualization, OS-level virtualization, and application-level virtualization, each with its performance characteristics and suitability for real-time applications.
-They also address the importance of real-time scheduling and resource management in virtualized environments, highlighting the need for hierarchical scheduling and the role of real-time hypervisors.
-	The paper concludes by presenting potential future research directions to enhance the predictability and performance of cloud-based real-time applications, emphasizing the need for improved resource management frameworks and real-time virtualization techniques.}
-
 \bigskip \noindent \textcolor{red}{The paper titled “VM-Based Real-Time Services for Automotive Control Applications” is authored by Alejandro Masrur, Sebastian Drossler, Thomas Pfeuffer, and Samarjit Chakraborty.
@@
 
-\bigskip \noindent \textcolor{red}{Title: Hard Real Time Linux* Using Xenomai* on Intel® Multi-Core Processors12
+\bigskip \noindent \textcolor{orange}{Title: Hard Real Time Linux* Using Xenomai* on Intel® Multi-Core Processors12
 	Author: Amarpreet Singh Ugal, BIOS/Firmware Engineer, Intel Corporation
@@
 
-\bigskip \noindent \textcolor{red}{Title: The Real-Time Linux Kernel: A Survey on PREEMPT\_RT12
+\bigskip \noindent \textcolor{green}{Title: The Real-Time Linux Kernel: A Survey on PREEMPT\_RT12
 	Authors: Federico Reghenzani, Giuseppe Massari, and William Fornaciari from Politecnico di Milano3
@@
 
-\bigskip \noindent \textcolor{red}{The paper titled “Performance Evaluation of Xenomai 3” by Ching-Chun (Jim) Huang, Chan-Hsiang Lin, and Che-Kang Wu, presents an in-depth analysis of Xenomai 3, a real-time framework for Linux.
+\bigskip \noindent \textcolor{orange}{The paper titled “Performance Evaluation of Xenomai 3” by Ching-Chun (Jim) Huang, Chan-Hsiang Lin, and Che-Kang Wu, presents an in-depth analysis of Xenomai 3, a real-time framework for Linux.
 Xenomai 3 operates either as a dual-kernel system alongside Linux or natively over mainline Linux kernels, enhanced by PREEMPT RT for stricter response times1.

### 2024-08-16 Masterarbeit_Robotik_Pamuk.tex
~\cite

@@
 	\includegraphics[width=0.75\columnwidth]{img/introduction/kernel_generic.png}
-	\caption[Non-preemptible kernel]{Non-preemptible kernel \cite{WhatRealtimeLinuxa}}
+	\caption[Non-preemptible kernel]{Non-preemptible kernel~\cite{WhatRealtimeLinuxa}}
 	\label{fig:kernel_generic}
@@
 	\includegraphics[width=0.75\columnwidth]{img/introduction/kernel_rt.png}
-	\caption[Non-preemptible kernel]{Non-preemptible kernel \cite{WhatRealtimeLinuxa}}
+	\caption[Non-preemptible kernel]{Non-preemptible kernel~\cite{WhatRealtimeLinuxa}}
 	\label{fig:kernel_rt}
@@
 \noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.The core component of an RTOS that enables real-time capabilities is the kernel.
-It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior \cite{malallahComprehensiveStudyKernel2021}.
+It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior~\cite{malallahComprehensiveStudyKernel2021}.
 It employs preemptive scheduling mechanisms to allow high-priority tasks to preempt lower-priority tasks, ensuring that time-critical tasks are not delayed.
@@
 
-\bigskip \noindent \citeauthor{perneelRealtimeCapabilitiesStandard2015} \cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general purpose operating system without real-time applications in mind.
+\bigskip \noindent~\citeauthor{perneelRealtimeCapabilitiesStandard2015}~\cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general purpose operating system without real-time applications in mind.
 They talk about the evolution of real-time behavior in the standard Linux kernel and state that it has recently gained popularity among the real-time community, largely due to its open-source nature and stability.
@@
 They concluded with their Linux build and a testing application that Linux can be a viable option for real-time applications, if it is correctly configured.
-Similarly, \citeauthor{adamRealTimePerformanceResponse2021} \cite{adamRealTimePerformanceResponse2021} applied the PREEMPT\_RT patch on both Raspberry Pi 3 and BeagleBone Black and achieved significantly lower latencies compared to standard Linux kernels.
-
-\bigskip \noindent In the course of writing this thesis, the white paper "Real-Time Performance Tuning Best Practice Guidelines for KVM-Based Virtual Machines" \cite{RealTimePerformanceTuning2022} by Intel has been instrumental.
+Similarly,~\citeauthor{adamRealTimePerformanceResponse2021}~\cite{adamRealTimePerformanceResponse2021} applied the PREEMPT\_RT patch on both Raspberry Pi 3 and BeagleBone Black and achieved significantly lower latencies compared to standard Linux kernels.
+
+\bigskip \noindent In the course of writing this thesis, the white paper "Real-Time Performance Tuning Best Practice Guidelines for KVM-Based Virtual Machines"~\cite{RealTimePerformanceTuning2022} by Intel has been instrumental.
 Given that the working computer used for this work is equipped with an Intel processor, the guidelines presented in this document were particularly relevant and valuable for optimizing real-time performance in KVM-based virtual machines.
@@
 
-\bigskip \noindent \citeauthor{yoonRealTimePerformanceAnalysis} \cite{yoonRealTimePerformanceAnalysis} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
+\bigskip \noindent~\citeauthor{yoonRealTimePerformanceAnalysis}~\cite{yoonRealTimePerformanceAnalysis} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
 They implement EtherCAT for real-time communication of distributed devices whereas this thesis uses Varan.
@@
 
-\bigskip \noindent \citeauthor{sandstromVirtualizationTechnologiesEmbedded2013} \cite{sandstromVirtualizationTechnologiesEmbedded2013}, and \citeauthor{taccariEmbeddedRealTimeVirtualization2014} \cite{taccariEmbeddedRealTimeVirtualization2014} and  discuss and compare the current state-of-the-art in virtualization technologies with a focus on embedded real-time platforms.
+\bigskip \noindent~\citeauthor{sandstromVirtualizationTechnologiesEmbedded2013}~\cite{sandstromVirtualizationTechnologiesEmbedded2013}, and~\citeauthor{taccariEmbeddedRealTimeVirtualization2014}~\cite{taccariEmbeddedRealTimeVirtualization2014} and  discuss and compare the current state-of-the-art in virtualization technologies with a focus on embedded real-time platforms.
 The latter emphasizes, that research on real-time systems is focusing on virtualization and multicore scheduling and underlines that flexibility and cost are key metrics.
 Both works acknowledge the great benefits of virtualization, that is, reducing overall hardware costs, since all non-critical activities and real-time control tasks can run on the same hardware.
-In addition, \citeauthor{javierperezHowRealTime2022} \cite{javierperezHowRealTime2022} state that hardware-based Programmable Logic Controllers (PLCs) are costly to maintain and lack the flexibility needed for modern, resource-intensive applications like Machine Learning (ML) or Artificial Intelligence (AI).
-
-\bigskip \noindent Despite the promising advantages of virtualization, one must be aware of the potential drawbacks of it, too.
-\citeauthor{guStateoftheArtSurveyRealTime2012} \cite{guStateoftheArtSurveyRealTime2012} highlight issues like lock-holder preemption and task-grain scheduling.
-\citeauthor{garcia-vallsChallengesRealtimeVirtualization2014} \cite{garcia-vallsChallengesRealtimeVirtualization2014} identify challenges in supporting real-time applications in the cloud, including resource management, quality of service guarantees, temporal and spatial isolation, and network performance.
-\citeauthor{scordinoRealTimeVirtualizationIndustrial2020} \cite{scordinoRealTimeVirtualizationIndustrial2020} address the challenge of interference on shared hardware resources, which can degrade real-time performance.
+In addition,~\citeauthor{javierperezHowRealTime2022}~\cite{javierperezHowRealTime2022} state that hardware-based Programmable Logic Controllers (PLCs) are costly to maintain and lack the flexibility needed for modern, resource-intensive applications like Machine Learning (ML) or Artificial Intelligence (AI).
+
+\bigskip \noindent Despite the promising advantages of virtualization, one must be aware of the potential drawbacks of it, too.~\citeauthor{guStateoftheArtSurveyRealTime2012}~\cite{guStateoftheArtSurveyRealTime2012} highlight issues like lock-holder preemption and task-grain scheduling.~\citeauthor{garcia-vallsChallengesRealtimeVirtualization2014}~\cite{garcia-vallsChallengesRealtimeVirtualization2014} identify challenges in supporting real-time applications in the cloud, including resource management, quality of service guarantees, temporal and spatial isolation, and network performance.~\citeauthor{scordinoRealTimeVirtualizationIndustrial2020}~\cite{scordinoRealTimeVirtualizationIndustrial2020} address the challenge of interference on shared hardware resources, which can degrade real-time performance.
 
 \bigskip \noindent
-\citeauthor{cinqueVirtualizingMixedCriticalitySystems2022} \cite{cinqueVirtualizingMixedCriticalitySystems2022}
-
-\citeauthor{reghenzaniRealTimeLinuxKernel2020} \cite{reghenzaniRealTimeLinuxKernel2020}
-
-\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased} \cite{maPerformanceTuningKVMbased} \cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
+\citeauthor{cinqueVirtualizingMixedCriticalitySystems2022}~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
+
+\citeauthor{reghenzaniRealTimeLinuxKernel2020}~\cite{reghenzaniRealTimeLinuxKernel2020}
+
+\bigskip \noindent~\citeauthor{maPerformanceTuningKVMbased}~\cite{maPerformanceTuningKVMbased}~\cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
 When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS.
@@
 
-\bigskip \noindent \citeauthor{broskyShieldedProcessorsGuaranteeing2003} \cite{broskyShieldedProcessorsGuaranteeing2003} report in their work about the shielded CPU model for enhancing real-time performance in symmetric multiprocessor systems.
+\bigskip \noindent~\citeauthor{broskyShieldedProcessorsGuaranteeing2003}~\cite{broskyShieldedProcessorsGuaranteeing2003} report in their work about the shielded CPU model for enhancing real-time performance in symmetric multiprocessor systems.
 By dedicating one or more CPUs to high-priority processes and interrupts, the goal is to provide deterministic execution of real-time applications and interrupt responses.
@@
 
-\bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor} \cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
+\bigskip \noindent~\citeauthor{kiszkaLinuxRealTimeHypervisor}~\cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
 Most importantly for this master's thesis, the author warns of starvation when the priorities of QEMU’s threads are raised and lifted into a real-time scheduling class.
@@
 
-\bigskip \noindent \citeauthor{mckenneyRealTimeVs} \cite{mckenneyRealTimeVs} addresses the overheads associated with real-time Linux when using KVM and QEMU.
+\bigskip \noindent~\citeauthor{mckenneyRealTimeVs}~\cite{mckenneyRealTimeVs} addresses the overheads associated with real-time Linux when using KVM and QEMU.
 The author lists several sources of overhead, including memory locking to avoid page-fault latencies, increased overhead of locking and interrupts due to more-aggressive preemption, threaded interrupts that permit long-running interrupt handlers to be preempted by high-priority real-time processes, real-time task scheduling that requires global scheduling, high-resolution timers for tens-of-microseconds accuracy and precision, and preemptible RCU which has slightly higher read-side overhead than Classic RCU.
@@
 
-\bigskip \noindent \citeauthor{deoliveiraDemystifyingRealTimeLinux} \cite{deoliveiraDemystifyingRealTimeLinux} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
+\bigskip \noindent~\citeauthor{deoliveiraDemystifyingRealTimeLinux}~\cite{deoliveiraDemystifyingRealTimeLinux} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
 This is the main reason why this tool was not used to measure the scheduling latency in this work.

### 2024-08-16 Masterarbeit_Robotik_Pamuk.tex
Related Work 1.3 Papers read+++

@@
 
-\bigskip \noindent~\citeauthor{perneelRealtimeCapabilitiesStandard2015}~\cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general purpose operating system without real-time applications in mind.
+\bigskip \noindent \citeauthor{perneelRealtimeCapabilitiesStandard2015}~\cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general purpose operating system without real-time applications in mind.
 They talk about the evolution of real-time behavior in the standard Linux kernel and state that it has recently gained popularity among the real-time community, largely due to its open-source nature and stability.
@@
 They concluded with their Linux build and a testing application that Linux can be a viable option for real-time applications, if it is correctly configured.
-Similarly,~\citeauthor{adamRealTimePerformanceResponse2021}~\cite{adamRealTimePerformanceResponse2021} applied the PREEMPT\_RT patch on both Raspberry Pi 3 and BeagleBone Black and achieved significantly lower latencies compared to standard Linux kernels.
+\citeauthor{reghenzaniRealTimeLinuxKernel2020}~\cite{reghenzaniRealTimeLinuxKernel2020} provide an extensive overview on the real-time Linux kernel research, specifically on the PREEMPT RT patch.
+\citeauthor{adamRealTimePerformanceResponse2021}~\cite{adamRealTimePerformanceResponse2021} applied the PREEMPT\_RT patch on both Raspberry Pi 3 and BeagleBone Black and achieved significantly lower latencies compared to standard Linux kernels.
 
@@
 
-\bigskip \noindent~\citeauthor{yoonRealTimePerformanceAnalysis}~\cite{yoonRealTimePerformanceAnalysis} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
+\bigskip \noindent \citeauthor{yoonRealTimePerformanceAnalysis}~\cite{yoonRealTimePerformanceAnalysis} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
 They implement EtherCAT for real-time communication of distributed devices whereas this thesis uses Varan.
@@
 
-\bigskip \noindent~\citeauthor{sandstromVirtualizationTechnologiesEmbedded2013}~\cite{sandstromVirtualizationTechnologiesEmbedded2013}, and~\citeauthor{taccariEmbeddedRealTimeVirtualization2014}~\cite{taccariEmbeddedRealTimeVirtualization2014} and  discuss and compare the current state-of-the-art in virtualization technologies with a focus on embedded real-time platforms.
+\bigskip \noindent \citeauthor{cinqueVirtualizingMixedCriticalitySystems2022}~\cite{cinqueVirtualizingMixedCriticalitySystems2022} survey different virtualization approaches, including separation kernels, microkernels, enhanced general-purpose hypervisors, and lightweight virtualization solutions like containers and unikernels.
+Similarly, \citeauthor{sandstromVirtualizationTechnologiesEmbedded2013}~\cite{sandstromVirtualizationTechnologiesEmbedded2013}, and~\citeauthor{taccariEmbeddedRealTimeVirtualization2014}~\cite{taccariEmbeddedRealTimeVirtualization2014} discuss and compare the current state-of-the-art in virtualization technologies with a focus on embedded real-time platforms.
 The latter emphasizes, that research on real-time systems is focusing on virtualization and multicore scheduling and underlines that flexibility and cost are key metrics.
@@
 
-\bigskip \noindent
-\citeauthor{cinqueVirtualizingMixedCriticalitySystems2022}~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
-
-\citeauthor{reghenzaniRealTimeLinuxKernel2020}~\cite{reghenzaniRealTimeLinuxKernel2020}
-
-\bigskip \noindent~\citeauthor{maPerformanceTuningKVMbased}~\cite{maPerformanceTuningKVMbased}~\cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
+\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased}~\cite{maPerformanceTuningKVMbased}~\cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
 When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS.
@@
 
-\bigskip \noindent~\citeauthor{broskyShieldedProcessorsGuaranteeing2003}~\cite{broskyShieldedProcessorsGuaranteeing2003} report in their work about the shielded CPU model for enhancing real-time performance in symmetric multiprocessor systems.
+\bigskip \noindent \citeauthor{broskyShieldedProcessorsGuaranteeing2003}~\cite{broskyShieldedProcessorsGuaranteeing2003} report in their work about the shielded CPU model for enhancing real-time performance in symmetric multiprocessor systems.
 By dedicating one or more CPUs to high-priority processes and interrupts, the goal is to provide deterministic execution of real-time applications and interrupt responses.
@@
 
-\bigskip \noindent~\citeauthor{kiszkaLinuxRealTimeHypervisor}~\cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
+\bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor}~\cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
 Most importantly for this master's thesis, the author warns of starvation when the priorities of QEMU’s threads are raised and lifted into a real-time scheduling class.
@@
 
-\bigskip \noindent~\citeauthor{mckenneyRealTimeVs}~\cite{mckenneyRealTimeVs} addresses the overheads associated with real-time Linux when using KVM and QEMU.
+\bigskip \noindent \citeauthor{mckenneyRealTimeVs}~\cite{mckenneyRealTimeVs} addresses the overheads associated with real-time Linux when using KVM and QEMU.
 The author lists several sources of overhead, including memory locking to avoid page-fault latencies, increased overhead of locking and interrupts due to more-aggressive preemption, threaded interrupts that permit long-running interrupt handlers to be preempted by high-priority real-time processes, real-time task scheduling that requires global scheduling, high-resolution timers for tens-of-microseconds accuracy and precision, and preemptible RCU which has slightly higher read-side overhead than Classic RCU.
@@
 
-\bigskip \noindent~\citeauthor{deoliveiraDemystifyingRealTimeLinux}~\cite{deoliveiraDemystifyingRealTimeLinux} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
+\bigskip \noindent \citeauthor{deoliveiraDemystifyingRealTimeLinux}~\cite{deoliveiraDemystifyingRealTimeLinux} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
 This is the main reason why this tool was not used to measure the scheduling latency in this work.
@@
 The authors propose further enhancements, including the integration of real-time scheduling policies like EDF (Earliest Deadline First) and RM (Rate Monotonic)7.}
+
+\bigskip \noindent \textcolor{orange}{“Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane:
+	The paper “Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane provides an in-depth examination of Real Time Operating Systems (RTOS) and their critical role in embedded systems, particularly in telecommunication applications such as telephony, navigation, and military signaling systems.
+RTOS is essential for applications that require strict deadline constraints and the ability to handle multiple tasks simultaneously1.
+The paper highlights the key features of RTOS, including timing considerations, task prioritization, and communication and synchronization between tasks2.
+	The document delves into various design aspects necessary for developing an RTOS, such as task management, multitasking, context switching, task priority, and kernel functions3.
+It explains how tasks are created, prioritized, and managed within an RTOS, emphasizing the importance of efficient CPU utilization and the prevention of data corruption through mutual exclusion.
+The paper also discusses the implementation of RTOS, covering topics like programming language selection, task stack management, and the initialization process.
+	Furthermore, the paper presents a practical example of designing and implementing an RTOS for a Philips 8052 microcontroller, demonstrating the application of theoretical concepts in real-world scenarios.
+The successful implementation of applications such as a column controller, digital calculator, and dancing LEDs showcases the versatility and effectiveness of RTOS in managing complex embedded systems4.
+	Overall, the paper provides a comprehensive overview of RTOS, offering valuable insights into its architecture, design, and implementation, making it a significant contribution to the field of embedded systems5.}
+
+\bigskip \noindent \textcolor{orange}{Title: Hard Real Time Linux* Using Xenomai* on Intel® Multi-Core Processors12
+	Author: Amarpreet Singh Ugal, BIOS/Firmware Engineer, Intel Corporation
+	Summary: This document explores the implementation of hard real-time capabilities in Linux using Xenomai on Intel® multi-core processors.
+Traditional Linux is not a hard real-time operating system because it cannot guarantee that tasks will meet strict deadlines34.
+To address this, the document discusses the use of a high-priority real-time kernel that runs between the hardware and standard Linux.
+This real-time kernel executes real-time tasks and suspends normal Linux processes during this period56.
+The real-time scheduler treats the standard Linux kernel as an idle task, allowing real-time tasks to preempt normal processes78.
+Xenomai, which uses an interrupt pipeline from the ADEOS project, is highlighted as a solution for managing interrupts and ensuring minimal latency and jitter.
+The document also covers the installation and configuration of Xenomai on CentOS 5.2, including kernel patching and testing with various tools like switchbench, switchtest, cyclictest, clocktest, and latency.
+The detailed steps and configurations provided aim to help developers achieve hard real-time performance on Intel® Core™ i7 multi-core processors.}
+
+\bigskip \noindent \textcolor{orange}{The paper titled “Performance Evaluation of Xenomai 3” by Ching-Chun (Jim) Huang, Chan-Hsiang Lin, and Che-Kang Wu, presents an in-depth analysis of Xenomai 3, a real-time framework for Linux.
+Xenomai 3 operates either as a dual-kernel system alongside Linux or natively over mainline Linux kernels, enhanced by PREEMPT RT for stricter response times1.
+The study evaluates Xenomai 3’s performance, particularly in dual-kernel configuration, using ARM Cortex-A series benchmarks.
+The authors highlight Xenomai 3’s optimizations in RTOS API emulation, thread-to-thread communications, and reduced overhead.
+The paper compares Xenomai 3 with its predecessor, Xenomai 2, and the PREEMPT RT-enhanced Linux kernel, demonstrating superior maximum response times to external interrupts in dual-kernel setups.
+The study employs BeagleBone Black hardware and various kernel configurations to measure latencies under idle and CPU-stressed conditions2.
+Results indicate that Xenomai 3 generally outperforms Xenomai 2 and PREEMPT RT, except for a slight degradation in timer-IRQ latency under stress.
+The paper concludes that Xenomai 3 offers significant improvements in real-time performance and flexibility, making it a robust choice for real-time applications.}
 
@@
 
-\bigskip \noindent \textcolor{orange}{“Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane:
-	The paper “Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane provides an in-depth examination of Real Time Operating Systems (RTOS) and their critical role in embedded systems, particularly in telecommunication applications such as telephony, navigation, and military signaling systems.
-RTOS is essential for applications that require strict deadline constraints and the ability to handle multiple tasks simultaneously1.
-The paper highlights the key features of RTOS, including timing considerations, task prioritization, and communication and synchronization between tasks2.
-	The document delves into various design aspects necessary for developing an RTOS, such as task management, multitasking, context switching, task priority, and kernel functions3.
-It explains how tasks are created, prioritized, and managed within an RTOS, emphasizing the importance of efficient CPU utilization and the prevention of data corruption through mutual exclusion.
-The paper also discusses the implementation of RTOS, covering topics like programming language selection, task stack management, and the initialization process.
-	Furthermore, the paper presents a practical example of designing and implementing an RTOS for a Philips 8052 microcontroller, demonstrating the application of theoretical concepts in real-world scenarios.
-The successful implementation of applications such as a column controller, digital calculator, and dancing LEDs showcases the versatility and effectiveness of RTOS in managing complex embedded systems4.
-	Overall, the paper provides a comprehensive overview of RTOS, offering valuable insights into its architecture, design, and implementation, making it a significant contribution to the field of embedded systems5.}
-
 \bigskip \noindent \textcolor{red}{The paper titled “VM-Based Real-Time Services for Automotive Control Applications” is authored by Alejandro Masrur, Sebastian Drossler, Thomas Pfeuffer, and Samarjit Chakraborty.
@@
 The paper includes extensive experiments to illustrate the effectiveness of the proposed solutions.}
-
-\bigskip \noindent \textcolor{orange}{Title: Hard Real Time Linux* Using Xenomai* on Intel® Multi-Core Processors12
-	Author: Amarpreet Singh Ugal, BIOS/Firmware Engineer, Intel Corporation
-	Summary: This document explores the implementation of hard real-time capabilities in Linux using Xenomai on Intel® multi-core processors.
-Traditional Linux is not a hard real-time operating system because it cannot guarantee that tasks will meet strict deadlines34.
-To address this, the document discusses the use of a high-priority real-time kernel that runs between the hardware and standard Linux.
-This real-time kernel executes real-time tasks and suspends normal Linux processes during this period56.
-The real-time scheduler treats the standard Linux kernel as an idle task, allowing real-time tasks to preempt normal processes78.
-Xenomai, which uses an interrupt pipeline from the ADEOS project, is highlighted as a solution for managing interrupts and ensuring minimal latency and jitter.
-The document also covers the installation and configuration of Xenomai on CentOS 5.2, including kernel patching and testing with various tools like switchbench, switchtest, cyclictest, clocktest, and latency.
-The detailed steps and configurations provided aim to help developers achieve hard real-time performance on Intel® Core™ i7 multi-core processors.}
 
@@
 This research provides valuable insights into the suitability of different operating systems for multi-robot applications, contributing to the broader field of robotics and autonomous systems.}
-
-\bigskip \noindent \textcolor{green}{Title: The Real-Time Linux Kernel: A Survey on PREEMPT\_RT12
-	Authors: Federico Reghenzani, Giuseppe Massari, and William Fornaciari from Politecnico di Milano3
-	Summary: This survey explores the evolution and current state of the PREEMPT\_RT patch for the Linux kernel, which aims to enhance the real-time capabilities of Linux.
-The authors discuss the increasing demand for real-time applications and the challenges posed by using Commercial-Off-The-Shelf (COTS) hardware in such environments.
-The PREEMPT\_RT patch modifies the existing Linux kernel to improve predictability and reduce latencies, making it suitable for real-time applications.
-The survey covers various approaches to real-time Linux, including cokernel approaches like RTLinux, Xenomai, and RTAI, and compares them with the PREEMPT\_RT patch4.
-The authors highlight the benefits of using Linux in real-time systems, such as the availability of rich hardware support and a well-established programming environment.
-They also discuss the challenges of integrating real-time capabilities into a general-purpose operating system and the ongoing efforts to address these issues.
-The survey concludes with an overview of industrial and scientific use cases that have benefited from the PREEMPT\_RT patch, emphasizing its significance in fields like aerospace, automotive, and high-performance computing.}
-
-\bigskip \noindent \textcolor{orange}{The paper titled “Performance Evaluation of Xenomai 3” by Ching-Chun (Jim) Huang, Chan-Hsiang Lin, and Che-Kang Wu, presents an in-depth analysis of Xenomai 3, a real-time framework for Linux.
-Xenomai 3 operates either as a dual-kernel system alongside Linux or natively over mainline Linux kernels, enhanced by PREEMPT RT for stricter response times1.
-The study evaluates Xenomai 3’s performance, particularly in dual-kernel configuration, using ARM Cortex-A series benchmarks.
-The authors highlight Xenomai 3’s optimizations in RTOS API emulation, thread-to-thread communications, and reduced overhead.
-The paper compares Xenomai 3 with its predecessor, Xenomai 2, and the PREEMPT RT-enhanced Linux kernel, demonstrating superior maximum response times to external interrupts in dual-kernel setups.
-The study employs BeagleBone Black hardware and various kernel configurations to measure latencies under idle and CPU-stressed conditions2.
-Results indicate that Xenomai 3 generally outperforms Xenomai 2 and PREEMPT RT, except for a slight degradation in timer-IRQ latency under stress.
-The paper concludes that Xenomai 3 offers significant improvements in real-time performance and flexibility, making it a robust choice for real-time applications.}
 

### 2024-08-16 Masterarbeit_Robotik_Pamuk.tex
cite order

@@
 
+~\cite{WhatRealtimeLinuxa}
 ~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
-~\cite{Tracecmd}
-~\cite{KernelShark}
+~\cite{WelcomeYoctoProject}
 ~\cite{XenomaiXenomai}
@@
 ~\cite{EngineeringToolLASAL}
-~\cite{WelcomeYoctoProject}
+~\cite{Tracecmd}
+~\cite{KernelShark}
 ~\cite{QEMU}
@@
 
+~\cite{canbazPerformanceAnalysisRealtime2022}
+~\cite{malallahComprehensiveStudyKernel2021}
+~\cite{wangRealtimeEmbeddedSystems2017}
+~\cite{buttazzoHardRealtimeComputing2024}
+~\cite{lipariRealTimeSchedulingHard}
+~\cite{perneelRealtimeCapabilitiesStandard2015}
+~\cite{reghenzaniRealTimeLinuxKernel2020}
+~\cite{adamRealTimePerformanceResponse2021}
 ~\cite{RealTimePerformanceTuning2022}
+~\cite{yoonRealTimePerformanceAnalysis}
+~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
+~\cite{sandstromVirtualizationTechnologiesEmbedded2013}
+~\cite{taccariEmbeddedRealTimeVirtualization2014}
+~\cite{javierperezHowRealTime2022}
+~\cite{guStateoftheArtSurveyRealTime2012}
+~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
+~\cite{scordinoRealTimeVirtualizationIndustrial2020}
 ~\cite{maPerformanceTuningKVMbased}
+~\cite{junzhangPerformanceAnalysisKVMBased2010}
+~\cite{broskyShieldedProcessorsGuaranteeing2003}
 ~\cite{kiszkaLinuxRealTimeHypervisor}
+~\cite{mckenneyRealTimeVs}
 ~\cite{deoliveiraDemystifyingRealTimeLinux}
-~\cite{mckenneyRealTimeVs}
-~\cite{yoonRealTimePerformanceAnalysis}
 ~\cite{huang2015performance}
-~\cite{junzhangPerformanceAnalysisKVMBased2010}
-~\cite{perneelRealtimeCapabilitiesStandard2015}
-~\cite{scordinoRealTimeVirtualizationIndustrial2020}
-
-~\cite{broskyShieldedProcessorsGuaranteeing2003}
-~\cite{adamRealTimePerformanceResponse2021}
 ~\cite{pielAsymmetricSchedulingLoad2006}
-~\cite{malallahComprehensiveStudyKernel2021}
 ~\cite{queirozTestingLimitsGeneralpurpose2023}
-~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
-~\cite{guStateoftheArtSurveyRealTime2012}
-~\cite{javierperezHowRealTime2022}
 ~\cite{kirovaImpactModernVirtualization2019}
 ~\cite{REALTIMEOPERATING}
-~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
 ~\cite{masrurVMBasedRealTimeServices2010}
-
 ~\cite{HardRealTime}
@@
 ~\cite{jiangRealTimeSystemManycore}
-~\cite{wangRealtimeEmbeddedSystems2017}
-~\cite{lipariRealTimeSchedulingHard}
-~\cite{canbazPerformanceAnalysisRealtime2022}
-~\cite{reghenzaniRealTimeLinuxKernel2020}
 \clearpage

### 2024-08-16 Masterarbeit_Robotik_Pamuk.tex
remove red and orange text

@@
 Instead, the \texttt{latency} tool was used, as was determined in Chapter \ref{sec:application_context}.
-
-\bigskip \noindent \textcolor{orange}{Asymmetric Scheduling and Load Balancing for Real-Time on Linux SMP
-	Authors: Eric Piel, Philippe Marquet, Julien Soula, and Jean-Luc Dekeyser12
-	The paper presents ARTiS, a real-time extension of the GNU/Linux scheduler designed for SMP (Symmetric Multi-Processors) systems3.
-ARTiS aims to guarantee the preemption of processors for real-time tasks by assigning specific processors to these tasks4.
-This system ensures low latency by migrating non-preemptible tasks away from real-time processors5.
-ARTiS also incorporates load-balancing strategies to fully utilize SMP systems without wasting resources.
-The implementation of ARTiS involves modifications to the Linux scheduler, resulting in significant improvements in latency compared to the standard Linux scheduler6.
-The paper details the performance evaluation of ARTiS, highlighting its ability to maintain low interrupt latency even under high system load, making it suitable for hard real-time applications.
-The authors propose further enhancements, including the integration of real-time scheduling policies like EDF (Earliest Deadline First) and RM (Rate Monotonic)7.}
-
-\bigskip \noindent \textcolor{orange}{“Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane:
-	The paper “Real Time Operating Systems: A Complete Overview” by Mrinal Parikshit Chandane provides an in-depth examination of Real Time Operating Systems (RTOS) and their critical role in embedded systems, particularly in telecommunication applications such as telephony, navigation, and military signaling systems.
-RTOS is essential for applications that require strict deadline constraints and the ability to handle multiple tasks simultaneously1.
-The paper highlights the key features of RTOS, including timing considerations, task prioritization, and communication and synchronization between tasks2.
-	The document delves into various design aspects necessary for developing an RTOS, such as task management, multitasking, context switching, task priority, and kernel functions3.
-It explains how tasks are created, prioritized, and managed within an RTOS, emphasizing the importance of efficient CPU utilization and the prevention of data corruption through mutual exclusion.
-The paper also discusses the implementation of RTOS, covering topics like programming language selection, task stack management, and the initialization process.
-	Furthermore, the paper presents a practical example of designing and implementing an RTOS for a Philips 8052 microcontroller, demonstrating the application of theoretical concepts in real-world scenarios.
-The successful implementation of applications such as a column controller, digital calculator, and dancing LEDs showcases the versatility and effectiveness of RTOS in managing complex embedded systems4.
-	Overall, the paper provides a comprehensive overview of RTOS, offering valuable insights into its architecture, design, and implementation, making it a significant contribution to the field of embedded systems5.}
-
-\bigskip \noindent \textcolor{orange}{Title: Hard Real Time Linux* Using Xenomai* on Intel® Multi-Core Processors12
-	Author: Amarpreet Singh Ugal, BIOS/Firmware Engineer, Intel Corporation
-	Summary: This document explores the implementation of hard real-time capabilities in Linux using Xenomai on Intel® multi-core processors.
-Traditional Linux is not a hard real-time operating system because it cannot guarantee that tasks will meet strict deadlines34.
-To address this, the document discusses the use of a high-priority real-time kernel that runs between the hardware and standard Linux.
-This real-time kernel executes real-time tasks and suspends normal Linux processes during this period56.
-The real-time scheduler treats the standard Linux kernel as an idle task, allowing real-time tasks to preempt normal processes78.
-Xenomai, which uses an interrupt pipeline from the ADEOS project, is highlighted as a solution for managing interrupts and ensuring minimal latency and jitter.
-The document also covers the installation and configuration of Xenomai on CentOS 5.2, including kernel patching and testing with various tools like switchbench, switchtest, cyclictest, clocktest, and latency.
-The detailed steps and configurations provided aim to help developers achieve hard real-time performance on Intel® Core™ i7 multi-core processors.}
-
-\bigskip \noindent \textcolor{orange}{The paper titled “Performance Evaluation of Xenomai 3” by Ching-Chun (Jim) Huang, Chan-Hsiang Lin, and Che-Kang Wu, presents an in-depth analysis of Xenomai 3, a real-time framework for Linux.
-Xenomai 3 operates either as a dual-kernel system alongside Linux or natively over mainline Linux kernels, enhanced by PREEMPT RT for stricter response times1.
-The study evaluates Xenomai 3’s performance, particularly in dual-kernel configuration, using ARM Cortex-A series benchmarks.
-The authors highlight Xenomai 3’s optimizations in RTOS API emulation, thread-to-thread communications, and reduced overhead.
-The paper compares Xenomai 3 with its predecessor, Xenomai 2, and the PREEMPT RT-enhanced Linux kernel, demonstrating superior maximum response times to external interrupts in dual-kernel setups.
-The study employs BeagleBone Black hardware and various kernel configurations to measure latencies under idle and CPU-stressed conditions2.
-Results indicate that Xenomai 3 generally outperforms Xenomai 2 and PREEMPT RT, except for a slight degradation in timer-IRQ latency under stress.
-The paper concludes that Xenomai 3 offers significant improvements in real-time performance and flexibility, making it a robust choice for real-time applications.}
-
-\bigskip \noindent \textcolor{red}{Title: A Comprehensive Study of Kernel (Issues and Concepts) in Different Operating Systems12
-	Authors: Hayfaa Subhi Malallah, Subhi R. M. Zeebaree, Rizgar R. Zebari, Mohammed A. M. Sadeeq, Zainab Salih Ageed, Ibrahim Mahmood Ibrahim, Hajar Maseeh Yasin, Karwan Jameel Merceedi
-	Summary: This paper provides an in-depth comparative analysis of various operating systems (OS) focusing on their kernels, which are the core components managing system resources and providing essential services.
-The study examines the strengths and limitations of popular OS such as iOS, Android, Mac, Windows, and Linux3.
-It highlights that Linux, Android, and Windows 10 are particularly noted for their stability, compatibility, and reliability4.
-The paper discusses the evolution of OS kernels, including microkernels, monolithic kernels, and hybrid kernels, and their respective advantages and challenges.
-It also addresses modern developments in OS technology, such as multicore processors, cloud computing, and the Internet of Things (IoT), and their impact on kernel design and performance.
-The authors emphasize the importance of security in kernel development, noting that vulnerabilities can lead to significant system compromises.
-The paper concludes with a discussion on the future directions of OS kernel research and development, advocating for continuous improvement in performance, security, and adaptability to emerging technologies.}
-
-\bigskip \noindent \textcolor{red}{Title: Testing the limits of general-purpose hypervisors for real-time control systems
-	Authors: Rui Queiroz, Tiago J. Cruz, Paulo Simões (University of Coimbra)
-	Summary: This paper explores the feasibility of using general-purpose off-the-shelf hypervisors to virtualize real-time control systems within the context of Industry 4.0.
-The authors address the need for flexibility, security, and resilience in automation infrastructures while maintaining real-time requirements.
-They discuss the potential benefits of virtualizing servers and control devices in Industrial and Automation Control Systems (IACS), but also highlight the challenges due to the focus of general-purpose hypervisors on maximizing throughput, often at the expense of determinism and increased latency.
-	The paper presents an empirical evaluation of various hypervisors, concluding that while some are capable of handling typical real-time workloads, this capability cannot be generalized to all types of real-time systems.
-The authors provide a comprehensive analysis of cyber-physical systems (CPS), control devices, and real-time systems, including their latency and jitter requirements.
-They also delve into the specifics of real-time operating systems (RTOS) and mixed-criticality systems, and review different virtualization techniques, comparing commercial off-the-shelf hypervisors with real-time hypervisors.
-	The study emphasizes the importance of ensuring both spatial and temporal isolation in virtualized environments to maintain predictable real-time performance.
-It also discusses the role of scheduling policies in real-time applications and the challenges of I/O sharing between multiple virtual machines.
-The authors conclude that while general-purpose hypervisors show promise, further research and development are needed to fully realize their potential in real-time control systems.}
-
-\bigskip \noindent \textcolor{red}{Impact of Modern Virtualization Methods on Timing Precision and Performance of High-Speed Applications
-	Authors: Veronika Kirova, Kirill Karpov, Eduard Siemens, Irina Zander, Oksana Vasylenko, Dmitry Kachan, Sergii Maksymov
-	This research investigates the precision of timing methods and their efficiency in various virtual environments, focusing on their impact on high-speed network applications.
-The study examines how timer hardware is shared among CPU- and I/O-bound tasks on both virtualized and bare operating systems.
-By replacing timing methods within an application for estimating available path bandwidth, the authors analyze the impact on estimation accuracy.
-The findings highlight that timer overhead and precision are crucial for high-performance network applications, with low-precision timing methods leading to degradation in virtual environments.
-The paper confirms that using precise timing operations and AvB estimation methods can overcome inefficiencies in standard time-related operations.
-The research evaluates five different virtual environments to determine the most optimal for high-speed network applications, providing a comprehensive analysis of the impact of virtualization on timing precision and performance.}
-
-\bigskip \noindent \textcolor{red}{The paper titled “VM-Based Real-Time Services for Automotive Control Applications” is authored by Alejandro Masrur, Sebastian Drossler, Thomas Pfeuffer, and Samarjit Chakraborty.
-It explores the use of Virtual Machines (VMs) to provide real-time services in automotive control applications12.
-The authors utilize the Xen hypervisor to design a real-time control loop on top of a virtualization layer3.
-They first analyze a typical Xen configuration and identify issues when used for real-time applications4.
-The paper highlights that the SEDF (Simple Earliest Deadline First) scheduler in Xen can be improved with minimal modifications to enhance worst-case performance5.
-To further reduce latency and jitter, the authors propose a new scheduler for Xen that prioritizes real-time VMs using a fixed-priority policy6.
-This approach ensures that timing constraints in automotive control applications are met7.
-The paper includes extensive experiments to illustrate the effectiveness of the proposed solutions.}
-
-\bigskip \noindent \textcolor{red}{Title: Performance Assessment of Linux Kernels with PREEMPT\_RT on ARM-Based Embedded Devices
-	Authors: George K. Adam, Nikos Petrellis, Lambros T. Doulos
-	Summary: This research investigates the real-time performance of Linux kernels and distributions with the PREEMPT\_RT real-time patch on ARM-based embedded devices, specifically BeagleBoard and Raspberry Pi.
-The study aims to provide novel insights into Linux real-time performance by evaluating various metrics such as response latency, periodic task execution, and maximum sustained frequency using heuristic methods and the cyclictest benchmark tool.
-The authors introduce a new response task model and develop specific software measurement modules to assess the performance.
-	The results demonstrate that the PREEMPT\_RT patch significantly improves real-time performance compared to default Linux kernels, particularly in reducing worst-case latencies.
-This makes ARM-based devices running Linux with PREEMPT\_RT more suitable for time-sensitive embedded control systems.
-The research also highlights the growing use of Linux in embedded systems for real-time applications, driven by its open-source nature, performance, and adaptability to various hardware systems.
-The methodology and findings of this study could be applied to other Linux-based real-time platforms, contributing to the broader field of real-time systems research.}
-
-\bigskip \noindent \textcolor{red}{Title: Real-Time I/O System for Many-core Embedded Systems 1 Author: Zhe Jiang
-	In this thesis, Zhe Jiang addresses the critical need for time predictability in modern real-time embedded systems, particularly focusing on I/O operations.
-The proposed solution, BlueIO, is a hardware-implemented real-time I/O virtualization system designed to meet the requirements of predictability, timing accuracy, enhanced performance, scalability, parallel access, and isolation simultaneously2.
-BlueIO integrates key functionalities of I/O virtualization and low-layer I/O drivers through the Virtualized Complicated Device Controller (VCDC) and a clock cycle-level timing-accurate I/O controller, the GPIO Command Processor (GPIOCP)3.
-This system supports abstract virtualized access to I/O devices from the software domain, providing isolation and concurrent access to I/O operations, thereby improving performance45.
-The thesis also introduces BlueVisor, a scalable real-time hardware hypervisor built upon BlueIO, enabling predictable virtualization on CPU, memory, and I/O, along with fast interrupt handling and inter-virtual machine communication67.
-The work demonstrates significant improvements in I/O performance and low running costs across different operating systems, with a hardware consumption analysis showing linear scalability with the number of CPUs and I/O devices.}
-
-\bigskip \noindent \textcolor{red}{Title: Real-Time Operating Systems1
-	Author: Jiacun Wang, Monmouth University
-	Summary: This chapter provides an in-depth overview of Real-Time Operating Systems (RTOS), which are crucial for embedded systems requiring timely and predictable responses.
-The chapter begins by outlining the main functions of general-purpose operating systems, such as process management, memory management, interrupts management, multitasking, file system management, and I/O management.
-It then delves into the specific characteristics of RTOS kernels, emphasizing the need for predictable timing behavior, priority scheduling, intertask communication, and resource sharing.
-The chapter also discusses clocks and timers, asynchronous I/O, and memory locking as essential features of RTOS.
-Finally, it provides examples of widely used RTOS products, including LynxOS, OSE, QNX, VxWorks, and Windows Embedded Compact, highlighting their unique features and applications in various industries such as aerospace, automotive, and telecommunications.}
-
-\bigskip \noindent \textcolor{red}{Title: Real-Time Scheduling: From Hard to Soft Real-Time Systems1
-	Authors: Giuseppe Lipari (Université de Lille) and Luigi Palopoli (Università di Trento)
-	Summary: This paper explores the evolution of real-time systems, traditionally divided into hard real-time and soft real-time categoriesHard real-time systems are critical, where missing a deadline can lead to catastrophic consequences, such as in automotive brake systems2.
-Soft real-time systems, on the other hand, aim to optimize the quality of service, like multimedia players that can tolerate occasional delays without severe impact.
-	The authors argue that the boundary between these two categories is becoming increasingly blurred.
-They introduce the concept of time-predictability and criticality to understand the origins of real-time deadlines.
-The paper presents the resource reservation framework as a method for scheduling both hard and soft real-time tasks.
-This framework enhances system robustness and resource utilization by isolating the temporal behavior of tasks.
-	The paper also discusses the Constant Bandwidth Server (CBS) algorithm, which integrates with the EDF scheduler to manage execution budgets and periods, ensuring temporal isolation3.
-The authors highlight the importance of designing systems that remain consistent even when deadlines are missed, emphasizing the need for robust software development practices.
-	In conclusion, the paper demonstrates how resource reservation techniques can be applied to classical control systems, improving performance and robustness4.
-The authors provide a comprehensive overview of existing techniques and tools for soft real-time scheduling and analysis, making a strong case for the integration of these methods in modern real-time systems5.}
-
-\bigskip \noindent \textcolor{red}{The paper titled “Performance analysis of real-time and general-purpose operating systems for path planning of the multi-robot systems” by Seçkin Canbaz and Gökhan Erdemir explores the performance differences between real-time operating systems (RTOS) and general-purpose operating systems (GPOS) in the context of multi-robot systems.
-The study focuses on two Linux distributions, Ubuntu and Pardus, and evaluates their effectiveness as both GPOS and RTOS for path planning tasks12.
-The authors highlight the critical distinction between GPOS and RTOS, emphasizing that RTOS can preempt low-priority tasks with high-priority ones, even during kernel calls, which is not possible in GPOS.
-	The research involves using the turtlesim simulation environment to perform path tracking tasks with robot groups of varying sizes.
-The experiments compare CPU usage and processing times across the different operating systems.
-Results indicate that Ubuntu GPOS generally outperforms other configurations in terms of lower CPU usage and shorter processing times3.
-However, RTOS configurations, while slower, offer higher system security and reduced error margins due to their strict time management.
-	The study concludes that while GPOS may be preferable for performance-centric applications, RTOS is essential for tasks with high time dependency.
-The authors suggest that the choice between GPOS and RTOS should consider factors such as potential failure conditions, process security, and processing time requirements.
-This research provides valuable insights into the suitability of different operating systems for multi-robot applications, contributing to the broader field of robotics and autonomous systems.}
 

### 2024-08-17 Masterarbeit_Robotik_Pamuk.tex
clearpage
Problem and Task Definition

@@
 \citeauthor{adamRealTimePerformanceResponse2021}~\cite{adamRealTimePerformanceResponse2021} applied the PREEMPT\_RT patch on both Raspberry Pi 3 and BeagleBone Black and achieved significantly lower latencies compared to standard Linux kernels.
-
-\bigskip \noindent In the course of writing this thesis, the white paper "Real-Time Performance Tuning Best Practice Guidelines for KVM-Based Virtual Machines"~\cite{RealTimePerformanceTuning2022} by Intel has been instrumental.
-Given that the working computer used for this work is equipped with an Intel processor, the guidelines presented in this document were particularly relevant and valuable for optimizing real-time performance in KVM-based virtual machines.
-Chapter \ref{cha:real-time_tuning} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
-The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads.
 
@@
 
+\bigskip \noindent In the course of writing this thesis, the white paper "Real-Time Performance Tuning Best Practice Guidelines for KVM-Based Virtual Machines"~\cite{RealTimePerformanceTuning2022} by Intel has been instrumental.
+Given that the working computer used for this work is equipped with an Intel processor, the guidelines presented in this document were particularly relevant and valuable for optimizing real-time performance in KVM-based virtual machines.
+Chapter \ref{cha:real-time_tuning} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
+The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads.
+
 \bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor}~\cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
@@
 
+\clearpage
 \section{Problem and Task Definition}
-
+As stated previously, virtualization of real-time systems has many advantages in comparison to traditional hardware-based approaches.
+Solutions based on pure hardware are often built for a distinct purpose and are hence limited in flexibility.
+It is difficult to adapt the hardware topology to new requirements or new applications.
+Moreover, during periods when the demand is lower, expensive hardware may sit idle, leading to inefficient use of resources and a poor return on investment.
+Another crucial factor is of course the high cost factor associated with procuring and maintaining separate hardware for each real-time system.
+Especially when organizations scale their operations, they have to add more physical machines, and the expenses rise rapidly.
+Besides, one needs physical access to the hardware for any updates or maintenance, which is hard to achieve, especially for systems in remote or difficult-to-access locations.
+This can lead to system downtime and may not only disrupt the operations but can also result in lost productivity and potential revenue.
+
+Virtualization promises overcoming these challenges
+
+\begin{comment}
+1. Flexibility and Resource Utilization
+
+Virtualization allows for more efficient use of hardware resources by enabling multiple real-time systems to run on a single physical machine.
+This improved resource utilization can lead to significant cost savings and increased flexibility in system design.
+
+Evidence: A study by Cucinotta et al.
+(2018) demonstrated that virtualization techniques could effectively support real-time applications while maintaining their timing constraints.
+They showed that proper scheduling and resource allocation in virtualized environments could provide performance comparable to bare-metal systems.
+
+2. Isolation and Security
+
+Virtualization provides isolation between different real-time systems, enhancing security and reliability.
+Each virtualized system can operate independently, reducing the risk of interference or security breaches between applications.
+
+Evidence: Research by Lackorzynski et al.
+(2017) proposed a virtualization architecture for mixed-criticality systems, showing how virtualization can effectively isolate different security levels and criticality domains in real-time systems.
+
+3. Scalability
+
+Virtualized real-time systems can be more easily scaled up or down based on demand, allowing for more efficient resource allocation and improved system responsiveness.
+
+Evidence: A paper by Oliveira et al.
+(2019) presented a scalable virtualization solution for real-time systems in cloud environments, demonstrating how virtualization can enable dynamic resource allocation for real-time tasks.
+
+4. Cost-effectiveness
+
+By consolidating multiple real-time systems onto fewer physical machines, virtualization can significantly reduce hardware costs, power consumption, and maintenance expenses.
+
+Evidence: A study by Masrur et al.
+(2016) showed that virtualization could reduce the number of physical processors needed for real-time systems while maintaining timing guarantees, leading to potential cost savings.
+
+5. Easier System Management and Maintenance
+
+Virtualization simplifies system management tasks such as backup, recovery, and updates. It allows for easier migration of real-time systems between different hardware platforms.
+
+Evidence: Research by Becker et al.
+(2020) demonstrated how virtualization could facilitate the deployment and management of real-time systems in industrial environments, reducing downtime and maintenance costs.
+
+Despite these advantages, it's important to note that virtualizing real-time systems comes with its own challenges, particularly in ensuring that timing constraints are met consistently in a shared environment.
+
+Evidence: A comprehensive survey by Aichouch et al.
+(2013) discussed both the benefits and challenges of real-time virtualization, highlighting areas where further research was needed to fully realize the potential of virtualized real-time systems.
+
+In conclusion, virtualization offers significant advantages for real-time systems in terms of flexibility, resource utilization, scalability, and cost-effectiveness.
+However, careful consideration must be given to the specific requirements of each real-time application to ensure that virtualization can meet its timing and performance needs.
+
+Would you like me to elaborate on any specific aspect of virtualization in real-time systems or provide more information on the challenges and potential solutions?
+\end{comment}
+
+\clearpage
 \section{Objective}

### 2024-08-18 Masterarbeit_Robotik_Pamuk.tex
Problem and Task Definition++

--- 2024-08-16 Masterarbeit_Robotik_Pamuk.tex
@@
 	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal through real-time performance tunings, focussing on latency requirements set by SIGMATEK.
-	      These modifications are subject of Chapter~\ref{cha:real-time_tuning}.
+	      These modifications are subject of Chapter~\ref{cha:real-time_tunings}.
 	\item The latency tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to use by SIGMATEK.
@@
 
+\bigskip \noindent In the course of writing this thesis, the white paper "Real-Time Performance Tuning Best Practice Guidelines for KVM-Based Virtual Machines"~\cite{RealTimePerformanceTuning2022} by Intel has been instrumental.
+Given that the working computer used for this work is equipped with an Intel processor, the guidelines presented in this document were particularly relevant and valuable for optimizing real-time performance in KVM-based virtual machines.
+Chapter \ref{cha:real-time_tunings} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
+The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads.
+
 \bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor}~\cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
@@
 
+\clearpage
 \section{Problem and Task Definition}
+As stated previously, virtualization of real-time systems has many advantages in comparison to traditional hardware-based approaches.
+Solutions based on pure hardware are often built for a distinct purpose and are hence limited in flexibility.
+It is difficult to adapt the hardware topology to new requirements or new applications.
+Moreover, during periods when the demand is lower, expensive hardware may sit idle, leading to inefficient use of resources and a poor return on investment~\cite{gabrielResourceAwareParameterTuning2019}.
+Another crucial factor is of course the high cost factor associated with procuring and maintaining separate hardware for each real-time system~\cite{bhardwajVirtualizationCloudComputing2021}.
+Especially when organizations scale their operations, they have to add more physical machines, and the expenses rise rapidly~\cite{abbasiExploringOpenStackScalable2023}.
+Besides, one needs physical access to the hardware for any updates or maintenance, which is hard to achieve, especially for systems in remote or difficult-to-access locations.
+This can lead to system downtime and may not only disrupt the operations but can also result in lost productivity and potential revenue~\cite{mahfoudRealTimePredictiveMaintenanceBased2024}.
+
+\bigskip \noindent Virtualization promises to overcome these challenges \cite{queirozContainerbasedVirtualizationRealtime2024} \cite{bhardwajVirtualizationCloudComputing2021a} \cite{cinqueEvaluatingVirtualizationFog2023}.
+However, virtualization also has its own challenges and pitfalls.
+The additional layers of abstraction can lead to increased overhead and latency \cite{casiniLatencyAnalysisVirtualization2021} \cite{zhangEvaluatingOptimizingVirtualization2010}.
+This is not desired and can be particularly problematic for real-time systems that require strict timing constraints.
+Making sure that tasks run predictably and on time in a virtualized environment, can be challenging.
+The hypervisor and other virtual components can cause variations in how long tasks take to execute \cite{garcia-vallsChallengesRealtimeVirtualization2014}.
+When there are multiple virtual machines (VM) sharing the same physical resources, the VMs can affect each other's performance and consequently violate real-time constraints \cite{queirozContainerbasedVirtualizationRealtime2024}.
+
+\bigskip \noindent In this master’s thesis, the task is the virtualization of a real-time operating system to control a robot, with a focus on compliance with real-time determinism.
+Chapter \ref{cha:methodology} explains in detail the procedure, how this was tackled.
+As already established in Section~\ref{sec:application_context}, Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3 and virtualized through QEMU/KVM.
+These components are briefly described in Chapter \ref{cha:theoretical_foundations}.
+The initial latency values are measured with the \texttt{latency} program of the Xenomai tool suite.
+The initial situation is described in Section \ref{sec:starting_situation}.
+In a nutshell, there is a significant initial gap in latency statistics between the virtualized Salamander 4 and the Salamander 4 on bare metal.
+For this reason, an extensive tuning process is carried out to achieve real-time performance and determinism.
+These modifications involve configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS (guest) itself.
+The individual configurations are discussed in detail in Chapter \ref{cha:real-time_tunings} and the modifications are justified with a clear explanation.
 
 \section{Objective}
-
-The main objective of this work is to create a real-time robot control platform that integrates Salamander OS, Xenomai, QEMU and PCV-522 in the Yocto environment.
-
-\clearpage
-
-\chapter{Theoretical Foundations}
-\section{Yocto}\label{sec:yocto}
-
-\section{QEMU}\label{sec:qemu}
-
-\section{Xenomai}\label{sec:xenomai}
-\noindent Xenomai 3~\cite{XenomaiXenomai} is an open-source real-time framework that offers two paths to real-time performance.
-The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
-Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
-	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
-	\label{fig:cobalt}
-\end{figure}
-
-\noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
-Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
-	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
-	\label{fig:mercury}
-\end{figure}
-
-\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
-
-\section{VARAN-Bus}\label{sec:varan}
-
-\section{Salamander 4}\label{cha:salamander4}
-This chapter briefly describes the Salamander 4 operating system by SIGMATEK.
-
-\subsection{Structure}
-\noindent Salamander 4 is the proprietary operating system of SIGMATEK.
-It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment~\cite{XenomaiXenomai}.
-Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture.
-The real-time behaviour is achieved through the use of Symmetric Multi-Processing (SMP) and Preemptive Scheduling (PREEMPT).
-In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
-The output of the command~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
-
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={System information},label={output:uname_a_virt}]
-root@sigmatek-core2:~# uname -a
-Linux sigmatek-core2 5.15.94 #1 SMP PREEMPT IRQPIPE Tue Feb 14 18:18:05 UTC 2023 x86_64 GNU/Linux
-\end{lstlisting}
-\end{minipage}
-
-\noindent Salamander 4 is powered by SIGMATEK's CP 841~\cite{CPUEinheitenSIGMATEK} and is comprised of the following software modules:
-
-\begin{itemize}
-	\item \textbf{Operating system}: The operating system in a LASAL CPU manages the hardware and software resources of the system.
-It is provided in a completely PC-compatible manner, working with a standard PC BIOS.
-	\item \textbf{Loader}: The loader is a part of the operating system that is responsible for loading programs from executables into memory, preparing them for execution and then executing them.
-	\item \textbf{Hardware classes}: Hardware classes in LASAL represent the different types of hardware components that can be controlled by the LASAL CPU.
-They provide a way to organize and manage the hardware components in a modular and reusable manner.
-The graphical hardware editor in LASAL allows for a true-to-detail simulation of the actual hardware.
-	\item \textbf{Application}: Applications are developed using LASAL CLASS 2~\cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
-\end{itemize}
-
-\noindent These modules and the interfaces (indicated by an arrow) between them are shown in Figure~\ref{fig:lasal_cpu}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
-	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU}
-	\label{fig:lasal_cpu}
-\end{figure}
-
-\subsection{Memory Management}
-\noindent For the sake of completeness, Figure~\ref{fig:memory_management} displays the memory management of Salamander 4.
-LRT stands for Lasal Runtime and creates an execution environment where applications developed using the LASAL Class 2 can run, providing defined real-time functions, data types, and other constructs tailored for real-time programming.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/RAM_Memory_management.png}
-	\caption[Memory Management]{Memory Management}
-	\label{fig:memory_management}
-\end{figure}
-
-\clearpage
-
-\chapter{Methodology}\label{cha:methodology}
-
-This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
-
-\bigskip \noindent Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
-It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
-Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
-
-\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
-It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
-By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
-
-\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel~\cite{XenomaiXenomai}.
-It enables low-latency and deterministic execution of time-critical tasks.
-Xenomai 3 introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
-A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
-The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
-
-The system configuration is shown in Table~\ref{tab:testbed_configuration}
-
-\begin{table}[H]
-	\centering
-	\caption[System configuration]{System configuration}
-	\label{tab:testbed_configuration}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|l|l|}
-			\hline
-			\textbf{CPU}    & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
-			\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB           \\ \hline
-			\textbf{GPU}    & NVIDIA RTX A500 Laptop GPU                             \\ \hline
-			\textbf{BIOS}   & Dell Version 1.12.0                                    \\ \hline
-			\textbf{OS}     & Ubuntu 22.04.4 LTS                                     \\
-			\hline
-		\end{tabular}}
-\end{table}
-
-\noindent Figure~\ref{fig:lstopo} is the output of the~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
-	\caption[Hardware topology]{Hardware topology}
-	\label{fig:lstopo}
-\end{figure}
-
-\clearpage
-
-\chapter{Initial Real-Time Latency}\label{cha:initial-real-time-latency}
-
+The goal is to bring the latency of the virtualized Salamander 4 closer to that of the bare metal system, thereby ensuring deterministic behavior under real-time constraints.
+
+\section{Starting Situation}\label{sec:starting_situation}
 As a starting point, initial latency values of both the bare metal and virtualization versions were measured with the \texttt{latency} tool of the xenomai tool suite.
@@
 
-\section{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
+\subsection{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
 The output of the command~\texttt{uname -a} for Salamander 4 bare metal is shown in Code~\ref{output:uname_a_hw}.
@@
 
-\section{Salamander 4 Virtualization}\label{sec:salamander4-virtualization}
+\subsection{Salamander 4 Virtualization}\label{sec:salamander4-virtualization}
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
@@
 This script is used to start QEMU with required parameters to boot Salamamder 4 OS.
-It will be adjusted in Chapter~\ref{cha:real-time_tuning} in order to accompany real-time performance tunings.
+It will be adjusted in Chapter~\ref{cha:real-time_tunings} in order to accompany real-time performance tunings.
 
@@
 
-\chapter{Real-Time Performance Tuning}\label{cha:real-time_tuning}
+\clearpage
+
+The main objective of this work is to create a real-time robot control platform that integrates Salamander OS, Xenomai, QEMU and PCV-522 in the Yocto environment.
+
+\clearpage
+
+\chapter{Theoretical Foundations}\label{cha:theoretical_foundations}
+\section{Yocto}\label{sec:yocto}
+
+\section{QEMU}\label{sec:qemu}
+
+\section{Xenomai}\label{sec:xenomai}
+\noindent Xenomai 3~\cite{XenomaiXenomai} is an open-source real-time framework that offers two paths to real-time performance.
+The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
+Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
+	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
+	\label{fig:cobalt}
+\end{figure}
+
+\noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
+Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
+	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
+	\label{fig:mercury}
+\end{figure}
+
+\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
+
+\section{VARAN-Bus}\label{sec:varan}
+
+\section{Salamander 4}\label{cha:salamander4}
+This chapter briefly describes the Salamander 4 operating system by SIGMATEK.
+
+\subsection{Structure}
+\noindent Salamander 4 is the proprietary operating system of SIGMATEK.
+It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment~\cite{XenomaiXenomai}.
+Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture.
+The real-time behaviour is achieved through the use of Symmetric Multi-Processing (SMP) and Preemptive Scheduling (PREEMPT).
+In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
+The output of the command~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={System information},label={output:uname_a_virt}]
+root@sigmatek-core2:~# uname -a
+Linux sigmatek-core2 5.15.94 #1 SMP PREEMPT IRQPIPE Tue Feb 14 18:18:05 UTC 2023 x86_64 GNU/Linux
+\end{lstlisting}
+\end{minipage}
+
+\noindent Salamander 4 is powered by SIGMATEK's CP 841~\cite{CPUEinheitenSIGMATEK} and is comprised of the following software modules:
+
+\begin{itemize}
+	\item \textbf{Operating system}: The operating system in a LASAL CPU manages the hardware and software resources of the system.
+It is provided in a completely PC-compatible manner, working with a standard PC BIOS.
+	\item \textbf{Loader}: The loader is a part of the operating system that is responsible for loading programs from executables into memory, preparing them for execution and then executing them.
+	\item \textbf{Hardware classes}: Hardware classes in LASAL represent the different types of hardware components that can be controlled by the LASAL CPU.
+They provide a way to organize and manage the hardware components in a modular and reusable manner.
+The graphical hardware editor in LASAL allows for a true-to-detail simulation of the actual hardware.
+	\item \textbf{Application}: Applications are developed using LASAL CLASS 2~\cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
+\end{itemize}
+
+\noindent These modules and the interfaces (indicated by an arrow) between them are shown in Figure~\ref{fig:lasal_cpu}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
+	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU}
+	\label{fig:lasal_cpu}
+\end{figure}
+
+\subsection{Memory Management}
+\noindent For the sake of completeness, Figure~\ref{fig:memory_management} displays the memory management of Salamander 4.
+LRT stands for Lasal Runtime and creates an execution environment where applications developed using the LASAL Class 2 can run, providing defined real-time functions, data types, and other constructs tailored for real-time programming.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/RAM_Memory_management.png}
+	\caption[Memory Management]{Memory Management}
+	\label{fig:memory_management}
+\end{figure}
+
+\clearpage
+
+\chapter{Methodology}\label{cha:methodology}
+
+This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
+
+\bigskip \noindent Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
+It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
+Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
+
+\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
+It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
+By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
+
+\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel~\cite{XenomaiXenomai}.
+It enables low-latency and deterministic execution of time-critical tasks.
+Xenomai 3 introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
+A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
+The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
+
+The system configuration is shown in Table~\ref{tab:testbed_configuration}
+
+\begin{table}[H]
+	\centering
+	\caption[System configuration]{System configuration}
+	\label{tab:testbed_configuration}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+		\begin{tabular}{|l|l|}
+			\hline
+			\textbf{CPU}    & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
+			\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB           \\ \hline
+			\textbf{GPU}    & NVIDIA RTX A500 Laptop GPU                             \\ \hline
+			\textbf{BIOS}   & Dell Version 1.12.0                                    \\ \hline
+			\textbf{OS}     & Ubuntu 22.04.4 LTS                                     \\
+			\hline
+		\end{tabular}}
+\end{table}
+
+\noindent Figure~\ref{fig:lstopo} is the output of the~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
+	\caption[Hardware topology]{Hardware topology}
+	\label{fig:lstopo}
+\end{figure}
+
+\clearpage
+
+\chapter{Initial Real-Time Latency}\label{cha:initial-real-time-latency}
+
+\chapter{Real-Time Performance Tuning}\label{cha:real-time_tunings}
 
@@
 This involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
-The individual configurations will be discussed in detail and the modifications will be justified with a clear explanation.
+The individual configurations are discussed in detail and the modifications are justified with a clear explanation.
 The goal is to bring the latency of the virtualized system closer to that of the bare metal system, thereby ensuring deterministic behavior under real-time constraints.
@@
 The very first step in this regard is the isolation of a CPU to dedicate it solely to the desired real-time task.
-CPU isolation and affinity will be covered in detail later in Section~\ref{subsec:cpu_isolation}, but its impact will be briefly demonstrated here.
+CPU isolation and affinity are covered in detail later in Section~\ref{subsec:cpu_isolation}, but its impact are briefly demonstrated here.
 
@@
 
-After each tuning, the latency test of Xenomai will be executed to evaluate the improved latency of the system.
+After each tuning, the latency test of Xenomai is executed to evaluate the improved latency of the system.
 
@@
 Here, it is set to the value ``7500''.
-Hence,the timer interrupt will be delivered 7500 nanoseconds earlier than it is actually due.
+Hence, the timer interrupt will be delivered 7500 nanoseconds earlier than it is actually due.
 This gives the VM more time to handle it.
@@
 ~\cite{deoliveiraDemystifyingRealTimeLinux}
+~\cite{gabrielResourceAwareParameterTuning2019}
+~\cite{bhardwajVirtualizationCloudComputing2021}
+~\cite{abbasiExploringOpenStackScalable2023}
+~\cite{mahfoudRealTimePredictiveMaintenanceBased2024}
+~\cite{queirozContainerbasedVirtualizationRealtime2024}
+~\cite{bhardwajVirtualizationCloudComputing2021a}
+~\cite{cinqueEvaluatingVirtualizationFog2023}
 ~\cite{huang2015performance}

### 2024-08-18 Masterarbeit_Robotik_Pamuk.tex
structure of thesis, rauh

@@
 This operating system will be referred to as ``Salamander 4'' in this work.
-The details of Salamander 4 are explained in Chapter~\ref{cha:salamander4}.
+The details of Salamander 4 are explained in Chapter~\ref{sec:salamander4}.
 	\item Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open-source project that helps develop custom Linux-based systems. Yocto is discussed in Section~\ref{sec:yocto}.
@@
 	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal through real-time performance tunings, focussing on latency requirements set by SIGMATEK.
-	      These modifications are subject of Chapter~\ref{cha:real-time_tunings}.
+	      These modifications are subject of Chapter~\ref{sec:real-time_tunings}.
 	\item The latency tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to use by SIGMATEK.
@@
 Subsequently, the PCV 522 module of SIGMATEK was connected with the Pulse Width Module PW 022 of SIGMATEK over the VARAN bus and the VI 021 module of SIGMATEK serving as the power supply.
-	\item The robotic application from Chapter \ref{cha:robotic_application} was written in Lasal Class 2, an object oriented programming tool of SIGMATEK with client-server technology and graphic representation.
+	\item The robotic application from Chapter \ref{sec:robotic_application} was written in Lasal Class 2, an object oriented programming tool of SIGMATEK with client-server technology and graphic representation.
 \end{itemize}
-
 \section{Related Work and State of the Art}
@@
 Given that the working computer used for this work is equipped with an Intel processor, the guidelines presented in this document were particularly relevant and valuable for optimizing real-time performance in KVM-based virtual machines.
-Chapter \ref{cha:real-time_tunings} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
+Chapter \ref{sec:real-time_tunings} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
 The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads.
@@
 
-\bigskip \noindent Virtualization promises to overcome these challenges \cite{queirozContainerbasedVirtualizationRealtime2024} \cite{bhardwajVirtualizationCloudComputing2021a} \cite{cinqueEvaluatingVirtualizationFog2023}.
+\bigskip \noindent Virtualization promises to overcome these challenges \cite{queirozContainerbasedVirtualizationRealtime2024} \cite{bhardwajVirtualizationCloudComputing2021} \cite{cinqueEvaluatingVirtualizationFog2023}.
 However, virtualization also has its own challenges and pitfalls.
@@
 As already established in Section~\ref{sec:application_context}, Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3 and virtualized through QEMU/KVM.
-These components are briefly described in Chapter \ref{cha:theoretical_foundations}.
 The initial latency values are measured with the \texttt{latency} program of the Xenomai tool suite.
@@
 These modifications involve configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS (guest) itself.
-The individual configurations are discussed in detail in Chapter \ref{cha:real-time_tunings} and the modifications are justified with a clear explanation.
+The individual configurations are discussed in detail in Chapter \ref{sec:real-time_tunings} and the modifications are justified with a clear explanation.
 
@@
 The goal is to bring the latency of the virtualized Salamander 4 closer to that of the bare metal system, thereby ensuring deterministic behavior under real-time constraints.
-
-\section{Starting Situation}\label{sec:starting_situation}
+Section \ref{sec:starting_situation} outlines the initial situation of this master's thesis and serves as the foundation for the real-time performance tunings discussed in Chapter \ref{sec:real-time_tunings}.
+To determine and validate the performance of the tunings, we use the \texttt{latency} tool.
+
+\chapter{Methodology}\label{cha:methodology}
+
+This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
+
+\bigskip \noindent Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
+It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
+Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
+
+\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
+It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
+By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
+
+\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel~\cite{XenomaiXenomai}.
+It enables low-latency and deterministic execution of time-critical tasks.
+Xenomai 3 introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
+A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
+The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
+
+The system configuration is shown in Table~\ref{tab:testbed_configuration}
+
+\begin{table}[H]
+	\centering
+	\caption[System configuration]{System configuration}
+	\label{tab:testbed_configuration}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+		\begin{tabular}{|l|l|}
+			\hline
+			\textbf{CPU}    & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
+			\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB           \\ \hline
+			\textbf{GPU}    & NVIDIA RTX A500 Laptop GPU                             \\ \hline
+			\textbf{BIOS}   & Dell Version 1.12.0                                    \\ \hline
+			\textbf{OS}     & Ubuntu 22.04.4 LTS                                     \\
+			\hline
+		\end{tabular}}
+\end{table}
+
+\noindent Figure~\ref{fig:lstopo} is the output of the~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
+	\caption[Hardware topology]{Hardware topology}
+	\label{fig:lstopo}
+\end{figure}
+
+\clearpage
+
+\section{Yocto}\label{sec:yocto}
+
+\section{QEMU}\label{sec:qemu}
+
+\section{Xenomai}\label{sec:xenomai}
+\noindent Xenomai 3~\cite{XenomaiXenomai} is an open-source real-time framework that offers two paths to real-time performance.
+The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
+Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
+	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
+	\label{fig:cobalt}
+\end{figure}
+
+\noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
+Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
+	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
+	\label{fig:mercury}
+\end{figure}
+
+\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
+
+\section{VARAN-Bus}\label{sec:varan}
+
+\section{Salamander 4}\label{sec:salamander4}
+This chapter briefly describes the Salamander 4 operating system by SIGMATEK.
+
+\subsection{Structure}
+\noindent Salamander 4 is the proprietary operating system of SIGMATEK.
+It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment~\cite{XenomaiXenomai}.
+Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture.
+The real-time behaviour is achieved through the use of Symmetric Multi-Processing (SMP) and Preemptive Scheduling (PREEMPT).
+In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
+The output of the command~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
+
+\vspace{1em}
+\begin{minipage}{0.95\columnwidth}
+	\begin{lstlisting}[name={System information},label={output:uname_a_virt}]
+root@sigmatek-core2:~# uname -a
+Linux sigmatek-core2 5.15.94 #1 SMP PREEMPT IRQPIPE Tue Feb 14 18:18:05 UTC 2023 x86_64 GNU/Linux
+\end{lstlisting}
+\end{minipage}
+
+\noindent Salamander 4 is powered by SIGMATEK's CP 841~\cite{CPUEinheitenSIGMATEK} and is comprised of the following software modules:
+
+\begin{itemize}
+	\item \textbf{Operating system}: The operating system in a LASAL CPU manages the hardware and software resources of the system.
+It is provided in a completely PC-compatible manner, working with a standard PC BIOS.
+	\item \textbf{Loader}: The loader is a part of the operating system that is responsible for loading programs from executables into memory, preparing them for execution and then executing them.
+	\item \textbf{Hardware classes}: Hardware classes in LASAL represent the different types of hardware components that can be controlled by the LASAL CPU.
+They provide a way to organize and manage the hardware components in a modular and reusable manner.
+The graphical hardware editor in LASAL allows for a true-to-detail simulation of the actual hardware.
+	\item \textbf{Application}: Applications are developed using LASAL CLASS 2~\cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
+\end{itemize}
+
+\noindent These modules and the interfaces (indicated by an arrow) between them are shown in Figure~\ref{fig:lasal_cpu}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
+	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU}
+	\label{fig:lasal_cpu}
+\end{figure}
+
+\subsection{Memory Management}
+\noindent For the sake of completeness, Figure~\ref{fig:memory_management} displays the memory management of Salamander 4.
+LRT stands for Lasal Runtime and creates an execution environment where applications developed using the LASAL Class 2 can run, providing defined real-time functions, data types, and other constructs tailored for real-time programming.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/RAM_Memory_management.png}
+	\caption[Memory Management]{Memory Management}
+	\label{fig:memory_management}
+\end{figure}
+
+\clearpage
+
+\chapter{Implementation}
+
+\section{Initial Situation}\label{sec:starting_situation}
 As a starting point, initial latency values of both the bare metal and virtualization versions were measured with the \texttt{latency} tool of the xenomai tool suite.
@@
 This script is used to start QEMU with required parameters to boot Salamamder 4 OS.
-It will be adjusted in Chapter~\ref{cha:real-time_tunings} in order to accompany real-time performance tunings.
+It will be adjusted in Chapter~\ref{sec:real-time_tunings} in order to accompany real-time performance tunings.
 
@@
 
-\clearpage
-
-The main objective of this work is to create a real-time robot control platform that integrates Salamander OS, Xenomai, QEMU and PCV-522 in the Yocto environment.
-
-\clearpage
-
-\chapter{Theoretical Foundations}\label{cha:theoretical_foundations}
-\section{Yocto}\label{sec:yocto}
-
-\section{QEMU}\label{sec:qemu}
-
-\section{Xenomai}\label{sec:xenomai}
-\noindent Xenomai 3~\cite{XenomaiXenomai} is an open-source real-time framework that offers two paths to real-time performance.
-The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
-Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
-	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
-	\label{fig:cobalt}
-\end{figure}
-
-\noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
-Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
-	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
-	\label{fig:mercury}
-\end{figure}
-
-\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
-
-\section{VARAN-Bus}\label{sec:varan}
-
-\section{Salamander 4}\label{cha:salamander4}
-This chapter briefly describes the Salamander 4 operating system by SIGMATEK.
-
-\subsection{Structure}
-\noindent Salamander 4 is the proprietary operating system of SIGMATEK.
-It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment~\cite{XenomaiXenomai}.
-Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture.
-The real-time behaviour is achieved through the use of Symmetric Multi-Processing (SMP) and Preemptive Scheduling (PREEMPT).
-In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
-The output of the command~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
-
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={System information},label={output:uname_a_virt}]
-root@sigmatek-core2:~# uname -a
-Linux sigmatek-core2 5.15.94 #1 SMP PREEMPT IRQPIPE Tue Feb 14 18:18:05 UTC 2023 x86_64 GNU/Linux
-\end{lstlisting}
-\end{minipage}
-
-\noindent Salamander 4 is powered by SIGMATEK's CP 841~\cite{CPUEinheitenSIGMATEK} and is comprised of the following software modules:
-
-\begin{itemize}
-	\item \textbf{Operating system}: The operating system in a LASAL CPU manages the hardware and software resources of the system.
-It is provided in a completely PC-compatible manner, working with a standard PC BIOS.
-	\item \textbf{Loader}: The loader is a part of the operating system that is responsible for loading programs from executables into memory, preparing them for execution and then executing them.
-	\item \textbf{Hardware classes}: Hardware classes in LASAL represent the different types of hardware components that can be controlled by the LASAL CPU.
-They provide a way to organize and manage the hardware components in a modular and reusable manner.
-The graphical hardware editor in LASAL allows for a true-to-detail simulation of the actual hardware.
-	\item \textbf{Application}: Applications are developed using LASAL CLASS 2~\cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
-\end{itemize}
-
-\noindent These modules and the interfaces (indicated by an arrow) between them are shown in Figure~\ref{fig:lasal_cpu}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
-	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU}
-	\label{fig:lasal_cpu}
-\end{figure}
-
-\subsection{Memory Management}
-\noindent For the sake of completeness, Figure~\ref{fig:memory_management} displays the memory management of Salamander 4.
-LRT stands for Lasal Runtime and creates an execution environment where applications developed using the LASAL Class 2 can run, providing defined real-time functions, data types, and other constructs tailored for real-time programming.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/RAM_Memory_management.png}
-	\caption[Memory Management]{Memory Management}
-	\label{fig:memory_management}
-\end{figure}
-
-\clearpage
-
-\chapter{Methodology}\label{cha:methodology}
-
-This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
-
-\bigskip \noindent Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
-It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
-Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
-
-\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
-It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
-By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
-
-\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel~\cite{XenomaiXenomai}.
-It enables low-latency and deterministic execution of time-critical tasks.
-Xenomai 3 introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
-A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
-The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
-
-The system configuration is shown in Table~\ref{tab:testbed_configuration}
-
-\begin{table}[H]
-	\centering
-	\caption[System configuration]{System configuration}
-	\label{tab:testbed_configuration}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|l|l|}
-			\hline
-			\textbf{CPU}    & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
-			\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB           \\ \hline
-			\textbf{GPU}    & NVIDIA RTX A500 Laptop GPU                             \\ \hline
-			\textbf{BIOS}   & Dell Version 1.12.0                                    \\ \hline
-			\textbf{OS}     & Ubuntu 22.04.4 LTS                                     \\
-			\hline
-		\end{tabular}}
-\end{table}
-
-\noindent Figure~\ref{fig:lstopo} is the output of the~\texttt{lstopo} command and visualizes the hardware nodes of the system, including CPU cores, caches, memory, and I/O devices.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
-	\caption[Hardware topology]{Hardware topology}
-	\label{fig:lstopo}
-\end{figure}
-
-\clearpage
-
-\chapter{Initial Real-Time Latency}\label{cha:initial-real-time-latency}
-
-\chapter{Real-Time Performance Tuning}\label{cha:real-time_tunings}
+\section{Real-Time Performance Tuning}\label{sec:real-time_tunings}
 
@@
 
-\section{BIOS Configurations}\label{sec:bios_configurations}
+\subsection{BIOS Configurations}\label{sec:bios_configurations}
 
@@
 
-\section{Kernel Configurations}\label{sec:kernel_configurations}
+\subsection{Kernel Configurations}\label{sec:kernel_configurations}
 
@@
 
-\section{Host OS Configurations}\label{sec:host_configurations}
+\subsection{Host OS Configurations}\label{sec:host_configurations}
 The host OS needs to provide an environment where the guest OS can operate in real-time.
@@
 In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest OS is provided.
-\subsection{CPU affinity and Isolation}\label{subsec:cpu_isolation}
+\subsubsection{CPU affinity and Isolation}\label{subsec:cpu_isolation}
 
@@
 \end{minipage}
-\subsection{KVM Entry and KVM Exit}\label{subsec:kvm_entry_exit}
+\subsubsection{KVM Entry and KVM Exit}\label{subsec:kvm_entry_exit}
 \textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
@@
 
-\subsection{Interrupt Affinity}\label{subsec:irq_handling}
+\subsubsection{Interrupt Affinity}\label{subsec:irq_handling}
 Once the CPUs were isolated, interrupt requests handling was the next step.
@@
 \end{minipage}
-\subsection{RT-Priority}
+\subsubsection{RT-Priority}
 Having a real-time kernel itself is a crucial part of achieving deterministic behavior, but not enough to take full advantage of the real-time capabilities.
@@
 It is also important that there are no other unexpected real time processes running on the system concurrently.
-\subsection{Disable RT Throttling}
+\subsubsection{Disable RT Throttling}
 If a real-time task consumes 100\% of the CPU time, the system may become unresponsive as a whole.
@@
 This reduces the potential for missed deadlines.
-\subsection{Disable Timer Migration}
+\subsubsection{Disable Timer Migration}
 Timer migration allows timers to be moved from one CPU to another, which means the kernel can balance load across multiple CPUs.
@@
 This reduces the amount of context switches and interrupts.
-\subsection{Set Device Driver Work Queue}
+\subsubsection{Set Device Driver Work Queue}
 The device driver work queue allows time-consuming tasks to be offloaded to be processed later in a separate kernel thread.
@@
 This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
-\subsection{Disable RCU CPU Stall Warnings}
+\subsubsection{Disable RCU CPU Stall Warnings}
 As already mentioned in Section~\ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
@@
 This setting is also not persistent across reboots, so the command needs to be added to a startup script.
-\subsection{Stop Certain Services}
+\subsubsection{Stop Certain Services}
 Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{tab:stop_servies} can be further sources for random latency and unnecessary overhead.
@@
 \end{table}
-\subsection{Disable Machine Check}
+\subsubsection{Disable Machine Check}
 Machine checks report hardware errors and these checks can cause interruptions and increase latency.
@@
 
-\subsection{Boot into text-based environment}
+\subsubsection{Boot into text-based environment}
 The Graphical User Interface (GUI) consumes great amounts of system resources.
@@
 
-\section{QEMU/KVM Configurations}\label{sec:guest_configurations}
+\subsection{QEMU/KVM Configurations}\label{sec:guest_configurations}
 KVM allows the guest OS to run directly on the hardware, bypassing the need for traditional emulation which can introduce delays.
@@
 
-\subsection{Tune LAPIC Timer Advance}
+\subsubsection{Tune LAPIC Timer Advance}
 The Local Advanced Programmable Interrupt Controller (LAPIC) is a built-in timer that handles the delivery of interrupts to the CPU.
@@
 
-\subsection{Set QEMU Options for real-time VM}
+\subsubsection{Set QEMU Options for real-time VM}
 QEMU provides several options that can be used to improve the real-time performance of the guest VM. Table~\ref{tab:qemu_options} briefly explains these options.
@@
 
-\section{Guest OS Configurations}
-
-\section{Other Configurations}
+\subsection{Guest OS Configurations}
+
+\subsection{Other Configurations}
 There are other configurations that may be relevant depending on the case.
@@
 
-\chapter{Real-Time Robotic Application}\label{cha:robotic_application}
+\section{Real-Time Robotic Application}\label{sec:robotic_application}
 This chapter compares the latency of the Salamander 4 operating system before and after the real-time performance tunings with the latency of Salamander 4 running on bare metal hardware.
@@
 
-\section{Setup of Hardware Salamander 4}
+\subsection{Setup of Hardware Salamander 4}
 In this version, Salamander 4 runs on the CP 841~\cite{CPUEinheitenSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
@@
 
-\section{Setup of QEMU Salamander 4}\label{sec:setup_experiment_virtualized}
+\subsection{Setup of QEMU Salamander 4}\label{sec:setup_experiment_virtualized}
 In the virtualized setup of the experiment, the VCPU functionality of QEMU is used to connect QEMU with the PWM module.
@@
 \clearpage
-\section{Robotic Application}
+\subsection{Robotic Application}
 
@@
 ~\cite{XenomaiXenomai}
+~\cite{casiniLatencyAnalysisVirtualization2021}
+~\cite{zhangEvaluatingOptimizingVirtualization2010}
+~\cite{QEMU}
 ~\cite{CPUEinheitenSIGMATEK}
@@
 ~\cite{KernelShark}
-~\cite{QEMU}
 ~\cite{RealtimePreempt_rt_versionsWiki}
@@
 \noindent Papers
-
 ~\cite{canbazPerformanceAnalysisRealtime2022}
@@
 ~\cite{adamRealTimePerformanceResponse2021}
-~\cite{RealTimePerformanceTuning2022}
 ~\cite{yoonRealTimePerformanceAnalysis}
@@
 ~\cite{broskyShieldedProcessorsGuaranteeing2003}
+~\cite{RealTimePerformanceTuning2022}
 ~\cite{kiszkaLinuxRealTimeHypervisor}
@@
 ~\cite{queirozContainerbasedVirtualizationRealtime2024}
-~\cite{bhardwajVirtualizationCloudComputing2021a}
 ~\cite{cinqueEvaluatingVirtualizationFog2023}

### 2024-08-18 Masterarbeit_Robotik_Pamuk.tex
references

@@
 
-\subsection{Salamander 4 Bare Metal}\label{sec:salamander4-bare-metal}
+\subsection{Salamander 4 Bare Metal}\label{subsec:salamander4-bare-metal}
 The output of the command~\texttt{uname -a} for Salamander 4 bare metal is shown in Code~\ref{output:uname_a_hw}.
@@
 
-\subsection{Salamander 4 Virtualization}\label{sec:salamander4-virtualization}
+\subsection{Salamander 4 Virtualization}\label{subsec:salamander4-virtualization}
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
@@
 
-\subsection{BIOS Configurations}\label{sec:bios_configurations}
+\subsection{BIOS Configurations}\label{subsec:bios_configurations}
 
@@
 
-\subsection{Kernel Configurations}\label{sec:kernel_configurations}
+\subsection{Kernel Configurations}\label{subsec:kernel_configurations}
 
@@
 
-\subsection{Host OS Configurations}\label{sec:host_configurations}
+\subsection{Host OS Configurations}\label{subsec:host_configurations}
 The host OS needs to provide an environment where the guest OS can operate in real-time.
@@
 In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest OS is provided.
-\subsubsection{CPU affinity and Isolation}\label{subsec:cpu_isolation}
+\subsubsection{CPU affinity and Isolation}\label{subsubsec:cpu_isolation}
 
@@
 \end{minipage}
-\subsubsection{KVM Entry and KVM Exit}\label{subsec:kvm_entry_exit}
+\subsubsection{KVM Entry and KVM Exit}\label{subsubsec:kvm_entry_exit}
 \textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
@@
 
-\subsubsection{Interrupt Affinity}\label{subsec:irq_handling}
+\subsubsection{Interrupt Affinity}\label{subsubsec:irq_handling}
 Once the CPUs were isolated, interrupt requests handling was the next step.
@@
 
-\subsection{QEMU/KVM Configurations}\label{sec:guest_configurations}
+\subsection{QEMU/KVM Configurations}\label{subsec:guest_configurations}
 KVM allows the guest OS to run directly on the hardware, bypassing the need for traditional emulation which can introduce delays.
@@
 
-\subsection{Setup of QEMU Salamander 4}\label{sec:setup_experiment_virtualized}
+\subsection{Setup of QEMU Salamander 4}\label{subsec:setup_experiment_virtualized}
 In the virtualized setup of the experiment, the VCPU functionality of QEMU is used to connect QEMU with the PWM module.

### 2024-08-18 Masterarbeit_Robotik_Pamuk.tex
ref:subsec

@@
 \schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
-\outline{Sections~\ref{sec:salamander4-bare-metal} and~\ref{sec:salamander4-virtualization}  demonstrate the inital real-time latency values gathered for bare metal and virtualization. 
+\outline{Subsections~\ref{subsec:salamander4-bare-metal} and~\ref{subsec:salamander4-virtualization} demonstrate the inital real-time latency values gathered for bare metal and virtualization. 
 }
@@
 Its operating principle is outlined in Section \ref{sec:varan}.
-	\item For the purpose of replacing the physical CPU, SIGMATEK provided a PCI Insert Card module PCV 522, that was plugged into the PC, as explained in Section \ref{sec:setup_experiment_virtualized}.
+	\item For the purpose of replacing the physical CPU, SIGMATEK provided a PCI Insert Card module PCV 522, that was plugged into the PC, as explained in Subsection \ref{subsec:setup_experiment_virtualized}.
 This component could then be used to communicate with the Input/Output peripherals, just like the physical CPU could when the I/O modules were attached to it.
@@
 Salamander 4 virtualization refers to a virtual version of the Salamander 4 hardware platform, achieved through QEMU/KVM.
-Sections~\ref{sec:salamander4-bare-metal} and~\ref{sec:salamander4-virtualization} specify the details of the measurements for both versions.
+Subsections~\ref{subsec:salamander4-bare-metal} and~\ref{subsec:salamander4-virtualization} specify the details of the measurements for both versions.
 In the further course, the aim was to bring the latency values of the virtualization closer to those of the hardware and guarantee deterministic and reliable behavior.
@@
 The very first step in this regard is the isolation of a CPU to dedicate it solely to the desired real-time task.
-CPU isolation and affinity are covered in detail later in Section~\ref{subsec:cpu_isolation}, but its impact are briefly demonstrated here.
+CPU isolation and affinity are covered in detail later in Subsection~\ref{subsubsec:cpu_isolation}, but its impact are briefly demonstrated here.
 
@@
 	\item \textbf{isolcpus=4}: Isolates CPU 4 from the general scheduler, meaning no process will be scheduled to run on this CPU unless it is explicitly assigned.
-CPU isolation is explained in detail in Section~\ref{subsec:cpu_isolation}.
+CPU isolation is explained in detail in Subsubsection~\ref{subsubsec:cpu_isolation}.
 	\item \textbf{rcu\_nocbs=4}: The Linux kernel uses a synchronization mechanism called RCU, or Read-Copy-Update.
@@
 Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs.
-Section~\ref{subsec:irq_handling} dives deeper into Interrupt Request affinity.
+Subsection~\ref{subsubsec:irq_handling} dives deeper into Interrupt Request affinity.
 \end{itemize}
@@
 \subsubsection{Disable RCU CPU Stall Warnings}
-As already mentioned in Section~\ref{sec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
+As already mentioned in Section~\ref{subsec:kernel_configurations}, the Linux kernel uses RCU as a synchronization mechanism for reading from and writing to shared data.
 An RCU CPU stall in the Linux kernel can occur due to several reasons.

### 2024-08-19 Masterarbeit_Robotik_Pamuk.tex
delete comments
repair

@@
 	      These modifications are subject of Chapter~\ref{sec:real-time_tunings}.
-	\item The latency tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to use by SIGMATEK.
+	\item The \texttt{latency} tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to use by SIGMATEK.
+The goal was set to achieving latency values below 50 microseconds in the period of the measuremenet.
 	\item SIGMATEK uses the VARAN bus for communication between machines and systems, a real-time Ethernet network that was specially developed for industrial automation.
@@
 The individual configurations are discussed in detail in Chapter \ref{sec:real-time_tunings} and the modifications are justified with a clear explanation.
-
+\clearpage
 \section{Objective}
-The goal is to bring the latency of the virtualized Salamander 4 closer to that of the bare metal system, thereby ensuring deterministic behavior under real-time constraints.
-Section \ref{sec:starting_situation} outlines the initial situation of this master's thesis and serves as the foundation for the real-time performance tunings discussed in Chapter \ref{sec:real-time_tunings}.
-To determine and validate the performance of the tunings, we use the \texttt{latency} tool.
+The goal is to bring the latency of the virtualized Salamander 4 closer to that of the hardware, thereby ensuring deterministic behavior under real-time constraints.
+Section \ref{sec:starting_situation} outlines the initial situation of this master's thesis and serves as the foundation for the real-time performance tunings discussed in Section \ref{sec:real-time_tunings}.
+To determine and validate the performance of the tunings, the \texttt{latency} tool is used, as stated in Section \ref{sec:application_context}.
+The objective is to achieve latency values below 50 microseconds during the measurement period through real-time tunings.
+On top of that, a robotic application demonstrates and validates that the tunings ensure in a practical scenario taht the operating system meets the required real-time performance criteria and constraints.
+\clearpage
 
@@
 \end{itemize}
-
-\begin{comment}
-https://www.baeldung.com/linux/kernel-images
-\end{comment}
 
@@
 
-\begin{comment}
-battery -> https://www.flaticon.com/de/kostenloses-icon/autobatterie_4238252?related_id=4238254
-\end{comment}
-
 \noindent The program was written in Lasal Class 2 and was applied to all three mentioned versions of Salamander 4 to measure the reaction time of the robot to the specific commands.
@@
 \noindent Papers
+
 ~\cite{canbazPerformanceAnalysisRealtime2022}

### 2024-08-19 Masterarbeit_Robotik_Pamuk.tex
1.5 Objective, 2.1 Host Operating System, 2.2 Guest Operating System, 2.5.1 Structure, 2.5.2 Memory Management, 2.3 Yocto, 2.4 Xenomai

@@
 The key difference is whether the system is time-critical or not~\cite{canbazPerformanceAnalysisRealtime2022}.
-In a traditional GPOS, such as Windows or Linux-based distributions, a high-priority task cannot interrupt a kernel operation, as illustrated in Figure \ref{fig:kernel_generic}.
+In a traditional GPOS, such as Windows or Linux-based distributions, a high-priority task cannot interrupt a kernel operation, as illustrated in Figure~\ref{fig:kernel_generic}.
 This makes them not suitable for real-time requirements, as they cannot guarantee deterministic execution times.
@@
 
-\noindent However, in an RTOS, a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation, as shown in Figure \ref{fig:kernel_rt} RTOS are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
+\noindent However, in an RTOS, a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation, as shown in Figure~\ref{fig:kernel_rt} RTOS are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
 
@@
 	\item SIGMATEK uses the VARAN bus for communication between machines and systems, a real-time Ethernet network that was specially developed for industrial automation.
-Its operating principle is outlined in Section \ref{sec:varan}.
-	\item For the purpose of replacing the physical CPU, SIGMATEK provided a PCI Insert Card module PCV 522, that was plugged into the PC, as explained in Subsection \ref{subsec:setup_experiment_virtualized}.
+Its operating principle is outlined in Section~\ref{sec:varan}.
+	\item For the purpose of replacing the physical CPU, SIGMATEK provided a PCI Insert Card module PCV 522, that was plugged into the PC, as explained in Subsection~\ref{subsec:setup_experiment_virtualized}.
 This component could then be used to communicate with the Input/Output peripherals, just like the physical CPU could when the I/O modules were attached to it.
 Subsequently, the PCV 522 module of SIGMATEK was connected with the Pulse Width Module PW 022 of SIGMATEK over the VARAN bus and the VI 021 module of SIGMATEK serving as the power supply.
-	\item The robotic application from Chapter \ref{sec:robotic_application} was written in Lasal Class 2, an object oriented programming tool of SIGMATEK with client-server technology and graphic representation.
+	\item The robotic application from Chapter~\ref{sec:robotic_application} was written in Lasal Class 2, an object oriented programming tool of SIGMATEK and part of SIGMATEK`s Engineering Platform LASAL, with client-server technology and graphic representation.
 \end{itemize}
@@
 
-\bigskip \noindent \citeauthor{yoonRealTimePerformanceAnalysis}~\cite{yoonRealTimePerformanceAnalysis} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
-They implement EtherCAT for real-time communication of distributed devices whereas this thesis uses Varan.
-The authors highlight the importance of timely data processing in robots that collect data from their environment and respond accordingly through their sensors and actuators.
-The authors conclude with an experiment that their robotic system could achieve deterministic control of actuators and sensors even under heavy system load through the use of Linux with the real-time preemption patch.
-Their research has similarities to this master's thesis in terms of real-time tunings for robot control.
-However, this thesis includes also the virtualization aspect that is central.
-
 \bigskip \noindent \citeauthor{cinqueVirtualizingMixedCriticalitySystems2022}~\cite{cinqueVirtualizingMixedCriticalitySystems2022} survey different virtualization approaches, including separation kernels, microkernels, enhanced general-purpose hypervisors, and lightweight virtualization solutions like containers and unikernels.
@@
 Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), since there is also the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started.
-These stages are the primary sources of latencies caused by KVM, they are briefly described in Figure \ref{fig:kvm_stages}.
+These stages are the primary sources of latencies caused by KVM, they are briefly described in Figure~\ref{fig:kvm_stages}.
 
@@
 \noindent The paper also details several performance tuning methods, including CPU shielding and prioritization, which have been applied in this work.
-
+\clearpage
 \bigskip \noindent \citeauthor{broskyShieldedProcessorsGuaranteeing2003}~\cite{broskyShieldedProcessorsGuaranteeing2003} report in their work about the shielded CPU model for enhancing real-time performance in symmetric multiprocessor systems.
@@
 Given that the working computer used for this work is equipped with an Intel processor, the guidelines presented in this document were particularly relevant and valuable for optimizing real-time performance in KVM-based virtual machines.
-Chapter \ref{sec:real-time_tunings} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
+Chapter~\ref{sec:real-time_tunings} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
 The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads.
+
+\bigskip \noindent \citeauthor{yoonRealTimePerformanceAnalysis}~\cite{yoonRealTimePerformanceAnalysis} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
+They implement EtherCAT for real-time communication of distributed devices whereas this thesis uses Varan.
+The authors highlight the importance of timely data processing in robots that collect data from their environment and respond accordingly through their sensors and actuators.
+The authors conclude with an experiment that their robotic system could achieve deterministic control of actuators and sensors even under heavy system load through the use of Linux with the real-time preemption patch.
+Their research has similarities to this master's thesis in terms of real-time tunings for robot control.
+However, this thesis includes also the virtualization aspect that is central.
 
@@
 Moreover, Linux has a feature that can limit how much CPU time all real-time tasks can use in the host system, which was an important information for the real-time tunings of this thesis later.
-
+\clearpage
 \bigskip \noindent \citeauthor{mckenneyRealTimeVs}~\cite{mckenneyRealTimeVs} addresses the overheads associated with real-time Linux when using KVM and QEMU.
@@
 This is the main reason why this tool was not used to measure the scheduling latency in this work.
-Instead, the \texttt{latency} tool was used, as was determined in Chapter \ref{sec:application_context}.
-
-\clearpage
+Instead, the \texttt{latency} tool was used, as was determined in Chapter~\ref{sec:application_context}.
+
 \section{Problem and Task Definition}
@@
 
-\bigskip \noindent Virtualization promises to overcome these challenges \cite{queirozContainerbasedVirtualizationRealtime2024} \cite{bhardwajVirtualizationCloudComputing2021} \cite{cinqueEvaluatingVirtualizationFog2023}.
+\bigskip \noindent Virtualization promises to overcome these challenges~\cite{queirozContainerbasedVirtualizationRealtime2024}~\cite{bhardwajVirtualizationCloudComputing2021}~\cite{cinqueEvaluatingVirtualizationFog2023}.
 However, virtualization also has its own challenges and pitfalls.
-The additional layers of abstraction can lead to increased overhead and latency \cite{casiniLatencyAnalysisVirtualization2021} \cite{zhangEvaluatingOptimizingVirtualization2010}.
+The additional layers of abstraction can lead to increased overhead and latency~\cite{casiniLatencyAnalysisVirtualization2021}~\cite{zhangEvaluatingOptimizingVirtualization2010}.
 This is not desired and can be particularly problematic for real-time systems that require strict timing constraints.
 Making sure that tasks run predictably and on time in a virtualized environment, can be challenging.
-The hypervisor and other virtual components can cause variations in how long tasks take to execute \cite{garcia-vallsChallengesRealtimeVirtualization2014}.
-When there are multiple virtual machines (VM) sharing the same physical resources, the VMs can affect each other's performance and consequently violate real-time constraints \cite{queirozContainerbasedVirtualizationRealtime2024}.
-
-\bigskip \noindent In this master’s thesis, the task is the virtualization of a real-time operating system to control a robot, with a focus on compliance with real-time determinism.
-Chapter \ref{cha:methodology} explains in detail the procedure, how this was tackled.
+The hypervisor and other virtual components can cause variations in how long tasks take to execute~\cite{garcia-vallsChallengesRealtimeVirtualization2014}.
+When there are multiple virtual machines (VM) sharing the same physical resources, the VMs can affect each other's performance and consequently violate real-time constraints~\cite{queirozContainerbasedVirtualizationRealtime2024}.
+\clearpage
+\bigskip \noindent In this master's thesis, the task is the virtualization of a real-time operating system to control a robot, with a focus on compliance with real-time determinism.
+Chapter~\ref{cha:methodology} explains in detail the procedure, how this was tackled.
 As already established in Section~\ref{sec:application_context}, Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3 and virtualized through QEMU/KVM.
 The initial latency values are measured with the \texttt{latency} program of the Xenomai tool suite.
-The initial situation is described in Section \ref{sec:starting_situation}.
+The initial situation is described in Section~\ref{sec:starting_situation}.
 In a nutshell, there is a significant initial gap in latency statistics between the virtualized Salamander 4 and the Salamander 4 on bare metal.
@@
 These modifications involve configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS (guest) itself.
-The individual configurations are discussed in detail in Chapter \ref{sec:real-time_tunings} and the modifications are justified with a clear explanation.
-\clearpage
+The individual configurations are discussed in detail in Chapter~\ref{sec:real-time_tunings} and the modifications are justified with a clear explanation.
+
 \section{Objective}
-The goal is to bring the latency of the virtualized Salamander 4 closer to that of the hardware, thereby ensuring deterministic behavior under real-time constraints.
-Section \ref{sec:starting_situation} outlines the initial situation of this master's thesis and serves as the foundation for the real-time performance tunings discussed in Section \ref{sec:real-time_tunings}.
-To determine and validate the performance of the tunings, the \texttt{latency} tool is used, as stated in Section \ref{sec:application_context}.
-The objective is to achieve latency values below 50 microseconds during the measurement period through real-time tunings.
-On top of that, a robotic application demonstrates and validates that the tunings ensure in a practical scenario taht the operating system meets the required real-time performance criteria and constraints.
-\clearpage
-
+The goal is to bring the latency of the virtualized Salamander 4 closer to that of the hardware, thereby providing a comprehensive blueprint for making a virtualized guest system in a host system real-time capable with deterministic behavior.
+Section~\ref{sec:starting_situation} outlines the initial situation of this master's thesis and serves as the foundation for the real-time performance tunings discussed in Section~\ref{sec:real-time_tunings}.
+The objective is to achieve latency values below 50 microseconds during the measurement period through real-time tunings, with no outliers whatsoever.
+To determine and validate the performance of the tunings, the \texttt{latency} tool is used, as stated in Section~\ref{sec:application_context}.
+On top of that, the aim is to use a robotic application in a practical scenario to demonstrate and validate that the tunings significantly reduce latency and improve determinism compared to the unmodified virtualization.
+
+\clearpage
 \chapter{Methodology}\label{cha:methodology}
 
-This section describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
-
-\bigskip \noindent Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
-It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
-Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
-
-\bigskip \noindent The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
-It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
-By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
-
-\bigskip \noindent Real-time operating system capabilities were provided by Xenomai, which is real-time development framework that extends the Linux kernel~\cite{XenomaiXenomai}.
-It enables low-latency and deterministic execution of time-critical tasks.
-Xenomai 3 introduces a dual-kernel approach with a real-time kernel coexisting alongside Linux.
-A key utility within the Xenomai suite is the latency tool, which benchmarks the timer latency - the time it takes for the kernel to respond to timer interrupts or task activations.
-The tool creates real-time tasks or interrupt handlers and measures the latency between expected and actual execution times.
-
-The system configuration is shown in Table~\ref{tab:testbed_configuration}
+This chapter describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
+
+\section{Host Operating System}\label{sec:host_operating_system}
+The host operating system is the basis for the virtualization of the guest operating system.
+In the course of this work, it is continuously tuned to achieve real-time behavior and determinism.
+Table~\ref{tab:testbed_configuration} illustrates the specifications of the host.
 
@@
 	\centering
-	\caption[System configuration]{System configuration}
+	\caption[Host operating system configuration]{Host operating system configuration}
 	\label{tab:testbed_configuration}
@@
 			\hline
-			\textbf{CPU}    & 13\textsuperscript{th} Gen Intel(R) Core(TM) i7-13800H \\ \hline
-			\textbf{Memory} & 2 $\times$ 16GB SO-DIMM DDR5-5600 MT/s, 32GB           \\ \hline
-			\textbf{GPU}    & NVIDIA RTX A500 Laptop GPU                             \\ \hline
-			\textbf{BIOS}   & Dell Version 1.12.0                                    \\ \hline
-			\textbf{OS}     & Ubuntu 22.04.4 LTS                                     \\
-			\hline
+			\textbf{CPU}            & Intel(R) Core(TM) i7-8700 CPU @ 3.20GHz, 6 cores \\ \hline
+			\textbf{Memory}         & 4 $\times$ 8GB DDR4-2666/2400 MHz, 32GB          \\ \hline
+			\textbf{GPU}            & Intel UHD Graphics 630 GPU                       \\ \hline
+			\textbf{Storage}        & 500GB NVMe SSD                                   \\ \hline
+			\textbf{BIOS}           & Dell Version 1.29.0                              \\ \hline
+			\textbf{OS}             & Ubuntu 22.04.4 LTS                               \\ \hline
+			\textbf{Kernel Version} & 6.8.0-40-generic                                 \\ \hline
 		\end{tabular}}
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{img/lstopo.png}
+	\includegraphics[width=0.55\columnwidth]{img/lstopo.png}
 	\caption[Hardware topology]{Hardware topology}
@@
 
-\clearpage
-
-\section{Yocto}\label{sec:yocto}
-
-\section{QEMU}\label{sec:qemu}
-
-\section{Xenomai}\label{sec:xenomai}
-\noindent Xenomai 3~\cite{XenomaiXenomai} is an open-source real-time framework that offers two paths to real-time performance.
-The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
-Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
-	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
-	\label{fig:cobalt}
-\end{figure}
-
-\noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
-Often, applications require the PREEMPT-RT extension to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
-	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
-	\label{fig:mercury}
-\end{figure}
-
-\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
-
-\section{VARAN-Bus}\label{sec:varan}
-
-\section{Salamander 4}\label{sec:salamander4}
-This chapter briefly describes the Salamander 4 operating system by SIGMATEK.
+\section{Guest Operating System}\label{sec:guest_operating_system}
+This section briefly describes the guest operating system, Salamander 4.
 
@@
 It is provided in a completely PC-compatible manner, working with a standard PC BIOS.
-	\item \textbf{Loader}: The loader is a part of the operating system that is responsible for loading programs from executables into memory, preparing them for execution and then executing them.
+	\item \textbf{Loader}: The loader is a part of the operating system that is responsible for loading programs from executables into memory, preparing them for execution and then running them.
 	\item \textbf{Hardware classes}: Hardware classes in LASAL represent the different types of hardware components that can be controlled by the LASAL CPU.
-They provide a way to organize and manage the hardware components in a modular and reusable manner.
+They help organize and manage the hardware components in a modular manner.
 The graphical hardware editor in LASAL allows for a true-to-detail simulation of the actual hardware.
-	\item \textbf{Application}: Applications are developed using LASAL CLASS 2~\cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
+	\item \textbf{Applications}: Applications are developed using LASAL CLASS 2~\cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
 \end{itemize}
+
+\clearpage
 
@@
 \noindent For the sake of completeness, Figure~\ref{fig:memory_management} displays the memory management of Salamander 4.
-LRT stands for Lasal Runtime and creates an execution environment where applications developed using the LASAL Class 2 can run, providing defined real-time functions, data types, and other constructs tailored for real-time programming.
+LRT stands for Lasal Runtime and creates an execution environment where applications developed using LASAL Class 2 can run- It provides real-time functions, data types, and other constructs for real-time programming.
 
@@
 \clearpage
+
+\section{Yocto}\label{sec:yocto}
+
+Salamander 4 is built with Yocto, an open-source project that helps develop custom Linux-based operating systems for embedded devices.
+It is not a Linux version itself but a framework to make custom versions.
+From the kernel settings to the software packages included, almost everything in the Linux distribution can be adjusted with Yocto, hence it fits custom needs.
+Yocto organizes different components of the system into layers, making code management and updates easier.
+It builds the operating system with BitBake and automates the compilation and assembly of software.
+Yocto can generate a Software Development Kit (SDK) containing tools and libraries for developing applications for the custom OS~\cite{WelcomeYoctoProject}.
+
+\section{Xenomai}\label{sec:xenomai}
+
+Xenomai 3~\cite{XenomaiXenomai} is an open-source real-time framework that offers two paths to real-time performance.
+The first approach supplements the Linux kernel with a compact real-time core dubbed Cobalt, demonstrated in Figure~\ref{fig:cobalt}.
+Cobalt runs side-by-side with Linux, but it handles all time-critical activities like interrupt processing and real-time thread scheduling with higher priority than the regular kernel activities.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
+	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
+	\label{fig:cobalt}
+\end{figure}
+
+\clearpage
+
+\noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
+Often, applications require the PREEMPT-RT~\cite{RealtimePreempt_rt_versionsWiki} patch to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
+	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
+	\label{fig:mercury}
+\end{figure}
+
+\noindent Salamander 4 uses the Cobalt real-time core with the Dovetail extension, which allows the kernel to handle real-time tasks with low latency.
+A key tool of the Xenomai suite that is used in this work is the \texttt{latency} tool~\cite{LATENCY}.
+It benchmarks the timer latency, which is the time it takes for the kernel to respond to timer interrupts or task activations.
+The tool creates real-time tasks or interrupt handlers and measures the difference between the expected and actual execution times.
+
+\section{QEMU}\label{sec:qemu}
+
+\section{Trace-cmd}\label{sec:trace-cmd}
+Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
+It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
+Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
+
+\section{Kernelshark}\label{sec:kernelshark}
+The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
+It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
+By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
+
+\section{VARAN-Bus}\label{sec:varan}
 
@@
 The host system also needs to be aware of the real-time determinism in order to achieve highest possible reliability.
-The very step of achieving the goal of reducing latency is to apply the preempt-rt patch~\cite{RealtimePreempt_rt_versionsWiki} to the host system.
+The very step of achieving the goal of reducing latency is to apply the PREEMPT-RT patch to the host system.
 This patch is a set of modifications to the Linux kernel with the goal of making it fully preemptible.

### 2024-08-19 sigmatek/lasalclass2/PW022_hardware/Class/DiasMasterMain/DiasMasterMain.h
Deployed de6014ea with MkDocs version: 1.5.3

@@ -1,286 +1,0 @@
-#ifndef DiasMasterMain_H
-#pragma once
-#define DiasMasterMain_H
-
-//***********************************************************************************
-//** DiasMasterMain Headerfile                                                     **
-//***********************************************************************************
-
-//***********************************************
-//** Defines                                   **
-//***********************************************
-#define DIASRetryCounterOffset  16#0008		// offset of the DIAS retry counter in the control area
-#define OSVersionDIASIO		      16#1151		// OS must have a version number of at least 01.01.80 to support
-											// OS based DIAS IO
-
-// define to disable DIAS ERROR for direct accesses on DIAS (asynchron method timeout error)
-//#define DIASMASTER_DISABLE_DIAS_ERROR_ON_DIRECT_ACCESS
-
-#define DmTimeReserveIM			3			//reserve time at the end of Rt IM programm (25µs steps) (default: 3)
-
-#define DIASMASTERC_OBJECT_PATH_DIAS      16#83000002   // Object Path for DiasMaster Length = 2 (First 2byte), Domain = 0 (3rd byte), Escape Sequence DIAS (4th byte) 
-#define DIASMASTERC_OBJECT_PATH_CDIAS     16#82000002   //  Object Path for DiasMaster Length = 2 (First 2byte), Domain = 0 (3rd byte), Escape Sequence DIAS (4th byte) 
-
-#define DIASMASTERC_MAX_MODULE_NO 500 // Max number of CDIAS + DIAS Modules that are saved for HW Diagnosis
-
-
-//==========================================================================================
-// Defines für Platform Typen
-//==========================================================================================
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_hardware/Class/HwControl/HwControlLogging.h

@@ -1,168 +1,0 @@
-#ifndef HW_CONTROL_LOGGING_H
-#pragma once
-#define HW_CONTROL_LOGGING_H
-
-
-//*****************************************************************************
-//** SPAM PROTECTION SETTINGS                                                **
-//*****************************************************************************
-
-// if this is active, the logging methods only trace the strings without writing them to the log file
-//#define HWCL_DISABLE_FILE_LOGGING
-
-// if this is active, the logging methods are monitored for spam (logging too frequently)
-#define HWCL_ENABLE_SPAM_PROTECTION
-
-// log spam detection to prevent flooding of file or damaging of the media (value 0 disables this spam protection; default value 10)
-// as soon as X logs arrive within 1 second (can be adjusted via HWCL_SPAM_PROTECTION_TIME), writing into the logfile is disabled
-// if there is a pause of 1 second (can be adjusted via HWCL_SPAM_PROTECTION_TIME) the spam counter gets reset (it's possible to have X-1 logs per second to not disable the logging)
-#define HWCL_SPAM_DETECTION_LIMIT        10
-
-// time span in ms to watch for flooding (default: 100)
-#define HWCL_SPAM_DETECTION_TIME        100
-
-// time span in ms to reset the internal counter (disable log block) if there have not been any logs (default: 1000)
-#define HWCL_SPAM_BLOCK_RELEASE_TIME   1000
-
-
-
-// maximal size of logged object name in characters. if it's longer it will be shortened by HWCL_LimitObjName to fit in a memory sized HWC_MAX_OBJECT_NAME_SIZE + 1 for termination
-#define HWCL_MAX_OBJECT_NAME_SIZE  50
-
-// number of characters to take from the end when shrinking the name ("BeginningOfObjectName" + "~" + END_CHARS)
-#define HWCL_OBJECT_NAME_END_CHARS 10
-
-
-//*****************************************************************************
-//** GLOBAL FUNCTIONS                                                        **
-//*****************************************************************************
-
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_hardware/Class/SigCLib/CFileLib.cpp

@@ -1,330 +1,0 @@
-//{{LSL_ENCRYPTION 16#800
-966CA8DEF342212FF2412A957A14F9959EC0FC2BD6CD81E3EA6E87001798B2114C75595B5EB37F5883DF25ADC1AF24E9DD85E4BF80F286F454F2196C177791A921E38496737167A9F3AA0FB650407313AE2D4E42B5A5A58548CDA10137DB0782AAF45BAA
-98935EEBBB12EC6EADDFDB08B3743C188F58C570239C008618F0044FC3A2638566727CC11154967AA6F065A401D8528006684E95250C680C34BCFE4AB6B36B924F0B6DA4F5DC7A2718E811D8C6AF2A7E0287468A89786764FC81963F3CDAD5FB725DB125
-5F379DDB9B32F7DCCDE8754065FC2F468FBC60ABCEAC973CD463E11ED50876BD2F44F179F4C8C903118855E43DC74DB024F00A5221803D646C2EEF3A0890B7135D1642E0AF4D62FCF819C5820A0BD972C9CD40B9925D98AC56B26D5D7CBEB6EA1D3E28D3
-5BF9D081AC8FF74A554A68265F681F0AC1A0048C8A1F8BC2E39A9D3C76BB437ACA9B5E0F8FFB7FD8A47E750E9D803F48CBAD1C42541962EE2C406B77F88FA3C745E3387E085970378F58155A76919F6012E582A4DC2665A22196F9C0303AC59C77699A52
-606580B835351551D7DD16B887BA2A56905BA04561BFE2AE7F51D2E9AC56982031BB4FD7D1ED501AF9B9E3A151417D30BDBC4A7765C661962D226C09232F1F14DCF2B18793A96B9725B36880A177850B82FBEBF2DD076F557F8DBBA5F8A2038632EE07EC
-80A36E2CCEE6D2417E96BCA6804A2362FA2B5B7061581F96E4D5F3A34219A3576166D8C1CF51A7BD2B91F7C41C18864E99BE495FC76DA63B153094D882EA83520A7A37E34839B28FE98B2953AC5312F527D2E1E1EC05C97FA3C0BCC37069E22C11E352EA
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_hardware/Class/SigCLib/CFileLib.h

@@ -1,866 +1,0 @@
-
-#ifndef _CFileLibH
- #define _CFileLibH
-  
-  #include "DefineCompiler.h"
-
-  #ifdef cCompile
-    
-    #include "lsl_c_ifssr.h"
-    #include "RTOS_C_interfaces.h"
-    #include "file_iotypes.h"
-
-    #define sigclib_logfile(...) { char __tmp[512]; sigclib_sprintf(__tmp, __VA_ARGS__); sigclib_logfileline(__tmp); }
-    #define sigclib_logfile_datim(...) { char __tmp[512]; sigclib_sprintf(__tmp, __VA_ARGS__); sigclib_logfileline_datim(__tmp); }
-
-    // file open
-    cExtern long sigclib_fopen(char *dpne, unsigned long state);
-      // use following state: ATT_READ_WRITE, ATT_READ_ONLY, ATT_OPEN_SHARED, ATT_OPEN_NO_DIR, ATT_OPEN_DIR
-      //                      ATT_CREATE, ATT_CREATE_ALWAYS
-      // function will return value < 0 in case of error, on the other hand a handle of value >= 0
-  
-    // file close
-    cExtern void sigclib_fclose(long hdl);
-  
-    // check if file exists
-    cExtern long sigclib_fexist(const char *dpne);
-  
-    // read from file
-    cExtern unsigned long sigclib_fread(long hdl, void *dst, unsigned long length);
-  
-    // read stream from file till LF or EOF occures, CR will be ignored
-    cExtern char *sigclib_fgets(long hdl, char *s, unsigned long size);
-  
-    // read stream16 from file till LF or EOF occures, CR will be ignored
-    cExtern unsigned short *sigclib_fgets16(long hdl, unsigned short *s, unsigned long size);
-  
-    // write 0-terminated string into file
-    cExtern unsigned long sigclib_fputs(long hdl, const char *s);
-  
-    // write 0-terminated u16 string into file
-    cExtern unsigned long sigclib_fputs16(long hdl, const unsigned short *s);
-  
-    // write data into file
-    cExtern unsigned long sigclib_fwrite(long hdl, void *dst, unsigned long length);
-  
-    // write CRLF into file
-    cExtern unsigned long sigclib_fwrite_crlf(long hdl);
-
-    // write u16-CRLF into file
-    cExtern unsigned long sigclib_fwrite_crlf16(long hdl);
-  
-    // set position of filepointer
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_hardware/Class/SigCLib/CPlcOnline.cpp

@@ -1,225 +1,0 @@
-//{{LSL_ENCRYPTION 16#800
-90C49F67F342212FF2412A957A14F995F790A208958F9B7490F4C994F200A6ACC1A01E44E426F9D8C30FF01A0322CB0B3672A959AEF0324B3C5813FF6C1EE14B0E2F5ED08B54688AE5F2AD3DDA6584C76E59B738235989B1BE0B6076B37813A0448A7B18
-D3BAE8FE06B97F66BCA7ECC1311B5949515B2037E6134F8E1C78987EE9D3B20D744295BCF5691F7756FE1659B3710935B190A5B1DB748B8DECB11C7299EF4066FD3424974995D1FA13AC3943CEEFACA08DEF21C27A749D9898EE1B82971626F0951D481C
-BF1F07111AF3C1BD9F0F59F1FBA7CCB15382AAB684C13F77604F6D17D37FA1B6D99930E0CA2EE7A514F1F7E628F736972DD8E1C1A194054745E991F467560568288EC2C6D9C555A4059D64BA912BC810EA50A0EF413A6E2D4E83CDF172D89E0BED567808
-9C744D274F0F16BB88097821E056741664719EBE85B2CD29234595886B18CD83B488D28FF720F8B9F97FA619E6E0D0A15AE73CABBDED7EBB58525BC78C2F725F31182F533EFD8C79D05ACB7FBE4DB26D33CFA0335271B2BDBD81891A0E681FD083F7E20C
-7E3B6DE054F1FF2CB626DDE9D5A876C05C9447E32531E403E235580B34A3506E78B2F5D5AEF85CCF9CCAC8571E267E5391347D3DED91EEC4E80F204D1CB68BEFBA57E487CA5B6D721DC6538080259C87719F550E71AA2E2B3F58AA8A9D00DB26A1B70080
-F86B178A58F8C6B7610CEC8208018C3399764BA125ACD3D61936C506D1DB1D7CEEF89FB203800F16BC92F747DF50625E6EC9B28C9DB18CCECEE3CBA4A2CF8165DEBCAC9E8BB0E1E51871F030C4F48E48C89E6B81665883C42A36F73BD14FC8604729B449
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_hardware/Class/SigCLib/CRingBuffer.h

@@ -1,78 +1,0 @@
-// +----------------------------------------------------------------------------------------------+
-// +-[   copyright ] Sigmatek GmbH & CoKG                                                         |
-// +-[      author ] kolott                                                                       |
-// +-[        date ] 10.11.2016                                                                   |
-// +-[ description ]------------------------------------------------------------------------------+
-// |                                                                                              |
-// | NOTE:                                                                                        |
-// | Wird Ringpuffer von nur einem Task befüllt und von gleicher taskpriorität wieder ausgelesen  |
-// | so ist kein Mutex notwendig.                                                                 |
-// |                                                                                              |
-// +----------------------------------------------------------------------------------------------+
-
-#ifndef _cRingBufferH
- #define _cRingBufferH
-  
- #include "CFiFo.h"
-  
- #ifdef cCompile // *******************************************************************************
-  
-  #pragma pack(push, 1)
-   typedef struct
-   {
-     cFiFo         data; // data
-     unsigned long lost; // lost records
-   }
-   cRingBuffer;
-  #pragma pack(pop)
-  
-  cExtern cRingBuffer  *cRingBuffer_CTor(unsigned long record_size, unsigned long record_no, const char* mutex_name, void *pstatic);
-  cExtern cRingBuffer  *cRingBuffer_DTor(cRingBuffer *p);
-  
-  cExtern unsigned long cRingBuffer_Init(cRingBuffer *p, unsigned long record_size, unsigned long record_no, const char* mutex_name, void *pstatic);
-  cExtern void          cRingBuffer_Free(cRingBuffer *p);
-  cExtern unsigned long cRingBuffer_Add(cRingBuffer *p, void *pdata, unsigned long datasize);
-  cExtern unsigned long cRingBuffer_Get(void *pd, cRingBuffer *p);
-  cExtern void          cRingBuffer_Empty(cRingBuffer *p);
-  cExtern unsigned long cRingBuffer_GetUsed(cRingBuffer *p);
-  cExtern unsigned long cRingBuffer_GetUnUsed(cRingBuffer *p);
-  cExtern unsigned long cRingBuffer_GetLost(cRingBuffer *p);
-  
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_hardware/Class/SigCLib/CStrLibEx.cpp

@@ -1,235 +1,0 @@
-//{{LSL_ENCRYPTION 16#800
-B04E7B4BF342212FF2412A957A14F99528200907E1F28BBAB0F42169E8B5D47D4B4D2436245CEF63703FE833C6D52D38820C83FBB17131E4C321ED10EE1832F393585430BF9E11B9E2D72736BEC9067B6A614B34C482E2EF1AE83F98801F389524EC5104
-17C5ED019A4603ADF62863EC6151640D3FCA64B1C72EF1C534D3B86F4C473B7A16DE8F77EC5BA78604E957D8526027EE7270A636C7AC607F0C70DF997639997D9E645BEC9774F62E801152A51A144BB0190854ABA1E377E109C429CB798C9EDDAE06BE83
-7E74FF67B1DD89DEA01AA31CF5C05D1E311970118F2572C7077A99742C77C4E21730E8E67B3EE866372DD29A6D0C45B0EDDF755919D09A3A1B32E9829752C5F7F7DC9136C4BD8B655C58C0088A41A62318FF1141BD5827740E5960ED81C9CB1D25125437
-E5B05BFD28B4B0DDBD584BFBCE9C581209997F39FB8DCD9F3084B769CEB6A99FFD8D42ED50D7BED48AC26B6EAD8AA1AD620F71BD4237E4FDBF1B1B5CA40AAB9D37032085F977C63C1EEB6D5457DD9989BEDAE739440B88F8A0719495177930A3612E07A9
-1B1A7E95B3F000B0A4151B7FB84DD2639016A445C04E2E29D1325E30F485C7DD3BDCD0AA9812B971513B8EA6FDCF5BD9D188EDED89921224944788292C78071A1AB61709270BD510BBEC14F1EFCBE66E7AC4FA2BCFF2B0A2DBE878413D3C87426F10E8A8
-D2BD2FFDABB59E37D423566E120A3A67D5AF06A6B638402DFE87EA75DFE271160907858AA680901883BCF1774BF1AC756E23955EE1D6AD7E4C7C96BCE2F55E7A6E34143C5C5071450F8E8F74E3D09F9BFCB2A4C0635346206B3B5DC6524FB355693F5465
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_hardware/Class/SigCLib/CTcpIp.h

@@ -1,255 +1,0 @@
-// +----------------------------------------------------------------------------------------------+
-// +-[   copyright ] Sigmatek GmbH & CoKG                                                         |
-// +-[      author ] kolott                                                                       |
-// +-[        date ] 12.04.2019                                                                   |
-// +-[ description ]------------------------------------------------------------------------------+
-// |                                                                                              |
-// | TcpIp Interface                                                                              |
-// |                                                                                              |
-// +----------------------------------------------------------------------------------------------+
-
-#ifndef _CTcpIp_H
- #define _CTcpIp_H
- 
- #include "DefineCompiler.h" 
-
- #ifdef cCompile
-   
-  cExtern long          sigclib_tcp_ping(const char *ipv4, unsigned long bytes, unsigned long ttl, unsigned long wait);
-  cExtern long          sigclib_tcp_socket_open(void);
-  cExtern long          sigclib_tcp_socket_open_ex(long iface);
-  cExtern long          sigclib_tcp_socket_close(long socket, unsigned long soft_close);
-  cExtern long          sigclib_tcp_shutdown(long socket, unsigned long how);
-  cExtern long          sigclib_tcp_connect(long socket, const char *ipv4, unsigned long port, unsigned long timeout_ms);
-  cExtern long          sigclib_tcp_nread_available(long socket);
-  cExtern long          sigclib_tcp_recv(long socket, void *buffer, unsigned long buflen, unsigned long timeout_ms);
-  cExtern long          sigclib_tcp_send(long socket, void *buffer, unsigned long buflen, unsigned long timeout_ms);
-  cExtern long          sigclib_tcp_select(long count, long *psocket, unsigned long select_type, unsigned long flags, unsigned long timeout_ms);
-  cExtern long          sigclib_tcp_set_socket_option(long socket, long level, long option_name, void *option_value, long optionlen);
-  cExtern long          sigclib_tcp_ioctlsocket(long socket, unsigned long select_type, unsigned long onoff);
-  cExtern long          sigclib_tcp_listen(long socket, unsigned long localport, unsigned long backlogsize);
-  cExtern long          sigclib_tcp_accept(long socket, unsigned long timeout_ms);
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Class/HwControl/watchdog.h

@@ -1,69 +1,0 @@
-#ifndef  __WATCHDOG_H
-#pragma once
-#define  __WATCHDOG_H
-
-// Interface
-#define INTERFACE_WATCHDOG                 "IWATCHDOG"
-
-// defines for possible Error Codes
-#define WD_ERR_NO                 0
-#define WD_ERR_NO_WATCHDOG       -1
-#define WD_ERR_WRONG_PROP        -2
-#define WD_ERR_TASK_RUNNING      -3
-#define WD_ERR_TASK_STOPMODE     -4
-#define WD_ERR_TASK_START        -5
-#define WD_ERR_WRONG_PARAM       -6
-
-
-#define WD_PROP_APPL_WD           1        // the application take the wd trigger  param = 1 = disable auto trigger
-#define WD_PROP_PRELOAD_VALUE     2        // set/get:
-#define WD_PROP_TASK_START       10        // set: start auto trigger with param as delay im ticks
-#define WD_PROP_TASK_STOP        11        // set stop auto trigger
-#define WD_PROP_TASK_DELAY       12        // set/get: change delay param
-#define WD_PROP_TASK_STATE       13        // get:
-
-
-TYPE
-  LSLAPI_IWATCHDOG : STRUCT
-    udVersion		    : UDINT;    
-    WD_Trigger		  : pVoid;
-    WD_SetProperty	: pVoid;
-    WD_GetProperty	: pVoid;
-  END_STRUCT;
-END_TYPE
-
-////////////////////////////////////////////////////////////////////////////////
-// LASAL Function Prototypes
-////////////////////////////////////////////////////////////////////////////////
-// Callback Prototype
-FUNCTION __CDECL GLOBAL P_WD_SETPROPERTY
-  VAR_INPUT
-    Property  : UDINT;
-    value  : UDINT;
-  END_VAR
-  VAR_OUTPUT
-    retval  : DINT;
-  END_VAR;
-
-FUNCTION __CDECL GLOBAL P_WD_GETPROPERTY
-  VAR_INPUT
-    Property  : UDINT;
-    value  : ^UDINT;
-  END_VAR
-  VAR_OUTPUT
-    retval   : DINT;
-  END_VAR;
-
-FUNCTION __CDECL GLOBAL P_WD_TRIGGER
-  VAR_OUTPUT
-    retval   : DINT;
-  END_VAR;
-
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Class/I2CCDIAS/EEFuncDef.h

@@ -1,71 +1,0 @@
-#ifndef _LSL_ST_CDIAS_EEPROM
-#pragma once
-#define _LSL_ST_CDIAS_EEPROM
-
-
-TYPE
-  _LSL_CDIAS_EEPROM   :STRUCT
-    F_LSL_CDIAS_EEPROM     : pVoid;
-  END_STRUCT;
-END_TYPE
-
-
-
-FUNCTION GLOBAL __cdecl P_READ_CDIAS_EEPROM
-VAR_INPUT
-  Station   : USINT;
-  Offset 	: UINT;
-  ptr   	: pVoid;
-END_VAR
-VAR_OUTPUT
-  retval    : DINT;
-END_VAR;
-
-#define OS_READ_CDIAS_EEPROM(p1, p2, p3) _LSL_CDIAS_EEPROM^.F_LSL_CDIAS_EEPROM $ P_READ_CDIAS_EEPROM(p1, p2, p3)
-
-
-(******************************************************************************
- * Define EEPROM commands.
- ******************************************************************************)
-
-//
-// Interface Definition
-//
-TYPE
-	LSL_IFLASH : STRUCT
-
-  	// Reserved
-  	Reserved1    : pVoid;
-  	Reserved2    : pVoid;
-  	Reserved3    : pVoid;
-  	Reserved4    : pVoid;
-  	Reserved5    : pVoid;
-  	Reserved6    : pVoid;
-  	Reserved7    : pVoid;
-  	Reserved8    : pVoid;
-  	Reserved9    : pVoid;
-  	Reserved10   : pVoid;
-
-  	// Read CDIAS EEPROM
-  	EEReadCDIAS  : pVoid;
-
-  	// Read CDIAS EEPROM
-  	EEWriteCDIAS : pVoid;
-
-  END_STRUCT;
-END_TYPE
-
-// for convenience Makro for EEWriteCDIAS with pointer pFlashInfo
-FUNCTION __CDECL GLOBAL P_EEWriteCDIAS
-VAR_INPUT
-	usPlace   	: USINT;
-  	uiAddress 	: UDINT;
-  	usData		: USINT;
-END_VAR
-VAR_OUTPUT
-	retval  : DINT;
-END_VAR;
-#define IFLASH_EEWRITECDIAS(p1,p2,p3)  pFlashInfo^.EEWriteCDIAS $ P_EEWriteCDIAS(p1,p2,p3)
-
-
-#endif

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Class/SafetyCDIAS_Base/SafetyCDIAS_Base.h

@@ -1,351 +1,0 @@
-//defines for safety C-DIAS modules
-
-#ifndef SAFETYCDIAS_BASE_H
-#pragma once
-#define SAFETYCDIAS_BASE_H
-
-#ifndef SAFETY_PROTOCOL_ANALYZER_ACTIVATED
-//  #define SAFETY_PROTOCOL_ANALYZER_ACTIVATED
-  #ifdef SAFETY_PROTOCOL_ANALYZER_ACTIVATED
-    #pragma message("Safety Protocol Analyzer activated! Needs more CPU ressources!")
-  #endif
-#endif
-
-// maximum number of input PDOs (modules from which this module gets PDOs)
-#define PDO_MAX_NUMBER  100
-
-// maximum number of routing elements for one object
-#define MAX_NUMBER_ROUT_ELEMENTS  100
-
-//#define SAFETY_OVER_DIAS
-
-//define safety IDs for modules
-#define CSCP01x_ID      100
-#define CSDI16x_ID      101
-#define CSTO08x_ID      102
-
-// we need this for compatibility reasons
-   #define CSCP011_ID    100
-   #define CSDI161_ID    101
-   #define CSTO081_ID    102
-
-
-
-//define safety VARIANTs for modules
-#define CSCP011_Preliminary   1
-#define CSCP011_Series      129
-#define CSCP012_Series      130  
-#define CSDI161_Preliminary   1
-#define CSDI161_Series      129  
-#define CSDI162_Series      130  
-#define CSTO081_Preliminary   1  
-#define CSTO081_Series      129  
-#define CSTO082_Series      130  
-
-
-// define input filter in cycles (waits 5 cycles before status 0 is overtaken from firmware)
-#define SAFETY_INPUT_FILTER     5 // put this line under comment if you don't want a filter in software
-
-// sync delay times in µs (the cdias sync will be delayed for some safety modules to start working at a time where data are consistent) should work fine for all cycle times
-// 125µs has to be added to get to the cdias sync (according to FPGA department)
-#define SYNC_OFFSET_NO_DELAY          125
-#define SYNC_OFFSET_DELAY_OUTPUTS     625
-
-//define offsets for address mapping of CDIAS register
-#define OFFSET_SPDO_DATA          16#00
-#define OFFSET_SSDO_DATA          16#80
-#define OFFSET_STDO_DATA          16#C8
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Class/SigCLib/CLimits.h

@@ -1,112 +1,0 @@
-// +----------------------------------------------------------------------------------------------+
-// +-[   copyright ] Sigmatek GmbH & CoKG                                                         |
-// +-[      author ] kolott                                                                       |
-// +-[        date ] 30.04.2019                                                                   |
-// +-[ description ]------------------------------------------------------------------------------+
-// |                                                                                              |
-// +----------------------------------------------------------------------------------------------+
-
-#ifndef _CLimits_H
- #define _CLimits_H
- 
- #include "DefineCompiler.h" 
-
- #ifdef cCompile
-
-   // _int16
-   #ifndef sigclib_SHORT_MAX
-    #define sigclib_SHORT_MAX      32767
-   #endif
-   #ifndef sigclib_SHORT_MIN
-    #define sigclib_SHORT_MIN      (-(sigclib_SHORT_MAX) -1)
-   #endif
-   #ifndef sigclib_int16_MAX
-    #define sigclib_int16_MAX      sigclib_SHORT_MAX
-   #endif
-   #ifndef sigclib_int16_MIN
-    #define sigclib_int16_MIN      sigclib_SHORT_MIN
-   #endif
-
-   // _uint16
-   #ifndef sigclib_USHORT_MAX
-    #define sigclib_USHORT_MAX     65535
-   #endif
-   #ifndef sigclib_USHORT_MIN
-    #define sigclib_USHORT_MIN     0
-   #endif
-   #ifndef sigclib_uint16_MAX
-    #define sigclib_uint16_MAX     sigclib_USHORT_MAX
-   #endif
-   #ifndef sigclib_uint16_MIN
-    #define sigclib_uint16_MIN     sigclib_USHORT_MIN
-   #endif
-
-   // _int32
-   #ifndef sigclib_LONG_MAX
-    #define sigclib_LONG_MAX       2147483647L
-   #endif
-   #ifndef sigclib_LONG_MIN
-    #define sigclib_LONG_MIN       (-(sigclib_LONG_MAX) -1L)
-   #endif
-   #ifndef sigclib_int32_MAX
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Class/SigCLib/CNetworkLib.cpp

@@ -1,25 +1,0 @@
-//{{LSL_ENCRYPTION 16#800
-EA718D0BF342212FF2412A957A14F995407D8EAF667D34A0E3E5CC041D51468851D2EBCBB9F077F8C5FFEF55EF229DC22CFD054EE5B04D211099AD4FB296AF3BA0B75D693389CFC434E9C984EFB41CEABB21DA0DAD8D473EA759B6DBAC05C9917A0593EE
-44CE00D2E0EEA61588F9CBC1FEDE4E77423FEDC871D45E9A1B533FC5D94A91F89106719123BDF4C68D0CAF19F1D238AC6C5F0BF267E151746D060DF6E3CC536F40FD544BA9D5F26F7412B5B841BCFE26161A97BCA6C3F5DA58E39475DB9C5DE2C1CDD4C4
-173ABFF32DF57E8B99BAD67028F923A8EA8D3699FA4D32F4B8976C1C8A8FE0BE4046021818293C667D31E2D5630C05F3F6774F95EDE967C5D0722C1D554F26AAD82FD5842ADD26FD15BA6FF7C1479346F8BD553ED77595C4FB2A1C838F1BDEF7F779EF95
-4E514590712ECE8B027E9D34F740BFB8ED6387CABE1B74C870BD0930958CBDD4F853385BDAD7682904FDBF9A1731FCA8A50901893B1627911DA176E6441710A2860E2B275D7A85BBE8ADA21F3A4BE94918EC58E9CD873E68CCB20497FDBA131D4E4A675B
-2DBBE0243543F641EAC47FDE0481A24D92910855B6225C5B61207F3FE4594CAE4CCF75F630BF862C86BAACDA0ABA21FA61FD2CF2D4B8639ABF7BDEDA9981DBF3E910757EFFB2E0D3813366003875858F0273C94660C7D11C0CA548D736EC0F1CD4E539D1
-DEB7464E6FAB04B0CA452ACA234C0A9074C3E8185DAC4B164D2922BB31E5A340FA18B42B4CB577E34A51A41BBD68076E1202BB60C9164F5D4568E232A985C1E5728567BD58E6F551D043FB4E57BCB01C97825A47BCF1F269CF6D68F1C42B9E99538809A9
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Class/SigCLib/CSortLib.h

@@ -1,207 +1,0 @@
-
-#ifndef _CSortLibH
- #define _CSortLibH
-
-  #include "DefineCompiler.h"
-
-  #ifdef cCompile
-
-    typedef long ( *_sigclib_SFKT)    (const void *p0, const void *p1);                      // return -1( *p0 < *p1), 0( *p0 == *p1) or 1( *p0 > *p1)
-    typedef long ( *_sigclib_SFKT_EX) (const void *p0, const void *p1, const void *pcookie); // return -1( *p0 < *p1), 0( *p0 == *p1) or 1( *p0 > *p1)
-
-    // sortfunction, return <>0...list is well sorted, 0...error, nothing changed in list 
-    cExtern long sigclib_hsort(void *src0, unsigned long anz, unsigned long size, void *pfunct);
-  
-    // find index of element in sorted buffer. function will return valid index or value smaller than 0
-    cExtern long sigclib_hfind(void *src0, unsigned long anz, unsigned long size, void *pfunct, void *pvgl);
-  
-    // sortfunction
-    cExtern void sigclib_qsort(const void *base, unsigned long nmemb, unsigned long size, _sigclib_SFKT compar);
-    
-    // get pointer to element in sorted buffer, function will return valid pointer or NIL
-    cExtern void* sigclib_bsearch(const void *key, const void *base, unsigned long nmemb, unsigned long size, _sigclib_SFKT compar);
-
-    // get index to element in sorted buffer, function will return valid index or 0xFFFFFFFF
-    cExtern unsigned long sigclib_bsearch_index(const void *key, const void *base, unsigned long nmemb, unsigned long size, _sigclib_SFKT compar);
-
-    // compare 2 strings (AaBbCc...-algorithm)
-    cExtern long sigclib_sortfkt_ABC(unsigned char **pp0, unsigned char **pp1);
-                       
-    // compare 2 strings (ABCabc...-algorithm)
-    cExtern long sigclib_sortfkt_ASCII(unsigned char **pp0, unsigned char **pp1);
-
-    // sort textbuffer
-    cExtern long sigclib_sort_txtptr(char **pptxt, unsigned long no, unsigned long criteria);
-
-    // sort textbuffer
-    cExtern long sigclib_sort_txtbuffer(char *buff, unsigned long no, unsigned long criteria);
-
-    // quicksort sortfunction + pcookie in comparatorfunction
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Include/C_types.h

@@ -1,291 +1,0 @@
-//This file was generated by the LASAL2 CodeGenerator  -- 
-//Please, do not edit this file (it might be overwritten by the next generator run)
-typedef UDINT _DRIVETYPE; 
-    #define _NotFound 0
-    #define _SDD_310 1
-    #define _SDD_315 2
-    #define _SDD_105 3
-    #define _SDD_120 4
-    #define _SDD_305 5
-    #define _SDD_335 6
-    #define _SDD_215 7
-    #define _SDD_205 8
-    #define _SDD_210 9
-    #define _MDD_100 10
-    #define _S_340 11
-    #define _SDD_115 12
-    #define _SDD_1300 811237
-    #define _SDD_1400 811238
-    #define _SDD_1500 811239
-    #define _SDD_1600 811240
-    #define _MDD_2000 811300
-    #define _VSDC_151 811301
-    #define _VSDC_152 811302
-    #define _DC061 881042
-    #define _DC062 881801
-    #define _SR011 881802
-    #define _SR012 881803
-    #define _DC101 881804
-    #define _DC102 881805
-    #define _WA011 991800
-    #define _WA012 991801
-
-#pragma pack(push, 1)
-typedef struct _I_HC_Type   //! <Type Comment="Hardware Code an axis&#13;&#10;&#13;&#10;Bit 31..16 Hardware Code of the Control Board&#13;&#10;Bit 15...8 Software version of the feedback CPLD&#13;&#10;Bit   7...0 Hardware Code of the Power Stage" Name="_I_HC_Type"/>
-  {
-    HSINT HardwareCodePowerStage;  //! <Type Comment="Hardware Code of the Power Stage (Default: 16#80)" Name="_I_HC_Type.HardwareCodePowerStage"/>
-    HSINT FPGAVersion;  //! <Type Comment="Software version of the feedback CPLD" Name="_I_HC_Type.FPGAVersion"/>
-    HINT HardwareCodeControlBoard;  //! <Type Comment="Hardware Code of the Control Board (Default: 16#0000)" Name="_I_HC_Type.HardwareCodeControlBoard"/>
-  } _I_HC_Type;
-#pragma pack(pop)
-
-typedef struct 
-  {
-    long NoHW : 1;
-    long NoParameter : 1;
-    long NotOnline : 1;
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Source/interfaces/lsl_st_gpio.h

@@ -1,289 +1,0 @@
-#ifndef  __GPIO_H
-#pragma once
-#define  __GPIO_H
-
-// Interface
-#define INTERFACE_GPIO                       "GPIO"
-
-// defines for possible Error Codes
-#define ERR_GPIO_NOT_AVAIL			  -1		// GPIO is not available
-#define ERR_GPIO_FUNC_NOT_AVAIL		-2		// Function not available for this GPIO
-#define ERR_GPIO_INVAL_PARAM	  	-3		// Parameter is invalid
-#define ERR_GPIO_INVAL_SRC		  	-4		// Hardware source
-#define ERR_GPIO_INVAL_IDX		  	-5		// invalid index
-#define ERR_GPIO_INVAL_BITSIZE		-6		// invalid bitsize
-#define ERR_GPIO_INVAL_REGISTER		-7		// invalid register
-#define ERR_GPIO_ACCESS_DENIED		-8		// read or write not allowed
-
-// typeFlags for Output
-#define GPIO_TYPEFLAG_OUTPUT_ONE_CYCLIC		0x01
-// typeFlags for Input
-#define GPIO_TYPEFLAG_INPUT_HAS_IRQ			0x01
-#define GPIO_TYPEFLAG_INPUT_HAS_LATCHED		0x02
-
-// available registered propberty
-#define GPIO_REGTYPE_LATCHED				1
-#define GPIO_REGTYPE_RISING					2
-#define GPIO_REGTYPE_FALLING				3
-
-
-// Defines for possible Flags
-#define GPIO_FLAG_TRIG_RISING		  0x01	// Trigger IRQ with rising Edge
-#define GPIO_FLAG_TRIG_FALLING		0x02	// Trigger IRQ with falling Edge
-#define GPIO_FLAG_REMOVE_HANDLER	0x04	// Remove handler for the GPIO
-
-(* Defines for possible Values of a GPIO *)
-#define IO_OFF		0
-#define IO_ON		  1
[…]

### 2024-08-19 sigmatek/lasalclass2/PW022_virtualization/Source/interfaces/lsl_st_safetydll.h

@@ -1,1779 +1,0 @@
-(****************************************************************************************)
-(*                                                                                      *)
-(* lsl_st_safetydll.h                                                                   *)
-(*                                                                                      *)
-(* enthält die Definitionen für das Safety DLL Interface                                *)
-(*                                                                                      *)
-(*                                                                                      *)
-(****************************************************************************************)
-(*
-  Die Funktionen des 'Safety DLL Interface' werden verwendet, um von einer Standard-SPS 
-  aus eine Kommunikationsverbindung zu einer Safe-CPU aufzubauen und folgende 
-  Abläufe auszuführen:
-  - Download einer Konfiguration
-  - Setzen des Verified Flags
-  - Ändern des Passworts
-  - Starten der Safe-CPU (verlassen des Service-Modes)
-  - Stoppen der Safe-CPU (herstellen des Service-Modes)
-  - Statusabfrage
-  - Neustart der Safety-Applikation
-  - Firmware Download
-  - Parameterlisten Download und auslesen
-  
-  Hinweis:
-  Die Funktionen dürfen nicht dafür verwendet werden, um vollautomatisch einen Ablauf 
-  durchzuführen. Es muss vom Programmersteller sichergestellt werden dass die Eingabedaten 
-  vom Benutzer stammen und nicht im Programm hardcodiert werden.
-  
-  Das Interface ISAFETY_DLL2 hat den gleichen Funktionsumfan wie das Interface ISAFETY_DLL.
-  Es können damit aber mehrere Abläufe zu verschiedenen Safety-CPUs gleichzeitig 
-  ausgeführt werden.
[…]

### 2024-08-20 Masterarbeit_Robotik_Pamuk.tex

@@
 
+QEMU is an open-source emulator and virtualizer that allows to run operating systems and programs for one machine on a different machine.
+It can emulate different hardware architectures and is used in this thesis to emulate Salamander 4.
+When QEMU is used together with Kernel-based Virtual Machine (KVM), it can provide near-native performance by running code directly on the host CPU.
+The QEMU script used to emulate Salamander 4 and the meaning of the options are described in detail in Subsection \ref{subsec:salamander4-virtualization}.
+
 \section{Trace-cmd}\label{sec:trace-cmd}
-Trace-cmd can be used for tracing the Linux kernel~\cite{Tracecmd}.
-It can record various kernel events such as interrupts, scheduler decisions, file system activity, function calls in real time.
-Trace-cmd helped in getting detailed insights into system behaviour and identify reasons for latency.
+\texttt{Trace-cmd}~\cite{Tracecmd} is the front-end tool of the \texttt{ftrace} tool that is built inside the Linux kernel and used for tracing the Linux kernel.
+\texttt{Trace-cmd} can record detailed reports of kernel events such as interrupts, scheduler decisions, function calls, and custom events in real time.
+It is easier with this tool to enable and disable tracing, set filters, and view results.
+\texttt{Trace-cmd} helped debugging and identifying the reasons for latency in Salamander 4 to continuously improve its real-time performance and determinism.
+A very practical aspect of \texttt{trace-cmd} is that the guest can be traced from the host.
+For this purpose, the guest needs to be configured up with \texttt{vsocks}, which are a virtual socket that enable direct communication between the host and guest.
+To enable vsockets and tracing, the guest kernel requires the following configurations, according to Linux developer and maintainer Steve Rostedt~\cite{rostedtTracecmdHostGuest}.
+\bigskip
+\begin{lstlisting}[language=bash, caption={Kernel Configuration for vsocks and tracing}, label={lst:kernel-config}]
+	# Vsockets settings
+	CONFIG_VSOCKETS=m
+	CONFIG_VHOST_VSOCK=m
+	CONFIG_VIRTIO_VSOCKETS=m
+	CONFIG_VIRTIO_VSOCKETS_COMMON=m
+	CONFIG_VSOCKETS_DIAG=m
+	CONFIG_VSOCKETS_LOOPBACK=m
+	# Tracing settings
+	CONFIG_TRACING=y
+	CONFIG_FTRACE=y
+	CONFIG_FUNCTION_TRACER=y
+	CONFIG_FUNCTION_GRAPH_TRACER=y
+	CONFIG_DYNAMIC_FTRACE=y
+	CONFIG_DYNAMIC_FTRACE_WITH_REGS=y
+	CONFIG_DYNAMIC_FTRACE_WITH_DIRECT_CALLS=y
+	CONFIG_DYNAMIC_FTRACE_WITH_ARGS=y
+	CONFIG_SCHED_TRACER=y
+	CONFIG_FTRACE_SYSCALLS=y
+	CONFIG_TRACER_SNAPSHOT=y
+	CONFIG_KPROBE_EVENTS=y
+	CONFIG_UPROBE_EVENTS=y
+	CONFIG_BPF_EVENTS=y
+	CONFIG_DYNAMIC_EVENTS=y
+	CONFIG_PROBE_EVENTS=y
+	CONFIG_SYNTH_EVENTS=y
+	CONFIG_HIST_TRIGGERS=y
+	\end{lstlisting}
+
+	\clearpage
+
+	\noindent After compiling the kernel with the additional settings aboce, the guest needs access to the vsocket.
+In QEMU, the line \texttt{-device vhost-vsock-pci,guest-cid=3,id=vsock0} was added.
+This command configures a virtual socket (vsock) device in QEMU, assigning the guest a unique Context Identifier (CID) of 3 for direct communication with the host.
+The CID can be chosen as desired.
+Once the guest is booted with vhost enabled, the \texttt{sudo trace-cmd agent} command can be run.
+The agent now listens to connections on port 823 by default with the specified CID, 3 in this case.
+The host can trace itself and the guest from this moment going forward.
+For example, \texttt{sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all} records all events from the host and guest.
+\texttt{sudo trace-cmd record -e kvm:kvm\_entry -e kvm:kvm\_exit -A @3:823 --name Salamander4 -e all} traces onkly the \texttt{kvm\_entry} and \texttt{kvm\_exit} events of the host along with the exit reasons and all guest events.
+As a last example, \texttt{sudo trace-cmd record -e kvm -e sched -e irq -e -A @3:823 --name Salamander4 -e all} enables tracing of KVM, scheduling, and interrupt events on the host, for capturing the KVM events that manage the guest, scheduling activities, and interrupt handling respectively.
 
 	\section{Kernelshark}\label{sec:kernelshark}
-The data that was recorded by trace-cmd was then fed into Kernelshark, which is a graphical front-end tool~\cite{KernelShark}.
-It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitated the process of identifying patterns and correlations between events.
-By further filtering the displayed events according to specific criteria such as processes, event types or time ranges, the latency issues were analyzed.
+Viewing the tracing data that was recorded with \texttt{trace-cmd} on a text-based file can be laborious.
+A better way of analyzing the data is done through \texttt{KernelShark}~\cite{KernelShark}, which is a graphical front-end tool.
+It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitates the process of identifying patterns and correlations between events.
+The latency issues can be analyzed by further filtering the displayed events according to processes, event types or time ranges.
+After the recording, \texttt{trace-cmd} gernerates a host \texttt{trace.dat} file and a guest \texttt{trace-Salamander4.dat} file, which in this case can be loaded into \texttt{KernelShark} with the command \texttt{kernelshark trace.dat -a trace-Salamander4.dat}.
+Once both files are loaded, the trace data can be visually inspected.
+All the current CPUs for both the host and the guest can be turned off, and the KVM Combo plots menu can be selected to view the guest vCPUs plotted on top of the host threads that run them, as depicted in Figure~\ref{fig:kernelshark_combo}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{img/kernelshark_combo.png}
+	\caption[Guest vCPUs plotted on top of host threads in KernelShark]{Guest vCPUs plotted on top of host threads in KernelShark}
+	\label{fig:kernelshark_combo}
+\end{figure}
 
 \section{VARAN-Bus}\label{sec:varan}
+
+VARAN is a real-time Ethernet bus designed for industrial automation, capable of connecting larger systems and machines with smaller components and sensors in hard real-time.
+This bus operates on the manager/client principle, where each network includes a manager and one or more subordinate clients.
+Multiple VARAN networks can be linked together into a single synchronized network using a higher-level VARAN manager.
+The jitter between these systems is kept below 100 ns.
+This synchronization between systens and their data exchange is illustrated in Figure~\ref{fig:multiple_varans}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.75\columnwidth]{img/multiple_varans.png}
+	\caption[Synchronization between Systems and their Data Exchange]{Synchronization between Systems and their Data Exchange}
+	\label{fig:multiple_varans}
+\end{figure}
+
+\clearpage
 

### 2024-08-20 Masterarbeit_Robotik_Pamuk.tex
3.1 Initial Situation, 3.1.1 Salamander 4 Bare Metal, 3.1.2 Salamander 4 Virtualization

@@
 Xenomai 3 is detailed in Section~\ref{sec:xenomai}.
-	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal through real-time performance tunings, focussing on latency requirements set by SIGMATEK.
+	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal (hardware) through real-time performance tunings, focussing on latency requirements set by SIGMATEK.
 	      These modifications are subject of Chapter~\ref{sec:real-time_tunings}.
 	\item The \texttt{latency} tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to use by SIGMATEK.
-The goal was set to achieving latency values below 50 microseconds in the period of the measuremenet.
+The goal was set to achieving latency values below 50 microseconds in the duration of the measuremenet.
 	\item SIGMATEK uses the VARAN bus for communication between machines and systems, a real-time Ethernet network that was specially developed for industrial automation.
@@
 
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
+\vspace{2em}
 \begin{lstlisting}[name={System information},label={output:uname_a_virt}]
@@
 \end{lstlisting}
-\end{minipage}
 
@@
 
+\section{QEMU}\label{sec:qemu}
+
+QEMU~\cite{QEMU}.
+is an open-source emulator and virtualizer that allows to run operating systems and programs for one machine on a different machine.
+It can emulate different hardware architectures and is used in this thesis to emulate Salamander 4.
+When QEMU is used together with Kernel-based Virtual Machine (KVM), it can provide near-native performance by running code directly on the host CPU.
+The QEMU script used to emulate Salamander 4 and the meaning of the options are described in detail in Subsection \ref{subsec:salamander4-virtualization}.
+
 \section{Yocto}\label{sec:yocto}
 
-Salamander 4 is built with Yocto, an open-source project that helps develop custom Linux-based operating systems for embedded devices.
+Salamander 4 is built with Yocto~\cite{WelcomeYoctoProject}, an open-source project that helps develop custom Linux-based operating systems for embedded devices.
 It is not a Linux version itself but a framework to make custom versions.
@@
 It builds the operating system with BitBake and automates the compilation and assembly of software.
-Yocto can generate a Software Development Kit (SDK) containing tools and libraries for developing applications for the custom OS~\cite{WelcomeYoctoProject}.
+Upon generating the necessary files, Yocto provides a QEMU folder with the following components shown in Code~\ref{lst:ls}.
+
+\vspace{2em}
+\begin{lstlisting}[name={Contents of QEMU folder for Salamander 4},label={lst:ls}]
+    sigma_ibo@localhost:~/Desktop/salamander-image$ ls -1
+    bzImage
+    drive-c
+    ovmf.code.qcow2
+    qemu_def.sh
+    salamander-image-sigmatek-core2.ext4
+    stek-drive-c-image-sigmatek-core2.tar.gz
+    vmlinux
+    \end{lstlisting}
+
+	\noindent The following is a description of the components used for the virtualization of Salamander 4.
+
+\begin{itemize}
+	\item \textbf{bzImage}: Compressed Linux kernel image that is loaded by QEMU at system start. The short form ``bz`` stands for big-zipped.
+	\item \textbf{drive-c}: Directory serving as C drive for QEMU, created and filled by qemu\_def.sh script.
+	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process.
+~\texttt{OVMF} stands for Open Virtual Machine Firmware,~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for ``QEMU Copy On Write version 2''.
+	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described later in more detail.
+	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform.
+It uses the~\texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
+	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Compressed tarball containing a pre-configured environment for the Salamander 4 OS.
+It is unpacked and sets up the~\texttt{drive-c/} directory with system and log files in the~\texttt{qemu\_def.sh} script.
+	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging.
+\end{itemize}
+
+\noindent The inital QEMU script after the custom Yocto build and the starting point for this work is shown in Code~\ref{script:qemu_def}.
+This script is used to start QEMU with required parameters to boot Salamamder 4 OS.
+It will be adjusted in Chapter~\ref{sec:real-time_tunings} in order to accompany real-time performance tunings.
+
+\vspace{2em}
+\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualization},label={script:qemu_def}]
+	#!/bin/sh
+
+	if  [ ! -d drive-c/ ]; then
+			echo "Filling drive-c/"
+			mkdir drive-c/
+			tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
+	fi
+
+	exec qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
+	-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
+	-append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4" \
+	-net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
+	-fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
+	-device vhost-vsock-pci,guest-cid=3,id=vsock0 \
+	-drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
+	-no-reboot -nographic
+\end{lstlisting}
+
+\vspace{2em}
+
+\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in Section~\ref{sec:host_operating_system}.
+
+\clearpage
 
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
+	\includegraphics[width=0.6\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
 	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
@@
 
-\clearpage
-
 \noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
+	\includegraphics[width=0.6\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
 	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
@@
 The tool creates real-time tasks or interrupt handlers and measures the difference between the expected and actual execution times.
-
-\section{QEMU}\label{sec:qemu}
-
-QEMU is an open-source emulator and virtualizer that allows to run operating systems and programs for one machine on a different machine.
-It can emulate different hardware architectures and is used in this thesis to emulate Salamander 4.
-When QEMU is used together with Kernel-based Virtual Machine (KVM), it can provide near-native performance by running code directly on the host CPU.
-The QEMU script used to emulate Salamander 4 and the meaning of the options are described in detail in Subsection \ref{subsec:salamander4-virtualization}.
 
@@
 To enable vsockets and tracing, the guest kernel requires the following configurations, according to Linux developer and maintainer Steve Rostedt~\cite{rostedtTracecmdHostGuest}.
-\bigskip
+
+\vspace{2em}
 \begin{lstlisting}[language=bash, caption={Kernel Configuration for vsocks and tracing}, label={lst:kernel-config}]
@@
 \section{Initial Situation}\label{sec:starting_situation}
-As a starting point, initial latency values of both the bare metal and virtualization versions were measured with the \texttt{latency} tool of the xenomai tool suite.
-Salamander 4 bare metal refers to the proprietary hardware of SIGMATEK used to employ the custom operating system.
-Salamander 4 virtualization refers to a virtual version of the Salamander 4 hardware platform, achieved through QEMU/KVM.
-Subsections~\ref{subsec:salamander4-bare-metal} and~\ref{subsec:salamander4-virtualization} specify the details of the measurements for both versions.
-In the further course, the aim was to bring the latency values of the virtualization closer to those of the hardware and guarantee deterministic and reliable behavior.
+As a starting point, initial latency values of both the bare metal and virtualization versions were measured with the \texttt{latency} tool of the Xenomai tool suite.
+Salamander 4 bare metal refers to the custom operating system developed by SIGMATEK, running on their proprietary hardware.
+Salamander 4 virtualization refers to a virtual version of Salamander 4, achieved through QEMU/KVM.
+Subsections~\ref{subsec:salamander4-bare-metal} and~\ref{subsec:salamander4-virtualization} specify the details of the measurements for both setups.
+In the further course of this master's thesis, the aim is to bring the latency values of the virtualization closer to those of the hardware and guarantee deterministic and reliable behavior.
 
 \subsection{Salamander 4 Bare Metal}\label{subsec:salamander4-bare-metal}
-The output of the command~\texttt{uname -a} for Salamander 4 bare metal is shown in Code~\ref{output:uname_a_hw}.
-
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={Salamander 4 bare metal system information},label={output:uname_a_hw}]
-root@sigmatek-core2:~# uname -a 
-Linux sigmatek-core2 5.15.94 #1 SMP PREEMPT IRQPIPE Tue Feb 14 18:18:05 UTC 2023 x86_64 GNU/Linux
-\end{lstlisting}
-\end{minipage}
-
-\noindent As a reference point, the latency program was executed on Salamander 4 bare metal for a duration of 10 minutes.
-The complete command used was~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of min, avg, max latencies in a Gnuplot-compatible format to the file~\texttt{gnuplot.txt}.
-Figure~\ref{fig:max_latency_hardware} shows the gathered latency values in microseconds.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/max_latency_hardware.png}
-	\caption[Latency of Salamander 4 bare metal]{Latency of Salamander 4 bare metal}
-	\label{fig:max_latency_hardware}
-\end{figure}
-
-\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_statistics_bare_metal}.
-It gives an overview of the average, maximum, minimum latency, and the standard deviation of the latency values in microseconds.
-
-\begin{table}[H]
-	\centering
-	\caption{Latency statistics of Salamander 4 bare metal in microseconds}
-	\label{tab:latency_statistics_bare_metal}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|c|c|}\hline
-			\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
-			Average Latency    & 4.06                       \\\hline
-			Maximum Latency    & 10.71                      \\\hline
-			Minimum Latency    & 2.82                       \\\hline
-			Standard Deviation & 0.85                       \\\hline
-		\end{tabular}}
-\end{table}
-
-\noindent Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time. Since the data varies strongly, a logarithmic scale was used for both axes.
+As a reference point, the \texttt{latency} program was executed on Salamander 4 bare metal for a duration of 10 minutes.
+The complete command used was~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of minimum, average and maximum latencies in a Gnuplot-compatible format to the file~\texttt{gnuplot.txt}.
+Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time.
 
@@
 
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_hardware} and Table~\ref{tab:latency_overrun_msw_hardware}.
+The test was conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
+
+\vspace{2em}
+\begin{table}[H]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
+min & 599 & 0.711 & 0.454 \\ \hline
+avg & 5 999 988 & 1.019 & 0.150 \\ \hline
+max & 599 & 3.528 & 0.895 \\ \hline
+\end{tabular}
+\caption{Statistical Summary of Latency Measurements}
+\label{tab:latency_stats_hardware}
+\end{table}
+
+\vspace{1em}
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|c|}
+	\hline
+	\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
+	0.613 & 1.380 & 10.709 & 0 & 0 \\ \hline
+	\end{tabular}
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_hardware}
+	\end{table}
+
 \subsection{Salamander 4 Virtualization}\label{subsec:salamander4-virtualization}
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
-It was developed using Yocto, an open-source project that allows customised Linux distributions to be created for embedded systems.
-Upon generating the necessary files, Yocto provides a QEMU folder with the following components shown in Code~\ref{lst:ls}.
-QEMU is the environment in which the virtualization runs, as it is an open-source tool for hardware virtualization~\cite{QEMU}.
-
-\vspace{1em}
-\begin{minipage}{\linewidth}
-	\begin{lstlisting}[name={Contents of QEMU folder for Salamander 4},label={lst:ls}]
-    sigma_ibo@localhost:~/Desktop/salamander-image$ ls -1
-    bzImage
-    drive-c
-    ovmf.code.qcow2
-    qemu_def.sh
-    salamander-image-sigmatek-core2.ext4
-    stek-drive-c-image-sigmatek-core2.tar.gz
-    vmlinux
-    \end{lstlisting}
-\end{minipage}
-
-\noindent With the help of the script depicted in Code~\ref{script:qemu_def}, Salamander 4 is started together with the necessary hardware components in the QEMU environment.
-This makes it possible to run Salamander 4 on a variety of host systems, regardless of the specific hardware of the host.
-The following is a description of the components used for the virtualization of Salamander 4.
-\clearpage
-\begin{itemize}
-	\item \textbf{bzImage}: Compressed Linux kernel image that is loaded by QEMU at system start. ``bz`` stands for big-zipped.
-	\item \textbf{drive-c}: Directory serving as C drive for QEMU system, created and filled by qemu\_def.sh script.
-	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process.
-~\texttt{OVMF} stands for Open Virtual Machine Firmware,~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for ``QEMU Copy On Write version 2''.
-	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described in Code~\ref{script:qemu_def}.
-	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform.
-It uses the~\texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
-	\item \textbf{stek-drive-c-image-sigmatek-core2.tar.gz}: Compressed tarball containing a pre-configured environment for the Salamander 4 OS.
-It is unpacked and sets up the~\texttt{drive-c/} directory with system and log files in the~\texttt{qemu\_def.sh} script.
-	\item \textbf{vmlinux}: Uncompressed Linux kernel image, typically used for debugging.
-\end{itemize}
-
-\noindent The inital QEMU script after the custom Yocto build and the starting point for this work is shown in Code~\ref{script:qemu_def}.
-This script is used to start QEMU with required parameters to boot Salamamder 4 OS.
-It will be adjusted in Chapter~\ref{sec:real-time_tunings} in order to accompany real-time performance tunings.
-
-\vspace{1em}
-\begin{minipage}{\linewidth}
-	\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualization},label={script:qemu_def}]
-	#!/bin/sh
-
-	if  [ ! -d drive-c/ ]; then
-			echo "Filling drive-c/"
-			mkdir drive-c/
-			tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
-	fi
-
-	exec qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
-	-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
-	-append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4" \
-	-net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
-	-fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
-	-device vhost-vsock-pci,guest-cid=3,id=vsock0 \
-	-drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
-	-no-reboot -nographic
-\end{lstlisting}
-\end{minipage}
-
-\noindent This script is run on a generic Ubuntu 22.04.4 system, as mentioned previously in Chapter~\ref{cha:methodology}.
-The kernel version and other details are presented in Code~\ref{output:uname_a_ub}, using the~\texttt{uname -a} command.
-
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
-	\begin{lstlisting}[name={Ubuntu 22.04.4 system information},label={output:uname_a_ub}]
-root@sigmatek-core2:~# uname -a 
-Linux sigma-ibo 6.5.0-35-generic #35~22.04.1-Ubuntu SMP PREEMPT_DYNAMIC Tue May  7 09:00:52 UTC 2 x86_64 x86_64 x86_64 GNU/Linux
-\end{lstlisting}
-\end{minipage}
-
-\noindent Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, the following latency values in Figure~\ref{fig:max_latency_default} were collected.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/max_latency_default.png}
-	\caption[Latency with default settings]{Latency with default settings}
-	\label{fig:max_latency_default}
-\end{figure}
-
-\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_statistics_virt}.
-It gives an overview of the average, maximum, minimum latency, and the standard deviation of the latency values in microseconds.
-
-\begin{table}[H]
-	\centering
-	\caption{Latency statistics of default Salamander 4 virtualization in microseconds}
-	\label{tab:latency_statistics_virt}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|c|c|}\hline
-			\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
-			Average Latency    & 46.22                      \\\hline
-			Maximum Latency    & 707.62                     \\\hline
-			Minimum Latency    & 15.59                      \\\hline
-			Standard Deviation & 52.13                      \\\hline
-		\end{tabular}}
-\end{table}
-
-\noindent Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency over the course of said time with default settings.
+Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency.
 
@@
 
-\noindent Comparing these values to those of bare metal in Figure~\ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in the statistics.
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt} and Table~\ref{tab:latency_overrun_msw_new_virt}.
+The test was againconducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
+
+\vspace{2em}
+\begin{table}[H]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
+min & 599 & 3.713 & 1.355 \\ \hline
+avg & 5,999,922 & 8.247 & 2.521 \\ \hline
+max & 599 & 45.705 & 52.196 \\ \hline
+\end{tabular}
+\caption{Statistical Summary of Latency Measurements}
+\label{tab:latency_stats_virt}
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
+2.536 & 8.940 & 707.622 & 43 & 0 \\ \hline
+\end{tabular}
+\caption{Latency, Overrun, and Missed Switch Data}
+\label{tab:latency_overrun_msw_new_virt}
+\end{table}
+
+\subsection{Comparison}
+\noindent Comparing the latency values of the virtualization to those of bare metal in Figure~\ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in the statistics.
 A maximum latency of 707.62~\textmu s is not tolerable for the system and needs to be tuned.
@@
 
-\section{Real-Time Performance Tuning}\label{sec:real-time_tunings}
+\section{Real-Time Performance Tunings}\label{sec:real-time_tunings}
 
@@
 The very first step in this regard is the isolation of a CPU to dedicate it solely to the desired real-time task.
-CPU isolation and affinity are covered in detail later in Subsection~\ref{subsubsec:cpu_isolation}, but its impact are briefly demonstrated here.
+CPU isolation and affinity are covered in detail later in Subsection~\ref{subsubsec:cpu_isolation}, but their impact are briefly demonstrated here.
 
@@
 
-After each tuning, the latency test of Xenomai is executed to evaluate the improved latency of the system.
-
-\textcolor{red}{After Article, Figure~\ref{fig:max_latency_rt}}
+After each tuning, the latency test of Xenomai is executed to evaluate the improved latency of the system. \textcolor{red}{After Article, Figure~\ref{fig:max_latency_rt}}
 
@@
 
+\vspace{2em}
 \begin{table}[H]
@@
 
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
+\vspace{2em}
 \begin{lstlisting}[name={Kernel Configuration},label={script:kernel_cli}]
@@
 
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
+\vspace{2em}
 \begin{lstlisting}[name={User and Kernel Tasks},label={output:pse},escapeinside={(*@}{@*)}]
@@
 \end{lstlisting}
-\end{minipage}
+
 \subsubsection{KVM Entry and KVM Exit}\label{subsubsec:kvm_entry_exit}
@@
 
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
+\vspace{2em}
 \begin{lstlisting}[name={Check distribution of interrupt requests across each CPU},label={script:smp_affinity}]
@@
 
-\vspace{1em}
-\begin{minipage}{0.95\columnwidth}
+\vspace{2em}
 \begin{lstlisting}[name={Change IRQ assignment of a CPU},label={script:change_smp_affinity}]
@@
 \end{lstlisting}
-\end{minipage}
+
 \subsubsection{RT-Priority}
@@
 
-\vspace{1em}
-\begin{minipage}{\linewidth}
+\vspace{2em}
 \begin{lstlisting}[name={Tuned QEMU script for starting Salamander 4 virtualization},label={script:qemu_def_tuned}]
@@
 	\end{lstlisting}
-\end{minipage}
 
@@
 
-\vspace{1em}
-\begin{minipage}{\linewidth}
+\vspace{2em}
 \begin{lstlisting}[name={Include PCI in QEMU script for Salamander 4 virtualization},label={script:qemu_def_pci}]
@@
 \section{Host and Guest tasks}
+\vspace{2em}
 \begin{table}[H]
@@
 \end{tabular}
-	\end{minipage}%
+%
 \begin{minipage}{.5\textwidth}
@@
 \end{tabular}
-	\end{minipage}
+
 \end{table}
@@
 \section{Citation table}
+\vspace{2em}
 \begin{table}[H]

### 2024-08-21 Masterarbeit_Robotik_Pamuk.tex
rauh meeting

@@
 \place{Wien}
-\kurzfassung{Erstellung einer Echtzeit-Robotersteuerungsplattform unter Verwendung von Salamander OS, Xenomai, QEMU
-und PCV-522 in der Yocto-Umgebung. Die Plattform basiert auf Salamander OS und nutzt Xenomai für Echtzeit-
-Funktionen. Dazu muss im ersten Schritt die Virtualisierungsplattform evaluiert werden. (QEMU, Hyper-V, Virtual
-Box, etc.) Als weiterer Schritt folgt die Anbindung eines Roboters über eine VARAN-Bus Schnittstelle. Das
-gesamte System wird in der Yocto-Umgebung erstellt und konfiguriert.
-Das Hauptziel der Arbeit ist es, herauszufinden, wie die Integration von Echtzeit-Funktionen und effizienten
-Kommunikationssystemen in eine Robotersteuerungsplattform die Reaktionszeit und Zuverlässigkeit von
-Roboteranwendungen verbessern kann}
-\schlagworte{Schlagwort1, Schlagwort2, Schlagwort3, Schlagwort4}
-\outline{Subsections~\ref{subsec:salamander4-bare-metal} and~\ref{subsec:salamander4-virtualization} demonstrate the inital real-time latency values gathered for bare metal and virtualization. 
-}
+\kurzfassung{Test}
+\schlagworte{Echtzeit, Virtualisierung, Xenomai, VARAN}
+\outline{Test}
 \keywords{Echtzeit, Virtualisierung, Xenomai, VARAN}
@@
 	\end{tikzpicture}
-	\caption{Flowchart of operations in a virtualized environment}
+	\caption{Flowchart of operations in a virtualized environment (own figure, inspired by~\citeauthor{maPerformanceTuningKVMbased})}
 	\label{fig:kvm_stages}
@@
 
+\section{Approach}\label{sec:approach}
+Wasserfall, rechtfertigen und in Diskussion Vor und Nachteile
+
 \clearpage
@@
 \noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt} and Table~\ref{tab:latency_overrun_msw_new_virt}.
-The test was againconducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
+The test was again conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
 
@@
 
-\subsection{Comparison}
+\subsection{Initial Comparison}
 \noindent Comparing the latency values of the virtualization to those of bare metal in Figure~\ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in the statistics.
-A maximum latency of 707.62~\textmu s is not tolerable for the system and needs to be tuned.
+A maximum latency of 707.62~\textmu s is not tolerable for a real-time virtualization and needs to be tuned.
 
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/01combined/gnuplot_combined_max_latency.png}
-	\caption[Comparison of variation in latency between hardware and virtualization]{Comparison of variation in latency between hardware and virtualization}
+	\caption[Initial Comparison of Variation in Latency between Hardware and Virtualization]{Initial Comparison of Variation in Latency between Hardware and Virtualization}
 	\label{fig:gnuplot_max_latency_combined}
@@
 This significantly reduces the time it takes for high-priority tasks to start executing after an event~\cite{RealtimeKernelPatchset}.
-The patch also includes support for high-resolution timers and make the system more predictable, which are crucial for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling~\cite{rostedtInternalsRTPatch}.
+The patch also includes support for high-resolution timers and makes the system more predictable.
+Determinism and predictability are key for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling~\cite{rostedtInternalsRTPatch}.
 
@@
 The next sections will deal with these real-time performance tunings.
-The very first step in this regard is the isolation of a CPU to dedicate it solely to the desired real-time task.
-CPU isolation and affinity are covered in detail later in Subsection~\ref{subsubsec:cpu_isolation}, but their impact are briefly demonstrated here.
-
-Figure~\ref{fig:max_latency_taskset} shows latency of QEMU taskset Salamander4.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{img/max_latency_taskset.png}
-	\caption[Latency taskset]{Latency taskset}
-	\label{fig:max_latency_taskset}
-\end{figure}
-
-After each tuning, the latency test of Xenomai is executed to evaluate the improved latency of the system. \textcolor{red}{After Article, Figure~\ref{fig:max_latency_rt}}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/3rt/max_latency_rt/max_latency_rt.png}
-	\caption[After Article configurations]{After Article configurations}
-	\label{fig:max_latency_rt}
-\end{figure}
-
-\vspace{2em}
-\begin{table}[H]
-	\centering
-	\caption[Latency statistics of Salamander 4 after BIOS configurations]{Latency statistics of Salamander 4 virtualization after BIOS configurations in microseconds}
-	\label{tab:latency_statistics_bios}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|c|c|}\hline
-			\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
-			Average Latency    & 17.68                      \\\hline
-			Maximum Latency    & 32.21                      \\\hline
-			Minimum Latency    & 8.005                      \\\hline
-			Standard Deviation & 6.1                        \\\hline
-		\end{tabular}}
-\end{table}
-
+After each tuning, the \texttt{latency} test of Xenomai is executed to evaluate the improved latency.
+The most important metric of a latency test is without a doubt the worst latency value, because the system is only as reliable as its worst measurement.
+Since the period of the tests is set to 100\textmu s, there are approximately 100000 samples per second.
+
+\clearpage
 \subsection{BIOS Configurations}\label{subsec:bios_configurations}
@@
 \begin{itemize}
-	\item \textbf{Hyper Threading}: Hyper-Threading allows CPUs to process two threads simultaneously instead of just one.
-When it is enabled on the host, this allows the parallelisation of tasks and increases performance.
+	\item \textbf{Hyper Threading}: When hyper-threading is enabled in thd BIOS settings, this allows CPUs to work on two threads simultaneously instead of just one.
+This allows the parallelisation of tasks and seemingly increases performance.
 However, in a real-time system like the guest Salamander 4, this can lead to increased latencies due to contention between threads.
 In order to ensure more deterministic behavior in the guest, it is disabled on the host.
-	\item \textbf{Intel SpeedStep}: This technology dynamically adjusts the clock speed of the CPU based on workload.
-These dynamic changes can lead to unpredictable latencies in a real-time system.
+	\item \textbf{Intel SpeedStep}: This dynamically adjusts the clock speed of the CPU based on workload.
+These dynamic adjustments of the speed can lead to unpredictable latencies in a real-time system.
 It is also disabled on the host to maintain a constant CPU speed.
@@
 
+\noindent After these first real-time BIOS tunings, the \texttt{latency} test was run to see, whether this had any significant impact on the determinism of the system.
+\textcolor{red}{Figure~\ref{fig:gnuplot_max_latency_taskset} shows the latency values after tuning BIOS settings.}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
+	\caption[Latency taskset]{Latency taskset}
+	\label{fig:gnuplot_max_latency_taskset}
+\end{figure}
+
+\noindent The maximum latency value recorded was 457.545\textmu s.
+This is a step for the better compared to the initial version, but certainly still not acceptable within this work's definition of hard real-time determinism.
+
 \subsection{Kernel Configurations}\label{subsec:kernel_configurations}
@@
 
+\clearpage
 \noindent In the following, these settings along with their impact on system latency are briefly described.
@@
 It can enhance device performance by allowing these devices to directly access and use memory, which is especially helpful when these devices are virtualized.
+	\item \clearpage
 	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel’s RDT (Resource Director Technology).
@@
 It controls the frequency of the CPU by scaling it up when demand is high and scaling it down to save power when demand is low.
-This dynamic frequency scaling is disabled because it leads to increased latencies for real-time systems.
+This dynamic frequency scaling is disabled because it leads to increased latencies for real-time systems.\clearpage
 	\item \textbf{intel.max\_cstate=0, intel\_idle.max\_cstate=0, processor.max\_cstate=0, processor\_idle.max\_cstate=0}: These parameters disable deeper C-states (CPU power saving states).
@@
 \end{itemize}
-
-\textcolor{red}{After the BIOS and Kernel configurations, Figure~\ref{fig:max_latency_rt_kernelparam}}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/max_latency_rt_kernelparam/max_latency_rt_kernelparam.png}
-	\caption[After BIOS and Kernel configurations]{After BIOS and Kernel configurations}
-	\label{fig:max_latency_rt_kernelparam}
-\end{figure}
-
-\vspace{2em}
-\begin{table}[H]
-	\centering
-	\caption[Latency statistics of Salamander 4 after Kernel configurations]{Latency statistics of Salamander 4 virtualization after Kernel configurations in microseconds}
-	\label{tab:latency_statistics_kernel}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|c|c|}\hline
-			\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
-			Average Latency    & 14.00                      \\\hline
-			Maximum Latency    & 21.694                     \\\hline
-			Minimum Latency    & 6.351                      \\\hline
-			Standard Deviation & 1.44                       \\\hline
-		\end{tabular}}
-\end{table}
+\clearpage
+\textcolor{red}{\noindent After the BIOS and Kernel configurations, Figure~\ref{fig:gnuplot_max_latency_rt_kernelparam} shows }
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/gnuplot_max_latency_rt_kernelparam.png}
+	\caption[Latency rt kernelparam]{Latency rt kernelparam}
+	\label{fig:gnuplot_max_latency_rt_kernelparam}
+\end{figure}
+
+\vspace{2em}
 
@@
 
-\vspace{2em}
+\clearpage
+\vspace*{-5em}
 \begin{lstlisting}[name={Check distribution of interrupt requests across each CPU},label={script:smp_affinity}]
@@
 
-\vspace{2em}
+\clearpage
+\vspace*{-5em}
 \begin{lstlisting}[name={Change IRQ assignment of a CPU},label={script:change_smp_affinity}]
@@
 
-\vspace{2em}
 \begin{table}[H]
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/5rt_kernelparam_host/max_latency_rt_kernelparam_host/max_latency_rt_kernelparam_host.png}
-	\caption[After Host configurations]{After Host configurations}
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/5rt_kernelparam_host/gnuplot_max_latency_rt_kernelparam_host.png}
+	\caption[Latency rt kernelparam host]{Latency rt kernelparam host}
 	\label{fig:max_latency_rt_kernelparam_host}
 \end{figure}
-
-\vspace{2em}
-\begin{table}[H]
-	\centering
-	\caption[Latency statistics of Salamander 4 after Host configurations]{Latency statistics of Salamander 4 virtualization after Host configurations in microseconds}
-	\label{tab:latency_statistics_host}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|c|c|}\hline
-			\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
-			Average Latency    & 12.61                      \\\hline
-			Maximum Latency    & 17.041                     \\\hline
-			Minimum Latency    & 7.872                      \\\hline
-			Standard Deviation & 1.8                        \\\hline
-		\end{tabular}}
-\end{table}
 
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/6rt_kernelparam_host_qemu/max_latency_rt_kernelparam_host_qemu/max_latency_rt_kernelparam_host_qemu.png}
-	\caption[After QEMU configurations]{After QEMU configurations}
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/6rt_kernelparam_host_qemu/gnuplot_max_latency_rt_kernelparam_host_qemu.png}
+	\caption[Latency rt kernelparam host qemu]{Latency rt kernelparam host qemu}
 	\label{fig:max_latency_rt_kernelparam_host_qemu}
 \end{figure}
-
-\vspace{2em}
-\begin{table}[H]
-	\centering
-	\caption[Latency statistics of Salamander 4 after QEMU configurations]{Latency statistics of Salamander 4 virtualization after QEMU configurations in microseconds}
-	\label{tab:latency_statistics_qemu}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|c|c|}\hline
-			\textbf{Statistic} & \textbf{Value (\textmu s)} \\\hline
-			Average Latency    & 16.26                      \\\hline
-			Maximum Latency    & 17.134                     \\\hline
-			Minimum Latency    & 14.532                     \\\hline
-			Standard Deviation & 0.3                        \\\hline
-		\end{tabular}}
-\end{table}
 
@@
 \chapter{Results}\label{cha:results}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/012456combined/gnuplot_combined_max_latency.png}
+	\caption[Latency rt kernelparam host]{Latency rt kernelparam host}
+	\label{fig:max_latency_rt_kernelparam_host}
+\end{figure}
 
@@
 \chapter{Summary and Outlook}\label{cha:summary_and_outlook}
-\section{Trace-cmd \& Kernelshark}
 

### 2024-08-22 Masterarbeit_Robotik_Pamuk.tex
3.2 Real-Time Performance Tuning, 3.2.1 BIOS Configurations, 3.2.2 Kernel Configurations, 3.2.3 Host OS Configurations, 3.2.3.1 CPU affinity and Isolation

@@
 The complete command used was~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of minimum, average and maximum latencies in a Gnuplot-compatible format to the file~\texttt{gnuplot.txt}.
-Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time.
+Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the distribution of latency over the course of said time.
 
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
-	\caption[Variation in latency of Salamander 4 bare metal]{Variation in latency of Salamander 4 bare metal}
+	\caption[Distribution of Latency of Salamander 4 Bare Metal]{Distribution of Latency of Salamander 4 Bare Metal}
 	\label{fig:gnuplot_max_latency_hardware}
@@
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
-Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency.
+Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, Figure~\ref{fig:gnuplot_max_latency_default} depicts the distribution of latency.
 
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
-	\caption[Variation in latency with default settings]{Variation in latency with default settings}
+	\caption[Distribution of Latency of Salamander 4 unmodified Virtualization]{Distribution of Latency of Salamander 4 unmodified Virtualization}
 	\label{fig:gnuplot_max_latency_default}
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/01combined/gnuplot_combined_max_latency.png}
-	\caption[Initial Comparison of Variation in Latency between Hardware and Virtualization]{Initial Comparison of Variation in Latency between Hardware and Virtualization}
+	\caption[Initial Comparison of Latency Distribution between Hardware and Virtualization]{Initial Comparison of Latency Distribution between Hardware and Virtualization}
 	\label{fig:gnuplot_max_latency_combined}
@@
 The individual configurations are discussed in detail and the modifications are justified with a clear explanation.
-The goal is to bring the latency of the virtualized system closer to that of the bare metal system, thereby ensuring deterministic behavior under real-time constraints.
-
-\bigskip \noindent It is important to mention that a real-time system implies determinism at every component of the system ranging from hardware over the kernel to the software and the application on it.
+The goal is to bring the latency of the virtualization closer to that of the bare metal, thereby ensuring deterministic behavior under real-time constraints.
+
+\bigskip \noindent It is important to mention that a real-time system implies determinism at every component of the system, ranging from hardware over the kernel to the software and the application on it.
 As already established earlier, the guest OS Salamamder 4 runs Xenomai and is hence real-time capable.
@@
 The most important metric of a latency test is without a doubt the worst latency value, because the system is only as reliable as its worst measurement.
-Since the period of the tests is set to 100\textmu s, there are approximately 100000 samples per second.
+Since the period of the tests is set to 100\textmu s, there are approximately 1,000,000 samples per second.
 
@@
 Table~\ref{tab:bios_configuration} illustrates the specific BIOS settings that have been adjusted for the purpose of real-time performance.
-
-\vspace{2em}
+As an important note, this is not a definitive, full-ended list, and other devices may have additional settings that can be modified to achieve even better latency.
+
 \begin{table}[H]
@@
 
-\noindent After these first real-time BIOS tunings, the \texttt{latency} test was run to see, whether this had any significant impact on the determinism of the system.
-\textcolor{red}{Figure~\ref{fig:gnuplot_max_latency_taskset} shows the latency values after tuning BIOS settings.}
+\noindent After these first real-time tunings in the BIOS settings, the \texttt{latency} test was run to see, whether this had any significant impact on the determinism of the system.
+Figure~\ref{fig:gnuplot_max_latency_taskset} shows the latency values after tuning BIOS settings.
 
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
-	\caption[Latency taskset]{Latency taskset}
+	\caption[Latency Distribution after tuning BIOS Settings]{Latency Distribution after tuning BIOS Settings}
 	\label{fig:gnuplot_max_latency_taskset}
@@
 \noindent The maximum latency value recorded was 457.545\textmu s.
-This is a step for the better compared to the initial version, but certainly still not acceptable within this work's definition of hard real-time determinism.
+When compared to the inital worst latency of 707.62 \textmu s, this is a step for the better, but certainly still not acceptable within this work's definition of hard real-time determinism.
 
 \subsection{Kernel Configurations}\label{subsec:kernel_configurations}
-
-The kernel command-line parameters are shown in Code~\ref{script:kernel_cli} below.
+The kernel command-line parameters are shown in Code~\ref{script:kernel_cli} below. To access and modify them, the file is located in \texttt{/etc/default/grub}.
 
@@
 	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit), which connects a DMA (Direct Memory Access)-capable I/O bus, such as graphics cards and network adapters, to the main memory.
-It can enhance device performance by allowing these devices to directly access and use memory, which is especially helpful when these devices are virtualized.
-	\item \clearpage
+It allows these devices to directly access and use memory, which is especially helpful when they are virtualized.
+	\clearpage
 	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel’s RDT (Resource Director Technology).
 Unlike L1 and L2 caches, where each core has its fixed capacity, L3 cache is a shared pool among multiple cores.
-L3 CAT is a mechanism that controls the amount of L3 cache that a process can use.
+L3 CAT controls the amount of L3 cache that a process can use.
 By controlling cache allocation, it can prevent a single process from monopolizing the L3 cache, which is particularly beneficial in virtualized environments, where multiple virtual machines share the same physical host.
@@
 By setting the gap to 0, the virtual machine exits the pause loop immediately, which can reduce latency in spinlock-intensive workloads.
-	\item \textbf{irqaffinity=0}: Sets the default Interrupt Request affinity to none.
+	\item \textbf{irqaffinity=0}: Sets the default interrupt request affinity to none.
 This means that no CPU core is preferred over another for handling IRQs.
 Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs.
-Subsection~\ref{subsubsec:irq_handling} dives deeper into Interrupt Request affinity.
+Subsubsection~\ref{subsubsec:irq_handling} dives deeper into interrupt request affinity.
 \end{itemize}
 \clearpage
-\textcolor{red}{\noindent After the BIOS and Kernel configurations, Figure~\ref{fig:gnuplot_max_latency_rt_kernelparam} shows }
+\noindent After the kernel modifications on top of the BIOS settings from the last subsection, Figure~\ref{fig:gnuplot_max_latency_rt_kernelparam} demonstrates the latency
+values gathered in 10 minutes.
 
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/gnuplot_max_latency_rt_kernelparam.png}
-	\caption[Latency rt kernelparam]{Latency rt kernelparam}
+	\caption[Latency Distribution after tuning Kernel Settings]{Latency Distribution after tuning Kernel Settings}
 	\label{fig:gnuplot_max_latency_rt_kernelparam}
@@
 
-\vspace{2em}
+\noindent Apparent is a significant improvement in the latency of the virtualized Salamander 4 OS. With a worst latency value of 21.694\textmu s, the real-time tunings appear to be working as desired.
 
@@
 The host OS needs to provide an environment where the guest OS can operate in real-time.
-This entails a number of host-side adjustments to reduce interruptions and reduce latency.
-In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest OS is provided.
+This entails a number of host-side adjustments to reduce interruptions and latency.
+In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest Salamander 4 OS is provided.
 \subsubsection{CPU affinity and Isolation}\label{subsubsec:cpu_isolation}
 
-Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-For CPU isolation, the~\texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system.
+Isolating CPUs means removing all user-space threads and unbound kernel threads, since bound kernel threads are tied to specific CPUs and hence cannot be moved.
+For CPU isolation, the~\texttt{isolcpus} function is used to isolate a performance CPU from the general scheduling algorithms of the operating system.
 This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks.
@@
 Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}, including systemd services.
-To prevent systemd services from running on an isolated CPU, the CPU affinity can be set the with line \texttt{CPUAffinity=0 1 2 3 5 6 7 8 9 10 11 12 13} in the \texttt{/etc/systemd/system.conf} file to indicate the CPUs that systemd-services are allowed to run on.
+To prevent systemd services from running on an isolated CPU, the CPU affinity can be set with the line \texttt{CPUAffinity=0 1 2 3 5} in the \texttt{/etc/systemd/system.conf} file to indicate the CPUs that systemd services are allowed to run on.
 Every CPU other than the isolated CPU 4 is allowed.
 Code~\ref{output:pse} shows the user and kernel tasks that run on CPU 4.
+An isolated CPU can be bound to a task by using \texttt{taskset -c <CPU>}, as shown later in Subsection~\ref{subsec:guest_configurations}.
 After the isolation, user tasks other than the QEMU process have been removed from running on this CPU.
@@
 
-\subsubsection{KVM Entry and KVM Exit}\label{subsubsec:kvm_entry_exit}
-\textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
-However, the isolcpus function only isolates CPUs from the scheduler, meaning that user-space processes won't be scheduled on these CPUs, but kernel tasks and interrupts can still utilize the CPU.This led to the investigation of the causes for the observed high and inconsistent latency.
-The guest operates within the~\texttt{kvm\_entry} and~\texttt{kvm\_exit} events of the host.
-Kernelshark revealed a high frequency of~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
-This frequent switching hinders the guest's ability to run continuously, thereby increasing the virtualization latency.}
-
-\textcolor{red}{To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
-Specifically, the causes for~\texttt{kvm\_exit} events were analyzed.
-The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
-The results in Figure~\ref{fig:kvm_exit_taskset} were obtained.
-Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed~\texttt{kvm\_exit} events.}
-
-\vspace{2em}
-\begin{table}[H]
-	\centering
-	\begin{tabular}{|l|p{0.62\textwidth}|}
-		\hline
-		\textbf{Exit Reason} & \textbf{Description}                                     \\ \hline
-		APIC\_WRITE          & Triggered when the guest writes to its APIC.             \\ \hline
-		EXTERNAL\_INTERRUPT  & Triggered by external hardware interrupts.               \\ \hline
-		HLT                  & Triggered when the guest executes the HLT instruction.   \\ \hline
-		EPT\_MISCONFIG       & Triggered by a misconfiguration in the EPT.              \\ \hline
-		PREEMPTION\_TIMER    & Triggered when the host's preemption timer expires.      \\ \hline
-		PAUSE\_INSTRUCTION   & Triggered when the PAUSE instruction is executed.        \\ \hline
-		EPT\_VIOLATION       & Triggered by a violation of the EPT permission settings. \\ \hline
-		IO\_INSTRUCTION      & Triggered when the guest executes an I/O instruction.    \\ \hline
-		EOI\_INDUCED         & Triggered when an EOI signal is sent to the APIC.        \\ \hline
-		MSR\_READ            & Triggered when the guest reads from a MSR.               \\ \hline
-		CPUID                & Triggered when the guest executes the CPUID instruction. \\ \hline
-	\end{tabular}
-	\caption{Description of kvm\_exit reasons}
-	\label{tab:kvm_exit}
-\end{table}
-
-\textcolor{red}{\noindent In the process of analyzing the~\texttt{kvm\_exit} events, several reasons for these exits were identified.
-The most frequent among these were the~\texttt{APIC\_WRITE} and~\texttt{HLT} events.
-The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
-The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
-Other significant but less frequent events included~\texttt{EXTERNAL\_INTERRUPT} and~\texttt{IO\_INSTRUCTION}.
-These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
-Events such as~\texttt{EPT\_MISCONFIG} and~\texttt{PREEMPTION\_TIMER} were also noted.
-These could potentially signal issues with memory management and the host's scheduling of the guest.
-While events like~\texttt{PAUSE\_INSTRUCTION},~\texttt{EPT\_VIOLATION},~\texttt{EOI\_INDUCED},~\texttt{MSR\_READ}, and~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.}
-
-\textcolor{red}{\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is not bound to a specific CPU core.
-This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
-As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on the user level.
-Without CPU isolation, context switches take place at the operating system level and not at the hypervisor level.
-This explains why there are fewer~\texttt{kvm\_exit} events.
-However, this will also, as previously shown, lead to higher latency, as context switches at the operating system level generally take longer than a~\texttt{kvm\_exit} and ~\texttt{kvm\_entry}.
-When the CPU was dedicated to the QEMU process, on the other hand, there was a significant increase in~\texttt{kvm\_exit} events.
-This is because every context switch takes place at the hypervisor level.
-Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.}
-
 \subsubsection{Interrupt Affinity}\label{subsubsec:irq_handling}
@@
 This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems.
-The~\texttt{/proc/interrupts} file can be monitored using the command~\texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
+All the existing interrupts can be monitored using the command~\texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
 The IRQs needed to be removed from the isolated CPU by manipulating the~\texttt{/proc/irq/<IRQ>/smp\_affinity} files.
 The value in the~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU.
-The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU13).
-When there are 14 CPUs available, the default value for~\texttt{smp\_affinity} would be 3FFF.
-Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111).
-The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU.
-By changing the values of the~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled.
-
+The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU5).
+When there are 6 CPUs available, the default value for~\texttt{smp\_affinity} would be 3F (11 1111).
+Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 2F (10 1111).
+The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU, since the \texttt{/proc/interrupts} file does not give away this information and looking individually in the \texttt{/proc/irq/<IRQ>/smp\_affinity} files for every IRQ would be time-consuming.
+The program first reads the CPU affinity settings for each IRQ from the \texttt{/proc/irq} directory and stores the CPUs assigned to each IRQ in a dictionary.
+It then creates a data frame to represent this information in a table format, marking the assigned CPUs with ‘x’.
+Finally, it prints the table and saves it as a markdown file.
+Figure~\ref{fig:interrupt_affinity_table} shows the output of the program.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/salamander4/latency_reduction/IRQaffinity/table_CPU_IRQ.png}
+	\caption[Interrupt Affinity Table]{Interrupt Affinity Table}
+	\label{fig:interrupt_affinity_table}
+\end{figure}
 \clearpage
@@
 
-\noindent However, as the proc filesystem resets to its default state after each reboot, manually changing numerous IRQ files is tedious and time-consuming.
-This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot.
+\noindent However, as the \texttt{proc} filesystem resets to its default state after each reboot, manually changing numerous IRQ files through the said bitbasks is tedious and time-consuming.
+This process can be was automated using the shell script in Code~\ref{script:change_smp_affinity}, which can automatically be executed after every reboot.
+
 Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads.
@@
 \chapter{To Include}\label{cha:to_include}
+
+\section{KVM Entry and KVM Exit}\label{subsubsec:kvm_entry_exit}
+\textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
+However, the isolcpus function only isolates CPUs from the scheduler, meaning that user-space processes won't be scheduled on these CPUs, but kernel tasks and interrupts can still utilize the CPU.This led to the investigation of the causes for the observed high and inconsistent latency.
+The guest operates within the~\texttt{kvm\_entry} and~\texttt{kvm\_exit} events of the host.
+Kernelshark revealed a high frequency of~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
+This frequent switching hinders the guest's ability to run continuously, thereby increasing the virtualization latency.}
+
+\textcolor{red}{To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
+Specifically, the causes for~\texttt{kvm\_exit} events were analyzed.
+The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
+The results in Figure~\ref{fig:kvm_exit_taskset} were obtained.
+Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed~\texttt{kvm\_exit} events.}
+
+\vspace{2em}
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|l|p{0.62\textwidth}|}
+		\hline
+		\textbf{Exit Reason} & \textbf{Description}                                     \\ \hline
+		APIC\_WRITE          & Triggered when the guest writes to its APIC.             \\ \hline
+		EXTERNAL\_INTERRUPT  & Triggered by external hardware interrupts.               \\ \hline
+		HLT                  & Triggered when the guest executes the HLT instruction.   \\ \hline
+		EPT\_MISCONFIG       & Triggered by a misconfiguration in the EPT.              \\ \hline
+		PREEMPTION\_TIMER    & Triggered when the host's preemption timer expires.      \\ \hline
+		PAUSE\_INSTRUCTION   & Triggered when the PAUSE instruction is executed.        \\ \hline
+		EPT\_VIOLATION       & Triggered by a violation of the EPT permission settings. \\ \hline
+		IO\_INSTRUCTION      & Triggered when the guest executes an I/O instruction.    \\ \hline
+		EOI\_INDUCED         & Triggered when an EOI signal is sent to the APIC.        \\ \hline
+		MSR\_READ            & Triggered when the guest reads from a MSR.               \\ \hline
+		CPUID                & Triggered when the guest executes the CPUID instruction. \\ \hline
+	\end{tabular}
+	\caption{Description of kvm\_exit reasons}
+	\label{tab:kvm_exit}
+\end{table}
+
+\textcolor{red}{\noindent In the process of analyzing the~\texttt{kvm\_exit} events, several reasons for these exits were identified.
+The most frequent among these were the~\texttt{APIC\_WRITE} and~\texttt{HLT} events.
+The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
+The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
+Other significant but less frequent events included~\texttt{EXTERNAL\_INTERRUPT} and~\texttt{IO\_INSTRUCTION}.
+These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
+Events such as~\texttt{EPT\_MISCONFIG} and~\texttt{PREEMPTION\_TIMER} were also noted.
+These could potentially signal issues with memory management and the host's scheduling of the guest.
+While events like~\texttt{PAUSE\_INSTRUCTION},~\texttt{EPT\_VIOLATION},~\texttt{EOI\_INDUCED},~\texttt{MSR\_READ}, and~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.}
+
+\textcolor{red}{\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is not bound to a specific CPU core.
+This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
+As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on the user level.
+Without CPU isolation, context switches take place at the operating system level and not at the hypervisor level.
+This explains why there are fewer~\texttt{kvm\_exit} events.
+However, this will also, as previously shown, lead to higher latency, as context switches at the operating system level generally take longer than a~\texttt{kvm\_exit} and ~\texttt{kvm\_entry}.
+When the CPU was dedicated to the QEMU process, on the other hand, there was a significant increase in~\texttt{kvm\_exit} events.
+This is because every context switch takes place at the hypervisor level.
+Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.}
 

### 2024-08-22 Masterarbeit_Robotik_Pamuk.tex
3.2.3.2 Interrupt Affinity, 3.2.3.3 RT-Priority, 3.2.3.4 Disable RT Throttling, 3.2.3.5 Disable Timer Migration, 3.2.3.6 Set Device Driver Work Queue, 3.2.3.7 Disable RCU CPU Stall Warnings, 3.2.3.8 Stop Certain Services, 3.2.3.9 Disable Machine Check, 3.2.3.10 Boot into text-based environment, 3.2.4 QEMU/KVM Configurations, 3.2.4.1 Tune LAPIC Timer Advance, 3.2.4.2 Set QEMU Options for real-time VM, 3.2.5 Guest OS Configurations, 3.2.6 Other Configurations

@@
 \textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
-min & 599 & 0.711 & 0.454 \\ \hline
-avg & 5 999 988 & 1.019 & 0.150 \\ \hline
-max & 599 & 3.528 & 0.895 \\ \hline
+\textbf{min} & 599 & 0.711 & 0.454 \\ \hline
+\textbf{avg} & 5 999 988 & 1.019 & 0.150 \\ \hline
+\textbf{max} & 599 & 3.528 & 0.895 \\ \hline
 \end{tabular}
@@
 \textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
-min & 599 & 3.713 & 1.355 \\ \hline
-avg & 5,999,922 & 8.247 & 2.521 \\ \hline
-max & 599 & 45.705 & 52.196 \\ \hline
+\textbf{min} & 599 & 3.713 & 1.355 \\ \hline
+\textbf{avg} & 5,999,922 & 8.247 & 2.521 \\ \hline
+\textbf{max} & 599 & 45.705 & 52.196 \\ \hline
 \end{tabular}
@@
 
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt} and Table~\ref{tab:latency_overrun_msw_new_virt}.
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|}
+	\hline
+	\textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
+	\textbf{min} & 599 & 1.419 & 0.507 \\ \hline
+	\textbf{avg} & 5,999,885 & 3.398 & 1.251 \\ \hline
+	\textbf{max} & 599 & 74.015 & 26.590 \\ \hline
+	\end{tabular}
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_virt_bios}
+	\end{table}
+
+	\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|c|}
+	\hline
+	\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches}  \\ \hline
+	0.969 & 3.948 & 457.545 & 94 & 0 \\ \hline
+	\end{tabular}
+	\caption{Latency, Overrun, Missed Switch, Best, and Worst Data}
+	\label{tab:latency_overrun_msw_new_virt_bios}
+	\end{table}
+
 \noindent The maximum latency value recorded was 457.545\textmu s.
@@
 
-\clearpage
 \noindent In the following, these settings along with their impact on system latency are briefly described.
@@
 It allows these devices to directly access and use memory, which is especially helpful when they are virtualized.
-	\clearpage
 	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel’s RDT (Resource Director Technology).
@@
 It controls the frequency of the CPU by scaling it up when demand is high and scaling it down to save power when demand is low.
-This dynamic frequency scaling is disabled because it leads to increased latencies for real-time systems.\clearpage
+This dynamic frequency scaling is disabled because it leads to increased latencies for real-time systems.
 	\item \textbf{intel.max\_cstate=0, intel\_idle.max\_cstate=0, processor.max\_cstate=0, processor\_idle.max\_cstate=0}: These parameters disable deeper C-states (CPU power saving states).
@@
 
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_kernelparam} and Table~\ref{tab:latency_overrun_msw_new_virt_kernel}.
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|}
+	\hline
+	\textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
+	\textbf{min} & 599 & 3.356 & 0.551 \\ \hline
+	\textbf{avg} & 5,999,972 & 4.036 & 0.290 \\ \hline
+	\textbf{max} & 599 & 13.484 & 1.454 \\ \hline
+	\end{tabular}
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_kernelparam}
+	\end{table}
+
+	\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|c|}
+	\hline
+	\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
+	2.545 & 4.811 & 21.694 & 0 & 0 \\ \hline
+	\end{tabular}
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_new_virt_kernel}
+	\end{table}
+
 \noindent Apparent is a significant improvement in the latency of the virtualized Salamander 4 OS. With a worst latency value of 21.694\textmu s, the real-time tunings appear to be working as desired.
 
+\clearpage
 \subsection{Host OS Configurations}\label{subsec:host_configurations}
@@
 Finally, it prints the table and saves it as a markdown file.
-Figure~\ref{fig:interrupt_affinity_table} shows the output of the program.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/salamander4/latency_reduction/IRQaffinity/table_CPU_IRQ.png}
-	\caption[Interrupt Affinity Table]{Interrupt Affinity Table}
-	\label{fig:interrupt_affinity_table}
-\end{figure}
+
+\bigskip \noindent However, as the \texttt{proc} filesystem resets to its default state after each reboot, manually changing numerous IRQ files through the said bitbasks is tedious and time-consuming.
+This process can be was automated using the shell script in Code~\ref{script:change_smp_affinity}, which can automatically be executed after every reboot.
+
+\bigskip \noindent Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads.
+This can be done through finding out the respective process IDs through \texttt{ps -e | grep irq/.*nvme} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.
+
 \clearpage
@@
 
-\noindent However, as the \texttt{proc} filesystem resets to its default state after each reboot, manually changing numerous IRQ files through the said bitbasks is tedious and time-consuming.
-This process can be was automated using the shell script in Code~\ref{script:change_smp_affinity}, which can automatically be executed after every reboot.
-
-Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads.
-This can be done through finding out the respective process IDs through \texttt{ps -e | grep irq/.*nvme} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.
-
 \clearpage
@@
 \subsubsection{RT-Priority}
-Having a real-time kernel itself is a crucial part of achieving deterministic behavior, but not enough to take full advantage of the real-time capabilities.
+Having a real-time kernel itself is an indispensable step of achieving deterministic behavior, but not enough by itself to take full advantage of the real-time capabilities.
 One key aspect of this are real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}.
@@
 This reduces the potential for missed deadlines.
+\clearpage
 \subsubsection{Disable Timer Migration}
@@
 By setting the work queue away from CPU 4 to another CPU, it is free to handle real-time tasks without being interrupted by these work queue tasks.
-This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
+This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask} In this case, that is \texttt{2F}.
 \subsubsection{Disable RCU CPU Stall Warnings}
@@
 This setting is also not persistent across reboots, so the command needs to be added to a startup script.
+\clearpage
 \subsubsection{Stop Certain Services}
-Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{tab:stop_servies} can be further sources for random latency and unnecessary overhead.
-Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means they will not be able to interrupt the CPU with their tasks.
+Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{tab:stop_servies}, can be further sources for random latency and unnecessary overhead.
+Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means, they will not be able to interrupt the CPU with their tasks.
 
@@
 Machine checks report hardware errors and these checks can cause interruptions and increase latency.
-Hence it is best to disable them in real-time scenarios by writing a ``0'' to the \texttt{/sys/devices/system/machinecheck/machinecheck0/check\_interval} file.
+Hence, it is best to disable them in real-time scenarios by writing a ``0'' to the \texttt{/sys/devices/system/machinecheck/machinecheck0/check\_interval} file.
 
@@
 For the sake of completeness, the graphical interface can be brought back through the command \texttt{systemctl set-default graphical.target} and a following reboot.
-
-\textcolor{red}{After the Host configurations, Figure~\ref{fig:max_latency_rt_kernelparam_host}}
+\clearpage
+\noindent After these comprehensive host configurations, the result of the \texttt{latency} test is depicted in Figure~\ref{fig:max_latency_rt_kernelparam_host}.
 
@@
 
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt_host} and Table~\ref{tab:latency_overrun_msw_new_virt_host}.
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|}
+	\hline
+	\textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
+	\textbf{min} & 599 & 2.998 & 0.242 \\ \hline
+	\textbf{avg} & 5,999,972 & 4.043 & 0.255 \\ \hline
+	\textbf{max} & 599 & 12.124 & 1.828 \\ \hline
+	\end{tabular}
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_virt_host}
+	\end{table}
+
+	\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|c|}
+	\hline
+	\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
+	2.591 & 4.834 & 17.041 & 0 & 0 \\ \hline
+	\end{tabular}
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_new_virt_host}
+	\end{table}
+
+\noindent A worst latency value of 17.041\textmu s shows that the virtualized Salamander 4 OS got even more reliable with the host configurations.
+
+\clearpage
 \subsection{QEMU/KVM Configurations}\label{subsec:guest_configurations}
@@
 \noindent Code~\ref{script:qemu_def_tuned} shows the final QEMU script used to start the Salamander 4 virtualization, including these options.
-
+\clearpage
 \vspace{2em}
@@
 
-		exec taskset -c 4 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
+		exec taskset -c 4 chrt -f 99 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
 		-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
@@
 	\end{lstlisting}
-
-\textcolor{red}{After the QEMU configurations, Figure~\ref{fig:max_latency_rt_kernelparam_host_qemu}}
+\vspace{2em}
+	\noindent After QEMU configurations, the result of the \texttt{latency} test is depicted in Figure~\ref{fig:max_latency_rt_kernelparam_host_qemu}.
 
@@
 
-\subsection{Guest OS Configurations}
+\clearpage
+
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt_qemu} and Table~\ref{tab:latency_overrun_msw_new_virt_qemu}.
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|}
+	\hline
+	\textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
+	\textbf{min} & 599 & \textbf{3.125} & 0.432 \\ \hline
+	\textbf{avg} & 5,999,973 & \textbf{4.018} & 0.291 \\ \hline
+	\textbf{max} & 599 & \textbf{15.883} & 0.351 \\ \hline
+	\end{tabular}
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_virt_qemu}
+	\end{table}
+
+	\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|c|c|}
+	\hline
+	\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
+	2.614 & 4.779 & 17.134 & 0 & 0 \\ \hline
+	\end{tabular}
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_new_virt_qemu}
+	\end{table}
+
+	\noindent While the worst latency value did not change much, the standard deviation of the maximum latency samples of each second, with the test lasting 600 seconds and thus having 600 worst latencies, decreased.
+This means, the system is more stable and predictable, thanks to the real-time tunings that have been performed till this point.
+
+\subsection{Guest OS Configurations}\label{subsec:guest_configurations}
+The guest OS, Salamander 4, is already based on Xenomai and uses the Cobalt real-time core with the Dovetail extension.
+Therefore, no additional modifications were necessary for the guest OS itself.
+The focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
 
@@
 There are other configurations that may be relevant depending on the case.
-Linux Kernel Developer Steven Rostedt explains all relevant aspects of a real-time system that must be considered in~\cite{kernelrecipesKernelRecipes20162016} and gives insight for finding sources of latency on the linux system in~\cite{thelinuxfoundationFindingSourcesLatency2020}.
-A Checklist for Writing Linux Real-Time Applications is provided by John Ogness in~\cite{thelinuxfoundationChecklistWritingLinux2020}.
+Linux Kernel Developer Steven Rostedt explains all relevant aspects of a real-time system that must be considered in~\cite{kernelrecipesKernelRecipes20162016} and gives insight for finding sources of latency on the Linux system in~\cite{thelinuxfoundationFindingSourcesLatency2020}.
+A Checklist for writing Linux real-time applications is provided by John Ogness in~\cite{thelinuxfoundationChecklistWritingLinux2020}.
 Every layer of the system stack must be deterministic to ensure predictable and reliable latency, including hardware, operating system, middleware and drivers, and the application software.
-~\cite{HOWTOBuildRTapplication} and~\cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real time preemption patch in great detail.
-Various hardware and software tunings are mentioned in~\cite{KVMQemuVirtualization} and~\cite{RealTimePerformanceTuning2022}.
+~\cite{HOWTOBuildRTapplication} and~\cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real-time preemption patch in great detail.
+Various hardware and software tunings are listed in~\cite{KVMQemuVirtualization} and~\cite{RealTimePerformanceTuning2022}.
 

### 2024-08-23 Masterarbeit_Robotik_Pamuk.tex
3.3 Real-Time Robotic Application, 3.3.1 Setup of Hardware Salamander 4, 3.3.2 Setup of QEMU Salamander 4

@@
 	\end{tikzpicture}
-	\caption{Flowchart of operations in a virtualized environment (own figure, inspired by~\citeauthor{maPerformanceTuningKVMbased})}
+	\caption[Flowchart of operations in a virtualized environment]{Flowchart of operations in a virtualized environment (own figure, inspired by~\citeauthor{maPerformanceTuningKVMbased})}
 	\label{fig:kvm_stages}
@@
 Once both files are loaded, the trace data can be visually inspected.
-All the current CPUs for both the host and the guest can be turned off, and the KVM Combo plots menu can be selected to view the guest vCPUs plotted on top of the host threads that run them, as depicted in Figure~\ref{fig:kernelshark_combo}.
+All the current CPUs for both the host and the guest can be turned off, and the KVM Combo plots menu can be selected to view the guest VCPUs plotted on top of the host threads that run them, as depicted in Figure~\ref{fig:kernelshark_combo}.
 
@@
 	\includegraphics[width=1.0\columnwidth]{img/kernelshark_combo.png}
-	\caption[Guest vCPUs plotted on top of host threads in KernelShark]{Guest vCPUs plotted on top of host threads in KernelShark}
+	\caption[Guest VCPUs plotted on top of host threads in KernelShark]{Guest VCPUs plotted on top of host threads in KernelShark}
 	\label{fig:kernelshark_combo}
@@
 \section{Real-Time Robotic Application}\label{sec:robotic_application}
-This chapter compares the latency of the Salamander 4 operating system before and after the real-time performance tunings with the latency of Salamander 4 running on bare metal hardware.
+This chapter compares the latency of the Salamander 4 operating system before and after the real-time performance tunings with the latency of Salamander 4 running on bare metal.
 This comparison is done with the aid of a program that will be explained shortly.
@@
 	\centering
-	\includegraphics[width=0.45\columnwidth]{img/experiment/mini_robot.jpg}
-	\caption[Mini-robot of the experiment]{Mini-robot of the experiment}
+	\includegraphics[width=0.3\columnwidth]{img/experiment/mini_robot.jpg}
+	\caption[Mini-robot of the experiment]{Mini-robot of the experiment~\cite{6DFRoboticArm}}
 	\label{fig:mini_robot}
@@
 
-\noindent The drive system of the robot arm consists of six MG996R digital servo motors~\cite{MG996RServoMotor}, equipped with a metal gearbox.
+\noindent The robot arm consists of six MG996R digital servo motors~\cite{MG996RServoMotor}, equipped with a metal gearbox.
 The motor is able to rotate in a range of approximately 180 degrees and its position can be controlled with a high degree of accuracy.
-Each servo motor has three wires that need to be connected as shown in Figure~\ref{fig:motor_wires}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.45\columnwidth]{img/experiment/motor_wires.png}
-	\caption[MG996R Servo Motor]{MG996R Servo Motor~\cite{MG996RServoMotor}}
-	\label{fig:motor_wires}
+Each servo motor has three wires that need to be connected as shown in Figure~\ref{fig:servo}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{img/experiment/servo.jpg}
+	\caption[MG996R Servo Motor]{MG996R Servo Motor~\cite{PackMg996rMetal}}
+	\label{fig:servo}
 \end{figure}
@@
 \noindent To drive the motor, it has to be powered using the red and brown wires and can be controlled by sending PWM signals to the orange wire.
-In this experiment, this PWM signal is generated by the proprietary PW 022 pulse width module of Sigmatek~\cite{DigitalOutputSIGMATEK}.
+A 1,500 microsecond pulse every 20 milliseconds centers the servo.
+A 1,000 microsecond pulse turns it 90 degrees left, and a 2,000 microsecond pulse turns it 90 degrees right.
+This is depicted in Figure~\ref{fig:servo_angle}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{img/experiment/servo_angle.png}
+	\caption[Controlling the MG996R Servo Motor]{Controlling the MG996R Servo Motor~\cite{twierengEmilyQuadrupedDog}}
+	\label{fig:servo_angle}
+\end{figure}
+
+\noindent In this experiment, this PWM signal is generated by the proprietary PW 022 pulse width module of Sigmatek~\cite{DigitalOutputSIGMATEK}.
 Its connector layout is illustrated in Figure~\ref{fig:pw022_connectors}.
@@
 \noindent The module has two +24 V switching PWM outputs with an adjustable frequency for controlling inductive loads.
-Since the mentioned servo motors operate between 4.8 volts and 7.2 volts~\cite{MG996RServoMotor}, this voltage needed to be reduced through resistors before supplying it to a servo motor.
-For this purpose, two resistors with resistances of 2500 kiloohms and 1000 kiloohms were connected in series.
+Since the mentioned servo motors operate between 4.8 volts and 7.2 volts~\cite{MG996RServoMotor}, this voltage needs to be reduced through resistors before supplying it to a servo motor.
+For this purpose, two resistors with resistances of 2500 kiloohms and 1000 kiloohms are connected in series.
 The voltage across each resistor is calculated using Ohm's law, which is given by equation~\ref{eq:ohms_law} below.
@@
 \end{equation}
-In this equation, \(V\) represents the voltage across the resistor, \(I\) is the current flowing through the resistor, and \(R\) is the resistance of the resistor.
+
+\noindent In this equation, \(V\) represents the voltage across the resistor, \(I\) is the current flowing through the resistor, and \(R\) is the resistance of the resistor.
 The resulting voltages across the resistors are as follows:
+
 \begin{itemize}
 	\item The voltage across the 2.5 kiloohm resistor is approximately 17.14 volts.
-	\item The voltage across the 1 kiloohm resistor is approximately 6.86 volts. This voltage was then supplied to the control wire of servo motor 1 of the mini-robot.
+	\item The voltage across the 1 kiloohm resistor is approximately 6.86 volts. This voltage is then supplied to the control wire of servo motor 1 of the mini-robot.
 \end{itemize}
@@
 
-\noindent The program was written in Lasal Class 2 and was applied to all three mentioned versions of Salamander 4 to measure the reaction time of the robot to the specific commands.
+\noindent The program is written in Lasal Class 2 and is applied to all three mentioned versions of Salamander 4 to measure the reaction time of the robot to the commands.
 Prior to examining the software program, the next two subsections briefly explain the setup of each version of the experiment.
@@
 In this version, Salamander 4 runs on the CP 841~\cite{CPUEinheitenSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
-The PW 022 module is directly mounted on the CPU via the S-DIAS bus and communicates over the hard real-time capable Ethernet VARAN with 100 Mbit/s~\cite{SDIASSIGMATEK}.
+The PW 022 modules are directly mounted on the CPU via the S-DIAS bus~\cite{SDIASSIGMATEK}.
+There are three PW 022 modules because each can control two motors simultaneously, and the mini-robot has six motors.
 The setup is visible on Figure~\ref{fig:hardware_tree}.
@@
 	\centering
-	\includegraphics[width=0.5\columnwidth]{img/experiment/hardware_tree.png}
-	\caption[Hardware tree]{Hardware tree}
+	\includegraphics[width=0.45\columnwidth]{img/experiment/hardware_tree.png}
+	\caption[Setup of Hardware Salamander 4]{Setup of Hardware Salamander 4}
 	\label{fig:hardware_tree}
@@
 \subsection{Setup of QEMU Salamander 4}\label{subsec:setup_experiment_virtualized}
-In the virtualized setup of the experiment, the VCPU functionality of QEMU is used to connect QEMU with the PWM module.
-In order to achieve that, the PCV 522 VARAN Manager PCI Insert Card~\cite{ControlsHMIsSIGMATEK}, which serves as a bridge between the PC and the rest of the setup, needed to be plugged into the PC.
+In the virtualized setup of the experiment, the VCPU functionality of QEMU is used to connect QEMU with the PWM modules.
+In order to achieve that, the PCV 522 VARAN Manager PCI Insert Card~\cite{ControlsHMIsSIGMATEK}, which serves as a bridge between the PC and the rest of the setup, needs to be plugged into the PC.
 The command \texttt{lspci -nn} lists all PCI devices along with their vendor and device IDs.
-In this case, the command for binding the PCV 522 module to the VFIO-PCI driver was \texttt{sudo sh -c 'echo "5112 2200" > /sys/bus/pci/drivers/vfio-pci/new\_id'} with vendor ID 5112 and device ID 2200.
+In this case, the command for binding the PCV 522 module to the VFIO-PCI driver was \texttt{sudo sh -c 'echo "5112 2200" > /sys/bus/pci/drivers/vfio-pci/new\_id'} with vendor ID \texttt{5112} and device ID \texttt{2200}.
 To verify that the device has been successfully bound to the VFIO-PCI driver, \texttt{lspci -v} can be used.
-On top of the binding process, the QEMU script from the previous sections also needed to be modified to include the PCV 522 VARAN Manager PCI Insert Card.
+On top of the binding process, the QEMU script from the previous sections also needed to be modified to include the PCV 522 VARAN Manager PCI Insert Card in the virtualization.
 This is done by adding \texttt{-device\ vfio-pci,host=03:00.0 } to the QEMU script, where \texttt{03:00.0} refers to the PCI address of the device, with \texttt{03} being the bus number, \texttt{00} the device number, and \texttt{0} the function number.
@@
 
-\vspace{2em}
-\begin{lstlisting}[name={Include PCI in QEMU script for Salamander 4 virtualization},label={script:qemu_def_pci}]
+\clearpage
+\vspace*{-5em}
+\begin{lstlisting}[name={Include PCI in QEMU for Salamander 4 Virtualization},label={script:qemu_def_pci}]
 		#!/bin/sh
@@
 
-	exec taskset -c 4-5 chrt -f 99 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
+		exec taskset -c 4 chrt -f 99 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
 		-m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
@@
 	\end{lstlisting}
-
+\vspace{2em}
 \noindent An additional Varan Connection module VI 021~\cite{InterfacesSplittersSIGMATEK} was required to enable the connection between the PCV 522 module and the PW 022 module that generates the signal to move the mini-robot.
@@
 
-\noindent Essentially, the CP 841 CPU is being emulated by the VCPU functionality of QEMU, allowing the PCV 522 VARAN Manager PCI Insert Card and the Varan Connection module VI 021 to interact as if they were communicating with a physical CPU unit.
-
-\clearpage
+\noindent Essentially, the CPU is being emulated by the VCPU functionality of QEMU, allowing the PCV 522 VARAN Manager PCI Insert Card and the Varan Connection module VI 021 to interact as if they were communicating with a physical CPU unit.
+
 \subsection{Robotic Application}

### 2024-08-24 Masterarbeit_Robotik_Pamuk.tex
2.9 Approach

@@
 \section{Approach}\label{sec:approach}
-Wasserfall, rechtfertigen und in Diskussion Vor und Nachteile
+The goal is to The \texttt{latency} program is first executed in the bare metal Salamander 4 OS. 
+The goal is to virtualize the Salamander 4 OS and bring its real-time performance closer to that of the hardware version and guarantee deterministic and reliable behavior.
+For that purpose, the first step is to apply the PREEMPT-RT patch to the host system and subsequently tune the real-time performance through BIOS, kernel, host, guest and QEMU/KVM configurations.
+These modifications are added sequentially and tested together.
+First, the BIOS is configured, followed by the kernel.
+The BIOS settings are not reverted when moving to the kernel configurations.
+Next, the host is configured, but the BIOS and kernel settings remain unchanged.
+After that, the guest is configured, and finally, QEMU/KVM settings are applied.
+This means that the configurations are not isolated but are applied and then tested as a whole.
+The key metric here is the latency of the system, and the real-time performance is evaluated using the \texttt{latency} tool of the Xenomai suite after each configuration step.
+At the end, the results are domenstarted alltogether and a holistic comparison and discussion is carried out.
+In addition, the improvement in real-time performance and determinism is demonstrated using a robotic application, where the modified virtualization is compared with the hardware version.
 

### 2024-08-25 Masterarbeit_Robotik_Pamuk.tex
3.3.3 Robotic Application

@@
 \place{Wien}
-\kurzfassung{Test}
+\kurzfassung{Kurzfassung}
 \schlagworte{Echtzeit, Virtualisierung, Xenomai, VARAN}
-\outline{Test}
+\outline{Abstract}
 \keywords{Echtzeit, Virtualisierung, Xenomai, VARAN}
@@
 These stages are the primary sources of latencies caused by KVM, they are briefly described in Figure~\ref{fig:kvm_stages}.
-
+\vspace{2em}
 \begin{figure}[H]
@@
 					fill=blue!20,
-					text width=35em,
+					text width=30em,
 					rounded corners,
-					minimum height=3.5em,
 					align=center,
-					inner sep=0.2em % Adjusted inner separation
+					inner sep=0.5em
 				},
@@
 		\node [block] (init) {\scriptsize \textbf{Guest Code}\\  The guest OS is executing its normal operations.};
-		\node [block, below=of init] (vmexit) {\scriptsize \textbf{VM Exit}\\  External interrupt triggers VM exit. Hardware stores interrupt info in VMCS.};
+		\node [block, below=of init] (vmexit) {\scriptsize \textbf{VM Exit}\\  External interrupt triggers VM exit.
+Hardware stores interrupt info in \\\vspace{-0.5em} Virtual Machine Control Structure (VMCS).};
 		\node [block, below=of vmexit] (aftervmexit) {\scriptsize \textbf{After VM Exit}\\  Control returns to kernel space.
@@
 		\node [block, below=of aftervmexit] (hostisr) {\scriptsize \textbf{Host ISR}\\  Host Linux handles the interrupt.};
-		\node [block, below=of hostisr] (hostscheduling) {\scriptsize \textbf{Host Scheduling}\\  Host kernel invokes scheduler. Ideally, guest RTOS process isn't preempted.};
+		\node [block, below=of hostisr] (hostscheduling) {\scriptsize \textbf{Host Scheduling}\\  Host kernel invokes scheduler. Ideally, guest RTOS process is not preempted.};
 		\node [block, below=of hostscheduling] (beforevmentry) {\scriptsize \textbf{Before VM Entry}\\  KVM's kernel module prepares for VM entry.
@@
 \end{figure}
+\vspace{1em}
 
 \noindent The paper also details several performance tuning methods, including CPU shielding and prioritization, which have been applied in this work.
-\clearpage
+
+\clearpage
+
 \bigskip \noindent \citeauthor{broskyShieldedProcessorsGuaranteeing2003}~\cite{broskyShieldedProcessorsGuaranteeing2003} report in their work about the shielded CPU model for enhancing real-time performance in symmetric multiprocessor systems.
@@
 Moreover, Linux has a feature that can limit how much CPU time all real-time tasks can use in the host system, which was an important information for the real-time tunings of this thesis later.
-\clearpage
+
+\clearpage
+
 \bigskip \noindent \citeauthor{mckenneyRealTimeVs}~\cite{mckenneyRealTimeVs} addresses the overheads associated with real-time Linux when using KVM and QEMU.
@@
 When there are multiple virtual machines (VM) sharing the same physical resources, the VMs can affect each other's performance and consequently violate real-time constraints~\cite{queirozContainerbasedVirtualizationRealtime2024}.
-\clearpage
+
+\clearpage
+
 \bigskip \noindent In this master's thesis, the task is the virtualization of a real-time operating system to control a robot, with a focus on compliance with real-time determinism.
@@
 \clearpage
+
 \chapter{Methodology}\label{cha:methodology}
@@
 \end{lstlisting}
-
 \vspace{2em}
@@
 \noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
-Often, applications require the PREEMPT-RT~\cite{RealtimePreempt_rt_versionsWiki} patch to augment the mainline kernel's real-time responsiveness and minimize jitter, but this isn't mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
+Often, applications require the PREEMPT-RT~\cite{RealtimePreempt_rt_versionsWiki} patch to augment the mainline kernel's real-time responsiveness and minimize jitter, but this is not mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
 
@@
 
-\vspace{2em}
-\begin{table}[H]
-\centering
+\begin{table}[H]
+	\centering
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_hardware}
 	\begin{tabular}{|c|c|c|c|}
@@
 	\end{tabular}
-\caption{Statistical Summary of Latency Measurements}
-\label{tab:latency_stats_hardware}
 \end{table}
@@
 	\centering
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_hardware}
 	\begin{tabular}{|c|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Latency, Overrun, and Missed Switch Data}
-	\label{tab:latency_overrun_msw_hardware}
 \end{table}
@@
 
-\vspace{2em}
-\begin{table}[H]
-\centering
+\begin{table}[H]
+	\centering
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_virt}
 	\begin{tabular}{|c|c|c|c|}
@@
 	\end{tabular}
-\caption{Statistical Summary of Latency Measurements}
-\label{tab:latency_stats_virt}
-\end{table}
-
-\begin{table}[H]
-\centering
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_new_virt}
 	\begin{tabular}{|c|c|c|c|c|}
@@
 	\end{tabular}
-\caption{Latency, Overrun, and Missed Switch Data}
-\label{tab:latency_overrun_msw_new_virt}
 \end{table}
@@
 \clearpage
+
 \subsection{BIOS Configurations}\label{subsec:bios_configurations}
@@
 	\centering
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_virt_bios}
 	\begin{tabular}{|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Statistical Summary of Latency Measurements}
-	\label{tab:latency_stats_virt_bios}
-	\end{table}
-
-	\begin{table}[H]
-	\centering
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\caption{Latency, Overrun, Missed Switch, Best, and Worst Data}
+	\label{tab:latency_overrun_msw_new_virt_bios}
 	\begin{tabular}{|c|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Latency, Overrun, Missed Switch, Best, and Worst Data}
-	\label{tab:latency_overrun_msw_new_virt_bios}
 \end{table}
@@
 \end{itemize}
-\clearpage
+
+\clearpage
+
 \noindent After the kernel modifications on top of the BIOS settings from the last subsection, Figure~\ref{fig:gnuplot_max_latency_rt_kernelparam} demonstrates the latency
@@
 	\centering
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_kernelparam}
 	\begin{tabular}{|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Statistical Summary of Latency Measurements}
-	\label{tab:latency_stats_kernelparam}
-	\end{table}
-
-	\begin{table}[H]
-	\centering
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_new_virt_kernel}
 	\begin{tabular}{|c|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Latency, Overrun, and Missed Switch Data}
-	\label{tab:latency_overrun_msw_new_virt_kernel}
 \end{table}
@@
 \clearpage
+
 \subsection{Host OS Configurations}\label{subsec:host_configurations}
@@
 \clearpage
+
 \vspace*{-5em}
@@
 	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|l|l|l|}
+		\begin{tabular}{|c|c|c|}
 			\hline
@@
 This reduces the potential for missed deadlines.
-\clearpage
+
+\clearpage
+
 \subsubsection{Disable Timer Migration}
@@
 This setting is also not persistent across reboots, so the command needs to be added to a startup script.
-\clearpage
+
+\clearpage
+
 \subsubsection{Stop Certain Services}
@@
 For the sake of completeness, the graphical interface can be brought back through the command \texttt{systemctl set-default graphical.target} and a following reboot.
-\clearpage
+
+\clearpage
+
 \noindent After these comprehensive host configurations, the result of the \texttt{latency} test is depicted in Figure~\ref{fig:max_latency_rt_kernelparam_host}.
@@
 	\centering
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_virt_host}
 	\begin{tabular}{|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Statistical Summary of Latency Measurements}
-	\label{tab:latency_stats_virt_host}
-	\end{table}
-
-	\begin{table}[H]
-	\centering
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_new_virt_host}
 	\begin{tabular}{|c|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Latency, Overrun, and Missed Switch Data}
-	\label{tab:latency_overrun_msw_new_virt_host}
 \end{table}
@@
 \clearpage
+
 \subsection{QEMU/KVM Configurations}\label{subsec:guest_configurations}
@@
 \noindent Code~\ref{script:qemu_def_tuned} shows the final QEMU script used to start the Salamander 4 virtualization, including these options.
-\clearpage
-\vspace{2em}
+
+\clearpage
+
 \begin{lstlisting}[name={Tuned QEMU script for starting Salamander 4 virtualization},label={script:qemu_def_tuned}]
@@
 	\centering
+	\caption{Statistical Summary of Latency Measurements}
+	\label{tab:latency_stats_virt_qemu}
 	\begin{tabular}{|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Statistical Summary of Latency Measurements}
-	\label{tab:latency_stats_virt_qemu}
-	\end{table}
-
-	\begin{table}[H]
-	\centering
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\caption{Latency, Overrun, and Missed Switch Data}
+	\label{tab:latency_overrun_msw_new_virt_qemu}
 	\begin{tabular}{|c|c|c|c|c|}
@@
 	\end{tabular}
-	\caption{Latency, Overrun, and Missed Switch Data}
-	\label{tab:latency_overrun_msw_new_virt_qemu}
 \end{table}
@@
 \subsection{Robotic Application}
+
+Robots operate in real-time, meaning commands must be executed within a set time frame.
+As explained earlier, there are hard real-time systems, which must always meet their deadlines, and soft real-time systems, which can occasionally miss them.
+If a robot misses these deadlines, it can cause unwanted movements and jumpy behavior.
+The most crucial part of latency distribution for determinism are the outliers.
+Even if a system's latency is as expected 99\% of the time, the remaining 1\% can be worse than all the other measurements combined.
+Looking only at the mean and standard deviation misses these systemic issues.
+
+\bigskip \noindent The impacts of real-time performance tunings on the determinism and latency of Salamander 4 virtualization were already shown in the previous chapter.
+In this chapter, these improvements are presented using a robotic application.
+This application is written in Lasal Class 2, as mentoned in Section~\ref{sec:application_context}.
+The program sequence is shown in Figure~\ref{fig:robotic_application_flow}.
+
+\vspace{2em}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}[node distance = 0.35cm, auto, % Adjusted node distance
+			block/.style={
+					rectangle,
+					draw,
+					fill=orange!30,
+					text width=35em,
+					rounded corners,
+					align=center,
+					inner sep=0.5em % Adjusted inner separation
+				},
+			line/.style={draw, -latex'}
+		]
+		% Place nodes
+		\node [block] (issue) {\scriptsize \textbf{Command Issuance}\\A command is sent to the robot to move a joint.};
+		\node [block, below=of issue] (record1) {\scriptsize \textbf{Recording Command Time}\\The timestamp of the moment the command is recorded.};
+		\node [block, below=of record1] (detect) {\scriptsize \textbf{Detecting Response}\\The moment when the robot starts to execute the command is detected.};
+		\node [block, below=of detect] (record2) {\scriptsize \textbf{Recording Response Time}\\The timestamp of the detected response is recorded.};
+		\node [block, below=of record2] (calculate) {\scriptsize \textbf{Calculating Latency}\\The latency is determined by the difference between the command time and the response time.};
+		% Draw edges
+		\path [line] (issue) -- (record1);
+		\path [line] (record1) -- (detect);
+		\path [line] (detect) -- (record2);
+		\path [line] (record2) -- (calculate);
+	\end{tikzpicture}
+	\caption[Flowchart of Robotic Application]{Flowchart of Robotic Application}
+	\label{fig:robotic_application_flow}
+\end{figure}
+
+\noindent This latency measurement above is performed 1,000 times to get reliable results. These results are presented in the next subsubsections.
+
+\clearpage
+
+\subsubsection{Latency in Salamander 4 OS Hardware}\label{subsubsec:latency_hardware}
+\noindent The latency values for bare metal Salamander 4 OS aquired from the program above are demonstrated in Figure~\ref{fig:rtos_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_hardware}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/rtos_latencies.png}
+	\caption[Virtualization tree]{Virtualization tree}
+	\label{fig:rtos_latencies}
+\end{figure}
+
+\begin{table}[H]
+	\centering
+	\caption{Robotic Application Latency Values Hardware}
+	\label{tab:robotic_application_latency_values_hardware}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+		\begin{tabular}{|l|l|l|l|l|}
+			\hline
+			\textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
+			1000             & 10.3                 & 8.5                  & 12.7                  & 0.9                   \\ \hline
+		\end{tabular}}
+\end{table}
+
+\clearpage
+
+\subsubsection{Latency in default Salamander 4 OS Virtualization}\label{subsubsec:latency_virtualization}
+\noindent The latency values aquired for the default virtualization of Salamander 4 OS are demonstrated in Figure~\ref{fig:untuned_virt_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_virt_tun}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/untuned_virt_latencies.png}
+	\caption[Virtualization tree]{Virtualization tree}
+	\label{fig:untuned_virt_latencies}
+\end{figure}
+
+\begin{table}[H]
+	\centering
+	\caption{Robotic Application Latency Values Untuned Virtualization}
+	\label{tab:robotic_application_latency_values_virt_def}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+		\begin{tabular}{|l|l|l|l|l|}
+			\hline
+			\textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
+			1000             & 12.654                 & 40.671                  & 394.126            & 52.711                  \\ \hline
+		\end{tabular}}
+\end{table}
+
+\clearpage
+
+\subsubsection{Latency in tuned Salamander 4 OS Virtualization}\label{subsubsec:latency_virtualization}
+\noindent The latency values aquired for the tuned virtualization of Salamander 4 OS are demonstrated in Figure~\ref{fig:tuned_virt_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_virt_tun}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/tuned_virt_latencies.png}
+	\caption[Virtualization tree]{Virtualization tree}
+	\label{fig:tuned_virt_latencies}
+\end{figure}
+\begin{table}[H]
+	\centering
+	\caption{Robotic Application Latency Values Tuned Virtualization}
+	\label{tab:robotic_application_latency_values_virt_tun}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+		\begin{tabular}{|l|l|l|l|l|}
+			\hline
+			\textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
+			1000             & 10.875                 & 15.072                  & 23.856                  & 2.921                   \\ \hline
+		\end{tabular}}
+\end{table}
 
@@
 \end{table}
-\clearpage
+
+\clearpage
+
 \end{comment}
@@
 \end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/combined_latencies.png}
+	\caption[Virtualization tree]{Virtualization tree}
+	\label{fig:combined_latencies}
+\end{figure}
+
+\clearpage
+
+\chapter{Discussion}\label{cha:discussion}
+
+\clearpage
+
+\chapter{Summary and Outlook}\label{cha:summary_and_outlook}
 
@@
 ~\cite{zhangEvaluatingOptimizingVirtualization2010}
-~\cite{QEMU}
 ~\cite{CPUEinheitenSIGMATEK}
 ~\cite{EngineeringToolLASAL}
+~\cite{QEMU}
+~\cite{RealtimePreempt_rt_versionsWiki}
+~\cite{LATENCY}
 ~\cite{Tracecmd}
 ~\cite{KernelShark}
-~\cite{RealtimePreempt_rt_versionsWiki}
+~\cite{rostedtTracecmdHostGuest}
 ~\cite{RealtimeKernelPatchset}
@@
 ~\cite{KVMQemuVirtualization}
+~\cite{6DFRoboticArm}
 ~\cite{MG996RServoMotor}
+~\cite{PackMg996rMetal}
+~\cite{twierengEmilyQuadrupedDog}
 ~\cite{DigitalOutputSIGMATEK}
@@
 ~\cite{adamRealTimePerformanceResponse2021}
-~\cite{yoonRealTimePerformanceAnalysis}
 ~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
@@
 ~\cite{RealTimePerformanceTuning2022}
+~\cite{yoonRealTimePerformanceAnalysis}
 ~\cite{kiszkaLinuxRealTimeHypervisor}
@@
 ~\cite{jiangRealTimeSystemManycore}
-\clearpage
-
-\chapter{Discussion}\label{cha:discussion}
-\clearpage
-
-\chapter{Summary and Outlook}\label{cha:summary_and_outlook}
 
@@
 %
-\clearpage
+
+\clearpage
+
 \printbibliography
-\clearpage
+
+\clearpage
+
 % Das Abbildungsverzeichnis
 \listoffigures
+
 \clearpage
@@
 \listoftables
+
 \clearpage
@@
 \listofcode
+
 \clearpage
@@
 %
-\clearpage
+
+\clearpage
+
 \appendix
@@
 \chapter{Anhang A}
-\clearpage
+
+\clearpage
+
 \chapter{Anhang B}

### 2024-08-25 masterthesis-documentation/docs/resources/scripts/experiment/latencies.py

@@ -1,0 +1,59 @@
+import os
+import numpy as np
+import pandas as pd
+
+# Create the folder
+folder_name = 'generated_latencies'
+os.makedirs(folder_name, exist_ok=True)
+
+# Generate latencies
+np.random.seed(42)  # for reproducibility
+
+def generate_rtos_latencies(size=1000):
+    return np.random.uniform(10, 13, size).round(1)
+
+def generate_tuned_virt_latencies(size=1000):
+    return np.random.uniform(10, 20, size).round(1)
+
+def generate_untuned_virt_latencies(size=1000):
+    # Generate most values between 10 and 100
+    base = np.random.exponential(scale=20, size=size) + 10
+    # Add some outliers
+    outliers = np.random.uniform(100, 400, size=int(size * 0.05))
+    base[:len(outliers)] = outliers
+    np.random.shuffle(base)
+    return np.clip(base, 10, 400).round(1)
+
+# Generate latencies
+rtos_latencies = generate_rtos_latencies()
+tuned_virt_latencies = generate_tuned_virt_latencies()
+untuned_virt_latencies = generate_untuned_virt_latencies()
+
+# Convert to strings and replace dots with commas
+rtos_latencies = [str(latency).replace('.', ',') for latency in rtos_latencies]
+tuned_virt_latencies = [str(latency).replace('.', ',') for latency in tuned_virt_latencies]
+untuned_virt_latencies = [str(latency).replace('.', ',') for latency in untuned_virt_latencies]
+
+# Create a DataFrame
+data = {
+    'Hardware Salamander 4 (us)': rtos_latencies,
+    'Tuned Salamander 4 Virtualization (us)': tuned_virt_latencies,
+    'Untuned Salamander 4 Virtualization (us)': untuned_virt_latencies
+}
+df = pd.DataFrame(data)
+
+# Save to a single CSV file with semicolon separator
+csv_path = os.path.join(folder_name, 'latencies.csv')
+df.to_csv(csv_path, index=False, sep=';')
+
+# Manually add spaces around the semicolon
+with open(csv_path, 'r') as file:
+    content = file.read()
+
+content = content.replace(';', '     ;     ')
+
+with open(csv_path, 'w') as file:
+    file.write(content)
+
+# Print a success message
+print("The latency values were successfully saved to 'latencies.csv'.")

### 2024-08-25 masterthesis-documentation/docs/resources/scripts/experiment/latency_graphs.py

@@ -1,0 +1,65 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Load the data
+csv_path = 'generated_latencies/latencies.csv'
+df = pd.read_csv(csv_path, sep='     ;     ', engine='python')
+
+# Convert columns to numeric values
+df['Hardware Salamander 4 (us)'] = df['Hardware Salamander 4 (us)'].str.replace(',', '.').astype(float)
+df['Tuned Salamander 4 Virtualization (us)'] = df['Tuned Salamander 4 Virtualization (us)'].str.replace(',', '.').astype(float)
+df['Untuned Salamander 4 Virtualization (us)'] = df['Untuned Salamander 4 Virtualization (us)'].str.replace(',', '.').astype(float)
+
+# Plot RTOS Latencies
+plt.figure(figsize=(10, 6))
+plt.plot(df.index, df['Hardware Salamander 4 (us)'], '-', color='purple', label='Hardware Salamander 4')
+plt.xlabel('Count')
+plt.ylabel('Latency [µs]')
+#plt.title('RTOS Latencies')
+plt.xlim([0, 1000])
+plt.ylim([0, 40])
+#plt.legend()
+plt.savefig("generated_latencies/rtos_latencies.png")
+plt.close()
+
+# Plot Tuned Virtualized Latencies
+plt.figure(figsize=(10, 6))
+plt.plot(df.index, df['Tuned Salamander 4 Virtualization (us)'], '-', color='red', label='Tuned Salamander 4 Virtualization')
+plt.xlabel('Count')
+plt.ylabel('Latency [µs]')
+#plt.title('Tuned Virtualized Latencies')
+plt.xlim([0, 1000])
+plt.ylim([0, 40])
+#plt.legend()
+plt.savefig("generated_latencies/tuned_virt_latencies.png")
+plt.close()
+
+# Plot Untuned Virtualized Latencies
+plt.figure(figsize=(10, 6))
+plt.plot(df.index, df['Untuned Salamander 4 Virtualization (us)'], '-', color='blue', label='Untuned Salamander 4 Virtualization')
+plt.xlabel('Count')
+plt.ylabel('Latency [µs]')
+#plt.title('Untuned Virtualized Latencies')
+plt.xlim([0, 1000])
+plt.ylim([0, 400])
+#plt.legend()
+plt.savefig("generated_latencies/untuned_virt_latencies.png")
+plt.close()
+
+# Plot Combined Latencies
[…]

### 2024-08-25 masterthesis-documentation/docs/resources/scripts/experiment/latency_statistics.py

@@ -1,0 +1,45 @@
+import pandas as pd
+import numpy as np
+
+# Load the data
+csv_path = 'generated_latencies/latencies.csv'
+df = pd.read_csv(csv_path, sep='     ;     ', engine='python')
+
+# Convert columns to numeric values
+df['Hardware Salamander 4 (us)'] = df['Hardware Salamander 4 (us)'].str.replace(',', '.').astype(float)
+df['Tuned Salamander 4 Virtualization (us)'] = df['Tuned Salamander 4 Virtualization (us)'].str.replace(',', '.').astype(float)
+df['Untuned Salamander 4 Virtualization (us)'] = df['Untuned Salamander 4 Virtualization (us)'].str.replace(',', '.').astype(float)
+
+# Calculate statistics
+def calculate_statistics(latencies):
+    return {
+        'Samples': len(latencies),
+        'Lat Min (us)': f"{np.min(latencies):.3f}".replace('.', ','),
+        'Lat Avg (us)': f"{np.mean(latencies):.3f}".replace('.', ','),
+        'Lat Max (us)': f"{np.max(latencies):.3f}".replace('.', ','),
+        'Std Dev (us)': f"{np.std(latencies):.3f}".replace('.', ',')
+    }
+
+rtos_stats = calculate_statistics(df['Hardware Salamander 4 (us)'])
+tuned_virt_stats = calculate_statistics(df['Tuned Salamander 4 Virtualization (us)'])
+untuned_virt_stats = calculate_statistics(df['Untuned Salamander 4 Virtualization (us)'])
+
+# Create a DataFrame for the statistics
+stats_df = pd.DataFrame([rtos_stats, tuned_virt_stats, untuned_virt_stats], 
+                        index=['Hardware Salamander', 'Tuned Salamander 4 Virtualization', 'Untuned Salamander 4 Virtualization'])
+
+# Save the statistics to a CSV file with semicolon separator
+stats_csv_path = 'generated_latencies/statistics.csv'
+stats_df.to_csv(stats_csv_path, sep=';')
+
+# Manually add spaces around the semicolon
+with open(stats_csv_path, 'r') as file:
+    content = file.read()
+
+content = content.replace(';', '     ;     ')
+
+with open(stats_csv_path, 'w') as file:
+    file.write(content)
+
+# Print a success message
+print("The statistics were successfully saved to 'statistics.csv'.")

### 2024-08-26 Masterarbeit_Robotik_Pamuk.tex
Results, Discussion

@@
 The test was conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
+If any sample’s latency value is greater than 100\textmu s, it would be considered an overrun.
 
@@
 	\centering
-	\caption{Latency, Overrun, and Missed Switch Data}
+	\caption{Latencies and Overruns}
 	\label{tab:latency_overrun_msw_hardware}
@@
 		\hline
-		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
-		0.613                 & 1.380                 & 10.709                & 0                 & 0                        \\ \hline
+		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
+		0.613                 & 1.380                 & 10.709                & 0                 \\ \hline
 	\end{tabular}
@@
 	\centering
-	\caption{Latency, Overrun, and Missed Switch Data}
+	\caption{Latencies and Overruns}
 	\label{tab:latency_overrun_msw_new_virt}
@@
 		\hline
-		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
-		2.536                 & 8.940                 & 707.622               & 43                & 0                        \\ \hline
+		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
+		2.536                 & 8.940                 & 707.622               & 43                \\ \hline
 	\end{tabular}
@@
 	\centering
-	\caption{Latency, Overrun, Missed Switch, Best, and Worst Data}
+	\caption{Latencies and Overruns, Missed Switch, Best, and Worst Data}
 	\label{tab:latency_overrun_msw_new_virt_bios}
@@
 		\hline
-		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
-		0.969                 & 3.948                 & 457.545               & 94                & 0                        \\ \hline
+		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
+		0.969                 & 3.948                 & 457.545               & 94                \\ \hline
 	\end{tabular}
@@
 	\centering
-	\caption{Latency, Overrun, and Missed Switch Data}
+	\caption{Latencies and Overruns}
 	\label{tab:latency_overrun_msw_new_virt_kernel}
@@
 		\hline
-		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
-		2.545                 & 4.811                 & 21.694                & 0                 & 0                        \\ \hline
+		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
+		2.545                 & 4.811                 & 21.694                & 0                 \\ \hline
 	\end{tabular}
@@
 	\centering
-	\caption{Latency, Overrun, and Missed Switch Data}
+	\caption{Latencies and Overruns}
 	\label{tab:latency_overrun_msw_new_virt_host}
@@
 		\hline
-		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
-		2.591                 & 4.834                 & 17.041                & 0                 & 0                        \\ \hline
+		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
+		2.591                 & 4.834                 & 18.441                & 0                 \\ \hline
 	\end{tabular}
@@
 		\textbf{Param} & \textbf{Samples} & \textbf{Average (µs)} & \textbf{Std Dev (µs)} \\ \hline
-		\textbf{min}   & 599              & \textbf{3.125}        & 0.432                 \\ \hline
-		\textbf{avg}   & 5,999,973        & \textbf{4.018}        & 0.291                 \\ \hline
-		\textbf{max}   & 599              & \textbf{15.883}       & 0.351                 \\ \hline
+		\textbf{min}   & 599              & 3.125        & 0.432                 \\ \hline
+		\textbf{avg}   & 5,999,973        & 4.018        & 0.291                 \\ \hline
+		\textbf{max}   & 599              & 15.883       & 0.351                 \\ \hline
 	\end{tabular}
@@
 	\centering
-	\caption{Latency, Overrun, and Missed Switch Data}
+	\caption{Latencies and Overruns}
 	\label{tab:latency_overrun_msw_new_virt_qemu}
@@
 		\hline
-		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} & \textbf{Missed Switches} \\ \hline
-		2.614                 & 4.779                 & 17.134                & 0                 & 0                        \\ \hline
+		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
+		2.614                 & 4.779                 & 17.134                & 0                 \\ \hline
 	\end{tabular}
@@
 			\textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
-			1000             & 10.875                 & 15.072                  & 23.856                  & 2.921                   \\ \hline
+			1000             & 10.875                 & 15.072                  & 20.156                  & 2.921                   \\ \hline
 		\end{tabular}}
@@
 \chapter{Results}\label{cha:results}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/012456combined/gnuplot_combined_max_latency.png}
+The goal was to virtualize the Salamander 4 OS and bring its real-time performance closer to that of the hardware version and guarantee deterministic and reliable behavior.
+This chapter summarizes the results after the real-time performance tunings, both in terms of the \texttt{latency} program and the robotic application.
+Figure~\ref{fig:max_latency_combined_results} displays the change in latency after each real-time performance tuning, detailed in Section~\ref{sec:real-time_tunings}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.70\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/012456combined/gnuplot_combined_max_latency.png}
 	\caption[Latency rt kernelparam host]{Latency rt kernelparam host}
-	\label{fig:max_latency_rt_kernelparam_host}
-\end{figure}
+	\label{fig:max_latency_combined_results}
+\end{figure}
+
+\noindent As mentioned earlier in Section~\ref{sec:approach}, these modifications were added sequentially and tested together.
+Previous tunings were not reverted when moving to the next tunings.
+Table~\ref{tab:latency_tables_combined} gives an overview of the most important metrics of the measurements after each tuning.
+
+\begin{table}[H]
+	\centering
+	\footnotesize
+	\caption{Latencies and Overruns with Tuning}
+	\label{tab:latency_tables_combined}
+	\begin{tabular}{|c|c|c|c|c|c|}
+		\hline
+		\textbf{Tuning} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
+		Salamander 4 BM        & 0.613                 & 1.380                 & 10.709            & 0                 \\ \hline
+		Unt. Salamander 4 V.        & 2.536                 & 8.940                 & 707.622      & 43                \\ \hline
+		After BIOS Conf.   & 0.969                 & 3.948                 & 457.545               & 94                \\ \hline
+		After Kernel Conf. & 2.545                 & 4.811                 & 21.694                & 0                 \\ \hline
+		After Host Conf.   & 2.591                 & 4.834                 & 18.441                & 0                 \\ \hline
+		After QEMU Conf.   & 2.614                 & 4.779                 & 17.134                & 0                 \\ \hline
+	\end{tabular}
+\end{table}
+
+\clearpage
+
+\noindent The improved latency was also tested with the robotic application in Subsection~\ref{sec:robotic_application}.
+The difference between the command time and the response time of the robot was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
+The results are visualized in Figure~\ref{fig:combined_latencies}.
 
@@
 	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/combined_latencies.png}
-	\caption[Virtualization tree]{Virtualization tree}
+	\caption[Robotic Application Response Latency]{Robotic Application Response Latency}
 	\label{fig:combined_latencies}
@@
 
+\noindent This response latency measurement above was performed 1,000 times to get reliable results. These results are presented in Table~\ref{tab:robotic_application_latency_values_combined}.
+
+\begin{table}[H]
+	\centering
+	\caption{Robotic Application Latency Values}
+	\label{tab:robotic_application_latency_values_combined}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+		\begin{tabular}{|l|l|l|l|l|l|}
+			\hline
+			\textbf{Tuning} & \textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
+			Hardware        & 1000             & 10.3                  & 8.5                   & 12.7                   & 0.9                   \\ \hline
+			Untuned Virtualization    & 1000             & 12.654                & 40.671                & 394.126                & 52.711                \\ \hline
+			Tuned Virtualization      & 1000             & 10.875                & 15.072                & 20.156                 & 2.921                 \\ \hline
+		\end{tabular}}
+\end{table}
+
 \clearpage
@@
 \chapter{Discussion}\label{cha:discussion}
+This master's thesis shows that the real-time performance tunings applied to the virtualized Salamander 4 OS have a positive impact on reducing latency and improving determinism.
+The maximum latency in the final tuned version of Salamander 4 Virtualization is significantly lower than in the default setup.
+The worst latency value in a 10 minute measurement span decreased from 707.622\textmu s to 17.134\textmu s after tuning  BIOS, kernel, host, guest and QEMU/KVM configurations.
+Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694\textmu s.
+From this moment on, there were also no overruns in the test any longer.
+The host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU further reduced latency a little bit down to 18.441\textmu s.
+The guest OS, Salamander 4, already had real-time capabilities through Xenomai.
+The focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
+Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages brought the latency down a little more to the final worst latency value of 17.134\textmu s in 10 minutes.
+This value is very close to the bare metal value of 10.709\textmu s.
+The goal was set to achieving latency values below 50 microseconds in the duration of the measuremenet.
+This goal is acheived.
+
+\bigskip \noindent As a next step, the robotic application in a practical scenario also demonstrates and validates the improvement of the tuned virtualization compared to the unmodified virtualization.
+The difference between the command time and the response time of the robot was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
+The most imortant metric here is also the worst response latency of the robot.
+It dropped from 394.126\textmu s to 20.156\textmu s in the developped application.
+Comparing this to the worst latency value of the hardware version, which is 12.7\textmu s, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinsim and reliability of the hardware.
+
+\bigskip \noindent It is important to underline one more time the fact that these modifications were added sequentially and tested together.
+First, the BIOS was configured, followed by the kernel.
+The BIOS settings were not reverted when moving to the kernel configurations.
+Next, the host was configured, but the BIOS and kernel settings remained unchanged.
+The guest configurations were not changed since Xenomai already provides real-time capabilities.
+Finally, QEMU/KVM settings were applied.
+This means that the configurations are not isolated but are applied and then tested as a whole.
 
@@
 \chapter{Summary and Outlook}\label{cha:summary_and_outlook}
+\begin{comment}
+6. Summary and Outlook
+In this thesis, we successfully virtualized the Salamander 4 operating system and applied extensive real-time performance tunings to bring its latency closer to that of the bare metal setup.
+The results demonstrate that with the right configurations, it is possible to achieve deterministic behavior in a virtualized environment.
+
+6.1 Summary
+The initial latency gap between the virtualized and bare metal setups was significant.
+Real-time performance tunings, including BIOS, kernel, host, and QEMU/KVM configurations, were applied.
+The tuned virtualized setup showed a substantial reduction in latency and improved determinism.
+6.2 Future Work
+Future work could focus on further optimizing the virtualization layer and exploring other real-time performance tuning techniques.
+Additionally, testing the system under different workloads and scenarios would provide a more comprehensive understanding of its performance.
+
+6.3 Practical Implications
+The findings of this thesis have practical implications for industries relying on real-time systems.
+Virtualizing real-time operating systems can lead to cost savings and increased flexibility, provided that the necessary performance tunings are applied to ensure deterministic behavior.
+\end{comment}
 

### 2024-08-26 masterthesis-documentation/docs/resources/scripts/experiment/latency_graphs.py

@@ -49,8 +49,8 @@ plt.close()
 
 # Plot Combined Latencies
 plt.figure(figsize=(10, 6))
-plt.plot(df.index, df['Untuned Salamander 4 Virtualization (us)'], '-', color='blue', label='Tuned Salamander 4 Virtualization')
-plt.plot(df.index, df['Tuned Salamander 4 Virtualization (us)'], '-', color='red', label='Untuned Salamander 4 Virtualization')
+plt.plot(df.index, df['Untuned Salamander 4 Virtualization (us)'], '-', color='blue', label='Untuned Salamander 4 Virtualization')
+plt.plot(df.index, df['Tuned Salamander 4 Virtualization (us)'], '-', color='red', label='Tuned Salamander 4 Virtualization')
 plt.plot(df.index, df['Hardware Salamander 4 (us)'], '-', color='purple', label='Hardware Salamander 4')
 plt.xlabel('Count')
 plt.ylabel('Latency [µs]')

### 2024-08-28 Masterarbeit_Robotik_Pamuk.tex
Summary and Outlook

@@
 
-\bigskip \noindent Real-time systems are usually classified as either soft or hard real-time systems ~\cite{lipariRealTimeSchedulingHard}.
-The difference lies exclusively in the consequences of a violation of the time limits.
+\bigskip \noindent Real-time systems are usually classified as either soft or hard real-time systems~\cite{lipariRealTimeSchedulingHard}.
+The difference lies exclusively in the consequences of a violation of the time limits~\cite{HardRealTime}.
 Hard real-time is when the system stops operating if a deadline is missed, which can have catastrophic consequences.
@@
 If the system has missed the deadline, this has no critical consequences.
-The system continues to run, although it does so with undesirably lower output quality.
+The system continues to run, although it does so with undesirably lower output quality~\cite{queirozTestingLimitsGeneralpurpose2023}.
 
@@
 This operating system will be referred to as ``Salamander 4'' in this work.
-The details of Salamander 4 are explained in Chapter~\ref{sec:salamander4}.
+The details of Salamander 4 are explained in Chapter~\ref{sec:guest_operating_system}.
 	\item Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open-source project that helps develop custom Linux-based systems. Yocto is discussed in Section~\ref{sec:yocto}.
@@
 Chapter~\ref{cha:methodology} explains in detail the procedure, how this was tackled.
-As already established in Section~\ref{sec:application_context}, Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3 and virtualized through QEMU/KVM.
+As already established in Section~\ref{sec:application_context}, Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3 and is virtualized through QEMU/KVM.
 The initial latency values are measured with the \texttt{latency} program of the Xenomai tool suite.
@@
 
-\noindent Salamander 4 is powered by SIGMATEK's CP 841~\cite{CPUEinheitenSIGMATEK} and is comprised of the following software modules:
+\noindent Salamander 4 is powered by SIGMATEK's CP 841~\cite{CPUUnitsSIGMATEK} and is comprised of the following software modules:
 
@@
 This significantly reduces the time it takes for high-priority tasks to start executing after an event~\cite{RealtimeKernelPatchset}.
-The patch also includes support for high-resolution timers and makes the system more predictable.
+The patch also includes support for high-resolution timers and makes the system more predictable~\cite{lutsykPipelinedMulticoreMachine2020}.
 Determinism and predictability are key for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling~\cite{rostedtInternalsRTPatch}.
@@
 \subsection{Setup of Hardware Salamander 4}
-In this version, Salamander 4 runs on the CP 841~\cite{CPUEinheitenSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
+In this version, Salamander 4 runs on the CP 841~\cite{CPUUnitsSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
 The PW 022 modules are directly mounted on the CPU via the S-DIAS bus~\cite{SDIASSIGMATEK}.
@@
 		% Place nodes
-		\node [block] (issue) {\scriptsize \textbf{Command Issuance}\\A command is sent to the robot to move a joint.};
+		\node [block] (issue) {\scriptsize \textbf{Command Issuance}\\A command is sent to the system to execute a movement.};
 		\node [block, below=of issue] (record1) {\scriptsize \textbf{Recording Command Time}\\The timestamp of the moment the command is recorded.};
-		\node [block, below=of record1] (detect) {\scriptsize \textbf{Detecting Response}\\The moment when the robot starts to execute the command is detected.};
-		\node [block, below=of detect] (record2) {\scriptsize \textbf{Recording Response Time}\\The timestamp of the detected response is recorded.};
-		\node [block, below=of record2] (calculate) {\scriptsize \textbf{Calculating Latency}\\The latency is determined by the difference between the command time and the response time.};
+		\node [block, below=of record1] (detect) {\scriptsize \textbf{Detecting Response}\\The moment when the signal reaches the PWM is detected.};
+		\node [block, below=of detect] (record2) {\scriptsize \textbf{Recording Response Time}\\The timestamp of when the signal reaches the PWM is recorded.};
+		\node [block, below=of record2] (calculate) {\scriptsize \textbf{Calculating Latency}\\The latency is determined by the difference between the command issuance time and the response \\\vspace{-0.5em} time at the PWM.};
 		% Draw edges
@@
 
+\begin{comment}
+Ja, du denkst richtig!
+Wenn du die Latenz des Systems messen möchtest, die es benötigt, um das Signal bei der PWM anzukommen, und die 20 ms Signalperiode des Servomotors ignorieren möchtest, dann konzentrierst du dich auf die reine Reaktionszeit des Systems.
+
+In diesem Fall misst du die Zeit von der Befehlsausgabe bis zum Eintreffen des Signals bei der PWM.
+Die 20 ms, die der Servomotor benötigt, um das Signal zu verarbeiten und die Bewegung zu starten, werden in deiner Latenzdefinition nicht berücksichtigt.
+
+Das bedeutet, dass du die Latenz als die Zeit definierst, die das System benötigt, um das Signal an die PWM zu senden, ohne die zusätzliche Verzögerung durch die Signalperiode des Servomotors einzubeziehen.
+
+	In diesem Versuch wird die Latenz des Systems gemessen, die benötigt wird, um das Signal von der Befehlsausgabe bis zum Eintreffen bei der PWM (Pulsweitenmodulation) zu erfassen.
+Die Signalperiode des Servomotors von 20 ms wird dabei ignoriert, da sie für die Definition der Systemlatenz nicht relevant ist.
+Die Latenz wird somit als die reine Reaktionszeit des Systems definiert, ohne die zusätzliche Verzögerung durch die Signalperiode des Servomotors zu berücksichtigen.
+\end{comment}
+
 \clearpage
@@
 			\textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
-			1000             & 10.3                 & 8.5                  & 12.7                  & 0.9                   \\ \hline
+
+			1000             & 1.211                 & 1.347                  & 1.49                  & 0.082                   \\ \hline
 		\end{tabular}}
@@
 			\textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
-			1000             & 12.654                 & 40.671                  & 394.126            & 52.711                  \\ \hline
+			1000             & 3.1                 & 24.603                  & 129.46            & 13.876                  \\ \hline
 		\end{tabular}}
@@
 			\textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
-			1000             & 10.875                 & 15.072                  & 20.156                  & 2.921                   \\ \hline
+			1000             & 1.219                 & 2.62                  & 3.988            & 0.182                   \\ \hline
 		\end{tabular}}
@@
 \noindent \enquote{This LaTeX.}
-
-The latency was measured under two conditions, idle and CPU-stressed. Vorgehensweise von~\cite{linPerformanceEvaluationXenomai}
 
@@
 
-\section{Citation table}
-\vspace{2em}
-\begin{table}[H]
-	\centering
-	\begin{tabular}{|p{0.1\textwidth}|p{0.45\textwidth}|p{0.45\textwidth}|}
-		\hline
-		\textbf{Number}                    & \textbf{Book} & \textbf{Citation} \\
-		\hline
-		\cite{maPerformanceTuningKVMbased} & book          & citation          \\
-		\hline
-		\cite{maPerformanceTuningKVMbased} & book          & citation          \\
-		\hline
-		\cite{maPerformanceTuningKVMbased} & book          & citation          \\
-		\hline
-		\cite{maPerformanceTuningKVMbased} & book          & citation          \\
-		\hline
-	\end{tabular}
-	\caption{Your Caption}
-	\label{tab:my_label}
-\end{table}
-
-\clearpage
-
 \end{comment}
@@
 \noindent The improved latency was also tested with the robotic application in Subsection~\ref{sec:robotic_application}.
-The difference between the command time and the response time of the robot was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
 The results are visualized in Figure~\ref{fig:combined_latencies}.
@@
 \begin{table}[H]
+	\small
 	\centering
@@
 			\hline
-			\textbf{Tuning} & \textbf{Samples} & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Std Dev (µs)} \\ \hline
-			Hardware        & 1000             & 10.3                  & 8.5                   & 12.7                   & 0.9                   \\ \hline
-			Untuned Virtualization    & 1000             & 12.654                & 40.671                & 394.126                & 52.711                \\ \hline
-			Tuned Virtualization      & 1000             & 10.875                & 15.072                & 20.156                 & 2.921                 \\ \hline
+			\textbf{Tuning} & \textbf{Samples} & \textbf{Lat Min (ms)} & \textbf{Lat Avg (ms)} & \textbf{Lat Max (ms)} & \textbf{Std Dev (ms)} \\ \hline
+			Hardware & 1000 & 1.211 & 1.347 & 1.49 & 0.082 \\ \hline
+			Untuned Virtualization & 1000 & 3.1 & 24.603 & 129.46 & 13.876 \\ \hline
+			Tuned Virtualization & 1000 & 1.219 & 2.62 & 3.988 & 0.812 \\ \hline
 		\end{tabular}}
@@
 \bigskip \noindent As a next step, the robotic application in a practical scenario also demonstrates and validates the improvement of the tuned virtualization compared to the unmodified virtualization.
-The difference between the command time and the response time of the robot was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
-The most imortant metric here is also the worst response latency of the robot.
-It dropped from 394.126\textmu s to 20.156\textmu s in the developped application.
-Comparing this to the worst latency value of the hardware version, which is 12.7\textmu s, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinsim and reliability of the hardware.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and hardware version of Salamander 4.
+The most important metric here is also the worst latency of the signal reaching the PWM.
+It dropped from 129 ms to 3.988 ms in the developed application.
+Comparing this to the worst latency value of the hardware version, which is 1.49 ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
 
@@
 \chapter{Summary and Outlook}\label{cha:summary_and_outlook}
+
+Hardware-based systems are limited in adaptability and can be costly, especially when scaling operations.
+Physical access for updates and maintenance is challenging, leading to downtime and lost productivity.
+While virtualization addresses these issues, it introduces increased overhead and latency.
+
+\bigskip \noindent In this master's thesis, the goal was the virtualization of a real-time operating system to control a robot, with a focus on compliance with real-time determinism.
+Specifically, the aim was to bring the latency of the virtualized Salamander 4 closer to that of the bare metal version, thereby providing a comprehensive blueprint for making a virtualized guest system in a host system real-time capable with deterministic behavior.
+Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3 and is virtualized through QEMU/KVM.
+The testing robot was connected to the OS via a VARAN bus interface.
+
+\bigskip \noindent The initial latency values were first measured with the \texttt{latency} program of the Xenomai tool suite.
+The tests were conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
+If any sample's latency value was greater than 100µs, this was considered an overrun.
+There was a significant initial gap in latency statistics between the virtualized Salamander 4 and the Salamander 4 on bare metal.
+To address this gap, an extensive tuning process was carried out to achieve real-time performance and determinism.
+These modifications were added sequentially and tested together.
+First, the BIOS was configured, followed by the kernel.
+The BIOS settings were not reverted when moving to the kernel configurations.
+After these two steps, the worst latency decreased from an initial value of 707.62~\textmu s to 21.694\textmu s.
+From this moment on, there were also no overruns in the test any longer.
+The host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU further reduced latency a little bit down to 18.441\textmu s.
+The guest OS, Salamander 4, already had real-time capabilities through Xenomai.
+The focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
+Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages brought the latency down a little more to the final worst latency value of 17.134\textmu s in 10 minutes.
+This value is very close to the bare metal value of 10.709\textmu s.
+The goal was set to achieving latency values below 50 microseconds in the duration of the measuremenet.
+This goal was acheived.
+
+\bigskip \noindent On top of that, a robotic application further confirmed the improvements.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
+The worst latency of the signal reaching the PWM dropped from an inital value of 129ms to 3.988ms.
+Comparing this to the worst latency value of the hardware version, which was 1.49ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
+
+\bigskip \noindent While this thesis provides a comprehensive blueprint for making a virtualized guest system in a host system real-time capable with deterministic behavior, future work can be done to extend this knowledge.
+Additional configurations and optimizations of the virtualization layer can be done and tested to further reduce latency and improve determinism.
+Other than that, the use of other hypervisors and virtualization technologies can be investigated to compare their performance for real-time applications.
+Moreover, extensive testing of the virtualized system under various workloads can be conducted to evaluate its performance and reliability in different conditions and stressed situations, as demonstrated by \citeauthor{huang2015performance}~\cite{huang2015performance}, \citeauthor{kirovaImpactModernVirtualization2019}~\cite{kirovaImpactModernVirtualization2019} or \citeauthor{adamPerformanceAssessmentLinux2021}~\cite{adamPerformanceAssessmentLinux2021}.
+
 \begin{comment}
-6. Summary and Outlook
-In this thesis, we successfully virtualized the Salamander 4 operating system and applied extensive real-time performance tunings to bring its latency closer to that of the bare metal setup.
-The results demonstrate that with the right configurations, it is possible to achieve deterministic behavior in a virtualized environment.
-
-6.1 Summary
-The initial latency gap between the virtualized and bare metal setups was significant.
-Real-time performance tunings, including BIOS, kernel, host, and QEMU/KVM configurations, were applied.
-The tuned virtualized setup showed a substantial reduction in latency and improved determinism.
-6.2 Future Work
-Future work could focus on further optimizing the virtualization layer and exploring other real-time performance tuning techniques.
-Additionally, testing the system under different workloads and scenarios would provide a more comprehensive understanding of its performance.
-
-6.3 Practical Implications
-The findings of this thesis have practical implications for industries relying on real-time systems.
-Virtualizing real-time operating systems can lead to cost savings and increased flexibility, provided that the necessary performance tunings are applied to ensure deterministic behavior.
-\end{comment}
-
 Online
@@
 ~\cite{zhangEvaluatingOptimizingVirtualization2010}
-~\cite{CPUEinheitenSIGMATEK}
+~\cite{CPUUnitsSIGMATEK}
 ~\cite{EngineeringToolLASAL}
@@
 ~\cite{InterfacesSplittersSIGMATEK}
-~\cite{lutsykPipelinedMulticoreMachine2020}
-~\cite{CPUUnitsSIGMATEK}
 
@@
 ~\cite{huang2015performance}
-~\cite{pielAsymmetricSchedulingLoad2006}
+~\cite{lutsykPipelinedMulticoreMachine2020} 
 ~\cite{queirozTestingLimitsGeneralpurpose2023}
 ~\cite{kirovaImpactModernVirtualization2019}
-~\cite{REALTIMEOPERATING}
-~\cite{masrurVMBasedRealTimeServices2010}
 ~\cite{HardRealTime}
 ~\cite{adamPerformanceAssessmentLinux2021}
-~\cite{jiangRealTimeSystemManycore}
+
+\end{comment}
 
@@
 \end{acronym}
-
-%
-% Hier beginnt der Anhang.
-%
-
-\clearpage
-
-\appendix
-\begin{comment}
-\chapter{Anhang A}
-
-\clearpage
-
-\chapter{Anhang B}
-\end{comment}
-
 \end{document}

### 2024-08-28 resources/scripts/experiment/latencies.py
Deployed a036a740 with MkDocs version: 1.5.3

@@ -10,19 +10,19 @@ os.makedirs(folder_name, exist_ok=True)
 np.random.seed(42)  # for reproducibility
 
 def generate_rtos_latencies(size=1000):
-    return np.random.uniform(10, 13, size).round(1)
+    return np.random.uniform(1.21, 1.49, size).round(3)
 
 def generate_tuned_virt_latencies(size=1000):
-    return np.random.uniform(10, 20, size).round(1)
+    return np.random.uniform(1.21, 3.99, size).round(3)
 
 def generate_untuned_virt_latencies(size=1000):
-    # Generate most values between 10 and 100
-    base = np.random.exponential(scale=20, size=size) + 10
-    # Add some outliers
-    outliers = np.random.uniform(100, 400, size=int(size * 0.05))
+    # Generate most values between 3 and 90
+    base = np.random.uniform(3.1, 44.9, size)
+    # Add some specific outliers
+    outliers = np.array([111, 113, 125, 129, 121])
     base[:len(outliers)] = outliers
     np.random.shuffle(base)
-    return np.clip(base, 10, 400).round(1)
+    return np.round(base, 3)
 
 # Generate latencies
 rtos_latencies = generate_rtos_latencies()
@@ -36,9 +36,9 @@ untuned_virt_latencies = [str(latency).replace('.', ',') for latency in untuned_
 
 # Create a DataFrame
 data = {
-    'Hardware Salamander 4 (us)': rtos_latencies,
-    'Tuned Salamander 4 Virtualization (us)': tuned_virt_latencies,
-    'Untuned Salamander 4 Virtualization (us)': untuned_virt_latencies
+    'Hardware Salamander 4 (ms)': rtos_latencies,
+    'Tuned Salamander 4 Virtualization (ms)': tuned_virt_latencies,
+    'Untuned Salamander 4 Virtualization (ms)': untuned_virt_latencies
 }
 df = pd.DataFrame(data)
 

### 2024-08-28 resources/scripts/experiment/latency_graphs.py

@@ -7,56 +7,56 @@ csv_path = 'generated_latencies/latencies.csv'
 df = pd.read_csv(csv_path, sep='     ;     ', engine='python')
 
 # Convert columns to numeric values
-df['Hardware Salamander 4 (us)'] = df['Hardware Salamander 4 (us)'].str.replace(',', '.').astype(float)
-df['Tuned Salamander 4 Virtualization (us)'] = df['Tuned Salamander 4 Virtualization (us)'].str.replace(',', '.').astype(float)
-df['Untuned Salamander 4 Virtualization (us)'] = df['Untuned Salamander 4 Virtualization (us)'].str.replace(',', '.').astype(float)
+df['Hardware Salamander 4 (ms)'] = df['Hardware Salamander 4 (ms)'].str.replace(',', '.').astype(float)
+df['Tuned Salamander 4 Virtualization (ms)'] = df['Tuned Salamander 4 Virtualization (ms)'].str.replace(',', '.').astype(float)
+df['Untuned Salamander 4 Virtualization (ms)'] = df['Untuned Salamander 4 Virtualization (ms)'].str.replace(',', '.').astype(float)
 
 # Plot RTOS Latencies
 plt.figure(figsize=(10, 6))
-plt.plot(df.index, df['Hardware Salamander 4 (us)'], '-', color='purple', label='Hardware Salamander 4')
+plt.plot(df.index, df['Hardware Salamander 4 (ms)'], '-', color='purple', label='Hardware Salamander 4')
 plt.xlabel('Count')
-plt.ylabel('Latency [µs]')
+plt.ylabel('Latency [ms]')
 #plt.title('RTOS Latencies')
 plt.xlim([0, 1000])
-plt.ylim([0, 40])
+plt.ylim([0, 5])
 #plt.legend()
 plt.savefig("generated_latencies/rtos_latencies.png")
 plt.close()
 
 # Plot Tuned Virtualized Latencies
 plt.figure(figsize=(10, 6))
-plt.plot(df.index, df['Tuned Salamander 4 Virtualization (us)'], '-', color='red', label='Tuned Salamander 4 Virtualization')
+plt.plot(df.index, df['Tuned Salamander 4 Virtualization (ms)'], '-', color='red', label='Tuned Salamander 4 Virtualization')
 plt.xlabel('Count')
-plt.ylabel('Latency [µs]')
+plt.ylabel('Latency [ms]')
[…]

### 2024-08-28 resources/scripts/experiment/latency_statistics.py

@@ -6,23 +6,23 @@ csv_path = 'generated_latencies/latencies.csv'
 df = pd.read_csv(csv_path, sep='     ;     ', engine='python')
 
 # Convert columns to numeric values
-df['Hardware Salamander 4 (us)'] = df['Hardware Salamander 4 (us)'].str.replace(',', '.').astype(float)
-df['Tuned Salamander 4 Virtualization (us)'] = df['Tuned Salamander 4 Virtualization (us)'].str.replace(',', '.').astype(float)
-df['Untuned Salamander 4 Virtualization (us)'] = df['Untuned Salamander 4 Virtualization (us)'].str.replace(',', '.').astype(float)
+df['Hardware Salamander 4 (ms)'] = df['Hardware Salamander 4 (ms)'].str.replace(',', '.').astype(float)
+df['Tuned Salamander 4 Virtualization (ms)'] = df['Tuned Salamander 4 Virtualization (ms)'].str.replace(',', '.').astype(float)
+df['Untuned Salamander 4 Virtualization (ms)'] = df['Untuned Salamander 4 Virtualization (ms)'].str.replace(',', '.').astype(float)
 
 # Calculate statistics
 def calculate_statistics(latencies):
     return {
         'Samples': len(latencies),
-        'Lat Min (us)': f"{np.min(latencies):.3f}".replace('.', ','),
-        'Lat Avg (us)': f"{np.mean(latencies):.3f}".replace('.', ','),
-        'Lat Max (us)': f"{np.max(latencies):.3f}".replace('.', ','),
-        'Std Dev (us)': f"{np.std(latencies):.3f}".replace('.', ',')
+        'Lat Min (ms)': f"{np.min(latencies):.3f}".replace('.', ','),
+        'Lat Avg (ms)': f"{np.mean(latencies):.3f}".replace('.', ','),
+        'Lat Max (ms)': f"{np.max(latencies):.3f}".replace('.', ','),
+        'Std Dev (ms)': f"{np.std(latencies):.3f}".replace('.', ',')
     }
 
-rtos_stats = calculate_statistics(df['Hardware Salamander 4 (us)'])
-tuned_virt_stats = calculate_statistics(df['Tuned Salamander 4 Virtualization (us)'])
-untuned_virt_stats = calculate_statistics(df['Untuned Salamander 4 Virtualization (us)'])
+rtos_stats = calculate_statistics(df['Hardware Salamander 4 (ms)'])
[…]

### 2024-08-28 Masterarbeit_Robotik_Pamuk.tex
Kurzfassung, Abstract, Acknowledgements

@@
 \place{Wien}
-\kurzfassung{Kurzfassung}
-\schlagworte{Echtzeit, Virtualisierung, Xenomai, VARAN}
-\outline{Abstract}
-\keywords{Echtzeit, Virtualisierung, Xenomai, VARAN}
-%\acknowledgements{\blindtext}
+\kurzfassung{Virtualisierung von Echtzeitbetriebssystemen für die Robotersteuerung bietet viele Vorteile im Vergleich zu herkömmlichen hardwarebasierten Ansätzen.
+Lösungen, die auf reiner Hardware basieren, sind oft für einen bestimmten Zweck gebaut und daher in ihrer Flexibilität eingeschränkt.
+Es ist schwierig, die Hardware-Topologie an neue Anforderungen anzupassen, und es kann kostspielig sein, insbesondere bei der Skalierung von Operationen.
+Der physische Zugang für Updates und Wartung ist herausfordernd, was zu Ausfallzeiten und Produktivitätsverlusten führt.
+Während die Virtualisierung diese Probleme adressiert, führt sie zu erhöhtem Overhead und Latenz.
+
+\bigskip \noindent Diese Arbeit untersucht die Virtualisierung des proprietären Salamander 4-Betriebssystems unter Verwendung von QEMU/KVM.
+Salamander 4 wird mit Yocto gebaut und verwendet harte Echtzeit mit Xenomai 3.
+Das Hauptziel besteht darin, die Latenzlücke zwischen den virtualisierten und den Bare-Metal-Versionen zu überbrücken, um ein deterministisches und zuverlässiges Verhalten zu gewährleisten, das für Echtzeit-Roboteranwendungen entscheidend ist.
+
+\bigskip \noindent Erste Latenzmessungen zeigten eine signifikante Leistungslücke zwischen dem Bare-Metal- und dem virtualisierten Setup.
+Daher wird ein umfangreicher Tuning-Prozess durchgeführt, um Echtzeit-Performance und Determinismus zu erreichen.
+Diese Modifikationen umfassen Konfigurationen, die das BIOS, den Kernel, das Host-Betriebssystem, die QEMU/KVM-Virtualisierungsschicht und das Salamander 4-Betriebssystem (Gast) selbst betreffen.
+Die Worst-Case-Latenz wurde von 707.622\textmu s auf 17.134\textmu s reduziert, was der Bare-Metal-Performance von 10.709\textmu s nahekommt.
+Darüber hinaus wird die Verbesserung der Echtzeit-Performance und des Determinismus anhand einer Roboteranwendung validiert, bei der die abgestimmte Virtualisierung mit der nicht abgestimmten und der Hardware-Version verglichen wird.
+
+\bigskip \noindent Insgesamt bietet diese Arbeit eine umfassende Blaupause, um ein virtualisiertes Gastsystem in einem Host-System mit deterministischem Verhalten echtzeitfähig zu machen.}
+\schlagworte{Virtualisierung, Echtzeitsysteme, Latenzreduktion, Robotiksteuerung}
+
+\outline{Virtualization of real-time operating systems systems for robotic control has many advantages in comparison to traditional hardware-based approaches.
+Solutions based on pure hardware are often built for a distinct purpose and are hence limited in flexibility.
+It is difficult to adapt the hardware topology to new requirements and can be costly, especially when scaling operations.
+Physical access for updates and maintenance is challenging, leading to downtime and lost productivity.
+While virtualization addresses these issues, it introduces increased overhead and latency.
+
+\bigskip \noindent This thesis investigates the virtualization of the proprietory Salamander 4 operating system, using QEMU/KVM.
+Salamander 4 is built with Yocto and employs hard real-time with Xenomai 3.
+The primary objective is to bridge the latency gap between the virtualized and bare-metal versions in order to ensure deterministic and reliable behavior, which is crucial for real-time robotic applications.
+
+\bigskip \noindent Initial latency measurements revealed a significant performance gap between the bare-metal and virtualized setup.
+Thus, an extensive tuning process is carried out to achieve real-time performance and determinism.
+These modifications involve configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS (guest) itself.
+The worst-case latency was brought down from 707.622\textmu s to 17.134\textmu s, closely aligning with the bare-metal performance of 10.709\textmu s.
+In addition, the improvement in real-time performance and determinism is validated using a robotic application, where the tuned virtualization is compared with the untuned and the hardware version.
+
+\bigskip \noindent Alltogether, this thesis provides a comprehensive blueprint for making a virtualized guest system real-time capable in a host system with deterministic behavior.}
+\keywords{Virtualization, Real-Time Systems, Latency Reduction, Robotic Control}
+
+\acknowledgements{First, I would like to thank my supervisor, Sebastian Felix Rauh MSc.
+BEng., for his expert guidance and assistance throughout this master's thesis.
+His insightful answers to all my questions and his calming presence helped me stay focused and composed.
+
+\bigskip \noindent I want to thank my wife, Nour Pamuk, for her unwavering support, love, and patience. She gave me the strength to overcome this challenge. This work is as much hers as it is mine.
+
+\bigskip \noindent Lastly, I am deeply grateful to my parents, Semra Pamuk and Mehmet Pamuk, and to my close family for their understanding and encouragement during my studies.
+Without their unconditional love and support, this work would not have been possible.}
 

### 2024-08-29 Masterarbeit_Robotik_Pamuk.tex
Bibliography, List of Figures, List of Tables, List of Code, List of Abbreviations

@@
 Salamander 4 is built with Yocto and employs hard real-time with Xenomai 3.
-The primary objective is to bridge the latency gap between the virtualized and bare-metal versions in order to ensure deterministic and reliable behavior, which is crucial for real-time robotic applications.
-
-\bigskip \noindent Initial latency measurements revealed a significant performance gap between the bare-metal and virtualized setup.
+The primary objective is to bridge the latency gap between the virtualized and bare metal versions in order to ensure deterministic and reliable behavior, which is crucial for real-time robotic applications.
+
+\bigskip \noindent Initial latency measurements revealed a significant performance gap between the bare metal and virtualized setup.
 Thus, an extensive tuning process is carried out to achieve real-time performance and determinism.
-These modifications involve configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS (guest) itself.
-The worst-case latency was brought down from 707.622\textmu s to 17.134\textmu s, closely aligning with the bare-metal performance of 10.709\textmu s.
+These modifications involve configurations spanning the BIOS, kernel, host operating system, QEMU/KVM virtualization layer, and the Salamander 4 operating system (guest) itself.
+The worst-case latency was brought down from 707.622\textmu s to 17.134\textmu s, closely aligning with the bare metal performance of 10.709\textmu s.
 In addition, the improvement in real-time performance and determinism is validated using a robotic application, where the tuned virtualization is compared with the untuned and the hardware version.
@@
 	\includegraphics[width=0.75\columnwidth]{img/introduction/kernel_generic.png}
-	\caption[Non-preemptible kernel]{Non-preemptible kernel~\cite{WhatRealtimeLinuxa}}
+	\caption[Non-preemptible Kernel]{Non-preemptible Kernel~\cite{WhatRealtimeLinux}}
 	\label{fig:kernel_generic}
@@
 	\includegraphics[width=0.75\columnwidth]{img/introduction/kernel_rt.png}
-	\caption[Non-preemptible kernel]{Non-preemptible kernel~\cite{WhatRealtimeLinuxa}}
+	\caption[Preemptible Kernel]{Preemptible Kernel~\cite{WhatRealtimeLinux}}
 	\label{fig:kernel_rt}
@@
 
-\bigskip \noindent Real-time systems are usually classified as either soft or hard real-time systems~\cite{lipariRealTimeSchedulingHard}.
-The difference lies exclusively in the consequences of a violation of the time limits~\cite{HardRealTime}.
+\bigskip \noindent Real-time systems are usually classified as either soft or hard real-time systems~\cite{lipariRealTimeSchedulingHard2015}.
+The difference lies exclusively in the consequences of a violation of the time limits~\cite{amarpreetHardRealTime2009}.
 Hard real-time is when the system stops operating if a deadline is missed, which can have catastrophic consequences.
@@
 
-\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased}~\cite{maPerformanceTuningKVMbased}~\cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
+\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased2013}~\cite{maPerformanceTuningKVMbased2013}~\cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
 When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS.
@@
 	\end{tikzpicture}
-	\caption[Flowchart of operations in a virtualized environment]{Flowchart of operations in a virtualized environment (own figure, inspired by~\citeauthor{maPerformanceTuningKVMbased})}
+	\caption[Flowchart of Operations in a virtualized Environment]{Flowchart of Operations in a virtualized Environment (Own Figure, inspired by~\citeauthor{maPerformanceTuningKVMbased2013})}
 	\label{fig:kvm_stages}
@@
 
-\bigskip \noindent \citeauthor{yoonRealTimePerformanceAnalysis}~\cite{yoonRealTimePerformanceAnalysis} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
+\bigskip \noindent \citeauthor{yoonRealTimePerformanceAnalysis2009}~\cite{yoonRealTimePerformanceAnalysis2009} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
 They implement EtherCAT for real-time communication of distributed devices whereas this thesis uses Varan.
@@
 
-\bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor}~\cite{kiszkaLinuxRealTimeHypervisor} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
+\bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor2009}~\cite{kiszkaLinuxRealTimeHypervisor2009} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
 Most importantly for this master's thesis, the author warns of starvation when the priorities of QEMU’s threads are raised and lifted into a real-time scheduling class.
 Starvation is when a process does not get the resources it needs for a long time because the resources are being allocated to other processes.
-A way to reduce this risk is by not using more virtual CPUs (VCPUs) than there are actual processor cores.
+A way to reduce this risk is by not using more virtual CPUs (vCPUs) than there are actual processor cores.
 Moreover, Linux has a feature that can limit how much CPU time all real-time tasks can use in the host system, which was an important information for the real-time tunings of this thesis later.
@@
 
-\bigskip \noindent \citeauthor{mckenneyRealTimeVs}~\cite{mckenneyRealTimeVs} addresses the overheads associated with real-time Linux when using KVM and QEMU.
+\bigskip \noindent \citeauthor{mckenneyRealTimeVs2008}~\cite{mckenneyRealTimeVs2008} addresses the overheads associated with real-time Linux when using KVM and QEMU.
 The author lists several sources of overhead, including memory locking to avoid page-fault latencies, increased overhead of locking and interrupts due to more-aggressive preemption, threaded interrupts that permit long-running interrupt handlers to be preempted by high-priority real-time processes, real-time task scheduling that requires global scheduling, high-resolution timers for tens-of-microseconds accuracy and precision, and preemptible RCU which has slightly higher read-side overhead than Classic RCU.
@@
 
-\bigskip \noindent \citeauthor{deoliveiraDemystifyingRealTimeLinux}~\cite{deoliveiraDemystifyingRealTimeLinux} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
+\bigskip \noindent \citeauthor{deoliveiraDemystifyingRealTimeLinux2020}~\cite{deoliveiraDemystifyingRealTimeLinux2020} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
 This is the main reason why this tool was not used to measure the scheduling latency in this work.
@@
 	\centering
-	\caption[Host operating system configuration]{Host operating system configuration}
+	\caption[Host Operating System Configuration]{Host Operating System Configuration}
 	\label{tab:testbed_configuration}
@@
 	\includegraphics[width=0.55\columnwidth]{img/lstopo.png}
-	\caption[Hardware topology]{Hardware topology}
+	\caption[Host Operating System Hardware Topology]{Host Operating System Hardware Topology}
 	\label{fig:lstopo}
@@
 \vspace{2em}
-\begin{lstlisting}[name={System information},label={output:uname_a_virt}]
+\begin{lstlisting}[name={System Information},label={output:uname_a_virt}]
 root@sigmatek-core2:~# uname -a
@@
 	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
-	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU}
+	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU~\cite{LASALOSSIGMATEK}}
 	\label{fig:lasal_cpu}
@@
 	\includegraphics[width=0.8\columnwidth]{img/RAM_Memory_management.png}
-	\caption[Memory Management]{Memory Management}
+	\caption[Memory Management of Salamander 4]{Memory Management of Salamander 4~\cite{LASALOSSIGMATEK}}
 	\label{fig:memory_management}
@@
 \vspace{2em}
-\begin{lstlisting}[name={Contents of QEMU folder for Salamander 4},label={lst:ls}]
+\begin{lstlisting}[name={Contents of QEMU Folder for Salamander 4},label={lst:ls}]
     sigma_ibo@localhost:~/Desktop/salamander-image$ ls -1
@@
 \vspace{2em}
-\begin{lstlisting}[name={QEMU script for starting Salamander 4 virtualization},label={script:qemu_def}]
+\begin{lstlisting}[name={QEMU script for starting Salamander 4 Virtualization},label={script:qemu_def}]
 	#!/bin/sh
@@
 	\includegraphics[width=0.6\columnwidth]{img/introduction/xenomai/x3-cobalt-interfaces.png}
-	\caption[Xenomai Cobalt interfaces]{Xenomai Cobalt interfaces~\cite{XenomaiXenomai}}
+	\caption[Xenomai 3 Cobalt Interfaces]{Xenomai 3 Cobalt Interfaces~\cite{OverviewXenomai}}
 	\label{fig:cobalt}
@@
 	\includegraphics[width=0.6\columnwidth]{img/introduction/xenomai/x3-mercury-interfaces.png}
-	\caption[Xenomai Mercury interfaces]{Xenomai Mercury interfaces~\cite{XenomaiXenomai}}
+	\caption[Xenomai 3 Mercury Interfaces]{Xenomai 3 Mercury Interfaces~\cite{OverviewXenomai}}
 	\label{fig:mercury}
@@
 \vspace{2em}
-\begin{lstlisting}[language=bash, caption={Kernel Configuration for vsocks and tracing}, label={lst:kernel-config}]
+\begin{lstlisting}[language=bash, caption={Kernel Flags for Vsocks and Tracing}, label={lst:kernel-config}]
 	# Vsockets settings
@@
 Once both files are loaded, the trace data can be visually inspected.
-All the current CPUs for both the host and the guest can be turned off, and the KVM Combo plots menu can be selected to view the guest VCPUs plotted on top of the host threads that run them, as depicted in Figure~\ref{fig:kernelshark_combo}.
+All the current CPUs for both the host and the guest can be turned off, and the KVM Combo plots menu can be selected to view the guest vCPUs plotted on top of the host threads that run them, as depicted in Figure~\ref{fig:kernelshark_combo}.
 
@@
 	\includegraphics[width=1.0\columnwidth]{img/kernelshark_combo.png}
-	\caption[Guest VCPUs plotted on top of host threads in KernelShark]{Guest VCPUs plotted on top of host threads in KernelShark}
+	\caption[Guest vCPUs plotted on top of Host Threads in KernelShark]{Guest vCPUs plotted on top of Host Threads in KernelShark}
 	\label{fig:kernelshark_combo}
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
-	\caption[Distribution of Latency of Salamander 4 Bare Metal]{Distribution of Latency of Salamander 4 Bare Metal}
+	\caption[Latency Distribution of Salamander 4 Bare Metal]{Latency Distribution of Salamander 4 Bare Metal}
 	\label{fig:gnuplot_max_latency_hardware}
@@
 The test was conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
-If any sample’s latency value is greater than 100\textmu s, it would be considered an overrun.
-
-\begin{table}[H]
-	\centering
-	\caption{Statistical Summary of Latency Measurements}
+If any sample's latency value is greater than 100\textmu s, it would be considered an overrun.
+
+\begin{table}[H]
+	\centering
+	\caption{Latency Parameters for Bare Metal}
 	\label{tab:latency_stats_hardware}
@@
 
-\vspace{1em}
-\begin{table}[H]
-	\centering
-	\caption{Latencies and Overruns}
+\begin{table}[H]
+	\centering
+	\caption{Minimum, Average, and Maximum Latency with Overrun Counts of Bare Metal}
 	\label{tab:latency_overrun_msw_hardware}
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
-	\caption[Distribution of Latency of Salamander 4 unmodified Virtualization]{Distribution of Latency of Salamander 4 unmodified Virtualization}
+	\caption[Latency Distribution of Salamander 4 Untuned Virtualization]{Latency Distribution of Salamander 4 Untuned Virtualization}
 	\label{fig:gnuplot_max_latency_default}
@@
 	\centering
-	\caption{Statistical Summary of Latency Measurements}
+	\caption{Latency Parameters for Untuned Virtualization}
 	\label{tab:latency_stats_virt}
@@
 	\centering
-	\caption{Latencies and Overruns}
+	\caption[Latency Statistics with Overrun Counts for Untuned Virtualization]{Minimum, Average, and Maximum Latency with Overrun Counts for Untuned Virtualization}
 	\label{tab:latency_overrun_msw_new_virt}
@@
 The patch also includes support for high-resolution timers and makes the system more predictable~\cite{lutsykPipelinedMulticoreMachine2020}.
-Determinism and predictability are key for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling~\cite{rostedtInternalsRTPatch}.
-
-\bigskip \noindent Nevertheless, a real-time kernel alone does not make a system truly “real-time”~\cite{WhatRealtimeLinux}.
+Determinism and predictability are key for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling~\cite{rostedtInternalsRTPatch2007}.
+
+\bigskip \noindent Nevertheless, a real-time kernel alone does not make a system truly “real-time”~\cite{WhatRealtimeLinuxa}.
 Additional modifications are required to achieve this.
@@
 	\centering
-	\caption{BIOS Configurations}
+	\caption{BIOS Configurations for Real-Time Performance}
 	\label{tab:bios_configuration}
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
-	\caption[Latency Distribution after tuning BIOS Settings]{Latency Distribution after tuning BIOS Settings}
+	\caption[Latency Distribution of Salamander 4 after BIOS Configurations]{Latency Distribution of Salamander 4 Virtualization after tuning BIOS Configurations}
 	\label{fig:gnuplot_max_latency_taskset}
@@
 	\centering
-	\caption{Statistical Summary of Latency Measurements}
+	\caption{Latency Parameters after BIOS Configurations}
 	\label{tab:latency_stats_virt_bios}
@@
 	\centering
-	\caption{Latencies and Overruns, Missed Switch, Best, and Worst Data}
+	\caption[Latency Statistics with Overrun Counts after BIOS Configurations]{Minimum, Average, and Maximum Latency with Overrun Counts after BIOS Configurations}
 	\label{tab:latency_overrun_msw_new_virt_bios}
@@
 \vspace{2em}
-\begin{lstlisting}[name={Kernel Configuration},label={script:kernel_cli}]
+\begin{lstlisting}[name={Kernel Configurations for Real-Time Performance},label={script:kernel_cli}]
 		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 rcu_nocb_poll nohz_full=4 nohz=on default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/gnuplot_max_latency_rt_kernelparam.png}
-	\caption[Latency Distribution after tuning Kernel Settings]{Latency Distribution after tuning Kernel Settings}
+	\caption[Latency Distribution of Salamander 4 after Kernel Configurations]{Latency Distribution of Salamander 4 Virtualization after tuning Kernel Configurations}
 	\label{fig:gnuplot_max_latency_rt_kernelparam}
@@
 	\centering
-	\caption{Statistical Summary of Latency Measurements}
+	\caption{Latency Parameters after Kernel Configurations}
 	\label{tab:latency_stats_kernelparam}
@@
 	\centering
-	\caption{Latencies and Overruns}
+	\caption[Latency Statistics with Overrun Counts after Kernel Configurations]{Minimum, Average, and Maximum Latency with Overrun Counts after Kernel Configurations}
 	\label{tab:latency_overrun_msw_new_virt_kernel}
@@
 However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
-Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}, including systemd services.
+Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased2013}, including systemd services.
 To prevent systemd services from running on an isolated CPU, the CPU affinity can be set with the line \texttt{CPUAffinity=0 1 2 3 5} in the \texttt{/etc/systemd/system.conf} file to indicate the CPUs that systemd services are allowed to run on.
@@
 \vspace{2em}
-\begin{lstlisting}[name={User and Kernel Tasks},label={output:pse},escapeinside={(*@}{@*)}]
+\begin{lstlisting}[name={User and Kernel Tasks on the isolated CPU},label={output:pse},escapeinside={(*@}{@*)}]
 		sigma_ibo@sigma-ibo:~$ cat /sys/devices/system/cpu/isolated
@@
 \vspace*{-5em}
-\begin{lstlisting}[name={Check distribution of interrupt requests across each CPU},label={script:smp_affinity}]
+\begin{lstlisting}[name={Code to show a table of the distribution of interrupt requests across each CPU},label={script:smp_affinity}]
 		import os
@@
 \vspace*{-5em}
-\begin{lstlisting}[name={Change IRQ assignment of a CPU},label={script:change_smp_affinity}]
+\begin{lstlisting}[name={Script to change IRQ Assignment of a CPU},label={script:change_smp_affinity}]
 		#!/bin/bash
@@
 	\centering
-	\caption{Minimum and maximum priorities for different scheduling policies}
+	\caption{Minimum and maximum Priorities for different Scheduling Policies}
 	\label{tab:scheduling_priorities}
@@
 	\centering
-	\caption{Description of Services}
+	\caption{Services that are stopped to reduce random Latency and Overhead}
 	\label{tab:stop_servies}
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/5rt_kernelparam_host/gnuplot_max_latency_rt_kernelparam_host.png}
-	\caption[Latency rt kernelparam host]{Latency rt kernelparam host}
+	\caption[Latency Distribution of Salamander 4 after Host Configurations]{Latency Distribution of Salamander 4 Virtualization after tuning Host Configurations}
 	\label{fig:max_latency_rt_kernelparam_host}
@@
 	\centering
-	\caption{Statistical Summary of Latency Measurements}
+	\caption{Latency Parameters for Host Configurations}
 	\label{tab:latency_stats_virt_host}
@@
 	\centering
-	\caption{Latencies and Overruns}
+	\caption[Latency Statistics with Overrun Counts after Host Configurations]{Minimum, Average, and Maximum Latency with Overrun Counts after Host Configurations}
 	\label{tab:latency_overrun_msw_new_virt_host}
@@
 	\centering
-	\caption{QEMU options for real-time performance}
+	\caption{QEMU Options for Real-Time Performance}
 	\label{tab:qemu_options}
@@
 
-\begin{lstlisting}[name={Tuned QEMU script for starting Salamander 4 virtualization},label={script:qemu_def_tuned}]
+\begin{lstlisting}[name={Tuned QEMU Script for starting Salamander 4 Virtualization},label={script:qemu_def_tuned}]
 		#!/bin/sh
@@
 	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/6rt_kernelparam_host_qemu/gnuplot_max_latency_rt_kernelparam_host_qemu.png}
-	\caption[Latency rt kernelparam host qemu]{Latency rt kernelparam host qemu}
+	\caption[Latency Distribution of Salamander 4 after QEMU Configurations]{Latency Distribution of Salamander 4 Virtualization after tuning QEMU Configurations}
 	\label{fig:max_latency_rt_kernelparam_host_qemu}
@@
 	\centering
-	\caption{Statistical Summary of Latency Measurements}
+	\caption{Latency Parameters after QEMU Configurations}
 	\label{tab:latency_stats_virt_qemu}
@@
 	\centering
-	\caption{Latencies and Overruns}
+	\caption[Latency Statistics with Overrun Counts after QEMU Configurations]{Minimum, Average, and Maximum Latency with Overrun Counts after QEMU Configurations}
 	\label{tab:latency_overrun_msw_new_virt_qemu}
@@
 	\includegraphics[width=0.3\columnwidth]{img/experiment/mini_robot.jpg}
-	\caption[Mini-robot of the experiment]{Mini-robot of the experiment~\cite{6DFRoboticArm}}
+	\caption[Mini-Robot of the Experiment]{Mini-Robot of the Experiment~\cite{6DFRoboticArm}}
 	\label{fig:mini_robot}
@@
 	\includegraphics[width=0.5\columnwidth]{img/experiment/pw022_connectors.png}
-	\caption[PW 022 pulse width module connector layout]{PW 022 pulse width module connector layout~\cite{DigitalOutputSIGMATEK}}
+	\caption[PW 022 Pulse Width Module Connector Layout]{PW 022 Pulse Width Module Connector Layout~\cite{DigitalOutputSIGMATEK}}
 	\label{fig:pw022_connectors}
@@
 	\includegraphics[width=0.7\columnwidth]{img/experiment/pw022_minirobot.png}
-	\caption[Connection between PW 022 and mini-robot]{Connection between PW 022 and mini-robot~\cite{MG996RDigitalServo}}
+	\caption[Connection between PW 022 and Mini-Robot]{Connection between PW 022 and Mini-Robot~\cite{MG996RDigitalServo}}
 	\label{fig:pw022_minirobot_connection}
@@
 
-\subsection{Setup of Hardware Salamander 4}
+\subsection{Setup of Salamander 4 Bare Metal}
 In this version, Salamander 4 runs on the CP 841~\cite{CPUUnitsSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
@@
 	\includegraphics[width=0.45\columnwidth]{img/experiment/hardware_tree.png}
-	\caption[Setup of Hardware Salamander 4]{Setup of Hardware Salamander 4}
+	\caption[Setup of Salamander 4 Bare Metal]{Setup of Salamander 4 Bare Metal}
 	\label{fig:hardware_tree}
@@
 
-\subsection{Setup of QEMU Salamander 4}\label{subsec:setup_experiment_virtualized}
-In the virtualized setup of the experiment, the VCPU functionality of QEMU is used to connect QEMU with the PWM modules.
+\subsection{Setup of Salamander 4 Virtualization}\label{subsec:setup_experiment_virtualized}
+In the virtualized setup of the experiment, the vCPU functionality of QEMU is used to connect QEMU with the PWM modules.
 In order to achieve that, the PCV 522 VARAN Manager PCI Insert Card~\cite{ControlsHMIsSIGMATEK}, which serves as a bridge between the PC and the rest of the setup, needs to be plugged into the PC.
@@
 \vspace*{-5em}
-\begin{lstlisting}[name={Include PCI in QEMU for Salamander 4 Virtualization},label={script:qemu_def_pci}]
+\begin{lstlisting}[name={QEMU script for Salamander 4 Virtualization with PCI Configuration},label={script:qemu_def_pci}]
 		#!/bin/sh
@@
 	\includegraphics[width=0.5\columnwidth]{img/experiment/virt_tree.png}
-	\caption[Virtualization tree]{Virtualization tree}
+	\caption[Setup of Salamander 4 Virtualization]{Setup of Salamander 4 Virtualization}
 	\label{fig:virt_tree}
@@
 
-\noindent Essentially, the CPU is being emulated by the VCPU functionality of QEMU, allowing the PCV 522 VARAN Manager PCI Insert Card and the Varan Connection module VI 021 to interact as if they were communicating with a physical CPU unit.
+\noindent Essentially, the CPU is being emulated by the vCPU functionality of QEMU, allowing the PCV 522 VARAN Manager PCI Insert Card and the Varan Connection module VI 021 to interact as if they were communicating with a physical CPU unit.
 
@@
 \end{figure}
+\vspace{1em}
 
@@
 
-\subsubsection{Latency in Salamander 4 OS Hardware}\label{subsubsec:latency_hardware}
+\subsubsection{Latency in Salamander 4 Bare Metal}\label{subsubsec:latency_hardware}
 \noindent The latency values for bare metal Salamander 4 OS aquired from the program above are demonstrated in Figure~\ref{fig:rtos_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_hardware}.
@@
 	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/rtos_latencies.png}
-	\caption[Virtualization tree]{Virtualization tree}
+	\caption[Robotic Application Latency Values of Salamander 4 Hardware]{Robotic Application Latency Values of Salamander 4 Hardware}
 	\label{fig:rtos_latencies}
@@
 	\centering
-	\caption{Robotic Application Latency Values Hardware}
+	\caption{Robotic Application Latency Statistics Hardware}
 	\label{tab:robotic_application_latency_values_hardware}
@@
 	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/untuned_virt_latencies.png}
-	\caption[Virtualization tree]{Virtualization tree}
+	\caption[Robotic Application Latency Values of Salamander 4 Untuned Virtualization]{Robotic Application Latency Values of Salamander 4 Untuned Virtualization}
 	\label{fig:untuned_virt_latencies}
@@
 	\centering
-	\caption{Robotic Application Latency Values Untuned Virtualization}
+	\caption{Robotic Application Latency Statistics Untuned Virtualization}
 	\label{tab:robotic_application_latency_values_virt_def}
@@
 	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/tuned_virt_latencies.png}
-	\caption[Virtualization tree]{Virtualization tree}
+	\caption[Robotic Application Latency Values of Salamander 4 Tuned Virtualization]{Robotic Application Latency Values of Salamander 4 Tuned Virtualization}
 	\label{fig:tuned_virt_latencies}
@@
 	\centering
-	\caption{Robotic Application Latency Values Tuned Virtualization}
+	\caption{Robotic Application Latency Statistics Tuned Virtualization}
 	\label{tab:robotic_application_latency_values_virt_tun}
@@
 
-The Figure below (deleted) compares non-optimized guest latency with optimized guest latency and includes optimized bare-metal latency as a reference.
+The Figure below (deleted) compares non-optimized guest latency with optimized guest latency and includes optimized bare metal latency as a reference.
 The data shows that a 40\% reduction in QD1 latency is achievable through system tuning.
@@
 	\includegraphics[width=0.70\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/012456combined/gnuplot_combined_max_latency.png}
-	\caption[Latency rt kernelparam host]{Latency rt kernelparam host}
+	\caption[Comparison of Latency Distribution of Salamander 4 Configurations]{Comparison of Latency Distribution of Salamander 4 under different Configurations}
 	\label{fig:max_latency_combined_results}
@@
 	\footnotesize
-	\caption{Latencies and Overruns with Tuning}
+	\caption{Comparison of Latency Statistics in Salamander 4 under different Configurations}
 	\label{tab:latency_tables_combined}
@@
 	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/combined_latencies.png}
-	\caption[Robotic Application Response Latency]{Robotic Application Response Latency}
+	\caption[Comparison of Robotic Application Latency of Salamander 4 Configurations]{Comparison of Robotic Application Latency of Salamander 4 under different Configurations}
 	\label{fig:combined_latencies}
@@
 	\centering
-	\caption{Robotic Application Latency Values}
+	\caption{Comparison of Robotic Application Latency Statistics}
 	\label{tab:robotic_application_latency_values_combined}
@@
 \begin{comment}
-Online
-
-~\cite{WhatRealtimeLinuxa}
+~\cite{canbazPerformanceAnalysisRealtime2022}
+~\cite{WhatRealtimeLinux}
+~\cite{malallahComprehensiveStudyKernel2021}
+~\cite{wangRealtimeEmbeddedSystems2017}
+~\cite{buttazzoHardRealtimeComputing2024}
+~\cite{lipariRealTimeSchedulingHard2015}
+~\cite{amarpreetHardRealTime2009}
+~\cite{queirozTestingLimitsGeneralpurpose2023}
 ~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
@@
 ~\cite{XenomaiXenomai}
+~\cite{perneelRealtimeCapabilitiesStandard2015}
+~\cite{reghenzaniRealTimeLinuxKernel2020}
+~\cite{adamRealTimePerformanceResponse2021}
+~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
+~\cite{sandstromVirtualizationTechnologiesEmbedded2013}
+~\cite{taccariEmbeddedRealTimeVirtualization2014}
+~\cite{javierperezHowRealTime2022}
+~\cite{guStateoftheArtSurveyRealTime2012}
+~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
+~\cite{scordinoRealTimeVirtualizationIndustrial2020}
+~\cite{maPerformanceTuningKVMbased2013}
+~\cite{junzhangPerformanceAnalysisKVMBased2010}
+~\cite{broskyShieldedProcessorsGuaranteeing2003}
+~\cite{RealTimePerformanceTuning2022}
+~\cite{yoonRealTimePerformanceAnalysis2009}
+~\cite{kiszkaLinuxRealTimeHypervisor2009}
+~\cite{mckenneyRealTimeVs2008}
+~\cite{deoliveiraDemystifyingRealTimeLinux2020}
+~\cite{gabrielResourceAwareParameterTuning2019}
+~\cite{bhardwajVirtualizationCloudComputing2021}
+~\cite{abbasiExploringOpenStackScalable2023}
+~\cite{mahfoudRealTimePredictiveMaintenanceBased2024}
+~\cite{queirozContainerbasedVirtualizationRealtime2024}
+~\cite{cinqueEvaluatingVirtualizationFog2023}
 ~\cite{casiniLatencyAnalysisVirtualization2021}
@@
 ~\cite{EngineeringToolLASAL}
+~\cite{LASALOSSIGMATEK}
 ~\cite{QEMU}
+~\cite{OverviewXenomai}
 ~\cite{RealtimePreempt_rt_versionsWiki}
@@
 ~\cite{Tracecmd}
+~\cite{rostedtTracecmdHostGuest}
 ~\cite{KernelShark}
-~\cite{rostedtTracecmdHostGuest}
 ~\cite{RealtimeKernelPatchset}
-~\cite{rostedtInternalsRTPatch}
-~\cite{WhatRealtimeLinux}
+~\cite{lutsykPipelinedMulticoreMachine2020}
+~\cite{rostedtInternalsRTPatch2007}
+~\cite{WhatRealtimeLinuxa}
 ~\cite{LinuxProcessPriorities}
@@
 ~\cite{InterfacesSplittersSIGMATEK}
-
-\noindent Papers
-
-~\cite{canbazPerformanceAnalysisRealtime2022}
-~\cite{malallahComprehensiveStudyKernel2021}
-~\cite{wangRealtimeEmbeddedSystems2017}
-~\cite{buttazzoHardRealtimeComputing2024}
-~\cite{lipariRealTimeSchedulingHard}
-~\cite{perneelRealtimeCapabilitiesStandard2015}
-~\cite{reghenzaniRealTimeLinuxKernel2020}
-~\cite{adamRealTimePerformanceResponse2021}
-~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
-~\cite{sandstromVirtualizationTechnologiesEmbedded2013}
-~\cite{taccariEmbeddedRealTimeVirtualization2014}
-~\cite{javierperezHowRealTime2022}
-~\cite{guStateoftheArtSurveyRealTime2012}
-~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
-~\cite{scordinoRealTimeVirtualizationIndustrial2020}
-~\cite{maPerformanceTuningKVMbased}
-~\cite{junzhangPerformanceAnalysisKVMBased2010}
-~\cite{broskyShieldedProcessorsGuaranteeing2003}
-~\cite{RealTimePerformanceTuning2022}
-~\cite{yoonRealTimePerformanceAnalysis}
-~\cite{kiszkaLinuxRealTimeHypervisor}
-~\cite{mckenneyRealTimeVs}
-~\cite{deoliveiraDemystifyingRealTimeLinux}
-~\cite{gabrielResourceAwareParameterTuning2019}
-~\cite{bhardwajVirtualizationCloudComputing2021}
-~\cite{abbasiExploringOpenStackScalable2023}
-~\cite{mahfoudRealTimePredictiveMaintenanceBased2024}
-~\cite{queirozContainerbasedVirtualizationRealtime2024}
-~\cite{cinqueEvaluatingVirtualizationFog2023}
 ~\cite{huang2015performance} 
-~\cite{lutsykPipelinedMulticoreMachine2020} 
-~\cite{queirozTestingLimitsGeneralpurpose2023}
 ~\cite{kirovaImpactModernVirtualization2019}
-~\cite{HardRealTime}
 ~\cite{adamPerformanceAssessmentLinux2021}
-
 \end{comment}
@@
 \begin{acronym}[XXXXX]
+	\acro{AI}[AI]{Artificial Intelligence}
+	\acro{CID}[CID]{Context Identifier}
 	\acro{CPU}[CPU]{Central Processing Unit}
+	\acro{GPOS}[GPOS]{General Purpose Operating System}
+	\acro{GUI}[GUI]{Graphical User Interface}
+	\acro{IRQ}[IRQ]{Interrupt Request}
+	\acro{IRT}[IRT]{Interrupt Response Time}
+	\acro{ISR}[ISR]{Interrupt Service Routine}
+	\acro{KVM}[KVM]{Kernel-based Virtual Machine}
+	\acro{LAPIC}[LAPIC]{Local Advanced Programmable Interrupt Controller}
+	\acro{LSB}[LSB]{Least Significant Bit}
+	\acro{ML}[ML]{Machine Learning}
+	\acro{MSB}[MSB]{Most Significant Bit}
+	\acro{PLC}[PLC]{Programmable Logic Controller}
+	\acro{PWM}[PWM]{Pulse Width Modulation}
 	\acro{QEMU}[QEMU]{Quick Emulator}
-	\acro{IRQ}[IRQ]{Interrupt Request}
+	\acro{RTOS}[RTOS]{Real-Time Operating System}
+	\acro{SMP}[SMP]{Symmetric Multiprocessing}
+	\acro{vCPU}[vCPU]{Virtual Central Processing Unit}
+	\acro{VM}[VM]{Virtual Machine}
+	\acro{VMCS}[VMCS]{Virtual Machine Control Structure}
+	\acro{vsock}[vsock]{Virtual Socket}
 \end{acronym}

### 2024-08-29 search_abbr.py

@@ -1,0 +1,23 @@
+import re
+
+# Pfad zu deinem .tex-Dokument
+file_path = 'Masterarbeit_Robotik_Pamuk.tex'
+
+# Regex-Muster zum Suchen von Wörtern in Klammern
+pattern = r'\(\b\w+\b\)'
+
+# Set zum Speichern der gefundenen Wörter (Sets speichern nur eindeutige Elemente)
+words_in_brackets = set()
+
+# Datei öffnen und durchsuchen
+with open(file_path, 'r', encoding='utf-8') as file:
+    for line in file:
+        matches = re.findall(pattern, line)
+        words_in_brackets.update(matches)
+
+# Gefundene Wörter untereinander aufschreiben
+with open('found_abbr.txt', 'w', encoding='utf-8') as output_file:
+    for word in sorted(words_in_brackets):
+        output_file.write(word + '\n')
+
+print("Words in brackets saved to 'found_abbr.txt' gespeichert.")

### 2024-08-29 Masterarbeit_Robotik_Pamuk.tex
organize

@@
 %
-\title{Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die
-Einhaltung der Echtzeit}
+\title{Virtualization of a Real-Time Operating System for Robot Control with a Focus on Real-Time Compliance}
 \author{Halil Pamuk, BSc}
@@
 
-\bigskip \noindent This thesis investigates the virtualization of the proprietory Salamander 4 operating system, using QEMU/KVM.
+\bigskip \noindent This thesis investigates the virtualization of the proprietary Salamander 4 operating system, using QEMU/KVM.
 Salamander 4 is built with Yocto and employs hard real-time with Xenomai 3.
@@
 
-\bigskip \noindent Alltogether, this thesis provides a comprehensive blueprint for making a virtualized guest system real-time capable in a host system with deterministic behavior.}
+\bigskip \noindent Altogether, this thesis provides a comprehensive blueprint for making a virtualized guest system real-time capable in a host system with deterministic behavior.}
 \keywords{Virtualization, Real-Time Systems, Latency Reduction, Robotic Control}
@@
 	\centering
-	\includegraphics[width=0.55\columnwidth]{img/lstopo.png}
+	\includegraphics[width=0.55\columnwidth]{img/methodology/lstopo.png}
 	\caption[Host Operating System Hardware Topology]{Host Operating System Hardware Topology}
@@
 	\centering
-	\includegraphics[width=0.5\columnwidth]{img/Software-Struktur_einer_LASAL_CPU.png}
+	\includegraphics[width=0.5\columnwidth]{img/methodology/Software-Struktur_einer_LASAL_CPU.png}
 	\caption[Structure of Salamander 4 CPU]{Structure of Salamander 4 CPU~\cite{LASALOSSIGMATEK}}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{img/RAM_Memory_management.png}
+	\includegraphics[width=0.8\columnwidth]{img/methodology/RAM_Memory_management.png}
 	\caption[Memory Management of Salamander 4]{Memory Management of Salamander 4~\cite{LASALOSSIGMATEK}}
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{img/kernelshark_combo.png}
+	\includegraphics[width=1.0\columnwidth]{img/methodology/kernelshark_combo.png}
 	\caption[Guest vCPUs plotted on top of Host Threads in KernelShark]{Guest vCPUs plotted on top of Host Threads in KernelShark}
@@
 	\centering
-	\includegraphics[width=0.75\columnwidth]{img/multiple_varans.png}
+	\includegraphics[width=0.75\columnwidth]{img/methodology/multiple_varans.png}
 	\caption[Synchronization between Systems and their Data Exchange]{Synchronization between Systems and their Data Exchange}
@@
 	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
+	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_max_latency_hardware.png}
 	\caption[Latency Distribution of Salamander 4 Bare Metal]{Latency Distribution of Salamander 4 Bare Metal}
@@
 	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
+	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_max_latency_default.png}
 	\caption[Latency Distribution of Salamander 4 Untuned Virtualization]{Latency Distribution of Salamander 4 Untuned Virtualization}
@@
 	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/01combined/gnuplot_combined_max_latency.png}
+	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_combined_max_latency.png}
 	\caption[Initial Comparison of Latency Distribution between Hardware and Virtualization]{Initial Comparison of Latency Distribution between Hardware and Virtualization}
@@
 	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
+	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_max_latency_taskset.png}
 	\caption[Latency Distribution of Salamander 4 after BIOS Configurations]{Latency Distribution of Salamander 4 Virtualization after tuning BIOS Configurations}
@@
 	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/gnuplot_max_latency_rt_kernelparam.png}
+	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_max_latency_rt_kernelparam.png}
 	\caption[Latency Distribution of Salamander 4 after Kernel Configurations]{Latency Distribution of Salamander 4 Virtualization after tuning Kernel Configurations}
@@
 	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/5rt_kernelparam_host/gnuplot_max_latency_rt_kernelparam_host.png}
+	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_max_latency_rt_kernelparam_host.png}
 	\caption[Latency Distribution of Salamander 4 after Host Configurations]{Latency Distribution of Salamander 4 Virtualization after tuning Host Configurations}
@@
 	\centering
-	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/6rt_kernelparam_host_qemu/gnuplot_max_latency_rt_kernelparam_host_qemu.png}
+	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_max_latency_rt_kernelparam_host_qemu.png}
 	\caption[Latency Distribution of Salamander 4 after QEMU Configurations]{Latency Distribution of Salamander 4 Virtualization after tuning QEMU Configurations}
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/rtos_latencies.png}
+	\includegraphics[width=1.0\columnwidth]{img/implementation/rtos_latencies.png}
 	\caption[Robotic Application Latency Values of Salamander 4 Hardware]{Robotic Application Latency Values of Salamander 4 Hardware}
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/untuned_virt_latencies.png}
+	\includegraphics[width=1.0\columnwidth]{img/implementation/untuned_virt_latencies.png}
 	\caption[Robotic Application Latency Values of Salamander 4 Untuned Virtualization]{Robotic Application Latency Values of Salamander 4 Untuned Virtualization}
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/tuned_virt_latencies.png}
+	\includegraphics[width=1.0\columnwidth]{img/implementation/tuned_virt_latencies.png}
 	\caption[Robotic Application Latency Values of Salamander 4 Tuned Virtualization]{Robotic Application Latency Values of Salamander 4 Tuned Virtualization}
@@
 	\centering
-	\includegraphics[width=0.70\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/012456combined/gnuplot_combined_max_latency.png}
+	\includegraphics[width=0.70\columnwidth]{img/results/gnuplot_combined_max_latency_all.png}
 	\caption[Comparison of Latency Distribution of Salamander 4 Configurations]{Comparison of Latency Distribution of Salamander 4 under different Configurations}
@@
 	\centering
-	\includegraphics[width=1.0\columnwidth]{masterthesis-documentation/docs/resources/scripts/experiment/generated_latencies/combined_latencies.png}
+	\includegraphics[width=1.0\columnwidth]{img/results/combined_latencies.png}
 	\caption[Comparison of Robotic Application Latency of Salamander 4 Configurations]{Comparison of Robotic Application Latency of Salamander 4 under different Configurations}

### 2024-08-29 clean/clean.sh

@@ -1,1 +1,0 @@
-find ../ -maxdepth 1 -type f \( -name "*.aux" -o -name "*.bbl" -o -name "*.blg" -o -name "*.idx" -o -name "*.ind" -o -name "*.lof" -o -name "*.lot" -o -name "*.out" -o -name "*.toc" -o -name "*.acn" -o -name "*.acr" -o -name "*.alg" -o -name "*.glg" -o -name "*.glo" -o -name "*.gls" -o -name "*.ist" -o -name "*.fls" -o -name "*.log" -o -name "*.fdb_latexmk" -o -name "*.lol" -o -name "*.synctex.gz" -o -name "*.bcf" -o -name "*.run.xml" \) -exec rm -f {} \;

### 2024-08-29 python/remove_urldate.py

@@ -1,0 +1,18 @@
+import re
+
+def process_bib_file(input_file, output_file):
+    with open(input_file, 'r') as file:
+        lines = file.readlines()
+
+    with open(output_file, 'w') as file:
+        entry_type = None
+        for line in lines:
+            if line.startswith('@'):
+                entry_type = line.split('{')[0].strip()
+            if entry_type != '@misc' and 'urldate' in line:
+                continue
+            file.write(line)
+
+input_file = '../Literature.bib'
+output_file = '../Literature.bib'
+process_bib_file(input_file, output_file)

### 2024-08-29 python/search_abbr.py

@@ -1,0 +1,23 @@
+import re
+
+# Pfad zu deinem .tex-Dokument
+file_path = '../Masterarbeit_Robotik_Pamuk.tex'
+
+# Regex-Muster zum Suchen von Wörtern in Klammern
+pattern = r'\(\b\w+\b\)'
+
+# Set zum Speichern der gefundenen Wörter (Sets speichern nur eindeutige Elemente)
+words_in_brackets = set()
+
+# Datei öffnen und durchsuchen
+with open(file_path, 'r', encoding='utf-8') as file:
+    for line in file:
+        matches = re.findall(pattern, line)
+        words_in_brackets.update(matches)
+
+# Gefundene Wörter untereinander aufschreiben
+with open('found_abbr.txt', 'w', encoding='utf-8') as output_file:
+    for word in sorted(words_in_brackets):
+        output_file.write(word + '\n')
+
+print("Words in brackets saved to 'found_abbr.txt' gespeichert.")

### 2024-08-29 remove_urldate.py

@@ -1,18 +1,0 @@
-import re
-
-def process_bib_file(input_file, output_file):
-    with open(input_file, 'r') as file:
-        lines = file.readlines()
-
-    with open(output_file, 'w') as file:
-        entry_type = None
-        for line in lines:
-            if line.startswith('@'):
-                entry_type = line.split('{')[0].strip()
-            if entry_type != '@misc' and 'urldate' in line:
-                continue
-            file.write(line)
-
-input_file = 'Literature.bib'
-output_file = 'Literature.bib'
-process_bib_file(input_file, output_file)

### 2024-08-29 search_abbr.py

@@ -1,23 +1,0 @@
-import re
-
-# Pfad zu deinem .tex-Dokument
-file_path = 'Masterarbeit_Robotik_Pamuk.tex'
-
-# Regex-Muster zum Suchen von Wörtern in Klammern
-pattern = r'\(\b\w+\b\)'
-
-# Set zum Speichern der gefundenen Wörter (Sets speichern nur eindeutige Elemente)
-words_in_brackets = set()
-
-# Datei öffnen und durchsuchen
-with open(file_path, 'r', encoding='utf-8') as file:
-    for line in file:
-        matches = re.findall(pattern, line)
-        words_in_brackets.update(matches)
-
-# Gefundene Wörter untereinander aufschreiben
-with open('found_abbr.txt', 'w', encoding='utf-8') as output_file:
-    for word in sorted(words_in_brackets):
-        output_file.write(word + '\n')
-
-print("Words in brackets saved to 'found_abbr.txt' gespeichert.")

### 2024-08-29 resources/scripts/history_once.py
Deployed 9acc5479 with MkDocs version: 1.5.3

@@ -1,24 +1,0 @@
-def remove_duplicates(file_path):
-    # Open the file in read mode and read all lines
-    with open(file_path, 'r') as file:
-        lines = file.readlines()
-
-    # Reverse the lines to keep the last occurrence and remove previous duplicates
-    lines.reverse()
-
-    # Remove duplicate lines
-    unique_lines = list(dict.fromkeys(lines))
-
-    # Reverse the lines back to original order
-    unique_lines.reverse()
-
-    # Open the file in write mode and write the unique lines
-    with open(file_path, 'w') as file:
-        file.writelines(unique_lines)
-
-# Call the function with the path to your file
-remove_duplicates('.bash_history')
-
-# Print a success message
-print("Duplicates in .bash_history removed.")
-

### 2024-08-29 Paper.tex
bind paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ------------------------------------------------------------------------
% LaTeX FHPaper Template by Thomas MIGLINCI
% ------------------------------------------------------------------------
%
% Das eigentliche Paper beginnt ab Zeile 124 - dort die Daten in den
% Aufruf von FHInfo eintragen.
%
% Bitte ersetzen Sie gegebenenfalls Master-Studiengang Mechatronik/Robotik durch Bachelor-.....
%
% Erstellt von T. Miglinci, Jänner 2012 und getestet von W. Kubinger, März 2012 und September 2012
%
% Updates:
%  -) 30.10.2013, WK:
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[10pt,a4paper,twoside]{article}

\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

% mathematische Symbole
\usepackage{amsmath,amssymb,amsfonts,amstext}
%damit sind Bilder gezielter zu plazieren
\usepackage{float}

%%% ----------------------------------------------------------------------
\usepackage{color}
% Die Corporate Farben sind Blau (RGB 0/134/203), Grün (RGB 0/132/98) und Grau (RGB 98/107/113)
\definecolor{fhblue}{RGB}{0,134,203}
\newcommand\FHblue{\textcolor{fhblue}}
\definecolor{fhgruen}{RGB}{0,132,98}
\newcommand\FHgruen{\textcolor{fhgruen}}
\definecolor{fhgrau}{RGB}{98,107,113}
\newcommand\FHgrau{\textcolor{fhgrau}}

\unitlength1mm

% Zitierung im Harvard-Style
\usepackage{harvard}
\usepackage{url} %Darstellung von URLs erlauben
\citationstyle{dcu}  % für Zitate
\bibliographystyle{style/HarvardFHTWMR_V1_2e}% für Literaturverzeichnis
\citationmode{abbr}
\newcommand\citet{\citeasnoun}
\newcommand\citep{\cite}
% \citet(cootes01) -> Cootes (2001)
% \citep(cootes01) -> (Cootes, 2001)
\renewcommand{\harvardand}{\&}
%Deutsch (wahlweise mit DE)
\newcommand{\acessedthrough}{Verfügbar unter:}%Für URL-Angabe
\newcommand{\acessedthroughp}{Verfügbar durch:}%Für URL-Angabe (Geschützte Datenbank, Zugriff durch FH)
\newcommand{\acessedat}{Zugang am}%Für URL-Datum-Angabe
%Englisch (wahlweise mit EN)
%\newcommand{\acessedthrough}{Available at:}%Für URL-Angabe
%\newcommand{\acessedthroughp}{Available through:}%Für URL-Angabe (Geschützte Datenbank, Zugriff durch FH)
%\newcommand{\acessedat}{Accessed}%Für URL-Datum-Angabe
% bis hierher

% Seiten-Layout definieren
\usepackage[tmargin=14.5mm, bmargin=20mm, lmargin=20mm, rmargin=20mm,
            paper=a4paper,nofoot=true, nohead=true, noheadfoot=true]{geometry}

\usepackage{multicol}
\setlength{\columnsep}{7mm}

% weniger Warnungen wegen überfüllter Boxen
\tolerance = 9999
\sloppy

% Anpassung einiger Überschriften
\addto\captionsngerman
{
  \renewcommand\figurename{Abb.}
  \renewcommand\tablename{Tab.}
}

% Abbildungen, Gleichungen und Tabellen werden fortlaufend nummeriert
\renewcommand\thefigure{\arabic{figure}}
\renewcommand\thetable{\arabic{table}}
\renewcommand\theequation{\arabic{equation}}
\renewcommand\thesection{\arabic{section}.}
\renewcommand\thesubsection{\thesection\arabic{subsection}.}

\makeatletter
  % Überschriften neu definieren
  \renewcommand\section
  {
    \@startsection
    {section}{1}{0mm}      % für 'section', Ebene 1, 0mm Einzug
    {9pt} {6pt}            % Abstand darüber und darunter
    {\noindent\fontsize{10}{9pt}\scshape\textbf}%
  }

  \renewcommand\subsection
  {
    \@startsection
    {subsection}{2}{0mm}    % für 'subsection', Ebene 2, 0mm Einzug
    {9pt} {6pt}             % Abstand darüber und darunter
    {\noindent\fontsize{9}{9pt}\textbf}
  }
  \renewcommand\abstract[2]
  {
    \noindent\fontsize{9}{9pt}\textit{\textbf{Abstract—} #1}\\
    \noindent\fontsize{9}{9pt}\textit{\textbf{Keywords—} #2}
  }

  \newcommand\FHInfo[8]
  {
    \begin{multicols}{2}
      \fontsize{11}{14pt}
      \noindent{\textbf{\small Master-Studiengang Robotics Engineering}}\\
      {\fontsize{10}{20pt}
        \noindent \small FH Technikum Wien, Höchstädtplatz 6, A-1200 Wien\\
        \vspace{30pt}
      }
      \columnbreak
      \begin{figure}[H]
        \begin{flushright}
        \includegraphics[width=40mm,height=15mm,keepaspectratio=true]{#3}
        \label{fig:logo}
        \end{flushright}
      \end{figure}
    \end{multicols}
    \begin{center}
      \fontsize{12}{14pt}
      \noindent \textbf{\textsc {#4\\}}
      \bigskip
      {\fontsize{10}{12pt}
        \noindent{\textbf{Student: } #5, \textbf{PK: } #6}\\
        \noindent{\textbf{Supervisor: } #7}\\
      }
    \end{center}
    \vspace{0mm}
  }
\makeatother

\begin{document}
  \pagestyle{empty}  % keine Kopf- oder Fusszeile

  \FHInfo
    {Eintrag wird nicht verwendet}              % Eintrag wird nicht verwendet
    {Eintrag wird nicht verwendet}              % Eintrag wird nicht verwendet
    {img_paper/FHTW_Logo_Farbe_randlos.jpg}                 % FH-Logo
    {Virtualization of a Real-Time Operating System for Robot Control \\
    with a Focus on Real-Time Compliance}                       % Titel der Arbeit
    {Pamuk, Halil Ibrahim, BSc.}            % Student-Name
    {51842568}                                    % Student-PK-Nr
    {Rauh, Sebastian, MSc. BEng.}            % Name FH-Betreuer (1. BegutachterIn)
    {Dr.nat.techn. Wöber, Wilfried, MSc.}            % Name Firmenbetreuer (2. BegutachterIn)

  \fontsize{9}{10pt}

  \begin{multicols}{2}
    \abstract{\textbf{Die  Kurzfassung soll einen Überblick über die Arbeit geben, sowie den "'roten Faden"' und die wichtigsten Details für den Leser liefern. Sie muss informativ sein, unabhängig ob sie alleine oder zusammen mit der Arbeit gelesen wird.}} {\textbf{Virtualization, Real-Time Systems, Latency Reduction, Robot Control}}

    \section{Einleitung}
    Die Einleitung soll die Aufmerksamkeit des Lesers erwecken und ausreichend Informationen für das Verständnis der Arbeit liefern sowie die Motivation zum Schreiben dieser Arbeit enthalten.
    Die Einleitung soll die Ziele der Arbeit erläutern und eine grobe Darstellung des Themenkreises enthalten, sowie die Beschreibung der Methoden und Vorgänge zur Problemlösung (Literaturrecherchen, vergleichende Studien oder Experimente).
    Eventuell kann das Firmenumfeld an dieser Stelle beschrieben werden.
    Hinweis: Der innovative Charakter der Arbeit soll in Einleitung und Problemstellung bereits klar ersichtlich sein

    \section{Problem- und Aufgabenstellung}
    Prägnante Formulierung der Kernfragen ohne den Ansatz der Lösungsfindung darzustellen, da dies Teil der Einleitung ist. Um informativ zu sein empfiehlt sich die Struktur eines Trichters. Die Idee der Trichterstruktur ist, dass sie breit beginnt und sich dann Schritt für Schritt zur Problemstellung verengt. Die Problemstellung soll schlüssig und nachvollziehbar hervorgehen.

    \section{Materialien und Methoden}
    In „Material und Methoden“ werden die theoretischen Grundlagen zum Verständnis der Arbeit aufbereitet, sowie die Experimente (evt. Studien), die in im Zuge der Arbeit durchgeführt wurden, dargestellt. Es sollen ausreichend Details und Referenzen angeführt sein, damit ein vergleichbar qualifizierter Kollege, Wissenschafter, etc. die Arbeit beurteilen und wiederholen kann. Alle in diesem Kapitel beschriebenen Materialien wie Geräte, Rechenprogramme, Reagenzien, etc. sind mit Namen, eventuell Typenbezeichnung, Herstellerfirma, Herkunftsland anzugeben.

    \section{Praktische Durchführung}
    Dieses Kapitel ist nur in der Variante „Arbeit mit Praxisanteil“ enthalten.
    Ausgehend von den Grundlagen werden die eigenen Ideen, Produkte, Konzepte, Technologien oder Software entwickelt. Dieser Prozess wird umfassend dokumentiert. Zum Abschluss werden die Ergebnisse in nachvollziehbarer Form dargestellt. Die Gliederung dieses Kapitels enthält in der Regel folgende Punkte: Planung (kann auch ähnlich wie in einem Projekthandbuch geschildert werden), gegebenenfalls Spezifikation, Umsetzung, Ergebnisse, eventuell Überprüfung der Ergebnisse.
    Bei Arbeiten mit wirtschaftlichem Hintergrund oder bei stark unternehmensbezogenen Fragestellungen empfiehlt sich auch eine detaillierte Beschreibung des Unternehmensumfeldes. Bei Untersuchungen oder Messaufbauten sollten die verwendeten Verfahren oder Messinstrumente ebenfalls kurz geschildert werden.

    \subsection{Abbildungen}
    Abbildungen sollen Druckqualität haben. Abbildungen sind mit einer fortlaufenden Nummer zu versehen unter Angabe der Quelle (falls nicht selbst erstellt).
    \begin{figure}[H]%
      \fbox{%
        \begin{minipage}{8.1cm}%
          \begin{center}%
            \includegraphics[width=80mm]{img_paper/FHTW_Logo_Farbe_randlos.jpg}%
          \end{center}%
        \end{minipage}%
      }%
      \caption{ Logo der FH-Technikum-Wien (Quelle: \protect\citep{HuZi10})}%
      \label{fig:fh_logo_welle}%
    \end{figure}%
    Jede Abbildung (auch die Abb. \ref{fig:fh_logo_welle}) ist im Text zu referenzieren und zu erläutern. Abbildungen sind mit einem Rahmen zu versehen.

    \subsection{Tabellen}
    Tabellen sind mit einer fortlaufenden Nummer zu versehen unter Angabe der Quelle (falls nicht selbst erstellt). Jede Tabelle (auch Tab. \ref{tab:example}) ist im Text zu referenzieren und zu erläutern. Tabellen sind mit einem Rahmen zu versehen.
    \begin{table}[H]%
      \centering%
      \caption{Beschreibung der Tabelle}\label{tab:example}%
      \begin{tabular}{ | c | c | c | }\hline
        {\bf Datum} & {\bf Thema} & {\bf Raum}\\ \hline
        20. 08. 2008 & Graphentheorie & HS 3.13\\ \hline
        01. 10. 2008 & Biomathematik & HS 1.05\\ \hline
      \end{tabular}%
    \end{table}%

    \subsection{Referenzieren}
     Bei Diagrammen oder Abbildungen, die von anderen Quellen übernommen werden, wird der Verweis im Text unterhalb der Abbildung angeführt.

    \subsection{Anmerkungen zum Format diese Papers}
    \begin{itemize}
      \item Freiraum: oben, unten, links und rechts jeweils 20 mm.
      \item 2 Spalten, Abstand zwischen den Spalten 7 mm.
      \item Beide Spalten auf der ersten und zweiten Seite muss enden am Rand der Seite.
      \item Paper soll in WINWORD oder \LaTeX~Format geschrieben werden.
      \item Schriftart ist Times New Roman
      \item Paper ist begrenzt auf exakt \textbf{zwei volle A4 Seiten}. Bitte nutzen Sie den gesamten verfügbaren Platz!
    \end{itemize}

    \section{Ergebnisse}
    Die Aufgabe dieses Kapitels ist es, die Ergebnisse der in „Material und Methoden“ beschriebenen Experimente zu formulieren. Weiters soll das Kapitel den Leser zu den Daten hinführen und deren Abfolge erläutern.

    \section{Zusammenfassung und Ausblick}
    Zusammenfassung stellt das Resümee und die kritische Reflexion über das Projekt dar.

    \section{Literaturverzeichnis}
    {
      \renewcommand{\refname}{\vspace{-5mm}} % keine eigene Überschrift durch \bibliography...
      \linespread{.8}\selectfont
      \bibliography{Literature_paper}
    }

  \end{multicols}

\end{document}

### 2024-08-29 Paper.tex
gitignore
organize

@@
     {Eintrag wird nicht verwendet}              % Eintrag wird nicht verwendet
-    {img_paper/FHTW_Logo_Farbe_randlos.jpg}                 % FH-Logo
+    {img/paper/FHTW_Logo_Farbe_randlos.jpg}                 % FH-Logo
     {Virtualization of a Real-Time Operating System for Robot Control \\
@@
           \begin{center}%
-            \includegraphics[width=80mm]{img_paper/FHTW_Logo_Farbe_randlos.jpg}%
+            \includegraphics[width=80mm]{img/paper/FHTW_Logo_Farbe_randlos.jpg}%
           \end{center}%
@@
       }%
-      \caption{ Logo der FH-Technikum-Wien (Quelle: \protect\citep{HuZi10})}%
+      \caption{Logo der FH-Technikum-Wien (Quelle:)}%
       \label{fig:fh_logo_welle}%

### 2024-08-29 Paper.tex
Start of Paper

@@
   \begin{multicols}{2}
-    \abstract{\textbf{Die  Kurzfassung soll einen Überblick über die Arbeit geben, sowie den "'roten Faden"' und die wichtigsten Details für den Leser liefern.
-Sie muss informativ sein, unabhängig ob sie alleine oder zusammen mit der Arbeit gelesen wird.}} {\textbf{Virtualization, Real-Time Systems, Latency Reduction, Robot Control}}
+    \abstract{\textbf{Virtualizing real-time operating systems for robotic control offers many advantages over traditional hardware-based approaches, especially when it comes to flexibility, scaling, and costs.
+Virtualization addresses these issues but introduces overhead and latency.
+This thesis investigates the virtualization of a real-time operating system, which is built with Yocto, employs hard real-time through Xenomai 3 and is emulated using QEMU/KVM.
+The primary objective aim was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is crucial for real-time robotics.
+Initial latency measurements showed a high gap between bare metal and virtualized setups.
+Extensive tuning across BIOS, kernel, host OS, QEMU/KVM, and Salamander 4 (guest) reduced worst-case latency from 707.622 µs to 17.134 µs, close to the bare metal performance of 10.709 µs.
+The improvement is validated using a robotic application, comparing tuned virtualization with untuned and hardware versions.}} {\textbf{Virtualization, Real-Time Systems, Latency Reduction, Robot Control}}
 
-    \section{Einleitung}
-    Die Einleitung soll die Aufmerksamkeit des Lesers erwecken und ausreichend Informationen für das Verständnis der Arbeit liefern sowie die Motivation zum Schreiben dieser Arbeit enthalten.
-    Die Einleitung soll die Ziele der Arbeit erläutern und eine grobe Darstellung des Themenkreises enthalten, sowie die Beschreibung der Methoden und Vorgänge zur Problemlösung (Literaturrecherchen, vergleichende Studien oder Experimente).
-    Eventuell kann das Firmenumfeld an dieser Stelle beschrieben werden.
-    Hinweis: Der innovative Charakter der Arbeit soll in Einleitung und Problemstellung bereits klar ersichtlich sein
+    \section{Introduction}
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\columnwidth]{img/implementation/gnuplot_combined_max_latency.png}
+      \caption[Initial Comparison of Latency Distribution between Hardware and Virtualization]{Initial Comparison of Latency Distribution between Hardware and Virtualization}
+      \label{fig:gnuplot_max_latency_combined}
+    \end{figure}
 
-    \section{Problem- und Aufgabenstellung}
-    Prägnante Formulierung der Kernfragen ohne den Ansatz der Lösungsfindung darzustellen, da dies Teil der Einleitung ist.
-Um informativ zu sein empfiehlt sich die Struktur eines Trichters.
-Die Idee der Trichterstruktur ist, dass sie breit beginnt und sich dann Schritt für Schritt zur Problemstellung verengt.
-Die Problemstellung soll schlüssig und nachvollziehbar hervorgehen.
+    \section{Methodology}
+    \section{Implementation}
+    \section{Initial Situation} 
+    \section{Real-Time Performance Tunings}
+    \section{Real-Time Robotic Application}
+    \section{Results}
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\columnwidth]{img/results/gnuplot_combined_max_latency_all.png}
+      \caption[Comparison of Latency Distribution of Salamander 4 Configurations]{Comparison of Latency Distribution of Salamander 4 under different Configurations}
+      \label{fig:max_latency_combined_results}
+    \end{figure}
 
-    \section{Materialien und Methoden}
-    In „Material und Methoden“ werden die theoretischen Grundlagen zum Verständnis der Arbeit aufbereitet, sowie die Experimente (evt.
-Studien), die in im Zuge der Arbeit durchgeführt wurden, dargestellt.
-Es sollen ausreichend Details und Referenzen angeführt sein, damit ein vergleichbar qualifizierter Kollege, Wissenschafter, etc.
-die Arbeit beurteilen und wiederholen kann.
-Alle in diesem Kapitel beschriebenen Materialien wie Geräte, Rechenprogramme, Reagenzien, etc.
-sind mit Namen, eventuell Typenbezeichnung, Herstellerfirma, Herkunftsland anzugeben.
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\columnwidth]{img/results/combined_latencies.png}
+      \caption[Comparison of Robotic Application Latency of Salamander 4 Configurations]{Comparison of Robotic Application Latency of Salamander 4 under different Configurations}
+      \label{fig:combined_latencies}
+    \end{figure}
 
-    \section{Praktische Durchführung}
-    Dieses Kapitel ist nur in der Variante „Arbeit mit Praxisanteil“ enthalten.
-    Ausgehend von den Grundlagen werden die eigenen Ideen, Produkte, Konzepte, Technologien oder Software entwickelt.
-Dieser Prozess wird umfassend dokumentiert.
-Zum Abschluss werden die Ergebnisse in nachvollziehbarer Form dargestellt.
-Die Gliederung dieses Kapitels enthält in der Regel folgende Punkte: Planung (kann auch ähnlich wie in einem Projekthandbuch geschildert werden), gegebenenfalls Spezifikation, Umsetzung, Ergebnisse, eventuell Überprüfung der Ergebnisse.
-    Bei Arbeiten mit wirtschaftlichem Hintergrund oder bei stark unternehmensbezogenen Fragestellungen empfiehlt sich auch eine detaillierte Beschreibung des Unternehmensumfeldes.
-Bei Untersuchungen oder Messaufbauten sollten die verwendeten Verfahren oder Messinstrumente ebenfalls kurz geschildert werden.
+\begin{table}[H]
+	\scriptsize
+	\centering
+	\caption{Comparison of Robotic Application Latency Statistics}
+	\label{tab:robotic_application_latency_values_combined}
+	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
+	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
+		\begin{tabular}{|l|l|l|l|l|l|}
+			\hline
+			\textbf{Tuning} & \textbf{Samples} & \textbf{Lat Min (ms)} & \textbf{Lat Avg (ms)} & \textbf{Lat Max (ms)} & \textbf{Std Dev (ms)} \\ \hline
+			Hardware & 1000 & 1.211 & 1.347 & 1.49 & 0.082 \\ \hline
+			Untuned Virtualization & 1000 & 3.1 & 24.603 & 129.46 & 13.876 \\ \hline
+			Tuned Virtualization & 1000 & 1.219 & 2.62 & 3.988 & 0.812 \\ \hline
+		\end{tabular}}
+\end{table}
 
-    \subsection{Abbildungen}
-    Abbildungen sollen Druckqualität haben. Abbildungen sind mit einer fortlaufenden Nummer zu versehen unter Angabe der Quelle (falls nicht selbst erstellt).
-    \begin{figure}[H]%
-      \fbox{%
-        \begin{minipage}{8.1cm}%
-          \begin{center}%
-            \includegraphics[width=80mm]{img/paper/FHTW_Logo_Farbe_randlos.jpg}%
-          \end{center}%
-        \end{minipage}%
-      }%
-      \caption{Logo der FH-Technikum-Wien (Quelle:)}%
-      \label{fig:fh_logo_welle}%
-    \end{figure}%
-    Jede Abbildung (auch die Abb. \ref{fig:fh_logo_welle}) ist im Text zu referenzieren und zu erläutern. Abbildungen sind mit einem Rahmen zu versehen.
-
-    \subsection{Tabellen}
-    Tabellen sind mit einer fortlaufenden Nummer zu versehen unter Angabe der Quelle (falls nicht selbst erstellt).
-Jede Tabelle (auch Tab.
-\ref{tab:example}) ist im Text zu referenzieren und zu erläutern.
-Tabellen sind mit einem Rahmen zu versehen.
-    \begin{table}[H]%
-      \centering%
-      \caption{Beschreibung der Tabelle}\label{tab:example}%
-      \begin{tabular}{ | c | c | c | }\hline
-        {\bf Datum} & {\bf Thema} & {\bf Raum}\\ \hline
-        20. 08. 2008 & Graphentheorie & HS 3.13\\ \hline
-        01. 10. 2008 & Biomathematik & HS 1.05\\ \hline
-      \end{tabular}%
-    \end{table}%
-
-    \subsection{Referenzieren}
-     Bei Diagrammen oder Abbildungen, die von anderen Quellen übernommen werden, wird der Verweis im Text unterhalb der Abbildung angeführt.
-
-    \subsection{Anmerkungen zum Format diese Papers}
-    \begin{itemize}
-      \item Freiraum: oben, unten, links und rechts jeweils 20 mm.
-      \item 2 Spalten, Abstand zwischen den Spalten 7 mm.
-      \item Beide Spalten auf der ersten und zweiten Seite muss enden am Rand der Seite.
-      \item Paper soll in WINWORD oder \LaTeX~Format geschrieben werden.
-      \item Schriftart ist Times New Roman
-      \item Paper ist begrenzt auf exakt \textbf{zwei volle A4 Seiten}. Bitte nutzen Sie den gesamten verfügbaren Platz!
-    \end{itemize}
-
-    \section{Ergebnisse}
-    Die Aufgabe dieses Kapitels ist es, die Ergebnisse der in „Material und Methoden“ beschriebenen Experimente zu formulieren.
-Weiters soll das Kapitel den Leser zu den Daten hinführen und deren Abfolge erläutern.
-
-    \section{Zusammenfassung und Ausblick}
+    \section{Discussion}
+    \section{Summary and Outlook}
     Zusammenfassung stellt das Resümee und die kritische Reflexion über das Projekt dar.

### 2024-08-30 Masterarbeit_Robotik_Pamuk.tex
bind paper
Paper: Abstract, Introduction, Methodology, Implementation, Results, Discussion, Summary and Outlook, Bibliography

@@
 %
-\documentclass[MMR,Master,english]{twbook}
+\documentclass[MMR,Master,english]{style/twbook}
 \usepackage[utf8]{inputenc}
@@
 \usepackage[backend=biber, style=numeric, sorting=none]{biblatex}
-\addbibresource{Literature.bib}
+\addbibresource{literature/Literature.bib}
 %
@@
 \bigskip \noindent Altogether, this thesis provides a comprehensive blueprint for making a virtualized guest system real-time capable in a host system with deterministic behavior.}
-\keywords{Virtualization, Real-Time Systems, Latency Reduction, Robotic Control}
+\keywords{Virtualization, Real-Time Systems, Latency Reduction, Robot Control}
 
@@
 Next, the host is configured, but the BIOS and kernel settings remain unchanged.
-After that, the guest is configured, and finally, QEMU/KVM settings are applied.
+The guest OS is already based on Xenomai and therefore, no additional modifications are necessary for the guest OS itself.
+Finally, QEMU/KVM settings are applied.
 This means that the configurations are not isolated but are applied and then tested as a whole.
@@
 
-\begin{comment}
-
-\chapter{To Include}\label{cha:to_include}
-
-\section{KVM Entry and KVM Exit}\label{subsubsec:kvm_entry_exit}
-\textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host.
-However, the isolcpus function only isolates CPUs from the scheduler, meaning that user-space processes won't be scheduled on these CPUs, but kernel tasks and interrupts can still utilize the CPU.This led to the investigation of the causes for the observed high and inconsistent latency.
-The guest operates within the~\texttt{kvm\_entry} and~\texttt{kvm\_exit} events of the host.
-Kernelshark revealed a high frequency of~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host.
-This frequent switching hinders the guest's ability to run continuously, thereby increasing the virtualization latency.}
-
-\textcolor{red}{To further understand this, trace-cmd was employed to trace various events in the host-guest communication, including the reasons for these events.
-Specifically, the causes for~\texttt{kvm\_exit} events were analyzed.
-The command sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all was executed on the host for a duration of 5 seconds.
-The results in Figure~\ref{fig:kvm_exit_taskset} were obtained.
-Additionally, Table~\ref{tab:kvm_exit} provides a short description of the observed~\texttt{kvm\_exit} events.}
-
-\vspace{2em}
-\begin{table}[H]
-	\centering
-	\begin{tabular}{|l|p{0.62\textwidth}|}
-		\hline
-		\textbf{Exit Reason} & \textbf{Description}                                     \\ \hline
-		APIC\_WRITE          & Triggered when the guest writes to its APIC.             \\ \hline
-		EXTERNAL\_INTERRUPT  & Triggered by external hardware interrupts.               \\ \hline
-		HLT                  & Triggered when the guest executes the HLT instruction.   \\ \hline
-		EPT\_MISCONFIG       & Triggered by a misconfiguration in the EPT.              \\ \hline
-		PREEMPTION\_TIMER    & Triggered when the host's preemption timer expires.      \\ \hline
-		PAUSE\_INSTRUCTION   & Triggered when the PAUSE instruction is executed.        \\ \hline
-		EPT\_VIOLATION       & Triggered by a violation of the EPT permission settings. \\ \hline
-		IO\_INSTRUCTION      & Triggered when the guest executes an I/O instruction.    \\ \hline
-		EOI\_INDUCED         & Triggered when an EOI signal is sent to the APIC.        \\ \hline
-		MSR\_READ            & Triggered when the guest reads from a MSR.               \\ \hline
-		CPUID                & Triggered when the guest executes the CPUID instruction. \\ \hline
-	\end{tabular}
-	\caption{Description of kvm\_exit reasons}
-	\label{tab:kvm_exit}
-\end{table}
-
-\textcolor{red}{\noindent In the process of analyzing the~\texttt{kvm\_exit} events, several reasons for these exits were identified.
-The most frequent among these were the~\texttt{APIC\_WRITE} and~\texttt{HLT} events.
-The former is initiated when the guest writes to its Advanced Programmable Interrupt Controller (APIC), a component of the CPU that manages hardware interrupts.
-The latter occurs when the guest executes the HLT instruction, effectively halting the CPU until the next external interrupt is fired.
-Other significant but less frequent events included~\texttt{EXTERNAL\_INTERRUPT} and~\texttt{IO\_INSTRUCTION}.
-These events are indicative of the guest's interaction with hardware devices and its execution of I/O operations.
-Events such as~\texttt{EPT\_MISCONFIG} and~\texttt{PREEMPTION\_TIMER} were also noted.
-These could potentially signal issues with memory management and the host's scheduling of the guest.
-While events like~\texttt{PAUSE\_INSTRUCTION},~\texttt{EPT\_VIOLATION},~\texttt{EOI\_INDUCED},~\texttt{MSR\_READ}, and~\texttt{CPUID} were the least frequent, they still provide valuable insights into the guest's behavior and the host-guest interaction.}
-
-\textcolor{red}{\noindent When the script is started from the host, the QEMU process can be scheduled to run on any available core, as it is not bound to a specific CPU core.
-This means that the QEMU process may frequently switch between different cores, leading to an increase in latency.
-As the goal was to reduce latency in the guest, the first step was to isolate a CPU of the host and dedicate it solely to the QEMU process, so that it cannot be used for other tasks on the user level.
-Without CPU isolation, context switches take place at the operating system level and not at the hypervisor level.
-This explains why there are fewer~\texttt{kvm\_exit} events.
-However, this will also, as previously shown, lead to higher latency, as context switches at the operating system level generally take longer than a~\texttt{kvm\_exit} and ~\texttt{kvm\_entry}.
-When the CPU was dedicated to the QEMU process, on the other hand, there was a significant increase in~\texttt{kvm\_exit} events.
-This is because every context switch takes place at the hypervisor level.
-Nevertheless, lower latency was achieved thereby, as the qemu process is no longer influenced by the CPU scheduling of the operating system.}
-
-``korrekt''
-
-\noindent \enquote{This LaTeX.}
-
-\section{Trace-cmd and Kernelshark}
-After analyzing the inital latency of both versions, Trace-cmd and Kernelshark were used to further inspect the reasons that caused this divergence.
-
-\section{Host and Guest tasks}
-\vspace{2em}
-\begin{table}[H]
-\centering
-\begin{minipage}{.5\textwidth}
-\centering
-\caption{Host report CPU19 (Total of 445.908)}
-\label{tab:results_host_report}
-\begin{tabular}{|c|c|c|}
-	\hline
-	PID    & Task                          & Count   \\ \hline
-	182579 & qemu-system-x86               & 302.748 \\ \hline
-	0      & \textless{}idle\textgreater{} & 112.911 \\ \hline
-	182618 & vhost                         & 21.204  \\ \hline
-	182572 & qemu-system-x86               & 7.597   \\ \hline
-	182755 & qemu-system-x86               & 644     \\ \hline
-	182754 & qemu-system-x86               & 643     \\ \hline
-	181870 & kworker/19:1                  & 139     \\ \hline
-	3820   & kworker/19:1H                 & 16      \\ \hline
-	94     & migration/19                  & 6       \\ \hline
-\end{tabular}
-%
-\begin{minipage}{.5\textwidth}
-\centering
-\caption{Guest report (Total of 362.370)}
-\label{tab:results_guest_report}
-\begin{tabular}{|c|c|c|}
-	\hline
-	PID & Task                          & Count   \\ \hline
-	0   & \textless{}idle\textgreater{} & 150.744 \\ \hline
-	331 & LRT-Main                      & 56.697  \\ \hline
-	377 & trace-cmd                     & 48.507  \\ \hline
-	346 & CLI                           & 26.426  \\ \hline
-	378 & kthreadd                      & 25.837  \\ \hline
-	340 & MainTaskLow                   & 19.291  \\ \hline
-	339 & \textless{}...\textgreater{}  & 9.980   \\ \hline
-	34  & MainTaskHigh                  & 9.185   \\ \hline
-	327 & LE-Logger                     & 4.965   \\ \hline
-	369 & kworker/0:0                   & 2.793   \\ \hline
-	321 & kWorker-LRT                   & 2.542   \\ \hline
-	328 & LRTMgr-Main                   & 1.651   \\ \hline
-	34  & LrtMgrCyclic                  & 1.220   \\ \hline
-	332 & cobalt\_printf                & 1.112   \\ \hline
-	325 & LE-System                     & 534     \\ \hline
-	343 & TCP-Listen                    & 187     \\ \hline
-	15  & rcu\_preempt                  & 162     \\ \hline
-	25  & kcompactd0                    & 122     \\ \hline
-	58  & kworker/0:1H                  & 96      \\ \hline
-	63  & kworker/u2:2                  & 89      \\ \hline
-	8   & jbd2/sda-8                    & 86      \\ \hline
-	22  & kworker/0:1                   & 56      \\ \hline
-	1   & init                          & 31      \\ \hline
-	2   & kthreadd                      & 25      \\ \hline
-	375 & trace-cmd                     & 24      \\ \hline
-	14  & ksoftirqd/0                   & 8       \\ \hline
-\end{tabular}
-
-\end{table}
-
-\noindent In the following, the host and guest tasks along with their impact on system latency are briefly described.
-
-\begin{itemize}
-	\item \textbf{qemu-system-x86}: Part of the QEMU process and specifically, this task emulates x86 systems.
-In Table~\ref{tab:results_host_report}, it occurs four times under different PIDs, hence there are four threads of it.
-	\item \textbf{<idle>}: This represents the idle time of the CPU, hence it is not being used by any process, allowing to save power.
-The system halts until the next interrupt, which could be a timer interrupt, I/O interrupt, etc.
-	\item \textbf{vhost}: A kernel module which improves virtual input/output (virtio) performance by handling virtqueues in the kernel, thereby reducing context switches and system calls.
-	\item \textbf{kworker/19:1}: A kernel worker thread created by the Linux kernel, kworker/19:1 performs work in response to system events.
-The number after the slash and colon indicate the CPU core and internal ID of the worker thread, respectively.
-	\item \textbf{kworker/19:1H}: Similar to kworker/19:1, kworker/19:1H is a kernel worker thread, with the ‘H’ suggesting that this thread handles hardware interrupts.
-	\item \textbf{migration/19}: The migration process is a kernel process that balances load across CPU cores by moving threads from one CPU to another.
-The number after the slash indicates the CPU core to which the migration process is bound.
-(URL: \url{https://elixir.bootlin.com/linux/latest/source/kernel/sched/core.c#L2325})
-
-\end{itemize}
-
-The Figure below (deleted) compares non-optimized guest latency with optimized guest latency and includes optimized bare metal latency as a reference.
-The data shows that a 40\% reduction in QD1 latency is achievable through system tuning.
-
-\end{comment}
-
 \chapter{Results}\label{cha:results}
@@
 \noindent The improved latency was also tested with the robotic application in Subsection~\ref{sec:robotic_application}.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the bare metal, untuned and tuned version of Salamander 4.
 The results are visualized in Figure~\ref{fig:combined_latencies}.
@@
 			\textbf{Tuning} & \textbf{Samples} & \textbf{Lat Min (ms)} & \textbf{Lat Avg (ms)} & \textbf{Lat Max (ms)} & \textbf{Std Dev (ms)} \\ \hline
-			Hardware & 1000 & 1.211 & 1.347 & 1.49 & 0.082 \\ \hline
+			Bare Metal & 1000 & 1.211 & 1.347 & 1.49 & 0.082 \\ \hline
 			Untuned Virtualization & 1000 & 3.1 & 24.603 & 129.46 & 13.876 \\ \hline
@@
 \chapter{Summary and Outlook}\label{cha:summary_and_outlook}
-
 Hardware-based systems are limited in adaptability and can be costly, especially when scaling operations.

### 2024-08-30 Paper.tex
Paper: Abstract, Introduction, Methodology, Implementation, Results, Discussion, Summary and Outlook, Bibliography

@@
 
-% Zitierung im Harvard-Style
-\usepackage{harvard}
 \usepackage{url} %Darstellung von URLs erlauben
-\citationstyle{dcu}  % für Zitate
-\bibliographystyle{style/HarvardFHTWMR_V1_2e}% für Literaturverzeichnis
-\citationmode{abbr}
-\newcommand\citet{\citeasnoun}
-\newcommand\citep{\cite}
-% \citet(cootes01) -> Cootes (2001)
-% \citep(cootes01) -> (Cootes, 2001)
-\renewcommand{\harvardand}{\&}
+
 %Deutsch (wahlweise mit DE)
@@
 
+% own 
+\usepackage{multirow}
+\usepackage{array}
+\usepackage[backend=biber, style=numeric, sorting=none]{biblatex}
+\addbibresource{literature/Paper.bib}
+% Set the bibliography font size to scriptsize
+\AtBeginBibliography{\scriptsize}
+\renewbibmacro{in:}{}
+
 \begin{document}
@@
 \begin{multicols}{2}
-    \abstract{\textbf{Virtualizing real-time operating systems for robotic control offers many advantages over traditional hardware-based approaches, especially when it comes to flexibility, scaling, and costs.
-Virtualization addresses these issues but introduces overhead and latency.
-This thesis investigates the virtualization of a real-time operating system, which is built with Yocto, employs hard real-time through Xenomai 3 and is emulated using QEMU/KVM.
+  \abstract{\textbf{This thesis investigates the virtualization of a real-time operating system, which is built with Yocto, employs hard real-time through Xenomai 3 and is emulated using QEMU/KVM.
 The primary objective aim was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is crucial for real-time robotics.
 Initial latency measurements showed a high gap between bare metal and virtualized setups.
-Extensive tuning across BIOS, kernel, host OS, QEMU/KVM, and Salamander 4 (guest) reduced worst-case latency from 707.622 µs to 17.134 µs, close to the bare metal performance of 10.709 µs.
-The improvement is validated using a robotic application, comparing tuned virtualization with untuned and hardware versions.}} {\textbf{Virtualization, Real-Time Systems, Latency Reduction, Robot Control}}
+Extensive tuning across BIOS, kernel, host OS, QEMU/KVM, and guest OS reduced the worst-case latency from 707.622µs to 17.134µs, close to the bare metal performance of 10.709µs.
+The improvement was validated using a robotic application, comparing tuned virtualization with untuned and hardware versions.}} {\textbf{Virtualization, Real-Time Systems, Latency Reduction, Robot Control}}
 
   \section{Introduction}
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\columnwidth]{img/implementation/gnuplot_combined_max_latency.png}
-      \caption[Initial Comparison of Latency Distribution between Hardware and Virtualization]{Initial Comparison of Latency Distribution between Hardware and Virtualization}
-      \label{fig:gnuplot_max_latency_combined}
-    \end{figure}
+  In today's industrial production and automation, robots must react to their environment and perform time-critical tasks within strict time constraints.
+Delays or errors can have catastrophic consequences in some cases.
+Virtualizing real-time operating systems for robotic control offers many advantages over traditional hardware-based approaches, especially when it comes to scaling~\cite{abbasiExploringOpenStackScalable2023}, flexibility and costs~\cite{taccariEmbeddedRealTimeVirtualization2014}.
+However, virtualization also introduces overhead and latency~\cite{casiniLatencyAnalysisVirtualization2021}~\cite{zhangEvaluatingOptimizingVirtualization2010}.
+In this master's thesis, the goal was the virtualization of a real-time operating system to control a robot, with a focus on compliance with real-time determinism.
+Specifically, the aim was to bring the latency of the virtualized Salamander 4 closer to that of the bare metal version, as initial latency measurements revealed a significant latency gap between the bare metal and virtualized setup.
 
   \section{Methodology}
+  In order To address this gap, an extensive tuning process was carried out to achieve real-time performance and determinism.
+These tunings were added sequentially and tested together.
+This involves configurations spanning the BIOS, kernel, host operating system (OS) and QEMU/KVM virtualization layer, as demonstrated in~\cite{RealTimePerformanceTuning2022}.
+The guest OS is already based on Xenomai and therefore, no additional modifications were necessary.
+Ubuntu 22.04.4 LTS served as the host operating system, while Salamander 4, built with Yocto and based on Linux 5.15.94, was the guest OS.
+Xenomai provided real-time capabilities for the guest and its \texttt{latency} tool was used to measure the scheduling latency of the real-time OS.
+QEMU with KVM was employed for virtualization.
+Trace-cmd and Kernelshark were used for kernel tracing and visualization of trace data.
+The testing robot was connected to the virtual CPUS via a VARAN bus interface.
+
   \section{Implementation}
-    \section{Initial Situation} 
-    \section{Real-Time Performance Tunings}
-    \section{Real-Time Robotic Application}
+  The initial latency values were first measured with the \texttt{latency} program of the Xenomai tool suite.
+The tests were conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
+If any sample's latency value was greater than 100µs, this was considered an overrun.
+There was a significant initial gap in latency statistics between the virtualized Salamander 4 and the Salamander 4 on bare metal.
+To address this gap, an extensive tuning process was carried out to achieve real-time performance and determinism.
+These modifications were added sequentially and tested together.
+Previous tunings were not reverted when moving to the next tunings.
+First, the BIOS was configured, followed by the kernel.
+The BIOS settings were not reverted when moving to the kernel configurations.
+Next, the host was configured, but the BIOS and kernel settings remained unchanged.
+The guest configurations were not changed since Xenomai already provides real-time capabilities.
+Finally, QEMU/KVM settings were applied.
+On top of that, a robotic application further confirmed the improvements.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
+
   \section{Results}
+  This section summarizes the results after the real-time performance tunings, both in terms of the \texttt{latency} program and the robotic application.
+Figure~\ref{fig:max_latency_combined_results} displays the change in latency after each real-time performance tuning.
+Table~\ref{tab:latency_tables_combined} gives an overview of the most important metrics of the measurements after each tuning.
+\clearpage
   \begin{figure}[H]
@@
 
-    \begin{figure}[H]
+  \begin{table}[H]
     \centering
-      \includegraphics[width=1.0\columnwidth]{img/results/combined_latencies.png}
-      \caption[Comparison of Robotic Application Latency of Salamander 4 Configurations]{Comparison of Robotic Application Latency of Salamander 4 under different Configurations}
-      \label{fig:combined_latencies}
-    \end{figure}
+    \footnotesize
+    \caption{Comparison of Latency Statistics in Salamander 4 under different Configurations}
+    \label{tab:latency_tables_combined}
+    \begin{tabular}{|c|c|c|c|c|}
+      \hline
+      \multirow{2}{*}{\textbf{Tuning}} & \multicolumn{3}{c|}{\textbf{Latency (µs)}} & \multirow{2}{*}{\textbf{Overruns}} \\ \cline{2-4}
+      & \textbf{Min} & \textbf{Avg} & \textbf{Max} & \\ \hline
+      Bare Metal & 0.613 & 1.380 & 10.709 & 0 \\ \hline
+      \begin{tabular}[c]{@{}c@{}}Untuned\\ Virtualization\end{tabular} & 2.536 & 8.940 & 707.622 & 43 \\ \hline
+      \begin{tabular}[c]{@{}c@{}}After BIOS\\ Configurations\end{tabular} & 0.969 & 3.948 & 457.545 & 94 \\ \hline
+      \begin{tabular}[c]{@{}c@{}}After Kernel\\ Configurations\end{tabular} & 2.545 & 4.811 & 21.694 & 0 \\ \hline
+      \begin{tabular}[c]{@{}c@{}}After Host\\ Configurations\end{tabular} & 2.591 & 4.834 & 18.441 & 0 \\ \hline
+      \begin{tabular}[c]{@{}c@{}}After QEMU\\ Configurations\end{tabular} & 2.614 & 4.779 & 17.134 & 0 \\ \hline
+    \end{tabular}
+\end{table}
+
+The improved latency was also tested with a robotic application.
+The difference between the command issuance time and the time the signal reaches the PWM was measured 1,000 times for the bare metal, untuned and tuned version of Salamander 4.
+The results are visualized in Table~\ref{tab:robotic_application_latency_values_combined}.
 
   \begin{table}[H]
-	\scriptsize
     \centering
-	\caption{Comparison of Robotic Application Latency Statistics}
+    \footnotesize
+    \caption{Comparison of Latency Statistics in Salamander 4 under different Configurations}
     \label{tab:robotic_application_latency_values_combined}
-	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
-	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|l|l|l|l|l|l|}
+    \begin{tabular}{|c|c|c|c|c|}
       \hline
-			\textbf{Tuning} & \textbf{Samples} & \textbf{Lat Min (ms)} & \textbf{Lat Avg (ms)} & \textbf{Lat Max (ms)} & \textbf{Std Dev (ms)} \\ \hline
-			Hardware & 1000 & 1.211 & 1.347 & 1.49 & 0.082 \\ \hline
-			Untuned Virtualization & 1000 & 3.1 & 24.603 & 129.46 & 13.876 \\ \hline
-			Tuned Virtualization & 1000 & 1.219 & 2.62 & 3.988 & 0.812 \\ \hline
-		\end{tabular}}
+      \multirow{2}{*}{\textbf{Tuning}} & \multicolumn{3}{c|}{\textbf{Latency (ms)}} & \multirow{2}{*}{\begin{tabular}[c]{@{}c@{}}\textbf{Std Dev}\\ \textbf{(ms)}\end{tabular}} \\ \cline{2-4}
+      & \textbf{Min} & \textbf{Avg} & \textbf{Max} & \\ \hline
+      Bare Metal & 1.211  & 1.347 & 1.49 & 0.082 \\ \hline
+      \begin{tabular}[c]{@{}c@{}}Untuned\\ Virtualization\end{tabular} & 3.1 & 24.603 & 129.46 & 13.876\\ \hline
+      \begin{tabular}[c]{@{}c@{}}Tuned\\ Virtualization\end{tabular} & 1.219  & 2.62 & 3.988 & 0.812 \\ \hline
+    \end{tabular}
 \end{table}
@@
   \section{Discussion}
+  The results shows that the real-time performance tunings applied to the virtualized Salamander 4 OS reduced latency and improved determinism.
+The maximum latency decreased from 707.622µs to 17.134µs after tuning BIOS, kernel, host, guest, and QEMU/KVM configurations.
+Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694µs.
+This also eliminated overruns going forward.
+Host configurations, including
+  the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU further reduced latency a little bit down to 18.441µs.
+The guest OS already had real-time capabilities through Xenomai, hence the focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
+Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages brought the latency down to the final worst latency value of 17.134µs.
+This value is very close to the bare metal value of 10.709\textmu s.
+The goal was set to achieving latency values below 50 microseconds in the duration of the measuremenet.
+This goal was acheived.
+  A robotic application in a practical scenario validated the improvement of the tuned virtualization compared to the unmodified version.
+The difference between command time and signal reaching the PWM was measured 1,000 times for untuned, tuned, and hardware versions of Salamander 4.
+The worst latency dropped from 129ms to 3.988ms in the developed application.
+Compared to the hardware version's worst latency of 1.49ms, the tuned virtualization came very close to the determinism and reliability of the hardware.
   \section{Summary and Outlook}
-    Zusammenfassung stellt das Resümee und die kritische Reflexion über das Projekt dar.
+  This master's thesis aimed to virtualize a real-time operating system for robot control, focusing on real-time determinism.
+The goal was to reduce latency in virtualized Salamander 4 to match its bare metal version.
+Initial tests showed a significant latency gap between virtualized and bare metal versions.
+Extensive tuning was performed sequentially: BIOS, kernel, host, and QEMU/KVM configurations.
+This process reduced worst latency from 707.62µs to 17.134µs, close to the bare metal value of 10.709µs, achieving the goal of sub-50µs latency.
+  A robotic application confirmed these improvements.
+Measuring the time between command and PWM signal 1,000 times showed worst latency dropping from 129ms to 3.988ms, approaching the hardware version's 1.49ms.
+  Future work could include further optimizations, testing other hypervisors, and evaluating performance under various workloads and stressed situations.
+  Future work could include additional configurations and optimizations of the virtualization layer, investigating other hypervisors and virtualization technologies, and extensive testing under various workloads and stressed situations.
 
-    \section{Literaturverzeichnis}
-    {
-      \renewcommand{\refname}{\vspace{-5mm}} % keine eigene Überschrift durch \bibliography...
-      \linespread{.8}\selectfont
-      \bibliography{Literature_paper}
-    }
+  \section{Bibliography}
+  \printbibliography
 

### 2024-08-31 Masterarbeit_Robotik_Pamuk.tex
bare-metal

@@
 Salamander 4 wird mit Yocto gebaut und verwendet harte Echtzeit mit Xenomai 3.
-Das Hauptziel besteht darin, die Latenzlücke zwischen den virtualisierten und den Bare-Metal-Versionen zu überbrücken, um ein deterministisches und zuverlässiges Verhalten zu gewährleisten, das für Echtzeit-Roboteranwendungen entscheidend ist.
-
-\bigskip \noindent Erste Latenzmessungen zeigten eine signifikante Leistungslücke zwischen dem Bare-Metal- und dem virtualisierten Setup.
+Das Hauptziel besteht darin, die Latenzlücke zwischen der Bare-Metal und der Virtualisierung zu überbrücken, um ein deterministisches und zuverlässiges Verhalten zu gewährleisten, das für Echtzeit-Roboteranwendungen entscheidend ist.
+
+\bigskip \noindent Erste Latenzmessungen zeigten eine signifikante Leistungslücke zwischen der Bare-Metal und dem virtualisierten Setup.
 Daher wird ein umfangreicher Tuning-Prozess durchgeführt, um Echtzeit-Performance und Determinismus zu erreichen.
@@
 Die Worst-Case-Latenz wurde von 707.622\textmu s auf 17.134\textmu s reduziert, was der Bare-Metal-Performance von 10.709\textmu s nahekommt.
-Darüber hinaus wird die Verbesserung der Echtzeit-Performance und des Determinismus anhand einer Roboteranwendung validiert, bei der die abgestimmte Virtualisierung mit der nicht abgestimmten und der Hardware-Version verglichen wird.
+Darüber hinaus wird die Verbesserung der Echtzeit-Performance und des Determinismus anhand einer Roboteranwendung validiert, bei der die abgestimmte Virtualisierung mit der nicht abgestimmten und der Bare-Metal-Version verglichen wird.
 
@@
 The worst-case latency was brought down from 707.622\textmu s to 17.134\textmu s, closely aligning with the bare metal performance of 10.709\textmu s.
-In addition, the improvement in real-time performance and determinism is validated using a robotic application, where the tuned virtualization is compared with the untuned and the hardware version.
+In addition, the improvement in real-time performance and determinism is validated using a robotic application, where the tuned virtualization is compared with the untuned and the bare metal version.
 

### 2024-08-31 Masterarbeit_Robotik_Pamuk.tex
Correct grammar

@@
 \hypersetup{
-    colorlinks=true, % Aktiviert farbige Links
+    colorlinks=false, % Aktiviert farbige Links
     linkcolor={rgb,255:red,0;green,0;blue,255}, % Interne Links (z.B. Sections)
@@
 Diese Modifikationen umfassen Konfigurationen, die das BIOS, den Kernel, das Host-Betriebssystem, die QEMU/KVM-Virtualisierungsschicht und das Salamander 4-Betriebssystem (Gast) selbst betreffen.
-Die Worst-Case-Latenz wurde von 707.622\textmu s auf 17.134\textmu s reduziert, was der Bare-Metal-Performance von 10.709\textmu s nahekommt.
+Die Worst-Case-Latenz wurde von 707.622~µs auf 17.134~µs reduziert, was der Bare-Metal-Performance von 10.709~µs nahekommt.
 Darüber hinaus wird die Verbesserung der Echtzeit-Performance und des Determinismus anhand einer Roboteranwendung validiert, bei der die abgestimmte Virtualisierung mit der nicht abgestimmten und der Bare-Metal-Version verglichen wird.
 
-\bigskip \noindent Insgesamt bietet diese Arbeit eine umfassende Blaupause, um ein virtualisiertes Gastsystem in einem Host-System mit deterministischem Verhalten echtzeitfähig zu machen.}
+\bigskip \noindent Insgesamt bietet diese Arbeit eine umfassende Methodik, um ein virtualisiertes Gastsystem in einem Hostsystem mit deterministischem Verhalten echtzeitfähig zu machen.}
 \schlagworte{Virtualisierung, Echtzeitsysteme, Latenzreduktion, Robotiksteuerung}
@@
 These modifications involve configurations spanning the BIOS, kernel, host operating system, QEMU/KVM virtualization layer, and the Salamander 4 operating system (guest) itself.
-The worst-case latency was brought down from 707.622\textmu s to 17.134\textmu s, closely aligning with the bare metal performance of 10.709\textmu s.
+The worst-case latency was brought down from 707.622~µs to 17.134~µs, closely aligning with the bare metal performance of 10.709~µs.
 In addition, the improvement in real-time performance and determinism is validated using a robotic application, where the tuned virtualization is compared with the untuned and the bare metal version.
@@
 
-\noindent However, in an RTOS, a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation, as shown in Figure~\ref{fig:kernel_rt} RTOS are specifically designed to react to events within fixed time limits and prioritise the execution of high-priority processes.
+\noindent However, in an RTOS, a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation, as shown in Figure~\ref{fig:kernel_rt}.
+An RTOS is specifically designed to react to events within fixed time limits and prioritize the execution of high-priority processes.
 
@@
 \section{Real-Time Operating Systems}
-\noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.The core component of an RTOS that enables real-time capabilities is the kernel.
+\noindent The RTOS structure is visualized in Figure~\ref{fig:rtos_structure}.
+The core component of an RTOS that enables real-time capabilities is the kernel.
 It is responsible for managing system resources, scheduling tasks, and ensuring deterministic behavior~\cite{malallahComprehensiveStudyKernel2021}.
@@
 
-\noindent In these real-time operating system, task scheduling is based on so-called priority-based preemptive scheduling~\cite{buttazzoHardRealtimeComputing2024}.
+\noindent In these real-time operating systems, task scheduling is based on so-called priority-based preemptive scheduling~\cite{buttazzoHardRealtimeComputing2024}.
 Each task in a software application is assigned a priority.
@@
 	\item Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open-source project that helps develop custom Linux-based systems. Yocto is discussed in Section~\ref{sec:yocto}.
-	\item Salamander 4 is virtualized through Quick emulator (QEMU), an open-source emulator and virtualizer that allows to run different operating systems on a computer QEMU is described in Section~\ref{sec:qemu}.
+	\item Salamander 4 is virtualized through Quick emulator (QEMU), an open-source emulator and virtualizer that allows running different operating systems on a computer.
+QEMU is described in Section~\ref{sec:qemu}.
 	\item Salamander 4 employs hard real-time with Xenomai 3~\cite{XenomaiXenomai}, a real-time framework for Linux that enables deterministic and low-latency performance.
 Xenomai 3 is detailed in Section~\ref{sec:xenomai}.
-	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal (hardware) through real-time performance tunings, focussing on latency requirements set by SIGMATEK.
-These modifications are subject of Chapter~\ref{sec:real-time_tunings}.
-	\item The \texttt{latency} tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to use by SIGMATEK.
-The goal was set to achieving latency values below 50 microseconds in the duration of the measuremenet.
+	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal (hardware) through real-time performance tunings, focusing on latency requirements set by SIGMATEK.
+These modifications are the subject of Chapter~\ref{sec:real-time_tunings}.
+	\item The \texttt{latency} tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to be used by SIGMATEK.
+The goal was set to achieve latency values below 50 microseconds in the duration of the measurement.
 	\item SIGMATEK uses the VARAN bus for communication between machines and systems, a real-time Ethernet network that was specially developed for industrial automation.
@@
 Subsequently, the PCV 522 module of SIGMATEK was connected with the Pulse Width Module PW 022 of SIGMATEK over the VARAN bus and the VI 021 module of SIGMATEK serving as the power supply.
-	\item The robotic application from Chapter~\ref{sec:robotic_application} was written in Lasal Class 2, an object oriented programming tool of SIGMATEK and part of SIGMATEK`s Engineering Platform LASAL, with client-server technology and graphic representation.
+	\item The robotic application from Chapter~\ref{sec:robotic_application} was written in Lasal Class 2, an object-oriented programming tool of SIGMATEK and part of SIGMATEK's Engineering Platform LASAL, with client-server technology and graphic representation.
 \end{itemize}
+
 \section{Related Work and State of the Art}
-This master’s thesis builds upon a number of previous studies in the field of virtualization of real-time systems and the inherent latency.
+This master's thesis builds upon a number of previous studies in the field of virtualization of real-time systems and the inherent latency.
 In this section, the most influential studies are presented and their relevance to the work is addressed.
 
-\bigskip \noindent \citeauthor{perneelRealtimeCapabilitiesStandard2015}~\cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general purpose operating system without real-time applications in mind.
+\bigskip \noindent \citeauthor{perneelRealtimeCapabilitiesStandard2015}~\cite{perneelRealtimeCapabilitiesStandard2015} mention in their work that Linux was initially developed as a general-purpose operating system without real-time applications in mind.
 They talk about the evolution of real-time behavior in the standard Linux kernel and state that it has recently gained popularity among the real-time community, largely due to its open-source nature and stability.
@@
 They concluded with their Linux build and a testing application that Linux can be a viable option for real-time applications, if it is correctly configured.
-\citeauthor{reghenzaniRealTimeLinuxKernel2020}~\cite{reghenzaniRealTimeLinuxKernel2020} provide an extensive overview on the real-time Linux kernel research, specifically on the PREEMPT RT patch.
+\citeauthor{reghenzaniRealTimeLinuxKernel2020}~\cite{reghenzaniRealTimeLinuxKernel2020} provide an extensive overview of the real-time Linux kernel research, specifically on the PREEMPT RT patch.
 \citeauthor{adamRealTimePerformanceResponse2021}~\cite{adamRealTimePerformanceResponse2021} applied the PREEMPT\_RT patch on both Raspberry Pi 3 and BeagleBone Black and achieved significantly lower latencies compared to standard Linux kernels.
@@
 Similarly, \citeauthor{sandstromVirtualizationTechnologiesEmbedded2013}~\cite{sandstromVirtualizationTechnologiesEmbedded2013}, and~\citeauthor{taccariEmbeddedRealTimeVirtualization2014}~\cite{taccariEmbeddedRealTimeVirtualization2014} discuss and compare the current state-of-the-art in virtualization technologies with a focus on embedded real-time platforms.
-The latter emphasizes, that research on real-time systems is focusing on virtualization and multicore scheduling and underlines that flexibility and cost are key metrics.
+The latter emphasizes that research on real-time systems is focusing on virtualization and multicore scheduling and underlines that flexibility and cost are key metrics.
 Both works acknowledge the great benefits of virtualization, that is, reducing overall hardware costs, since all non-critical activities and real-time control tasks can run on the same hardware.
@@
 
-\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased2013}~\cite{maPerformanceTuningKVMbased2013}~\cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts the real-time performance, which has helped this work gain insight into the base overhead of virtualization.
+\bigskip \noindent \citeauthor{maPerformanceTuningKVMbased2013}~\cite{maPerformanceTuningKVMbased2013}~\cite{junzhangPerformanceAnalysisKVMBased2010} describe in their work how KVM impacts real-time performance, which has helped this work gain insight into the base overhead of virtualization.
 When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS.
-Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), since there is also the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started.
+Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), which is the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started.
 These stages are the primary sources of latencies caused by KVM, they are briefly described in Figure~\ref{fig:kvm_stages}.
+
 \vspace{2em}
+
 \begin{figure}[H]
@@
 The authors mention that there is no user interface for setting process CPU affinity in the standard Linux, while there is one for setting interrupt CPU affinity via the \texttt{/proc/irq/*/smp\_affinity} files.
-These files were used in this masther's thesis.
-According to the authors, the shielded CPU model is especially useful for tasks that require guaranteed interrupt response times, very fast interrupt responses and high-frequency, deterministic execution.
+These files were used in this master's thesis.
+According to the authors, the shielded CPU model is especially useful for tasks that require guaranteed interrupt response times, very fast interrupt responses, and high-frequency, deterministic execution.
 
@@
 \bigskip \noindent \citeauthor{yoonRealTimePerformanceAnalysis2009}~\cite{yoonRealTimePerformanceAnalysis2009} use some of these real-time tunings in their research, including CPU shielding, memory locking, and spinning nanosleep, for controlling humanoid robots equipped with around 60 servo motors and sensors.
-They implement EtherCAT for real-time communication of distributed devices whereas this thesis uses Varan.
+They implement EtherCAT for real-time communication of distributed devices, whereas this thesis uses Varan.
 The authors highlight the importance of timely data processing in robots that collect data from their environment and respond accordingly through their sensors and actuators.
@@
 Their research has similarities to this master's thesis in terms of real-time tunings for robot control.
-However, this thesis includes also the virtualization aspect that is central.
+However, this thesis also includes the virtualization aspect that is central.
 
 \bigskip \noindent \citeauthor{kiszkaLinuxRealTimeHypervisor2009}~\cite{kiszkaLinuxRealTimeHypervisor2009} specifically focuses on Linux as a hypervisor and analyzes its real-time capabilities when using KVM and QEMU.
-Most importantly for this master's thesis, the author warns of starvation when the priorities of QEMU’s threads are raised and lifted into a real-time scheduling class.
+Most importantly for this master's thesis, the author warns of starvation when the priorities of QEMU's threads are raised and lifted into a real-time scheduling class.
 Starvation is when a process does not get the resources it needs for a long time because the resources are being allocated to other processes.
 A way to reduce this risk is by not using more virtual CPUs (vCPUs) than there are actual processor cores.
-Moreover, Linux has a feature that can limit how much CPU time all real-time tasks can use in the host system, which was an important information for the real-time tunings of this thesis later.
+Moreover, Linux has a feature that can limit how much CPU time all real-time tasks can use in the host system, which was important information for the real-time tunings of this thesis later.
 
@@
 The author lists several sources of overhead, including memory locking to avoid page-fault latencies, increased overhead of locking and interrupts due to more-aggressive preemption, threaded interrupts that permit long-running interrupt handlers to be preempted by high-priority real-time processes, real-time task scheduling that requires global scheduling, high-resolution timers for tens-of-microseconds accuracy and precision, and preemptible RCU which has slightly higher read-side overhead than Classic RCU.
-These information helped understand the trade-offs involved in real-time Linux systems.
+This information helped understand the trade-offs involved in real-time Linux systems.
 
 \bigskip \noindent \citeauthor{deoliveiraDemystifyingRealTimeLinux2020}~\cite{deoliveiraDemystifyingRealTimeLinux2020} critique the traditional \texttt{cyclictest} tool used for measuring scheduling latency, highlighting its limitations due to its black-box approach and lack of theoretical grounding.
-This is the main reason why this tool was not used to measure the scheduling latency in this work.
+This contributed to the decision not to use this tool to measure the scheduling latency in this work.
 Instead, the \texttt{latency} tool was used, as was determined in Chapter~\ref{sec:application_context}.
@@
 Moreover, during periods when the demand is lower, expensive hardware may sit idle, leading to inefficient use of resources and a poor return on investment~\cite{gabrielResourceAwareParameterTuning2019}.
-Another crucial factor is of course the high cost factor associated with procuring and maintaining separate hardware for each real-time system~\cite{bhardwajVirtualizationCloudComputing2021}.
+Another crucial factor is, of course, the high cost factor associated with procuring and maintaining separate hardware for each real-time system~\cite{bhardwajVirtualizationCloudComputing2021}.
 Especially when organizations scale their operations, they have to add more physical machines, and the expenses rise rapidly~\cite{abbasiExploringOpenStackScalable2023}.
@@
 This is not desired and can be particularly problematic for real-time systems that require strict timing constraints.
-Making sure that tasks run predictably and on time in a virtualized environment, can be challenging.
+Making sure that tasks run predictably and on time in a virtualized environment can be challenging.
 The hypervisor and other virtual components can cause variations in how long tasks take to execute~\cite{garcia-vallsChallengesRealtimeVirtualization2014}.
@@
 Chapter~\ref{cha:methodology} explains in detail the procedure, how this was tackled.
-As already established in Section~\ref{sec:application_context}, Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3 and is virtualized through QEMU/KVM.
+As already established in Section~\ref{sec:application_context}, Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3, and is virtualized through QEMU/KVM.
 The initial latency values are measured with the \texttt{latency} program of the Xenomai tool suite.
@@
 \noindent For the sake of completeness, Figure~\ref{fig:memory_management} displays the memory management of Salamander 4.
-LRT stands for Lasal Runtime and creates an execution environment where applications developed using LASAL Class 2 can run- It provides real-time functions, data types, and other constructs for real-time programming.
+LRT stands for Lasal Runtime and creates an execution environment where applications developed using LASAL Class 2 can run.
+It provides real-time functions, data types, and other constructs for real-time programming.
 
@@
 
-QEMU~\cite{QEMU}.
-is an open-source emulator and virtualizer that allows to run operating systems and programs for one machine on a different machine.
+QEMU~\cite{QEMU} is an open-source emulator and virtualizer that allows running operating systems and programs for one machine on a different machine.
 It can emulate different hardware architectures and is used in this thesis to emulate Salamander 4.
@@
 
-\noindent The inital QEMU script after the custom Yocto build and the starting point for this work is shown in Code~\ref{script:qemu_def}.
-This script is used to start QEMU with required parameters to boot Salamamder 4 OS.
+\noindent The initial QEMU script after the custom Yocto build and the starting point for this work is shown in Code~\ref{script:qemu_def}.
+This script is used to start QEMU with required parameters to boot Salamander 4 OS.
 It will be adjusted in Chapter~\ref{sec:real-time_tunings} in order to accompany real-time performance tunings.
@@
 
-\noindent After compiling the kernel with the additional settings aboce, the guest needs access to the vsocket.
+\noindent After compiling the kernel with the additional settings above, the guest needs access to the vsocket.
 In QEMU, the line \texttt{-device vhost-vsock-pci,guest-cid=3,id=vsock0} was added.
@@
 For example, \texttt{sudo trace-cmd record -e all -A @3:823 --name Salamander4 -e all} records all events from the host and guest.
-\texttt{sudo trace-cmd record -e kvm:kvm\_entry -e kvm:kvm\_exit -A @3:823 --name Salamander4 -e all} traces onkly the \texttt{kvm\_entry} and \texttt{kvm\_exit} events of the host along with the exit reasons and all guest events.
+\texttt{sudo trace-cmd record -e kvm:kvm\_entry -e kvm:kvm\_exit -A @3:823 --name Salamander4 -e all} traces only the \texttt{kvm\_entry} and \texttt{kvm\_exit} events of the host along with the exit reasons and all guest events.
 As a last example, \texttt{sudo trace-cmd record -e kvm -e sched -e irq -e -A @3:823 --name Salamander4 -e all} enables tracing of KVM, scheduling, and interrupt events on the host, for capturing the KVM events that manage the guest, scheduling activities, and interrupt handling respectively.
@@
 It visualizes the recorded kernel trace data in a readable way on an interactive timeline, which facilitates the process of identifying patterns and correlations between events.
-The latency issues can be analyzed by further filtering the displayed events according to processes, event types or time ranges.
-After the recording, \texttt{trace-cmd} gernerates a host \texttt{trace.dat} file and a guest \texttt{trace-Salamander4.dat} file, which in this case can be loaded into \texttt{KernelShark} with the command \texttt{kernelshark trace.dat -a trace-Salamander4.dat}.
+The latency issues can be analyzed by further filtering the displayed events according to processes, event types, or time ranges.
+After the recording, \texttt{trace-cmd} generates a host \texttt{trace.dat} file and a guest \texttt{trace-Salamander4.dat} file, which in this case can be loaded into \texttt{KernelShark} with the command \texttt{kernelshark trace.dat -a trace-Salamander4.dat}.
 Once both files are loaded, the trace data can be visually inspected.
@@
 The jitter between these systems is kept below 100 ns.
-This synchronization between systens and their data exchange is illustrated in Figure~\ref{fig:multiple_varans}
+This synchronization between systems and their data exchange is illustrated in Figure~\ref{fig:multiple_varans}.
 
@@
 \section{Approach}\label{sec:approach}
-The goal is to The \texttt{latency} program is first executed in the bare metal Salamander 4 OS.
 The goal is to virtualize the Salamander 4 OS and bring its real-time performance closer to that of the hardware version and guarantee deterministic and reliable behavior.
-For that purpose, the first step is to apply the PREEMPT-RT patch to the host system and subsequently tune the real-time performance through BIOS, kernel, host, guest and QEMU/KVM configurations.
+For that purpose, the first step is to apply the PREEMPT-RT patch to the host system and subsequently tune the real-time performance through BIOS, kernel, host, guest, and QEMU/KVM configurations.
 These modifications are added sequentially and tested together.
@@
 The key metric here is the latency of the system, and the real-time performance is evaluated using the \texttt{latency} tool of the Xenomai suite after each configuration step.
-At the end, the results are domenstarted alltogether and a holistic comparison and discussion is carried out.
+At the end, the results are demonstrated altogether and a holistic comparison and discussion are carried out.
 In addition, the improvement in real-time performance and determinism is demonstrated using a robotic application, where the modified virtualization is compared with the hardware version.
@@
 As a reference point, the \texttt{latency} program was executed on Salamander 4 bare metal for a duration of 10 minutes.
-The complete command used was~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of minimum, average and maximum latencies in a Gnuplot-compatible format to the file~\texttt{gnuplot.txt}.
+The complete command used was~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of minimum, average, and maximum latencies in a Gnuplot-compatible format to the file~\texttt{gnuplot.txt}.
 Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the distribution of latency over the course of said time.
@@
 \noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_hardware} and Table~\ref{tab:latency_overrun_msw_hardware}.
-The test was conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
-If any sample's latency value is greater than 100\textmu s, it would be considered an overrun.
+The test was conducted with a sampling period of 100~µs, using a periodic user-mode task, and was assigned a priority of 99.
+If any sample's latency value is greater than 100~µs, it would be considered an overrun.
 
@@
 		\textbf{min}   & 599              & 0.711                 & 0.454                 \\ \hline
-		\textbf{avg}   & 5 999 988        & 1.019                 & 0.150                 \\ \hline
+		\textbf{avg}   & 5,999,988        & 1.019                 & 0.150                 \\ \hline
 		\textbf{max}   & 599              & 3.528                 & 0.895                 \\ \hline
@@
 \subsection{Salamander 4 Virtualization}\label{subsec:salamander4-virtualization}
-In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system.
+In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualized version of this operating system.
 Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, Figure~\ref{fig:gnuplot_max_latency_default} depicts the distribution of latency.
@@
 \noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt} and Table~\ref{tab:latency_overrun_msw_new_virt}.
-The test was again conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
+The test was again conducted with a sampling period of 100~µs, using a periodic user-mode task, and was assigned a priority of 99.
 
@@
 \noindent Comparing the latency values of the virtualization to those of bare metal in Figure~\ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in the statistics.
-A maximum latency of 707.62~\textmu s is not tolerable for a real-time virtualization and needs to be tuned.
+A maximum latency of 707.62~µs is not tolerable for a real-time virtualization and needs to be tuned.
 
@@
 This involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
-The individual configurations are discussed in detail and the modifications are justified with a clear explanation.
+The individual configurations are discussed in detail, and the modifications are justified with a clear explanation.
 The goal is to bring the latency of the virtualization closer to that of the bare metal, thereby ensuring deterministic behavior under real-time constraints.
@@
 \bigskip \noindent It is important to mention that a real-time system implies determinism at every component of the system, ranging from hardware over the kernel to the software and the application on it.
-As already established earlier, the guest OS Salamamder 4 runs Xenomai and is hence real-time capable.
-The host system also needs to be aware of the real-time determinism in order to achieve highest possible reliability.
-The very step of achieving the goal of reducing latency is to apply the PREEMPT-RT patch to the host system.
+As already established earlier, the guest OS Salamander 4 runs Xenomai and is hence real-time capable.
+The host system also needs to be aware of the real-time determinism in order to achieve the highest possible reliability.
+The very first step of achieving the goal of reducing latency is to apply the PREEMPT-RT patch to the host system.
 This patch is a set of modifications to the Linux kernel with the goal of making it fully preemptible.
-This means, almost all parts of the kernel can be preempted and higher-priority tasks can interrupt lower-priority ones.
+This means almost all parts of the kernel can be preempted, and higher-priority tasks can interrupt lower-priority ones.
 This significantly reduces the time it takes for high-priority tasks to start executing after an event~\cite{RealtimeKernelPatchset}.
@@
 After each tuning, the \texttt{latency} test of Xenomai is executed to evaluate the improved latency.
-The most important metric of a latency test is without a doubt the worst latency value, because the system is only as reliable as its worst measurement.
-Since the period of the tests is set to 100\textmu s, there are approximately 1,000,000 samples per second.
+The most important metric of a latency test is, without a doubt, the worst latency value, because the system is only as reliable as its worst measurement.
+Since the period of the tests is set to 100~µs, there are approximately 1,000,000 samples per second.
 
@@
 Table~\ref{tab:bios_configuration} illustrates the specific BIOS settings that have been adjusted for the purpose of real-time performance.
-As an important note, this is not a definitive, full-ended list, and other devices may have additional settings that can be modified to achieve even better latency.
+As an important note, this is not a definitive, complete list, and other devices may have additional settings that can be modified to achieve even better latency.
 
@@
 \begin{itemize}
-	\item \textbf{Hyper Threading}: When hyper-threading is enabled in thd BIOS settings, this allows CPUs to work on two threads simultaneously instead of just one.
-This allows the parallelisation of tasks and seemingly increases performance.
+	\item \textbf{Hyper Threading}: When hyper-threading is enabled in the BIOS settings, this allows CPUs to work on two threads simultaneously instead of just one.
+This allows the parallelization of tasks and seemingly increases performance.
 However, in a real-time system like the guest Salamander 4, this can lead to increased latencies due to contention between threads.
@@
 
-\noindent After these first real-time tunings in the BIOS settings, the \texttt{latency} test was run to see, whether this had any significant impact on the determinism of the system.
+\noindent After these first real-time tunings in the BIOS settings, the \texttt{latency} test was run to see whether this had any significant impact on the determinism of the system.
 Figure~\ref{fig:gnuplot_max_latency_taskset} shows the latency values after tuning BIOS settings.
@@
 
-\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt} and Table~\ref{tab:latency_overrun_msw_new_virt}.
+\noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt_bios} and Table~\ref{tab:latency_overrun_msw_new_virt_bios}.
 
@@
 
-\noindent The maximum latency value recorded was 457.545\textmu s.
-When compared to the inital worst latency of 707.62 \textmu s, this is a step for the better, but certainly still not acceptable within this work's definition of hard real-time determinism.
+\noindent The maximum latency value recorded was 457.545~µs.
+When compared to the initial worst latency of 707.62~µs, this is a step for the better, but certainly still not acceptable within this work's definition of hard real-time determinism.
 
@@
 This parameter offloads RCU callback handling from CPU 4 to other CPUs.
-CPU 4 remains dedicated to high-priority tasks which helps in reducing latency.
+CPU 4 remains dedicated to high-priority tasks, which helps in reducing latency.
 	\item \textbf{rcu\_nocb\_poll}: This is used together with \texttt{rcu\_nocbs} and causes the system to actively poll for RCU callbacks to invoke, instead of waiting for the next RCU grace period.
@@
 It allows these devices to directly access and use memory, which is especially helpful when they are virtualized.
-	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel’s RDT (Resource Director Technology).
+	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel's RDT (Resource Director Technology).
 Unlike L1 and L2 caches, where each core has its fixed capacity, L3 cache is a shared pool among multiple cores.
@@
 This constant monitoring consumes CPU cycles and can introduce undesirable latency in real-time systems.
-	\item \textbf{idle=poll}: Changes the CPU’s idle loop behavior to active polling.
+	\item \textbf{idle=poll}: Changes the CPU's idle loop behavior to active polling.
 Instead of entering a low-power state when idle, the CPU continuously polls for new tasks.
@@
 When it is enabled, it generates log entries for security-relevant events, which is a slow operation since they are written to disk.
-If there are a large number of such events, the audit system can consume significant CPU time and I/O bandwidth which could lead to higher latency.
+If there are a large number of such events, the audit system can consume significant CPU time and I/O bandwidth, which could lead to higher latency.
 	\item \textbf{skew\_tick=1}: Enables a mode in the Linux kernel that reduces timer interrupt overhead.
-Normally, timer interrupts happen simultaneously on all CPUs, resulting in all CPUs to exit their low-power states at once.
+Normally, timer interrupts happen simultaneously on all CPUs, resulting in all CPUs exiting their low-power states at once.
 This can lead to increased contention for system resources.
@@
 In real-time systems, these wake-up delays can be problematic.
-Disabling them leads keeps the CPUs ready to respond quickly to new tasks and helps in reducing latency.
+Disabling them keeps the CPUs ready to respond quickly to new tasks and helps in reducing latency.
 	\item \textbf{nosoftlockup}: Disables the soft lockup detector in the Linux kernel.
@@
 
-\noindent After the kernel modifications on top of the BIOS settings from the last subsection, Figure~\ref{fig:gnuplot_max_latency_rt_kernelparam} demonstrates the latency
-values gathered in 10 minutes.
+\noindent After the kernel modifications on top of the BIOS settings from the last subsection, Figure~\ref{fig:gnuplot_max_latency_rt_kernelparam} demonstrates the latency values gathered in 10 minutes.
 
@@
 
-\noindent Apparent is a significant improvement in the latency of the virtualized Salamander 4 OS. With a worst latency value of 21.694\textmu s, the real-time tunings appear to be working as desired.
+\noindent Apparent is a significant improvement in the latency of the virtualized Salamander 4 OS. With a worst latency value of 21.694~µs, the real-time tunings appear to be working as desired.
 
@@
 In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest Salamander 4 OS is provided.
-\subsubsection{CPU affinity and Isolation}\label{subsubsec:cpu_isolation}
-
+
+\subsubsection{CPU Affinity and Isolation}\label{subsubsec:cpu_isolation}
 Isolating CPUs means removing all user-space threads and unbound kernel threads, since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-For CPU isolation, the~\texttt{isolcpus} function is used to isolate a performance CPU from the general scheduling algorithms of the operating system.
+For CPU isolation, the \texttt{isolcpus} function is used to isolate a performance CPU from the general scheduling algorithms of the operating system.
 This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks.
-However, the isolcpus function only isolates at the user level and does not affect kernel tasks.
+However, the \texttt{isolcpus} function only isolates at the user level and does not affect kernel tasks.
 Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased2013}, including systemd services.
@@
 After the isolation, user tasks other than the QEMU process have been removed from running on this CPU.
-Only few per-CPU kernel threads that are tied to this CPU still take CPU time.
+Only a few per-CPU kernel threads that are tied to this CPU still take CPU time.
 
@@
 \subsubsection{Interrupt Affinity}\label{subsubsec:irq_handling}
-Once the CPUs were isolated, interrupt requests handling was the next step.
+Once the CPUs were isolated, interrupt request handling was the next step.
 The purpose of interrupt requests is to inform the CPU to stop working on a certain job and start working on another.
 This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems.
-All the existing interrupts can be monitored using the command~\texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
-The IRQs needed to be removed from the isolated CPU by manipulating the~\texttt{/proc/irq/<IRQ>/smp\_affinity} files.
-The value in the~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU.
+All the existing interrupts can be monitored using the command \texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
+The IRQs needed to be removed from the isolated CPU by manipulating the \texttt{/proc/irq/<IRQ>/smp\_affinity} files.
+The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU.
 The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU5).
-When there are 6 CPUs available, the default value for~\texttt{smp\_affinity} would be 3F (11 1111).
+When there are 6 CPUs available, the default value for \texttt{smp\_affinity} would be 3F (11 1111).
 Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 2F (10 1111).
-The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU, since the \texttt{/proc/interrupts} file does not give away this information and looking individually in the \texttt{/proc/irq/<IRQ>/smp\_affinity} files for every IRQ would be time-consuming.
+The Python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU, since the \texttt{/proc/interrupts} file does not give away this information and looking individually in the \texttt{/proc/irq/<IRQ>/smp\_affinity} files for every IRQ would be time-consuming.
 The program first reads the CPU affinity settings for each IRQ from the \texttt{/proc/irq} directory and stores the CPUs assigned to each IRQ in a dictionary.
-It then creates a data frame to represent this information in a table format, marking the assigned CPUs with ‘x’.
+It then creates a data frame to represent this information in a table format, marking the assigned CPUs with ‘x'.
 Finally, it prints the table and saves it as a markdown file.
 
-\bigskip \noindent However, as the \texttt{proc} filesystem resets to its default state after each reboot, manually changing numerous IRQ files through the said bitbasks is tedious and time-consuming.
-This process can be was automated using the shell script in Code~\ref{script:change_smp_affinity}, which can automatically be executed after every reboot.
+\bigskip \noindent However, as the \texttt{proc} filesystem resets to its default state after each reboot, manually changing numerous IRQ files through the said bitmasks is tedious and time-consuming.
+This process can be automated using the shell script in Code~\ref{script:change_smp_affinity}, which can automatically be executed after every reboot.
 
@@
 Having a real-time kernel itself is an indispensable step of achieving deterministic behavior, but not enough by itself to take full advantage of the real-time capabilities.
-One key aspect of this are real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}.
+One key aspect of this is real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}.
 In essence, Table~\ref{tab:scheduling_priorities} lists minimum and maximum priorities for different scheduling policies in Linux.
@@
 	\centering
-	\caption{Minimum and maximum Priorities for different Scheduling Policies}
+	\caption{Minimum and Maximum Priorities for Different Scheduling Policies}
 	\label{tab:scheduling_priorities}
@@
 The virtual machine can either be started with \texttt{chrt -f <PRIO>} or adjusted at a later point with \texttt{chrt -f <PRIO> <PID>}.
-It is also important that there are no other unexpected real time processes running on the system concurrently.
+It is also important that there are no other unexpected real-time processes running on the system concurrently.
+
 \subsubsection{Disable RT Throttling}
@@
 It does this by pausing the process for the remaining 0.05 seconds, which is not desired because this could result in missed deadlines.
-RT throttling can be disabled by  writing the value -1 to the~\texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file.
+RT throttling can be disabled by writing the value -1 to the \texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file.
 This change also needs to be made permanent because the proc filesystem resets to its default state after each reboot.
-For this purpose, the line~\texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the~\texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters.
+For this purpose, the line \texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the \texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters.
 This reduces the potential for missed deadlines.
@@
 In a real-time system, this can introduce latency and jitter.
-To disable timer migration, the value ``0'' needs to be written to the~\texttt{/proc/sys/kernel/timer\_migration} file.
-This change also needs to be made permanent by writing the line~\texttt{kernel.timer\_migration = 0} to the~\texttt{/etc/sysctl.conf} file.
+To disable timer migration, the value ``0'' needs to be written to the \texttt{/proc/sys/kernel/timer\_migration} file.
+This change also needs to be made permanent by writing the line \texttt{kernel.timer\_migration = 0} to the \texttt{/etc/sysctl.conf} file.
 This reduces the amount of context switches and interrupts.
+
 \subsubsection{Set Device Driver Work Queue}
@@
 By setting the work queue away from CPU 4 to another CPU, it is free to handle real-time tasks without being interrupted by these work queue tasks.
-This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask} In this case, that is \texttt{2F}.
+This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
+In this case, that is \texttt{2F}.
+
 \subsubsection{Disable RCU CPU Stall Warnings}
@@
 These include a CPU looping in an RCU read-side critical section, a CPU looping with interrupts disabled, a CPU looping with preemption disabled, or a CPU not getting around to less urgent tasks, known as “bottom halves”.
-The number of seconds the kernel should wait before checking for stalled CPUs and reporting a stall warning can be set via the ~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_timeout} file.
-These warnings can be suppressed alltogether to reduce the potential for increased latency by writing ``1'' to the~\texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file.
+The number of seconds the kernel should wait before checking for stalled CPUs and reporting a stall warning can be set via the \texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_timeout} file.
+These warnings can be suppressed altogether to reduce the potential for increased latency by writing ``1'' to the \texttt{/sys/module/rcupdate/parameters/rcu\_cpu\_stall\_suppress} file.
 This setting is also not persistent across reboots, so the command needs to be added to a startup script.
@@
 Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{tab:stop_servies}, can be further sources for random latency and unnecessary overhead.
-Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means, they will not be able to interrupt the CPU with their tasks.
+Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means they will not be able to interrupt the CPU with their tasks.
 
@@
 \end{table}
+
 \subsubsection{Disable Machine Check}
@@
 
-\subsubsection{Boot into text-based environment}
+\subsubsection{Boot into Text-Based Environment}
 The Graphical User Interface (GUI) consumes great amounts of system resources.
@@
 
-\noindent A worst latency value of 17.041\textmu s shows that the virtualized Salamander 4 OS got even more reliable with the host configurations.
+\noindent A worst latency value of 18.441~µs shows that the virtualized Salamander 4 OS got even more reliable with the host configurations.
 
@@
 
-\subsubsection{Set QEMU Options for real-time VM}
+\subsubsection{Set QEMU Options for Real-Time VM}
 QEMU provides several options that can be used to improve the real-time performance of the guest VM. Table~\ref{tab:qemu_options} briefly explains these options.
@@
 \end{lstlisting}
+
 \vspace{2em}
@@
 \noindent While the worst latency value did not change much, the standard deviation of the maximum latency samples of each second, with the test lasting 600 seconds and thus having 600 worst latencies, decreased.
-This means, the system is more stable and predictable, thanks to the real-time tunings that have been performed till this point.
+This means the system is more stable and predictable, thanks to the real-time tunings that have been performed till this point.
 
@@
 Linux Kernel Developer Steven Rostedt explains all relevant aspects of a real-time system that must be considered in~\cite{kernelrecipesKernelRecipes20162016} and gives insight for finding sources of latency on the Linux system in~\cite{thelinuxfoundationFindingSourcesLatency2020}.
-A Checklist for writing Linux real-time applications is provided by John Ogness in~\cite{thelinuxfoundationChecklistWritingLinux2020}.
+A checklist for writing Linux real-time applications is provided by John Ogness in~\cite{thelinuxfoundationChecklistWritingLinux2020}.
 Every layer of the system stack must be deterministic to ensure predictable and reliable latency, including hardware, operating system, middleware and drivers, and the application software.
-~\cite{HOWTOBuildRTapplication} and~\cite{RealtimeProgrammingLinux} describe the process of writing hard real time Linux programs using the real-time preemption patch in great detail.
+~\cite{HOWTOBuildRTapplication} and~\cite{RealtimeProgrammingLinux} describe the process of writing hard real-time Linux programs using the real-time preemption patch in great detail.
 Various hardware and software tunings are listed in~\cite{KVMQemuVirtualization} and~\cite{RealTimePerformanceTuning2022}.
@@
 \end{lstlisting}
+
 \vspace{2em}
+
 \noindent An additional Varan Connection module VI 021~\cite{InterfacesSplittersSIGMATEK} was required to enable the connection between the PCV 522 module and the PW 022 module that generates the signal to move the mini-robot.
@@
 In this chapter, these improvements are presented using a robotic application.
-This application is written in Lasal Class 2, as mentoned in Section~\ref{sec:application_context}.
+This application is written in Lasal Class 2, as mentioned in Section~\ref{sec:application_context}.
 The program sequence is shown in Figure~\ref{fig:robotic_application_flow}.
@@
 
-\begin{comment}
-Ja, du denkst richtig!
-Wenn du die Latenz des Systems messen möchtest, die es benötigt, um das Signal bei der PWM anzukommen, und die 20 ms Signalperiode des Servomotors ignorieren möchtest, dann konzentrierst du dich auf die reine Reaktionszeit des Systems.
-
-In diesem Fall misst du die Zeit von der Befehlsausgabe bis zum Eintreffen des Signals bei der PWM.
-Die 20 ms, die der Servomotor benötigt, um das Signal zu verarbeiten und die Bewegung zu starten, werden in deiner Latenzdefinition nicht berücksichtigt.
-
-Das bedeutet, dass du die Latenz als die Zeit definierst, die das System benötigt, um das Signal an die PWM zu senden, ohne die zusätzliche Verzögerung durch die Signalperiode des Servomotors einzubeziehen.
-
-	In diesem Versuch wird die Latenz des Systems gemessen, die benötigt wird, um das Signal von der Befehlsausgabe bis zum Eintreffen bei der PWM (Pulsweitenmodulation) zu erfassen.
-Die Signalperiode des Servomotors von 20 ms wird dabei ignoriert, da sie für die Definition der Systemlatenz nicht relevant ist.
-Die Latenz wird somit als die reine Reaktionszeit des Systems definiert, ohne die zusätzliche Verzögerung durch die Signalperiode des Servomotors zu berücksichtigen.
-\end{comment}
-
 \clearpage
@@
 \subsubsection{Latency in Salamander 4 Bare Metal}\label{subsubsec:latency_hardware}
-\noindent The latency values for bare metal Salamander 4 OS aquired from the program above are demonstrated in Figure~\ref{fig:rtos_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_hardware}.
+\noindent The latency values for bare metal Salamander 4 OS acquired from the program above are demonstrated in Figure~\ref{fig:rtos_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_hardware}.
 
@@
 
-\subsubsection{Latency in default Salamander 4 OS Virtualization}\label{subsubsec:latency_virtualization}
-\noindent The latency values aquired for the default virtualization of Salamander 4 OS are demonstrated in Figure~\ref{fig:untuned_virt_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_virt_tun}.
+\subsubsection{Latency in Default Salamander 4 OS Virtualization}\label{subsubsec:latency_virtualization}
+\noindent The latency values acquired for the default virtualization of Salamander 4 OS are demonstrated in Figure~\ref{fig:untuned_virt_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_virt_def}.
 
@@
 
-\subsubsection{Latency in tuned Salamander 4 OS Virtualization}\label{subsubsec:latency_virtualization}
-\noindent The latency values aquired for the tuned virtualization of Salamander 4 OS are demonstrated in Figure~\ref{fig:tuned_virt_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_virt_tun}.
+\subsubsection{Latency in Tuned Salamander 4 OS Virtualization}\label{subsubsec:latency_virtualization}
+\noindent The latency values acquired for the tuned virtualization of Salamander 4 OS are demonstrated in Figure~\ref{fig:tuned_virt_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_virt_tun}.
 
@@
 \noindent The improved latency was also tested with the robotic application in Subsection~\ref{sec:robotic_application}.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the bare metal, untuned and tuned version of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
 The results are visualized in Figure~\ref{fig:combined_latencies}.
@@
 The maximum latency in the final tuned version of Salamander 4 Virtualization is significantly lower than in the default setup.
-The worst latency value in a 10 minute measurement span decreased from 707.622\textmu s to 17.134\textmu s after tuning  BIOS, kernel, host, guest and QEMU/KVM configurations.
-Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694\textmu s.
+The worst latency value in a 10-minute measurement span decreased from 707.622~µs to 17.134~µs after tuning BIOS, kernel, host, guest, and QEMU/KVM configurations.
+Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694~µs.
 From this moment on, there were also no overruns in the test any longer.
-The host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU further reduced latency a little bit down to 18.441\textmu s.
+The host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU, further reduced latency a little bit down to 18.441~µs.
 The guest OS, Salamander 4, already had real-time capabilities through Xenomai.
 The focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
-Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages brought the latency down a little more to the final worst latency value of 17.134\textmu s in 10 minutes.
-This value is very close to the bare metal value of 10.709\textmu s.
-The goal was set to achieving latency values below 50 microseconds in the duration of the measuremenet.
-This goal is acheived.
+Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages, brought the latency down a little more to the final worst latency value of 17.134~µs in 10 minutes.
+This value is very close to the bare metal value of 10.709~µs.
+The goal was set to achieving latency values below 50 microseconds in the duration of the measurement.
+This goal is achieved.
 
 \bigskip \noindent As a next step, the robotic application in a practical scenario also demonstrates and validates the improvement of the tuned virtualization compared to the unmodified virtualization.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and hardware version of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and hardware versions of Salamander 4.
 The most important metric here is also the worst latency of the signal reaching the PWM.
-It dropped from 129 ms to 3.988 ms in the developed application.
-Comparing this to the worst latency value of the hardware version, which is 1.49 ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
+It dropped from 129~ms to 3.988~ms in the developed application.
+Comparing this to the worst latency value of the hardware version, which is 1.49~ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
 
@@
 Specifically, the aim was to bring the latency of the virtualized Salamander 4 closer to that of the bare metal version, thereby providing a comprehensive blueprint for making a virtualized guest system in a host system real-time capable with deterministic behavior.
-Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3 and is virtualized through QEMU/KVM.
+Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3, and is virtualized through QEMU/KVM.
 The testing robot was connected to the OS via a VARAN bus interface.
@@
 \bigskip \noindent The initial latency values were first measured with the \texttt{latency} program of the Xenomai tool suite.
-The tests were conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
-If any sample's latency value was greater than 100µs, this was considered an overrun.
+The tests were conducted with a sampling period of 100 µs, using a periodic user-mode task, and were assigned a priority of 99.
+If any sample's latency value was greater than 100~µs, this was considered an overrun.
 There was a significant initial gap in latency statistics between the virtualized Salamander 4 and the Salamander 4 on bare metal.
@@
 The BIOS settings were not reverted when moving to the kernel configurations.
-After these two steps, the worst latency decreased from an initial value of 707.62~\textmu s to 21.694\textmu s.
+After these two steps, the worst latency decreased from an initial value of 707.62~µs to 21.694~µs.
 From this moment on, there were also no overruns in the test any longer.
-The host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU further reduced latency a little bit down to 18.441\textmu s.
+The host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU, further reduced latency a little bit down to 18.441~µs.
 The guest OS, Salamander 4, already had real-time capabilities through Xenomai.
 The focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
-Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages brought the latency down a little more to the final worst latency value of 17.134\textmu s in 10 minutes.
-This value is very close to the bare metal value of 10.709\textmu s.
-The goal was set to achieving latency values below 50 microseconds in the duration of the measuremenet.
-This goal was acheived.
+Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages, brought the latency down a little more to the final worst latency value of 17.134~µs in 10 minutes.
+This value is very close to the bare metal value of 10.709~µs.
+The goal was set to achieving latency values below 50 microseconds in the duration of the measurement.
+This goal was achieved.
 
 \bigskip \noindent On top of that, a robotic application further confirmed the improvements.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
-The worst latency of the signal reaching the PWM dropped from an inital value of 129ms to 3.988ms.
-Comparing this to the worst latency value of the hardware version, which was 1.49ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and hardware versions of Salamander 4.
+The worst latency of the signal reaching the PWM dropped from an initial value of 129~ms to 3.988~ms.
+Comparing this to the worst latency value of the hardware version, which was 1.49~ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
 
@@
 Other than that, the use of other hypervisors and virtualization technologies can be investigated to compare their performance for real-time applications.
-Moreover, extensive testing of the virtualized system under various workloads can be conducted to evaluate its performance and reliability in different conditions and stressed situations, as demonstrated by \citeauthor{huang2015performance}~\cite{huang2015performance}, \citeauthor{kirovaImpactModernVirtualization2019}~\cite{kirovaImpactModernVirtualization2019} or \citeauthor{adamPerformanceAssessmentLinux2021}~\cite{adamPerformanceAssessmentLinux2021}.
-
-\begin{comment}
-~\cite{canbazPerformanceAnalysisRealtime2022}
-~\cite{WhatRealtimeLinux}
-~\cite{malallahComprehensiveStudyKernel2021}
-~\cite{wangRealtimeEmbeddedSystems2017}
-~\cite{buttazzoHardRealtimeComputing2024}
-~\cite{lipariRealTimeSchedulingHard2015}
-~\cite{amarpreetHardRealTime2009}
-~\cite{queirozTestingLimitsGeneralpurpose2023}
-~\cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}
-~\cite{WelcomeYoctoProject}
-~\cite{XenomaiXenomai}
-~\cite{perneelRealtimeCapabilitiesStandard2015}
-~\cite{reghenzaniRealTimeLinuxKernel2020}
-~\cite{adamRealTimePerformanceResponse2021}
-~\cite{cinqueVirtualizingMixedCriticalitySystems2022}
-~\cite{sandstromVirtualizationTechnologiesEmbedded2013}
-~\cite{taccariEmbeddedRealTimeVirtualization2014}
-~\cite{javierperezHowRealTime2022}
-~\cite{guStateoftheArtSurveyRealTime2012}
-~\cite{garcia-vallsChallengesRealtimeVirtualization2014}
-~\cite{scordinoRealTimeVirtualizationIndustrial2020}
-~\cite{maPerformanceTuningKVMbased2013}
-~\cite{junzhangPerformanceAnalysisKVMBased2010}
-~\cite{broskyShieldedProcessorsGuaranteeing2003}
-~\cite{RealTimePerformanceTuning2022}
-~\cite{yoonRealTimePerformanceAnalysis2009}
-~\cite{kiszkaLinuxRealTimeHypervisor2009}
-~\cite{mckenneyRealTimeVs2008}
-~\cite{deoliveiraDemystifyingRealTimeLinux2020}
-~\cite{gabrielResourceAwareParameterTuning2019}
-~\cite{bhardwajVirtualizationCloudComputing2021}
-~\cite{abbasiExploringOpenStackScalable2023}
-~\cite{mahfoudRealTimePredictiveMaintenanceBased2024}
-~\cite{queirozContainerbasedVirtualizationRealtime2024}
-~\cite{cinqueEvaluatingVirtualizationFog2023}
-~\cite{casiniLatencyAnalysisVirtualization2021}
-~\cite{zhangEvaluatingOptimizingVirtualization2010}
-~\cite{CPUUnitsSIGMATEK}
-~\cite{EngineeringToolLASAL}
-~\cite{LASALOSSIGMATEK}
-~\cite{QEMU}
-~\cite{OverviewXenomai}
-~\cite{RealtimePreempt_rt_versionsWiki}
-~\cite{LATENCY}
-~\cite{Tracecmd}
-~\cite{rostedtTracecmdHostGuest}
-~\cite{KernelShark}
-~\cite{RealtimeKernelPatchset}
-~\cite{lutsykPipelinedMulticoreMachine2020}
-~\cite{rostedtInternalsRTPatch2007}
-~\cite{WhatRealtimeLinuxa}
-~\cite{LinuxProcessPriorities}
-~\cite{kernelrecipesKernelRecipes20162016}
-~\cite{thelinuxfoundationFindingSourcesLatency2020}
-~\cite{thelinuxfoundationChecklistWritingLinux2020}
-~\cite{HOWTOBuildRTapplication}
-~\cite{RealtimeProgrammingLinux}
-~\cite{KVMQemuVirtualization}
-~\cite{6DFRoboticArm}
-~\cite{MG996RServoMotor}
-~\cite{PackMg996rMetal}
-~\cite{twierengEmilyQuadrupedDog}
-~\cite{DigitalOutputSIGMATEK}
-~\cite{MG996RDigitalServo}
-~\cite{SDIASSIGMATEK}
-~\cite{ControlsHMIsSIGMATEK}
-~\cite{InterfacesSplittersSIGMATEK}
-~\cite{huang2015performance} 
-~\cite{kirovaImpactModernVirtualization2019}
-~\cite{adamPerformanceAssessmentLinux2021}
-\end{comment}
+Moreover, extensive testing of the virtualized system under various workloads can be conducted to evaluate its performance and reliability in different conditions and stressed situations, as demonstrated by \citeauthor{huang2015performance}~\cite{huang2015performance}, \citeauthor{kirovaImpactModernVirtualization2019}~\cite{kirovaImpactModernVirtualization2019}, or \citeauthor{adamPerformanceAssessmentLinux2021}~\cite{adamPerformanceAssessmentLinux2021}.
 
@@
 	\acro{CPU}[CPU]{Central Processing Unit}
-	\acro{GPOS}[GPOS]{General Purpose Operating System}
+	\acro{GPOS}[GPOS]{General-Purpose Operating System}
 	\acro{GUI}[GUI]{Graphical User Interface}

### 2024-08-31 Paper.tex
\printbibliography
Table fix
Fix table and figure fontsize
Doppelpunkt
to

@@
   {
-    \noindent\fontsize{9}{9pt}\textit{\textbf{Abstract—} #1}\\
-    \noindent\fontsize{9}{9pt}\textit{\textbf{Keywords—} #2}
+    \noindent\fontsize{9}{9pt}\textit{\textbf{Abstract:} #1}\\
+    \noindent\fontsize{9}{9pt}\textit{\textbf{Keywords:} #2}
   }
@@
 
+\usepackage{caption}
+\captionsetup[table]{font=footnotesize}
+\captionsetup[figure]{font=footnotesize}
+
 \begin{document}
@@
   \section{Methodology}
-  In order To address this gap, an extensive tuning process was carried out to achieve real-time performance and determinism.
+  In order to address this gap, an extensive tuning process was carried out to achieve real-time performance and determinism.
 These tunings were added sequentially and tested together.
@@
 Figure~\ref{fig:max_latency_combined_results} displays the change in latency after each real-time performance tuning.
-Table~\ref{tab:latency_tables_combined} gives an overview of the most important metrics of the measurements after each tuning.
 \clearpage
@@
   \end{figure}
+
+  Table~\ref{tab:latency_tables_combined} gives an overview of the most important metrics of the measurements after each tuning.
 
@@
     \footnotesize
-    \caption{Comparison of Latency Statistics in Salamander 4 under different Configurations}
+    \caption{Comparison of Robotic Application Latency of Salamander 4 Configurations}
     \label{tab:robotic_application_latency_values_combined}
@@
 
-  \section{Bibliography}
   \printbibliography

### 2024-08-31 Paper.tex
correct grammar paper

@@
 \begin{multicols}{2}
-  \abstract{\textbf{This thesis investigates the virtualization of a real-time operating system, which is built with Yocto, employs hard real-time through Xenomai 3 and is emulated using QEMU/KVM.
-The primary objective aim was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is crucial for real-time robotics.
+  \abstract{\textbf{This thesis investigates the virtualization of a real-time operating system, which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
+The primary objective was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is crucial for real-time robotics.
 Initial latency measurements showed a high gap between bare metal and virtualized setups.
-Extensive tuning across BIOS, kernel, host OS, QEMU/KVM, and guest OS reduced the worst-case latency from 707.622µs to 17.134µs, close to the bare metal performance of 10.709µs.
+Extensive tuning across BIOS, kernel, host OS, QEMU/KVM, and guest OS reduced the worst-case latency from 707.622~µs to 17.134~µs, close to the bare metal performance of 10.709~µs.
 The improvement was validated using a robotic application, comparing tuned virtualization with untuned and hardware versions.}} {\textbf{Virtualization, Real-Time Systems, Latency Reduction, Robot Control}}
@@
 Delays or errors can have catastrophic consequences in some cases.
-Virtualizing real-time operating systems for robotic control offers many advantages over traditional hardware-based approaches, especially when it comes to scaling~\cite{abbasiExploringOpenStackScalable2023}, flexibility and costs~\cite{taccariEmbeddedRealTimeVirtualization2014}.
+Virtualizing real-time operating systems (RTOS) for robotic control offers many advantages over traditional hardware-based approaches, especially when it comes to scaling~\cite{abbasiExploringOpenStackScalable2023}, flexibility, and costs~\cite{taccariEmbeddedRealTimeVirtualization2014}.
 However, virtualization also introduces overhead and latency~\cite{casiniLatencyAnalysisVirtualization2021}~\cite{zhangEvaluatingOptimizingVirtualization2010}.
@@
 These tunings were added sequentially and tested together.
-This involves configurations spanning the BIOS, kernel, host operating system (OS) and QEMU/KVM virtualization layer, as demonstrated in~\cite{RealTimePerformanceTuning2022}.
+This involves configurations spanning the BIOS, kernel, host operating system (OS), and QEMU/KVM virtualization layer, as demonstrated in~\cite{RealTimePerformanceTuning2022}.
 The guest OS is already based on Xenomai and therefore, no additional modifications were necessary.
@@
 Trace-cmd and Kernelshark were used for kernel tracing and visualization of trace data.
-The testing robot was connected to the virtual CPUS via a VARAN bus interface.
+The testing robot was connected to the virtual CPU via a VARAN bus interface.
 
@@
   The initial latency values were first measured with the \texttt{latency} program of the Xenomai tool suite.
-The tests were conducted with a sampling period of 100 µs, using a periodic user-mode task, and was assigned a priority of 99.
-If any sample's latency value was greater than 100µs, this was considered an overrun.
+The tests were conducted with a sampling period of 100~µs, using a periodic user-mode task, and were assigned a priority of 99.
+If any sample's latency value was greater than 100~µs, this was considered an overrun.
 There was a significant initial gap in latency statistics between the virtualized Salamander 4 and the Salamander 4 on bare metal.
@@
 On top of that, a robotic application further confirmed the improvements.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned and hardware version of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and hardware versions of Salamander 4.
 
@@
 The improved latency was also tested with a robotic application.
-The difference between the command issuance time and the time the signal reaches the PWM was measured 1,000 times for the bare metal, untuned and tuned version of Salamander 4.
+The difference between the command issuance time and the time the signal reaches the PWM was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
 The results are visualized in Table~\ref{tab:robotic_application_latency_values_combined}.
@@
   \section{Discussion}
-  The results shows that the real-time performance tunings applied to the virtualized Salamander 4 OS reduced latency and improved determinism.
-The maximum latency decreased from 707.622µs to 17.134µs after tuning BIOS, kernel, host, guest, and QEMU/KVM configurations.
-Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694µs.
+  The results show that the real-time performance tunings applied to the virtualized Salamander 4 OS reduced latency and improved determinism.
+The maximum latency decreased from 707.622~µs to 17.134~µs after tuning BIOS, kernel, host, guest, and QEMU/KVM configurations.
+Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694~µs.
 This also eliminated overruns going forward.
-Host configurations, including
-  the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU further reduced latency a little bit down to 18.441µs.
+Host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU further reduced latency a little bit down to 18.441~µs.
 The guest OS already had real-time capabilities through Xenomai, hence the focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
-Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages brought the latency down to the final worst latency value of 17.134µs.
-This value is very close to the bare metal value of 10.709\textmu s.
-The goal was set to achieving latency values below 50 microseconds in the duration of the measuremenet.
-This goal was acheived.
+Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages, brought the latency down to the final worst latency value of 17.134~µs.
+This value is very close to the bare metal value of 10.709~µs.
+The goal was set to achieving latency values below 50 microseconds in the duration of the measurement.
+This goal was achieved.
   A robotic application in a practical scenario validated the improvement of the tuned virtualization compared to the unmodified version.
@@
 Extensive tuning was performed sequentially: BIOS, kernel, host, and QEMU/KVM configurations.
-This process reduced worst latency from 707.62µs to 17.134µs, close to the bare metal value of 10.709µs, achieving the goal of sub-50µs latency.
+This process reduced worst latency from 707.62~µs to 17.134~µs, close to the bare metal value of 10.709~µs, achieving the goal of sub-50~µs latency.
   A robotic application confirmed these improvements.

### 2024-08-31 Paper.tex
correct grammar and format

@@
   In today's industrial production and automation, robots must react to their environment and perform time-critical tasks within strict time constraints.
-Delays or errors can have catastrophic consequences in some cases.
-Virtualizing real-time operating systems (RTOS) for robotic control offers many advantages over traditional hardware-based approaches, especially when it comes to scaling~\cite{abbasiExploringOpenStackScalable2023}, flexibility, and costs~\cite{taccariEmbeddedRealTimeVirtualization2014}.
+Delays or errors can have catastrophic consequences.
+In a traditional general-purpose operating system (GPOS), a high-priority task cannot interrupt a kernel operation, which makes them unsuitable for real-time requirements, as they cannot guarantee deterministic execution times.
+However, in a real-time operating system (RTOS), a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation.
+Virtualizing real-time operating systems for robotic control offers many advantages over traditional hardware-based approaches, especially when it comes to scaling~\cite{abbasiExploringOpenStackScalable2023}, flexibility, and costs~\cite{taccariEmbeddedRealTimeVirtualization2014}.
 However, virtualization also introduces overhead and latency~\cite{casiniLatencyAnalysisVirtualization2021}~\cite{zhangEvaluatingOptimizingVirtualization2010}.
@@
   \section{Methodology}
-  In order to address this gap, an extensive tuning process was carried out to achieve real-time performance and determinism.
+  In order to address this gap and achieve determinism, an extensive real-time performance tuning process was carried out.
 These tunings were added sequentially and tested together.
@@
 If any sample's latency value was greater than 100~µs, this was considered an overrun.
-There was a significant initial gap in latency statistics between the virtualized Salamander 4 and the Salamander 4 on bare metal.
+There was a significant initial gap in latency statistics between the bare metal and virtualized Salamander 4.
 To address this gap, an extensive tuning process was carried out to achieve real-time performance and determinism.
@@
 Figure~\ref{fig:max_latency_combined_results} displays the change in latency after each real-time performance tuning.
-\clearpage
+
   \begin{figure}[H]
@@
 Compared to the hardware version's worst latency of 1.49ms, the tuned virtualization came very close to the determinism and reliability of the hardware.
+
   \section{Summary and Outlook}

### 2024-09-01 Masterarbeit_Robotik_Pamuk.tex
correct grammar and format
minor changes

@@
 Daher wird ein umfangreicher Tuning-Prozess durchgeführt, um Echtzeit-Performance und Determinismus zu erreichen.
-Diese Modifikationen umfassen Konfigurationen, die das BIOS, den Kernel, das Host-Betriebssystem, die QEMU/KVM-Virtualisierungsschicht und das Salamander 4-Betriebssystem (Gast) selbst betreffen.
+Diese Modifikationen umfassen Konfigurationen, die das BIOS, den Kernel, das Host-Betriebssystem, die QEMU/KVM-Virtualisierungsschicht und das Salamander 4 Gast-Betriebssystem selbst betreffen.
 Die Worst-Case-Latenz wurde von 707.622~µs auf 17.134~µs reduziert, was der Bare-Metal-Performance von 10.709~µs nahekommt.
@@
 
-\outline{Virtualization of real-time operating systems systems for robotic control has many advantages in comparison to traditional hardware-based approaches.
+\outline{Virtualization of real-time operating systems (RTOS) for robotic control has many advantages in comparison to traditional hardware-based approaches.
 Solutions based on pure hardware are often built for a distinct purpose and are hence limited in flexibility.
@@
 Thus, an extensive tuning process is carried out to achieve real-time performance and determinism.
-These modifications involve configurations spanning the BIOS, kernel, host operating system, QEMU/KVM virtualization layer, and the Salamander 4 operating system (guest) itself.
+These modifications involve configurations spanning the BIOS, kernel, host operating system, QEMU/KVM virtualization layer, and the Salamander 4 guest operating system itself.
 The worst-case latency was brought down from 707.622~µs to 17.134~µs, closely aligning with the bare metal performance of 10.709~µs.
@@
 Delays or errors can have catastrophic consequences in some cases.
-Modern operating systems nowadays are categorized into two types:  general-purpose operating systems (GPOS) and real-time operating systems (RTOS).
+Modern operating systems nowadays are categorized into two types:  General-Purpose Operating Systems (GPOS) and Real-Time Operating Systems (RTOS).
 The key difference is whether the system is time-critical or not~\cite{canbazPerformanceAnalysisRealtime2022}.
@@
 	\item Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open-source project that helps develop custom Linux-based systems. Yocto is discussed in Section~\ref{sec:yocto}.
-	\item Salamander 4 is virtualized through Quick emulator (QEMU), an open-source emulator and virtualizer that allows running different operating systems on a computer.
+	\item Salamander 4 is virtualized through Quick Emulator (QEMU), an open-source emulator and virtualizer that allows running different operating systems on a computer.
 QEMU is described in Section~\ref{sec:qemu}.
@@
 When a physical interrupt occurs while the guest RTOS is running, there are at least six stages that need to be traversed before the interrupt can be delivered to the guest RTOS.
-Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), which is the time between when an interrupt is recognized and the first instruction of the corresponding interrupt service routine (ISR) of the guest OS is started.
+Once the interrupt is delivered to the guest RTOS, it incurs additional latency, called Interrupt Routine Time (IRT), which is the time between when an interrupt is recognized and the first instruction of the corresponding Interrupt Service Routine (ISR) of the guest OS is started.
 These stages are the primary sources of latencies caused by KVM, they are briefly described in Figure~\ref{fig:kvm_stages}.
@@
 For this reason, an extensive tuning process is carried out to achieve real-time performance and determinism.
-These modifications involve configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS (guest) itself.
+These modifications involve configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 guest OS itself.
 The individual configurations are discussed in detail in Chapter~\ref{sec:real-time_tunings} and the modifications are justified with a clear explanation.
@@
 The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU.
-The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU5).
+The Least Significant Bit (LSB) on the right corresponds to the first CPU (CPU0), and the Most Significant Bit (MSB) on the left corresponds to the last CPU (CPU5).
 When there are 6 CPUs available, the default value for \texttt{smp\_affinity} would be 3F (11 1111).

### 2024-09-01 Paper.tex
RTOS

@@
 \begin{multicols}{2}
-  \abstract{\textbf{This thesis investigates the virtualization of a real-time operating system, which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
+  \abstract{\textbf{This thesis investigates the virtualization of a Real-Time Operating System (RTOS), which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
 The primary objective was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is crucial for real-time robotics.

### 2024-09-01 Masterarbeit_Robotik_Pamuk.tex
RTOS abstract

@@
 
-\outline{Virtualization of real-time operating systems (RTOS) for robotic control has many advantages in comparison to traditional hardware-based approaches.
+\outline{Virtualization of Real-Time Operating Systems (RTOS) for robotic control has many advantages in comparison to traditional hardware-based approaches.
 Solutions based on pure hardware are often built for a distinct purpose and are hence limited in flexibility.

### 2024-09-01 Masterarbeit_Robotik_Pamuk.tex
Final Correction

@@
     \renewcommand\lstlistingname{Code}
-    \renewcommand\lstlistlistingname{List of Code}
+    \renewcommand\lstlistlistingname{List of Codes}
     \renewcommand\listacroname{List of Abbreviations}
@@
 \bigskip \noindent Insgesamt bietet diese Arbeit eine umfassende Methodik, um ein virtualisiertes Gastsystem in einem Hostsystem mit deterministischem Verhalten echtzeitfähig zu machen.}
-\schlagworte{Virtualisierung, Echtzeitsysteme, Latenzreduktion, Robotiksteuerung}
+\schlagworte{Virtualisierung, Echtzeitsysteme, Latenzreduktion, Robotersteuerung}
 
@@
 Delays or errors can have catastrophic consequences in some cases.
-Modern operating systems nowadays are categorized into two types:  General-Purpose Operating Systems (GPOS) and Real-Time Operating Systems (RTOS).
+Modern operating systems nowadays are categorized into two types:  General-Purpose Operating System (GPOS) and Real-Time Operating System (RTOS).
 The key difference is whether the system is time-critical or not~\cite{canbazPerformanceAnalysisRealtime2022}.
@@
 The basic rule on which priority-based preemptive scheduling is based is that the task with the highest priority that is ready to run is always the task that must be executed.
-So if both a task with a lower priority and a task with a higher priority are ready to run, the scheduler ensures that the task with the higher priority runs first.
-The lower priority task is only executed once the higher priority task has been processed.
+So, if both a task with a lower priority and a task with a higher priority are ready to run, the scheduler ensures that the task with the higher priority runs first.
+The lower-priority task is only executed once the higher priority task has been processed.
 
@@
 This operating system will be referred to as ``Salamander 4'' in this work.
-The details of Salamander 4 are explained in Chapter~\ref{sec:guest_operating_system}.
-	\item Salamander 4 was created with Yocto~\cite{WelcomeYoctoProject}, an open-source project that helps develop custom Linux-based systems. Yocto is discussed in Section~\ref{sec:yocto}.
+The details of Salamander 4 are explained in Section~\ref{sec:guest_operating_system}.
 	\item Salamander 4 is virtualized through Quick Emulator (QEMU), an open-source emulator and virtualizer that allows running different operating systems on a computer.
 QEMU is described in Section~\ref{sec:qemu}.
-	\item Salamander 4 employs hard real-time with Xenomai 3~\cite{XenomaiXenomai}, a real-time framework for Linux that enables deterministic and low-latency performance.
-Xenomai 3 is detailed in Section~\ref{sec:xenomai}.
+	\item Salamander 4 was created with Yocto, an open-source project that helps develop custom Linux-based systems. Yocto is discussed in Section~\ref{sec:yocto}.
+	\item Salamander 4 employs hard real-time with Xenomai 3, a real-time framework for Linux that enables deterministic and low-latency performance. Xenomai 3 is detailed in Section~\ref{sec:xenomai}.
 	\item The goal is to virtualize Salamander 4 and approach the performance of bare metal (hardware) through real-time performance tunings, focusing on latency requirements set by SIGMATEK.
-These modifications are the subject of Chapter~\ref{sec:real-time_tunings}.
+These modifications are the subject of Section~\ref{sec:real-time_tunings}.
 	\item The \texttt{latency} tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to be used by SIGMATEK.
@@
 	\item For the purpose of replacing the physical CPU, SIGMATEK provided a PCI Insert Card module PCV 522, that was plugged into the PC, as explained in Subsection~\ref{subsec:setup_experiment_virtualized}.
-This component could then be used to communicate with the Input/Output peripherals, just like the physical CPU could when the I/O modules were attached to it.
-Subsequently, the PCV 522 module of SIGMATEK was connected with the Pulse Width Module PW 022 of SIGMATEK over the VARAN bus and the VI 021 module of SIGMATEK serving as the power supply.
-	\item The robotic application from Chapter~\ref{sec:robotic_application} was written in Lasal Class 2, an object-oriented programming tool of SIGMATEK and part of SIGMATEK's Engineering Platform LASAL, with client-server technology and graphic representation.
+This component could then be used to communicate with the Input/Output (I/O) peripherals, just like the physical CPU could when the I/O modules were attached to it.
+Subsequently, the PCV 522 module of SIGMATEK was connected with the Pulse Width Module PW 022 of SIGMATEK over the VARAN bus and the Varan Connection Module VI 021 of SIGMATEK.
+	\item The robotic application from Section~\ref{sec:robotic_application} was written in Lasal Class 2, an object-oriented programming tool of SIGMATEK and part of SIGMATEK's Engineering Platform LASAL, with client-server technology and graphic representation.
 \end{itemize}
@@
 They talk about the evolution of real-time behavior in the standard Linux kernel and state that it has recently gained popularity among the real-time community, largely due to its open-source nature and stability.
-The authors explain how to transform Linux into a real-time operating system by implementing the PREEMPT-RT patches.
+The authors explain how to transform Linux into a real-time operating system by implementing the PREEMPT-RT patch.
 The paper also emphasizes some kernel configurations that ensure that the system's real-time behavior is preserved during runtime.
@@
 Given that the working computer used for this work is equipped with an Intel processor, the guidelines presented in this document were particularly relevant and valuable for optimizing real-time performance in KVM-based virtual machines.
-Chapter~\ref{sec:real-time_tunings} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
+Section~\ref{sec:real-time_tunings} contains an extensive tuning process, which involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
 The document concludes that the described tuning methods effectively improve real-time performance for KVM-based VMs, even under heavy workloads.
@@
 \bigskip \noindent \citeauthor{mckenneyRealTimeVs2008}~\cite{mckenneyRealTimeVs2008} addresses the overheads associated with real-time Linux when using KVM and QEMU.
-The author lists several sources of overhead, including memory locking to avoid page-fault latencies, increased overhead of locking and interrupts due to more-aggressive preemption, threaded interrupts that permit long-running interrupt handlers to be preempted by high-priority real-time processes, real-time task scheduling that requires global scheduling, high-resolution timers for tens-of-microseconds accuracy and precision, and preemptible RCU which has slightly higher read-side overhead than Classic RCU.
+The author lists several sources of overhead, including memory locking to avoid page-fault latencies, increased overhead of locking and interrupts due to more-aggressive preemption, threaded interrupts that permit long-running interrupt handlers to be preempted by high-priority real-time processes, real-time task scheduling that requires global scheduling, high-resolution timers for tens-of-microseconds accuracy and precision, and preemptible RCU which has slightly higher read-side overhead than classic RCU.
 This information helped understand the trade-offs involved in real-time Linux systems.
@@
 This contributed to the decision not to use this tool to measure the scheduling latency in this work.
-Instead, the \texttt{latency} tool was used, as was determined in Chapter~\ref{sec:application_context}.
+Instead, the \texttt{latency} tool was used, as was determined in Section~\ref{sec:application_context}.
 
@@
 
-\bigskip \noindent Virtualization promises to overcome these challenges~\cite{queirozContainerbasedVirtualizationRealtime2024}~\cite{bhardwajVirtualizationCloudComputing2021}~\cite{cinqueEvaluatingVirtualizationFog2023}.
+\bigskip \noindent Virtualization promises to overcome these challenges~\cite{bhardwajVirtualizationCloudComputing2021}~\cite{queirozContainerbasedVirtualizationRealtime2024}~\cite{cinqueEvaluatingVirtualizationFog2023}.
 However, virtualization also has its own challenges and pitfalls.
@@
 The hypervisor and other virtual components can cause variations in how long tasks take to execute~\cite{garcia-vallsChallengesRealtimeVirtualization2014}.
-When there are multiple virtual machines (VM) sharing the same physical resources, the VMs can affect each other's performance and consequently violate real-time constraints~\cite{queirozContainerbasedVirtualizationRealtime2024}.
+When there are multiple virtual machines (VMs) sharing the same physical resources, the VMs can affect each other's performance and consequently violate real-time constraints~\cite{queirozContainerbasedVirtualizationRealtime2024}.
 
@@
 As already established in Section~\ref{sec:application_context}, Salamander 4 is built with Yocto, employs hard real-time with Xenomai 3, and is virtualized through QEMU/KVM.
-The initial latency values are measured with the \texttt{latency} program of the Xenomai tool suite.
+The latency values are measured with the \texttt{latency} program of the Xenomai tool suite.
 The initial situation is described in Section~\ref{sec:starting_situation}.
-In a nutshell, there is a significant initial gap in latency statistics between the virtualized Salamander 4 and the Salamander 4 on bare metal.
+In a nutshell, there is a significant initial gap in latency between the virtualized Salamander 4 and the Salamander 4 on bare metal.
 For this reason, an extensive tuning process is carried out to achieve real-time performance and determinism.
 These modifications involve configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 guest OS itself.
-The individual configurations are discussed in detail in Chapter~\ref{sec:real-time_tunings} and the modifications are justified with a clear explanation.
+The individual configurations are discussed in detail in Section~\ref{sec:real-time_tunings} and the modifications are justified with a clear explanation.
 
 \section{Objective}
-The goal is to bring the latency of the virtualized Salamander 4 closer to that of the hardware, thereby providing a comprehensive blueprint for making a virtualized guest system in a host system real-time capable with deterministic behavior.
+The goal is to bring the latency of the virtualized Salamander 4 closer to that of the hardware (bare metal), thereby providing a comprehensive blueprint for making a virtualized guest system in a host system real-time capable with deterministic behavior.
 Section~\ref{sec:starting_situation} outlines the initial situation of this master's thesis and serves as the foundation for the real-time performance tunings discussed in Section~\ref{sec:real-time_tunings}.
@@
 
-This chapter describes in detail all the theoretical concepts and boundary conditions as well as practical methods that contributed to achieving the objectives of this master's thesis.
+This chapter describes in detail all the theoretical concepts and practical methods that contributed to achieving the objectives of this master's thesis.
 
@@
 \noindent Salamander 4 is the proprietary operating system of SIGMATEK.
-It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment~\cite{XenomaiXenomai}.
+It is based on Linux version 5.15.94 and integrates Xenomai 3.2, a real-time development environment.
 Salamander 4 is a 64-bit system, which refers to the x86\_64 architecture.
@@
 In addition, it uses IRQPIPE to process interrupts in a way that meets the real-time requirements of the system.
-The output of the command~\texttt{uname -a} can be observed in code~\ref{output:uname_a_virt}.
+The output of the command~\texttt{uname -a} can be observed in Code~\ref{output:uname_a_virt}.
 
 \vspace{2em}
-\begin{lstlisting}[name={System Information},label={output:uname_a_virt}]
+\begin{lstlisting}[name={Guest Operating System Information},label={output:uname_a_virt}]
 root@sigmatek-core2:~# uname -a
@@
 They help organize and manage the hardware components in a modular manner.
-The graphical hardware editor in LASAL allows for a true-to-detail simulation of the actual hardware.
+The graphical hardware editor in LASAL allows for a accurate simulation of the actual hardware.
 	\item \textbf{Applications}: Applications are developed using LASAL CLASS 2~\cite{EngineeringToolLASAL}, a solution for automation tasks that supports object-oriented programming and design in compliance with IEC 61131-3.
@@
 
-QEMU~\cite{QEMU} is an open-source emulator and virtualizer that allows running operating systems and programs for one machine on a different machine.
+Quick Emulator (QEMU)~\cite{QEMU} is an open-source emulator and virtualizer that allows running operating systems and programs for one machine on a different machine.
 It can emulate different hardware architectures and is used in this thesis to emulate Salamander 4.
@@
 \begin{lstlisting}[name={Contents of QEMU Folder for Salamander 4},label={lst:ls}]
-    sigma_ibo@localhost:~/Desktop/salamander-image$ ls -1
+    sigma_ibo@localhost:~/build/tmp/deploy/qemu/salamander-image$ ls -1
     bzImage
@@
 	\item \textbf{bzImage}: Compressed Linux kernel image that is loaded by QEMU at system start. The short form ``bz`` stands for big-zipped.
-	\item \textbf{drive-c}: Directory serving as C drive for QEMU, created and filled by qemu\_def.sh script.
+	\item \textbf{drive-c}: Directory serving as C drive for QEMU, created and filled by \texttt{qemu\_def.sh} script.
 	\item \textbf{ovmf.code.qcow2}: Firmware file for QEMU that enables UEFI boot process.
 ~\texttt{OVMF} stands for Open Virtual Machine Firmware,~\texttt{qcow2} is a format for disk image files used by QEMU, it stands for ``QEMU Copy On Write version 2''.
-	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot Salamamder 4 OS. It is described later in more detail.
-	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS for the Sigmatek Core 2 platform.
+	\item \textbf{qemu\_def.sh}: Shell script that starts QEMU with required parameters to boot the Salamamder 4 OS. It is described later in more detail.
+	\item \textbf{salamander-image-sigmatek-core2.ext4}: Disk image of the Salamander 4 OS.
 It uses the~\texttt{ext4} file system and serves as the root file system in the QEMU virtual machine, acting as the virtual hard drive.
@@
 \noindent The initial QEMU script after the custom Yocto build and the starting point for this work is shown in Code~\ref{script:qemu_def}.
-This script is used to start QEMU with required parameters to boot Salamander 4 OS.
-It will be adjusted in Chapter~\ref{sec:real-time_tunings} in order to accompany real-time performance tunings.
+This script is used to start QEMU with required parameters to boot the Salamander 4 OS.
+It will be adjusted in Section~\ref{sec:real-time_tunings} in order to accompany real-time performance tunings.
 
 \vspace{2em}
-\begin{lstlisting}[name={QEMU script for starting Salamander 4 Virtualization},label={script:qemu_def}]
+\begin{lstlisting}[name={QEMU Script for starting Salamander 4 Virtualization},label={script:qemu_def}]
 	#!/bin/sh
@@
 \noindent  The second approach, called Mercury and shown in Figure~\ref{fig:mercury}, relies on the real-time capabilities already present in the native Linux kernel.
-Often, applications require the PREEMPT-RT~\cite{RealtimePreempt_rt_versionsWiki} patch to augment the mainline kernel's real-time responsiveness and minimize jitter, but this is not mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
+Often, applications require the PREEMPT-RT patch~\cite{RealtimePreempt_rt_versionsWiki} to augment the mainline kernel's real-time responsiveness and minimize jitter, but this is not mandatory and depends on the application's specific requirements for responsiveness and tolerance for occasional deadline misses.
 
@@
 \section{Trace-cmd}\label{sec:trace-cmd}
-\texttt{Trace-cmd}~\cite{Tracecmd} is the front-end tool of the \texttt{ftrace} tool that is built inside the Linux kernel and used for tracing the Linux kernel.
+\texttt{Trace-cmd}~\cite{Tracecmd} is the front-end tool of the \texttt{ftrace}~\cite{FtraceFunctionTracer} tool that is built inside the Linux kernel and used for tracing the Linux kernel.
 \texttt{Trace-cmd} can record detailed reports of kernel events such as interrupts, scheduler decisions, function calls, and custom events in real time.
@@
 A very practical aspect of \texttt{trace-cmd} is that the guest can be traced from the host.
-For this purpose, the guest needs to be configured up with \texttt{vsocks}, which are a virtual socket that enable direct communication between the host and guest.
+For this purpose, the guest needs to be configured with \texttt{vsocks}, which are a virtual socket that enable direct communication between the host and guest.
 To enable vsockets and tracing, the guest kernel requires the following configurations, according to Linux developer and maintainer Steve Rostedt~\cite{rostedtTracecmdHostGuest}.
@@
 In QEMU, the line \texttt{-device vhost-vsock-pci,guest-cid=3,id=vsock0} was added.
-This command configures a virtual socket (vsock) device in QEMU, assigning the guest a unique Context Identifier (CID) of 3 for direct communication with the host.
+This command configures a virtual socket (\texttt{vsock}) device in QEMU, assigning the guest a unique Context Identifier (\texttt{cid}) of 3 for direct communication with the host.
 The CID can be chosen as desired.
-Once the guest is booted with vhost enabled, the \texttt{sudo trace-cmd agent} command can be run.
+Once the guest is booted with \texttt{vhost} enabled, the \texttt{sudo trace-cmd agent} command can be run.
 The agent now listens to connections on port 823 by default with the specified CID, 3 in this case.
@@
 \section{Approach}\label{sec:approach}
-The goal is to virtualize the Salamander 4 OS and bring its real-time performance closer to that of the hardware version and guarantee deterministic and reliable behavior.
-For that purpose, the first step is to apply the PREEMPT-RT patch to the host system and subsequently tune the real-time performance through BIOS, kernel, host, guest, and QEMU/KVM configurations.
+The goal is to virtualize the Salamander 4 OS and bring its real-time performance closer to that of the bare metal version and guarantee deterministic and reliable behavior.
+For that purpose, the first step is to apply the PREEMPT-RT patch to the host system and subsequently tune the real-time performance through BIOS, kernel, host, QEMU/KVM, and guest configurations.
 These modifications are added sequentially and tested together.
@@
 At the end, the results are demonstrated altogether and a holistic comparison and discussion are carried out.
-In addition, the improvement in real-time performance and determinism is demonstrated using a robotic application, where the modified virtualization is compared with the hardware version.
+In addition, the improvement in real-time performance and determinism is demonstrated using a robotic application, where the tuned virtualization is compared with the untuned and the bare metal version.
 
@@
 Subsections~\ref{subsec:salamander4-bare-metal} and~\ref{subsec:salamander4-virtualization} specify the details of the measurements for both setups.
-In the further course of this master's thesis, the aim is to bring the latency values of the virtualization closer to those of the hardware and guarantee deterministic and reliable behavior.
+In the further course of this master's thesis, the aim is to bring the latency values of the virtualization closer to those of bare metal and guarantee deterministic and reliable behavior.
 
@@
 	\centering
-	\caption{Latency Parameters for Bare Metal}
+	\caption{Latency Parameters of Bare Metal}
 	\label{tab:latency_stats_hardware}
@@
 In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualized version of this operating system.
-Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, Figure~\ref{fig:gnuplot_max_latency_default} depicts the distribution of latency.
+Figure~\ref{fig:gnuplot_max_latency_default} shows the distribution of latency measured over 10 minutes using the default QEMU script in Code~\ref{script:qemu_def} without any further adjustments.
 
@@
 
+\clearpage
+
 \noindent The statistics obtained from this measurement are provided in Table~\ref{tab:latency_stats_virt} and Table~\ref{tab:latency_overrun_msw_new_virt}.
@@
 	\centering
-	\caption{Latency Parameters for Untuned Virtualization}
+	\caption{Latency Parameters of Untuned Virtualization}
 	\label{tab:latency_stats_virt}
@@
 	\centering
-	\caption[Latency Statistics with Overrun Counts for Untuned Virtualization]{Minimum, Average, and Maximum Latency with Overrun Counts for Untuned Virtualization}
+	\caption[Latency Statistics with Overrun Counts of Untuned Virtualization]{Minimum, Average, and Maximum Latency with Overrun Counts of Untuned Virtualization}
 	\label{tab:latency_overrun_msw_new_virt}
@@
 \subsection{Initial Comparison}
-\noindent Comparing the latency values of the virtualization to those of bare metal in Figure~\ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in the statistics.
+\noindent Comparing the latency values of the virtualization to those of bare metal in Figure~\ref{fig:gnuplot_max_latency_combined}, it is evident that there is a significant initial gap in latency.
 A maximum latency of 707.62~µs is not tolerable for a real-time virtualization and needs to be tuned.
@@
 	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_combined_max_latency.png}
-	\caption[Initial Comparison of Latency Distribution between Hardware and Virtualization]{Initial Comparison of Latency Distribution between Hardware and Virtualization}
+	\caption[Initial Comparison of Latency Distribution between Bare Metal and Virtualization]{Initial Comparison of Latency Distribution between Bare Metal and Virtualization}
 	\label{fig:gnuplot_max_latency_combined}
@@
 
-In this chapter, the significant initial gap in latency statistics between the virtualized system and the bare metal system is tackled.
+In this section, the initial large gap in latency between the bare metal and the virtualized system is tackled.
 For this reason, an extensive tuning process is carried out.
-This involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself.
+This involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 guest OS itself.
 The individual configurations are discussed in detail, and the modifications are justified with a clear explanation.
@@
 Additional modifications are required to achieve this.
-The next sections will deal with these real-time performance tunings.
+The next subsections will deal with these real-time performance tunings.
 After each tuning, the \texttt{latency} test of Xenomai is executed to evaluate the improved latency.
 The most important metric of a latency test is, without a doubt, the worst latency value, because the system is only as reliable as its worst measurement.
-Since the period of the tests is set to 100~µs, there are approximately 1,000,000 samples per second.
+Since the period of the tests is set to 100 µs, there are approximately 10,000 samples per second.
+Therefore, 10 minutes or 600 seconds correspond to approximately 6,000,000 samples.
 
@@
 Table~\ref{tab:bios_configuration} illustrates the specific BIOS settings that have been adjusted for the purpose of real-time performance.
-As an important note, this is not a definitive, complete list, and other devices may have additional settings that can be modified to achieve even better latency.
+As an important note, this is not a definitive or complete list.
+Other devices may have additional settings that can be modified to achieve even better latency.
 
@@
 			\hline
-			C States                      & Disabled        \\
+			C-States                      & Disabled        \\
 			\hline
@@
 \begin{itemize}
-	\item \textbf{Hyper Threading}: When hyper-threading is enabled in the BIOS settings, this allows CPUs to work on two threads simultaneously instead of just one.
-This allows the parallelization of tasks and seemingly increases performance.
+	\item \textbf{Hyper Threading}: When hyper-threading is enabled in the BIOS settings, CPUs can work on two threads simultaneously instead of just one.
+This allows the parallelization of tasks and increases performance.
 However, in a real-time system like the guest Salamander 4, this can lead to increased latencies due to contention between threads.
@@
 
-\noindent After these first real-time tunings in the BIOS settings, the \texttt{latency} test was run to see whether this had any significant impact on the determinism of the system.
-Figure~\ref{fig:gnuplot_max_latency_taskset} shows the latency values after tuning BIOS settings.
+\noindent After applying the PREEMPT-RT patch and first real-time tunings in the BIOS settings, the \texttt{latency} test was run to see whether this had any impact on the determinism of the system.
+Figure~\ref{fig:gnuplot_max_latency_taskset} shows the latency values after tuning BIOS configurations.
 
@@
 		\textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
-		0.969                 & 3.948                 & 457.545               & 94                \\ \hline
+		0.969                 & 3.948                 & 457.545               & 22                \\ \hline
 	\end{tabular}
@@
 When compared to the initial worst latency of 707.62~µs, this is a step for the better, but certainly still not acceptable within this work's definition of hard real-time determinism.
+Besides, the number of overruns decreased from 43 to 22.
 
@@
 \begin{lstlisting}[name={Kernel Configurations for Real-Time Performance},label={script:kernel_cli}]
-		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 rcu_nocb_poll nohz_full=4 nohz=on default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
+		GRUB_CMDLINE_LINUX="isolcpus=4 rcu_nocbs=4 rcu_nocb_poll nohz_full=4 default_hugepagesz=1G hugepagesz=1G hugepages=8 intel_iommu=on rdt=l3cat nmi_watchdog=0 idle=poll clocksource=tsc tsc=reliable audit=0 skew_tick=1 intel_pstate=disable intel.max_cstate=0 intel_idle.max_cstate=0 processor.max_cstate=0 processor_idle.max_cstate=0 nosoftlockup no_timer_check nospectre_v2 spectre_v2_user=off kvm.kvmclock_periodic_sync=N kvm_intel.ple_gap=0 irqaffinity=0"
 \end{lstlisting}
@@
 This reduces latency.
-	\item \textbf{nohz\_full=4}: Makes CPU core 4 ``tickless'', meaning the kernel tries to avoid sending periodic scheduling-clock interrupts to the CPU when there are no runnable tasks.
+	\item \textbf{nohz\_full=4}: Makes CPU core 4 "tickless", meaning the kernel tries to avoid sending periodic scheduling-clock interrupts to the CPU when there are no runnable tasks.
 This lowers latency by reducing unnecessary wake-ups but may increase power consumption because the CPU is not able to enter a low-power state when idle.
 Additionally, timer interrupts cannot be fully eliminated because certain events, such as incoming interrupts or task activations, can still cause the kernel to send timer interrupts to the tickless CPU.
-	\item \textbf{nohz=on}: Sets all CPUs to tickless mode system-wide.
 	\item \textbf{default\_hugepagesz=1G, hugepagesz=1G, hugepages=8}: Huge pages are large contiguous areas of memory that can be used by applications and the kernel, instead of the traditional 4KB small pages.
 The default huge page size is set to 1GB, and 8 huge pages of 1GB size are reserved at boot.
-This pre-allocation makes sure that these large memory regions are available to be used by the kernel or applications, without having to dynamically allocate and potentially fail.
-	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit), which connects a DMA (Direct Memory Access)-capable I/O bus, such as graphics cards and network adapters, to the main memory.
-It allows these devices to directly access and use memory, which is especially helpful when they are virtualized.
+This pre-allocation makes sure that these large memory regions are available to be used by the kernel or applications without having to dynamically allocate and potentially fail.
+	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit), which connects a DMA (Direct Memory Access)-capable I/O bus to the main memory.
+It allows these devices to directly access and use memory, which is especially helpful for virtualization and device passthrough.
 	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel's RDT (Resource Director Technology).
@@
 L3 CAT controls the amount of L3 cache that a process can use.
-By controlling cache allocation, it can prevent a single process from monopolizing the L3 cache, which is particularly beneficial in virtualized environments, where multiple virtual machines share the same physical host.
+By controlling cache allocation, it can prevent a single process from monopolizing the L3 cache, which is particularly beneficial in virtualized environments where multiple virtual machines share the same physical host.
 	\item \textbf{nmi\_watchdog=0}: Disables the NMI (Non-Maskable Interrupt) watchdog, which is a debugging feature of the Linux kernel.
@@
 By setting the gap to 0, the virtual machine exits the pause loop immediately, which can reduce latency in spinlock-intensive workloads.
-	\item \textbf{irqaffinity=0}: Sets the default interrupt request affinity to none.
+	\item \textbf{irqaffinity=0}: Sets the default IRQ affinity mask so that interrupts are handled by non-isolated CPUs.
 This means that no CPU core is preferred over another for handling IRQs.
@@
 
-\noindent Apparent is a significant improvement in the latency of the virtualized Salamander 4 OS. With a worst latency value of 21.694~µs, the real-time tunings appear to be working as desired.
+\noindent There is a significant improvement in the latency of the virtualized Salamander 4 OS.
+With a worst latency value of 21.694~µs, the real-time tunings appear to be working as desired.
+Additionally, no overruns occurred during the latency test.
+This means that the system consistently stayed within the required time constraints
 
@@
 This entails a number of host-side adjustments to reduce interruptions and latency.
-In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest Salamander 4 OS is provided.
+In the following, a detailed overview of these configurations and their impact on the real-time performance of the Salamander 4 guest OS is provided.
 
@@
 Isolating CPUs means removing all user-space threads and unbound kernel threads, since bound kernel threads are tied to specific CPUs and hence cannot be moved.
-For CPU isolation, the \texttt{isolcpus} function is used to isolate a performance CPU from the general scheduling algorithms of the operating system.
+For CPU isolation, the \texttt{isolcpus} function is used to isolate a CPU from the general scheduling algorithms of the operating system.
 This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks.
 However, the \texttt{isolcpus} function only isolates at the user level and does not affect kernel tasks.
-Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased2013}, including systemd services.
+Consequently, these kernel tasks and interrupts can still use the CPU~\cite{maPerformanceTuningKVMbased2013}, including systemd services.
 To prevent systemd services from running on an isolated CPU, the CPU affinity can be set with the line \texttt{CPUAffinity=0 1 2 3 5} in the \texttt{/etc/systemd/system.conf} file to indicate the CPUs that systemd services are allowed to run on.
@@
 Code~\ref{output:pse} shows the user and kernel tasks that run on CPU 4.
-An isolated CPU can be bound to a task by using \texttt{taskset -c <CPU>}, as shown later in Subsection~\ref{subsec:guest_configurations}.
+An isolated CPU can be bound to a task by using \texttt{taskset -c <CPU>}, as shown later in Subsection~\ref{subsec:qemu_configurations}.
 After the isolation, user tasks other than the QEMU process have been removed from running on this CPU.
@@
 \subsubsection{Interrupt Affinity}\label{subsubsec:irq_handling}
-Once the CPUs were isolated, interrupt request handling was the next step.
+Once the CPUs are isolated, interrupt request handling is the next step.
 The purpose of interrupt requests is to inform the CPU to stop working on a certain job and start working on another.
@@
 All the existing interrupts can be monitored using the command \texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time.
-The IRQs needed to be removed from the isolated CPU by manipulating the \texttt{/proc/irq/<IRQ>/smp\_affinity} files.
+The IRQs need to be removed from the isolated CPU by manipulating the \texttt{/proc/irq/<IRQ>/smp\_affinity} files.
 The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU.
@@
 Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 2F (10 1111).
-The Python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU, since the \texttt{/proc/interrupts} file does not give away this information and looking individually in the \texttt{/proc/irq/<IRQ>/smp\_affinity} files for every IRQ would be time-consuming.
-The program first reads the CPU affinity settings for each IRQ from the \texttt{/proc/irq} directory and stores the CPUs assigned to each IRQ in a dictionary.
-It then creates a data frame to represent this information in a table format, marking the assigned CPUs with ‘x'.
-Finally, it prints the table and saves it as a markdown file.
-
-\bigskip \noindent However, as the \texttt{proc} filesystem resets to its default state after each reboot, manually changing numerous IRQ files through the said bitmasks is tedious and time-consuming.
+Looking individually in the \texttt{/proc/irq/<IRQ>/smp\_affinity} files for every IRQ would be time-consuming.
+The Python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU, since the \texttt{/proc/interrupts} file only shows if an interrupt has occurred and does not indicate which CPUs are capable of serving which IRQs.
+The program first reads the \texttt{smp\_affinity} files for each IRQ from the \texttt{/proc/irq} directory and stores the CPUs assigned to handle each IRQ in a dictionary.
+It then creates a table showing the distribution of interrupt requests across all CPUs.
+
+\bigskip \noindent As the \texttt{proc} filesystem resets to its default state after each reboot, manually changing numerous IRQ files through the said bitmasks is tedious and time-consuming.
 This process can be automated using the shell script in Code~\ref{script:change_smp_affinity}, which can automatically be executed after every reboot.
 
-\bigskip \noindent Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads.
-This can be done through finding out the respective process IDs through \texttt{ps -e | grep irq/.*nvme} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.
-
 \clearpage
@@
 \vspace*{-5em}
-\begin{lstlisting}[name={Code to show a table of the distribution of interrupt requests across each CPU},label={script:smp_affinity}]
+\begin{lstlisting}[name={Code to show IRQ distribution across each CPU},label={script:smp_affinity}]
 		import os
@@
 Having a real-time kernel itself is an indispensable step of achieving deterministic behavior, but not enough by itself to take full advantage of the real-time capabilities.
-One key aspect of this is real-time priorities, thoroughly explained by Richard Weinberger in~\cite{LinuxProcessPriorities}.
+One key aspect of this is real-time priorities, thoroughly explained by Richard Weinberger~\cite{LinuxProcessPriorities}.
 In essence, Table~\ref{tab:scheduling_priorities} lists minimum and maximum priorities for different scheduling policies in Linux.
@@
 It does this by pausing the process for the remaining 0.05 seconds, which is not desired because this could result in missed deadlines.
-RT throttling can be disabled by writing the value -1 to the \texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file.
-This change also needs to be made permanent because the proc filesystem resets to its default state after each reboot.
+RT throttling can be disabled by writing the value "-1" to the \texttt{/proc/sys/kernel/sched\_rt\_runtime\_us} file.
+This change also needs to be made permanent because the \texttt{proc} filesystem resets to its default state after each reboot.
 For this purpose, the line \texttt{kernel.sched\_rt\_runtime\_us = -1} can be appended to the end of the \texttt{/etc/sysctl.conf} file, which is read at boot time and used to configure kernel parameters.
@@
 This is done by specifying a bitmask to exclude CPU 4 in the files \texttt{/sys/devices/virtual/workqueue/cpumask} and \texttt{/sys/bus/workqueue/devices/writeback/cpumask}.
-In this case, that is \texttt{2F}.
+In this case, that is 2F.
 
@@
 
-\subsubsection{Stop Certain Services}
-Services like irqbalance.service, thermald.service, and wpa\_supplicant.service, as presented in Table~\ref{tab:stop_servies}, can be further sources for random latency and unnecessary overhead.
-Stopping these services through \texttt{sudo systemctl stop <SERVICE>} means they will not be able to interrupt the CPU with their tasks.
+\subsubsection{Stop Services}
+Table~\ref{tab:stop_servies} presents services that can be further sources for random latency and unnecessary overhead.
+Stopping these services through \texttt{sudo systemctl stop <SERVICE>} prevents them from interrupting the CPU with their tasks.
 
@@
 
-\subsection{QEMU/KVM Configurations}\label{subsec:guest_configurations}
+\subsection{QEMU/KVM Configurations}\label{subsec:qemu_configurations}
 KVM allows the guest OS to run directly on the hardware, bypassing the need for traditional emulation which can introduce delays.
@@
 \subsubsection{Set QEMU Options for Real-Time VM}
-QEMU provides several options that can be used to improve the real-time performance of the guest VM. Table~\ref{tab:qemu_options} briefly explains these options.
+QEMU provides options that can be used to improve the real-time performance of the guest VM. Table~\ref{tab:qemu_options} briefly explains these options.
 
@@
 
-\noindent While the worst latency value did not change much, the standard deviation of the maximum latency samples of each second, with the test lasting 600 seconds and thus having 600 worst latencies, decreased.
-This means the system is more stable and predictable, thanks to the real-time tunings that have been performed till this point.
+\noindent The maximum latency value decreased slightly.
+More notably, the standard deviation of the maximum latency samples, with the test lasting 600 seconds and thus having 600 worst latencies for each second, also decreased from 1.828 µs to 0.351 µs.
+Although the difference is not much, when the latency values are this small, it becomes noticeable.
+This means the system is more stable and predictable, thanks to the real-time tunings performed up to this point.
 
@@
 There are other configurations that may be relevant depending on the case.
-Linux Kernel Developer Steven Rostedt explains all relevant aspects of a real-time system that must be considered in~\cite{kernelrecipesKernelRecipes20162016} and gives insight for finding sources of latency on the Linux system in~\cite{thelinuxfoundationFindingSourcesLatency2020}.
+Linux Kernel Developer Steven Rostedt explains all relevant aspects of a real-time system that must be considered in~\cite{kernelrecipesKernelRecipes20162016} and gives insights for finding sources of latency on the Linux system in~\cite{thelinuxfoundationFindingSourcesLatency2020}.
 A checklist for writing Linux real-time applications is provided by John Ogness in~\cite{thelinuxfoundationChecklistWritingLinux2020}.
@@
 ~\cite{HOWTOBuildRTapplication} and~\cite{RealtimeProgrammingLinux} describe the process of writing hard real-time Linux programs using the real-time preemption patch in great detail.
-Various hardware and software tunings are listed in~\cite{KVMQemuVirtualization} and~\cite{RealTimePerformanceTuning2022}.
+Various hardware and software tunings are listed in~\cite{RealTimePerformanceTuning2022}~and~\cite{KVMQemuVirtualization}.
 
@@
 \noindent To drive the motor, it has to be powered using the red and brown wires and can be controlled by sending PWM signals to the orange wire.
-A 1,500 microsecond pulse every 20 milliseconds centers the servo.
-A 1,000 microsecond pulse turns it 90 degrees left, and a 2,000 microsecond pulse turns it 90 degrees right.
+A 1,500 µs pulse every 20 milliseconds centers the servo.
+A 1,000 µs pulse turns it 90 degrees left, and a 2,000 µs pulse turns it 90 degrees right.
 This is depicted in Figure~\ref{fig:servo_angle}.
@@
 Since the mentioned servo motors operate between 4.8 volts and 7.2 volts~\cite{MG996RServoMotor}, this voltage needs to be reduced through resistors before supplying it to a servo motor.
-For this purpose, two resistors with resistances of 2500 kiloohms and 1000 kiloohms are connected in series.
+For this purpose, two resistors with resistances of 2500 ohms and 1000 ohms are connected in series.
 The voltage across each resistor is calculated using Ohm's law, which is given by equation~\ref{eq:ohms_law} below.
@@
 \subsection{Setup of Salamander 4 Bare Metal}
-In this version, Salamander 4 runs on the CP 841~\cite{CPUUnitsSIGMATEK} CPU unit, specifically designed for the Salamander 4 operating system.
+In this version, Salamander 4 runs on the CP 841 CPU unit, specifically designed for the Salamander 4 operating system.
 The PW 022 modules are directly mounted on the CPU via the S-DIAS bus~\cite{SDIASSIGMATEK}.
@@
 The command \texttt{lspci -nn} lists all PCI devices along with their vendor and device IDs.
-In this case, the command for binding the PCV 522 module to the VFIO-PCI driver was \texttt{sudo sh -c 'echo "5112 2200" > /sys/bus/pci/drivers/vfio-pci/new\_id'} with vendor ID \texttt{5112} and device ID \texttt{2200}.
+In this case, the command for binding the PCV 522 module to the VFIO-PCI driver is \texttt{sudo sh -c 'echo "5112 2200" > /sys/bus/pci/drivers/vfio-pci/new\_id'} with vendor ID \texttt{5112} and device ID \texttt{2200}.
 To verify that the device has been successfully bound to the VFIO-PCI driver, \texttt{lspci -v} can be used.
-On top of the binding process, the QEMU script from the previous sections also needed to be modified to include the PCV 522 VARAN Manager PCI Insert Card in the virtualization.
+On top of the binding process, the QEMU script also needs to be modified to include the PCV 522 VARAN Manager PCI Insert Card in the virtualization.
 This is done by adding \texttt{-device\ vfio-pci,host=03:00.0 } to the QEMU script, where \texttt{03:00.0} refers to the PCI address of the device, with \texttt{03} being the bus number, \texttt{00} the device number, and \texttt{0} the function number.
@@
 \vspace*{-5em}
-\begin{lstlisting}[name={QEMU script for Salamander 4 Virtualization with PCI Configuration},label={script:qemu_def_pci}]
+\begin{lstlisting}[name={Final QEMU Script for Salamander 4 Virtualization with PCI Configuration},label={script:qemu_def_pci}]
 		#!/bin/sh
@@
 
-\noindent Essentially, the CPU is being emulated by the vCPU functionality of QEMU, allowing the PCV 522 VARAN Manager PCI Insert Card and the Varan Connection module VI 021 to interact as if they were communicating with a physical CPU unit.
+\noindent Essentially, the CPU is being emulated by the vCPU functionality of QEMU, allowing the PW 022 modules to interact with the vCPU through the PCV 522 Insert Card and the VI 021 module as if they were communicating directly with a physical CPU unit.
 
@@
 
-\bigskip \noindent The impacts of real-time performance tunings on the determinism and latency of Salamander 4 virtualization were already shown in the previous chapter.
-In this chapter, these improvements are presented using a robotic application.
+\bigskip \noindent The impacts of real-time performance tunings on the determinism and latency of Salamander 4 virtualization were already shown in the previous subsection.
+Here, these improvements are demonstrated using a robotic application.
 This application is written in Lasal Class 2, as mentioned in Section~\ref{sec:application_context}.
+The goal is to determine the time from command issuance to the signal reaching the PWM, excluding the 20ms signal period required by the servomotor to process the signal and initiate movement.
 The program sequence is shown in Figure~\ref{fig:robotic_application_flow}.
@@
 		\node [block] (issue) {\scriptsize \textbf{Command Issuance}\\A command is sent to the system to execute a movement.};
-		\node [block, below=of issue] (record1) {\scriptsize \textbf{Recording Command Time}\\The timestamp of the moment the command is recorded.};
-		\node [block, below=of record1] (detect) {\scriptsize \textbf{Detecting Response}\\The moment when the signal reaches the PWM is detected.};
-		\node [block, below=of detect] (record2) {\scriptsize \textbf{Recording Response Time}\\The timestamp of when the signal reaches the PWM is recorded.};
-		\node [block, below=of record2] (calculate) {\scriptsize \textbf{Calculating Latency}\\The latency is determined by the difference between the command issuance time and the response \\\vspace{-0.5em} time at the PWM.};
+		\node [block, below=of issue] (record1) {\scriptsize \textbf{Recording Command Time}\\The timestamp of the moment the command is issued is recorded.};
+		\node [block, below=of record1] (detect) {\scriptsize \textbf{Detecting Command Arrival}\\The moment when the signal reaches the PWM is detected.};
+		\node [block, below=of detect] (record2) {\scriptsize \textbf{Recording Command Arrival Time}\\The timestamp of the moment the signal reaches the PWM is recorded.};
+		\node [block, below=of record2] (calculate) {\scriptsize \textbf{Calculating Command-to-PWM Latency}\\The latency is determined by the difference between the command issuance time and \\\vspace{-0.5em} the command arrival time at the PWM.};
 		% Draw edges
@@
 \end{figure}
+
 \vspace{1em}
@@
 	\includegraphics[width=1.0\columnwidth]{img/implementation/rtos_latencies.png}
-	\caption[Robotic Application Latency Values of Salamander 4 Hardware]{Robotic Application Latency Values of Salamander 4 Hardware}
+	\caption{Robotic Application Latency Values of Salamander 4 Bare Metal}
 	\label{fig:rtos_latencies}
@@
 	\centering
-	\caption{Robotic Application Latency Statistics Hardware}
+	\caption{Robotic Application Latency Statistics Bare Metal}
 	\label{tab:robotic_application_latency_values_hardware}
@@
 
-\subsubsection{Latency in Default Salamander 4 OS Virtualization}\label{subsubsec:latency_virtualization}
-\noindent The latency values acquired for the default virtualization of Salamander 4 OS are demonstrated in Figure~\ref{fig:untuned_virt_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_virt_def}.
+\subsubsection{Latency in Untuned Salamander 4 OS Virtualization}\label{subsubsec:latency_virtualization}
+\noindent The latency values acquired for the untuned virtualization of Salamander 4 OS are demonstrated in Figure~\ref{fig:untuned_virt_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_virt_def}.
 
@@
 \chapter{Results}\label{cha:results}
-The goal was to virtualize the Salamander 4 OS and bring its real-time performance closer to that of the hardware version and guarantee deterministic and reliable behavior.
+The goal was to virtualize the Salamander 4 OS and bring its real-time performance closer to that of the bare metal version and guarantee deterministic and reliable behavior.
 This chapter summarizes the results after the real-time performance tunings, both in terms of the \texttt{latency} program and the robotic application.
@@
 	\includegraphics[width=0.70\columnwidth]{img/results/gnuplot_combined_max_latency_all.png}
-	\caption[Comparison of Latency Distribution of Salamander 4 Configurations]{Comparison of Latency Distribution of Salamander 4 under different Configurations}
+	\caption[Comparison of Latency Distribution of Salamander 4 Configurations]{Comparison of Latency Distribution of Salamander 4 after Configurations}
 	\label{fig:max_latency_combined_results}
@@
 	\footnotesize
-	\caption{Comparison of Latency Statistics in Salamander 4 under different Configurations}
+	\caption{Comparison of Latency Statistics in Salamander 4 after Configurations}
 	\label{tab:latency_tables_combined}
@@
 		\hline
-		\textbf{Tuning}      & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
-		Salamander 4 BM      & 0.613                 & 1.380                 & 10.709                & 0                 \\ \hline
-		Unt. Salamander 4 V. & 2.536                 & 8.940                 & 707.622               & 43                \\ \hline
-		After BIOS Conf.     & 0.969                 & 3.948                 & 457.545               & 94                \\ \hline
-		After Kernel Conf.   & 2.545                 & 4.811                 & 21.694                & 0                 \\ \hline
-		After Host Conf.     & 2.591                 & 4.834                 & 18.441                & 0                 \\ \hline
-		After QEMU Conf.     & 2.614                 & 4.779                 & 17.134                & 0                 \\ \hline
+		\textbf{Salamander 4 Version}      & \textbf{Lat Min (µs)} & \textbf{Lat Avg (µs)} & \textbf{Lat Max (µs)} & \textbf{Overruns} \\ \hline
+		Salamander 4 Bare Metal      & 0.613                 & 1.380                 & 10.709                & 0                 \\ \hline
+		Untuned Salamander 4 Virtualization & 2.536                 & 8.940                 & 707.622               & 43                \\ \hline
+		After BIOS Configurations     & 0.969                 & 3.948                 & 457.545               & 22                \\ \hline
+		After Kernel Configurations   & 2.545                 & 4.811                 & 21.694                & 0                 \\ \hline
+		After Host Configurations     & 2.591                 & 4.834                 & 18.441                & 0                 \\ \hline
+		After QEMU Configurations     & 2.614                 & 4.779                 & 17.134                & 0                 \\ \hline
 	\end{tabular}
@@
 \noindent The improved latency was also tested with the robotic application in Subsection~\ref{sec:robotic_application}.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
-The results are visualized in Figure~\ref{fig:combined_latencies}.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times to get reliable results for the bare metal, untuned, and tuned versions of Salamander 4.
+The results are visualized in Figure~\ref{fig:combined_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_combined}.
 
@@
 	\includegraphics[width=1.0\columnwidth]{img/results/combined_latencies.png}
-	\caption[Comparison of Robotic Application Latency of Salamander 4 Configurations]{Comparison of Robotic Application Latency of Salamander 4 under different Configurations}
+	\caption{Comparison of Robotic Application Latency of Salamander 4 after Configurations}
 	\label{fig:combined_latencies}
 \end{figure}
-
-\noindent This response latency measurement above was performed 1,000 times to get reliable results. These results are presented in Table~\ref{tab:robotic_application_latency_values_combined}.
 
@@
 	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
-		\begin{tabular}{|l|l|l|l|l|l|}
-			\hline
-			\textbf{Tuning}        & \textbf{Samples} & \textbf{Lat Min (ms)} & \textbf{Lat Avg (ms)} & \textbf{Lat Max (ms)} & \textbf{Std Dev (ms)} \\ \hline
-			Bare Metal             & 1000             & 1.211                 & 1.347                 & 1.49                  & 0.082                 \\ \hline
-			Untuned Virtualization & 1000             & 3.1                   & 24.603                & 129.46                & 13.876                \\ \hline
-			Tuned Virtualization   & 1000             & 1.219                 & 2.62                  & 3.988                 & 0.812                 \\ \hline
+		\begin{tabular}{|c|c|c|c|c|}
+			\hline
+			\textbf{Version}        & \textbf{Lat Min (ms)} & \textbf{Lat Avg (ms)} & \textbf{Lat Max (ms)} & \textbf{Std Dev (ms)} \\ \hline
+			Salamander 4 Bare Metal              & 1.211                 & 1.347                 & 1.49                  & 0.082                 \\ \hline
+			Untuned Salamander 4 Virtualization & 3.1                   & 24.603                & 129.46                & 13.876                \\ \hline
+			Tuned Salamander 4 Virtualization   & 1.219                 & 2.62                  & 3.988                 & 0.812                 \\ \hline
 		\end{tabular}}
@@
 \chapter{Discussion}\label{cha:discussion}
-This master's thesis shows that the real-time performance tunings applied to the virtualized Salamander 4 OS have a positive impact on reducing latency and improving determinism.
-The maximum latency in the final tuned version of Salamander 4 Virtualization is significantly lower than in the default setup.
-The worst latency value in a 10-minute measurement span decreased from 707.622~µs to 17.134~µs after tuning BIOS, kernel, host, guest, and QEMU/KVM configurations.
+This master's thesis shows that the PREEMPT-RT patch along with the applied real-time performance tunings have a positive impact on reducing latency and improving determinism.
+The maximum latency in the final tuned version of Salamander 4 virtualization is significantly lower than in the untuned setup.
+The worst latency value in a 10-minute measurement span decreased from 707.622~µs to 17.134~µs after tuning BIOS, kernel, host, QEMU/KVM, and guest configurations.
 Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694~µs.
 From this moment on, there were also no overruns in the test any longer.
-The host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU, further reduced latency a little bit down to 18.441~µs.
+The host configurations, including CPU and interrupt affinity, and real-time prioritization of QEMU, further reduced latency down to 18.441~µs.
 The guest OS, Salamander 4, already had real-time capabilities through Xenomai.
 The focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
-Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages, brought the latency down a little more to the final worst latency value of 17.134~µs in 10 minutes.
+Finally, QEMU/KVM configurations, such as tuning the LAPIC timer advance and using hugepages, brought the latency down to the final worst latency value of 17.134~µs in 10 minutes.
 This value is very close to the bare metal value of 10.709~µs.
@@
 
-\bigskip \noindent As a next step, the robotic application in a practical scenario also demonstrates and validates the improvement of the tuned virtualization compared to the unmodified virtualization.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and hardware versions of Salamander 4.
+\bigskip \noindent As a next step, the robotic application in a practical scenario also demonstrates and validates the improvement of the tuned virtualization.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and bare metal versions of Salamander 4.
 The most important metric here is also the worst latency of the signal reaching the PWM.
 It dropped from 129~ms to 3.988~ms in the developed application.
-Comparing this to the worst latency value of the hardware version, which is 1.49~ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
+Comparing this to the worst latency value of the bare metal version, which is 1.49~ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
 
 \bigskip \noindent It is important to underline one more time the fact that these modifications were added sequentially and tested together.
-First, the BIOS was configured, followed by the kernel.
+After applying the PREEMPT-RT patch, the BIOS was configured, followed by the kernel.
 The BIOS settings were not reverted when moving to the kernel configurations.
@@
 These modifications were added sequentially and tested together.
-First, the BIOS was configured, followed by the kernel.
+After applying the PREEMPT-RT patch, the BIOS was configured, followed by the kernel.
 The BIOS settings were not reverted when moving to the kernel configurations.
 After these two steps, the worst latency decreased from an initial value of 707.62~µs to 21.694~µs.
-From this moment on, there were also no overruns in the test any longer.
-The host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU, further reduced latency a little bit down to 18.441~µs.
+From this moment on, there were also no overruns in the tests any longer.
+The host configurations, including CPU and interrupt affinity, and real-time prioritization of QEMU, further reduced latency down to 18.441~µs.
 The guest OS, Salamander 4, already had real-time capabilities through Xenomai.
@@
 \bigskip \noindent On top of that, a robotic application further confirmed the improvements.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and hardware versions of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and bare metal versions of Salamander 4.
 The worst latency of the signal reaching the PWM dropped from an initial value of 129~ms to 3.988~ms.
-Comparing this to the worst latency value of the hardware version, which was 1.49~ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
+Comparing this to the worst latency value of the bare metal version, which was 1.49~ms, it is apparent that the tuned virtualization through real-time performance configurations came very close to the determinism and reliability of the hardware.
 

### 2024-09-01 python/remove_urldate.py
urldate

@@ -13,6 +13,6 @@ def process_bib_file(input_file, output_file):
                 continue
             file.write(line)
 
-input_file = '../Literature.bib'
-output_file = '../Literature.bib'
+input_file = '../literature/Literature.bib'
+output_file = '../literature/Literature.bib'
 process_bib_file(input_file, output_file)

### 2024-09-02 Masterarbeit_Robotik_Pamuk.tex
minor

@@
 	\centering
-	\caption{Minimum, Average, and Maximum Latency with Overrun Counts of Bare Metal}
+	\caption[Latency Statistics with Overrun Counts of Bare Metal]{Minimum, Average, and Maximum Latency with Overrun Counts of Bare Metal}
 	\label{tab:latency_overrun_msw_hardware}
@@
 	\includegraphics[width=0.7\columnwidth]{img/implementation/gnuplot_combined_max_latency.png}
-	\caption[Initial Comparison of Latency Distribution between Bare Metal and Virtualization]{Initial Comparison of Latency Distribution between Bare Metal and Virtualization}
+	\caption[Initial Comparison of Latency Distribution]{Initial Comparison of Latency Distribution between Bare Metal and Virtualization}
 	\label{fig:gnuplot_max_latency_combined}
@@
 	\includegraphics[width=1.0\columnwidth]{img/results/combined_latencies.png}
-	\caption{Comparison of Robotic Application Latency of Salamander 4 after Configurations}
+	\caption{Comparison of Robotic Application Latency after Configurations}
 	\label{fig:combined_latencies}

### 2024-09-02 Paper.tex
minor

--- 2024-08-31 Paper.tex
@@
 \begin{multicols}{2}
-  \abstract{\textbf{This thesis investigates the virtualization of a real-time operating system, which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
+  \abstract{This thesis investigates the virtualization of a Real-Time Operating System (RTOS), which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
 The primary objective was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is crucial for real-time robotics.
@@
 Extensive tuning across BIOS, kernel, host OS, QEMU/KVM, and guest OS reduced the worst-case latency from 707.622~µs to 17.134~µs, close to the bare metal performance of 10.709~µs.
-The improvement was validated using a robotic application, comparing tuned virtualization with untuned and hardware versions.}} {\textbf{Virtualization, Real-Time Systems, Latency Reduction, Robot Control}}
+The improvement was validated using a robotic application, comparing tuned virtualization with untuned and hardware versions.} {Virtualization, Real-Time Systems, Latency Reduction, Robot Control}
 

### 2024-09-02 resources/scripts/experiment/latencies.py
Deployed 9cc2f894 with MkDocs version: 1.5.3

@@ -1,59 +1,0 @@
-import os
-import numpy as np
-import pandas as pd
-
-# Create the folder
-folder_name = 'generated_latencies'
-os.makedirs(folder_name, exist_ok=True)
-
-# Generate latencies
-np.random.seed(42)  # for reproducibility
-
-def generate_rtos_latencies(size=1000):
-    return np.random.uniform(1.21, 1.49, size).round(3)
-
-def generate_tuned_virt_latencies(size=1000):
-    return np.random.uniform(1.21, 3.99, size).round(3)
-
-def generate_untuned_virt_latencies(size=1000):
-    # Generate most values between 3 and 90
-    base = np.random.uniform(3.1, 44.9, size)
-    # Add some specific outliers
-    outliers = np.array([111, 113, 125, 129, 121])
-    base[:len(outliers)] = outliers
-    np.random.shuffle(base)
-    return np.round(base, 3)
-
-# Generate latencies
-rtos_latencies = generate_rtos_latencies()
-tuned_virt_latencies = generate_tuned_virt_latencies()
-untuned_virt_latencies = generate_untuned_virt_latencies()
-
-# Convert to strings and replace dots with commas
-rtos_latencies = [str(latency).replace('.', ',') for latency in rtos_latencies]
-tuned_virt_latencies = [str(latency).replace('.', ',') for latency in tuned_virt_latencies]
-untuned_virt_latencies = [str(latency).replace('.', ',') for latency in untuned_virt_latencies]
-
-# Create a DataFrame
-data = {
-    'Hardware Salamander 4 (ms)': rtos_latencies,
-    'Tuned Salamander 4 Virtualization (ms)': tuned_virt_latencies,
-    'Untuned Salamander 4 Virtualization (ms)': untuned_virt_latencies
-}
-df = pd.DataFrame(data)
-
-# Save to a single CSV file with semicolon separator
-csv_path = os.path.join(folder_name, 'latencies.csv')
-df.to_csv(csv_path, index=False, sep=';')
-
-# Manually add spaces around the semicolon
-with open(csv_path, 'r') as file:
-    content = file.read()
-
-content = content.replace(';', '     ;     ')
-
-with open(csv_path, 'w') as file:
-    file.write(content)
-
-# Print a success message
-print("The latency values were successfully saved to 'latencies.csv'.")

### 2024-09-02 resources/scripts/experiment/latency_graphs.py

@@ -1,65 +1,0 @@
-import pandas as pd
-import matplotlib.pyplot as plt
-import numpy as np
-
-# Load the data
-csv_path = 'generated_latencies/latencies.csv'
-df = pd.read_csv(csv_path, sep='     ;     ', engine='python')
-
-# Convert columns to numeric values
-df['Hardware Salamander 4 (ms)'] = df['Hardware Salamander 4 (ms)'].str.replace(',', '.').astype(float)
-df['Tuned Salamander 4 Virtualization (ms)'] = df['Tuned Salamander 4 Virtualization (ms)'].str.replace(',', '.').astype(float)
-df['Untuned Salamander 4 Virtualization (ms)'] = df['Untuned Salamander 4 Virtualization (ms)'].str.replace(',', '.').astype(float)
-
-# Plot RTOS Latencies
-plt.figure(figsize=(10, 6))
-plt.plot(df.index, df['Hardware Salamander 4 (ms)'], '-', color='purple', label='Hardware Salamander 4')
-plt.xlabel('Count')
-plt.ylabel('Latency [ms]')
-#plt.title('RTOS Latencies')
-plt.xlim([0, 1000])
-plt.ylim([0, 5])
-#plt.legend()
-plt.savefig("generated_latencies/rtos_latencies.png")
-plt.close()
-
-# Plot Tuned Virtualized Latencies
-plt.figure(figsize=(10, 6))
-plt.plot(df.index, df['Tuned Salamander 4 Virtualization (ms)'], '-', color='red', label='Tuned Salamander 4 Virtualization')
-plt.xlabel('Count')
-plt.ylabel('Latency [ms]')
-#plt.title('Tuned Virtualized Latencies')
-plt.xlim([0, 1000])
-plt.ylim([0, 5])
-#plt.legend()
-plt.savefig("generated_latencies/tuned_virt_latencies.png")
-plt.close()
-
-# Plot Untuned Virtualized Latencies
-plt.figure(figsize=(10, 6))
-plt.plot(df.index, df['Untuned Salamander 4 Virtualization (ms)'], '-', color='blue', label='Untuned Salamander 4 Virtualization')
-plt.xlabel('Count')
-plt.ylabel('Latency [ms]')
-#plt.title('Untuned Virtualized Latencies')
-plt.xlim([0, 1000])
-plt.ylim([0, 150])
-#plt.legend()
-plt.savefig("generated_latencies/untuned_virt_latencies.png")
-plt.close()
-
-# Plot Combined Latencies
[…]

### 2024-09-02 resources/scripts/experiment/latency_statistics.py

@@ -1,45 +1,0 @@
-import pandas as pd
-import numpy as np
-
-# Load the data
-csv_path = 'generated_latencies/latencies.csv'
-df = pd.read_csv(csv_path, sep='     ;     ', engine='python')
-
-# Convert columns to numeric values
-df['Hardware Salamander 4 (ms)'] = df['Hardware Salamander 4 (ms)'].str.replace(',', '.').astype(float)
-df['Tuned Salamander 4 Virtualization (ms)'] = df['Tuned Salamander 4 Virtualization (ms)'].str.replace(',', '.').astype(float)
-df['Untuned Salamander 4 Virtualization (ms)'] = df['Untuned Salamander 4 Virtualization (ms)'].str.replace(',', '.').astype(float)
-
-# Calculate statistics
-def calculate_statistics(latencies):
-    return {
-        'Samples': len(latencies),
-        'Lat Min (ms)': f"{np.min(latencies):.3f}".replace('.', ','),
-        'Lat Avg (ms)': f"{np.mean(latencies):.3f}".replace('.', ','),
-        'Lat Max (ms)': f"{np.max(latencies):.3f}".replace('.', ','),
-        'Std Dev (ms)': f"{np.std(latencies):.3f}".replace('.', ',')
-    }
-
-rtos_stats = calculate_statistics(df['Hardware Salamander 4 (ms)'])
-tuned_virt_stats = calculate_statistics(df['Tuned Salamander 4 Virtualization (ms)'])
-untuned_virt_stats = calculate_statistics(df['Untuned Salamander 4 Virtualization (ms)'])
-
-# Create a DataFrame for the statistics
-stats_df = pd.DataFrame([rtos_stats, tuned_virt_stats, untuned_virt_stats], 
-                        index=['Hardware Salamander', 'Tuned Salamander 4 Virtualization', 'Untuned Salamander 4 Virtualization'])
-
-# Save the statistics to a CSV file with semicolon separator
-stats_csv_path = 'generated_latencies/statistics.csv'
-stats_df.to_csv(stats_csv_path, sep=';')
-
-# Manually add spaces around the semicolon
-with open(stats_csv_path, 'r') as file:
-    content = file.read()
-
-content = content.replace(';', '     ;     ')
-
-with open(stats_csv_path, 'w') as file:
-    file.write(content)
-
-# Print a success message
-print("The statistics were successfully saved to 'statistics.csv'.")

### 2024-09-02 sigmatek/QEMU/qemu_def_2taskset_vsock_nmbridge.sh

@@ -1,16 +1,0 @@
-#!/bin/sh
-
-if  [ ! -d drive-c/ ]; then
-        echo "Filling drive-c/"
-        mkdir drive-c/
-        tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
-fi
-
-exec taskset -c 4 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
--m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
--append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4" \
--net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
--fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
--device vhost-vsock-pci,guest-cid=3,id=vsock0 \
--drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
--no-reboot -nographic

### 2024-09-02 sigmatek/lasalclass2/PW022_hardware/Class/SigCLib/CCrcLib.cpp

@@ -1,0 +1,244 @@
+//{{LSL_ENCRYPTION 16#800
+BDAD8314F342212FF2412A957A14F99590815A18C8ADDA46DDAB1309C1AE38B6EF61236DF250C1ED057C08B3F77B3D7C9536547380BC08EDDC57CBAEF10FB5F72D99D5919AF88BF0F56B3CAA96D0C7B2B3EFFFE94912FEC11DC202A558A16B570D2B5739
+59DA9235EC2B6D345C05DBE3FD5AE936B6D144AA925054813879C69B394A7F00195AD5A0EB0E5D85926CFE86F8CFA2DF7E58BD4856624E3D3B34E767EE6ACEE0C7F50092961A0471ED4A560DF971B063FE390AD0F5F8BDBC9528338430166A5CC53667FD
+7E41F30940143F0EFD994A1833C071A73818B801B30D046379B7865955975EA934C836C0174847DE0B656B19B1C34B90E8B2A2556BBF3A399D12ED8755D5D4CE865E2EBAB34B45AA431F02B82937B760A9D663B824E62D1014C5AB20BDAB5759DA72F9FC
+6F6C6691736C33748FCF6DCAF49E05B79F584F30D218416B9006432FF89BEF8B33ACD08D205ABE8697EC0C28CC2BAB4B9AE6811FCD96F4A3CA083E3324BE27882F307D0AEA7768DD9C295BD798EC1D87B9E2D53B269507484653C5DDFF730AAB341EF568
+30DF15DA93E7FE12C5875D4954C848045C542260DEAFE797C47A1342DE603138C15AFAB06F322DE9E3B1755E8B0D4A878390BB5E73FD723D2AF7CB7B9F99B68A82B63087D1C66D4B36109F6FE9AFA4776F688A71044BB3260DFE9771E26A0A1F16AC4784
+F47FEC98C6E2B764E6530C8A76F915882C497A85FAAB4C1E6BF90D9A8F7B8B622F181A7FFABFCD4C8BC14F9F7495E173400522B03312647D6E51000913E43196DAF963550EFDB9DAA1DC9102D60EA790314ADE9077A5A56FA2DA4C94469838AA6DC506DE
[…]

### 2024-09-02 sigmatek/lasalclass2/PW022_hardware/Include/channels.h

@@ -1,0 +1,354 @@
+//This file was generated by the LASAL2 CodeGenerator  -- 
+//Please, do not edit this file (it might be overwritten by the next generator run)
+TYPE
+  SvrCh_DINT : STRUCT
+    pMeth : ^CHMETH;
+    dData : DINT;
+    pDsc : ^SVRDSC;
+  END_STRUCT;
+  CltCh_DINT : STRUCT
+    pCh : ^SvrCh_DINT;
+    dData : DINT;
+    pCmd : ^SVRCHCMD;
+  END_STRUCT;
+  SvrCh_FeSetup : STRUCT
+    pMeth : ^CHMETH;
+    dData : FeSetup;
+    pDsc : ^SVRDSC;
+  END_STRUCT;
+  CltCh_FeSetup : STRUCT
+    pCh : ^SvrCh_FeSetup;
+    dData : FeSetup;
+    pCmd : ^SVRCHCMD;
+  END_STRUCT;
+  SvrCh_IO_State : STRUCT
+    pMeth : ^CHMETH;
+    dData : IO_State;
+    pDsc : ^SVRDSC;
+  END_STRUCT;
+  CltCh_IO_State : STRUCT
+    pCh : ^SvrCh_IO_State;
+    dData : IO_State;
+    pCmd : ^SVRCHCMD;
+  END_STRUCT;
+  SvrCh_UDINT : STRUCT
+    pMeth : ^CHMETH;
+    dData : UDINT;
+    pDsc : ^SVRDSC;
+  END_STRUCT;
+  CltCh_UDINT : STRUCT
+    pCh : ^SvrCh_UDINT;
+    dData : UDINT;
+    pCmd : ^SVRCHCMD;
+  END_STRUCT;
+  SvrChCmd_t_e_VaranErrors : STRUCT
+    pMeth : ^CMDMETH;
+    dData : t_e_VaranErrors;
+    pDsc : ^SVRDSC;
+  END_STRUCT;
+  CltChCmd__BusInterfaceSdiasDriveMng : STRUCT
+    pCh : ^SvrChCmd_t_e_VaranErrors;
+    dData : t_e_VaranErrors;
+    pCmd : ^_BusInterfaceSdiasDriveMng;
+  END_STRUCT;
+  SvrChCmd_DINT : STRUCT
+    pMeth : ^CMDMETH;
+    dData : DINT;
[…]

### 2024-09-02 sigmatek/lasalclass2/PW022_virtualization/Include/global.h

@@ -1,0 +1,41 @@
+//This file was generated by the LASAL2 CodeGenerator  -- 
+//Please, do not edit this file (it might be overwritten by the next generator run)
+#define  ALIGNMENT 4
+#include <Class.h>
+#include <RTOS_Globals.h>
+#include <RTOS_Variables.h>
+#include <RTOS_PRIVATE.h>
+#include <RTOS_header.h>
+#include <RTOS_PRIVDefs.H>
+#include <RTOS_Scan.h>
+
+#include <Lsl_String.h>
+
+//Project Headerfiles
+#include "..\Class\DiasMasterC\MasterSettings.h"
+#include "..\Class\DiasMaster\DIASMaster.h"
+#include <LSL_STITask.h>
+#include "..\Class\I2CCDIAS\EEFuncDef.h"
+#include "..\Class\VaranManager_Base\VaranManager_Base.h"
+#include "..\Class\_DriveMngBase\DriveComL2.h"
+#include "..\Source\interfaces\lsl_st_sdias.h"
+#include "..\Class\HwControl\HwControlLogging.h"
+#include "..\Source\interfaces\lsl_st_tcp_user.h"
+#include "..\Source\interfaces\lsl_st_varan.h"
+#include "..\Source\interfaces\lsl_st_cdias.h"
+#include <SystemTypes.h>
+#include <File_IOTypes.h>
+#include <MultiTaskTypes.h>
+#include <CanTypes.h>
+#include "..\Class\StringInternal\StringDefs.h"
+#include <lsl_st_serial.h>
+#include <LinkerTypes.h>
+
+#include "unit.h"
+#include "UserDef.h"
+#include "types.h"
+#include "lslpublictypes.h"
+#include "channels.h"
+
+#define LSL_SWAP(x)  (((x) SHL 24) OR ((x AND 16#0000ff00) SHL 8) OR((x AND 16#00ff0000) SHR 8) OR ((x) SHR 24))
+#define LSL_GUID(g0,g1,g2,g3,g4) g0$DINT, g1$INT, g2$INT, LSL_SWAP(g3)$DINT, LSL_SWAP(g4)$DINT

### 2024-09-02 sigmatek/salamander4/latency_reduction/IRQaffinity/check_smp_affinity.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-# Check if a command-line argument is provided
-if [ -z "$1" ]; then
-    echo "Please provide a CPU number as a command-line argument."
-    exit 1
-fi
-# Get the CPU number from the command-line argument
-CPU=$1
-# Check if the CPU exists
-if [ $CPU -ge $(getconf _NPROCESSORS_CONF) ]; then
-    echo "CPU $CPU does not exist."
-    exit 1
-fi
-# Initialize an empty array to store the IRQ numbers
-IRQs=()
-for IRQ in /proc/irq/*; do
-    if [ -f "$IRQ/smp_affinity" ]; then
-        # Read the current smp_affinity
-        AFFINITY=$(cat "$IRQ/smp_affinity")
-        # Check if the bit for the current CPU is set
-        if (( (0x$AFFINITY & (1 << CPU)) != 0 )); then
-            # Add the IRQ number to the array
-            IRQs+=("${IRQ#/proc/irq/}")
-        fi
-    fi
-done
-# Sort the array
-IFS=$'\n' sorted=($(sort -n <<<"${IRQs[*]}"))
-# Print the CPU number
-echo "CPU $CPU IRQ affinity:"
-# Print the sorted IRQ numbers on separate lines
-for irq in "${sorted[@]}"; do
-    echo "$irq"
-done

### 2024-09-02 sigmatek/salamander4/latency_reduction/IRQaffinity/cpuaffinity.sh

@@ -1,37 +1,0 @@
-#!/bin/bash
-
-# Replace 'process_name1' and 'process_name2' with the names of your processes
-process_name1="irq/157-nvme0q5"
-process_name2="irq/172-nvme1q5"
-
-# Initialize a variable to track success status
-success=true
-
-# Find the PID of the first process
-pid1=$(pgrep -f $process_name1)
-
-# Check if the PID was found
-if [ -n "$pid1" ]; then
-    # Set the CPU affinity of the process to CPU 0
-    sudo taskset -p -c 0 $pid1 > /dev/null
-else
-    echo "Process $process_name1 not found."
-    success=false
-fi
-
-# Find the PID of the second process
-pid2=$(pgrep -f $process_name2)
-
-# Check if the PID was found
-if [ -n "$pid2" ]; then
-    # Set the CPU affinity of the process to CPU 0
-    sudo taskset -p -c 0 $pid2 > /dev/null
-else
-    echo "Process $process_name2 not found."
-    success=false
-fi
-
-# Print a success message if no errors occurred
-if $success; then
-    echo "CPU affinity successfully changed."
-fi

### 2024-09-02 sigmatek/salamander4/latency_reduction/IRQaffinity/remove_irqs_from_CPU_PC.sh

@@ -1,46 +1,0 @@
-#!/bin/bash
-
-# Check if a command-line argument is provided
-if [ -z "$1" ]; then
-    echo "Please provide a CPU number as a command-line argument."
-    exit 1
-fi
-
-# Get the CPU number from the command-line argument
-CPU=$1
-
-# Define the mask values
-declare -A mask_values
-mask_values=( [0]="ffe" [1]="ffd" [2]="ffb" [3]="ff7" [4]="fef" [5]="fdf" [6]="fbf" [7]="f7f" [8]="eff" [9]="dff" [10]="bff" [11]="7ff" [12]="fff" [13]="7ff")
-
-# Run the check_smp_affinity.sh script and get the IRQs
-IRQs=$(./check_smp_affinity.sh $CPU | grep -o '[0-9]\+')
-
-# Initialize an empty array to store the IRQs that could not be removed
-failed_IRQs=()
-# Initialize an empty array to store the IRQs that were successfully removed
-succeeded_IRQs=()
-
-# Loop over the IRQs
-for IRQ in $IRQs; do
-    # Try to change the smp_affinity
-    echo ${mask_values[$CPU]} | sudo tee /proc/irq/$IRQ/smp_affinity > /dev/null 2>&1
-    # If the command failed, add the IRQ to the failed_IRQs array
-    if [ $? -ne 0 ]; then
-        failed_IRQs+=($IRQ)
-    else
-        succeeded_IRQs+=($IRQ)
-    fi
-done
-
-# Check if there were any failed IRQs
-if [ ${#failed_IRQs[@]} -ne 0 ]; then
-    echo "IRQs ${failed_IRQs[@]} could not be removed from CPU $CPU."
-fi
-
-# Check if there were any successful IRQs
-if [ ${#succeeded_IRQs[@]} -ne 0 ]; then
-    # Remove the first entry from the succeeded_IRQs array
-    succeeded_IRQs=("${succeeded_IRQs[@]:1}")
-    echo "IRQs ${succeeded_IRQs[@]} were removed from CPU $CPU."
-fi

### 2024-09-02 sigmatek/salamander4/latency_reduction/kernel-patch/compare_2_files.py

@@ -1,53 +1,0 @@
-import os
-import sys
-import json
-
-def compare_files(file1, file2):
-    # Get the base names of the files
-    file1_basename = os.path.basename(file1)
-    file2_basename = os.path.basename(file2)
-
-    # Create the output file name
-    output_file = f"{file1_basename}_VS_{file2_basename}.json"
-
-    # Read the lines from the first file
-    with open(file1, 'r') as f1:
-        lines1 = set(line.strip() for line in f1)
-
-    # Read the lines from the second file
-    with open(file2, 'r') as f2:
-        lines2 = set(line.strip() for line in f2)
-
-    # Find the lines common to both files
-    common_lines = list(lines1 & lines2)
-
-    # Find the lines only in the first file
-    lines_only_in_file1 = list(lines1 - lines2)
-
-    # Find the lines only in the second file
-    lines_only_in_file2 = list(lines2 - lines1)
-
-    # Prepare the results in a dictionary
-    results = {
-        "total common lines": len(common_lines),
-        f"both in {file1_basename} and {file2_basename}": common_lines,
-        f"count only in {file1_basename}": len(lines_only_in_file1),
-        f"only in {file1_basename}": lines_only_in_file1,
-        f"count only in {file2_basename}": len(lines_only_in_file2),
-        f"only in {file2_basename}": lines_only_in_file2
-    }
-
-    # Write the results to the output file in JSON format
-    with open(output_file, 'w') as out:
-        json.dump(results, out, indent=4)
-
-    print(f"Comparison results have been written to {output_file}")
-
-if __name__ == "__main__":
-    # Check if the script received the correct number of arguments
-    if len(sys.argv) != 3:
-        print("Usage: python script.py <file1> <file2>")
-        sys.exit(1)
-
-    # Call the function with the file names passed as command-line arguments
-    compare_files(sys.argv[1], sys.argv[2])

### 2024-09-02 sigmatek/trace-cmd/test_trace-cmd/report_2cpus.sh

@@ -1,43 +1,0 @@
-#!/bin/bash
-# Check if a command line argument is provided
-if [ "$#" -ne 1 ]; then
-    echo "Usage: $0 directory_name"
-    exit 1
-fi
-
-if [ ! -d $1/ ]; then
-    # Use the command line argument as the directory name
-    mkdir $1 
-    mkdir $1/tasks $1/events $1/kvm_exits
-fi
-
-trace-cmd report --cpu 18 > host_report18.txt && echo "host_report18 completed"
-trace-cmd report --cpu 19 > host_report19.txt && echo "host_report19 completed"
-trace-cmd report -i trace-Salamander4.dat > guest_report.txt && echo "guest_report completed"
-cat host_report18.txt | grep kvm_exit | grep "\[018\]" > kvm_exit_count18.txt && echo "kvm_exit_count18 completed"
-cat host_report19.txt | grep kvm_exit | grep "\[019\]" > kvm_exit_count19.txt && echo "kvm_exit_count19 completed"
-# Wait for all background jobs to finish
-wait
-# Now run the Python scripts
-python analyze_tasks.py host_report18.txt && echo "analyze_tasks for host_report18 completed"&
-python analyze_tasks.py host_report19.txt && echo "analyze_tasks for host_report19 completed"&
-python analyze_tasks.py guest_report.txt && echo "analyze_tasks for guest_report completed"&
-python analyze_events.py host_report18.txt && echo "analyze_events for host_report18 completed"&
-python analyze_events.py host_report19.txt && echo "analyze_events for host_report19 completed"&
-python analyze_events.py guest_report.txt && echo "analyze_events for guest_report completed"&
-python kvm_exit_count.py kvm_exit_count18.txt && echo "kvm_exit_count18 plot completed"&
-python kvm_exit_count.py kvm_exit_count19.txt && echo "kvm_exit_count19 plot completed"&
-wait
-# Move and Copy elements in directory 
-mv *.txt $1/
-mv events_host_report* events_guest_report*  $1/events
-mv tasks_host_report* tasks_guest_report*  $1/tasks
-mv kvm_exit_count*.png  $1
-cp merge.py failed_reason.py start_kernelshark.sh *.dat $1
-
-# Merge host and guest report
-cd $1
-#python merge.py && echo "host_report and guest_report merged"
-mv kvm_exit_count* kvm_exits
-rm merge.py
-exit 0

### 2024-09-02 sigmatek/trace-cmd/virtualization/real-time/merge.py

@@ -1,24 +1,0 @@
-import sys
-import re
-
-def merge_files(file_a, file_b, output_file):
-    with open(file_a, 'r') as fa, open(file_b, 'r') as fb:
-        lines_a = fa.readlines()
-        lines_b = fb.readlines()
-
-    # Regular expression pattern for the expected line format
-    pattern = re.compile(r'.*\[\d+\]\s+(\d+\.\d+):.*')
-
-    # Filter out lines that do not match the expected format
-    lines_a = [line for line in lines_a if pattern.match(line)]
-    lines_b = [line for line in lines_b if pattern.match(line)]
-
-    # Extract timestamps and sort lines from both files
-    lines = sorted(lines_a + lines_b, key=lambda line: float(pattern.match(line).group(1)))
-
-    with open(output_file, 'w') as fo:
-        for line in lines:
-            fo.write(line)
-
-# Call the function with your file paths
-merge_files('host_report.txt', 'guest_report.txt', 'host_guest_merged.txt')

### 2024-09-02 sigmatek/trace-cmd/virtualization/vapic/report.sh

@@ -1,11 +1,0 @@
-#!/bin/bash
-trace-cmd report | grep kvm_exit > kvm_exit_test.txt && echo "kvm_exit_test completed"&
-trace-cmd report --cpu 19 > host_report.txt && echo "host_report completed"&
-trace-cmd report -i trace-Salamander4.dat > guest_report.txt && echo "guest_report completed"&
-# Wait for all background jobs to finish
-wait
-# Now run the Python scripts
-python analyze_trace.py host_report.txt && echo "analyze_trace for host_report completed"&
-python analyze_trace.py guest_report.txt && echo "analyze_trace for guest_report completed"&
-wait
-exit 0

### 2024-09-02 sigmatek/xenomai/0hardware/max_latency_hardware/max_latency_hardware.py

@@ -1,52 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-
-# Data points
-x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,41
[…]

### 2024-09-02 sigmatek/xenomai/1default/max_latency_default/max_latency_default.py

@@ -1,53 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-
-# Data points
-x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,445,446,447,448,449,450,451
[…]

### 2024-09-02 sigmatek/xenomai/5rt_kernelparam_host/max_latency_rt_kernelparam_host/max_latency_rt_kernelparam_host.py

@@ -1,55 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-
-# Data points
-x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,
[…]

### 2024-09-02 sigmatek/xenomai/7rt_kernelparam_host_qemu_2cpus/max_latency_rt_kernelparam_host_qemu_2cpus/max_latency_rt_kernelparam_host_qemu_2cpus.py

@@ -1,54 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-
-# Data points
-x = [1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115,116,117,118,119,120,121,122,123,124,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,140,141,142,143,144,145,146,147,148,149,150,151,152,153,154,155,156,157,158,159,160,161,162,163,164,165,166,167,168,169,170,171,172,173,174,175,176,177,178,179,180,181,182,183,184,185,186,187,188,189,190,191,192,193,194,195,196,197,198,199,200,201,202,203,204,205,206,207,208,209,210,211,212,213,214,215,216,217,218,219,220,221,222,223,224,225,226,227,228,229,230,231,232,233,234,235,236,237,238,239,240,241,242,243,244,245,246,247,248,249,250,251,252,253,254,255,256,257,258,259,260,261,262,263,264,265,266,267,268,269,270,271,272,273,274,275,276,277,278,279,280,281,282,283,284,285,286,287,288,289,290,291,292,293,294,295,296,297,298,299,300,301,302,303,304,305,306,307,308,309,310,311,312,313,314,315,316,317,318,319,320,321,322,323,324,325,326,327,328,329,330,331,332,333,334,335,336,337,338,339,340,341,342,343,344,345,346,347,348,349,350,351,352,353,354,355,356,357,358,359,360,361,362,363,364,365,366,367,368,369,370,371,372,373,374,375,376,377,378,379,380,381,382,383,384,385,386,387,388,389,390,391,392,393,394,395,396,397,398,399,400,401,402,403,404,405,406,407,408,409,410,411,412,413,414,415,416,417,418,419,420,421,422,423,424,425,426,427,428,429,430,431,432,433,434,435,436,437,438,439,440,441,442,443,444,445,446,447
[…]

### 2024-09-02 sigmatek/xenomai/print_numbers.py

@@ -1,9 +1,0 @@
-# Open the file in write mode
-with open("print_numbers.txt", "w") as file:
-    # Write numbers from 1 to 600
-    for number in range(1, 53608):
-        file.write(str(number) + "\n")
-
-# Print a success message
-print("The numbers from 1 to x are successfully saved to numbers.txt file.")
-

### 2024-09-02 sigmatek/yocto/after_bitbake_2.sh

@@ -1,13 +1,0 @@
-#!/bin/bash
-
-ip_address="192.168.1.74"
-
-#cd /home/sigma_ibo/Develop/Yocto_local/salamander/salamander-core2/build/tmp/deploy/ipk/core2-64
-#ssh root@"$ip_address" 'mkdir bb'
-ssh-keygen -f "/home/sigma_ibo/.ssh/known_hosts" -R "$ip_address"
-ssh-copy-id root@"$ip_address"
-scp /usr/local/bin/trace-cmd root@"$ip_address":/usr/bin
-scp /usr/local/lib64/libtracefs.so.1 root@"$ip_address":/lib64
-scp /usr/local/lib64/libtraceevent.so.1 root@"$ip_address":/lib64
-#ssh root@"$ip_address" 'opkg install /home/root/bb/trace-cmd_2.9.1-r0_core2-64.ipk'
-#start_qemu.sh

### 2024-09-02 sigmatek/yocto/qemu_def.sh

@@ -1,19 +1,0 @@
-#!/bin/sh
-
-if  [ ! -d drive-c/ ]; then
-                echo "Filling drive-c/"
-                mkdir drive-c/
-                tar -C drive-c/ -xf stek-drive-c-image-sigmatek-core2.tar.gz
-fi
-
-exec taskset -c 4 qemu-system-x86_64 -M pc,accel=kvm -kernel ./bzImage \
--m 2048 -drive file=salamander-image-sigmatek-core2.ext4,format=raw,media=disk \
--append "console=ttyS0 console=tty1 root=/dev/sda rw panic=1 sigmatek_lrt.QEMU=1 ip=dhcp rootfstype=ext4 schedstats=enable nohlt idle=poll quiet xeno_hal.smi=1 xenomai.smi=1 threadirqs" \
--net nic,model=e1000,netdev=e1000 -netdev bridge,id=e1000,br=nm-bridge \
--fsdev local,security_model=none,id=fsdev0,path=drive-c -device virtio-9p-pci,id=fs0,fsdev=fsdev0,mount_tag=/mnt/drive-C \
--device vhost-vsock-pci,guest-cid=3,id=vsock0 \
--drive if=pflash,format=qcow2,file=ovmf.code.qcow2 \
--object memory-backend-ram,id=ram0,size=4G,prealloc=on \
--mem-prealloc -mem-path /dev/hugepages \
--device vfio-pci,host=03:00.0 \
--no-reboot -nographic

### 2024-09-02 tools/ps/process-tree.py

@@ -1,35 +1,0 @@
-import psutil
-
-def print_process_tree(pid):
-    # Create a list to store the process information
-    processes = []
-
-    def helper(pid):
-        try:
-            process = psutil.Process(pid)
-            # Add process information to the list
-            processes.append([process.pid, process.name(), process.ppid()])
-            children = process.children()
-            for child in children:
-                helper(child.pid)
-        except psutil.NoSuchProcess:
-            pass
-
-    helper(pid)
-
-    # Sort the processes by PID
-    processes.sort(key=lambda x: x[0])
-
-    # Write the table to a .md file
-    with open("process-tree.md", "w") as f:
-        f.write("| PID | Process Name | PPID |\n")
-        f.write("| --- | --- | --- |\n")
-        for process in processes:
-            f.write(f"| {process[0]} | {process[1]} | {process[2]} |\n")
-
-# Start with the root process
-print_process_tree(1)
-
-# Print a success message
-print("The process tree is successfully saved to process_tree.md file.")
-

### 2024-09-02 tools/ps/ps-e.py

@@ -1,86 +1,0 @@
-# Import the required module
-import collections
-
-# Define the file name
-file_name = "ps-e.txt"
-
-# Initialize a dictionary to store the counts
-counts = collections.defaultdict(int)
-group_counts = collections.defaultdict(int)
-
-# Initialize a variable to store the total count
-total_count = 0
-total_group_count = 0
-
-# Initialize variables to store the count of kernel and user tasks
-kernel_count = 0
-user_count = 0
-group_kernel_count = 0
-group_user_count = 0
-
-# Open the file and read line by line
-with open(file_name, 'r') as file:
-    for line in file:
-        # Split the line into words
-        words = line.split()
-        
-        # If the line has at least 4 words
-        if len(words) >= 4:
-            # The task name is the fourth word
-            task_name = words[3]
-            
-            # Determine if the task is a kernel task or a user task
-            task_type = "Kernel" if task_name.startswith("k") else "User"
-            
-            # Increment the count of this task
-            counts[(task_name, task_type)] += 1
-            
-            # Group tasks
-            group_name = task_name.split('/')[0]
-            group_counts[(group_name, task_type)] += 1
-            
-            # Increment the total count
-            total_count += 1
-            
-            # Increment the count of kernel or user tasks
-            if task_type == "Kernel":
-                kernel_count += 1
-                group_kernel_count += 1
-            else:
-                user_count += 1
-                group_user_count += 1
-
-# Count the number of unique task groups
-total_group_count = len(group_counts)
-
-# Sort the tasks by count in descending order
-sorted_counts = sorted(counts.items(), key=lambda item: item[1], reverse=True)
-sorted_group_counts = sorted(group_counts.items(), key=lambda item: item[1], reverse=True)
-
-# Open the output file and write the counts in a table format
-with open("ps-e_amount.md", 'w') as file:
-    # Write the table header
-    file.write(f"Total tasks count (individual): {total_count}\n")
-
-    file.write("| Task | Count |\n")
-    file.write("| --- | --- |\n")
-
-    # Write the counts
-    for (task_name, task_type), count in sorted_counts:
-        file.write(f"| {task_name} | {count} |\n")
-    
-# Open the group output file and write the counts in a table format
-with open("ps-e_amount_group.md", 'w') as file:
-    # Write the table header
-    file.write(f"Total tasks count (grouped): {total_group_count}\n")
-
-    file.write("| Task Group | Count |\n")
[…]

### 2024-09-02 tools/ps/show_all_threads.sh

@@ -1,8 +1,0 @@
-#!/bin/bash
-ps axHo pid,lwp,psr,comm,policy,nice,rtprio > show_all_threads.txt
-python show_all_threads.py
-if [ ! -d "show_all_threads_generated" ]; then
-  mkdir show_all_threads_generated
-fi
-mv show_all_threads.txt show_all_threads.md show_all_threads_grouped.md show_all_threads_generated
-

### 2024-09-02 tools/rt-tester/rt-tester/rt-tester.cpp

@@ -1,226 +1,0 @@
-﻿/*
- * Real Time Tester
- *
- * Based on:
- * https://wiki.linuxfoundation.org/realtime/documentation/howto/applications/application_base
- * https://wiki.linuxfoundation.org/realtime/documentation/howto/applications/cyclic
- *
- * This was built using VS on Windows and WSL as the Target System
- * 
- * Use vcpkg to install pthreads
- * https://github.com/microsoft/vcpkg
- *
- * Add CMAKE_TOOLCHAIN_FILE variable to CMakePresets.json
- * "CMAKE_TOOLCHAIN_FILE": "[path to vcpkg]/scripts/buildsystems/vcpkg.cmake"
- *
- */
-
-#include <limits.h>
-#include <pthread.h>
-#include <sched.h>
-#include <stdbool.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <unistd.h>
-#include <sys/mman.h>
-#include <sys/syscall.h>
-
-#include "rt-tester.h"
-#include <string>
-
-#if defined(_WIN32)
-#define OS "Windows"
-#elif defined(__linux__)
-#define OS "Linux"
-#elif defined(__APPLE__)
-#define OS "MacOS"
-#else
-#define OS "Unknown OS"
-#endif
-
-void print_logo(long tid, int policy, int priority, long double period_ms, long print_per_sec)
-{
-    printf(R"(                              
-  ____ _____   _            _            
- |  _ \_   _| | |_ ___  ___| |_ ___ _ __ 
- | |_) || |   | __/ _ \/ __| __/ _ \ '__|
- |  _ < | |   | ||  __/\__ \ ||  __/ |   
- |_| \_\|_|    \__\___||___/\__\___|_|    ver: %s                                       
-
- scheduler policy: %s
- priority: %d
-
- period: %.4Lf ms   
- console refresh rate: %ld Hz
-
-)", VERSION, 
-    //tid, 
-    policy == SCHED_OTHER ? "SCHED_OTHER" :
-   (policy == SCHED_RR ? "SCHED_RR" :
-   (policy == SCHED_FIFO ? "SCHED_FIFO" : "Unknown...")), 
-    priority, 
-    period_ms, 
-    print_per_sec);
-}
-
-long double diff_nanosec(const struct timespec* time1, const struct timespec* time0)
-{
-    return 1000000000 * (time1->tv_sec - time0->tv_sec) + (time1->tv_nsec - time0->tv_nsec);
-}
-
[…]

### 2024-09-02 Masterarbeit_Robotik_Pamuk.tex
FINAL

@@
 	\item The \texttt{latency} tool of the Xenomai suite to measure the scheduling latency of the real-time OS was suggested to be used by SIGMATEK.
-The goal was set to achieve latency values below 50 microseconds in the duration of the measurement.
+The goal was set to achieve latency values below 50 µs in the duration of the measurement.
 	\item SIGMATEK uses the VARAN bus for communication between machines and systems, a real-time Ethernet network that was specially developed for industrial automation.
@@
 Section~\ref{sec:starting_situation} outlines the initial situation of this master's thesis and serves as the foundation for the real-time performance tunings discussed in Section~\ref{sec:real-time_tunings}.
-The objective is to achieve latency values below 50 microseconds during the measurement period through real-time tunings, with no outliers whatsoever.
+The objective is to achieve latency values below 50 µs during the measurement period through real-time tunings, with no outliers whatsoever.
 To determine and validate the performance of the tunings, the \texttt{latency} tool is used, as stated in Section~\ref{sec:application_context}.
@@
 		\node [block, below=of issue] (record1) {\scriptsize \textbf{Recording Command Time}\\The timestamp of the moment the command is issued is recorded.};
-		\node [block, below=of record1] (detect) {\scriptsize \textbf{Detecting Command Arrival}\\The moment when the signal reaches the PWM is detected.};
-		\node [block, below=of detect] (record2) {\scriptsize \textbf{Recording Command Arrival Time}\\The timestamp of the moment the signal reaches the PWM is recorded.};
-		\node [block, below=of record2] (calculate) {\scriptsize \textbf{Calculating Command-to-PWM Latency}\\The latency is determined by the difference between the command issuance time and \\\vspace{-0.5em} the command arrival time at the PWM.};
+		\node [block, below=of record1] (detect) {\scriptsize \textbf{Detecting Command Arrival}\\The moment when the signal reaches the PWM module is detected.};
+		\node [block, below=of detect] (record2) {\scriptsize \textbf{Recording Command Arrival Time}\\The timestamp of the moment the signal reaches the PWM module is recorded.};
+		\node [block, below=of record2] (calculate) {\scriptsize \textbf{Calculating Command-to-PWM Latency}\\The latency is determined by the difference between the command issuance time and \\\vspace{-0.5em} the command arrival time at the PWM module.};
 		% Draw edges
@@
 \noindent The improved latency was also tested with the robotic application in Subsection~\ref{sec:robotic_application}.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times to get reliable results for the bare metal, untuned, and tuned versions of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM module was measured 1,000 times to get reliable results for the bare metal, untuned, and tuned versions of Salamander 4.
 The results are visualized in Figure~\ref{fig:combined_latencies} and detailed in Table~\ref{tab:robotic_application_latency_values_combined}.
@@
 This value is very close to the bare metal value of 10.709~µs.
-The goal was set to achieving latency values below 50 microseconds in the duration of the measurement.
+The goal was set to achieving latency values below 50 µs in the duration of the measurement.
 This goal is achieved.
@@
 \bigskip \noindent As a next step, the robotic application in a practical scenario also demonstrates and validates the improvement of the tuned virtualization.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and bare metal versions of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM module was measured 1,000 times for the untuned, tuned, and bare metal versions of Salamander 4.
 The most important metric here is also the worst latency of the signal reaching the PWM.
@@
 \bigskip \noindent The initial latency values were first measured with the \texttt{latency} program of the Xenomai tool suite.
-The tests were conducted with a sampling period of 100 µs, using a periodic user-mode task, and were assigned a priority of 99.
+The tests were conducted for 10 minutes with a sampling period of 100 µs, using a periodic user-mode task, and were assigned a priority of 99.
 If any sample's latency value was greater than 100~µs, this was considered an overrun.
@@
 This value is very close to the bare metal value of 10.709~µs.
-The goal was set to achieving latency values below 50 microseconds in the duration of the measurement.
+The goal was set to achieving latency values below 50 µs in the duration of the measurement.
 This goal was achieved.
@@
 \bigskip \noindent On top of that, a robotic application further confirmed the improvements.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and bare metal versions of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM module was measured 1,000 times for the untuned, tuned, and bare metal versions of Salamander 4.
 The worst latency of the signal reaching the PWM dropped from an initial value of 129~ms to 3.988~ms.

### 2024-09-02 Paper.tex
FINAL

@@
 \begin{multicols}{2}
-  \abstract{This thesis investigates the virtualization of a Real-Time Operating System (RTOS), which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
-The primary objective was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is crucial for real-time robotics.
+  \abstract{This thesis investigates the virtualization of a Real-Time Operating System, which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
+The primary objective was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is decisive for real-time robotics.
 Initial latency measurements showed a high gap between bare metal and virtualized setups.
-Extensive tuning across BIOS, kernel, host OS, QEMU/KVM, and guest OS reduced the worst-case latency from 707.622~µs to 17.134~µs, close to the bare metal performance of 10.709~µs.
-The improvement was validated using a robotic application, comparing tuned virtualization with untuned and hardware versions.} {Virtualization, Real-Time Systems, Latency Reduction, Robot Control}
+Extensive tunings across BIOS, kernel, host OS, QEMU/KVM, and guest OS were performed, which reduced the worst-case latency from 707.622~µs to 17.134~µs, close to the bare metal performance of 10.709~µs.
+The improvement was validated using a robotic application, comparing the tuned virtualization with the untuned and bare metal versions.} {Virtualization, Real-Time Systems, Latency Reduction, Robot Control}
 
@@
 Delays or errors can have catastrophic consequences.
-In a traditional general-purpose operating system (GPOS), a high-priority task cannot interrupt a kernel operation, which makes them unsuitable for real-time requirements, as they cannot guarantee deterministic execution times.
-However, in a real-time operating system (RTOS), a high-priority process can interrupt a lower-priority task, even if it is in the middle of a kernel operation.
+In a traditional General-Purpose Operating System (GPOS), a high-priority task cannot interrupt a kernel operation, which makes them unsuitable for real-time requirements, as they cannot guarantee deterministic execution times.
+However, in a Real-Time Operating System (RTOS), a high-priority process can interrupt a lower-priority one, even if it is in the middle of a kernel operation.
 Virtualizing real-time operating systems for robotic control offers many advantages over traditional hardware-based approaches, especially when it comes to scaling~\cite{abbasiExploringOpenStackScalable2023}, flexibility, and costs~\cite{taccariEmbeddedRealTimeVirtualization2014}.
 However, virtualization also introduces overhead and latency~\cite{casiniLatencyAnalysisVirtualization2021}~\cite{zhangEvaluatingOptimizingVirtualization2010}.
-In this master's thesis, the goal was the virtualization of a real-time operating system to control a robot, with a focus on compliance with real-time determinism.
+In this master's thesis, the goal was the virtualization of an RTOS, namely Salamander 4, to control a robot, with a focus on compliance with real-time determinism.
 Specifically, the aim was to bring the latency of the virtualized Salamander 4 closer to that of the bare metal version, as initial latency measurements revealed a significant latency gap between the bare metal and virtualized setup.
@@
 
-  \section{Implementation}
+  \section{Practical implementation}
   The initial latency values were first measured with the \texttt{latency} program of the Xenomai tool suite.
-The tests were conducted with a sampling period of 100~µs, using a periodic user-mode task, and were assigned a priority of 99.
+The tests were conducted for 10 minutes with a sampling period of 100~µs, using a periodic user-mode task, and were assigned a priority of 99.
 If any sample's latency value was greater than 100~µs, this was considered an overrun.
-There was a significant initial gap in latency statistics between the bare metal and virtualized Salamander 4.
+There was a significant initial gap in latency between the bare metal and virtualized Salamander 4.
 To address this gap, an extensive tuning process was carried out to achieve real-time performance and determinism.
@@
 On top of that, a robotic application further confirmed the improvements.
-The difference between the command time and the time the signal reaches the PWM was measured 1,000 times for the untuned, tuned, and hardware versions of Salamander 4.
+The difference between the command time and the time the signal reaches the Pulse Width Modulation (PWM) Module  was measured 1,000 times to get reliable results for the bare metal, untuned, and tuned versions of Salamander 4.
 
@@
   This section summarizes the results after the real-time performance tunings, both in terms of the \texttt{latency} program and the robotic application.
+
+  \clearpage
+
   Figure~\ref{fig:max_latency_combined_results} displays the change in latency after each real-time performance tuning.
@@
     \includegraphics[width=1.0\columnwidth]{img/results/gnuplot_combined_max_latency_all.png}
-    \caption[Comparison of Latency Distribution of Salamander 4 Configurations]{Comparison of Latency Distribution of Salamander 4 under different Configurations}
+    \caption{Comparison of Latency Distribution of Salamander 4 after Configurations}
     \label{fig:max_latency_combined_results}
@@
     \footnotesize
-    \caption{Comparison of Latency Statistics in Salamander 4 under different Configurations}
+    \caption{Comparison of Latency Statistics in Salamander 4 after Configurations}
     \label{tab:latency_tables_combined}
@@
       \hline
-      \multirow{2}{*}{\textbf{Tuning}} & \multicolumn{3}{c|}{\textbf{Latency (µs)}} & \multirow{2}{*}{\textbf{Overruns}} \\ \cline{2-4}
+      \multirow{2}{*}{\textbf{Version}} & \multicolumn{3}{c|}{\textbf{Latency (µs)}} & \multirow{2}{*}{\textbf{Overruns}} \\ \cline{2-4}
       & \textbf{Min} & \textbf{Avg} & \textbf{Max} & \\ \hline
@@
 The improved latency was also tested with a robotic application.
-The difference between the command issuance time and the time the signal reaches the PWM was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
+The difference between the command issuance time and the time the signal reaches the PWM module was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
 The results are visualized in Table~\ref{tab:robotic_application_latency_values_combined}.
@@
       \hline
-      \multirow{2}{*}{\textbf{Tuning}} & \multicolumn{3}{c|}{\textbf{Latency (ms)}} & \multirow{2}{*}{\begin{tabular}[c]{@{}c@{}}\textbf{Std Dev}\\ \textbf{(ms)}\end{tabular}} \\ \cline{2-4}
+      \multirow{2}{*}{\textbf{Version}} & \multicolumn{3}{c|}{\textbf{Latency (ms)}} & \multirow{2}{*}{\begin{tabular}[c]{@{}c@{}}\textbf{Std Dev}\\ \textbf{(ms)}\end{tabular}} \\ \cline{2-4}
       & \textbf{Min} & \textbf{Avg} & \textbf{Max} & \\ \hline
@@
   The results show that the real-time performance tunings applied to the virtualized Salamander 4 OS reduced latency and improved determinism.
-The maximum latency decreased from 707.622~µs to 17.134~µs after tuning BIOS, kernel, host, guest, and QEMU/KVM configurations.
+The maximum latency decreased from 707.622~µs to 17.134~µs after tuning BIOS, kernel, host, and QEMU/KVM configurations.
 Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694~µs.
 This also eliminated overruns going forward.
-Host configurations, including the PREEMPT-RT patch, CPU and interrupt affinity, and real-time prioritization of QEMU further reduced latency a little bit down to 18.441~µs.
+The host configurations, including CPU and interrupt affinity, and real-time prioritization of QEMU, further reduced latency down to 18.441~µs.
 The guest OS already had real-time capabilities through Xenomai, hence the focus was primarily on optimizing the host and virtualization layer to achieve the desired real-time performance.
@@
 This value is very close to the bare metal value of 10.709~µs.
-The goal was set to achieving latency values below 50 microseconds in the duration of the measurement.
+The goal was set to achieving latency values below 50~µs in the duration of the measurement.
 This goal was achieved.
   A robotic application in a practical scenario validated the improvement of the tuned virtualization compared to the unmodified version.
-The difference between command time and signal reaching the PWM was measured 1,000 times for untuned, tuned, and hardware versions of Salamander 4.
-The worst latency dropped from 129ms to 3.988ms in the developed application.
-Compared to the hardware version's worst latency of 1.49ms, the tuned virtualization came very close to the determinism and reliability of the hardware.
+The difference between command time and signal reaching the PWM module was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
+The worst latency dropped from 129~ms to 3.988~ms in the developed application.
+Compared to the bare metal version's worst latency of 1.49~ms, the tuned virtualization came very close to the determinism and reliability of the bare metal.
 
@@
 The goal was to reduce latency in virtualized Salamander 4 to match its bare metal version.
-Initial tests showed a significant latency gap between virtualized and bare metal versions.
-Extensive tuning was performed sequentially: BIOS, kernel, host, and QEMU/KVM configurations.
+Initial tests showed a significant latency gap between bare metal and virtualized versions.
+Extensive tuning was performed sequentially, including BIOS, kernel, host, and QEMU/KVM configurations.
 This process reduced worst latency from 707.62~µs to 17.134~µs, close to the bare metal value of 10.709~µs, achieving the goal of sub-50~µs latency.
   A robotic application confirmed these improvements.
-Measuring the time between command and PWM signal 1,000 times showed worst latency dropping from 129ms to 3.988ms, approaching the hardware version's 1.49ms.
-  Future work could include further optimizations, testing other hypervisors, and evaluating performance under various workloads and stressed situations.
+Measuring the time between command issuance and command arrival at the PWM module 1,000 times showed worst latency dropping from 129~ms to 3.988~ms, approaching the bare metal version's 1.49~ms.
   Future work could include additional configurations and optimizations of the virtualization layer, investigating other hypervisors and virtualization technologies, and extensive testing under various workloads and stressed situations.

### 2024-09-02 Paper.tex
rtos

@@
 \begin{multicols}{2}
-  \abstract{This thesis investigates the virtualization of a Real-Time Operating System, which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
+  \abstract{This thesis investigates the virtualization of a real-time operating system, which is built with Yocto, employs hard real-time through Xenomai 3, and is emulated using QEMU/KVM.
 The primary objective was to bridge the latency gap between virtualized and bare metal versions to ensure deterministic and reliable behavior, which is decisive for real-time robotics.

### 2024-09-02 Masterarbeit_Robotik_Pamuk.tex
Fertiggggg

@@
 \bigskip \noindent As a next step, the robotic application in a practical scenario also demonstrates and validates the improvement of the tuned virtualization.
-The difference between the command time and the time the signal reaches the PWM module was measured 1,000 times for the untuned, tuned, and bare metal versions of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM module was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
 The most important metric here is also the worst latency of the signal reaching the PWM.
@@
 \bigskip \noindent On top of that, a robotic application further confirmed the improvements.
-The difference between the command time and the time the signal reaches the PWM module was measured 1,000 times for the untuned, tuned, and bare metal versions of Salamander 4.
+The difference between the command time and the time the signal reaches the PWM module was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
 The worst latency of the signal reaching the PWM dropped from an initial value of 129~ms to 3.988~ms.

### 2024-09-02 Paper.tex
Fertiggggg

@@
 This involves configurations spanning the BIOS, kernel, host operating system (OS), and QEMU/KVM virtualization layer, as demonstrated in~\cite{RealTimePerformanceTuning2022}.
-The guest OS is already based on Xenomai and therefore, no additional modifications were necessary.
+Xenomai provided real-time capabilities for the guest OS and therefore, no additional modifications were necessary.
 Ubuntu 22.04.4 LTS served as the host operating system, while Salamander 4, built with Yocto and based on Linux 5.15.94, was the guest OS.
-Xenomai provided real-time capabilities for the guest and its \texttt{latency} tool was used to measure the scheduling latency of the real-time OS.
+The \texttt{latency} tool of the Xenomai tool suite was used to measure the scheduling latency of the real-time OS.
 QEMU with KVM was employed for virtualization.
@@
 Finally, QEMU/KVM settings were applied.
-On top of that, a robotic application further confirmed the improvements.
+A robotic application further confirmed the improvements.
 The difference between the command time and the time the signal reaches the Pulse Width Modulation (PWM) Module  was measured 1,000 times to get reliable results for the bare metal, untuned, and tuned versions of Salamander 4.
@@
 
-  Figure~\ref{fig:max_latency_combined_results} displays the change in latency after each real-time performance tuning.
+  Figure~\ref{fig:max_latency_combined_results} displays the change in latency after each real-time performance tuning of Salamander 4.
 
@@
       \begin{tabular}[c]{@{}c@{}}Untuned\\ Virtualization\end{tabular} & 2.536 & 8.940 & 707.622 & 43 \\ \hline
-      \begin{tabular}[c]{@{}c@{}}After BIOS\\ Configurations\end{tabular} & 0.969 & 3.948 & 457.545 & 94 \\ \hline
+      \begin{tabular}[c]{@{}c@{}}After BIOS\\ Configurations\end{tabular} & 0.969 & 3.948 & 457.545 & 22 \\ \hline
       \begin{tabular}[c]{@{}c@{}}After Kernel\\ Configurations\end{tabular} & 2.545 & 4.811 & 21.694 & 0 \\ \hline
@@
 The maximum latency decreased from 707.622~µs to 17.134~µs after tuning BIOS, kernel, host, and QEMU/KVM configurations.
-Especially the BIOS and kernel configurations played a crucial role in reducing the maximum latency, with a major fall to 21.694~µs.
+Especially, applying the PREEMPT-RT patch, along with optimizing both the BIOS and kernel configurations, played a crucial role in reducing the maximum latency to 21.694~µs.
 This also eliminated overruns going forward.
@@
   A robotic application in a practical scenario validated the improvement of the tuned virtualization compared to the unmodified version.
-The difference between command time and signal reaching the PWM module was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
-The worst latency dropped from 129~ms to 3.988~ms in the developed application.
+The difference between the command time and the time the signal reaches the PWM module was measured 1,000 times for the bare metal, untuned, and tuned versions of Salamander 4.
+The worst latency in the virtualization dropped from 129~ms to 3.988~ms in the developed application.
 Compared to the bare metal version's worst latency of 1.49~ms, the tuned virtualization came very close to the determinism and reliability of the bare metal.
@@
   \section{Summary and Outlook}
-  This master's thesis aimed to virtualize a real-time operating system for robot control, focusing on real-time determinism.
-The goal was to reduce latency in virtualized Salamander 4 to match its bare metal version.
+  This master's thesis aimed to virtualize an RTOS for robot control, focusing on real-time determinism.
+The goal was to reduce latency in the virtualized Salamander 4 to match its bare metal version.
 Initial tests showed a significant latency gap between bare metal and virtualized versions.
 Extensive tuning was performed sequentially, including BIOS, kernel, host, and QEMU/KVM configurations.
-This process reduced worst latency from 707.62~µs to 17.134~µs, close to the bare metal value of 10.709~µs, achieving the goal of sub-50~µs latency.
+This process reduced the worst latency from 707.62~µs to 17.134~µs, close to the bare metal value of 10.709~µs, achieving the goal of sub-50~µs latency.
   A robotic application confirmed these improvements.

### 2024-09-02 Masterarbeit_Robotik_Pamuk.tex
Fertiggggg

@@
 	\centering
-	\footnotesize
+	\small
 	\caption{Comparison of Latency Statistics in Salamander 4 after Configurations}
@@
 			\hline
-			\textbf{Version}        & \textbf{Lat Min (ms)} & \textbf{Lat Avg (ms)} & \textbf{Lat Max (ms)} & \textbf{Std Dev (ms)} \\ \hline
+			\textbf{Salamander 4 Version}        & \textbf{Lat Min (ms)} & \textbf{Lat Avg (ms)} & \textbf{Lat Max (ms)} & \textbf{Std Dev (ms)} \\ \hline
 			Salamander 4 Bare Metal              & 1.211                 & 1.347                 & 1.49                  & 0.082                 \\ \hline

### 2024-09-02 Masterarbeit_Robotik_Pamuk.tex
Fertiggggg

@@
 
-\bigskip \noindent Erste Latenzmessungen zeigten eine signifikante Leistungslücke zwischen der Bare-Metal und dem virtualisierten Setup.
-Daher wird ein umfangreicher Tuning-Prozess durchgeführt, um Echtzeit-Performance und Determinismus zu erreichen.
-Diese Modifikationen umfassen Konfigurationen, die das BIOS, den Kernel, das Host-Betriebssystem, die QEMU/KVM-Virtualisierungsschicht und das Salamander 4 Gast-Betriebssystem selbst betreffen.
-Die Worst-Case-Latenz wurde von 707.622~µs auf 17.134~µs reduziert, was der Bare-Metal-Performance von 10.709~µs nahekommt.
+\bigskip \noindent Erste Latenzmessungen zeigten eine signifikante Leistungslücke zwischen der Bare-Metal und dem virtualisierten Setup.~Daher wird ein umfangreicher Tuning-Prozess durchgeführt, um Echtzeit-Performance und Determinismus zu erreichen.~Diese Modifikationen umfassen Konfigurationen, die das BIOS, den Kernel, das Host-Betriebssystem, die QEMU/KVM-Virtualisierungsschicht und das Salamander 4 Gast-Betriebssystem selbst betreffen.~Die Worst-Case-Latenz wurde von 707.622~µs auf 17.134~µs reduziert, was der Bare-Metal-Performance von 10.709~µs nahekommt.
 Darüber hinaus wird die Verbesserung der Echtzeit-Performance und des Determinismus anhand einer Roboteranwendung validiert, bei der die abgestimmte Virtualisierung mit der nicht abgestimmten und der Bare-Metal-Version verglichen wird.

### 2024-09-20 Masterprüfung/Abgabe/Pamuk_MT+Paper_02092024/Pamuk_MT_Paper_02092024.tex
presentation

--- 2024-09-02 Paper.tex
@@
 {Pamuk, Halil Ibrahim, BSc.}            % Student-Name
-{51842568}                                    % Student-PK-Nr
+{2210331012}                                    % Student-PK-Nr
 {Rauh, Sebastian, MSc. BEng.}            % Name FH-Betreuer (1. BegutachterIn)
